### 2022-12-30 data_cleaning/Data_cleaning_and_covariates.qmd
Created repository and README file

---
title: "Fields COVID-19 dataset exploration"
author: "Ariel Mundo"
format: pdf
toc: true
toc-depth: 2
editor: visual
urlcolor: blue
#bibliography: refs.bib
---

```{r, setup, echo=FALSE}
#| warning: false
library(tidyverse)
library(here)
library(lubridate)
library(patchwork)
library(visdat)
library(gt)
library(scales)
library(scico)

thm1<-scale_fill_scico_d(palette="buda",begin=0.3, end=0.8, direction = -1, aesthetics = "fill")

thm2<-scale_fill_scico_d(palette="oleron",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")

thm3<-scale_fill_scico_d(palette="fes",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")
```

# Background

This dataset is about vaccination and other COVID measures in different cities in Ontario. The covariates available are:

-   Age (only above 16, if age above 75 then it appears as 98)
-   Age-group (generated from age)
-   Employment status
-   Remote work within the last month
-   If person receives paid sick leave
-   Number of people in household
-   Number of people from household that attend school
-   Chronic illnesses within the household (*I will not analyze this one as the question is too open and asks about age and disease at the same time*)
-   Race
-   Three first digits of postal code
-   Day, month and year the survey was accessed

There are many responses from the survey on the dataset, but there are some that I would be interesed in analyzing:

-   "Have you received the first dose of the COVID vaccine?" (y/n)

-   (If answered "yes" above) "Have you received the second dose of the COVID vaccine?" (y/n)

-   (If answer was "no" to the first question) "If a vaccine was made available to you you would:"

          - definitely get
          - definitely not get
          - probably get
          - probably not get

**Note: As of Dec 30, 2022 I no longer intend to analyze this last question as after cleaning the dataset it has a missing rate of about 76% and therefore it has too few observations compared to the first two questions**.

# Preliminary Analyses

This document focuses on some preliminary analyses of the data I undertook to assess its representativeness. I performed five major tasks for this preliminary analysis:

-   Choosing covariates to assess their missigness rates
-   Cleaning the data in order to keep all the covariates with a 0% missigness rate
-   Race and ethnicity representativeness of the clean data
-   Trends in the survey answers by geographical location
-   Trends in survey answers by group age

Overall the goal of this analysis was to see what adjustments would need to be made to formally analyze the data.

As way of comparison, I used the 2016 census data for Ontario.

```{r,load-data,warning=FALSE,echo=FALSE}

#the dataframe has missing answers as blankspaces, need to change them all to NA

data<-tibble(read.csv(here("data","Fields_data.csv"),na.strings = c("")))

#set the time and date fields in the format that is required by lubridate

data$time_date_code<-ymd(data$time_date_code)
```

```{r, rename-columns,echo=FALSE,warning=FALSE}

#get a new dataframe and extract variables of interest and make them factors

#get the variables of interest, convert to factors those required
d1<-data%>%select(
  time_date_code,
  city,
  age,
  age_group,
  q02_first_dose,
  q02a_second_dose,
  q11_employed,
  q13_hh_size,
  q09_hh_income,
  q02b_vaccine,
  q16_race,
  q11a_work_remotely,
  q11b_sick_leave,
  q14_hh_attending_school,
  q24a_postal_first,
  q24b_postal_second,
  q24c_postal_third)%>%
  rename(date=time_date_code,
         first_dose=q02_first_dose,
         second_dose=q02a_second_dose,
         get_vaccine=q02b_vaccine,
         h_size=q13_hh_size,
         income=q09_hh_income,
         employed=q11_employed,
         race=q16_race,
         remote_work=q11a_work_remotely,
         sick_leave=q11b_sick_leave,
         school=q14_hh_attending_school,
         pc_1=q24a_postal_first,
         pc_2=q24b_postal_second,
         pc_3=q24c_postal_third
         )%>%
  mutate(across(c("age_group",
            "city",
            "employed",
            "h_size",
            "income",
            "race",
            "remote_work",
            "sick_leave",
            "school"),as.factor))

```

## Choosing covariates

The next step is to see from the covariates what are the ones with the highest number of missing observations, in order to decide what can we include in the model.

```{r}
#| echo: false
#| warning: false
#| tbl-cap: Percentage of missing observations \n all covariates

t1<-d1%>%
  select(age_group,
         employed,
         h_size,
         income,race,
         remote_work,
         sick_leave,
         school,
         pc_1,
         pc_2,
         pc_3)%>%
 summarise_all(funs(sum(!is.na(.)/length(.))))%>%#count percentage of NAs
  pivot_longer(everything())%>%
  arrange(desc(value))%>%
  rename(observations=value,
         variable=name)

#get actual observations
t2<-d1%>%
  select(age_group,
         employed,
         h_size,
         income,race,
         remote_work,
         sick_leave,
         school,
         pc_1,
         pc_2,
         pc_3)%>%
 summarise_all(funs(sum(is.na(.)/length(.))))%>%#count percentage of NAs
  pivot_longer(everything())%>%
  arrange(desc(value))%>%
  rename(missigness=value,
         variable=name)

#joining both to make table

t3<-left_join(t1,t2,by="variable")

#make table

t3%>%
  gt()%>%
  fmt_percent(
   columns=missigness,
    decimals=1)

```

From the table, it can be seen that the covariate with the least amount of observations is "sick_leave".

However, there are a couple of things we want to consider: The answer to "sick_leave" was answered only by those that reported to be employed (the survey design made this response conditional). Therefore, those unemployed would be excluded in an analysis that considers this variable.

At the very least, it would be interesting to analyze how the reported status of vaccination changed over geographical location and time by:

-   age group
-   income
-   race
-   employment status

From these covariates, the one with the highest missing rate of observations is the employment status. I next cleaned the data in order to have complete observations about employment, and to see how the missing rates looked for the other covariates.

```{r}
#| echo: false
#| warning: false
#| tbl-cap: Missing observations and histograms of the data
#| fig-width: 10

clean_data<-d1%>%
  select(age_group,
         employed,
         age,
         income,
         race,
         city,
         h_size,
         date,
         first_dose,
         second_dose,
         get_vaccine)%>%
  drop_na(employed)

### histogram plots to see issues about income, age and household #####

# clean_data%>%
#   ggplot(aes(x=age))+
#   geom_histogram(binwidth=10)+
#   facet_wrap(~income)

a<-d1%>%
  mutate(income_ord = fct_relevel(income,
                                   "under_15000",
                                   "15000_24999",
                                   "25000_39999",
                                   "40000_59999",
                                   "60000_89999",
                                   "90000_109999",
                                  "over_110000"
                                   ))%>%
 #drop_na(age,income,h_size)%>%
  ggplot(aes(x=age))+
  geom_histogram(binwidth=10)+
  facet_wrap(~income_ord)+
  ylim(c(0,1200))+
  theme(strip.text.x = element_text(size = 8))+
  theme_classic()+
  thm1

b<-d1%>%
  mutate(h_size_ord = fct_relevel(h_size,
                                   "1",
                                   "2",
                                   "3",
                                   "4",
                                   "5",
                                   "6",
                                   "7",
                                   "8",
                                   "9",
                                   "10_or_more"
                                   ))%>%
  #drop_na(age,income,h_size_ord)%>%
  ggplot(aes(x=age))+
  geom_histogram(binwidth=10)+
  facet_wrap(~h_size_ord)+
  ylim(c(0,200))+
  theme(strip.text.x = element_text(size = 8))+
  theme_classic()+
  thm1

c<-clean_data%>%
   mutate(income_ord = fct_relevel(income,
                                   "under_15000",
                                   "15000_24999",
                                   "25000_39999",
                                   "40000_59999",
                                   "60000_89999",
                                   "90000_109999",
                                  "over_110000"
                                   ))%>%
  ggplot(aes(x=age))+
  geom_histogram(binwidth=10)+
  facet_wrap(~income_ord)+
  theme(strip.text.x = element_text(size = 5))+
  theme_classic()+
  thm1

d<-clean_data%>%
   mutate(h_size_ord = fct_relevel(h_size,
                                   "1",
                                   "2",
                                   "3",
                                   "4",
                                   "5",
                                   "6",
                                   "7",
                                   "8",
                                   "9",
                                   "10_or_more"
                                   ))%>%
  #drop_na(age,income,h_size_ord)%>%
  ggplot(aes(x=age))+
  geom_histogram(binwidth=10)+
  facet_wrap(~h_size_ord)+
  ylim(c(0,200))+
  theme_classic()+
  theme(strip.text.x = element_text(size = 8))+
  thm1

a+b+plot_annotation(title="Original dataset")

c+d+plot_annotation(title="Clean dataset")

e<-clean_data%>%
  #drop_na(age,income,h_size_ord)%>%
  ggplot(aes(x=age,fill=employed))+
  geom_histogram(stat="count",binwidth=10)+
  #ylim(c(0,200))+
facet_wrap(~income)+
  theme_classic()+
  ggtitle("Clean dataset")+
  #theme(strip.text.x = element_text(size = 8))+
  thm1

f<-d1%>%
  #drop_na(age,income,h_size_ord)%>%
  ggplot(aes(x=age,fill=employed))+
  geom_histogram(stat="count",binwidth=10)+
  #ylim(c(0,200))+
facet_wrap(~income)+
  theme_classic()+
  ylim(c(0,100))+
  ggtitle("Original dataset")+
  #theme(strip.text.x = element_text(size = 8))+
  thm1

f+e

t4<-clean_data%>%
 summarise_all(funs(sum(is.na(.)/length(.))))%>% #count percentage of NAs
  pivot_longer(everything())%>%
  arrange(desc(value))%>%
  rename(missigness=value,
         variable=name)%>%
  gt()%>%
  fmt_percent(
   columns=missigness,
    decimals=1)
t4

```

There seem to be outliers where people \<25 y of age report having an income \>110k while living in a household of 1. Explore the reported household composition of those \<25 with income\>110k.

```{r,outliers}
test<-clean_data %>%
  filter(age_group=="16_24" &
           #h_size==1 &
           income=="over_110000"

  )

test%>%
  ggplot(aes(x=h_size))+
  geom_histogram(stat="count")

```

From the plot it can be seen that about 20 observations are the ones that report a household of 1. The Census data indicates that the average household size is 2.6, and that he median total income in the province is \$33,539. Therefore, this would mean that an income of \>110 could be expected from a household of 4 people and above.

This table shows something interesting due to the nature of the dataset: the missing observations across variables are different, which is why now, although in the original dataset there were more observations of "race" than "employment", when we select complete observations for the latter we lose some observations for race. The second step of cleaning would be to remove the missing observations from race as well, and see what the proportions of covariates and responses are when compared to the census data.

```{r,complete-obs}
#| echo: false
#| warning: false
#| tbl-cap: Clean dataset missigness

t7<-clean_data%>%
  drop_na(race)%>%
 summarise_all(funs(sum(is.na(.)/length(.))))%>% #count percentage of NAs
  pivot_longer(everything())%>%
  arrange(desc(value))%>%
  rename(missigness=value,
         variable=name)%>%
  gt()%>%
  fmt_percent(
   columns=missigness,
    decimals=1)
t7
```

It can be seen that after removing all missing observations in the covariates, the only response that still has missing observations is the answers for the second dose of the vaccine, with 27% of missing observations.

## Race and Ethnicity

Once the data was clean, I explored how the race and ethnicity information from the dataset compared to the data from the Census.

```{r,ethnic information}
#| echo: false
#| warning: false
#| tbl-cap: Ethnic information from the clean dataset
  t6<-clean_data%>%
    select(race)%>%
    group_by(race)%>%
    summarise(observations = n()) %>%
  mutate(percentage = observations / sum(observations))%>%
    gt()%>%
      fmt_percent(
   columns=percentage,
    decimals=1)

t6
```

Below is the information provided by the 2016 Canada Census for Ontario for different visible minorities and ethnic origins. [Link to Census Data](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Ethnic%20origin&TABID=1&type=1).

| Ethnicity/Race | percentage |
|----------------|:-----------|
| Arab           | 1.6%       |
| Black          | 4.7%       |
| East Asian     | 6.6%       |
| Indigenous     | 3.9%       |
| Latin American | 1.5%       |
| Mixed          | 1.0%       |
| Other          | 0.7%       |
| South Asian    | 8.7%       |

: Data from the 2016 Census for Ontario

In the table above, the following data were obtained from the "Visible Minorities" from the Census:

-   Arab
-   Black
-   South Asian
-   East Asian (computed by adding the Chinese, Korean and Japanese entries)
-   Mixed ("multiple visible minorities" entry in the census)
-   Other (from the "visible minority n.i.e" entry in the census)

The following data were obtained from the "ethnic origin" data from the census: - Indigenous - European Origins

However, according to the census "ethnicity" and "visible minority" seem to have different values. For example, in the Visible Minority entry of the census Latin American is 1.5%, but in the "Ethnic Origin" data the value is 2.4%.

Moreover, I would need to adjust the data for Pacific Islanders, as the survey included East Asian and Pacific Islanders in the same question but they are categorized differently in the census.

Lastly, the survey had "white caucasian" as one of the options but that category is not present in the census data. The most approximate entry on the census is "European origins" (61.6%).

**Overall, we will have to weight the data because it is not representative of the population distribution in Ontario**.

## Geographical location

Next, I was interested in analyzing the geographical distribution of the participants over time, to see the percentage of answers from Toronto. This is important to see where the majority of the answers came from (is there such thing as geographical weighting?). The same will be done with the age groups of the participants, as this is data that is reported in the census.

```{r,plot-location}
#| echo: false
#| warning: false
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Percentage of survey answers by city

p1<-clean_data%>%
  select(date,city)%>%
  mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
                            city!="Toronto"~ "All other cities"
                          )))%>%
  group_by(date,location)%>%
  summarize(Freq=n())%>%
  ggplot(aes(x=date, y=Freq,fill=factor(location)
                ))+
  geom_bar(position="fill",stat="identity",show.legend = TRUE)+
  theme_classic()+
  ylab("Percentage")+
  theme(legend.title=element_blank())+
  scale_x_date(date_labels = "%b-%Y")+
  thm1

# p2<-clean_data%>%
#   select(date,age_group)%>%
#   mutate(ag=as.numeric(case_when(age_group=="16_24"~1,
#                             age_group=="25_34"~ 2,
#                             age_group=="35_44"~ 3,
#                             age_group=="45_54"~ 4,
#                             age_group=="55_64"~ 5,
#                             age_group=="65_and_over"~6
#                             )))%>%
#   group_by(date,ag)%>%
#   summarise(n = n()) %>%
#   mutate(perc = n / sum(n)
#          )%>%
#   ggplot(aes(x=date,y=perc,color=factor(ag)))+
#   geom_point()+
#   geom_line(aes(x=date,y=perc,group=factor(ag)))+
#   facet_wrap(~ag)+
#   scale_x_date(date_breaks = "2 months")+
#   ggtitle("Age groups")+
#   theme_classic()+
#   thm1

p1

```

The proportion of responses from Toronto is somewhat uniform across time, around 20%-25%. The Census data indicates that Toronto has around 38% of the total population province (5,433,590/14,223,942). This would mean that a correction would be needed also for the proportion of answers. Note that I only analyzed the proportion of answers from Toronto as it is the major urban centre of the province, but also because the dataset contains at least 500 different city entries. I did not explore geographically if some of these cities lie within what is considered Toronto (the dataset is also not explicit as the method used to assign the cities).

## Age groups

Another important aspect to consider is how the trends in the responses look over time by age group. This is because from the census data, the age groups for the province are as follows:

| Group age   | Percentage of population |
|-------------|:-------------------------|
| 15-24       | 12.7%                    |
| 25-34       | 12.9%                    |
| 35-44       | 12.8%                    |
| 45-54       | 14.9%                    |
| 55-64       | 13.7%                    |
| 65 and over | 16.7                     |

: Age distributions from the 2016 Census for Ontario

However, when I plotted the proportion of answers at each time point among groups, the distribution looks rather different:

```{r,trends-by-age}
#| echo: false
#| warning: false
#| fig-height: 4
#| fig-width: 8
#| fig-cap: Age group distributions from the dataset

p2<-clean_data%>%
  select(date,age_group)%>%
   mutate(age_group_ord = fct_relevel(age_group,
                                  "65_and_over",
                                  "55_64",
                                  "45_54",
                                  "35_44",
                                  "25_34",
                                  "16_24"))%>%
  group_by(date,age_group_ord)%>%
  summarize(Freq=n())%>%
  ggplot(aes(x=date, y=Freq,fill=age_group_ord))+
  geom_bar(position="fill",stat="identity",show.legend = TRUE)+
  theme_classic()+
  # scale_fill_discrete(name="Age group",
  #                        labels=c("16-24", 
  #                                 "25-34",
  #                                 "35-44",
  #                                 "45-54",
  #                                 "55-64",
  #                                 "65 and over"))+
  ylab("Percentage")+
  scale_y_continuous(breaks = seq(0, 1, by = 0.1))+
  theme(legend.title = element_blank())+
   scale_x_date(date_labels = "%b-%Y")+
  thm2

p2

# a4<-clean_data%>%
#   group_by(age_group)%>%
#   count(date,first_dose)%>%
#   ggplot(aes(x=date,y=n,color=first_dose))+
#   geom_point(show.legend = FALSE)+
#   geom_line(aes(x=date,y=n,group=first_dose),show.legend = TRUE)+
#   facet_wrap(~age_group)+
#   ggtitle("First dose")+
#   theme_classic()+
#   thm1
# 
# a5<-clean_data%>%
#   group_by(age_group)%>%
#   count(date,second_dose)%>%
#   ggplot(aes(x=date,y=n,color=second_dose))+
#   geom_point()+
#   geom_line(aes(x=date,y=n,group=second_dose))+
#   facet_wrap(~age_group)+
#   ggtitle("Second dose")+
#   theme_classic()
# 
# layout<-"AAAAA
#          AAAAA
#          AAAAA
#          AAAAA
#          BBBBB
#          BBBBB
#          BBBBB
#          BBBBB"
# 
# a4+a5+plot_layout(design = layout)
# 
# 
# 
# p3<-clean_data%>%
#   select(date,age_group)%>%
#   mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
#                             city!="Toronto"~ "All other cities"
#                           )))%>%
#   group_by(date,location)%>%
#   summarize(Freq=n())%>%
#   ggplot(aes(x=date, y=Freq,fill=factor(location)
#                 ))+
#   geom_bar(position="fill",stat="identity",show.legend = TRUE)+
#   ggtitle("Geographical location of survey answers")+
#   theme_classic()+
#   ylab("Percentage")+
#   theme(legend.title=element_blank())+
#   scale_x_date(date_labels = "%b-%Y")+
#   thm1

```

It can be seen from the graph that the trends in answers by group age do not follow the overall distribution from the province. In particular, the 16-24 group represents in most cases at least 20% of the answers, whereas the 35-44 group is less than 10% in some cases.

## Income

The final aspect I wanted to explore was the income distribution on the responses over time. In this regard the question from the survey was "What is your household annual income?". To compare the results of this answer, I used the "Household total income groups in 2015 for private households" from the Census data. The information from the census appears next.

| Household income range (CAD) | Percentage |
|------------------------------|:-----------|
| \< 15,000                    | 5.7%       |
| 15,000 - 24,999              | 7.5%       |
| 25,000 - 39,999              | 11.6%      |
| 40,000- 59,999               | 15.4%      |
| 60,000 - 89,999              | 19.5%      |
| \>90,000                     | 40.4%      |

: Income percentages from the 2016 Census for Ontario

One thing to keep in mind in this case is that the brackets for income in the census data are different than the brackets used in the survey. The census does CAD 4,999 brackets (e.g., CAD 5,000- CAD 9,9999) up to CAD 49,999 and then it does CAD 9,999 brackets up to CAD 99,999 but after that, the brackets increase to CAD 24,999, and therefore, one cannot obtain percentages for the 90,000-109,999 and \>110,000 brackets from the survey.

```{r,income}
#| echo: false
#| warning: false
#| fig-width: 8
#| fig-height: 4

p3<-clean_data%>%
  select(date,income)%>%
  mutate(income_ord = fct_relevel(income,
                                  "over_110000",
                                  "90000_109999",
                                  "60000_89999",
                                   "40000_59999",
                                  "25000_39999",
                                     "15000_24999",
                                   "under_15000"))%>%
  group_by(date,income_ord)%>%
  summarize(Freq=n())%>%
  ggplot(aes(x=date, y=Freq,fill=income_ord))+
  geom_bar(position="fill",stat="identity",show.legend = TRUE)+
  theme_classic()+
  ylab("Percentage")+
  scale_y_continuous(breaks = seq(0, 1, by = 0.1))+
  theme(legend.title = element_blank())+
   scale_x_date(date_labels = "%b-%Y")+
  thm3

p3

```

Again, there are differences here in the patterns of response. The \<15,000 bracket accounts for about 20% of the responses, a much higher rate than the percentage they represent from the census, and there is variation within the other brackets as well.

# Conclusions

Overall, it seems that poststratification corrections would be needed for the dataset with regard to race, ethnicity, age groups

### 2022-12-30 protocol/Background_and_methods_Dec_30.qmd
Created repository and README file

---
title: "Covid Fields Data Protocol"
author:
  - name: Ariel Mundo Ortiz
    orcid: 0000-0002-6014-4538
    email: ariel.mundo.ortiz@umontreal.ca
    affiliations:
      - name: École de santé publique - Université de Montréal
        address: 7101 Park Avenue
        city: Montreal
        state: QC
        postal-code: H3N 1X9
format:
  pdf:
    link-citations: true
    number-sections: true
editor: visual
bibliography: bibliography/refs.bib
keep-tex: true
csl: bibliography/jama.csl
knitr:
  opts_chunk:
    dev: "ragg_png"
---

# Background

The COVID-19 pandemic continues around the world with more than 600 million confirmed cases as of November of 2022 [@WHO-Covid]. During the first months of the pandemic in early 2020, non-pharmaceutical interventions (e.g., masking, social distancing) were the only methods available to manage the spread of the disease, but the rapid development of vaccines against the virus permitted their approval and use in some countries towards the last month part of 2020. For example, in the US and Canada vaccine campaigns began in mid-December of 2020 [@tanne2020; @bogoch2022]. Although it has been estimated that vaccines against COVID-19 have prevented around 14 millions of deaths worldwide [@watson2022], the rollout of COVID-19 vaccines has faced multiple challenges since its inception.

In this regard, vaccination efforts have faced multiple: Inequalities with regard to vaccine access due to socio-economic factors, vaccine hesitancy, and differences in vaccination rates across different segments of the population are among the challenges identified in the administration of COVID-19 vaccines [@gerretsen2021; @nafilyan2021; @malik2020]. In the case of Canada, lower vaccine uptake has been associated with socio-economic factors such as younger age, educational level, presence of children in the household, lack of a regular healthcare provider, ethnic origin, and financial instability [@guay2022; @muhajarine2021; @carter2022].

Additionally, it has been shown that geography also plays a crucial role in vaccination rates, as they vary due to spatial differences in attitudes towards vaccination [@malik2020], geographical differences in vaccine access and supply, vaccination location availability, and lack of prioritization of vulnerable groups [@bogoch2022; @nguyen2021].

Studies that analyze geographical variations in vaccine uptake can help inform public health decision-makers to design policies to that are aimed at addressing vaccination disparities. In this regad, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governamental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].

In Canada, studies that have used a spatial approach to analyze vaccine uptake have shown disparities in vaccination rates across low and high income neighborhoods in the city of Toronto [@choi2021], among adolescents from deprived neighborhoods in the city of Montreal [@mckinnon2021], and highlighted disparities in vaccination status depending on age, income, and ethnic origin in all of the Canadian provinces[@guay2022]. However, to the best of our knowledge, there are no studies that have analyzed vaccination status within a province at the district/rural municipality level, which can be useful to identify inequalities over these geographical areas, thus providing a dissagregated view that can help understand the barriers for vaccine delivery in the case of visible minorities, which have been disproportionately impacted in the pandemic [@hussain2022].

# Research Question

This study will examine self-reported COVID-19 vaccination status in the province of Ontario in order to determine the influence in vaccination status due to socio-economic (e.g., ethnic origin, age, income) and geographical factors (at the municipal level).

# Methods

## Data source: survey overview

We obtained data from the Fields Institute for Research in Mathematical Sciences' (henceforth Fields) *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario which ran from Sept 30, 2021 until January 17, 2022. This survey was commissioned by Fields and the Mathematical Modelling of COVID-19 Task Force, under the supervision of Dr. Kumar Murty, the Director of Fields with funding from the Canadian Institutes of Health Research. The survey was conducted by a third-party service provider (RIWI Corp.), under ethical guidance from University of Toronto.

The survey was deployed using random domain intercept technology. Briefly, when web users clicked on a registered but commercially inactive web link or typed in a web address for a site that was dormant, they had a random chance of that link being temporarily managed by the company that administered the survey (RIWI Corp). Thus, instead of coming across a notification about the status of the site("this page does not exist"), the survey was deployed to the user. Web users then decided whether to anonymously participate, exiting the survey at any time if desired [@sargent2022].

Respondents who wished to participate were asked to select their age from a matrix of values, and subsequent questions were displayed one at a time, after the respondent confirmed their selection by answering and selecting "next". Those who do not wished to participate were asked to either close the browser window or navigate away from the domain. After the survey closed (complete or incomplete) no one from that internet protocol (IP) address could access the survey again and the domain entry point rotated such that if a respondent were to attempt to access the survey again, share the link, or enter via the same address using an alternative IP address, the survey would not render.

Additionally, respondents who indicated they were under the age of 16 were exited from the survey. No record was created in this case and due to domain cycling these users were unable to navigate back to the "age select" screen. The personal identifier information from each respondent was automatically scrubbed and replaced by a unique ID. Respondents were drawn exclusively from the province of Ontario, as per their devices meta-data.

## Survey responses

### Socio-demographic factors {#sec-socio-demographic-factors}

From the different answers provided by the survey respondents, we selected the age group which they belonged to, income bracket, race/ethnicity, and employment status. The original survey included additional questions (e.g., sick leave, remote work, presence of minors in the household) but the survey design, which permitted respondents to exit the survey at any point resulted in a high rates of missing data for most of these answers. The socio-economic factors chosen for this study were the ones that had both the lowest rates of missingness and that provided an adequate level socio-economic and demographic information for our analysis. Information about the chosen socio-economic factors from the survey is provided in @tbl-covariates.

| Variable             | Values                                                                                                                          |
|:---------------------|---------------------------------------------------------------------------------------------------------------------------------|
| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
| Employment status    | yes, no                                                                                                                         |

: Selected socio-economic factors from the survey {#tbl-covariates}

### Vaccination status

We selected two of the questions that were asked in the survey regarding vaccination status:

-   "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no"

-   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no"

<!-- -   (If answer was "no" to the first question) "If a vaccine was made available to you you would:" with possible answers "definitely get", "definitely not get", "probably get", and "probably not get" -->

<!--           - definitely get -->

<!--           - definitely not get -->

<!--           - probably get -->

<!--           - probably not get -->

## Data cleaning

The original dataset contained 39,029 entries (where each entry corresponded to a set of answers provided by a unique respondent). Following a preliminary analysis to identify the missing rates across the different answers within each entry, it was identified that many of the answers had high missing rates (\>80%) (note that the graph will be in the Appendix). Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variables.

The cleaning process also included removing outliers that were identified during the preliminary analyses. Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000. After cleaning the dataset contained 5,247 entries (unique respondents).

## Corrections

Differences between the clean dataset and the 2016 Census data for Ontario regarding proportions of age groups, income, and ethnicity/race were identified. Therefore, the proportions of each of these variables were corrected using an iterative proportional fitting procedure (also known as *raking*) [@deming1940] in R using the `survey` package. Proportions for the correction were obtained from the 2016 Census Data for Ontario. Because the categories provided by the survey in some cases (e.g., race/ethnicity categories) did not match the categories from the Census, we aggregated some of them to obtain an approximation to the categories in the Census. The aggregation is described in detail in the Appendix.

## Geographical location {#sec-geographical-location}

For each survey participant certain data was automatically captured. This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset. Because our interest lies in exploring vaccination status within a governamental administrative unit that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities].

## Statistical model

We used a multivariable logistic regression with the socio-economic factors described in @sec-socio-demographic-factors and the aggregated demographical locations from @sec-geographical-location.

## References

::: {#refs}
:::

### 2022-12-30 protocol/Background_and_methods_Dec_30.tex
Created repository and README file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Covid Fields Data Protocol},
  pdfauthor={Ariel Mundo Ortiz},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Covid Fields Data Protocol}
\author{Ariel Mundo Ortiz}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, frame hidden, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi

\hypertarget{background}{%
\section{Background}\label{background}}

The COVID-19 pandemic continues around the world with more than 600
million confirmed cases as of November of
2022\textsuperscript{\protect\hyperlink{ref-WHO-Covid}{1}}. During the
first months of the pandemic in early 2020, non-pharmaceutical
interventions (e.g., masking, social distancing) were the only methods
available to manage the spread of the disease, but the rapid development
of vaccines against the virus permitted their approval and use in some
countries towards the last month part of 2020. For example, in the US
and Canada vaccine campaigns began in mid-December of
2020\textsuperscript{\protect\hyperlink{ref-tanne2020}{2},\protect\hyperlink{ref-bogoch2022}{3}}.
Although it has been estimated that vaccines against COVID-19 have
prevented around 14 millions of deaths
worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{4}}, the
rollout of COVID-19 vaccines has faced multiple challenges since its
inception.

In this regard, vaccination efforts have faced multiple: Inequalities
with regard to vaccine access due to socio-economic factors, vaccine
hesitancy, and differences in vaccination rates across different
segments of the population are among the challenges identified in the
administration of COVID-19
vaccines\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{5}--\protect\hyperlink{ref-malik2020}{7}}.
In the case of Canada, lower vaccine uptake has been associated with
socio-economic factors such as younger age, educational level, presence
of children in the household, lack of a regular healthcare provider,
ethnic origin, and financial
instability\textsuperscript{\protect\hyperlink{ref-guay2022}{8}--\protect\hyperlink{ref-carter2022}{10}}.

Additionally, it has been shown that geography also plays a crucial role
in vaccination rates, as they vary due to spatial differences in
attitudes towards
vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{7}},
geographical differences in vaccine access and supply, vaccination
location availability, and lack of prioritization of vulnerable
groups\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3},\protect\hyperlink{ref-nguyen2021}{11}}.

Studies that analyze geographical variations in vaccine uptake can help
inform public health decision-makers to design policies to that are
aimed at addressing vaccination disparities. In this regad, previous
geographical (spatial) analyses of vaccination rates have shown that
variations in vaccine uptake can occur within small governamental
administrative units (e.g., counties in the case of the
US)\textsuperscript{\protect\hyperlink{ref-mollalo2021}{12}--\protect\hyperlink{ref-bhuiyan2022}{15}},
and that geographical analyses can be predictive of booster uptake
patterns\textsuperscript{\protect\hyperlink{ref-wood2022}{16}}.

In Canada, studies that have used a spatial approach to analyze vaccine
uptake have shown disparities in vaccination rates across low and high
income neighborhoods in the city of
Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{17}}, among
adolescents from deprived neighborhoods in the city of
Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{18}}, and
highlighted disparities in vaccination status depending on age, income,
and ethnic origin in all of the Canadian
provinces\textsuperscript{\protect\hyperlink{ref-guay2022}{8}}. However,
to the best of our knowledge, there are no studies that have analyzed
vaccination status within a province at the district/rural municipality
level, which can be useful to identify inequalities over these
geographical areas, thus providing a dissagregated view that can help
understand the barriers for vaccine delivery in the case of visible
minorities, which have been disproportionately impacted in the
pandemic\textsuperscript{\protect\hyperlink{ref-hussain2022}{19}}.

\hypertarget{research-question}{%
\section{Research Question}\label{research-question}}

This study will examine self-reported COVID-19 vaccination status in the
province of Ontario in order to determine the influence in vaccination
status due to socio-economic (e.g., ethnic origin, age, income) and
geographical factors (at the municipal level).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-source-survey-overview}{%
\subsection{Data source: survey
overview}\label{data-source-survey-overview}}

We obtained data from the Fields Institute for Research in Mathematical
Sciences' (henceforth Fields) \emph{Survey of COVID-19 related
Behaviours and Attitudes}, a repeated cross sectional survey focused on
the Canadian province of Ontario which ran from Sept 30, 2021 until
January 17, 2022. This survey was commissioned by Fields and the
Mathematical Modelling of COVID-19 Task Force, under the supervision of
Dr.~Kumar Murty, the Director of Fields with funding from the Canadian
Institutes of Health Research. The survey was conducted by a third-party
service provider (RIWI Corp.), under ethical guidance from University of
Toronto.

The survey was deployed using random domain intercept technology.
Briefly, when web users clicked on a registered but commercially
inactive web link or typed in a web address for a site that was dormant,
they had a random chance of that link being temporarily managed by the
company that administered the survey (RIWI Corp). Thus, instead of
coming across a notification about the status of the site(``this page
does not exist''), the survey was deployed to the user. Web users then
decided whether to anonymously participate, exiting the survey at any
time if
desired\textsuperscript{\protect\hyperlink{ref-sargent2022}{20}}.

Respondents who wished to participate were asked to select their age
from a matrix of values, and subsequent questions were displayed one at
a time, after the respondent confirmed their selection by answering and
selecting ``next''. Those who do not wished to participate were asked to
either close the browser window or navigate away from the domain. After
the survey closed (complete or incomplete) no one from that internet
protocol (IP) address could access the survey again and the domain entry
point rotated such that if a respondent were to attempt to access the
survey again, share the link, or enter via the same address using an
alternative IP address, the survey would not render.

Additionally, respondents who indicated they were under the age of 16
were exited from the survey. No record was created in this case and due
to domain cycling these users were unable to navigate back to the ``age
select'' screen. The personal identifier information from each
respondent was automatically scrubbed and replaced by a unique ID.
Respondents were drawn exclusively from the province of Ontario, as per
their devices meta-data.

\hypertarget{survey-responses}{%
\subsection{Survey responses}\label{survey-responses}}

\hypertarget{sec-socio-demographic-factors}{%
\subsubsection{Socio-demographic
factors}\label{sec-socio-demographic-factors}}

From the different answers provided by the survey respondents, we
selected the age group which they belonged to, income bracket,
race/ethnicity, and employment status. The original survey included
additional questions (e.g., sick leave, remote work, presence of minors
in the household) but the survey design, which permitted respondents to
exit the survey at any point resulted in a high rates of missing data
for most of these answers. The socio-economic factors chosen for this
study were the ones that had both the lowest rates of missingness and
that provided an adequate level socio-economic and demographic
information for our analysis. Information about the chosen
socio-economic factors from the survey is provided in
Table~\ref{tbl-covariates}.

\hypertarget{tbl-covariates}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1457}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8543}}@{}}
\caption{\label{tbl-covariates}Selected socio-economic factors from the
survey}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Values
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Values
\end{minipage} \\
\midrule()
\endhead
Age group & 15-24,25-34,35-44,45-54,55-64, 65+ \\
Income bracket (CAD) & \textless15,000, 15,000-24,999, 25,000-39,999,
40,000-59,999, 60,000-89,999, \textgreater90,000 \\
Race/ethnicity & Arab/Middle Eastern, Black, East Asian/Pacific
Islander, Indigenous, Latin American, Mixed, South Asian, White
Caucasian, other \\
Employment status & yes, no \\
\bottomrule()
\end{longtable}

\hypertarget{vaccination-status}{%
\subsubsection{Vaccination status}\label{vaccination-status}}

We selected two of the questions that were asked in the survey regarding
vaccination status:

\begin{itemize}
\item
  ``Have you received the first dose of the COVID vaccine?'', with
  possible answers ``yes'' and ``no''
\item
  (If answered ``yes'' to the previous question) ``Have you received the
  second dose of the COVID vaccine?'' with possible answers ``yes'' and
  ``no''
\end{itemize}

\hypertarget{data-cleaning}{%
\subsection{Data cleaning}\label{data-cleaning}}

The original dataset contained 39,029 entries (where each entry
corresponded to a set of answers provided by a unique respondent).
Following a preliminary analysis to identify the missing rates across
the different answers within each entry, it was identified that many of
the answers had high missing rates (\textgreater80\%) (note that the
graph will be in the Appendix). Therefore, the dataset was cleaned in
order to contain only the independent variables of interest with the
lowest missing rates (Table~\ref{tbl-covariates}) and the dependent
variables.

The cleaning process also included removing outliers that were
identified during the preliminary analyses. Specifically, we removed
those respondents that indicated to be below 25 years of age, living in
a household of size 1, and that reported an income above CAD 110,000.
After cleaning the dataset contained 5,247 entries (unique respondents).

\hypertarget{corrections}{%
\subsection{Corrections}\label{corrections}}

Differences between the clean dataset and the 2016 Census data for
Ontario regarding proportions of age groups, income, and ethnicity/race
were identified. Therefore, the proportions of each of these variables
were corrected using an iterative proportional fitting procedure (also
known as
\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{21}}
in R using the \texttt{survey} package. Proportions for the correction
were obtained from the 2016 Census Data for Ontario. Because the
categories provided by the survey in some cases (e.g., race/ethnicity
categories) did not match the categories from the Census, we aggregated
some of them to obtain an approximation to the categories in the Census.
The aggregation is described in detail in the Appendix.

\hypertarget{sec-geographical-location}{%
\subsection{Geographical location}\label{sec-geographical-location}}

For each survey participant certain data was automatically captured.
This included the nearest municipality, which resulted in a total of 578
different municipalities within the dataset. Because our interest lies
in exploring vaccination status within a governamental administrative
unit that encompasses multiple cities, we grouped the municipalities
from the dataset according to the regional government aggregations
provided by the Association of Municipalities of Ontario, which groups
municipalities within regions, counties, districts, and
single-tiers\textsuperscript{\protect\hyperlink{ref-ontario-municipalities}{22}}.

\hypertarget{statistical-model}{%
\subsection{Statistical model}\label{statistical-model}}

We used a multivariable logistic regression with the socio-economic
factors described in Section~\ref{sec-socio-demographic-factors} and the
aggregated demographical locations from
Section~\ref{sec-geographical-location}.

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-WHO-Covid}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{{World Health Organization Coronavirus (COVID-19)
Dashboard}. Accessed November 27, 2022. \url{https://covid19.who.int/}}

\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
2020:m4799.
doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}

\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
social and policy view of vaccination efforts in ontario, canada.
\emph{Cambridge Journal of Regions, Economy and Society}. Published
online November 2022.
doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}

\leavevmode\vadjust pre{\hypertarget{ref-watson2022}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Watson OJ, Barnsley G, Toor J, Hogan AB, Winskill P,
Ghani AC. Global impact of the first year of {COVID}-19 vaccination: A
mathematical modelling study. \emph{The Lancet Infectious Diseases}.
2022;22(9):1293-1302.
doi:\href{https://doi.org/10.1016/s1473-3099(22)00320-6}{10.1016/s1473-3099(22)00320-6}}

\leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
determinants of {COVID}-19 vaccine hesitancy. Inbaraj LR, ed.
\emph{{PLOS} {ONE}}. 2021;16(11):e0258462.
doi:\href{https://doi.org/10.1371/journal.pone.0258462}{10.1371/journal.pone.0258462}}

\leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
inequality in {COVID}-19 vaccination coverage among elderly adults in
england: A national linked data study. \emph{{BMJ} Open}.
2021;11(7):e053402.
doi:\href{https://doi.org/10.1136/bmjopen-2021-053402}{10.1136/bmjopen-2021-053402}}

\leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
of {COVID}-19 vaccine acceptance in the {US}.
\emph{{EClinicalMedicine}}. 2020;26:100495.
doi:\href{https://doi.org/10.1016/j.eclinm.2020.100495}{10.1016/j.eclinm.2020.100495}}

\leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
inequalities in {COVID}-19 vaccination uptake and intent: Results from
the canadian community health survey 2021. \emph{{BMC} Public Health}.
2022;22(1).
doi:\href{https://doi.org/10.1186/s12889-022-14090-z}{10.1186/s12889-022-14090-z}}

\leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
Fahlman M, Kallio N. {COVID}-19 vaccine hesitancy and refusal and
associated factors in an adult population in saskatchewan, canada:
Evidence from predictive modelling. Gesser-Edelsburg A, ed. \emph{{PLOS}
{ONE}}. 2021;16(11):e0259513.
doi:\href{https://doi.org/10.1371/journal.pone.0259513}{10.1371/journal.pone.0259513}}

\leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
{COVID}-19 vaccine uptake in southeastern ontario, canada: Monitoring
and addressing health inequities. \emph{Journal of Public Health
Management and Practice}. 2022;28(6):615-623.
doi:\href{https://doi.org/10.1097/phh.0000000000001565}{10.1097/phh.0000000000001565}}

\leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
Changes in {COVID}-19 vaccination receipt and intention to vaccinate by
socioeconomic characteristics and geographic area, united states,
january 6 {\textendash} march 29, 2021. \emph{Annals of Medicine}.
2021;53(1):1419-1428.
doi:\href{https://doi.org/10.1080/07853890.2021.1957998}{10.1080/07853890.2021.1957998}}

\leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
vaccine hesitancy in the united states. \emph{International Journal of
Environmental Research and Public Health}. 2021;18(18):9488.
doi:\href{https://doi.org/10.3390/ijerph18189488}{10.3390/ijerph18189488}}

\leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
spatial process: {COVID}-19 fully vaccinated rates in u.s. counties.
\emph{American Journal of Preventive Medicine}. 2022;63(6):954-961.
doi:\href{https://doi.org/10.1016/j.amepre.2022.06.006}{10.1016/j.amepre.2022.06.006}}

\leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
the spatiotemporal heterogeneity of the {COVID}-19 vaccination
landscape. \emph{American Journal of Epidemiology}.
2022;191(10):1792-1802.
doi:\href{https://doi.org/10.1093/aje/kwac080}{10.1093/aje/kwac080}}

\leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
social vulnerability index to assess {COVID}-19 vaccine uptake in
louisiana. \emph{{GeoJournal}}. Published online December 2022.
doi:\href{https://doi.org/10.1007/s10708-022-10802-5}{10.1007/s10708-022-10802-5}}

\leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
future spatial patterns in {COVID}-19 booster vaccine uptake. Published
online September 2022.
doi:\href{https://doi.org/10.1101/2022.08.30.22279415}{10.1101/2022.08.30.22279415}}

\leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
trajectories. \emph{Socius: Sociological Research for a Dynamic World}.
2021;7:237802312110529.
doi:\href{https://doi.org/10.1177/23780231211052946}{10.1177/23780231211052946}}

\leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
Social inequalities in {COVID}-19 vaccine acceptance and uptake for
children and adolescents in montreal, canada. \emph{Vaccine}.
2021;39(49):7140-7145.
doi:\href{https://doi.org/10.1016/j.vaccine.2021.10.077}{10.1016/j.vaccine.2021.10.077}}

\leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
Overcoming {COVID}-19 vaccine hesitancy among ethnic minorities: A
systematic review of {UK} studies. \emph{Vaccine}.
2022;40(25):3413-3432.
doi:\href{https://doi.org/10.1016/j.vaccine.2022.04.030}{10.1016/j.vaccine.2022.04.030}}

\leavevmode\vadjust pre{\hypertarget{ref-sargent2022}{}}%
\CSLLeftMargin{20. }%
\CSLRightInline{Sargent RH, Laurie S, Weakland LF, et al. Use of random
domain intercept technology to track {COVID}-19 vaccination rates in
real time across the united states: Survey study. \emph{Journal of
Medical Internet Research}. 2022;24(7):e37920.
doi:\href{https://doi.org/10.2196/37920}{10.2196/37920}}

\leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
\CSLLeftMargin{21. }%
\CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
a sampled frequency table when the expected marginal totals are known.
\emph{The Annals of Mathematical Statistics}. 1940;11(4):427-444.
doi:\href{https://doi.org/10.1214/aoms/1177731829}{10.1214/aoms/1177731829}}

\leavevmode\vadjust pre{\hypertarget{ref-ontario-municipalities}{}}%
\CSLLeftMargin{22. }%
\CSLRightInline{Ontario {M}unicipalities. Accessed December 30, 2022.
\url{https://www.amo.on.ca/about-us/municipal-101/ontario-municipalities}}

\end{CSLReferences}

\end{document}

### 2023-01-04 data_cleaning/Data_cleaning_and_covariates_Jan_03.qmd
Updated geographical data

--- 2022-12-30 data_cleaning/Data_cleaning_and_covariates.qmd
@@
 toc-depth: 2
-editor: visual
+editor: source
 urlcolor: blue
@@
 ```{r,outliers}
-test<-clean_data %>%
+outliers<-clean_data %>%
   filter(age_group=="16_24" &
-           #h_size==1 &
+           h_size==1 &
            income=="over_110000"
@@
 
-test%>%
+outliers%>%
   ggplot(aes(x=h_size))+
@@
 
+#From the plot it is about 21 entries that are outliers. Now, remove these outliers
+
+clean_data<-anti_join(clean_data,outliers)
+
 ```
-
-From the plot it can be seen that about 20 observations are the ones that report a household of 1.
-The Census data indicates that the average household size is 2.6, and that he median total income in the province is \$33,539.
-Therefore, this would mean that an income of \>110 could be expected from a household of 4 people and above.
 
@@
 
-**Overall, we will have to weight the data because it is not representative of the population distribution in Ontario**.
+**Overall, we weighted the data because it is not representative of the population distribution in Ontario**.
 
@@
 p1<-clean_data%>%
-  select(date,city)%>%
+  select(city)%>%
   mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
                             city!="Toronto"~ "All other cities"
-                          )))%>%
-  group_by(date,location)%>%
-  summarize(Freq=n())%>%
-  ggplot(aes(x=date, y=Freq,fill=factor(location)
-                ))+
-  geom_bar(position="fill",stat="identity",show.legend = TRUE)+
+                          )))
+
+p1$location<-as.factor(p1$location)
+
+ggplot(p1 %>% count(location) %>%    # then count number in each group
+         mutate(pct=n/sum(n),               # Calculate percent within each group
+                ypos = 200),  # Calculate label positions
+       aes(location, n, fill=location)) +
+  geom_bar(stat="identity") +
+  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
   theme_classic()+
-  ylab("Percentage")+
-  theme(legend.title=element_blank())+
-  scale_x_date(date_labels = "%b-%Y")+
-  thm1
-
-# p2<-clean_data%>%
-#   select(date,age_group)%>%
-#   mutate(ag=as.numeric(case_when(age_group=="16_24"~1,
-#                             age_group=="25_34"~ 2,
-#                             age_group=="35_44"~ 3,
-#                             age_group=="45_54"~ 4,
-#                             age_group=="55_64"~ 5,
-#                             age_group=="65_and_over"~6
-#                             )))%>%
-#   group_by(date,ag)%>%
-#   summarise(n = n()) %>%
-#   mutate(perc = n / sum(n)
-#          )%>%
-#   ggplot(aes(x=date,y=perc,color=factor(ag)))+
-#   geom_point()+
-#   geom_line(aes(x=date,y=perc,group=factor(ag)))+
-#   facet_wrap(~ag)+
-#   scale_x_date(date_breaks = "2 months")+
-#   ggtitle("Age groups")+
-#   theme_classic()+
-#   thm1
-
-p1
+  thm2
 
@@
 
-The proportion of responses from Toronto is somewhat uniform across time, around 20%-25%.
+The proportion of responses from Toronto is 30.9% of the total.
 The Census data indicates that Toronto has around 38% of the total population province (5,433,590/14,223,942).
 This would mean that a correction would be needed also for the proportion of answers.
-Note that I only analyzed the proportion of answers from Toronto as it is the major urban centre of the province, but also because the dataset contains at least 500 different city entries.
-I did not explore geographically if some of these cities lie within what is considered Toronto (the dataset is also not explicit as the method used to assign the cities).
+
+Additionally, the data needs to be cleaned as some of the responses do not have an assigned city, and the entry only reads "None". These entries will be removed as they lack geographical information.
+
+```{r, remove-entries}
+
+clean_data<-clean_data%>%
+  filter(city!="None")
+
+```
 
@@
 
-However, when I plotted the proportion of answers at each time point among groups, the distribution looks rather different:
+However, when plotted the proportion of answers at each time point among groups, the distribution looks rather different.
 
@@
 p2<-clean_data%>%
-  select(date,age_group)%>%
+  select(age_group)%>%
    mutate(age_group_ord = fct_relevel(age_group,
-                                  "65_and_over",
+                                      "16_24",
+                                      "25_34",
+                                      "35_44",
+                                      "45_54",
                                       "55_64",
-                                  "45_54",
-                                  "35_44",
-                                  "25_34",
-                                  "16_24"))%>%
-  group_by(date,age_group_ord)%>%
-  summarize(Freq=n())%>%
-  ggplot(aes(x=date, y=Freq,fill=age_group_ord))+
-  geom_bar(position="fill",stat="identity",show.legend = TRUE)+
+                                  "65_and_over"
+                                          ))
+
+  ggplot(p2%>%count(age_group_ord) %>%    # G count number in each group
+         mutate(pct=n/sum(n),               # Calculate percent within each group
+                ypos = 200),  # Calculate label positions
+       aes(age_group_ord, n, fill=age_group_ord)) +
+  geom_bar(stat="identity") +
+  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
   theme_classic()+
-  # scale_fill_discrete(name="Age group",
-  #                        labels=c("16-24", 
-  #                                 "25-34",
-  #                                 "35-44",
-  #                                 "45-54",
-  #                                 "55-64",
-  #                                 "65 and over"))+
-  ylab("Percentage")+
-  scale_y_continuous(breaks = seq(0, 1, by = 0.1))+
-  theme(legend.title = element_blank())+
-   scale_x_date(date_labels = "%b-%Y")+
-  thm2
-
-p2
-
-# a4<-clean_data%>%
-#   group_by(age_group)%>%
-#   count(date,first_dose)%>%
-#   ggplot(aes(x=date,y=n,color=first_dose))+
-#   geom_point(show.legend = FALSE)+
-#   geom_line(aes(x=date,y=n,group=first_dose),show.legend = TRUE)+
-#   facet_wrap(~age_group)+
-#   ggtitle("First dose")+
-#   theme_classic()+
-#   thm1
-# 
-# a5<-clean_data%>%
-#   group_by(age_group)%>%
-#   count(date,second_dose)%>%
-#   ggplot(aes(x=date,y=n,color=second_dose))+
-#   geom_point()+
-#   geom_line(aes(x=date,y=n,group=second_dose))+
-#   facet_wrap(~age_group)+
-#   ggtitle("Second dose")+
-#   theme_classic()
-# 
-# layout<-"AAAAA
-#          AAAAA
-#          AAAAA
-#          AAAAA
-#          BBBBB
-#          BBBBB
-#          BBBBB
-#          BBBBB"
-# 
-# a4+a5+plot_layout(design = layout)
-# 
-# 
-# 
-# p3<-clean_data%>%
-#   select(date,age_group)%>%
-#   mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
-#                             city!="Toronto"~ "All other cities"
-#                           )))%>%
-#   group_by(date,location)%>%
-#   summarize(Freq=n())%>%
-#   ggplot(aes(x=date, y=Freq,fill=factor(location)
-#                 ))+
-#   geom_bar(position="fill",stat="identity",show.legend = TRUE)+
-#   ggtitle("Geographical location of survey answers")+
-#   theme_classic()+
-#   ylab("Percentage")+
-#   theme(legend.title=element_blank())+
-#   scale_x_date(date_labels = "%b-%Y")+
-#   thm1
+  thm1
 
@@
 
-It can be seen from the graph that the trends in answers by group age do not follow the overall distribution from the province.
-In particular, the 16-24 group represents in most cases at least 20% of the answers, whereas the 35-44 group is less than 10% in some cases.
+It can be seen from the graph that the proportion of answers by group age do not follow the overall distribution from the province.
 
@@
 
-The final aspect I wanted to explore was the income distribution on the responses over time.
+The final aspect I wanted to explore was the income distribution.
 In this regard the question from the survey was "What is your household annual income?".
@@
   mutate(income_ord = fct_relevel(income,
-                                  "over_110000",
+                                  "under_15000",
+                                  "15000_24999",
+                                  "25000_39999",
+                                  "40000_59999",
+                                  "60000_89999",
                                   "90000_109999",
-                                  "60000_89999",
-                                   "40000_59999",
-                                  "25000_39999",
-                                     "15000_24999",
-                                   "under_15000"))%>%
-  group_by(date,income_ord)%>%
-  summarize(Freq=n())%>%
-  ggplot(aes(x=date, y=Freq,fill=income_ord))+
-  geom_bar(position="fill",stat="identity",show.legend = TRUE)+
+                                  "over_110000"
+                                ))
+
+ggplot(p3%>% count(income_ord)%>%  # count number in each group
+         mutate(pct=n/sum(n),               # Calculate percent within each group
+                ypos = 200),  # Calculate label positions
+       aes(income_ord, n, fill=income_ord)) +
+  geom_bar(stat="identity") +
+  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
   theme_classic()+
-  ylab("Percentage")+
-  scale_y_continuous(breaks = seq(0, 1, by = 0.1))+
-  theme(legend.title = element_blank())+
-   scale_x_date(date_labels = "%b-%Y")+
   thm3
-
-p3
 
@@
 
-# Conclusions
+# Joining the municipalities data
 
-Overall, it seems that poststratification corrections would be needed for the dataset with regard to race, ethnicity, age groups
+We have obtained and cleaned the data of the municipalities from the province of Ontario website, and cleaned the dataset to obtain the city names and geographical locations.
+Further details can be found in `municipalities.qmd` in the `data_cleaning` directory.
+Next, we will join and match the geographical location from the municipalities dataset to the clean dataset we have obtained in this document so far.
+
+```{r,join-datasets}
+
+municipalities<-read.csv(here("data","municipalities_clean.csv"))
+
+municipalities$Municipal.status<-as.factor(municipalities$Municipal.status)
+
+municipalities$Geographic.area<-as.factor(municipalities$Geographic.area)
+
+municipalities$city<-as.factor(municipalities$city)
+
+clean_data<-left_join(clean_data,municipalities,by="city")
+
+```
+
+Need to explore which entries were left without a geographical region:
+
+```{r, missing-geog-area}
+
+test<-clean_data %>%
+  filter(is.na(Geographic.area))%>%
+  distinct(city)
+
+```
+
+There are 756 observations that did not get a geographical region, and they correspond to 177 unique cities.
+These will be exported to a csv file in order to manually write their geographical areas following the Association of Municipalities of Ontario divisions, and using Wikipedia to check the status of each municipality.
+
+```{r,write-file}
+#| echo: false
+
+write.csv(test,here("data","missing_municipalities.csv"),row.names = FALSE)
+
+```
+
+The code chunk above was ran once to get the names of the municipalities.
+After manually entering the geographic areas, the file was saved as `missing_municipalities_updated.csv`, and this will be the file to be used for the next steps.
+
+Note: there is one city "Kinburn", but there are two communities with such name, one in Huron County and one in Carelton County, for the time being Huron County is assigned, will check later if is better to remove it.
+
+Also one occurence of "Sydenham", which can be a ward in Kingston or a community in Frotenac. Will assign Frotenac for the time being.
+
+## Merging missing municipalities and geographic area names
+
+After manually assigning the geographical regions to the municipalities that were missing (which can be found in `missing_municipalities.csv` in the `data` directory), these will be merged as the dataset `geographic_areas.csv` (also in `data`), which contains the titles for each region (e.g., "County", "Region").
+
+```{r,merge-datasets-2}
+
+```

### 2023-01-04 data_cleaning/municipalities.qmd
Updated geographical data

---
title: "Municipalities dataset cleaning"
format: pdf
editor: visual
---

This document is about how the municipality data was obtained and the steps used to clean it.

The dataset was obtained from the website of the province of Ontario (https://www.ontario.ca/page/list-ontario-municipalities#section-3).

The file obtained is a CSV but it needed to be cleaned to remove the innecesary information found in the name of each city so it can be combined with the survey dataset.

```{r,loading}
#| echo: false
#| warning: false

library(tidyverse)
library(here)

municipalities<-read.csv(here("data","municipalities_list.csv"))

municipalities_clean <- municipalities %>%
#get only the city part (see https://stackoverflow.com/questions/66212344/extract-word-in-quotes-from-string)

  mutate(city=str_extract(Municipality,'(?<=(?<!\\()")[^"]+(?=(?!\\))")'))%>%
#keep only the name of the city
mutate(city=word(city,sep=",")) %>%
  #remove the first column as it is no longer needed
  subset(select=-c(Municipality))

```

One additional step is to remove from the dataset duplicate entries. In the dataset there are some instances that appear as upper tier division (counties, regions) that have the same name of a municipality, thus causing duplicates. The ones to remove are identified below.

```{r}

municipalities_clean[duplicated(municipalities_clean$city),]

```

R detected the "duplicate" entries because the names of the cities we are interested in sometimes appear after the entry we want to remove. After examining the original municipalities dataset (`municipalities_list.csv`), and comparing it with the list provided above, the duplicate entries in the dataset that need to be removed correspond to:

-   County of Essex
-   Township of Hamilton (it is an aggregation according to https://en.wikipedia.org/wiki/Hamilton_Township,\_Ontario)
-   County of Perth
-   County of Peterborough
-   County of Renfrew
-   Regional Municipality of Waterloo

These entries need to be removed as we are interested in municipalities within larger geographical areas, and these are aggregations that do not fit that criteria. The final step is to remove the duplicates. Because the word that identifies them as aggregations (i.e., "County", "Township") is lost in the process of getting the names of the cities to match them with the survey dataset, we will remove them by rownumber (after comparing with the `municipalities_list.csv`) dataset. The rownumbers are:

-   County of Essex: 119
-   Township of Hamilton: 155
-   County of Perth: 313
-   County of Peterborough: 317
-   County of Renfrew: 338
-   Regional Municipality of Waterloo: 420

```{r,save-file}

#| echo: false

municipalities_clean <- municipalities_clean[-c(119,155,313,317,338,420),]

#save the file in csv
write.csv(municipalities_clean,here("data","municipalities_clean.csv"),row.names = FALSE)
```

### 2023-01-04 protocol/Background_and_methods_Dec_30.qmd
Updated geographical data

@@
 | Variable             | Values                                                                                                                          |
-|:---------------------|---------------------------------------------------------------------------------------------------------------------------------|
+|:--------------|---------------------------------------------------------|
 | Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
@@
 Because our interest lies in exploring vaccination status within a governamental administrative unit that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities].
+The geographical regions were obtained from the websie of the province of Ontario [https://www.ontario.ca/page/list-ontario-municipalities#section-3](https://www.ontario.ca/page/list-ontario-municipalities#section-3)
 

### 2023-01-04 protocol/Background_and_methods_Dec_30.tex
Updated geographical data

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, frame hidden, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, sharp corners, frame hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1457}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8543}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2083}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7917}}@{}}
 \caption{\label{tbl-covariates}Selected socio-economic factors from the
@@
 single-tiers\textsuperscript{\protect\hyperlink{ref-ontario-municipalities}{22}}.
+The geographical regions were obtained from the websie of the province
+of Ontario
+\url{https://www.ontario.ca/page/list-ontario-municipalities\#section-3}
 

### 2023-01-04 data_cleaning/Data_cleaning_and_covariates_Jan_03.qmd
Clean dataset created

@@
 
-write.csv(test,here("data","missing_municipalities.csv"),row.names = FALSE)
+#write.csv(test,here("data","missing_municipalities.csv"),row.names = FALSE)
 
@@
 ```{r,merge-datasets-2}
+#load missing municipalities dataset
+
+missing_municipalities<-read.csv(here("data","missing_municipalities_updated.csv"))
+
+#combining geographical regions
+
+clean_data<-clean_data %>%
+  left_join(missing_municipalities, by = c("city")) %>%
+  mutate(Geographic.area = coalesce(Geographic.area.x,Geographic.area.y)) %>%
+  select(-c(Geographic.area.x,Geographic.area.y))
+
+clean_data$Geographic.area<-as.factor(clean_data$Geographic.area)
+
+# load the the titles for each region
+
+geographic_areas<-read.csv(here("data","geographic_areas.csv"))
+
+geographic_areas$Geographic.area<-as.factor(geographic_areas$Geographic.area)
+
+geographic_areas$Geographic.area.title <-as.factor(geographic_areas$Geographic.area.title)
+
+#merge the datasets
+
+clean_data <-left_join(clean_data,geographic_areas,by="Geographic.area")
 

### 2023-01-04 data_cleaning/Data_cleaning_and_covariates_Jan_03.qmd
Updated
need to check proportions for weights

@@
 ```
+
+# Raking
+
+The next step is to adjust the variables according to the Census data (raking).
+I will follow the steps provided by [https://sdaza.com/blog/2012/raking/](https://sdaza.com/blog/2012/raking/) https://www.r-bloggers.com/2018/12/survey-raking-an-illustration/
+
+The variables to define raking weights are:
+
+  - income
+  - age groups 
+  - geographical location
+  - race and ethnicity: check proportions here https://www.ontario.ca/document/2016-census-highlights/fact-sheet-9-ethnic-origin-and-visible-minorities
+
+```{r,define-target-list}
+
+library(weights)
+
+#get percentages
+wpct(clean_data$income)
+wpct(clean_data$age_group)
+wpct(clean_data$race)
+
+#create target list for weights
+
+age_group<-c(0.127,0.129,0.128,0.149,0.137,0.167)
+
+race<-c(0.016,0.047,0.066,)
+
+income<-c(0.075,0.116,0.154,0.195,)
+
+```

### 2023-01-05 code/raking_and_model.qmd
need to write model equation, fix model and read about random effects

---
title: "Raking and model"
format: html
editor: visual
---

This document pertains the corrections on the clean dataset created in `Data_cleaning_and_covariates_Jan_05.qmd`. There are differences in the proportions of the dataset with regard to income, age groups, race/ethnicity, and geographical location when compared to the Census data. The dataset will be corrected using _raking_ with the package `survey`.

# Raking

This section follows the steps provided by [https://sdaza.com/blog/2012/raking/](https://sdaza.com/blog/2012/raking/) https://www.r-bloggers.com/2018/12/survey-raking-an-illustration/

The variables to define raking weights are:

  - income
  - age groups 
  - geographical location
  - race and ethnicity

  The population proportions in each case have been already presented in the data cleaning document.

  The dataset has missing observations in the race (28% missing), and second_dose (26.3%) covariates. First attempt for raking and model will after removing all missing observations.

```{r,define-target-list}

library(weights)
library(survey)
library(magrittr)

clean_data<-read.csv(here("data","clean_dataset.csv"))

test<-clean_data%>%drop_na(race)

#remove municipal status as well to avoid problems

test<-test%>%
  select(-Municipal.status)

#make factors
test<-test %<>%
  mutate_at(c("income_ord","race","age_group","location","Geographic.area"),factor)

test<-test %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))

#get percentages
wpct(test$income_ord)
wpct(test$age_group)
wpct(test$race)
wpct(test$location)

#create dataframes for weights

t1<-svydesign(id=~1,data=test)

ages<-data.frame(age_group=c("16_24",
                             "25_34",
                             "35_44",
                             "45_54",
                             "55_64",
                             "65_and_over"),
                 Freq=c(0.127,
                        0.129,
                        0.128,
                        0.149,
                        0.137,
                        0.167))

races<-data.frame(race=c("arab_middle_eastern",
                         "black",
                         "east_asian_pacific_islander",
                         "indigenous",
                         "latin_american",
                         "mixed",
                         "other",
                         "south_asian",
                         "white_caucasian"),
                  Freq=c(0.016,
                         0.047,
                         0.066,
                         0.028,
                         0.015,
                         0.01,
                         0.052,
                         0.087,
                         0.678
                  ))

income_ords<-data.frame(income_ord=c("under_15000",
                                     "15000_24999",
                                     "25000_39999",
                                     "40000_59999",
                                     "60000_89999",
                                     "over_90000"),
                        Freq=c(0.057,
                               0.075,
                               0.116,
                               0.154,
                               0.195,
                               0.403
                        ))

locations<-data.frame(location=c("All other cities",
                                 "Toronto"),
                      Freq=c(0.687,
                             0.313))

t.rake<-rake(t1,
             sample=list(~age_group,
                         ~race,
                         ~income_ord,
                         ~location),
             population=list(ages,races,income_ords,locations))

model<-svyglm(first_dose_m~income_ord+age_group+race,design=t1,family = quasibinomial(), control= list(maxit=50))

summary(model)

```

Test example for raking from `survey` package

```{r,test-raking}
# data(api)
# 
# dclus1 <- svydesign(id=~dnum, weights=~pw, data=apiclus1, fpc=~fpc) #no clusters
# rclus1 <- as.svrepdesign(dclus1) #create replicate weights survey design
# 
# svymean(~api00, rclus1)
# svytotal(~enroll, rclus1)
# 
# pop.types <- data.frame(stype=c("E","H","M"), Freq=c(4421,755,1018))
# pop.schwide <- data.frame(sch.wide=c("No","Yes"), Freq=c(1072,5122))
# 
# rclus1r <- rake(rclus1, list(~stype,~sch.wide), list(pop.types, pop.schwide))
# 
# svymean(~api00, rclus1r)
# svytotal(~enroll, rclus1r)
test$first_dose<-as.factor(test$first_dose)

 ggplot(clean_data %>% count(first_dose,race) %>%    # G count number in each group
         mutate(pct=n/sum(n)),  # Calculate label positions
       aes(first_dose, n,fill=race)) +
  geom_bar(stat="identity") +
   facet_wrap(~as.factor(clean_data$Geographic.area))+
  theme_classic()
```

### 2023-01-05 data_cleaning/Data_cleaning_and_covariates_Jan_05.qmd
need to write model equation, fix model and read about random effects

--- 2023-01-04 data_cleaning/Data_cleaning_and_covariates_Jan_03.qmd
@@
 
-Below is the information provided by the 2016 Canada Census for Ontario for different visible minorities and ethnic origins.
-[Link to Census Data](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Ethnic%20origin&TABID=1&type=1).
+Because the categories for race/ethnicity in the survey did not match the categories from the Canadian Census, we used a combination of sources to obtain the corresponding percentages to match the survey categories.
+Sources are indicated below:
+
+- Fact sheet from the Provice of Ontario for Visible Minorities [link](https://www.ontario.ca/document/2016-census-highlights/fact-sheet-9-ethnic-origin-and-visible-minorities): Used to obtain percentages for Arab, Black, East Asian/Pacific Islander (adding Chinese, Korean, and Japanese percentages), Latin American, Mixed (using the percentage for "multiple visible minority"), Other (obtained by adding the Southeast Asian, Filipino, West Asian, and Minority not identified elsewhere percentages), South Asian.
+_Accessed on January 05, 2022_
+
+- Census Profile for Ontario [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Aboriginal%20peoples&TABID=1&type=1): Used to obtain percentage of Aboriginal population.
+_Accessed on January 05, 2022_
+
+- Wikipedia entry for Ontario demographics [link](https://en.wikipedia.org/wiki/Demographics_of_Ontario): To corroborate that the percentage of population reported as "European" in this website matched the percentage obtained for "White Caucasian" that was independently obtained by obtaining the difference in population proportion after subtracting the sum of Visible Minorities and Aboriginal Population percentages.
 
@@
 | Black          | 4.7%       |
-| East Asian     | 6.6%       |
-| Indigenous     | 3.9%       |
+| East Asian/Pacific Islander     | 6.6%       |
+| Indigenous     | 2.8%       |
 | Latin American | 1.5%       |
 | Mixed          | 1.0%       |
-| Other          | 0.7%       |
+| Other          | 5.2%       |
 | South Asian    | 8.7%       |
+| White Caucasian| 67.8%      |
+| **Total**      | 99.9%      |
+: Reference Data from Race/Ethnicity
 
-: Data from the 2016 Census for Ontario
-
-In the table above, the following data were obtained from the "Visible Minorities" from the Census:
-
--   Arab
--   Black
--   South Asian
--   East Asian (computed by adding the Chinese, Korean and Japanese entries)
--   Mixed ("multiple visible minorities" entry in the census)
--   Other (from the "visible minority n.i.e" entry in the census)
-
-The following data were obtained from the "ethnic origin" data from the census: - Indigenous - European Origins
-
-However, according to the census "ethnicity" and "visible minority" seem to have different values.
-For example, in the Visible Minority entry of the census Latin American is 1.5%, but in the "Ethnic Origin" data the value is 2.4%.
-
-Moreover, I would need to adjust the data for Pacific Islanders, as the survey included East Asian and Pacific Islanders in the same question but they are categorized differently in the census.
-
-Lastly, the survey had "white caucasian" as one of the options but that category is not present in the census data. The most approximate entry on the census is "European origins" (61.6%).
-
-**Overall, we weighted the data because it is not representative of the population distribution in Ontario**.
-
-## Geographical location
-
-Next, I was interested in analyzing the geographical distribution of the participants over time, to see the percentage of answers from Toronto.
-This is important to see where the majority of the answers came from (is there such thing as geographical weighting?).
-The same will be done with the age groups of the participants, as this is data that is reported in the census.
-
-```{r,plot-location}
-#| echo: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 4
-#| fig-cap: Percentage of survey answers by city
-
-p1<-clean_data%>%
-  select(city)%>%
-  mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
-                            city!="Toronto"~ "All other cities"
-                          )))
-
-p1$location<-as.factor(p1$location)
-
-ggplot(p1 %>% count(location) %>%    # then count number in each group
-         mutate(pct=n/sum(n),               # Calculate percent within each group
-                ypos = 200),  # Calculate label positions
-       aes(location, n, fill=location)) +
-  geom_bar(stat="identity") +
-  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
-  theme_classic()+
-  thm2
-
-```
-
-The proportion of responses from Toronto is 30.9% of the total.
-The Census data indicates that Toronto has around 38% of the total population province (5,433,590/14,223,942).
-This would mean that a correction would be needed also for the proportion of answers.
-
-Additionally, the data needs to be cleaned as some of the responses do not have an assigned city, and the entry only reads "None". These entries will be removed as they lack geographical information.
-
-```{r, remove-entries}
-
-clean_data<-clean_data%>%
-  filter(city!="None")
-
-```
+Visible Minorities: 29.3%
+Not a Visible Minority: 70.7% (Aboriginal 2.8%, White Caucasian 67.8%)
 
@@
 
-The final aspect I wanted to explore was the income distribution.
-In this regard the question from the survey was "What is your household annual income?".
-To compare the results of this answer, I used the "Household total income groups in 2015 for private households" from the Census data.
-The information from the census appears next.
+The question from the survey was "What is your household annual income?".
+To compare the results of this answer, we used the "Household total income groups in 2015 for private households" from the Census data (available in the Census Data for Ontario website [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Income&TABID=1&type=1)).
 
@@
 | 60,000 - 89,999              | 19.5%      |
-| \>90,000                     | 40.4%      |
+| \>90,000                     | 40.3%      |
 
@@
 
-One thing to keep in mind in this case is that the brackets for income in the census data are different than the brackets used in the survey.
-The census does CAD 4,999 brackets (e.g., CAD 5,000- CAD 9,9999) up to CAD 49,999 and then it does CAD 9,999 brackets up to CAD 99,999 but after that, the brackets increase to CAD 24,999, and therefore, one cannot obtain percentages for the 90,000-109,999 and \>110,000 brackets from the survey.
+The brackets for income in the census data are different than the brackets used in the survey.
+The census does CAD 4,999 brackets (e.g., CAD 5,000- CAD 9,9999) up to CAD 49,999, followed by CAD 9,999 brackets up to CAD 99,999.
+After that, the brackets increase to CAD 24,999, and therefore, one cannot obtain percentages for the 90,000-109,999 and \>110,000 brackets from the survey.
+
+Therefore, an additional category for income will be created in the dataset so it matches the information from the Census.
 
@@
 
+clean_data<-clean_data%>%
+    mutate(income_ord = case_when(
+                                  income %in% "90000_109999" ~ "over_90000",
+                                  income %in%"over_110000" ~ "over_90000",
+                                  TRUE ~ as.character(income)
+                                ))%>%
+  mutate(income_ord=as.factor(income_ord))
+
 p3<-clean_data%>%
-  select(date,income)%>%
-  mutate(income_ord = fct_relevel(income,
+  mutate(income_ord = fct_relevel(income_ord,
                                   "under_15000",
@@
                                   "60000_89999",
-                                  "90000_109999",
-                                  "over_110000"
+                                  "over_90000"
                                 ))
@@
   theme_classic()+
-  thm3
+  thm2
 
@@
 
-# Joining the municipalities data
+## Geographical location
+
+Next, we are interested in analyzing the geographical distribution of the answers the participants within the geographical areas defined by the Association of Municipalities from Ontario.
+There are two parts to the geographical analysis:
+
+1) Add to the dataset the data for the municipalities so each city has the corresponding region it belongs to.
+1) Determine percentage of answers that correspond to Toronto.
+This is important because although the Census regions do not match the exact geographical boundaries of the municipalities and thus a direct correction for this cannot be easily obtained, Toronto is the largest urban area in the province and it is important to identify if corrections are needed with regard to the proportions from this place in the survey.
+
+### Joining the municipalities data
 
@@
 Further details can be found in `municipalities.qmd` in the `data_cleaning` directory.
-Next, we will join and match the geographical location from the municipalities dataset to the clean dataset we have obtained in this document so far.
+Next, we will join and match the geographical location from the municipalities dataset to the clean dataset we have obtained in this document so far after removing the entries that do not have a corresponding geographical location.
 
 ```{r,join-datasets}
+
+clean_data<-clean_data%>%
+  filter(city!="None")
 
@@
 
-## Merging missing municipalities and geographic area names
+### Merging missing municipalities and geographic area names
 
@@
 ```
+
+## Proportion of answers within Toronto
+
+The final step is of the geographical analysis is to determine the proportion of answers from Toronto.
+
+```{r,plot-location}
+#| echo: false
+#| warning: false
+#| fig-width: 8
+#| fig-height: 4
+#| fig-cap: Percentage of survey answers by city
+
+clean_data<-clean_data%>%
+  mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
+                            city!="Toronto"~ "All other cities"
+                          )))%>%
+  mutate(location=as.factor(location))
+
+ggplot(clean_data %>% count(location) %>%    # then count number in each group
+         mutate(pct=n/sum(n),               # Calculate percent within each group
+                ypos = 200),  # Calculate label positions
+       aes(location, n, fill=location)) +
+  geom_bar(stat="identity") +
+  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
+  theme_classic()+
+  thm1
+
+```
+
+The proportion of responses from the city of Toronto is 31.3% of the total in the clean dataset.
+From the (Census data)[https://www.toronto.ca/wp-content/uploads/2022/02/92e3-City-Planning-2021-Census-Backgrounder-Population-Dwellings-Backgrounder.pdf], the city of Toronto as of 2021 has a population of 2,794,356, corresponding to 19.6% of the total population of the Province.
+
+At this point, the data has been cleaned and prepared for formal analysis.
+
+The clean dataset will be saved as a *.csv file.
+
+```{r,save-clean-dataset}
+#| echo: false
+
+#write.csv(clean_data,here("data","clean_dataset.csv"),row.names = FALSE)
+```

### 2023-01-06 data_cleaning/raking_and_model.qmd
Updated raking and models

--- 2023-01-05 code/raking_and_model.qmd
@@
 There are differences in the proportions of the dataset with regard to income, age groups, race/ethnicity, and geographical location when compared to the Census data.
-The dataset will be corrected using _raking_ with the package `survey`.
+The dataset will be corrected using *raking* with the package `survey`.
 
@@
 
-This section follows the steps provided by [https://sdaza.com/blog/2012/raking/](https://sdaza.com/blog/2012/raking/) https://www.r-bloggers.com/2018/12/survey-raking-an-illustration/
+This section follows the steps provided by <https://sdaza.com/blog/2012/raking/> https://www.r-bloggers.com/2018/12/survey-raking-an-illustration/
 
@@
 library(weights)
+library(here)
 library(survey)
-library(magrittr)
+library(tidyverse)
 
@@
 
-t1<-svydesign(id=~1,data=test)
+# t1<-svydesign(id=~1,data=test)
+# 
+# ages<-data.frame(age_group=c("16_24",
+#                              "25_34",
+#                              "35_44",
+#                              "45_54",
+#                              "55_64",
+#                              "65_and_over"),
+#                  Freq=c(0.127,
+#                         0.129,
+#                         0.128,
+#                         0.149,
+#                         0.137,
+#                         0.167))
+# 
+# 
+# 
+# races<-data.frame(race=c("arab_middle_eastern",
+#                          "black",
+#                          "east_asian_pacific_islander",
+#                          "indigenous",
+#                          "latin_american",
+#                          "mixed",
+#                          "other",
+#                          "south_asian",
+#                          "white_caucasian"),
+#                   Freq=c(0.016,
+#                          0.047,
+#                          0.066,
+#                          0.028,
+#                          0.015,
+#                          0.01,
+#                          0.052,
+#                          0.087,
+#                          0.678
+#                   ))
+# 
+# income_ords<-data.frame(income_ord=c("under_15000",
+#                                      "15000_24999",
+#                                      "25000_39999",
+#                                      "40000_59999",
+#                                      "60000_89999",
+#                                      "over_90000"),
+#                         Freq=c(0.057,
+#                                0.075,
+#                                0.116,
+#                                0.154,
+#                                0.195,
+#                                0.403
+#                         ))
+# 
+# locations<-data.frame(location=c("All other cities",
+#                                  "Toronto"),
+#                       Freq=c(0.687,
+#                              0.313))
+
+########### use population totals ################
+########## this is the approach used in the survey package examples #######
 
@@
                              "65_and_over"),
-                 Freq=c(0.127,
-                        0.129,
-                        0.128,
-                        0.149,
-                        0.137,
-                        0.167))
+                 Freq=c(1707959,
+                        1734856,
+                        1721407,
+                        2003826,
+                        1842444,
+                        2245898))
 
@@
                          "white_caucasian"),
-                  Freq=c(0.016,
-                         0.047,
-                         0.066,
-                         0.028,
-                         0.015,
-                         0.01,
-                         0.052,
-                         0.087,
-                         0.678
-                  ))
+                  Freq=c(212782,
+                         638346,
+                         886592,
+                         376558,
+                         197020,
+                         130033,
+                         705333,
+                         1166361,
+                         9118079 ))
 
@@
                                      "over_90000"),
-                        Freq=c(0.057,
-                               0.075,
-                               0.116,
-                               0.154,
-                               0.195,
-                               0.403
+                        Freq=c(294643,
+                               387688,
+                               599624,
+                               796052,
+                               1007988,
+                               2083176
+
                                   ))
@@
                                  "Toronto"),
-                      Freq=c(0.687,
-                             0.313))
+                      Freq=c(10812589,
+                              2635905
+                                    ))
 
@@
 
-summary(model)
+model.rake<-svyglm(first_dose_m~+Geographic.area*race,design=t.rake,family = quasibinomial(), control= list(maxit=50))
+
+summary(model.rake)
+
+model.test<-glmer(first_dose_m~Geographic.area*race+income_ord+age_group+Geographic.area,data=model.frame(t.rake),family=binomial)
 

### 2023-01-06 code/raking_and_model.R

@@ -1,0 +1,181 @@
+#### this script uses the clean dataset to: 1) correct the observations in the survey
+#### response using raking and 2) compute a model of logistic regression
+
+#load packages
+
+library(weights)
+library(here)
+library(survey)
+library(tidyverse)
+library(lme4)
+library(gtsummary)
+
+data<-read.csv(here("data","clean_dataset.csv"))
+
+# remove rows with missing observations for race
+
+clean_data<-data%>%drop_na(race)
+
+#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
+
+clean_data<-clean_data%>%
+  select(-Municipal.status)
+
+#make variables of interest factors
+clean_data<-clean_data %<>%
+  mutate_at(c("income_ord","race","age_group","location","Geographic.area","Geographic.area.title"),factor)
+
+# recode the response of interest for categorical analysis
+clean_data<-clean_data %>%
+  mutate(first_dose_m=case_when(
+    first_dose=="yes"~1,
+    first_dose=="no"~0
+  ))
+
+
+# get percentages to see how different the data is from the Census (details of the Census values are in
+# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")
+
+wpct(clean_data$income_ord)
+wpct(clean_data$age_group)
+wpct(clean_data$race)
+wpct(clean_data$location)
+
+
+
+################ RAKING ###################################
+## I had originally used the percentages from the census, but the "population values"
+## that need to be provided in the examples of the survey package are all based on total counts.
+## This would make sense as to estimate the sampling probability we input from the dataset are counts as well.
+## I transformed all the percentages using the total population for Ontario (from the 2016 census) and 
+## household counts (also from the 2016 Census).
+
+## Each of these dataframes provide the population totals for the different variables and their categories
+
+ages<-data.frame(age_group=c("16_24",
+                             "25_34",
+                             "35_44",
+                             "45_54",
+                             "55_64",
+                             "65_and_over"),
+                 Freq=c(1707959,
+                        1734856,
+                        1721407,
+                        2003826,
+                        1842444,
+                        2245898))
+
+
+
+races<-data.frame(race=c("arab_middle_eastern",
+                         "black",
+                         "east_asian_pacific_islander",
+                         "indigenous",
+                         "latin_american",
+                         "mixed",
+                         "other",
+                         "south_asian",
+                         "white_caucasian"),
+                  Freq=c(212782,
+                         638346,
+                         886592,
+                         376558,
+                         197020,
[…]

### 2023-01-08 data_cleaning/Data_cleaning_and_covariates_Jan_05.qmd
Desgracia

@@
 Next, we are interested in analyzing the geographical distribution of the answers the participants within the geographical areas defined by the Association of Municipalities from Ontario.
+
+**Update: the populations by census division/city were found on Jan 8, 2023 (from https://www12.statcan.gc.ca/census-recensement/2021/dp-pd/prof/index.cfm?Lang=E).
+The correction will be done using these numbers instead.
+The numbers can be found in `census_population_by_division.csv` in `data` directory.**
+
 There are two parts to the geographical analysis:
@@
 
+**This section needs to be modified as the correction has been done by geographica area on Jan 8, 2023**
+
 The final step is of the geographical analysis is to determine the proportion of answers from Toronto.

### 2023-01-08 code/raking_and_model.R

@@ -6,7 +6,7 @@
 library(weights)
 library(here)
 library(survey)
-library(tidyverse)
+library(magrittr)
 library(lme4)
 library(gtsummary)
 
@@ -107,6 +107,108 @@ locations<-data.frame(location=c("All other cities",
                              2635905
                       ))
 
+geographic_areas <- data.frame(Geographic.area=c("Algoma",
+                                                "Brant",
+                                                #"Brantford",
+                                                "Bruce",
+                                                "Chatham-Kent",
+                                                "Cochrane",
+                                                "Dufferin",
+                                                "Durham",
+                                                "Elgin",
+                                                "Essex",
+                                                "Frontenac",
+                                                "Greater Sudbury",
+                                                "Grey",
+                                                "Haldimand",
+                                                #"Haliburton",
+                                                "Halton",
+                                                "Hamilton",
+                                                "Hastings",
+                                                "Huron",
+                                                "Kawartha Lakes",
+                                                "Kenora",
+                                                "Lambton",
+                                                "Lanark",
+                                                "Leeds and Grenville",
+                                                "Lennox and Addington",
+                                                "Manitoulin",
+                                                "Middlesex",
+                                                "Muskoka",
+                                                "Niagara",
+                                                "Nipissing",
+                                                "Norfolk",
+                                                "Northumberland",
+                                                "Ottawa",
+                                                "Oxford",
+                                                "Parry Sound",
+                                                "Peel",
+                                                "Perth",
+                                                "Peterborough",
+                                                "Prescott and Russell",
+                                                "Prince Edward",
+                                                "Rainy River",
+                                                "Renfrew",
+                                                "Simcoe",
+                                                "Stormont, Dundas and Glengarry",
+                                                "Sudbury",
+                                                "Thunder Bay",
+                                                "Timiskaming",
+                                                "Toronto",
+                                                "Waterloo",
[…]

### 2023-01-09 code/raking_and_model.R
Updated

@@ -7,6 +7,7 @@ library(weights)
 library(here)
 library(survey)
 library(magrittr)
+library(tidyverse)
 library(lme4)
 library(gtsummary)
 
@@ -333,15 +334,40 @@ model3 <- mix(first_dose_m ~ race + income_ord+age_group+ (1|Geographic_area), d
 
 ### differences by race on percentage of yes and no
 
-clean_data %>%
-  filter(race=="latin_american",
-         #Geographic.area=="Toronto"
+test<-clean_data %>%
+  subset(
+         Geographic.area=="Toronto"
          )%>%
+  group_by(age_group,race)%>%
+ count()
+
+test
+
+test1<-clean_data %>%
+  subset(
+    Geographic.area=="Toronto"
+  )%>%
+  filter(first_dose=="yes")%>%
+  group_by(age_group,race)%>%
+summarise(val=n(),.groups="drop")
+
+test2<-left_join(test,test1)
+
+test2%>%
+  mutate(perc=val/n)%>%
+  ggplot(aes(x=age_group,y=perc,fill=race))+
+  geom_bar(stat="identity",position="stack")
+
+
+
+  mutate(total=count(first_dose=="yes"))
 ggplot(mapping = aes(x = as.factor(first_dose),
-                     y = after_stat(count/sum(count)))) +
+                     y = after_stat(count/sum(count)))
+       ) +
   geom_bar() +
  #xlab(paste0(unique(race)))+
-  scale_y_continuous(labels = scales::percent)
+  scale_y_continuous(labels = scales::percent)+
+  facet_wrap(~age_group)
 
 
 clean_data %>%

### 2023-01-11 data_cleaning/Data_cleaning_and_covariates_Jan_09.qmd
Added data for the health region approach

--- 2023-01-08 data_cleaning/Data_cleaning_and_covariates_Jan_05.qmd
@@
 
-We have obtained and cleaned the data of the municipalities from the province of Ontario website, and cleaned the dataset to obtain the city names and geographical locations.
+We have obtained and cleaned the data of the municipalities from the province of Ontario [website](https://www.ontario.ca/page/list-ontario-municipalities#section-3), and cleaned the dataset to obtain the city names and geographical locations.
 Further details can be found in `municipalities.qmd` in the `data_cleaning` directory.
-Next, we will join and match the geographical location from the municipalities dataset to the clean dataset we have obtained in this document so far after removing the entries that do not have a corresponding geographical location.
+We will join and match the geographical location from the municipalities dataset to the clean dataset we have obtained in this document so far after removing the entries that do not have a corresponding geographical location.
 
@@
 
-There are 756 observations that did not get a geographical region, and they correspond to 177 unique cities.
+There are 2744 observations that did not get a geographical region, and they correspond to 187 unique cities.
 These will be exported to a csv file in order to manually write their geographical areas following the Association of Municipalities of Ontario divisions, and using Wikipedia to check the status of each municipality.
@@
 
-geographic_areas$Geographic.area<-as.factor(geographic_areas$Geographic.area)
+geographic_areas$Geographic_area<-as.factor(geographic_areas$Geographic_area)
 
-geographic_areas$Geographic.area.title <-as.factor(geographic_areas$Geographic.area.title)
+geographic_areas$Geographic_area_title <-as.factor(geographic_areas$Geographic_area_title)
 
@@
 
-clean_data <-left_join(clean_data,geographic_areas,by="Geographic.area")
+clean_data <-left_join(clean_data,geographic_areas,by=c("Geographic.area"="Geographic_area")) %>%
+   subset(select=-c(Full_title.y))
 
@@
 
-**This section needs to be modified as the correction has been done by geographica area on Jan 8, 2023**
+**This section needs to be modified as the correction has been done by geographical area on Jan 8, 2023**
 
@@
 ```
+
+# Health Regions approach
+
+Following on a suggestion by BN on Jan 09, the idea is to use broader regions for geographical analysis.
+The Health Regions of Ontario will be used.
+However, these Health Regions do not match the census divisions.
+I have not been able to find a dataset from Health Ontario that lists each municipality and its corresponding Health Region.
+I will use a multi-stage approach to incorporate the information into the dataset:
+
+- Get dataset from Paul Allen regarding long-term care homes in Ontario (https://paulallen.ca/consolidated-dataset-of-ltc-homes-in-ontario/).
+This dataset has communities and the Local Health Integration Network (LHINs) where long-term care homes were located.
+I will keep the community and LHIN information.
+
+- Add column for Health Region based on LHIN: The Health Regions substituted LHINs, a list of their correspondence can be found here: https://www.ontariohealth.ca/about-us/our-people.
+I will add this information.
+
+- Merge the dataset and check if there are missing municipalities. Manually add missing municipalities.
+
+## Steps
+
+The dataset from Paul Allen was downloaded and saved as `Consolidated_LTC_dataset.csv` in the `data` directory.
+One thing to note is that there is a missing observation (coded as "Not provided") for one of the entries (city of Napanee, located in the Lennox and Addington County).
+Under the LHIN divisions, Napanee was in the South East LHIN.
+Also, there is one entry that says "244 Main Street East" as the community entry but it should be "Stayner" (the address provided belongs to Stayner.) The information is fixed before adding the Health Region.
+
+```{r,get-health-data}
+
+#load data
+ltc_data <- read.csv(here("data","Consolidated_LTC_dataset.csv"))
+
+## checking missing observations
+
+sum(ltc_data$X_LHIN=="Not provided")
+
+ltc_data <- ltc_data %>%
+  mutate(X_LHIN=replace(X_LHIN,X_LHIN=="Not provided","South East"))
+
+ltc_data <- ltc_data %>%
+  mutate(COMMUNITY=replace(COMMUNITY,COMMUNITY==" 244 Main Street East","Stayner"))
+
+#keep relevant columns and add Health region information
+ltc_data <-ltc_data %>%
+ select(c("COMMUNITY","X_LHIN"))%>%
+  rename(city=COMMUNITY,
+         LHIN=X_LHIN)%>%
+  mutate(Health_Region=
+    case_when(LHIN=="Central"~ "Central",
+              LHIN=="Central West"~"Central",
+              LHIN=="Mississauga Halton"~"Central",
+              LHIN=="North Simcoe Muskoka"~"Central",
+              LHIN=="Central East"~"East",
+              LHIN=="South East"~"East",
+              LHIN=="Champlain"~"East",
+              LHIN=="North East"~"North East",
+              LHIN=="North West"~"North West",
+              LHIN=="Toronto Central"~"Toronto",
+              LHIN=="South West"~"West",
+              LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)"~"West",
+              LHIN=="Waterloo Wellington"~"West",
+              LHIN=="Erie St. Clair"~"West"
+    )
+  )%>%
+  distinct(city,.keep_all = TRUE)
+
+```
+
+Now, combine datasets and write csv file for those cities that were not assigned a Health Region.
+
+```{r,combine}
+
+clean_data<-left_join(clean_data,ltc_data,by="city")
+
+ missing_health_regions<-clean_data %>% filter(is.na(LHIN)) %>%
+   distinct(city,.keep_all = TRUE)%>%
+   select(city,Geographic.area,Geographic_area_title,LHIN,Health_Region)
+
+#write.csv(missing_health_regions,here("data","missing_health_regions.csv"),row.names = FALSE)
+
+```
+
+To complete the missing LHIN information, the correspoding LHINs were obtained from the LHIN websites:
+
+https://www.southeasthealthline.ca/aboutourregion.aspx?region=RuralFrontenacLennoxAddington
+
+https://www.nsmhealthline.ca/aboutourregion.aspx
+
+https://www.champlainhealthline.ca/aboutourregion.aspx?region=StormontDundasGlen
+
+https://www.wwhealthline.ca/aboutourregion.aspx?region=KitchenerWaterlooWellesleyWilmotWoolwich
+
+Some cities did not belong entirely to a LHIN.
+For example, Etobicoke was divided between the Central, Central East, and Toronto Central LHINs.
+We chose in these cases the LHIN that covered the larger geographical region of each city (Toronto Central LHIN in the case of Etobicoke).
+We now assign LHINs to the dataset and Health Regions.
+The dataset with the information of the missing LHINs is called `missing_health_regions_updated.csv`
+
+```{r,merge-health-regions}
+
+mhru<-read.csv(here("data","missing_health_regions_updated.csv"))
+
+```

### 2023-01-11 data_cleaning/municipalities.qmd
Added data for the health region approach

@@
 
-The file obtained is a CSV but it needed to be cleaned to remove the innecesary information found in the name of each city so it can be combined with the survey dataset.
+The file obtained is a CSV but it needed to be cleaned to remove the unnecessary information found in the name of each city so it can be combined with the survey dataset.First, we only keep the name of the city and the municipal region they belong to, and later remove duplicate entries.
 
-```{r,loading}
+```{r,loading-first-clean}
 #| echo: false
@@
 municipalities_clean <- municipalities %>%
-#get only the city part (see https://stackoverflow.com/questions/66212344/extract-word-in-quotes-from-string)
+#get only the city/county part (see https://stackoverflow.com/questions/66212344/extract-word-in-quotes-from-string)
+  mutate(Full_title=str_extract(Municipality,'(?<=(?<!\\()")[^"]+(?=(?!\\))")'))
 
-  mutate(city=str_extract(Municipality,'(?<=(?<!\\()")[^"]+(?=(?!\\))")'))%>%
-#keep only the name of the city
-mutate(city=word(city,sep=",")) %>%
+```
+
+The second step of cleaning this dataset is the fact that it contains the names of the regions (e.g., the Regional Municipality of Peel appears listed along the geographical area of Peel, when they are the same thing).
+These entries need to be removed as in some cases some towns have the same name as a geographical region (e.g., Durham in Grey County, having the same name as the Regional Municipality of Durham).
+To do this, we will use the geographical regions listed on the [website](https://www.amo.on.ca/about-us/municipal-101/ontario-municipalities) of the Association of Municipalities of Ontario (AMO).
+As a clarification, because the "Districts" listed in this website do not serve any municipal government purpose, except on the case of the District Municipality of Muskoka only this last one needs to be removed from the municipality list.
+Lastly, on the AMO website some municipalities have their most recent status (e.g., Brant, which appears as city, but still appears listed as "County") in the website of the province of Ontario.
+We will load the AMO dataset (called `geographic_areas.csv`) and use to remove those entries from `municipalities_clean` that match the entries.
+Later, we will manually remove entries that do not have the same name but correspond to the same region.
+
+```{r,clean-second-step}
+
+# load AMO dataset
+geographic_areas <- read.csv(here("data","geographic_areas.csv"))
+
+#remove entries with the same name in the Ontario dataset
+
+municipalities_clean <- anti_join(municipalities_clean,geographic_areas,by="Full_title")
+
+# manually remove entries
+
+exclude <-tibble(Full_title=c("Brant, County of", 
+                                       "Haldimand County", 
+                                       "Norfolk County",
+                                       "Prince Edward, County of",
+                                       "Chatham-Kent, Municipality of"))
+
+ #remove entries
+
+municipalities_clean <-anti_join(municipalities_clean,exclude, by="Full_title")%>%
   #remove the first column as it is no longer needed
@@
 
-One additional step is to remove from the dataset duplicate entries.
-In the dataset there are some instances that appear as upper tier division (counties, regions) that have the same name of a municipality, thus causing duplicates.
-The ones to remove are identified below.
+One additional step is to create a column that only has the city name (as the Fields Dataset only has the city name on it without the full name of "Township of", or "City of").
+After that, the dataset is saved as a csv file.
 
@@
 
-municipalities_clean[duplicated(municipalities_clean$city),]
-
-```
-
-R detected the "duplicate" entries because the names of the cities we are interested in sometimes appear after the entry we want to remove.
-After examining the original municipalities dataset (`municipalities_list.csv`), and comparing it with the list provided above, the duplicate entries in the dataset that need to be removed correspond to:
-
--   County of Essex
--   Township of Hamilton (it is an aggregation according to https://en.wikipedia.org/wiki/Hamilton_Township,\_Ontario)
--   County of Perth
--   County of Peterborough
--   County of Renfrew
--   Regional Municipality of Waterloo
-
-These entries need to be removed as we are interested in municipalities within larger geographical areas, and these are aggregations that do not fit that criteria.
-The final step is to remove the duplicates.
-Because the word that identifies them as aggregations (i.e., "County", "Township") is lost in the process of getting the names of the cities to match them with the survey dataset, we will remove them by rownumber (after comparing with the `municipalities_list.csv`) dataset.
-The rownumbers are:
-
--   County of Essex: 119
--   Township of Hamilton: 155
--   County of Perth: 313
--   County of Peterborough: 317
--   County of Renfrew: 338
--   Regional Municipality of Waterloo: 420
-
-```{r,save-file}
-
-#| echo: false
-
-municipalities_clean <- municipalities_clean[-c(119,155,313,317,338,420),]
+municipalities_clean <- municipalities_clean %>%
+  mutate(city=sub(",.*","",Full_title))
 
@@
 write.csv(municipalities_clean,here("data","municipalities_clean.csv"),row.names = FALSE)
+
 ```

### 2023-01-11 data_cleaning/Data_cleaning_and_covariates_Jan_09.qmd
No missing obs. fixed model works

@@
 
-At this point, the data has been cleaned and prepared for formal analysis.
-
-The clean dataset will be saved as a *.csv file.
-
-```{r,save-clean-dataset}
-#| echo: false
-
-#write.csv(clean_data,here("data","clean_dataset.csv"),row.names = FALSE)
-```
-
 # Health Regions approach
@@
 
+#load health regions with missing entries fixed
+
 mhru<-read.csv(here("data","missing_health_regions_updated.csv"))
 
+#rename columns
+mhru <- mhru %>%
+  rename(Geographic_area_title=Geographic.area.title)
+
+clean_data<-clean_data %>%
+  left_join(mhru, by = c("city")) %>%
+   mutate(LHIN = coalesce(LHIN.x,LHIN.y)) %>%
+  select(-c(LHIN.x,LHIN.y,Full_title.x,Geographic.area.y,Geographic_area_title.y))%>%
+  mutate(Health_Region=
+    case_when(LHIN=="Central"~ "Central",
+              LHIN=="Central West"~"Central",
+              LHIN=="Mississauga Halton"~"Central",
+              LHIN=="North Simcoe Muskoka"~"Central",
+              LHIN=="Central East"~"East",
+              LHIN=="South East"~"East",
+              LHIN=="Champlain"~"East",
+              LHIN=="North East"~"North East",
+              LHIN=="North West"~"North West",
+              LHIN=="Toronto Central"~"Toronto",
+              LHIN=="South West"~"West",
+              LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)"~"West",
+              LHIN=="Waterloo Wellington"~"West",
+              LHIN=="Erie St. Clair"~"West"
+    ))%>%
+  rename(Geographic_area=Geographic.area.x,Geographic_area_title=Geographic_area_title.x)
+
 ```
+
+At this point, the data has been cleaned and prepared for formal analysis.
+
+The clean dataset will be saved as a *.csv file.
+
+```{r,save-clean-dataset}
+#| echo: false
+
+write.csv(clean_data,here("data","clean_dataset.csv"),row.names = FALSE)
+```

### 2023-01-11 code/raking_and_model.R

@@ -24,8 +24,7 @@ clean_data<-clean_data%>%
 
 #make variables of interest factors
 clean_data<-clean_data %<>%
-  mutate_at(c("income_ord","race","age_group","location","Geographic.area","Geographic.area.title"),factor)
-
+  mutate_at(c("income_ord","race","age_group","location","Geographic_area","Geographic_area_title"),factor)
 # recode the response of interest for categorical analysis
 clean_data<-clean_data %>%
   mutate(first_dose_m=case_when(
@@ -108,9 +107,9 @@ locations<-data.frame(location=c("All other cities",
                              2635905
                       ))
 
-geographic_areas <- data.frame(Geographic.area=c("Algoma",
+geographic_areas <- data.frame(Geographic_area=c("Algoma",
                                                 "Brant",
-                                                #"Brantford",
+                                                "Brantford",
                                                 "Bruce",
                                                 "Chatham-Kent",
                                                 "Cochrane",
@@ -161,7 +160,7 @@ geographic_areas <- data.frame(Geographic.area=c("Algoma",
                                                 "York"),
                               Freq=c(113777,
                                      39474,
-                                     #104688,
+                                     104688,
                                      73396,
                                      103988,
                                      77963,
@@ -211,6 +210,25 @@ geographic_areas <- data.frame(Geographic.area=c("Algoma",
                                      241026,
                                      1173334))
 
+
+# corrections for health regions
+
+health_regions<-data.frame(Health_Region=c(
+  "North West",
+  "North East",
+  "West",
+  "East",
+  "Central",
+  "Toronto"),
+  Freq=c(232299,
+    557000,
+    4000000,
+    3700000,
+    5000000,
+    1400000
+  )
+)
+
 ### Raking ####
 
 ## First, provide a survey design object. This is done using the dataset. The syntax
@@ -231,7 +249,7 @@ a1_rake<-rake(a1,
              sample=list(~age_group,
                          ~race,
                          ~income_ord,
-                         ~Geographic.area),
+                         ~Geographic_area),
              population=list(ages,
                              races,
                              income_ords,
@@ -242,9 +260,23 @@ a1_rake%>%
                  percent = "row", 
[…]

### 2023-01-13 protocol/Background_and_methods_Jan_12.qmd
updated protocol

--- 2022-12-30 protocol/Background_and_methods_Dec_30.qmd
@@
 ---
-title: "Covid Fields Data Protocol"
+title: "Covid-19 Fields Survey Data Protocol"
 author:
@@
         postal-code: H3N 1X9
+date: "`r Sys.Date()`"
 format:
@@
 Studies that analyze geographical variations in vaccine uptake can help inform public health decision-makers to design policies to that are aimed at addressing vaccination disparities.
-In this regad, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governamental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].
+In this regard, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governmental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].
 
 In Canada, studies that have used a spatial approach to analyze vaccine uptake have shown disparities in vaccination rates across low and high income neighborhoods in the city of Toronto [@choi2021], among adolescents from deprived neighborhoods in the city of Montreal [@mckinnon2021], and highlighted disparities in vaccination status depending on age, income, and ethnic origin in all of the Canadian provinces[@guay2022].
-However, to the best of our knowledge, there are no studies that have analyzed vaccination status within a province at the district/rural municipality level, which can be useful to identify inequalities over these geographical areas, thus providing a dissagregated view that can help understand the barriers for vaccine delivery in the case of visible minorities, which have been disproportionately impacted in the pandemic [@hussain2022].
+However, to the best of our knowledge, there are no studies that have analyzed vaccination status within a province to identify inequalities that may exist within these geographical areas.
+A dissagregated view of variability within a province can help understand the barriers for vaccine delivery in the case of visible minorities, which have been disproportionately impacted in the pandemic [@hussain2022].
 
@@
 
-This study will examine self-reported COVID-19 vaccination status in the province of Ontario in order to determine the influence in vaccination status due to socio-economic (e.g., ethnic origin, age, income) and geographical factors (at the municipal level).
+This study will examine self-reported COVID-19 vaccination status in the province of Ontario in order to determine the influence in vaccination status due to socio-economic (e.g., ethnic origin, age, income) and geographical factors (at level of the Health Regions of Ontario).
 
@@
 | Variable             | Values                                                                                                                          |
-|:---------------------|---------------------------------------------------------------------------------------------------------------------------------|
+|:----------------|-------------------------------------------------------|
 | Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
@@
 | Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
-| Employment status    | yes, no                                                                                                                         |
+                                                                                                                      |
 
@@
 
-We selected two of the questions that were asked in the survey regarding vaccination status:
+From the survey, we selected the question regarding vaccination status:
 
@@
 
--   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no"
+<!-- -   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no" -->
 
@@
 Following a preliminary analysis to identify the missing rates across the different answers within each entry, it was identified that many of the answers had high missing rates (\>80%) (note that the graph will be in the Appendix).
-Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variables.
+Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variable.
 
@@
 
-## Corrections
-
-Differences between the clean dataset and the 2016 Census data for Ontario regarding proportions of age groups, income, and ethnicity/race were identified.
-Therefore, the proportions of each of these variables were corrected using an iterative proportional fitting procedure (also known as *raking*) [@deming1940] in R using the `survey` package.
-Proportions for the correction were obtained from the 2016 Census Data for Ontario.
-Because the categories provided by the survey in some cases (e.g., race/ethnicity categories) did not match the categories from the Census, we aggregated some of them to obtain an approximation to the categories in the Census.
-The aggregation is described in detail in the Appendix.
-
-## Geographical location {#sec-geographical-location}
+### Geographical location {#sec-geographical-location}
 
@@
 This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset.
-Because our interest lies in exploring vaccination status within a governamental administrative unit that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities].
+Because our interest was to analyze the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset.
+Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN), which were the geographical divisions for health in Ontario before the adoption of the Health Regions, to match each city with its corresponding LHIN.
+Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset.
+Finally, we assigned the corresponding Health Regio to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).
+
+## Corrections
+
+We identified differences between the proportions of various socio-economic factors the survey respondents when compared to the 2016 Census data for Ontario.
+These factors included age groups, income, and ethnicity/race identified.
+Additionally, because the Census divisions do not match the exact boundaries of the Census, we also obtained population estimates for each Health Region from the Ontario Health website in order to correct for the population within each Health Region.
+We corrected for all of these factors (age group, race/ethnicity, income, Health Region Population) using an iterative proportional fitting procedure (also known as *raking*) [@deming1940] in R using the `survey` package.
+The proportions of each of these factors and the data from the 2016 Census Data for Ontario used for the corrections can be found in the Appendix.
+Of notice, because the categories provided by the survey in some cases (e.g., race/ethnicity categories) did not match the categories from the Census, we aggregated them where appropriate to obtain an approximation to the categories in the Census.
+
+<!-- ## Geographical location {#sec-geographical-location} -->
+
+<!-- For each survey participant certain data was automatically captured.
+This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset.
+Due to our interest in exploring vaccination status within a health region that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities].
+The geographical regions were obtained from the websie of the province of Ontario <https://www.ontario.ca/page/list-ontario-municipalities#section-3> -->
 
@@
 
-We used a multivariable logistic regression with the socio-economic factors described in @sec-socio-demographic-factors and the aggregated demographical locations from @sec-geographical-location.
+We used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location.
+We examined the interaction of income and Health Region as it has been shown that there has been an increase in income inequality within the provinces of Canada over time [@marchand2020], and the relatively recent implementation of the Health Regions motivated us to explore if income disparity may be an important factor to consider within each of the Health Regions.
+Our model appears in @eq-model.
+
+$$
+\log \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}}= \beta_0+ \beta_{1}\textrm{(age group)} +\beta_{2} \textrm{ race} + \beta_3 \textrm{(Health Region} \times \textrm{income)} 
+$$ {#eq-model}
+
+Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine.
+
+# Results
+
+## Descriptive statistics of the Fields Covid-19 Survey
+
+@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed.
+
+```{r, load-prepare-data}
+#| echo: false
+#| warning: false
+#| message: false
+
+#load packages
+
+library(weights)
+library(here)
+library(survey)
+library(magrittr)
+library(tidyverse)
+library(lme4)
+library(gtsummary)
+
+data<-read.csv(here("data","clean_dataset.csv"))
+
+# remove rows with missing observations for race
+
+clean_data<-data%>%drop_na(race)
+
+#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
+
+clean_data<-clean_data%>%
+  select(-Municipal.status)
+
+#make variables of interest factors
+clean_data<-clean_data %<>%
+  mutate_at(c("income_ord","race","age_group","location","Geographic_area","Geographic_area_title"),factor)
+# recode the response of interest for categorical analysis
+clean_data<-clean_data %>%
+  mutate(first_dose_m=case_when(
+    first_dose=="yes"~1,
+    first_dose=="no"~0
+  ))
+
+# get percentages to see how different the data is from the Census (details of the Census values are in
+# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")
+
+##### Summary of the dataset####
+
+data_summary<-clean_data %>%
+  tbl_summary(percent="row",
+    by= first_dose,
+    include=c(income_ord,
+    age_group,
+    Health_Region,
+    race)
+  ) %>% bold_labels()
+
+```
+
+```{r,summary-table}
+#| label: tbl-descriptive-stats
+#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey"
+#| message: false
+#| echo: false
+
+data_summary
+
+```
 

### 2023-01-13 protocol/Background_and_methods_Jan_12.tex
updated protocol

--- 2022-12-30 protocol/Background_and_methods_Dec_30.tex
@@
 \hypersetup{
-  pdftitle={Covid Fields Data Protocol},
+  pdftitle={Covid-19 Fields Survey Data Protocol},
   pdfauthor={Ariel Mundo Ortiz},
@@
 
-\title{Covid Fields Data Protocol}
+\title{Covid-19 Fields Survey Data Protocol}
 \author{Ariel Mundo Ortiz}
-\date{}
+\date{1/12/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, frame hidden, boxrule=0pt, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 inform public health decision-makers to design policies to that are
-aimed at addressing vaccination disparities. In this regad, previous
+aimed at addressing vaccination disparities. In this regard, previous
 geographical (spatial) analyses of vaccination rates have shown that
-variations in vaccine uptake can occur within small governamental
+variations in vaccine uptake can occur within small governmental
 administrative units (e.g., counties in the case of the
@@
 to the best of our knowledge, there are no studies that have analyzed
-vaccination status within a province at the district/rural municipality
-level, which can be useful to identify inequalities over these
-geographical areas, thus providing a dissagregated view that can help
-understand the barriers for vaccine delivery in the case of visible
-minorities, which have been disproportionately impacted in the
+vaccination status within a province to identify inequalities that may
+exist within these geographical areas. A dissagregated view of
+variability within a province can help understand the barriers for
+vaccine delivery in the case of visible minorities, which have been
+disproportionately impacted in the
 pandemic\textsuperscript{\protect\hyperlink{ref-hussain2022}{19}}.
@@
 status due to socio-economic (e.g., ethnic origin, age, income) and
-geographical factors (at the municipal level).
+geographical factors (at level of the Health Regions of Ontario).
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1457}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8543}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2361}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7639}}@{}}
 \caption{\label{tbl-covariates}Selected socio-economic factors from the
@@
 Caucasian, other \\
-Employment status & yes, no \\
+& \\
 \bottomrule()
@@
 
-We selected two of the questions that were asked in the survey regarding
-vaccination status:
+From the survey, we selected the question regarding vaccination status:
 
 \begin{itemize}
+\tightlist
 \item
@@
   possible answers ``yes'' and ``no''
-\item
-  (If answered ``yes'' to the previous question) ``Have you received the
-  second dose of the COVID vaccine?'' with possible answers ``yes'' and
-  ``no''
 \end{itemize}
@@
 lowest missing rates (Table~\ref{tbl-covariates}) and the dependent
-variables.
+variable.
 
@@
 
+\hypertarget{sec-geographical-location}{%
+\subsubsection{Geographical location}\label{sec-geographical-location}}
+
+For each survey participant certain data was automatically captured.
+This included the nearest municipality, which resulted in a total of 578
+different municipalities within the dataset. Because our interest was to
+analyze the differences between Health Regions, we assigned the location
+(city) of each entry to its correspondent Health Region following a
+multi-step process: First, we used the municipality information from the
+website of the Association of Municipalities of Ontario (AMO) to assign
+geographical regions to each entry on the dataset. Secondly, we used a
+publicly available dataset of long-term care homes by Local Health
+Integrated Network (LHIN), which were the geographical divisions for
+health in Ontario before the adoption of the Health Regions, to match
+each city with its corresponding LHIN. Cities that did not have a LHIN
+entry on the dataset were searched in the websites of the LHINs and
+manually added to the dataset. Finally, we assigned the corresponding
+Health Regio to each city by matching the LHINs to the Health Regions
+that now encompass them, following the information provided by Ontario
+Health (details of the complete process are in the Appendix).
+
 \hypertarget{corrections}{%
@@
 
-Differences between the clean dataset and the 2016 Census data for
-Ontario regarding proportions of age groups, income, and ethnicity/race
-were identified. Therefore, the proportions of each of these variables
-were corrected using an iterative proportional fitting procedure (also
-known as
+We identified differences between the proportions of various
+socio-economic factors the survey respondents when compared to the 2016
+Census data for Ontario. These factors included age groups, income, and
+ethnicity/race identified. Additionally, because the Census divisions do
+not match the exact boundaries of the Census, we also obtained
+population estimates for each Health Region from the Ontario Health
+website in order to correct for the population within each Health
+Region. We corrected for all of these factors (age group,
+race/ethnicity, income, Health Region Population) using an iterative
+proportional fitting procedure (also known as
 \emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{21}}
-in R using the \texttt{survey} package. Proportions for the correction
-were obtained from the 2016 Census Data for Ontario. Because the
+in R using the \texttt{survey} package. The proportions of each of these
+factors and the data from the 2016 Census Data for Ontario used for the
+corrections can be found in the Appendix. Of notice, because the
 categories provided by the survey in some cases (e.g., race/ethnicity
 categories) did not match the categories from the Census, we aggregated
-some of them to obtain an approximation to the categories in the Census.
-The aggregation is described in detail in the Appendix.
-
-\hypertarget{sec-geographical-location}{%
-\subsection{Geographical location}\label{sec-geographical-location}}
-
-For each survey participant certain data was automatically captured.
-This included the nearest municipality, which resulted in a total of 578
-different municipalities within the dataset. Because our interest lies
-in exploring vaccination status within a governamental administrative
-unit that encompasses multiple cities, we grouped the municipalities
-from the dataset according to the regional government aggregations
-provided by the Association of Municipalities of Ontario, which groups
-municipalities within regions, counties, districts, and
-single-tiers\textsuperscript{\protect\hyperlink{ref-ontario-municipalities}{22}}.
+them where appropriate to obtain an approximation to the categories in
+the Census.
 
@@
 
-We used a multivariable logistic regression with the socio-economic
-factors described in Section~\ref{sec-socio-demographic-factors} and the
-aggregated demographical locations from
-Section~\ref{sec-geographical-location}.
+We used a fixed-effects logistic regression to estimate the probability
+of vaccination depending on the socio-economic factors described in
+Section~\ref{sec-socio-demographic-factors} and the Health Regions from
+Section~\ref{sec-geographical-location}. We examined the interaction of
+income and Health Region as it has been shown that there has been an
+increase in income inequality within the provinces of Canada over
+time\textsuperscript{\protect\hyperlink{ref-marchand2020}{22}}, and the
+relatively recent implementation of the Health Regions motivated us to
+explore if income disparity may be an important factor to consider
+within each of the Health Regions. Our model appears in
+Equation~\ref{eq-model}.
+
+\begin{equation}\protect\hypertarget{eq-model}{}{
+\log \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}}= \beta_0+ \beta_{1}\textrm{(age group)} +\beta_{2} \textrm{ race} + \beta_3 \textrm{(Health Region} \times \textrm{income)} 
+}\label{eq-model}\end{equation}
+
+Where \(p\textrm{(vac)}\) indicates the probability of having received
+the first dose of a Covid-19 vaccine.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{descriptive-statistics-of-the-fields-covid-19-survey}{%
+\subsection{Descriptive statistics of the Fields Covid-19
+Survey}\label{descriptive-statistics-of-the-fields-covid-19-survey}}
+
+Table~\ref{tbl-descriptive-stats} shows the descriptive statistics
+(uncorrected) from the Fields Covid-19 survey data for vaccination
+status and each of the covariates analyzed.
+
+\hypertarget{tbl-descriptive-stats}{}
+\begin{longtable}[]{@{}lcc@{}}
+\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the
+Fields Covid-19 Survey}\tabularnewline
+\toprule()
+\textbf{Characteristic} & \textbf{no}, N = 1,000 & \textbf{yes}, N =
+2,709 \\
+\midrule()
+\endfirsthead
+\toprule()
+\textbf{Characteristic} & \textbf{no}, N = 1,000 & \textbf{yes}, N =
+2,709 \\
+\midrule()
+\endhead
+\textbf{income\_ord} & & \\
+15000\_24999 & 167 (34\%) & 326 (66\%) \\
+25000\_39999 & 144 (31\%) & 316 (69\%) \\
+40000\_59999 & 116 (25\%) & 348 (75\%) \\
+60000\_89999 & 71 (17\%) & 343 (83\%) \\
+over\_90000 & 251 (25\%) & 752 (75\%) \\
+under\_15000 & 251 (29\%) & 624 (71\%) \\
+\textbf{age\_group} & & \\
+16\_24 & 230 (27\%) & 634 (73\%) \\
+25\_34 & 193 (27\%) & 518 (73\%) \\
+35\_44 & 136 (25\%) & 400 (75\%) \\
+45\_54 & 130 (28\%) & 335 (72\%) \\
+55\_64 & 80 (20\%) & 317 (80\%) \\
+65\_and\_over & 231 (31\%) & 505 (69\%) \\
+\textbf{Health\_Region} & & \\
+Central & 224 (28\%) & 581 (72\%) \\
+East & 135 (23\%) & 448 (77\%) \\
+North East & 36 (34\%) & 71 (66\%) \\
+North West & 6 (12\%) & 45 (88\%) \\
+Toronto & 371 (28\%) & 953 (72\%) \\
+West & 228 (27\%) & 611 (73\%) \\
+\textbf{race} & & \\
+arab\_middle\_eastern & 77 (35\%) & 141 (65\%) \\
+black & 120 (39\%) & 186 (61\%) \\
+east\_asian\_pacific\_islander & 72 (23\%) & 236 (77\%) \\
+indigenous & 82 (38\%) & 134 (62\%) \\
+latin\_american & 70 (38\%) & 112 (62\%) \\
+mixed & 111 (34\%) & 213 (66\%) \\
+other & 134 (34\%) & 255 (66\%) \\
+south\_asian & 89 (23\%) & 294 (77\%) \\
+white\_caucasian & 245 (18\%) & 1,138 (82\%) \\
+\bottomrule()
+\end{longtable}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-municipalities}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-marchand2020}{}}%
 \CSLLeftMargin{22. }%
-\CSLRightInline{Ontario {M}unicipalities. Accessed December 30, 2022.
-\url{https://www.amo.on.ca/about-us/municipal-101/ontario-municipalities}}
+\CSLRightInline{Marchand Y, Dubé J, Breau S. Exploring the causes and
+consequences of regional income inequality in canada. \emph{Economic
+Geography}. 2020;96(2):83-107.
+doi:\href{https://doi.org/10.1080/00130095.2020.1715793}{10.1080/00130095.2020.1715793}}
 

### 2023-01-13 code/raking_and_model.R

@@ -41,7 +41,17 @@ wpct(clean_data$age_group)
 wpct(clean_data$race)
 wpct(clean_data$location)
 
+##### Summary of the dataset####
 
+
+data_summary<-clean_data %>%
+  tbl_summary(percent="row",
+    by= first_dose,
+    include=c(income_ord,
+    age_group,
+    Health_Region,
+    race)
+  )
 
 ################ RAKING ###################################
 ## I had originally used the percentages from the census, but the "population values"
@@ -277,11 +287,22 @@ a2_rake<-rake(a1,
                               income_ords,
                               health_regions))
 
-
+raking_summary <-a2_rake %>% tbl_svysummary(by = first_dose, 
+                           percent = "row", 
+                           include=c(income_ord,
+                                     age_group,
+                                     Health_Region,
+                                     race))%>%
+  bold_labels()
 ## The sampling probabilities have now changed after the correction
 
-a1_rake$prob #probabilities have changed
+tbl_merge_raking<-
+  tbl_merge(
+    tbls = list(data_summary,raking_summary),
+    tab_spanner = c("**Uncorrected data**","**Corrected by raking**")
+  )
 
+tbl_merge_raking
 
 ### Model ###
 
@@ -290,7 +311,14 @@ a1_rake$prob #probabilities have changed
 ## The model uses the response to first dose vaccine status (yes or no) using income, age group,
 ## race, and the geographic areas
 
-model1.hr<-svyglm(first_dose_m~+age_group+Health_Region*income_ord,
+ #model with raking for geographical region
+
+model1<-svyglm(first_dose_m~+age_group+race+Health_Region*income_ord,
+                  design=a1_rake,
+                  family = quasibinomial(), 
+                  control= list(maxit=25))
+
+model1.hr<-svyglm(first_dose_m~+age_group+race+Health_Region*income_ord,
               design=a2_rake,
               family = quasibinomial(), 
               control= list(maxit=25))
@@ -299,42 +327,44 @@ model1.hr<-svyglm(first_dose_m~+age_group+Health_Region*income_ord,
 
 summary(model1.hr)
 
+
 ## helps visualize the model output
 
-x1<-model1 %>% tbl_regression(exponentiate = FALSE)
+x1<-model1 %>% tbl_regression(exponentiate = TRUE)
 
 
[…]

### 2023-01-13 protocol/Background_and_methods_Jan_12.qmd
updatd models

@@
     number-sections: true
+    include-in-header: 
+      text: |
+        \usepackage{placeins}
 editor: visual
@@
 ---
+
+```{r, load-prepare-data}
+#| echo: false
+#| warning: false
+#| message: false
+
+#load packages
+
+library(weights)
+library(here)
+library(survey)
+library(magrittr)
+library(tidyverse)
+library(lme4)
+library(kableExtra)
+library(modelsummary)
+library(gtsummary)
+
+data<-read.csv(here("data","clean_dataset.csv"))
+
+# remove rows with missing observations for race
+
+clean_data<-data%>%drop_na(race)
+
+#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
+
+clean_data<-clean_data %>%
+  select(-Municipal.status) %>%
+  rename(Race=race,Age_group=age_group,Income=income_ord)
+
+#make variables of interest factors
+clean_data<-clean_data %<>%
+  mutate_at(c("Income","Race","Age_group","location","Geographic_area","Geographic_area_title"),factor)
+# recode the response of interest for categorical analysis
+clean_data<-clean_data %>%
+  mutate(first_dose_m=case_when(
+    first_dose=="yes"~1,
+    first_dose=="no"~0
+  ))
+
+# get percentages to see how different the data is from the Census (details of the Census values are in
+# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")
+
+##### Summary of the dataset####
+
+data_summary<-clean_data %>%
+  tbl_summary(percent="row",
+    by= first_dose,
+    include=c(Income,
+    Age_group,
+    Health_Region,
+    Race)
+  ) %>% bold_labels()
+
+# Each of these dataframes provide the population totals for the different variables and their categories
+
+Ages<-data.frame(Age_group=c("16_24",
+                             "25_34",
+                             "35_44",
+                             "45_54",
+                             "55_64",
+                             "65_and_over"),
+                 Freq=c(1707959,
+                        1734856,
+                        1721407,
+                        2003826,
+                        1842444,
+                        2245898))
+
+Races<-data.frame(Race=c("arab_middle_eastern",
+                         "black",
+                         "east_asian_pacific_islander",
+                         "indigenous",
+                         "latin_american",
+                         "mixed",
+                         "other",
+                         "south_asian",
+                         "white_caucasian"),
+                  Freq=c(212782,
+                         638346,
+                         886592,
+                         376558,
+                         197020,
+                         130033,
+                         705333,
+                         1166361,
+                         9118079 ))
+
+Incomes<-data.frame(Income=c("under_15000",
+                                     "15000_24999",
+                                     "25000_39999",
+                                     "40000_59999",
+                                     "60000_89999",
+                                     "over_90000"),
+                        Freq=c(294643,
+                               387688,
+                               599624,
+                               796052,
+                               1007988,
+                               2083176
+
+                        ))
+
+# corrections for health regions
+
+Health_Regions<-data.frame(Health_Region=c(
+  "North West",
+  "North East",
+  "West",
+  "East",
+  "Central",
+  "Toronto"),
+  Freq=c(232299,
+    557000,
+    4000000,
+    3700000,
+    5000000,
+    1400000
+  )
+)
+
+### Raking ####
+
+## First, provide a survey design object. This is done using the dataset. The syntax
+## means that there is no stratification (id=~1). When the line below is run it will throw
+## a warning because we do not have "design" weights in the dataset (other datasets do have this).
+
+a1<-svydesign(id=~1,data=clean_data)
+
+## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage.
+
+#a1$prob #all probs are 1
+
+## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
+## from the data frames from above.
+
+a1_rake<-rake(a1,
+             sample=list(~Age_group,
+                         ~Race,
+                         ~Income,
+                         ~Health_Region),
+             population=list(Ages,
+                             Races,
+                             Incomes,
+                             Health_Regions))
+
+```
 
@@
 
-This study will examine self-reported COVID-19 vaccination status in the province of Ontario in order to determine the influence in vaccination status due to socio-economic (e.g., ethnic origin, age, income) and geographical factors (at level of the Health Regions of Ontario).
+This study will examine self-reported COVID-19 vaccination status within the province of Ontario in order to determine how socio-economic (e.g., ethnic origin, age, income) and geographical factors (at level of the Health Regions of Ontario) influence vaccination within the province.
 
@@
 Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variable.
+The final clean dataset contained 3,709 unique entries.
 
@@
 
-## Statistical model
+## Statistical models
 
-We used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location.
-We examined the interaction of income and Health Region as it has been shown that there has been an increase in income inequality within the provinces of Canada over time [@marchand2020], and the relatively recent implementation of the Health Regions motivated us to explore if income disparity may be an important factor to consider within each of the Health Regions.
-Our model appears in @eq-model.
+Because of the binary outcome of the survey answer of our interest (vaccination status as "yes" or "no") we used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location.
+Previous studies have shown that socio-economic factors, and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021;@cnat2022a].
+At the same time in the case of Canada, others have indicated an ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].
+
+Therefore, we built different logistic regression models that accounted for different levels of interaction between the available socio-economic factors obtained from the Fields Covid survey.
+The different models are described below.
+
+### Model building and selection
+
+The first model (`model1`) examined the interaction of income and Health Region as it has been shown that there has been an increase in income inequality within the provinces of Canada over time [@marchand2020], and the relatively recent implementation of the Health Regions motivated us to explore if income disparity may be an important factor to consider within each of the Health Regions.
+
+The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking. The model appears in @eq-model1.
 
 $$
-\log \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}}= \beta_0+ \beta_{1}\textrm{(age group)} +\beta_{2} \textrm{ race} + \beta_3 \textrm{(Health Region} \times \textrm{income)} 
-$$ {#eq-model}
+\begin{aligned}
+\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} +\\ \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income} + \beta_5\textrm{(Health Region} \times \textrm{Income)}
+\end{aligned}
+$$ {#eq-model1}
 
 Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine.
+
+The second model (`model2`) incorporated an interaction effect between age group and income to examine if a significant interaction occured between these factors as well.
+The model appears in @eq-model2.
+
+$$
+\begin{aligned}
+\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \\\beta_4 \textrm{ Income}+ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)}
+\end{aligned}
+$$ {#eq-model2}
+
+A third model (`model3`) incorporated an interaction effect between race and income, and appears in @eq-model3.
+
+$$
+\begin{aligned}
+\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)} + \beta_7 \textrm{ (Race} \times \textrm{Income)}
+\end{aligned}
+$$ {#eq-model3}
+
+Although it could have been interesting to examine the interaction of Health Region and Race, or 3-way interactions between socio-economic factors (e.g., age,income,race), these models had singularity issues.
+Next, we compared the AIC of the three models indicated above.
+The AICs are very similar and appear in @tbl-AIC.
+
+```{r,models}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-AIC
+#| tbl-cap: "AIC for the three models"
+
+### Model ###
+
+## the first model is a simple logistic regression with fixed effects only svyglm uses the glm
+## function to calculate, so it does not allow for random effects.
+
+model1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region,
+              design=a1_rake,
+              family = quasibinomial(), 
+              control= list(maxit=25))
+
+model2<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income,
+              design=a1_rake,
+              family = quasibinomial(), 
+              control= list(maxit=25))
+
+model3<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income+Race*Income,
+              design=a1_rake,
+              family = quasibinomial(), 
+              control= list(maxit=25))
+
+# model4<-svyglm(first_dose_m~Age_group+Income+Race+Health_Region+Health_Region*Income+Income*Age_group+Income*Race+Income*Age_group*Race,
+#               design=a1_rake,
+#               family = quasibinomial(), 
+#               control= list(maxit=25))
+# 
+# 
+# model5<-svyglm(first_dose_m~Age_group+Income+Race+Health_Region+Health_Region*Income+Income*Age_group+Income*Race+Race*Health_Region,
+#               design=a1_rake,
+#               family = quasibinomial(), 
+#               control= list(maxit=25))
+
+#summary(model1.test)
+#summary(model1.hr)
+
+model_comp<-as_tibble(AIC(model1,model2,model3))%>%
+  select(-c(eff.p,deltabar))%>%
+  add_column(model=c("model1","model2","model3"))
+
+model_comp%>%
+  kbl(booktabs=T,
+      col.names = c("AIC","model"))
+
+## helps visualize the model output
+
+###############
+
+```
 
@@
 
-```{r, load-prepare-data}
-#| echo: false
-#| warning: false
-#| message: false
-
-#load packages
-
-library(weights)
-library(here)
-library(survey)
-library(magrittr)
-library(tidyverse)
-library(lme4)
-library(gtsummary)
-
-data<-read.csv(here("data","clean_dataset.csv"))
-
-# remove rows with missing observations for race
-
-clean_data<-data%>%drop_na(race)
-
-#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
-
-clean_data<-clean_data%>%
-  select(-Municipal.status)
-
-#make variables of interest factors
-clean_data<-clean_data %<>%
-  mutate_at(c("income_ord","race","age_group","location","Geographic_area","Geographic_area_title"),factor)
-# recode the response of interest for categorical analysis
-clean_data<-clean_data %>%
-  mutate(first_dose_m=case_when(
-    first_dose=="yes"~1,
-    first_dose=="no"~0
-  ))
-
-# get percentages to see how different the data is from the Census (details of the Census values are in
-# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")
-
-##### Summary of the dataset####
-
-data_summary<-clean_data %>%
-  tbl_summary(percent="row",
-    by= first_dose,
-    include=c(income_ord,
-    age_group,
-    Health_Region,
-    race)
-  ) %>% bold_labels()
-
-```
-
 ```{r,summary-table}
@@
 
-data_summary
+data_summary %>%
+  modify_header(label = "**Variable**") %>%
+  as_flex_table()
 
 ```
+
+## Model Comparisons
+
+The output from the three models is too long to give a summarized view. I am still trying to get the important parts out but in the meantime here are the summary findings:
+
+- **Model 1**
+  Significant values on:
+- Age: None
+- Race: East Asian/Pacific Islander, South Asian, White Caucasian
+- Income: 60000-89000 CAD, Over 90000
+- Interaction of Income and Health Region: Toronto and 60000-899999, North West and over_90000, North West  and under_15000
+
+- **Model 2**
+    Significant values on:
+- Age: None
+- Race: East Asian/Pacific Islander, South Asian, White Caucasian
+- Health Region: East
+- Income: 60000-89000 CAD, Over 90000
+- Interaction of Income and Health Region: East and 60000_89999, Toronto and 60000-899999, North West and over_90000, North West  and under_15000
+- Interaction of Age and income: A couple of low values, but none below 0.05
+
+including here the results of the third model, which I believe is the most appropriate as it has a similar AIC to the other ones, and it is able to provide interaction information on three different instances.
+
+- **Model 3**
+    Significant values on:
+- Age: None
+- Race: Indigenous
+- Health Region: East, Toronto
+- Income: 60000-89000 CAD
+- Interaction of Income and Health Region: East and 40000_59999, East and 60000_89999, Toronto and 60000-899999, East and  over_90000, North West and over_90000, North West  and under_15000
+- Interaction of Age and income: A couple of low values, but none below 0.05
+- Interaction of Race and Income: indigenous and 25000_39999, indigenous and over_90000, mixed and over_90000, white_caucasian and over_90000
+
+So far, considering that the interaction of socioeconomic factors are important predictors of vaccination status, I believe the model that provides a more cohesive picture of the data is model 3 as it shows differences between race and income which could be important if we consider the types of occupations that people in certain regions do, and how these influences their vaccination status.
+
+<!-- @tbl-model shows the odds-ratio, 95\% confidence interval, and p-value for the logistic regression model for the first dose of vaccination by age group, race, income, Health Region, and the interaction of income and Health Region.
+-->
+
+```{r, model-table}
+#| label: tbl-model
+#| tbl-cap: "Logistic Regression Model Estimates"
+#| message: false
+#| echo: false
+#| warning: false
+
+# models<-list(model1,model2,model3)
+# # 
+# modelsummary(models,exponentiate = TRUE, stars = T, statistic = c("conf.int",
+#                           "p = {p.value}"))
+# 
+#  
+# modelsummary(model1,exponentiate = TRUE, stars = T, statistic = c("conf.int",
+#                           coef_omit = c(1),
+#                           "p = {p.value}"))
+
+# model_tbl1<-model1 %>%
+#   tbl_regression(exponentiate = TRUE) %>%
+#   bold_labels()%>%
+#   modify_header(label = "**Variable**")
+# # 
+# # test<-summary(model1)
+# # 
+# model_tbl1<-model1 %>%
+#   tbl_regression(exponentiate = TRUE) %>%
+#   select(variable, estimate, p.value) %>%
+#   filter(p.value <= 0.05)%>%
+#   kbl()%>%
+#   as_kable_extra() %>%
+#   kableExtra::kable_styling(
+#     font_size = 8
+#   )
+# 
+# # 
+# model_tbl2<-model2 %>%
+#   tbl_regression(exponentiate = TRUE) %>%
+#   bold_labels()%>%
+#   modify_header(label = "**Variable**")%>%
+#   as_kable_extra() %>%
+#   kableExtra::kable_styling(
+#     font_size = 8
+#   )
+# 
+# model_tbl3<-model3 %>%
+#   tbl_regression(exponentiate = TRUE) %>%
+#   bold_labels()%>%
+#   modify_header(label = "**Variable**")%>%
+#   as_kable_extra() %>%
+#   kableExtra::kable_styling(
+#     font_size = 10
+#   )
+
+# model_tbl3
+# 
+# model_tbl3
+# 
+# tbl_merge_model<-
+#   tbl_merge(
+#     tbls = list(model_tbl1,model_tbl2,model_tbl3),
+#     tab_spanner = c("**Model 1**","**Model 2**", "**Model 3**")
+#   )%>%
+#   as_flex_table()
+# 
+# tbl_merge_model
+#
+
+# modelsummary(model3,exponentiate = TRUE, stars = T, statistic = c("conf.int",
+#                           "p = {p.value}"))
+
+```
+
+\FloatBarrier
 

### 2023-01-13 protocol/Background_and_methods_Jan_12.tex
updatd models

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{fontspec}
+\usepackage{multicol}
+\usepackage{hhline}
+\usepackage{hyperref}
+\usepackage{placeins}
 \KOMAoption{captions}{tableheading}
@@
 \author{Ariel Mundo Ortiz}
-\date{1/12/23}
+\date{1/13/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, enhanced, breakable, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-This study will examine self-reported COVID-19 vaccination status in the
-province of Ontario in order to determine the influence in vaccination
-status due to socio-economic (e.g., ethnic origin, age, income) and
-geographical factors (at level of the Health Regions of Ontario).
+This study will examine self-reported COVID-19 vaccination status within
+the province of Ontario in order to determine how socio-economic (e.g.,
+ethnic origin, age, income) and geographical factors (at level of the
+Health Regions of Ontario) influence vaccination within the province.
 
@@
 lowest missing rates (Table~\ref{tbl-covariates}) and the dependent
-variable.
+variable. The final clean dataset contained 3,709 unique entries.
 
@@
 
-\hypertarget{statistical-model}{%
-\subsection{Statistical model}\label{statistical-model}}
+\hypertarget{statistical-models}{%
+\subsection{Statistical models}\label{statistical-models}}
 
-We used a fixed-effects logistic regression to estimate the probability
-of vaccination depending on the socio-economic factors described in
+Because of the binary outcome of the survey answer of our interest
+(vaccination status as ``yes'' or ``no'') we used a fixed-effects
+logistic regression to estimate the probability of vaccination depending
+on the socio-economic factors described in
 Section~\ref{sec-socio-demographic-factors} and the Health Regions from
-Section~\ref{sec-geographical-location}. We examined the interaction of
-income and Health Region as it has been shown that there has been an
-increase in income inequality within the provinces of Canada over
-time\textsuperscript{\protect\hyperlink{ref-marchand2020}{22}}, and the
+Section~\ref{sec-geographical-location}. Previous studies have shown
+that socio-economic factors, and their interactions are significant
+predictors of intent of vaccination and vaccination
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{22}--\protect\hyperlink{ref-cnat2022a}{24}}.
+At the same time in the case of Canada, others have indicated an ongoing
+need of socio-economic information that can provide a rationale for the
+disparities in vaccination observed within some racial
+groups\textsuperscript{\protect\hyperlink{ref-cnat2022b}{25}}.
+
+Therefore, we built different logistic regression models that accounted
+for different levels of interaction between the available socio-economic
+factors obtained from the Fields Covid survey. The different models are
+described below.
+
+\hypertarget{model-building-and-selection}{%
+\subsubsection{Model building and
+selection}\label{model-building-and-selection}}
+
+The first model (\texttt{model1}) examined the interaction of income and
+Health Region as it has been shown that there has been an increase in
+income inequality within the provinces of Canada over
+time\textsuperscript{\protect\hyperlink{ref-marchand2020}{26}}, and the
 relatively recent implementation of the Health Regions motivated us to
 explore if income disparity may be an important factor to consider
-within each of the Health Regions. Our model appears in
-Equation~\ref{eq-model}.
+within each of the Health Regions.
 
-\begin{equation}\protect\hypertarget{eq-model}{}{
-\log \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}}= \beta_0+ \beta_{1}\textrm{(age group)} +\beta_{2} \textrm{ race} + \beta_3 \textrm{(Health Region} \times \textrm{income)} 
-}\label{eq-model}\end{equation}
+The the model was fitted using the function \texttt{svyglm} from the
+\texttt{survey} R package in order to incorporate the correction in
+sampling probability obtained from raking. The model appears in
+Equation~\ref{eq-model1}.
+
+\begin{equation}\protect\hypertarget{eq-model1}{}{
+\begin{aligned}
+\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} +\\ \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income} + \beta_5\textrm{(Health Region} \times \textrm{Income)}
+\end{aligned}
+}\label{eq-model1}\end{equation}
 
@@
 the first dose of a Covid-19 vaccine.
+
+The second model (\texttt{model2}) incorporated an interaction effect
+between age group and income to examine if a significant interaction
+occured between these factors as well. The model appears in
+Equation~\ref{eq-model2}.
+
+\begin{equation}\protect\hypertarget{eq-model2}{}{
+\begin{aligned}
+\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \\\beta_4 \textrm{ Income}+ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)}
+\end{aligned}
+}\label{eq-model2}\end{equation}
+
+A third model (\texttt{model3}) incorporated an interaction effect
+between race and income, and appears in Equation~\ref{eq-model3}.
+
+\begin{equation}\protect\hypertarget{eq-model3}{}{
+\begin{aligned}
+\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)} + \beta_7 \textrm{ (Race} \times \textrm{Income)}
+\end{aligned}
+}\label{eq-model3}\end{equation}
+
+Although it could have been interesting to examine the interaction of
+Health Region and Race, or 3-way interactions between socio-economic
+factors (e.g., age,income,race), these models had singularity issues.
+Next, we compared the AIC of the three models indicated above. The AICs
+are very similar and appear in Table~\ref{tbl-AIC}.
+
+\hypertarget{tbl-AIC}{}
+\begin{table}
+\caption{\label{tbl-AIC}AIC for the three models }\tabularnewline
+
+\centering
+\begin{tabular}[t]{rl}
+\toprule
+AIC & model\\
+\midrule
+3580.878 & model1\\
+3614.465 & model2\\
+3633.575 & model3\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 \hypertarget{tbl-descriptive-stats}{}
-\begin{longtable}[]{@{}lcc@{}}
-\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the
-Fields Covid-19 Survey}\tabularnewline
-\toprule()
-\textbf{Characteristic} & \textbf{no}, N = 1,000 & \textbf{yes}, N =
-2,709 \\
-\midrule()
-\endfirsthead
-\toprule()
-\textbf{Characteristic} & \textbf{no}, N = 1,000 & \textbf{yes}, N =
-2,709 \\
-\midrule()
-\endhead
-\textbf{income\_ord} & & \\
-15000\_24999 & 167 (34\%) & 326 (66\%) \\
-25000\_39999 & 144 (31\%) & 316 (69\%) \\
-40000\_59999 & 116 (25\%) & 348 (75\%) \\
-60000\_89999 & 71 (17\%) & 343 (83\%) \\
-over\_90000 & 251 (25\%) & 752 (75\%) \\
-under\_15000 & 251 (29\%) & 624 (71\%) \\
-\textbf{age\_group} & & \\
-16\_24 & 230 (27\%) & 634 (73\%) \\
-25\_34 & 193 (27\%) & 518 (73\%) \\
-35\_44 & 136 (25\%) & 400 (75\%) \\
-45\_54 & 130 (28\%) & 335 (72\%) \\
-55\_64 & 80 (20\%) & 317 (80\%) \\
-65\_and\_over & 231 (31\%) & 505 (69\%) \\
-\textbf{Health\_Region} & & \\
-Central & 224 (28\%) & 581 (72\%) \\
-East & 135 (23\%) & 448 (77\%) \\
-North East & 36 (34\%) & 71 (66\%) \\
-North West & 6 (12\%) & 45 (88\%) \\
-Toronto & 371 (28\%) & 953 (72\%) \\
-West & 228 (27\%) & 611 (73\%) \\
-\textbf{race} & & \\
-arab\_middle\_eastern & 77 (35\%) & 141 (65\%) \\
-black & 120 (39\%) & 186 (61\%) \\
-east\_asian\_pacific\_islander & 72 (23\%) & 236 (77\%) \\
-indigenous & 82 (38\%) & 134 (62\%) \\
-latin\_american & 70 (38\%) & 112 (62\%) \\
-mixed & 111 (34\%) & 213 (66\%) \\
-other & 134 (34\%) & 255 (66\%) \\
-south\_asian & 89 (23\%) & 294 (77\%) \\
-white\_caucasian & 245 (18\%) & 1,138 (82\%) \\
-\bottomrule()
+\begin{table}[H]
+
+\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+
+\setlength{\tabcolsep}{0pt}
+
+\renewcommand*{\arraystretch}{1.5}
+
+\begin{longtable}[c]{|p{2.35in}|p{1.21in}|p{1.27in}}
+
+\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields Covid-19 Survey } \\
+
+\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Variable}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{no,\ N\ =\ 1,000}}\textcolor[HTML]{000000}{\textsuperscript{\fontsize{11}{11}\selectfont{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{yes,\ N\ =\ 2,709}}\textcolor[HTML]{000000}{\textsuperscript{\fontsize{11}{11}\selectfont{1}}}} \\
+
+\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}\endhead
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Income}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{15000\_24999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{167\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{326\ (66\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{25000\_39999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{144\ (31\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{316\ (69\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{40000\_59999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{116\ (25\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{348\ (75\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{60000\_89999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{71\ (17\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{343\ (83\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{over\_90000}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{251\ (25\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{752\ (75\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{under\_15000}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{251\ (29\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{624\ (71\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Age\_group}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{16\_24}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{230\ (27\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{634\ (73\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{25\_34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{193\ (27\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{518\ (73\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{35\_44}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{136\ (25\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{400\ (75\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{45\_54}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{130\ (28\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{335\ (72\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{55\_64}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{80\ (20\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{317\ (80\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{65\_and\_over}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{231\ (31\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{505\ (69\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Health\_Region}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Central}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{224\ (28\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{581\ (72\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{East}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{135\ (23\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{448\ (77\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{North\ East}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{36\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{71\ (66\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{North\ West}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{6\ (12\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{45\ (88\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Toronto}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{371\ (28\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{953\ (72\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{West}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{228\ (27\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{611\ (73\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Race}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{arab\_middle\_eastern}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{77\ (35\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{141\ (65\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{black}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120\ (39\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{186\ (61\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{east\_asian\_pacific\_islander}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{72\ (23\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{236\ (77\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{indigenous}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{82\ (38\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{134\ (62\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{latin\_american}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{70\ (38\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{112\ (62\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{mixed}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{111\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{213\ (66\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{other}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{134\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{255\ (66\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{south\_asian}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{89\ (23\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{294\ (77\%)}}} \\
+
+\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{white\_caucasian}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{245\ (18\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{1,138\ (82\%)}}} \\
+
+\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}
+
+\multicolumn{3}{!{\color[HTML]{FFFFFF}\vrule width 0pt}>{\raggedright}p{\dimexpr 4.83in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{FFFFFF}\vrule width 0pt}}{\textcolor[HTML]{000000}{\textsuperscript{\fontsize{11}{11}\selectfont{1}}}\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{n\ (\%)}}} \\
+
 \end{longtable}
+
+\end{table}
+
+\hypertarget{model-comparisons}{%
+\subsection{Model Comparisons}\label{model-comparisons}}
+
+The output from the three models is too long to give a summarized view.
+I am still trying to get the important parts out but in the meantime
+here are the summary findings:
+
+\begin{itemize}
+\item
+  \textbf{Model 1} Significant values on:
+\item
+  Age: None
+\item
+  Race: East Asian/Pacific Islander, South Asian, White Caucasian
+\item
+  Income: 60000-89000 CAD, Over 90000
+\item
+  Interaction of Income and Health Region: Toronto and 60000-899999,
+  North West and over\_90000, North West and under\_15000
+\item
+  \textbf{Model 2} Significant values on:
+\item
+  Age: None
+\item
+  Race: East Asian/Pacific Islander, South Asian, White Caucasian
+\item
+  Health Region: East
+\item
+  Income: 60000-89000 CAD, Over 90000
+\item
+  Interaction of Income and Health Region: East and 60000\_89999,
+  Toronto and 60000-899999, North West and over\_90000, North West and
+  under\_15000
+\item
+  Interaction of Age and income: A couple of low values, but none below
+  0.05
+\end{itemize}
+
+including here the results of the third model, which I believe is the
+most appropriate as it has a similar AIC to the other ones, and it is
+able to provide interaction information on three different instances.
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{Model 3} Significant values on:
+\item
+  Age: None
+\item
+  Race: Indigenous
+\item
+  Health Region: East, Toronto
+\item
+  Income: 60000-89000 CAD
+\item
+  Interaction of Income and Health Region: East and 40000\_59999, East
+  and 60000\_89999, Toronto and 60000-899999, East and over\_90000,
+  North West and over\_90000, North West and under\_15000
+\item
+  Interaction of Age and income: A couple of low values, but none below
+  0.05
+\item
+  Interaction of Race and Income: indigenous and 25000\_39999,
+  indigenous and over\_90000, mixed and over\_90000, white\_caucasian
+  and over\_90000
+\end{itemize}
+
+So far, considering that the interaction of socioeconomic factors are
+important predictors of vaccination status, I believe the model that
+provides a more cohesive picture of the data is model 3 as it shows
+differences between race and income which could be important if we
+consider the types of occupations that people in certain regions do, and
+how these influences their vaccination status.
+
+\begin{table}
+
+\end{table}
+
+\FloatBarrier
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
+\CSLLeftMargin{22. }%
+\CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
+Corlin L. Disparities in national and state estimates of {COVID}-19
+vaccination receipt and intent to vaccinate by race/ethnicity, income,
+and age group among adults~\(\geq\)~18~years, united states.
+\emph{Vaccine}. 2022;40(1):107-113.
+doi:\href{https://doi.org/10.1016/j.vaccine.2021.11.040}{10.1016/j.vaccine.2021.11.040}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
+\CSLLeftMargin{23. }%
+\CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
+Zikmund-Fisher BJ. Vaccine hesitancy and rejection of a vaccine for the
+novel coronavirus in the united states. \emph{Frontiers in Immunology}.
+2021;12.
+doi:\href{https://doi.org/10.3389/fimmu.2021.558270}{10.3389/fimmu.2021.558270}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
+\CSLLeftMargin{24. }%
+\CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
+and factors related to {COVID}-19 vaccine hesitancy and unwillingness in
+canada: A systematic review and meta-analysis. \emph{Journal of Medical
+Virology}. 2022;95(1).
+doi:\href{https://doi.org/10.1002/jmv.28156}{10.1002/jmv.28156}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
+\CSLLeftMargin{25. }%
+\CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
+review on vaccine hesitancy in black communities in canada: Critical
+issues and research failures. \emph{Vaccines}. 2022;10(11):1937.
+doi:\href{https://doi.org/10.3390/vaccines10111937}{10.3390/vaccines10111937}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-marchand2020}{}}%
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Marchand Y, Dubé J, Breau S. Exploring the causes and

### 2023-01-15 protocol/Background_and_methods_Jan_12.qmd
Working to fix figure

@@
 library(gtsummary)
+library(patchwork)
 
@@
 ```{r, model-table}
-#| label: tbl-model
-#| tbl-cap: "Logistic Regression Model Estimates"
 #| message: false
@@
 #| warning: false
+#| fig-width: 8
+#| fig-height: 8
 
-# models<-list(model1,model2,model3)
-# # 
+ # models<-list("First Model"=model1,
+ #              "Second Model"=model2,
+ #              "Third Model"=model3)
+
+a<-modelplot(model1,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+
+b<-modelplot(model2,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("Second Model")
+
+c<-modelplot(model3,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("Third Model")
+
+a+b+c+plot_layout(guides = 'collect')+plot_annotation(title="Fixed effects of all models")& theme(legend.position = 'bottom')
+
+d<-modelplot(model1,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+
+e<-modelplot(model2,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+
+f<-modelplot(model3,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+
+d+e+f+plot_layout(guides = 'collect')+plot_annotation(title="Interaction effects of all models") & theme(legend.position = 'bottom')
+
 #  modelsummary(models,exponentiate = TRUE, stars = T, statistic = c("conf.int",
@@
 # 
-# modelsummary(model1,exponentiate = TRUE, stars = T, statistic = c("conf.int",
-#                           coef_omit = c(1),
-#                           "p = {p.value}"))
-
+#  modelsummary(model1,exponentiate = TRUE, stars = T, statistic ="conf.int")
+# 
 #  model_tbl1<-model1 %>%
@@
 #    modify_header(label = "**Variable**")
-# # 
-# # test<-summary(model1)
-# # 
+# 
+#   test<-summary(model1)
+# 
 #  model_tbl1<-model1 %>%
@@
 # 
+# 
+# 
+# # # 
+# # model_tbl2<-model2 %>%
+# #   tbl_regression(exponentiate = TRUE) %>%
+# #   bold_labels()%>%
+# #   modify_header(label = "**Variable**")%>%
+# #   as_kable_extra() %>%
+# #   kableExtra::kable_styling(
+# #     font_size = 8
+# #   )
 # # 
-# model_tbl2<-model2 %>%
-#   tbl_regression(exponentiate = TRUE) %>%
-#   bold_labels()%>%
-#   modify_header(label = "**Variable**")%>%
-#   as_kable_extra() %>%
-#   kableExtra::kable_styling(
-#     font_size = 8
-#   )
-# 
 # model_tbl3<-model3 %>%
@@
 #   )
-
-# model_tbl3
 # 
-# model_tbl3
+# # model_tbl3
+# # 
+# # model_tbl3
+# # 
+# # tbl_merge_model<-
+# #   tbl_merge(
+# #     tbls = list(model_tbl1,model_tbl2,model_tbl3),
+# #     tab_spanner = c("**Model 1**","**Model 2**", "**Model 3**")
+# #   )%>%
+# #   as_flex_table()
+# # 
+# # tbl_merge_model
+# # 
 # 
-# tbl_merge_model<-
-#   tbl_merge(
-#     tbls = list(model_tbl1,model_tbl2,model_tbl3),
-#     tab_spanner = c("**Model 1**","**Model 2**", "**Model 3**")
-#   )%>%
-#   as_flex_table()
-# 
-# tbl_merge_model
-#
-
-# modelsummary(model3,exponentiate = TRUE, stars = T, statistic = c("conf.int",
-#                           "p = {p.value}"))
+#  modelsummary(model3,exponentiate = TRUE, 
+#               stars = T, 
+#               estimate = "{estimate} [{conf.low}, {conf.high}]",
+#               #statistic = NULL
+#               )
 

### 2023-01-15 protocol/Background_and_methods_Jan_12.tex
Working to fix figure

@@
 \author{Ariel Mundo Ortiz}
-\date{1/13/23}
+\date{1/15/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, enhanced, breakable, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{table}
+\includegraphics{Background_and_methods_Jan_12_files/figure-pdf/model-table-1.png}
 
-\end{table}
+\includegraphics{Background_and_methods_Jan_12_files/figure-pdf/model-table-2.png}
 

### 2023-01-15 protocol/Background_and_methods_Jan_12.qmd
need to fix labels

@@
 d<-modelplot(model1,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+  scale_color_manual(values = c("black", "red"))+ggtitle("1st Model")
 
 e<-modelplot(model2,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+  scale_color_manual(values = c("black", "red"))+ggtitle("2nd Model")
 
 f<-modelplot(model3,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+  scale_color_manual(values = c("black", "red"))+ggtitle("3rd Model")
 

### 2023-01-16 protocol/Background_and_methods_Jan_12.qmd
Updated

@@
         \usepackage{placeins}
+  html:
+    toc: true
 editor: visual
@@
 - Interaction of Age and income: A couple of low values, but none below 0.05
-- Interaction of Race and Income: indigenous and 25000_39999, indigenous and over_90000, mixed and over_90000, white_caucasian and over_90000
+- Interaction of Race and Income: indigenous and 25000_39999, indigenous and over_90000, mixed and over_90000, white_caucasian and over_90000.
+A visual representation of the estimates of the models can be see below, where @fig-fixed-effects shows the odds-ratio, 95\% confidence interval for the single fixed effects of the three models.
+@fig-interaction-effects shows the odds ratio and 95\% confidence interval for the interaction effects of models 1 and 2, and @fig-interaction-effects-mod3 shows the interaction effects for model 3.
 
-So far, considering that the interaction of socioeconomic factors are important predictors of vaccination status, I believe the model that provides a more cohesive picture of the data is model 3 as it shows differences between race and income which could be important if we consider the types of occupations that people in certain regions do, and how these influences their vaccination status.
-
-<!-- @tbl-model shows the odds-ratio, 95\% confidence interval, and p-value for the logistic regression model for the first dose of vaccination by age group, race, income, Health Region, and the interaction of income and Health Region.
--->
-
-```{r, model-table}
+```{r, model-fixed-eff}
 #| message: false
+#| label: fig-fixed-effects
 #| echo: false
@@
 #| fig-width: 8
-#| fig-height: 8
+#| fig-cap: "Fixed effects of all models"
+#| fig-height: 12
 
@@
 
+# cm<-c('Incomeunder_15000'='Under CAD 15000',
+#       'Income25000_39999'='CAD 25000-39999',
+#       'Income40000_59999'='CAD 40000-59999',
+#       'Income60000_89999'='CAD 60000-89999',
+#       'Incomeover_90000'='Over CAD 90000',
+#       'Age_group25_24'='Age 25-34',
+#       'Age_group35_44'='Age 35-44',
+#       'Age_group45-54'='Age 45-54',
+#       'Age_group55_64'='Age 55-64',
+#       'Age_group65_and_over'='Age 65 and over',
+#       'Raceblack'='Black',
+#       'Raceeast_asian_pacific_islander'='East Asian/Pacific Islander',
+#       'Raceindigenous'="Indigenous",
+#       'Racelatin_american'='Latin American',
+#       'Racemixed'='mixed',
+#       'Raceother'='Other',
+#       'Racesouth_asian'='South Asian',
+#       'Racewhite_caucasian'='White Caucasian',
+#       'Health_RegionNorth East'='North East',
+#       'Health_RegionNorth West'='North West',
+#       'Health_RegionToronto'='Toronto',
+#       'Health_RegionWest'='West',
+#       
+#       )
+
 a<-modelplot(model1,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
+  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")+xlab("")
 
 b<-modelplot(model2,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("Second Model")
+  scale_color_manual(values = c("black", "red"))+ggtitle("Second Model")+xlab("")
 
@@
 
-a+b+c+plot_layout(guides = 'collect')+plot_annotation(title="Fixed effects of all models")& theme(legend.position = 'bottom')
-
-d<-modelplot(model1,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("1st Model")
-
-e<-modelplot(model2,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("2nd Model")
-
-f<-modelplot(model3,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("3rd Model")
-
-d+e+f+plot_layout(guides = 'collect')+plot_annotation(title="Interaction effects of all models") & theme(legend.position = 'bottom')
+a+b+c+plot_layout(guides = 'collect',ncol=1)+plot_annotation(title="Fixed effects of all models")& theme(legend.position = 'bottom')
 
@@
 
+```{r,model-interactions}
+#| message: false
+#| echo: false
+#| label: fig-interaction-effects
+#| warning: false
+#| fig-cap: "Interaction Effects of models 1 and 2"
+#| fig-width: 10
+#| fig-height: 10
+
+#cm<-c('Incomeunder_15000'='Under 15000')
+
+d<-modelplot(model1,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant"))+
+  scale_color_manual(values = c("black", "red"))+ggtitle("1st Model")+xlab("")
+
+e<-modelplot(model2,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("2nd Model")
+
+f<-modelplot(model3,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
+  scale_color_manual(values = c("black", "red"))+ggtitle("3rd Model")
+
+d+e+plot_layout(guides = 'collect',ncol=1)+plot_annotation(title="Interaction effects for models 1 and 2") & theme(legend.position = 'bottom')
+
+```
+
+```{r,third-model}
+#| message: false
+#| echo: false
+#| label: fig-interaction-effects-mod3
+#| warning: false
+#| fig-cap: "Interaction Effects of model 3"
+#| fig-width: 10
+#| fig-height: 15
+
+f+plot_layout(guides = 'collect',ncol=1)+plot_annotation(title="Interaction effects for model 3") & theme(legend.position = 'bottom')
+
+```
+
+**So far, considering that the interaction of socioeconomic factors are important predictors of vaccination status, I believe the model that provides a more cohesive picture of the data is model 3 as it shows differences between race and income which could be important if we consider the types of occupations that people in certain regions do, and how these influences their vaccination status.**
+
 \FloatBarrier

### 2023-01-16 protocol/Background_and_methods_Jan_12.tex
Updated

@@
 \author{Ariel Mundo Ortiz}
-\date{1/15/23}
+\date{1/16/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, breakable, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi
 

### 2023-01-23 protocol/Background_and_methods_Jan_12.qmd
Idea on changing data to run analysis

@@
 
+clean_data <-clean_data %>%
+  filter(Health_Region !="North East", Health_Region !="North West")
+
 # Each of these dataframes provide the population totals for the different variables and their categories
+
+clean_data %>%
+  tbl_summary(
+    include=c(
+    Health_Region,)
+  ) %>% bold_labels()
 
@@
 Health_Regions<-data.frame(Health_Region=c(
-  "North West",
-  "North East",
+  #"North West",
+  #"North East",
   "West",
@@
   "Toronto"),
-  Freq=c(232299,
-    557000,
+  Freq=c(#232299,
+    #557000,
     4000000,
@@
 )
+
+# probably best strategy is to remove the North West and North East regions as they only constitute 5% of the total population of Ontario.
 

### 2023-01-25 protocol/Background_and_methods_Jan_24.qmd
Working on model and results

--- 2023-01-15 protocol/Background_and_methods_Jan_12.qmd
@@
 clean_data<-clean_data %<>%
-  mutate_at(c("Income","Race","Age_group","location","Geographic_area","Geographic_area_title"),factor)
+  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)
+
 # recode the response of interest for categorical analysis
@@
   ))
+
+#get health regions to eliminate
+clean_data <-clean_data %>%
+   filter(Health_Region!="North East" , Health_Region!="North West")
+
+#eliminate the levels of the removed health regions
+clean_data <- droplevels(clean_data)
+
+## relevel for reference groups in the model
+
+ clean_data<- within(clean_data, Age_group <- relevel(Age_group, ref = "16_24"))
+ clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
+ clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
+ clean_data<- within(clean_data, Income <- relevel(Income, ref = "over_90000"))
 
@@
 # Each of these dataframes provide the population totals for the different variables and their categories
+
+# clean_data %>%
+#   tbl_summary(
+#     include=c(
+#     Health_Region,)
+#   ) %>% bold_labels()
 
@@
   "Toronto"),
-  Freq=c(232299,
-    557000,
-    4000000,
-    3700000,
-    5000000,
-    1400000
+  Freq=c(#232299,
+    #557000,
+    4095589,
+    3742520,
+    5032410,
+    1440644
   )
 )
+
+# probably best strategy is to remove the North West and North East regions as they only constitute 5% of the total population of Ontario.
 
@@
 | Variable             | Values                                                                                                                          |
-|:----------------|-------------------------------------------------------|
+|:-----------------|------------------------------------------------------|
 | Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
@@
 | Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
-                                                                                                                      |
+|                      |                                                                                                                                 |
 
@@
 Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000.
-After cleaning the dataset contained 5,247 entries (unique respondents).
+After cleaning, the dataset contained 5,247 entries (each entry corresponding to a unique respondent).
 
@@
 Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset.
-Finally, we assigned the corresponding Health Regio to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).
+Finally, we assigned the corresponding Health Region to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).
+
+Following an assessment of the number of entries corresponding to each Health Region in the clean dataset, there were a relatively low number of observations from the North East and North West regions (2.9% and 1.4% of the total clean dataset, respectively).
+Due to the low number of entries in each case, we omitted the entries from these regions from further analyses.
+Therefore, the total number of unique entries used for analysis was 3,551.
 
@@
 
-## Statistical models
+## Statistical model
 
 Because of the binary outcome of the survey answer of our interest (vaccination status as "yes" or "no") we used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location.
-Previous studies have shown that socio-economic factors, and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021;@cnat2022a].
+Previous studies have shown that socio-economic factors, and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
 At the same time in the case of Canada, others have indicated an ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].
 
-Therefore, we built different logistic regression models that accounted for different levels of interaction between the available socio-economic factors obtained from the Fields Covid survey.
-The different models are described below.
-
-### Model building and selection
-
-The first model (`model1`) examined the interaction of income and Health Region as it has been shown that there has been an increase in income inequality within the provinces of Canada over time [@marchand2020], and the relatively recent implementation of the Health Regions motivated us to explore if income disparity may be an important factor to consider within each of the Health Regions.
-
-The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking. The model appears in @eq-model1.
-
-$$
-\begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} +\\ \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income} + \beta_5\textrm{(Health Region} \times \textrm{Income)}
-\end{aligned}
-$$ {#eq-model1}
-
-Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine.
-
-The second model (`model2`) incorporated an interaction effect between age group and income to examine if a significant interaction occured between these factors as well.
-The model appears in @eq-model2.
-
-$$
-\begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \\\beta_4 \textrm{ Income}+ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)}
-\end{aligned}
-$$ {#eq-model2}
-
-A third model (`model3`) incorporated an interaction effect between race and income, and appears in @eq-model3.
+Therefore, we built a logistic regression model that accounted for interactions between the available socio-economic factors obtained from the Fields Covid survey. The model appears in @eq-model3,
 
@@
 
-Although it could have been interesting to examine the interaction of Health Region and Race, or 3-way interactions between socio-economic factors (e.g., age,income,race), these models had singularity issues.
-Next, we compared the AIC of the three models indicated above.
-The AICs are very similar and appear in @tbl-AIC.
+Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, $\beta_1...\beta_7$ indicate the coefficients for each of the fixed and interaction effects of age group, race, Health Region, and income.
+The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.
 
@@
 ## function to calculate, so it does not allow for random effects.
-
-model1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region,
-              design=a1_rake,
-              family = quasibinomial(), 
-              control= list(maxit=25))
-
-model2<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income,
-              design=a1_rake,
-              family = quasibinomial(), 
-              control= list(maxit=25))
+# 
+# model1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region,
+#               design=a1_rake,
+#               family = quasibinomial(), 
+#               control= list(maxit=25))
+# 
+# 
+# model2<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income,
+#               design=a1_rake,
+#               family = quasibinomial(), 
+#               control= list(maxit=25))
 
@@
 #summary(model1.hr)
-
-model_comp<-as_tibble(AIC(model1,model2,model3))%>%
-  select(-c(eff.p,deltabar))%>%
-  add_column(model=c("model1","model2","model3"))
-
-model_comp%>%
-  kbl(booktabs=T,
-      col.names = c("AIC","model"))
-
+# 
+# model_comp<-as_tibble(AIC(model1,model2,model3))%>%
+#   select(-c(eff.p,deltabar))%>%
+#   add_column(model=c("model1","model2","model3"))
+# 
+# model_comp%>%
+#   kbl(booktabs=T,
+#       col.names = c("AIC","model"))
+#   
 ## helps visualize the model output
@@
   modify_header(label = "**Variable**") %>%
-  as_flex_table()
+  as_kable_extra(booktabs = TRUE) %>%
+  # reduce font size to make table fit. 
+  # you may also use the `latex_options = "scale_down"` argument here.
+  kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)
 
@@
 
-## Model Comparisons
+## Vaccination status, socio-economic and geographical variables
 
-The output from the three models is too long to give a summarized view. I am still trying to get the important parts out but in the meantime here are the summary findings:
+@tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors.
+The reference groups in each case were set as follow:  16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
+Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.58), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.17) and Other Race/Ethnicity (a group  that included minorities not found elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000 and CAD 39,999.
 
-- **Model 1**
-  Significant values on:
-- Age: None
-- Race: East Asian/Pacific Islander, South Asian, White Caucasian
-- Income: 60000-89000 CAD, Over 90000
-- Interaction of Income and Health Region: Toronto and 60000-899999, North West and over_90000, North West  and under_15000
+Regarding the interaction of socio-economic and geographical factors, statistically significant odds ratios were the Central Health Region and Income of 15000 to 24999 CAD (OR=0.4), Age Group of 65 years and older with a household income between 25000 and 39999 CAD (OR=0.001), and with an income between 40000 and 59999 (OR=3.12)
 
-- **Model 2**
-    Significant values on:
-- Age: None
-- Race: East Asian/Pacific Islander, South Asian, White Caucasian
-- Health Region: East
-- Income: 60000-89000 CAD, Over 90000
-- Interaction of Income and Health Region: East and 60000_89999, Toronto and 60000-899999, North West and over_90000, North West  and under_15000
-- Interaction of Age and income: A couple of low values, but none below 0.05
-
-including here the results of the third model, which I believe is the most appropriate as it has a similar AIC to the other ones, and it is able to provide interaction information on three different instances.
-
-- **Model 3**
-    Significant values on:
-- Age: None
-- Race: Indigenous
-- Health Region: East, Toronto
-- Income: 60000-89000 CAD
-- Interaction of Income and Health Region: East and 40000_59999, East and 60000_89999, Toronto and 60000-899999, East and  over_90000, North West and over_90000, North West  and under_15000
-- Interaction of Age and income: A couple of low values, but none below 0.05
-- Interaction of Race and Income: indigenous and 25000_39999, indigenous and over_90000, mixed and over_90000, white_caucasian and over_90000
-
-So far, considering that the interaction of socioeconomic factors are important predictors of vaccination status, I believe the model that provides a more cohesive picture of the data is model 3 as it shows differences between race and income which could be important if we consider the types of occupations that people in certain regions do, and how these influences their vaccination status.
-
-<!-- @tbl-model shows the odds-ratio, 95\% confidence interval, and p-value for the logistic regression model for the first dose of vaccination by age group, race, income, Health Region, and the interaction of income and Health Region.
--->
-
-```{r, model-table}
+```{r,model-table}
 #| message: false
 #| echo: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 8
+#| label: tbl-model
+#| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"
 
- # models<-list("First Model"=model1,
- #              "Second Model"=model2,
- #              "Third Model"=model3)
-
-a<-modelplot(model1,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("First Model")
-
-b<-modelplot(model2,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("Second Model")
-
-c<-modelplot(model3,coef_omit = ':')+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("Third Model")
-
-a+b+c+plot_layout(guides = 'collect')+plot_annotation(title="Fixed effects of all models")& theme(legend.position = 'bottom')
-
-d<-modelplot(model1,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("1st Model")
-
-e<-modelplot(model2,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("2nd Model")
-
-f<-modelplot(model3,coef_omit = "^(?!.*:)")+aes(color = ifelse(p.value < 0.05, "Significant", "Not significant")) +
-  scale_color_manual(values = c("black", "red"))+ggtitle("3rd Model")
-
-d+e+f+plot_layout(guides = 'collect')+plot_annotation(title="Interaction effects of all models") & theme(legend.position = 'bottom')
-
-#  modelsummary(models,exponentiate = TRUE, stars = T, statistic = c("conf.int",
-#                            "p = {p.value}"))
+# modelsummary(model3, shape=term~response)
 # 
 # 
-#  modelsummary(model1,exponentiate = TRUE, stars = T, statistic ="conf.int")
-# 
-#  model_tbl1<-model1 %>%
-#    tbl_regression(exponentiate = TRUE) %>%
-#    bold_labels()%>%
-#    modify_header(label = "**Variable**")
-# 
-#   test<-summary(model1)
-# 
-#  model_tbl1<-model1 %>%
-#    tbl_regression(exponentiate = TRUE) %>%
-#    select(variable, estimate, p.value) %>%
-#    filter(p.value <= 0.05)%>%
-#    kbl()%>%
-#    as_kable_extra() %>%
-#    kableExtra::kable_styling(
-#      font_size = 8
+# panels <- list(
+#   "Health Regions" = list(
+#     "(I)" = lm(mpg ~ 1, data = mtcars),
+#     "(II)" = lm(mpg ~ qsec, data = mtcars)
+#   ),
+#   "Outcome: hp" = list(
+#     "(I)" = lm(hp ~ 1, data = mtcars),
+#     "(II)" = lm(hp ~ qsec, data = mtcars)
+#   )
 # )
 # 
-# 
-# 
-# # # 
-# # model_tbl2<-model2 %>%
-# #   tbl_regression(exponentiate = TRUE) %>%
-# #   bold_labels()%>%
-# #   modify_header(label = "**Variable**")%>%
-# #   as_kable_extra() %>%
-# #   kableExtra::kable_styling(
-# #     font_size = 8
-# #   )
-# # 
-# model_tbl3<-model3 %>%
-#   tbl_regression(exponentiate = TRUE) %>%
-#   bold_labels()%>%
-#   modify_header(label = "**Variable**")%>%
-#   as_kable_extra() %>%
-#   kableExtra::kable_styling(
-#     font_size = 10
-#   )
-# 
-# # model_tbl3
-# # 
-# # model_tbl3
-# # 
-# # tbl_merge_model<-
-# #   tbl_merge(
-# #     tbls = list(model_tbl1,model_tbl2,model_tbl3),
-# #     tab_spanner = c("**Model 1**","**Model 2**", "**Model 3**")
-# #   )%>%
-# #   as_flex_table()
-# # 
-# # tbl_merge_model
-# # 
-# 
-#  modelsummary(model3,exponentiate = TRUE, 
-#               stars = T, 
-#               estimate = "{estimate} [{conf.low}, {conf.high}]",
-#               #statistic = NULL
-#               )
+# modelsummary(
+#   panels,
+#   shape = "rbind",
+#   gof_map = gm)
 
+model3 %>%
+  tbl_regression(exponentiate = TRUE,
+                 label=list(Health_Region ~ "Health Region",
+                 Age_group ~ "Age Group")) %>%
+    bold_labels()%>%
+  as_kable_extra(booktabs = TRUE, longtable=T) %>%
+  # reduce font size to make table fit. 
+  # you may also use the `latex_options = "scale_down" argument here.
+  kableExtra::kable_styling(latex_options = "repeat_header")
 ```

### 2023-01-25 protocol/Background_and_methods_Jan_24.tex
Working on model and results

--- 2023-01-13 protocol/Background_and_methods_Jan_12.tex
@@
 \usepackage{xcolor}
-\usepackage{fontspec}
-\usepackage{multicol}
-\usepackage{hhline}
-\usepackage{hyperref}
 \usepackage{placeins}
@@
 \author{Ariel Mundo Ortiz}
-\date{1/13/23}
+\date{1/24/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, enhanced, breakable, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 \begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2361}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7639}}@{}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2500}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7500}}@{}}
 \caption{\label{tbl-covariates}Selected socio-economic factors from the
@@
 a household of size 1, and that reported an income above CAD 110,000.
-After cleaning the dataset contained 5,247 entries (unique respondents).
+After cleaning, the dataset contained 5,247 entries (each entry
+corresponding to a unique respondent).
 
@@
 manually added to the dataset. Finally, we assigned the corresponding
-Health Regio to each city by matching the LHINs to the Health Regions
+Health Region to each city by matching the LHINs to the Health Regions
 that now encompass them, following the information provided by Ontario
 Health (details of the complete process are in the Appendix).
+
+Following an assessment of the number of entries corresponding to each
+Health Region in the clean dataset, there were a relatively low number
+of observations from the North East and North West regions (2.9\% and
+1.4\% of the total clean dataset, respectively). Due to the low number
+of entries in each case, we omitted the entries from these regions from
+further analyses. Therefore, the total number of unique entries used for
+analysis was 3,551.
 
@@
 
-\hypertarget{statistical-models}{%
-\subsection{Statistical models}\label{statistical-models}}
+\hypertarget{statistical-model}{%
+\subsection{Statistical model}\label{statistical-model}}
 
@@
 
-Therefore, we built different logistic regression models that accounted
-for different levels of interaction between the available socio-economic
-factors obtained from the Fields Covid survey. The different models are
-described below.
-
-\hypertarget{model-building-and-selection}{%
-\subsubsection{Model building and
-selection}\label{model-building-and-selection}}
-
-The first model (\texttt{model1}) examined the interaction of income and
-Health Region as it has been shown that there has been an increase in
-income inequality within the provinces of Canada over
-time\textsuperscript{\protect\hyperlink{ref-marchand2020}{26}}, and the
-relatively recent implementation of the Health Regions motivated us to
-explore if income disparity may be an important factor to consider
-within each of the Health Regions.
-
-The the model was fitted using the function \texttt{svyglm} from the
-\texttt{survey} R package in order to incorporate the correction in
-sampling probability obtained from raking. The model appears in
-Equation~\ref{eq-model1}.
-
-\begin{equation}\protect\hypertarget{eq-model1}{}{
-\begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} +\\ \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income} + \beta_5\textrm{(Health Region} \times \textrm{Income)}
-\end{aligned}
-}\label{eq-model1}\end{equation}
-
-Where \(p\textrm{(vac)}\) indicates the probability of having received
-the first dose of a Covid-19 vaccine.
-
-The second model (\texttt{model2}) incorporated an interaction effect
-between age group and income to examine if a significant interaction
-occured between these factors as well. The model appears in
-Equation~\ref{eq-model2}.
-
-\begin{equation}\protect\hypertarget{eq-model2}{}{
-\begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \\\beta_4 \textrm{ Income}+ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)}
-\end{aligned}
-}\label{eq-model2}\end{equation}
-
-A third model (\texttt{model3}) incorporated an interaction effect
-between race and income, and appears in Equation~\ref{eq-model3}.
+Therefore, we built a logistic regression model that accounted for
+interactions between the available socio-economic factors obtained from
+the Fields Covid survey. The model appears in Equation~\ref{eq-model3},
 
@@
 
-Although it could have been interesting to examine the interaction of
-Health Region and Race, or 3-way interactions between socio-economic
-factors (e.g., age,income,race), these models had singularity issues.
-Next, we compared the AIC of the three models indicated above. The AICs
-are very similar and appear in Table~\ref{tbl-AIC}.
+Where \(p\textrm{(vac)}\) indicates the probability of having received
+the first dose of a Covid-19 vaccine, \(\beta_0\) indicates the
+population intercept, \(\beta_1...\beta_7\) indicate the coefficients
+for each of the fixed and interaction effects of age group, race, Health
+Region, and income. The the model was fitted using the function
+\texttt{svyglm} from the \texttt{survey} R package in order to
+incorporate the correction in sampling probability obtained from raking.
 
-\hypertarget{tbl-AIC}{}
 \begin{table}
-\caption{\label{tbl-AIC}AIC for the three models }\tabularnewline
 
-\centering
-\begin{tabular}[t]{rl}
-\toprule
-AIC & model\\
-\midrule
-3580.878 & model1\\
-3614.465 & model2\\
-3633.575 & model3\\
-\bottomrule
-\end{tabular}
 \end{table}
@@
 \hypertarget{tbl-descriptive-stats}{}
-\begin{table}[H]
+\begin{table}
+\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields Covid-19 Survey }\tabularnewline
 
-\providecommand{\docline}[3]{\noalign{\global\setlength{\arrayrulewidth}{#1}}\arrayrulecolor[HTML]{#2}\cline{#3}}
+\centering\begingroup\fontsize{9}{11}\selectfont
 
-\setlength{\tabcolsep}{0pt}
-
-\renewcommand*{\arraystretch}{1.5}
-
-\begin{longtable}[c]{|p{2.35in}|p{1.21in}|p{1.27in}}
-
-\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields Covid-19 Survey } \\
-
-\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Variable}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{no,\ N\ =\ 1,000}}\textcolor[HTML]{000000}{\textsuperscript{\fontsize{11}{11}\selectfont{1}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{yes,\ N\ =\ 2,709}}\textcolor[HTML]{000000}{\textsuperscript{\fontsize{11}{11}\selectfont{1}}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}\endhead
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Income}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{15000\_24999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{167\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{326\ (66\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{25000\_39999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{144\ (31\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{316\ (69\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{40000\_59999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{116\ (25\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{348\ (75\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{60000\_89999}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{71\ (17\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{343\ (83\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{over\_90000}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{251\ (25\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{752\ (75\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{under\_15000}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{251\ (29\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{624\ (71\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Age\_group}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{16\_24}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{230\ (27\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{634\ (73\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{25\_34}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{193\ (27\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{518\ (73\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{35\_44}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{136\ (25\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{400\ (75\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{45\_54}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{130\ (28\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{335\ (72\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{55\_64}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{80\ (20\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{317\ (80\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{65\_and\_over}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{231\ (31\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{505\ (69\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Health\_Region}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Central}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{224\ (28\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{581\ (72\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{East}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{135\ (23\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{448\ (77\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{North\ East}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{36\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{71\ (66\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{North\ West}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{6\ (12\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{45\ (88\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{Toronto}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{371\ (28\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{953\ (72\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{West}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{228\ (27\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{611\ (73\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{\textbf{Race}}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{arab\_middle\_eastern}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{77\ (35\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{141\ (65\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{black}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{120\ (39\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{186\ (61\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{east\_asian\_pacific\_islander}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{72\ (23\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{236\ (77\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{indigenous}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{82\ (38\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{134\ (62\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{latin\_american}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{70\ (38\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{112\ (62\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{mixed}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{111\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{213\ (66\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{other}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{134\ (34\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{255\ (66\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{south\_asian}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{89\ (23\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{294\ (77\%)}}} \\
-
-\multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\raggedright}p{\dimexpr 2.35in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{white\_caucasian}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.21in+0\tabcolsep+0\arrayrulewidth}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{245\ (18\%)}}} & \multicolumn{1}{!{\color[HTML]{000000}\vrule width 0pt}>{\centering}p{\dimexpr 1.27in+0\tabcolsep+0\arrayrulewidth}!{\color[HTML]{000000}\vrule width 0pt}}{\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{1,138\ (82\%)}}} \\
-
-\hhline{>{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}->{\arrayrulecolor[HTML]{000000}\global\arrayrulewidth=1pt}-}
-
-\multicolumn{3}{!{\color[HTML]{FFFFFF}\vrule width 0pt}>{\raggedright}p{\dimexpr 4.83in+4\tabcolsep+2\arrayrulewidth}!{\color[HTML]{FFFFFF}\vrule width 0pt}}{\textcolor[HTML]{000000}{\textsuperscript{\fontsize{11}{11}\selectfont{1}}}\textcolor[HTML]{000000}{\fontsize{11}{11}\selectfont{n\ (\%)}}} \\
-
-\end{longtable}
-
+\begin{tabular}{lcc}
+\toprule
+\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,593\\
+\midrule
+\textbf{Income} &  & \\
+\hspace{1em}over\_90000 & 238 (25\%) & 722 (75\%)\\
+\hspace{1em}15000\_24999 & 160 (34\%) & 315 (66\%)\\
+\hspace{1em}25000\_39999 & 140 (32\%) & 302 (68\%)\\
+\hspace{1em}40000\_59999 & 113 (25\%) & 334 (75\%)\\
+\addlinespace
+\hspace{1em}60000\_89999 & 67 (17\%) & 327 (83\%)\\
+\hspace{1em}under\_15000 & 240 (29\%) & 593 (71\%)\\
+\textbf{Age\_group} &  & \\
+\hspace{1em}16\_24 & 223 (27\%) & 614 (73\%)\\
+\hspace{1em}25\_34 & 186 (27\%) & 498 (73\%)\\
+\addlinespace
+\hspace{1em}35\_44 & 129 (25\%) & 385 (75\%)\\
+\hspace{1em}45\_54 & 123 (27\%) & 327 (73\%)\\
+\hspace{1em}55\_64 & 75 (20\%) & 299 (80\%)\\
+\hspace{1em}65\_and\_over & 222 (32\%) & 470 (68\%)\\
+\textbf{Health\_Region} &  & \\
+\addlinespace
+\hspace{1em}Toronto & 371 (28\%) & 953 (72\%)\\
+\hspace{1em}Central & 224 (28\%) & 581 (72\%)\\
+\hspace{1em}East & 135 (23\%) & 448 (77\%)\\
+\hspace{1em}West & 228 (27\%) & 611 (73\%)\\
+\textbf{Race} &  & \\
+\addlinespace
+\hspace{1em}white\_caucasian & 233 (18\%) & 1,080 (82\%)\\
+\hspace{1em}arab\_middle\_eastern & 76 (36\%) & 138 (64\%)\\
+\hspace{1em}black & 114 (38\%) & 184 (62\%)\\
+\hspace{1em}east\_asian\_pacific\_islander & 69 (23\%) & 234 (77\%)\\
+\hspace{1em}indigenous & 76 (40\%) & 115 (60\%)\\
+\addlinespace
+\hspace{1em}latin\_american & 69 (38\%) & 111 (62\%)\\
+\hspace{1em}mixed & 105 (34\%) & 205 (66\%)\\
+\hspace{1em}other & 128 (35\%) & 239 (65\%)\\
+\hspace{1em}south\_asian & 88 (23\%) & 287 (77\%)\\
+\bottomrule
+\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\end{tabular}
+\endgroup{}
 \end{table}
 
-\hypertarget{model-comparisons}{%
-\subsection{Model Comparisons}\label{model-comparisons}}
+\hypertarget{vaccination-status-socio-economic-and-geographical-variables}{%
+\subsection{Vaccination status, socio-economic and geographical
+variables}\label{vaccination-status-socio-economic-and-geographical-variables}}
 
-The output from the three models is too long to give a summarized view.
-I am still trying to get the important parts out but in the meantime
-here are the summary findings:
+Table~\ref{tbl-model} shows the results of the logistic regression on
+vaccination status using socio-economic, geographical factors (by Health
+Region) and their interactions as predictors. The reference groups in
+each case were set as follow: 16 to 24 years (age group), White
+Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
+Socio-economic factors with statistically significant odds ratios were
+the age group of 65 years and over (OR=0.58), persons who identified as
+Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27),
+Latin American (OR=0.24), South Asian (OR=0.17) and Other Race/Ethnicity
+(a group that included minorities not found elsewhere, OR=0.17), and
+those that lived in a household with an income bracket between CAD
+25,000 and CAD 39,999.
 
-\begin{itemize}
-\item
-  \textbf{Model 1} Significant values on:
-\item
-  Age: None
-\item
-  Race: East Asian/Pacific Islander, South Asian, White Caucasian
-\item
-  Income: 60000-89000 CAD, Over 90000
-\item
-  Interaction of Income and Health Region: Toronto and 60000-899999,
-  North West and over\_90000, North West and under\_15000
-\item
-  \textbf{Model 2} Significant values on:
-\item
-  Age: None
-\item
-  Race: East Asian/Pacific Islander, South Asian, White Caucasian
-\item
-  Health Region: East
-\item
-  Income: 60000-89000 CAD, Over 90000
-\item
-  Interaction of Income and Health Region: East and 60000\_89999,
-  Toronto and 60000-899999, North West and over\_90000, North West and
-  under\_15000
-\item
-  Interaction of Age and income: A couple of low values, but none below
-  0.05
-\end{itemize}
+Regarding the interaction of socio-economic and geographical factors,
+statistically significant odds ratios were the Central Health Region and
+Income of 15000 to 24999 CAD (OR=0.4), Age Group of 65 years and older
+with a household income between 25000 and 39999 CAD (OR=0.001), and with
+an income between 40000 and 59999 (OR=3.12)
 
-including here the results of the third model, which I believe is the
-most appropriate as it has a similar AIC to the other ones, and it is
-able to provide interaction information on three different instances.
+\hypertarget{tbl-model}{}
+\begin{longtable}{lccc}
+\caption{\label{tbl-model}Multiple Regression Analysis-Predictors of Vaccination Status }\tabularnewline
 
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Model 3} Significant values on:
-\item
-  Age: None
-\item
-  Race: Indigenous
-\item
-  Health Region: East, Toronto
-\item
-  Income: 60000-89000 CAD
-\item
-  Interaction of Income and Health Region: East and 40000\_59999, East
-  and 60000\_89999, Toronto and 60000-899999, East and over\_90000,
-  North West and over\_90000, North West and under\_15000
-\item
-  Interaction of Age and income: A couple of low values, but none below
-  0.05
-\item
-  Interaction of Race and Income: indigenous and 25000\_39999,
-  indigenous and over\_90000, mixed and over\_90000, white\_caucasian
-  and over\_90000
-\end{itemize}
+\toprule
+\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
+\midrule
+\endfirsthead
+\multicolumn{4}{@{}l}{\textit{(continued)}}\\
+\toprule
+\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
+\midrule
+\endhead
 
-So far, considering that the interaction of socioeconomic factors are
-important predictors of vaccination status, I believe the model that
-provides a more cohesive picture of the data is model 3 as it shows
-differences between race and income which could be important if we
-consider the types of occupations that people in certain regions do, and
-how these influences their vaccination status.
-
-\begin{table}
-
-\end{table}
+\endfoot
+\bottomrule
+\endlastfoot
+\textbf{Age Group} &  &  & \\
+\hspace{1em}16\_24 & — & — & \\
+\hspace{1em}25\_34 & 0.58 & 0.26, 1.28 & 0.2\\
+\hspace{1em}35\_44 & 0.72 & 0.33, 1.57 & 0.4\\
+\hspace{1em}45\_54 & 0.53 & 0.25, 1.13 & 0.10\\
+\addlinespace
+\hspace{1em}55\_64 & 0.94 & 0.42, 2.09 & 0.9\\
+\hspace{1em}65\_and\_over & 0.44 & 0.22, 0.86 & 0.017\\
+\textbf{Race} &  &  & \\
+\hspace{1em}white\_caucasian & — & — & \\
+\hspace{1em}arab\_middle\_eastern & 0.15 & 0.07, 0.35 & <0.001\\
+\addlinespace
+\hspace{1em}black & 0.17 & 0.09, 0.34 & <0.001\\
+\hspace{1em}east\_asian\_pacific\_islander & 0.83 & 0.40, 1.75 & 0.6\\
+\hspace{1em}indigenous & 0.27 & 0.11, 0.62 & 0.002\\
+\hspace{1em}latin\_american & 0.24 & 0.10, 0.58 & 0.002\\
+\hspace{1em}mixed & 0.49 & 0.23, 1.03 & 0.059\\
+\addlinespace
+\hspace{1em}other & 0.17 & 0.10, 0.31 & <0.001\\
+\hspace{1em}south\_asian & 0.40 & 0.20, 0.79 & 0.009\\
+\textbf{Health Region} &  &  & \\
+\hspace{1em}Toronto & — & — & \\
+\hspace{1em}Central & 1.12 & 0.69, 1.82 & 0.6\\
+\addlinespace
+\hspace{1em}East & 1.07 & 0.59, 1.93 & 0.8\\
+\hspace{1em}West & 1.21 & 0.72, 2.04 & 0.5\\
+\textbf{Income} &  &  & \\
+\hspace{1em}over\_90000 & — & — & \\
+\hspace{1em}15000\_24999 & 0.42 & 0.15, 1.19 & 0.10\\
+\addlinespace
+\hspace{1em}25000\_39999 & 0.23 & 0.08, 0.69 & 0.009\\
+\hspace{1em}40000\_59999 & 0.49 & 0.16, 1.43 & 0.2\\
+\hspace{1em}60000\_89999 & 0.84 & 0.26, 2.72 & 0.8\\
+\hspace{1em}under\_15000 & 0.40 & 0.16, 1.02 & 0.055\\
+\textbf{Health Region * Income} &  &  & \\
+\addlinespace
+\hspace{1em}Central * 15000\_24999 & 0.40 & 0.17, 0.97 & 0.043\\
+\hspace{1em}East * 15000\_24999 & 1.43 & 0.54, 3.74 & 0.5\\
+\hspace{1em}West * 15000\_24999 & 0.63 & 0.28, 1.43 & 0.3\\
+\hspace{1em}Central * 25000\_39999 & 0.99 & 0.42, 2.31 & >0.9\\
+\hspace{1em}East * 25000\_39999 & 0.96 & 0.36, 2.57 & >0.9\\
+\addlinespace
+\hspace{1em}West * 25000\_39999 & 0.97 & 0.41, 2.29 & >0.9\\
+\hspace{1em}Central * 40000\_59999 & 1.01 & 0.42, 2.40 & >0.9\\
+\hspace{1em}East * 40000\_59999 & 0.86 & 0.33, 2.22 & 0.8\\
+\hspace{1em}West * 40000\_59999 & 0.64 & 0.27, 1.50 & 0.3\\
+\hspace{1em}Central * 60000\_89999 & 2.92 & 0.91, 9.33 & 0.071\\
+\addlinespace
+\hspace{1em}East * 60000\_89999 & 1.58 & 0.48, 5.18 & 0.4\\
+\hspace{1em}West * 60000\_89999 & 1.48 & 0.54, 4.04 & 0.4\\
+\hspace{1em}Central * under\_15000 & 0.80 & 0.39, 1.63 & 0.5\\
+\hspace{1em}East * under\_15000 & 1.11 & 0.46, 2.73 & 0.8\\
+\hspace{1em}West * under\_15000 & 0.66 & 0.31, 1.39 & 0.3\\
+\addlinespace
+\textbf{Age Group * Income} &  &  & \\
+\hspace{1em}25\_34 * 15000\_24999 & 0.97 & 0.30, 3.09 & >0.9\\
+\hspace{1em}35\_44 * 15000\_24999 & 0.68 & 0.21, 2.19 & 0.5\\
+\hspace{1em}45\_54 * 15000\_24999 & 1.17 & 0.33, 4.10 & 0.8\\
+\hspace{1em}55\_64 * 15000\_24999 & 2.81 & 0.63, 12.5 & 0.2\\
+\addlinespace
+\hspace{1em}65\_and\_over * 15000\_24999 & 1.54 & 0.50, 4.69 & 0.5\\
+\hspace{1em}25\_34 * 25000\_39999 & 2.68 & 0.80, 8.95 & 0.11\\
+\hspace{1em}35\_44 * 25000\_39999 & 2.37 & 0.64, 8.78 & 0.2\\
+\hspace{1em}45\_54 * 25000\_39999 & 2.12 & 0.63, 7.11 & 0.2\\
+\hspace{1em}55\_64 * 25000\_39999 & 1.59 & 0.42, 6.05 & 0.5\\
+\addlinespace
+\hspace{1em}65\_and\_over * 25000\_39999 & 7.21 & 2.20, 23.6 & 0.001\\
+\hspace{1em}25\_34 * 40000\_59999 & 2.18 & 0.66, 7.19 & 0.2\\
+\hspace{1em}35\_44 * 40000\_59999 & 1.48 & 0.43, 5.04 & 0.5\\
+\hspace{1em}45\_54 * 40000\_59999 & 2.57 & 0.72, 9.13 & 0.15\\
+\hspace{1em}55\_64 * 40000\_59999 & 1.09 & 0.28, 4.18 & >0.9\\
+\addlinespace
+\hspace{1em}65\_and\_over * 40000\_59999 & 3.12 & 1.03, 9.50 & 0.045\\
+\hspace{1em}25\_34 * 60000\_89999 & 0.47 & 0.12, 1.89 & 0.3\\
+\hspace{1em}35\_44 * 60000\_89999 & 1.05 & 0.17, 6.64 & >0.9\\
+\hspace{1em}45\_54 * 60000\_89999 & 0.93 & 0.20, 4.37 & >0.9\\
+\hspace{1em}55\_64 * 60000\_89999 & 0.56 & 0.13, 2.38 & 0.4\\
+\addlinespace
+\hspace{1em}65\_and\_over * 60000\_89999 & 2.51 & 0.66, 9.56 & 0.2\\
+\hspace{1em}25\_34 * under\_15000 & 1.95 & 0.70, 5.45 & 0.2\\
+\hspace{1em}35\_44 * under\_15000 & 1.32 & 0.45, 3.84 & 0.6\\
+\hspace{1em}45\_54 * under\_15000 & 1.58 & 0.53, 4.73 & 0.4\\
+\hspace{1em}55\_64 * under\_15000 & 0.81 & 0.25, 2.68 & 0.7\\
+\addlinespace
+\hspace{1em}65\_and\_over * under\_15000 & 1.20 & 0.47, 3.04 & 0.7\\
+\textbf{Race * Income} &  &  & \\
+\hspace{1em}arab\_middle\_eastern * 15000\_24999 & 5.68 & 1.56, 20.6 & 0.008\\
+\hspace{1em}black * 15000\_24999 & 3.35 & 1.15, 9.75 & 0.026\\
+\hspace{1em}east\_asian\_pacific\_islander * 15000\_24999 & 1.82 & 0.52, 6.39 & 0.4\\
+\addlinespace
+\hspace{1em}indigenous * 15000\_24999 & 0.93 & 0.26, 3.31 & >0.9\\
+\hspace{1em}latin\_american * 15000\_24999 & 3.44 & 0.70, 16.9 & 0.13\\
+\hspace{1em}mixed * 15000\_24999 & 1.03 & 0.29, 3.67 & >0.9\\
+\hspace{1em}other * 15000\_24999 & 11.2 & 2.79, 45.2 & <0.001\\
+\hspace{1em}south\_asian * 15000\_24999 & 1.92 & 0.59, 6.24 & 0.3\\
+\addlinespace
+\hspace{1em}arab\_middle\_eastern * 25000\_39999 & 1.54 & 0.38, 6.31 & 0.5\\
+\hspace{1em}black * 25000\_39999 & 1.90 & 0.61, 5.88 & 0.3\\
+\hspace{1em}east\_asian\_pacific\_islander * 25000\_39999 & 0.63 & 0.17, 2.27 & 0.5\\
+\hspace{1em}indigenous * 25000\_39999 & 1.86 & 0.47, 7.41 & 0.4\\
+\hspace{1em}latin\_american * 25000\_39999 & 1.55 & 0.39, 6.25 & 0.5\\
+\addlinespace
+\hspace{1em}mixed * 25000\_39999 & 0.95 & 0.23, 3.95 & >0.9\\
+\hspace{1em}other * 25000\_39999 & 4.67 & 1.33, 16.4 & 0.016\\
+\hspace{1em}south\_asian * 25000\_39999 & 1.78 & 0.53, 5.98 & 0.4\\
+\hspace{1em}arab\_middle\_eastern * 40000\_59999 & 2.90 & 0.69, 12.1 & 0.14\\
+\hspace{1em}black * 40000\_59999 & 1.87 & 0.58, 6.03 & 0.3\\
+\addlinespace
+\hspace{1em}east\_asian\_pacific\_islander * 40000\_59999 & 0.42 & 0.13, 1.35 & 0.15\\
+\hspace{1em}indigenous * 40000\_59999 & 1.12 & 0.29, 4.36 & 0.9\\
+\hspace{1em}latin\_american * 40000\_59999 & 0.88 & 0.20, 3.82 & 0.9\\
+\hspace{1em}mixed * 40000\_59999 & 0.70 & 0.19, 2.64 & 0.6\\
+\hspace{1em}other * 40000\_59999 & 9.58 & 2.18, 42.1 & 0.003\\
+\addlinespace
+\hspace{1em}south\_asian * 40000\_59999 & 1.35 & 0.41, 4.45 & 0.6\\
+\hspace{1em}arab\_middle\_eastern * 60000\_89999 & 3.83 & 0.82, 17.9 & 0.088\\
+\hspace{1em}black * 60000\_89999 & 1.83 & 0.42, 7.89 & 0.4\\
+\hspace{1em}east\_asian\_pacific\_islander * 60000\_89999 & 2.59 & 0.59, 11.4 & 0.2\\
+\hspace{1em}indigenous * 60000\_89999 & 1.67 & 0.23, 12.2 & 0.6\\
+\addlinespace
+\hspace{1em}latin\_american * 60000\_89999 & 0.73 & 0.14, 3.90 & 0.7\\
+\hspace{1em}mixed * 60000\_89999 & 1.17 & 0.20, 6.90 & 0.9\\
+\hspace{1em}other * 60000\_89999 & 0.97 & 0.21, 4.45 & >0.9\\
+\hspace{1em}south\_asian * 60000\_89999 & 1.87 & 0.45, 7.74 & 0.4\\
+\hspace{1em}arab\_middle\_eastern * under\_15000 & 4.21 & 1.13, 15.7 & 0.033\\
+\addlinespace
+\hspace{1em}black * under\_15000 & 4.26 & 1.45, 12.5 & 0.008\\
+\hspace{1em}east\_asian\_pacific\_islander * under\_15000 & 1.04 & 0.35, 3.08 & >0.9\\
+\hspace{1em}indigenous * under\_15000 & 1.66 & 0.52, 5.33 & 0.4\\
+\hspace{1em}latin\_american * under\_15000 & 2.54 & 0.72, 8.98 & 0.15\\
+\hspace{1em}mixed * under\_15000 & 0.93 & 0.20, 4.26 & >0.9\\
+\addlinespace
+\hspace{1em}other * under\_15000 & 3.73 & 1.64, 8.45 & 0.002\\
+\hspace{1em}south\_asian * under\_15000 & 3.10 & 1.10, 8.74 & 0.032\\*
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\
+\end{longtable}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-marchand2020}{}}%
-\CSLLeftMargin{26. }%
-\CSLRightInline{Marchand Y, Dubé J, Breau S. Exploring the causes and
-consequences of regional income inequality in canada. \emph{Economic
-Geography}. 2020;96(2):83-107.
-doi:\href{https://doi.org/10.1080/00130095.2020.1715793}{10.1080/00130095.2020.1715793}}
-
 \end{CSLReferences}

### 2023-01-26 protocol/Background_and_methods_Jan_24.qmd
updated

@@
 @tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors.
-The reference groups in each case were set as follow:  16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
+The reference groups in each case were set as follows:  16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
 Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.58), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.17) and Other Race/Ethnicity (a group  that included minorities not found elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000 and CAD 39,999.
 
-Regarding the interaction of socio-economic and geographical factors, statistically significant odds ratios were the Central Health Region and Income of 15000 to 24999 CAD (OR=0.4), Age Group of 65 years and older with a household income between 25000 and 39999 CAD (OR=0.001), and with an income between 40000 and 59999 (OR=3.12)
+Regarding the interaction effects, the interaction of the Central Health Region and income bracket between 15000 to 24999 CAD was statistically significant (OR=0.4).
+In the age group and income interaction effect, significant effects were estimated for those 65 years and older with a household income between 25000-39999 CAD (OR=0.001), and with an income between 40000 and 59999 (OR=3.12).
+In the case of race and income, significant effects were estimated for individuals that identified as Arab/Middle Easterners with a family income between 15000 and 24000 (OR=5.68) or income under 15000 CAD (OR=4.21), Black individuals with a family income under 15000 CAD (OR=4.26) or between 15000 and 24999 CAD (OR=3.35); additionally, there were statistically significant interaction effects in the case of those that identified as belonging to a race not identified elsewhere with an income under 15000 CAD (OR=3.73), between 15000 and 24999 (OR 11.2), between 25000 and 39999 CAD (OR=4.67), or with an income between 40000 and 59999 CAD (OR=9.58).
+Finally, there was also a significant interaction effect for South Asian individuals with an income under 15000 CAD (OR=3.10).
 

### 2023-01-26 protocol/Background_and_methods_Jan_24.tex
updated

@@
 \author{Ariel Mundo Ortiz}
-\date{1/24/23}
+\date{1/25/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 Region) and their interactions as predictors. The reference groups in
-each case were set as follow: 16 to 24 years (age group), White
+each case were set as follows: 16 to 24 years (age group), White
 Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
@@
 
-Regarding the interaction of socio-economic and geographical factors,
-statistically significant odds ratios were the Central Health Region and
-Income of 15000 to 24999 CAD (OR=0.4), Age Group of 65 years and older
-with a household income between 25000 and 39999 CAD (OR=0.001), and with
-an income between 40000 and 59999 (OR=3.12)
+Regarding the interaction effects, the interaction of the Central Health
+Region and income bracket between 15000 to 24999 CAD was statistically
+significant (OR=0.4). In the age group and income interaction effect,
+significant effects were estimated for those 65 years and older with a
+household income between 25000-39999 CAD (OR=0.001), and with an income
+between 40000 and 59999 (OR=3.12). In the case of race and income,
+significant effects were estimated for individuals that identified as
+Arab/Middle Easterners with a family income between 15000 and 24000
+(OR=5.68) or income under 15000 CAD (OR=4.21), Black individuals with a
+family income under 15000 CAD (OR=4.26) or between 15000 and 24999 CAD
+(OR=3.35); additionally, there were statistically significant
+interaction effects in the case of those that identified as belonging to
+a race not identified elsewhere with an income under 15000 CAD
+(OR=3.73), between 15000 and 24999 (OR 11.2), between 25000 and 39999
+CAD (OR=4.67), or with an income between 40000 and 59999 CAD (OR=9.58).
+Finally, there was also a significant interaction effect for South Asian
+individuals with an income under 15000 CAD (OR=3.10).
 

### 2023-01-30 protocol/Background_and_methods_Jan_24.qmd
updated

--- 2023-01-25 protocol/Background_and_methods_Jan_24.qmd
@@
 @tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors.
-The reference groups in each case were set as follow:  16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
-Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.58), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.17) and Other Race/Ethnicity (a group  that included minorities not found elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000 and CAD 39,999.
+The reference groups in each case were set as follows:  16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
+Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.44), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.4) and Other Race/Ethnicity (a group  that included minorities not found elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000-39,999 (OR=0.23).
 
-Regarding the interaction of socio-economic and geographical factors, statistically significant odds ratios were the Central Health Region and Income of 15000 to 24999 CAD (OR=0.4), Age Group of 65 years and older with a household income between 25000 and 39999 CAD (OR=0.001), and with an income between 40000 and 59999 (OR=3.12)
+Regarding the interaction effects, the interaction of the Central Health Region and income bracket between 15000 to 24999 CAD was statistically significant (OR=0.4).
+In age group and income, significant effects were estimated for those 65 years and older with a household income between CAD 25000-39999 (OR=7.21), and with an income between CAD 40000-59999 (OR=3.12).
+In the case of race and income, significant effects were estimated for individuals that identified as Arab/Middle Easterners with a family income between 15000 and 24999 (OR=5.68) or income under CAD 15000 (OR=4.21), Black individuals with a family income under CAD 15000 (OR=4.26) or between CAD 15000-24999 (OR=3.35); additionally, there were statistically significant interaction effects in the case of those that identified as belonging to a race not identified elsewhere ("Other") within multiple income brackets: under CAD 15000 (OR=3.73), between 15000 and 24999 (OR 11.2), between 25000 and 39999 CAD (OR=4.67), and with income CAD 40000-59999 (OR=9.58).
+Finally, there was a significant interaction effect for South Asian individuals with an income under CAD 15000 (OR=3.10).
 

### 2023-01-30 protocol/Background_and_methods_Jan_24.tex
updated

@@
 \author{Ariel Mundo Ortiz}
-\date{1/25/23}
+\date{1/30/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, boxrule=0pt, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 Region and income bracket between 15000 to 24999 CAD was statistically
-significant (OR=0.4). In the age group and income interaction effect,
-significant effects were estimated for those 65 years and older with a
-household income between 25000-39999 CAD (OR=0.001), and with an income
-between 40000 and 59999 (OR=3.12). In the case of race and income,
-significant effects were estimated for individuals that identified as
-Arab/Middle Easterners with a family income between 15000 and 24000
-(OR=5.68) or income under 15000 CAD (OR=4.21), Black individuals with a
-family income under 15000 CAD (OR=4.26) or between 15000 and 24999 CAD
-(OR=3.35); additionally, there were statistically significant
-interaction effects in the case of those that identified as belonging to
-a race not identified elsewhere with an income under 15000 CAD
-(OR=3.73), between 15000 and 24999 (OR 11.2), between 25000 and 39999
-CAD (OR=4.67), or with an income between 40000 and 59999 CAD (OR=9.58).
-Finally, there was also a significant interaction effect for South Asian
-individuals with an income under 15000 CAD (OR=3.10).
+significant (OR=0.4). In age group and income, significant effects were
+estimated for those 65 years and older with a household income between
+25000-39999 CAD (OR=0.001), and with an income between 40000 and 59999
+(OR=3.12). In the case of race and income, significant effects were
+estimated for individuals that identified as Arab/Middle Easterners with
+a family income between 15000 and 24000 (OR=5.68) or income under 15000
+CAD (OR=4.21), Black individuals with a family income under 15000 CAD
+(OR=4.26) or between 15000 and 24999 CAD (OR=3.35); additionally, there
+were statistically significant interaction effects in the case of those
+that identified as belonging to a race not identified elsewhere with an
+income under 15000 CAD (OR=3.73), between 15000 and 24999 (OR 11.2),
+between 25000 and 39999 CAD (OR=4.67), or with an income between 40000
+and 59999 CAD (OR=9.58). Finally, in race and income there was a
+significant interaction effect for South Asian individuals with an
+income under 15000 CAD (OR=3.10).
 

### 2023-02-02 protocol/Background_and_methods_Jan_24.qmd
updated file

@@
 This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset.
-Because our interest was to analyze the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset.
-Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN), which were the geographical divisions for health in Ontario before the adoption of the Health Regions, to match each city with its corresponding LHIN.
+Due to our interest in analyzing the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset.
+Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN, the previous geographical divisions for health in Ontario before the adoption of the Health Regions), to match each city with its corresponding LHIN.
 Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset.
@@
 
-Following an assessment of the number of entries corresponding to each Health Region in the clean dataset, there were a relatively low number of observations from the North East and North West regions (2.9% and 1.4% of the total clean dataset, respectively).
-Due to the low number of entries in each case, we omitted the entries from these regions from further analyses.
-Therefore, the total number of unique entries used for analysis was 3,551.
+Following an assessment of the number of entries corresponding to each Health Region in the clean dataset, we removed entries that corresponded to the North East and North West Health Regions as they comprised only 107 observations (4.3% of the total).
+Due to the low number of entries in these regions, we omitted these Health Regions from further analyses.
+Therefore, the total number of unique entries used for analysis was 3,551 comprising the East, Central, Toronto, and West Health Regions.
 

### 2023-02-02 protocol/Background_and_methods_Jan_24.tex
updated file

--- 2023-01-25 protocol/Background_and_methods_Jan_24.tex
@@
 \author{Ariel Mundo Ortiz}
-\date{1/24/23}
+\date{1/31/23}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, breakable, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 This included the nearest municipality, which resulted in a total of 578
-different municipalities within the dataset. Because our interest was to
-analyze the differences between Health Regions, we assigned the location
-(city) of each entry to its correspondent Health Region following a
-multi-step process: First, we used the municipality information from the
-website of the Association of Municipalities of Ontario (AMO) to assign
-geographical regions to each entry on the dataset. Secondly, we used a
-publicly available dataset of long-term care homes by Local Health
-Integrated Network (LHIN), which were the geographical divisions for
-health in Ontario before the adoption of the Health Regions, to match
-each city with its corresponding LHIN. Cities that did not have a LHIN
-entry on the dataset were searched in the websites of the LHINs and
-manually added to the dataset. Finally, we assigned the corresponding
-Health Region to each city by matching the LHINs to the Health Regions
-that now encompass them, following the information provided by Ontario
-Health (details of the complete process are in the Appendix).
+different municipalities within the dataset. Due to our interest in
+analyzing the differences between Health Regions, we assigned the
+location (city) of each entry to its correspondent Health Region
+following a multi-step process: First, we used the municipality
+information from the website of the Association of Municipalities of
+Ontario (AMO) to assign geographical regions to each entry on the
+dataset. Secondly, we used a publicly available dataset of long-term
+care homes by Local Health Integrated Network (LHIN, the previous
+geographical divisions for health in Ontario before the adoption of the
+Health Regions), to match each city with its corresponding LHIN. Cities
+that did not have a LHIN entry on the dataset were searched in the
+websites of the LHINs and manually added to the dataset. Finally, we
+assigned the corresponding Health Region to each city by matching the
+LHINs to the Health Regions that now encompass them, following the
+information provided by Ontario Health (details of the complete process
+are in the Appendix).
 
 Following an assessment of the number of entries corresponding to each
-Health Region in the clean dataset, there were a relatively low number
-of observations from the North East and North West regions (2.9\% and
-1.4\% of the total clean dataset, respectively). Due to the low number
-of entries in each case, we omitted the entries from these regions from
-further analyses. Therefore, the total number of unique entries used for
-analysis was 3,551.
+Health Region in the clean dataset, we removed entries that corresponded
+to the North East and North West Health Regions as they comprised only
+107 observations (4.3\% of the total). Due to the low number of entries
+in these regions, we omitted these Health Regions from further analyses.
+Therefore, the total number of unique entries used for analysis was
+3,551 comprising the East, Central, Toronto, and West Health Regions.
 
@@
 Region) and their interactions as predictors. The reference groups in
-each case were set as follow: 16 to 24 years (age group), White
+each case were set as follows: 16 to 24 years (age group), White
 Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
 Socio-economic factors with statistically significant odds ratios were
-the age group of 65 years and over (OR=0.58), persons who identified as
+the age group of 65 years and over (OR=0.44), persons who identified as
 Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27),
-Latin American (OR=0.24), South Asian (OR=0.17) and Other Race/Ethnicity
+Latin American (OR=0.24), South Asian (OR=0.4) and Other Race/Ethnicity
 (a group that included minorities not found elsewhere, OR=0.17), and
 those that lived in a household with an income bracket between CAD
-25,000 and CAD 39,999.
+25,000-39,999 (OR=0.23).
 
-Regarding the interaction of socio-economic and geographical factors,
-statistically significant odds ratios were the Central Health Region and
-Income of 15000 to 24999 CAD (OR=0.4), Age Group of 65 years and older
-with a household income between 25000 and 39999 CAD (OR=0.001), and with
-an income between 40000 and 59999 (OR=3.12)
+Regarding the interaction effects, the interaction of the Central Health
+Region and income bracket between 15000 to 24999 CAD was statistically
+significant (OR=0.4). In age group and income, significant effects were
+estimated for those 65 years and older with a household income between
+CAD 25000-39999 (OR=7.21), and with an income between CAD 40000-59999
+(OR=3.12). In the case of race and income, significant effects were
+estimated for individuals that identified as Arab/Middle Easterners with
+a family income between 15000 and 24999 (OR=5.68) or income under CAD
+15000 (OR=4.21), Black individuals with a family income under CAD 15000
+(OR=4.26) or between CAD 15000-24999 (OR=3.35); additionally, there were
+statistically significant interaction effects in the case of those that
+identified as belonging to a race not identified elsewhere (``Other'')
+within multiple income brackets: under CAD 15000 (OR=3.73), between
+15000 and 24999 (OR 11.2), between 25000 and 39999 CAD (OR=4.67), and
+with income CAD 40000-59999 (OR=9.58). Finally, there was a significant
+interaction effect for South Asian individuals with an income under CAD
+15000 (OR=3.10).
 

### 2023-02-06 01-background.qmd
Re-organized files and re-wrote background

# Background

The COVID-19 pandemic continues around the world with more than 600 million confirmed cases as of November of 2022 [@WHO-Covid]. During the first months of the pandemic in early 2020, non-pharmaceutical interventions such as masking and social distancing were the only methods available to manage the spread of the disease; however, the rapid development of vaccines permitted their approval and use in some countries towards the last month part of 2020. For example, in the US and Canada vaccine campaigns began in mid-December of 2020 [@tanne2020; @bogoch2022]. Although it has been estimated that vaccines against COVID-19 have prevented around 14 millions of deaths worldwide [@watson2022], the rollout of COVID-19 vaccines has faced multiple challenges since its inception.

Indeed, multiple obstacles that have complicated vaccination efforts against COVID-19 have been identified: inequality in vaccine access, vaccine hesitancy, and differences in vaccination rates across different segments of the population [@gerretsen2021; @nafilyan2021; @malik2020]. In the case of Canada, lower vaccine uptake has been associated with socio-economic factors such as younger age, educational level, presence of children in the household, lack of a regular healthcare provider, ethnic origin, and financial instability [@guay2022;@muhajarine2021;@carter2022;@hussain2022].

Additionally, vaccination is influenced by changes in geography. In this regard, it has been shown that there have been spatial differences in COVID-19 vaccination rates due to regional differences in attitudes towards vaccination [@malik2020], geographical differences in vaccine access and supply, vaccination location availability, and lack of prioritization of vulnerable groups [@bogoch2022; @nguyen2021].

Studies that analyze geographical variations in vaccine uptake are important as they can help inform public health decision-makers to design policies to that consider spatial changes to address vaccination disparities. In this regard, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governmental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].

In Canada, studies that have used a spatial approach to analyze vaccine uptake have shown disparities in vaccination rates across low and high income neighborhoods in the city of Toronto [@choi2021], among adolescents from deprived neighborhoods in the city of Montreal [@mckinnon2021], and have also highlighted disparities in vaccination status depending on age, income, and ethnic origin at the provincial level [@guay2022]. However, there is limited information on differences in vaccination status inside the provinces. Such analysis is important as it can help identify inequalities that may exist within these geographical areas while providing a granular view of intra-provincial differences that can help understand the barriers for vaccine uptake.

In this study we examined self-reported COVID-19 vaccination status within the province of Ontario using a combination of socio-economic factors(such as ethnic origin, age, and income) and geographical analysis (at the level of the Health Regions)  in order to determine intra-provincial disparities in vaccine uptake and address the ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].

### 2023-02-06 03-results.qmd
Re-organized files and re-wrote background

--- 2023-01-30 protocol/Background_and_methods_Jan_24.qmd
@@
----
-title: "Covid-19 Fields Survey Data Protocol"
-author:
-  - name: Ariel Mundo Ortiz
-    orcid: 0000-0002-6014-4538
-    email: ariel.mundo.ortiz@umontreal.ca
-    affiliations:
-      - name: École de santé publique - Université de Montréal
-        address: 7101 Park Avenue
-        city: Montreal
-        state: QC
-        postal-code: H3N 1X9
-date: "`r Sys.Date()`"
-format:
-  pdf:
-    link-citations: true
-    number-sections: true
-    include-in-header: 
-      text: |
-        \usepackage{placeins}
-  html:
-    toc: true
-editor: visual
-bibliography: bibliography/refs.bib
-keep-tex: true
-csl: bibliography/jama.csl
-knitr:
-  opts_chunk:
-    dev: "ragg_png"
----
+# Results
 
@@
 ```
-
-# Background
-
-The COVID-19 pandemic continues around the world with more than 600 million confirmed cases as of November of 2022 [@WHO-Covid].
-During the first months of the pandemic in early 2020, non-pharmaceutical interventions (e.g., masking, social distancing) were the only methods available to manage the spread of the disease, but the rapid development of vaccines against the virus permitted their approval and use in some countries towards the last month part of 2020.
-For example, in the US and Canada vaccine campaigns began in mid-December of 2020 [@tanne2020; @bogoch2022].
-Although it has been estimated that vaccines against COVID-19 have prevented around 14 millions of deaths worldwide [@watson2022], the rollout of COVID-19 vaccines has faced multiple challenges since its inception.
-
-In this regard, vaccination efforts have faced multiple: Inequalities with regard to vaccine access due to socio-economic factors, vaccine hesitancy, and differences in vaccination rates across different segments of the population are among the challenges identified in the administration of COVID-19 vaccines [@gerretsen2021; @nafilyan2021; @malik2020].
-In the case of Canada, lower vaccine uptake has been associated with socio-economic factors such as younger age, educational level, presence of children in the household, lack of a regular healthcare provider, ethnic origin, and financial instability [@guay2022; @muhajarine2021; @carter2022].
-
-Additionally, it has been shown that geography also plays a crucial role in vaccination rates, as they vary due to spatial differences in attitudes towards vaccination [@malik2020], geographical differences in vaccine access and supply, vaccination location availability, and lack of prioritization of vulnerable groups [@bogoch2022; @nguyen2021].
-
-Studies that analyze geographical variations in vaccine uptake can help inform public health decision-makers to design policies to that are aimed at addressing vaccination disparities.
-In this regard, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governmental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].
-
-In Canada, studies that have used a spatial approach to analyze vaccine uptake have shown disparities in vaccination rates across low and high income neighborhoods in the city of Toronto [@choi2021], among adolescents from deprived neighborhoods in the city of Montreal [@mckinnon2021], and highlighted disparities in vaccination status depending on age, income, and ethnic origin in all of the Canadian provinces[@guay2022].
-However, to the best of our knowledge, there are no studies that have analyzed vaccination status within a province to identify inequalities that may exist within these geographical areas.
-A dissagregated view of variability within a province can help understand the barriers for vaccine delivery in the case of visible minorities, which have been disproportionately impacted in the pandemic [@hussain2022].
-
-# Research Question
-
-This study will examine self-reported COVID-19 vaccination status within the province of Ontario in order to determine how socio-economic (e.g., ethnic origin, age, income) and geographical factors (at level of the Health Regions of Ontario) influence vaccination within the province.
-
-# Methods
-
-## Data source: survey overview
-
-We obtained data from the Fields Institute for Research in Mathematical Sciences' (henceforth Fields) *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario which ran from Sept 30, 2021 until January 17, 2022.
-This survey was commissioned by Fields and the Mathematical Modelling of COVID-19 Task Force, under the supervision of Dr.
-Kumar Murty, the Director of Fields with funding from the Canadian Institutes of Health Research.
-The survey was conducted by a third-party service provider (RIWI Corp.), under ethical guidance from University of Toronto.
-
-The survey was deployed using random domain intercept technology.
-Briefly, when web users clicked on a registered but commercially inactive web link or typed in a web address for a site that was dormant, they had a random chance of that link being temporarily managed by the company that administered the survey (RIWI Corp).
-Thus, instead of coming across a notification about the status of the site("this page does not exist"), the survey was deployed to the user.
-Web users then decided whether to anonymously participate, exiting the survey at any time if desired [@sargent2022].
-
-Respondents who wished to participate were asked to select their age from a matrix of values, and subsequent questions were displayed one at a time, after the respondent confirmed their selection by answering and selecting "next".
-Those who do not wished to participate were asked to either close the browser window or navigate away from the domain.
-After the survey closed (complete or incomplete) no one from that internet protocol (IP) address could access the survey again and the domain entry point rotated such that if a respondent were to attempt to access the survey again, share the link, or enter via the same address using an alternative IP address, the survey would not render.
-
-Additionally, respondents who indicated they were under the age of 16 were exited from the survey.
-No record was created in this case and due to domain cycling these users were unable to navigate back to the "age select" screen.
-The personal identifier information from each respondent was automatically scrubbed and replaced by a unique ID.
-Respondents were drawn exclusively from the province of Ontario, as per their devices meta-data.
-
-## Survey responses
-
-### Socio-demographic factors {#sec-socio-demographic-factors}
-
-From the different answers provided by the survey respondents, we selected the age group which they belonged to, income bracket, race/ethnicity, and employment status.
-The original survey included additional questions (e.g., sick leave, remote work, presence of minors in the household) but the survey design, which permitted respondents to exit the survey at any point resulted in a high rates of missing data for most of these answers.
-The socio-economic factors chosen for this study were the ones that had both the lowest rates of missingness and that provided an adequate level socio-economic and demographic information for our analysis.
-Information about the chosen socio-economic factors from the survey is provided in @tbl-covariates.
-
-| Variable             | Values                                                                                                                          |
-|:-----------------|------------------------------------------------------|
-| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
-| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
-| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
-|                      |                                                                                                                                 |
-
-: Selected socio-economic factors from the survey {#tbl-covariates}
-
-### Vaccination status
-
-From the survey, we selected the question regarding vaccination status:
-
--   "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no"
-
-<!-- -   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no" -->
-
-<!-- -   (If answer was "no" to the first question) "If a vaccine was made available to you you would:" with possible answers "definitely get", "definitely not get", "probably get", and "probably not get" -->
-
-<!--           - definitely get -->
-
-<!--           - definitely not get -->
-
-<!--           - probably get -->
-
-<!--           - probably not get -->
-
-## Data cleaning
-
-The original dataset contained 39,029 entries (where each entry corresponded to a set of answers provided by a unique respondent).
-Following a preliminary analysis to identify the missing rates across the different answers within each entry, it was identified that many of the answers had high missing rates (\>80%) (note that the graph will be in the Appendix).
-Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variable.
-The final clean dataset contained 3,709 unique entries.
-
-The cleaning process also included removing outliers that were identified during the preliminary analyses.
-Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000.
-After cleaning, the dataset contained 5,247 entries (each entry corresponding to a unique respondent).
-
-### Geographical location {#sec-geographical-location}
-
-For each survey participant certain data was automatically captured.
-This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset.
-Because our interest was to analyze the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset.
-Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN), which were the geographical divisions for health in Ontario before the adoption of the Health Regions, to match each city with its corresponding LHIN.
-Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset.
-Finally, we assigned the corresponding Health Region to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).
-
-Following an assessment of the number of entries corresponding to each Health Region in the clean dataset, there were a relatively low number of observations from the North East and North West regions (2.9% and 1.4% of the total clean dataset, respectively).
-Due to the low number of entries in each case, we omitted the entries from these regions from further analyses.
-Therefore, the total number of unique entries used for analysis was 3,551.
-
-## Corrections
-
-We identified differences between the proportions of various socio-economic factors the survey respondents when compared to the 2016 Census data for Ontario.
-These factors included age groups, income, and ethnicity/race identified.
-Additionally, because the Census divisions do not match the exact boundaries of the Census, we also obtained population estimates for each Health Region from the Ontario Health website in order to correct for the population within each Health Region.
-We corrected for all of these factors (age group, race/ethnicity, income, Health Region Population) using an iterative proportional fitting procedure (also known as *raking*) [@deming1940] in R using the `survey` package.
-The proportions of each of these factors and the data from the 2016 Census Data for Ontario used for the corrections can be found in the Appendix.
-Of notice, because the categories provided by the survey in some cases (e.g., race/ethnicity categories) did not match the categories from the Census, we aggregated them where appropriate to obtain an approximation to the categories in the Census.
-
-<!-- ## Geographical location {#sec-geographical-location} -->
-
-<!-- For each survey participant certain data was automatically captured.
-This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset.
-Due to our interest in exploring vaccination status within a health region that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities].
-The geographical regions were obtained from the websie of the province of Ontario <https://www.ontario.ca/page/list-ontario-municipalities#section-3> -->
-
-## Statistical model
-
-Because of the binary outcome of the survey answer of our interest (vaccination status as "yes" or "no") we used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location.
-Previous studies have shown that socio-economic factors, and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
-At the same time in the case of Canada, others have indicated an ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].
-
-Therefore, we built a logistic regression model that accounted for interactions between the available socio-economic factors obtained from the Fields Covid survey. The model appears in @eq-model3,
-
-$$
-\begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)} + \beta_7 \textrm{ (Race} \times \textrm{Income)}
-\end{aligned}
-$$ {#eq-model3}
-
-Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, $\beta_1...\beta_7$ indicate the coefficients for each of the fixed and interaction effects of age group, race, Health Region, and income.
-The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.
-
-```{r,models}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: tbl-AIC
-#| tbl-cap: "AIC for the three models"
-
-### Model ###
-
-## the first model is a simple logistic regression with fixed effects only svyglm uses the glm
-## function to calculate, so it does not allow for random effects.
-# 
-# model1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-# 
-# 
-# model2<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-
-model3<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income+Race*Income,
-              design=a1_rake,
-              family = quasibinomial(), 
-              control= list(maxit=25))
-
-# model4<-svyglm(first_dose_m~Age_group+Income+Race+Health_Region+Health_Region*Income+Income*Age_group+Income*Race+Income*Age_group*Race,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-# 
-# 
-# model5<-svyglm(first_dose_m~Age_group+Income+Race+Health_Region+Health_Region*Income+Income*Age_group+Income*Race+Race*Health_Region,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-
-#summary(model1.test)
-#summary(model1.hr)
-# 
-# model_comp<-as_tibble(AIC(model1,model2,model3))%>%
-#   select(-c(eff.p,deltabar))%>%
-#   add_column(model=c("model1","model2","model3"))
-# 
-# model_comp%>%
-#   kbl(booktabs=T,
-#       col.names = c("AIC","model"))
-#   
-## helps visualize the model output
-
-###############
-
-```
-
-# Results
-
-## Descriptive statistics of the Fields Covid-19 Survey
-
-@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed.
-
-```{r,summary-table}
-#| label: tbl-descriptive-stats
-#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey"
-#| message: false
-#| echo: false
-
-data_summary %>%
-  modify_header(label = "**Variable**") %>%
-  as_kable_extra(booktabs = TRUE) %>%
-  # reduce font size to make table fit. 
-  # you may also use the `latex_options = "scale_down"` argument here.
-  kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)
-
-```
-
-## Vaccination status, socio-economic and geographical variables
-
-@tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors.
-The reference groups in each case were set as follows:  16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
-Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.44), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.4) and Other Race/Ethnicity (a group  that included minorities not found elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000-39,999 (OR=0.23).
-
-Regarding the interaction effects, the interaction of the Central Health Region and income bracket between 15000 to 24999 CAD was statistically significant (OR=0.4).
-In age group and income, significant effects were estimated for those 65 years and older with a household income between CAD 25000-39999 (OR=7.21), and with an income between CAD 40000-59999 (OR=3.12).
-In the case of race and income, significant effects were estimated for individuals that identified as Arab/Middle Easterners with a family income between 15000 and 24999 (OR=5.68) or income under CAD 15000 (OR=4.21), Black individuals with a family income under CAD 15000 (OR=4.26) or between CAD 15000-24999 (OR=3.35); additionally, there were statistically significant interaction effects in the case of those that identified as belonging to a race not identified elsewhere ("Other") within multiple income brackets: under CAD 15000 (OR=3.73), between 15000 and 24999 (OR 11.2), between 25000 and 39999 CAD (OR=4.67), and with income CAD 40000-59999 (OR=9.58).
-Finally, there was a significant interaction effect for South Asian individuals with an income under CAD 15000 (OR=3.10).
-
-```{r,model-table}
-#| message: false
-#| echo: false
-#| label: tbl-model
-#| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"
-
-# modelsummary(model3, shape=term~response)
-# 
-# 
-# panels <- list(
-#   "Health Regions" = list(
-#     "(I)" = lm(mpg ~ 1, data = mtcars),
-#     "(II)" = lm(mpg ~ qsec, data = mtcars)
-#   ),
-#   "Outcome: hp" = list(
-#     "(I)" = lm(hp ~ 1, data = mtcars),
-#     "(II)" = lm(hp ~ qsec, data = mtcars)
-#   )
-# )
-# 
-# modelsummary(
-#   panels,
-#   shape = "rbind",
-#   gof_map = gm)
-
-model3 %>%
-  tbl_regression(exponentiate = TRUE,
-                 label=list(Health_Region ~ "Health Region",
-                 Age_group ~ "Age Group")) %>%
-    bold_labels()%>%
-  as_kable_extra(booktabs = TRUE, longtable=T) %>%
-  # reduce font size to make table fit. 
-  # you may also use the `latex_options = "scale_down" argument here.
-  kableExtra::kable_styling(latex_options = "repeat_header")
-```
-
-\FloatBarrier
-
-## References
-
-::: {#refs}
-:::

### 2023-02-06 main.qmd
Re-organized files and re-wrote background

--- 2023-02-06 03-results.qmd
@@
-# Results
+---
+title: '**Socio-economic and geographic determinants of Covid-19 vaccination**'
+author: Ariel Mundo Ortiz
+format: pdf
+bibliography: refs.bib
+csl: jama.csl
+link-citations: yes #adds links to the citations
+---
 
-```{r, load-prepare-data}
+# Abstract
+
+# Keywords {-}
+
+```{r setup, message = FALSE,include=FALSE}
 #| echo: false
@@
 ```
+```
+
+```{r} 
+#| child: 01-background.qmd
+```
+
+<!-- \FloatBarrier -->
+
+```{r} 
+#| child: 02-methods.qmd
+```
+
+# References

### 2023-02-06 manuscript/02-methods.qmd
Re-organized files and re-wrote background

# Methods

## Data source: survey overview

We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario which ran from Sept 30, 2021 until January 17, 2022 and that was comissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force.  The survey was conducted by a third-party service provider (RIWI Corp.), under ethical guidance from the University of Toronto.

Briefly, the survey was deployed used random domain intercept technology, where if users clicked on a registered but commercially inactive web link or typed in a web address for a site that was dormant, they had a random chance of that link being temporarily managed by the company that administered the survey and instead of coming across a notification about the status of the site("this page does not exist"), the survey was deployed to the user [@sargent2022]. Users then decided whether to anonymously participate, and those that participated were able to exit the survey at any time. After the survey closed, regardless if it was complete or incomplete, access was denied to any further users with the same internet protocol (IP) address and the domain entry point rotated such that if a user were to attempt to access the survey again, share the link, or enter via the same address using an alternative IP address, the survey would not deploy. This effectively meant that a user could participate only once in the survey.

Additionally, users who indicated they were under the age of 16 were exited from the survey, no record was created, and these users were unable to navigate back to the "age select" screen. The personal identifier information from each user was automatically scrubbed and replaced by a unique ID. Respondents were drawn exclusively from the province of Ontario, as per their devices meta-data.

## Survey responses

### Socio-demographic factors {#sec-socio-demographic-factors}

Socio-economic information selected for analysis from the survey answers included age group, income bracket, and race/ethnicity. Information about these socio-economic factors in the survey is provided in @tbl-covariates.

| Variable             | Values                                                                                                                          |
|:------------------|-----------------------------------------------------|
| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, Other |
|                      |                                                                                                                                 |

: Selected socio-economic factors from the survey {#tbl-covariates}

Furthermore, the information on vaccination status was provided by survey participants who answered the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".

<!-- -   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no" -->

<!-- -   (If answer was "no" to the first question) "If a vaccine was made available to you you would:" with possible answers "definitely get", "definitely not get", "probably get", and "probably not get" -->

<!--           - definitely get -->

<!--           - definitely not get -->

<!--           - probably get -->

<!--           - probably not get -->

## Data cleaning

The original dataset contained 39,029 entries (where each entry corresponded to a set of answers provided by a unique respondent). Following a preliminary analysis, the dataset was cleaned in order to only contain the socio-economic information provided in @tbl-covariates and vaccination status. The cleaning process also included removing outliers that were identified during the preliminary analyses. Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000. After cleaning, the dataset contained 3,709 unique entries.

### Geographical location {#sec-geographical-location}

For each survey participant certain data was automatically captured, including the nearest municipality (city). This resulted in a total of 578 different municipalities within the clean dataset. Due to our interest in analyzing the differences between Health Regions, we assigned the city of each entry to its correspondent Health Region following a multi-step process. Briefly, we used Local Health Integrated Networks (LHINs) to assign a Health Region to each entry in the survey. LHINs were the geographical divisions for health used by  Ontario before the adoption of the Health Regions; because of the lack of a publicly available list of all municipalities within each Health Region, we used a dataset of long-term care homes and LHINs to match each city to LHIN, followed by matching each LHIN to a Health Region following the information provided on the Ontario Health Website, where the list of LHINs and corresponding Health Regions is available. In the case of municipalities that did not appear in the long-term dataset, we manually searched each city in the LHINs websites in order to provide geographical information. Details of the complete data cleaning process are in the Appendix.

Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions. Due to the low number of entries, we omitted these Health Regions from further analyses. Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions.

## Corrections

We identified differences between the proportions of all the socio-economic factors included in the analysis (@tbl-covariates) and the 2016 Canada Census data for Ontario. Additionally, because the Census divisions do not match the exact boundaries of the Health Regions, we also obtained population estimates for each Health Region from the Ontario Health website in order to correct for the population totals. We used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct for socio-economic factors and Health Region populations using the R `survey` package. Details about the correction can be found in the Appendix.

## Statistical analyses

We used a fixed-effects logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a]. The model appears in @eq-model3,

$$
\begin{aligned}
\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)} + \beta_7 \textrm{ (Race} \times \textrm{Income)}
\end{aligned}
$$ {#eq-model3}

Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, $\beta_1...\beta_7$ indicate the coefficients for each of the fixed and interaction effects of age group, race, Health Region, and income. The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.

All analyses were conducted in RStudio (2022.12.0 Build 353), using R 4.2.2 and the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], and `quarto` [@quarto].

### 2023-02-06 manuscript/03-results.qmd
Re-organized files and re-wrote background

--- 2023-01-30 protocol/Background_and_methods_Jan_24.qmd
@@
----
-title: "Covid-19 Fields Survey Data Protocol"
-author:
-  - name: Ariel Mundo Ortiz
-    orcid: 0000-0002-6014-4538
-    email: ariel.mundo.ortiz@umontreal.ca
-    affiliations:
-      - name: École de santé publique - Université de Montréal
-        address: 7101 Park Avenue
-        city: Montreal
-        state: QC
-        postal-code: H3N 1X9
-date: "`r Sys.Date()`"
-format:
-  pdf:
-    link-citations: true
-    number-sections: true
-    include-in-header: 
-      text: |
-        \usepackage{placeins}
-  html:
-    toc: true
-editor: visual
-bibliography: bibliography/refs.bib
-keep-tex: true
-csl: bibliography/jama.csl
-knitr:
-  opts_chunk:
-    dev: "ragg_png"
----
-
-```{r, load-prepare-data}
-#| echo: false
-#| warning: false
-#| message: false
-
-#load packages
-
-library(weights)
-library(here)
-library(survey)
-library(magrittr)
-library(tidyverse)
-library(lme4)
-library(kableExtra)
-library(modelsummary)
-library(gtsummary)
-library(patchwork)
-
-data<-read.csv(here("data","clean_dataset.csv"))
-
-# remove rows with missing observations for race
-
-clean_data<-data%>%drop_na(race)
-
-#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
-
-clean_data<-clean_data %>%
-  select(-Municipal.status) %>%
-  rename(Race=race,Age_group=age_group,Income=income_ord)
-
-#make variables of interest factors
-clean_data<-clean_data %<>%
-  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)
-
-# recode the response of interest for categorical analysis
-clean_data<-clean_data %>%
-  mutate(first_dose_m=case_when(
-    first_dose=="yes"~1,
-    first_dose=="no"~0
-  ))
-
-#get health regions to eliminate
-clean_data <-clean_data %>%
-   filter(Health_Region!="North East" , Health_Region!="North West")
-
-#eliminate the levels of the removed health regions
-clean_data <- droplevels(clean_data)
-
-## relevel for reference groups in the model
-
- clean_data<- within(clean_data, Age_group <- relevel(Age_group, ref = "16_24"))
- clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
- clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
- clean_data<- within(clean_data, Income <- relevel(Income, ref = "over_90000"))
-
-# get percentages to see how different the data is from the Census (details of the Census values are in
-# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")
-
-##### Summary of the dataset####
-
-data_summary<-clean_data %>%
-  tbl_summary(percent="row",
-    by= first_dose,
-    include=c(Income,
-    Age_group,
-    Health_Region,
-    Race)
-  ) %>% bold_labels()
-
-# Each of these dataframes provide the population totals for the different variables and their categories
-
-# clean_data %>%
-#   tbl_summary(
-#     include=c(
-#     Health_Region,)
-#   ) %>% bold_labels()
-
-Ages<-data.frame(Age_group=c("16_24",
-                             "25_34",
-                             "35_44",
-                             "45_54",
-                             "55_64",
-                             "65_and_over"),
-                 Freq=c(1707959,
-                        1734856,
-                        1721407,
-                        2003826,
-                        1842444,
-                        2245898))
-
-Races<-data.frame(Race=c("arab_middle_eastern",
-                         "black",
-                         "east_asian_pacific_islander",
-                         "indigenous",
-                         "latin_american",
-                         "mixed",
-                         "other",
-                         "south_asian",
-                         "white_caucasian"),
-                  Freq=c(212782,
-                         638346,
-                         886592,
-                         376558,
-                         197020,
-                         130033,
-                         705333,
-                         1166361,
-                         9118079 ))
-
-Incomes<-data.frame(Income=c("under_15000",
-                                     "15000_24999",
-                                     "25000_39999",
-                                     "40000_59999",
-                                     "60000_89999",
-                                     "over_90000"),
-                        Freq=c(294643,
-                               387688,
-                               599624,
-                               796052,
-                               1007988,
-                               2083176
-
-                        ))
-
-# corrections for health regions
-
-Health_Regions<-data.frame(Health_Region=c(
-  #"North West",
-  #"North East",
-  "West",
-  "East",
-  "Central",
-  "Toronto"),
-  Freq=c(#232299,
-    #557000,
-    4095589,
-    3742520,
-    5032410,
-    1440644
-  )
-)
-
-# probably best strategy is to remove the North West and North East regions as they only constitute 5% of the total population of Ontario.
-
-### Raking ####
-
-## First, provide a survey design object. This is done using the dataset. The syntax
-## means that there is no stratification (id=~1). When the line below is run it will throw
-## a warning because we do not have "design" weights in the dataset (other datasets do have this).
-
-a1<-svydesign(id=~1,data=clean_data)
-
-## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage.
-
-#a1$prob #all probs are 1
-
-## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
-## from the data frames from above.
-
-a1_rake<-rake(a1,
-             sample=list(~Age_group,
-                         ~Race,
-                         ~Income,
-                         ~Health_Region),
-             population=list(Ages,
-                             Races,
-                             Incomes,
-                             Health_Regions))
-
-```
-
-# Background
-
-The COVID-19 pandemic continues around the world with more than 600 million confirmed cases as of November of 2022 [@WHO-Covid].
-During the first months of the pandemic in early 2020, non-pharmaceutical interventions (e.g., masking, social distancing) were the only methods available to manage the spread of the disease, but the rapid development of vaccines against the virus permitted their approval and use in some countries towards the last month part of 2020.
-For example, in the US and Canada vaccine campaigns began in mid-December of 2020 [@tanne2020; @bogoch2022].
-Although it has been estimated that vaccines against COVID-19 have prevented around 14 millions of deaths worldwide [@watson2022], the rollout of COVID-19 vaccines has faced multiple challenges since its inception.
-
-In this regard, vaccination efforts have faced multiple: Inequalities with regard to vaccine access due to socio-economic factors, vaccine hesitancy, and differences in vaccination rates across different segments of the population are among the challenges identified in the administration of COVID-19 vaccines [@gerretsen2021; @nafilyan2021; @malik2020].
-In the case of Canada, lower vaccine uptake has been associated with socio-economic factors such as younger age, educational level, presence of children in the household, lack of a regular healthcare provider, ethnic origin, and financial instability [@guay2022; @muhajarine2021; @carter2022].
-
-Additionally, it has been shown that geography also plays a crucial role in vaccination rates, as they vary due to spatial differences in attitudes towards vaccination [@malik2020], geographical differences in vaccine access and supply, vaccination location availability, and lack of prioritization of vulnerable groups [@bogoch2022; @nguyen2021].
-
-Studies that analyze geographical variations in vaccine uptake can help inform public health decision-makers to design policies to that are aimed at addressing vaccination disparities.
-In this regard, previous geographical (spatial) analyses of vaccination rates have shown that variations in vaccine uptake can occur within small governmental administrative units (e.g., counties in the case of the US) [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that geographical analyses can be predictive of booster uptake patterns [@wood2022].
-
-In Canada, studies that have used a spatial approach to analyze vaccine uptake have shown disparities in vaccination rates across low and high income neighborhoods in the city of Toronto [@choi2021], among adolescents from deprived neighborhoods in the city of Montreal [@mckinnon2021], and highlighted disparities in vaccination status depending on age, income, and ethnic origin in all of the Canadian provinces[@guay2022].
-However, to the best of our knowledge, there are no studies that have analyzed vaccination status within a province to identify inequalities that may exist within these geographical areas.
-A dissagregated view of variability within a province can help understand the barriers for vaccine delivery in the case of visible minorities, which have been disproportionately impacted in the pandemic [@hussain2022].
-
-# Research Question
-
-This study will examine self-reported COVID-19 vaccination status within the province of Ontario in order to determine how socio-economic (e.g., ethnic origin, age, income) and geographical factors (at level of the Health Regions of Ontario) influence vaccination within the province.
-
-# Methods
-
-## Data source: survey overview
-
-We obtained data from the Fields Institute for Research in Mathematical Sciences' (henceforth Fields) *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario which ran from Sept 30, 2021 until January 17, 2022.
-This survey was commissioned by Fields and the Mathematical Modelling of COVID-19 Task Force, under the supervision of Dr.
-Kumar Murty, the Director of Fields with funding from the Canadian Institutes of Health Research.
-The survey was conducted by a third-party service provider (RIWI Corp.), under ethical guidance from University of Toronto.
-
-The survey was deployed using random domain intercept technology.
-Briefly, when web users clicked on a registered but commercially inactive web link or typed in a web address for a site that was dormant, they had a random chance of that link being temporarily managed by the company that administered the survey (RIWI Corp).
-Thus, instead of coming across a notification about the status of the site("this page does not exist"), the survey was deployed to the user.
-Web users then decided whether to anonymously participate, exiting the survey at any time if desired [@sargent2022].
-
-Respondents who wished to participate were asked to select their age from a matrix of values, and subsequent questions were displayed one at a time, after the respondent confirmed their selection by answering and selecting "next".
-Those who do not wished to participate were asked to either close the browser window or navigate away from the domain.
-After the survey closed (complete or incomplete) no one from that internet protocol (IP) address could access the survey again and the domain entry point rotated such that if a respondent were to attempt to access the survey again, share the link, or enter via the same address using an alternative IP address, the survey would not render.
-
-Additionally, respondents who indicated they were under the age of 16 were exited from the survey.
-No record was created in this case and due to domain cycling these users were unable to navigate back to the "age select" screen.
-The personal identifier information from each respondent was automatically scrubbed and replaced by a unique ID.
-Respondents were drawn exclusively from the province of Ontario, as per their devices meta-data.
-
-## Survey responses
-
-### Socio-demographic factors {#sec-socio-demographic-factors}
-
-From the different answers provided by the survey respondents, we selected the age group which they belonged to, income bracket, race/ethnicity, and employment status.
-The original survey included additional questions (e.g., sick leave, remote work, presence of minors in the household) but the survey design, which permitted respondents to exit the survey at any point resulted in a high rates of missing data for most of these answers.
-The socio-economic factors chosen for this study were the ones that had both the lowest rates of missingness and that provided an adequate level socio-economic and demographic information for our analysis.
-Information about the chosen socio-economic factors from the survey is provided in @tbl-covariates.
-
-| Variable             | Values                                                                                                                          |
-|:-----------------|------------------------------------------------------|
-| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
-| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
-| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, other |
-|                      |                                                                                                                                 |
-
-: Selected socio-economic factors from the survey {#tbl-covariates}
-
-### Vaccination status
-
-From the survey, we selected the question regarding vaccination status:
-
--   "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no"
-
-<!-- -   (If answered "yes" to the previous question) "Have you received the second dose of the COVID vaccine?" with possible answers "yes" and "no" -->
-
-<!-- -   (If answer was "no" to the first question) "If a vaccine was made available to you you would:" with possible answers "definitely get", "definitely not get", "probably get", and "probably not get" -->
-
-<!--           - definitely get -->
-
-<!--           - definitely not get -->
-
-<!--           - probably get -->
-
-<!--           - probably not get -->
-
-## Data cleaning
-
-The original dataset contained 39,029 entries (where each entry corresponded to a set of answers provided by a unique respondent).
-Following a preliminary analysis to identify the missing rates across the different answers within each entry, it was identified that many of the answers had high missing rates (\>80%) (note that the graph will be in the Appendix).
-Therefore, the dataset was cleaned in order to contain only the independent variables of interest with the lowest missing rates (@tbl-covariates) and the dependent variable.
-The final clean dataset contained 3,709 unique entries.
-
-The cleaning process also included removing outliers that were identified during the preliminary analyses.
-Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000.
-After cleaning, the dataset contained 5,247 entries (each entry corresponding to a unique respondent).
-
-### Geographical location {#sec-geographical-location}
-
-For each survey participant certain data was automatically captured.
-This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset.
-Because our interest was to analyze the differences between Health Regions, we assigned the location (city) of each entry to its correspondent Health Region following a multi-step process: First, we used the municipality information from the website of the Association of Municipalities of Ontario (AMO) to assign geographical regions to each entry on the dataset.
-Secondly, we used a publicly available dataset of long-term care homes by Local Health Integrated Network (LHIN), which were the geographical divisions for health in Ontario before the adoption of the Health Regions, to match each city with its corresponding LHIN.
-Cities that did not have a LHIN entry on the dataset were searched in the websites of the LHINs and manually added to the dataset.
-Finally, we assigned the corresponding Health Region to each city by matching the LHINs to the Health Regions that now encompass them, following the information provided by Ontario Health (details of the complete process are in the Appendix).
-
-Following an assessment of the number of entries corresponding to each Health Region in the clean dataset, there were a relatively low number of observations from the North East and North West regions (2.9% and 1.4% of the total clean dataset, respectively).
-Due to the low number of entries in each case, we omitted the entries from these regions from further analyses.
-Therefore, the total number of unique entries used for analysis was 3,551.
-
-## Corrections
-
-We identified differences between the proportions of various socio-economic factors the survey respondents when compared to the 2016 Census data for Ontario.
-These factors included age groups, income, and ethnicity/race identified.
-Additionally, because the Census divisions do not match the exact boundaries of the Census, we also obtained population estimates for each Health Region from the Ontario Health website in order to correct for the population within each Health Region.
-We corrected for all of these factors (age group, race/ethnicity, income, Health Region Population) using an iterative proportional fitting procedure (also known as *raking*) [@deming1940] in R using the `survey` package.
-The proportions of each of these factors and the data from the 2016 Census Data for Ontario used for the corrections can be found in the Appendix.
-Of notice, because the categories provided by the survey in some cases (e.g., race/ethnicity categories) did not match the categories from the Census, we aggregated them where appropriate to obtain an approximation to the categories in the Census.
-
-<!-- ## Geographical location {#sec-geographical-location} -->
-
-<!-- For each survey participant certain data was automatically captured.
-This included the nearest municipality, which resulted in a total of 578 different municipalities within the dataset.
-Due to our interest in exploring vaccination status within a health region that encompasses multiple cities, we grouped the municipalities from the dataset according to the regional government aggregations provided by the Association of Municipalities of Ontario, which groups municipalities within regions, counties, districts, and single-tiers [@ontario-municipalities].
-The geographical regions were obtained from the websie of the province of Ontario <https://www.ontario.ca/page/list-ontario-municipalities#section-3> -->
-
-## Statistical model
-
-Because of the binary outcome of the survey answer of our interest (vaccination status as "yes" or "no") we used a fixed-effects logistic regression to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location.
-Previous studies have shown that socio-economic factors, and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
-At the same time in the case of Canada, others have indicated an ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].
-
-Therefore, we built a logistic regression model that accounted for interactions between the available socio-economic factors obtained from the Fields Covid survey. The model appears in @eq-model3,
-
-$$
-\begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)} + \beta_7 \textrm{ (Race} \times \textrm{Income)}
-\end{aligned}
-$$ {#eq-model3}
-
-Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, $\beta_1...\beta_7$ indicate the coefficients for each of the fixed and interaction effects of age group, race, Health Region, and income.
-The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.
-
-```{r,models}
-#| echo: false
-#| warning: false
-#| message: false
-#| label: tbl-AIC
-#| tbl-cap: "AIC for the three models"
-
-### Model ###
-
-## the first model is a simple logistic regression with fixed effects only svyglm uses the glm
-## function to calculate, so it does not allow for random effects.
-# 
-# model1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-# 
-# 
-# model2<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-
-model3<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income+Race*Income,
-              design=a1_rake,
-              family = quasibinomial(), 
-              control= list(maxit=25))
-
-# model4<-svyglm(first_dose_m~Age_group+Income+Race+Health_Region+Health_Region*Income+Income*Age_group+Income*Race+Income*Age_group*Race,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-# 
-# 
-# model5<-svyglm(first_dose_m~Age_group+Income+Race+Health_Region+Health_Region*Income+Income*Age_group+Income*Race+Race*Health_Region,
-#               design=a1_rake,
-#               family = quasibinomial(), 
-#               control= list(maxit=25))
-
-#summary(model1.test)
-#summary(model1.hr)
-# 
-# model_comp<-as_tibble(AIC(model1,model2,model3))%>%
-#   select(-c(eff.p,deltabar))%>%
-#   add_column(model=c("model1","model2","model3"))
-# 
-# model_comp%>%
-#   kbl(booktabs=T,
-#       col.names = c("AIC","model"))
-#   
-## helps visualize the model output
-
-###############
-
-```
-
 # Results
 
-## Descriptive statistics of the Fields Covid-19 Survey
+## Survey Results
 
 @tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed.
+The total number of entries from the survey in the dataset after cleaning was 3,551.
+Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it.
+Regarding socio-economic factors, respondents who reported living in a household with an income under CAD 15,000 or above 90,000 represented approximately 50\% of the total number of entries, with 71\% and 75\% of the respondents in each income bracket indicating to have received the first dose of the vaccine, respectively.
+The age group between 16-24 years had the highest representation in the survey responses (837, 23.6\% of all responses), followed by the age group of those between 65 and over (692, 19.5\% of the total); 73\% of those in the 16 to 24 years age group reported to have received the vaccine, whereas 68\% of those above 65 years indicated the same.
+The Health Region with the highest number of participants in the survey was Toronto, accounting for 1,342 entries (37.7\%) and a 72\% vaccination rate.
+Regarding race, individuals that identified as White/Caucasian represented 1313 (37\%) of all entries and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine.
+On the other hand, those that identified as Indigenous had the lowest vaccine uptake with 60\% indicating to have received the vaccine.
 
@@
 #| echo: false
+
+source(here::here("code","data_preparation.R"))
 
@@
 
-## Vaccination status, socio-economic and geographical variables
+## Multivariate Regression
 
@@
 #| message: false
+#| warning: false
 #| echo: false
@@
 #| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"
+source(here::here("code","raking_logistic_regression.R"))
 
-# modelsummary(model3, shape=term~response)
-# 
-# 
-# panels <- list(
-#   "Health Regions" = list(
-#     "(I)" = lm(mpg ~ 1, data = mtcars),
-#     "(II)" = lm(mpg ~ qsec, data = mtcars)
-#   ),
-#   "Outcome: hp" = list(
-#     "(I)" = lm(hp ~ 1, data = mtcars),
-#     "(II)" = lm(hp ~ qsec, data = mtcars)
-#   )
-# )
-# 
-# modelsummary(
-#   panels,
-#   shape = "rbind",
-#   gof_map = gm)
-
-model3 %>%
+m1 %>%
   tbl_regression(exponentiate = TRUE,
@@
 ```
-
-\FloatBarrier
-
-## References
-
-::: {#refs}
-:::

### 2023-02-06 manuscript/main.qmd
Re-organized files and re-wrote background

--- 2023-02-06 main.qmd
@@
 ```{r setup, message = FALSE,include=FALSE}
-#| echo: false
-#| warning: false
-#| message: false
-
-#load packages
+#the packages needed to run the code
 
@@
 
-data<-read.csv(here("data","clean_dataset.csv"))
-
-# remove rows with missing observations for race
-
-clean_data<-data%>%drop_na(race)
-
-#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
-
-clean_data<-clean_data %>%
-  select(-Municipal.status) %>%
-  rename(Race=race,Age_group=age_group,Income=income_ord)
-
-#make variables of interest factors
-clean_data<-clean_data %<>%
-  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)
-
-# recode the response of interest for categorical analysis
-clean_data<-clean_data %>%
-  mutate(first_dose_m=case_when(
-    first_dose=="yes"~1,
-    first_dose=="no"~0
-  ))
-
-#get health regions to eliminate
-clean_data <-clean_data %>%
-   filter(Health_Region!="North East" , Health_Region!="North West")
-
-#eliminate the levels of the removed health regions
-clean_data <- droplevels(clean_data)
-
-## relevel for reference groups in the model
-
- clean_data<- within(clean_data, Age_group <- relevel(Age_group, ref = "16_24"))
- clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
- clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
- clean_data<- within(clean_data, Income <- relevel(Income, ref = "over_90000"))
-
-# get percentages to see how different the data is from the Census (details of the Census values are in
-# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")
-
-##### Summary of the dataset####
-
-data_summary<-clean_data %>%
-  tbl_summary(percent="row",
-    by= first_dose,
-    include=c(Income,
-    Age_group,
-    Health_Region,
-    Race)
-  ) %>% bold_labels()
-
-# Each of these dataframes provide the population totals for the different variables and their categories
-
-# clean_data %>%
-#   tbl_summary(
-#     include=c(
-#     Health_Region,)
-#   ) %>% bold_labels()
-
-Ages<-data.frame(Age_group=c("16_24",
-                             "25_34",
-                             "35_44",
-                             "45_54",
-                             "55_64",
-                             "65_and_over"),
-                 Freq=c(1707959,
-                        1734856,
-                        1721407,
-                        2003826,
-                        1842444,
-                        2245898))
-
-Races<-data.frame(Race=c("arab_middle_eastern",
-                         "black",
-                         "east_asian_pacific_islander",
-                         "indigenous",
-                         "latin_american",
-                         "mixed",
-                         "other",
-                         "south_asian",
-                         "white_caucasian"),
-                  Freq=c(212782,
-                         638346,
-                         886592,
-                         376558,
-                         197020,
-                         130033,
-                         705333,
-                         1166361,
-                         9118079 ))
-
-Incomes<-data.frame(Income=c("under_15000",
-                                     "15000_24999",
-                                     "25000_39999",
-                                     "40000_59999",
-                                     "60000_89999",
-                                     "over_90000"),
-                        Freq=c(294643,
-                               387688,
-                               599624,
-                               796052,
-                               1007988,
-                               2083176
-
-                        ))
-
-# corrections for health regions
-
-Health_Regions<-data.frame(Health_Region=c(
-  #"North West",
-  #"North East",
-  "West",
-  "East",
-  "Central",
-  "Toronto"),
-  Freq=c(#232299,
-    #557000,
-    4095589,
-    3742520,
-    5032410,
-    1440644
-  )
-)
-
-# probably best strategy is to remove the North West and North East regions as they only constitute 5% of the total population of Ontario.
-
-### Raking ####
-
-## First, provide a survey design object. This is done using the dataset. The syntax
-## means that there is no stratification (id=~1). When the line below is run it will throw
-## a warning because we do not have "design" weights in the dataset (other datasets do have this).
-
-a1<-svydesign(id=~1,data=clean_data)
-
-## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage.
-
-#a1$prob #all probs are 1
-
-## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
-## from the data frames from above.
-
-a1_rake<-rake(a1,
-             sample=list(~Age_group,
-                         ~Race,
-                         ~Income,
-                         ~Health_Region),
-             population=list(Ages,
-                             Races,
-                             Incomes,
-                             Health_Regions))
-
-```
 ```
@@
 
+```{r}
+#| child: 03-results.qmd
+```
+
 # References
+
+::: {#refs}
+:::

### 2023-02-06 protocol/Background_and_methods_Jan_24.qmd
Re-organized files and re-wrote background

@@
 | Variable             | Values                                                                                                                          |
-|:-----------------|------------------------------------------------------|
+|:-------------------|----------------------------------------------------|
 | Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |

### 2023-02-06 code/data_preparation.R

@@ -1,0 +1,51 @@
+## this script takes the clean data set and does some preparatory steps such as recoding
+## variables to create a table summary
+
+data<-read.csv(here("data","clean_dataset.csv"))
+
+# remove rows with missing observations for race
+
+clean_data<-data%>%drop_na(race)
+
+#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
+
+clean_data<-clean_data %>%
+  select(-Municipal.status) %>%
+  rename(Race=race,Age_group=age_group,Income=income_ord)
+
+#make variables of interest factors
+clean_data<-clean_data %<>%
+  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)
+
+# recode the response of interest for categorical analysis
+clean_data<-clean_data %>%
+  mutate(first_dose_m=case_when(
+    first_dose=="yes"~1,
+    first_dose=="no"~0
+  ))
+
+#get health regions to eliminate
+clean_data <-clean_data %>%
+  filter(Health_Region!="North East" , Health_Region!="North West")
+
+
+#eliminate the levels of the removed health regions
+clean_data <- droplevels(clean_data)
+
+## relevel for reference groups in the model
+
+clean_data<- within(clean_data, Age_group <- relevel(Age_group, ref = "16_24"))
+clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
+clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
+clean_data<- within(clean_data, Income <- relevel(Income, ref = "over_90000"))
+
+#create summary table
+
+data_summary<-clean_data %>%
+  tbl_summary(percent="row",
+              by= first_dose,
+              include=c(Income,
+                        Age_group,
+                        Health_Region,
+                        Race)
+  ) %>% bold_labels()

### 2023-02-06 code/raking_logistic_regression.R

@@ -1,0 +1,111 @@
+#Each of these dataframes provides the population totals for the different variables and their categories
+
+# clean_data %>%
+#   tbl_summary(
+#     include=c(
+#     Health_Region,)
+#   ) %>% bold_labels()
+
+
+Ages<-data.frame(Age_group=c("16_24",
+                             "25_34",
+                             "35_44",
+                             "45_54",
+                             "55_64",
+                             "65_and_over"),
+                 Freq=c(1707959,
+                        1734856,
+                        1721407,
+                        2003826,
+                        1842444,
+                        2245898))
+
+
+
+Races<-data.frame(Race=c("arab_middle_eastern",
+                         "black",
+                         "east_asian_pacific_islander",
+                         "indigenous",
+                         "latin_american",
+                         "mixed",
+                         "other",
+                         "south_asian",
+                         "white_caucasian"),
+                  Freq=c(212782,
+                         638346,
+                         886592,
+                         376558,
+                         197020,
+                         130033,
+                         705333,
+                         1166361,
+                         9118079 ))
+
+Incomes<-data.frame(Income=c("under_15000",
+                             "15000_24999",
+                             "25000_39999",
+                             "40000_59999",
+                             "60000_89999",
+                             "over_90000"),
+                    Freq=c(294643,
+                           387688,
+                           599624,
+                           796052,
+                           1007988,
+                           2083176
+                           
+                    ))
+
+
+# corrections for Health Regions, without the North East and North West regions
+
+Health_Regions<-data.frame(Health_Region=c(
+  #"North West",
+  #"North East",
+  "West",
+  "East",
+  "Central",
+  "Toronto"),
+  Freq=c(#232299,
+    #557000,
+    4095589,
+    3742520,
+    5032410,
+    1440644
+  )
+)
+
+
+
+### Raking ####
+
+## First, provide a survey design object. This is done using the dataset. The syntax
+## means that there is no stratification (id=~1). When the line below is run it will throw
+## a warning because we do not have "design" weights in the dataset (other datasets do have this).
+
+
+a1<-svydesign(id=~1,data=clean_data)
+
+## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage. 
+
+
+## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
+## from the data frames from above.
+
+a1_rake<-rake(a1,
+              sample=list(~Age_group,
+                          ~Race,
[…]

### 2023-02-07 data_cleaning/Data_cleaning_final.qmd
created file with details of data cleaning

--- 2023-01-11 data_cleaning/Data_cleaning_and_covariates_Jan_09.qmd
@@
 ---
-title: "Fields COVID-19 dataset exploration"
+title: "**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**"
+subtitle: Data Cleaning
 author: "Ariel Mundo"
@@
 
-This dataset is about vaccination and other COVID measures in different cities in Ontario. The covariates available are:
+The original Fields Covid-19 survey contained information about Covid-19 vaccination status and other COVID measures in different cities in Ontario. Information provided by respondents included:
 
@@
 -   Number of people from household that attend school
--   Chronic illnesses within the household (*I will not analyze this one as the question is too open and asks about age and disease at the same time*)
+-   Chronic illnesses within the household same time
 -   Race
@@
 
-There are many responses from the survey on the dataset, but there are some that I would be interesed in analyzing:
+Respondents provided multiple answers regarding vaccination:
 
@@
 
-**Note: As of Dec 30, 2022 I no longer intend to analyze this last question as after cleaning the dataset it has a missing rate of about 76% and therefore it has too few observations compared to the first two questions**.
-
 # Preliminary Analyses
 
-This document focuses on some preliminary analyses of the data I undertook to assess its representativeness. I performed five major tasks for this preliminary analysis:
-
--   Choosing covariates to assess their missigness rates
--   Cleaning the data in order to keep all the covariates with a 0% missigness rate
--   Race and ethnicity representativeness of the clean data
--   Trends in the survey answers by geographical location
--   Trends in survey answers by group age
-
-Overall the goal of this analysis was to see what adjustments would need to be made to formally analyze the data.
+This document focuses on exploratory analyses, covariate selection, and outlier identification and removal. The goal of the exploratory analyses was to clean the dataset for formal analyses.
 
 As way of comparison, I used the 2016 census data for Ontario.
+
+## Data Loading
+
+The first task was to load the raw data, extract the majority of the responses from the survey that could be used for regression, and if they were categorical, make them factors to do the exploratory analysis.
 
@@
 
-#set the time and date fields in the format that is required by lubridate
-
-data$time_date_code<-ymd(data$time_date_code)
-```
-
-```{r, rename-columns,echo=FALSE,warning=FALSE}
-
-#get a new dataframe and extract variables of interest and make them factors
-
-#get the variables of interest, convert to factors those required
+#get the variables of interest, convert to factors when required
 d1<-data%>%select(
@@
 
-The next step is to see from the covariates what are the ones with the highest number of missing observations, in order to decide what can we include in the model.
+The next step consisted in identify the missing rates of the covariates, and determinewhich of those could be included in the model.
+The following code chunk creates a table with the number of missing observations and percentages.
 
@@
 
-However, there are a couple of things we want to consider: The answer to "sick_leave" was answered only by those that reported to be employed (the survey design made this response conditional).
+However, per the dictionary in the original dataset, "sick_leave" was answered only by those that reported to be employed (the survey design made this response conditional).
 Therefore, those unemployed would be excluded in an analysis that considers this variable.
 
-At the very least, it would be interesting to analyze how the reported status of vaccination changed over geographical location and time by:
+Therefore, we decided to select the following covariates from the original dataset:
 
@@
 From these covariates, the one with the highest missing rate of observations is the employment status.
-I next cleaned the data in order to have complete observations about employment, and to see how the missing rates looked for the other covariates.
+Next, we cleaned the data in order to have complete observations about employment, and to see how the missing rates looked for the other covariates.
+The following code chunk creates histograms for the different covariates and a table with missing rates in the the `clean_data` data frame which was created from the raw data.
 
@@
                                    ))%>%
- #drop_na(age,income,h_size)%>%
   ggplot(aes(x=age))+
@@
                                    ))%>%
-  #drop_na(age,income,h_size_ord)%>%
   ggplot(aes(x=age))+
@@
 e<-clean_data%>%
-  #drop_na(age,income,h_size_ord)%>%
   ggplot(aes(x=age,fill=employed))+
   geom_histogram(stat="count",binwidth=10)+
-  #ylim(c(0,200))+
 facet_wrap(~income)+
@@
   ggtitle("Clean dataset")+
-  #theme(strip.text.x = element_text(size = 8))+
   thm1
@@
 f<-d1%>%
-  #drop_na(age,income,h_size_ord)%>%
   ggplot(aes(x=age,fill=employed))+
@@
   ggtitle("Original dataset")+
-  #theme(strip.text.x = element_text(size = 8))+
   thm1
@@
 
-There seem to be outliers where people \<25 y of age report having an income \>110k while living in a household of 1. Explore the reported household composition of those \<25 with income\>110k.
+At this stage in the analysis, we identified that outliers were present in the data.
+Specifically, individuals that were under 25 years of age and that reported having an income \>110k while living in a household of 1.
+We therefore, created some diagnostic plots to explore the reported household composition of those \<25 with income\>110k, and we removed the outliers from the dataset.
 
@@
 
-This table shows something interesting due to the nature of the dataset: the missing observations across variables are different, which is why now, although in the original dataset there were more observations of "race" than "employment", when we select complete observations for the latter we lose some observations for race.
-The second step of cleaning would be to remove the missing observations from race as well, and see what the proportions of covariates and responses are when compared to the census data.
+The next code chunk creates another table with covariate missing rates in the clean dataset.
+It can be seen that at this stage the answer about the first dose of the vaccine has no missing observations, the answer about having received the second dose of the vaccine has 27% of missing observations, and the answer about if people would get a vaccine has 76% of missing observations.
 
@@
 
-It can be seen that after removing all missing observations in the covariates, the only response that still has missing observations is the answers for the second dose of the vaccine, with 27% of missing observations.
-
 ## Race and Ethnicity
 
-Once the data was clean, I explored how the race and ethnicity information from the dataset compared to the data from the Census.
+Next, we explored how the race and ethnicity information from the dataset compared to the data from the Census. The following chunk creates a summary table for the Race variable.
 
@@
 
-Because the categories for race/ethnicity in the survey did not match the categories from the Canadian Census, we used a combination of sources to obtain the corresponding percentages to match the survey categories.
-Sources are indicated below:
+It is important to mention that the categories for race/ethnicity provided in the survey did not match the categories used in the Census.
+Therefore, we used a combination of sources to obtain the information presented in @tbl-races, where the categories used could be matched to the Census Information.
+The data sources for the table were:
 
@@
 | **Total**      | 99.9%      |
-: Reference Data from Race/Ethnicity
+: Reference Data from Race/Ethnicity {#tbl-races}
 
@@
 
-Another important aspect to consider is how the trends in the responses look over time by age group. This is because from the census data, the age groups for the province are as follows:
+According to the Census, the distribution of the different the age groups for the province of Ontario are as follows:
 
@@
 
-However, when plotted the proportion of answers at each time point among groups, the distribution looks rather different.
+To compare the survey data to the Census, the next chunk creates a barplot of the age groups in the survey.
+Here, it can be noticed that the age-group distribution from the survey data is different from the Census data.
 
@@
 
-It can be seen from the graph that the proportion of answers by group age do not follow the overall distribution from the province.
-
 ## Income
 
-The question from the survey was "What is your household annual income?".
-To compare the results of this answer, we used the "Household total income groups in 2015 for private households" from the Census data (available in the Census Data for Ontario website [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Income&TABID=1&type=1)).
+Survey respondents answered the question "What is your household annual income?".
+To compare the distribution of responses to this question from the survey and the Census data, we used the "Household total income groups in 2015 for private households" from the Census data (available in the Census Data for Ontario website [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Income&TABID=1&type=1)).
 
@@
 
-The brackets for income in the census data are different than the brackets used in the survey.
+One difference that is noticeable is that the brackets for income in the census data are different than the brackets used in the survey.
 The census does CAD 4,999 brackets (e.g., CAD 5,000- CAD 9,9999) up to CAD 49,999, followed by CAD 9,999 brackets up to CAD 99,999.
-After that, the brackets increase to CAD 24,999, and therefore, one cannot obtain percentages for the 90,000-109,999 and \>110,000 brackets from the survey.
+After that, the brackets increase to CAD 24,999, and therefore, it is not possible to obtain percentages for the 90,000-109,999 and \>110,000 brackets from the survey.
 
-Therefore, an additional category for income will be created in the dataset so it matches the information from the Census.
+Therefore, in the following code chunk, we created an additional category for income in the dataset that matched the information from the Census, and barplots to visualize the proportion of each income bracket in the dataset.
 
@@
 
-Again, there are differences here in the patterns of response.
-The \<15,000 bracket accounts for about 20% of the responses, a much higher rate than the percentage they represent from the census, and there is variation within the other brackets as well.
+We identified differences again between the survey data and the Census data.
+For example, the \<15,000 bracket accounts for about 25% of the responses, a much higher rate than the proportion in the Census.
+The other brackets were different as well from the Census data.
 
-## Geographical location
+## Geographical Information
 
-Next, we are interested in analyzing the geographical distribution of the answers the participants within the geographical areas defined by the Association of Municipalities from Ontario.
-
-**Update: the populations by census division/city were found on Jan 8, 2023 (from https://www12.statcan.gc.ca/census-recensement/2021/dp-pd/prof/index.cfm?Lang=E).
-The correction will be done using these numbers instead.
-The numbers can be found in `census_population_by_division.csv` in `data` directory.**
+Because each of the respondents of the survey was assigned a geographical location (city), we were interested in accounting for geographical location in our analysis.
+We used a multi-step process to assign geographical information to the entries in the dataset.
 
@@
 
-1) Add to the dataset the data for the municipalities so each city has the corresponding region it belongs to.
-1) Determine percentage of answers that correspond to Toronto.
-This is important because although the Census regions do not match the exact geographical boundaries of the municipalities and thus a direct correction for this cannot be easily obtained, Toronto is the largest urban area in the province and it is important to identify if corrections are needed with regard to the proportions from this place in the survey.
+1) Assign to each entry in the dataset municipalities the geographical region it belongs to using municipality and geographical region information.
 
-### Joining the municipalities data
+2) Assign a Health Region to each survey entry using the geographical and Local Integrated Health Network (LHIN) information.
+
+The details of each step are outlined below.
+
+### Municipality Data
 
@@
 
-Need to explore which entries were left without a geographical region:
+The following chunk identified which entries were left without a geographical region:
 
@@
 
-There are 2744 observations that did not get a geographical region, and they correspond to 187 unique cities.
-These will be exported to a csv file in order to manually write their geographical areas following the Association of Municipalities of Ontario divisions, and using Wikipedia to check the status of each municipality.
+This analysis identified that 2744 entries did not get a geographical region.
+These 2744 entries corresponded to 187 unique cities.
+These cities without a region were exported to a csv file in order to manually write their geographical areas following the Association of Municipalities of Ontario divisions, using Wikipedia to check the status of each municipality.
+The following chunk created the csv file.
+The code is commented now as it was run once.
 
@@
 
-The code chunk above was ran once to get the names of the municipalities.
-After manually entering the geographic areas, the file was saved as `missing_municipalities_updated.csv`, and this will be the file to be used for the next steps.
+After searching and manually entering the geographic area for each city, the file was saved as `missing_municipalities_updated.csv`, and this file was used for the next steps.
 
-Note: there is one city "Kinburn", but there are two communities with such name, one in Huron County and one in Carelton County, for the time being Huron County is assigned, will check later if is better to remove it.
-
-Also one occurence of "Sydenham", which can be a ward in Kingston or a community in Frotenac. Will assign Frotenac for the time being.
+Note: there is one city "Kinburn", but there are two communities with such name, one in Huron County and one in Carelton County, we assigned it to Huron County.
+In the case of "Sydenham", which can be a ward in Kingston or a community in Frotenac, we assigned it to Frotenac.
 
@@
 
-After manually assigning the geographical regions to the municipalities that were missing (which can be found in `missing_municipalities.csv` in the `data` directory), these will be merged as the dataset `geographic_areas.csv` (also in `data`), which contains the titles for each region (e.g., "County", "Region").
+After manually assigning the geographical regions to the municipalities that were missing (which can be found in `missing_municipalities.csv` in the `data` directory), these were merged as the dataset `geographic_areas.csv` (also in `data`), which contains the titles for each region (e.g., "County", "Region").
 
@@
 
-## Proportion of answers within Toronto
+### Health Regions
 
-**This section needs to be modified as the correction has been done by geographical area on Jan 8, 2023**
+We sought to geographically analyze the information in the survey using the The Health Regions of Ontario.
+However, these Health Regions do not match the divisions from the census, and there is no publicly available dataset from Health Ontario that lists each municipality and its corresponding Health Region.
+We used therefore a multi-stage approach to incorporate the information into the dataset:
 
-The final step is of the geographical analysis is to determine the proportion of answers from Toronto.
+- First, we used the dataset from Paul Allen regarding long-term care homes in Ontario (https://paulallen.ca/consolidated-dataset-of-ltc-homes-in-ontario/) to obtain information about communities and the Local Health Integration Network (LHINs) where long-term care homes were located.
 
-```{r,plot-location}
-#| echo: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 4
-#| fig-cap: Percentage of survey answers by city
+- Second, using the LHIN information, we added the Health Region each entry corresponded to using the information on LHIN and Health Region correspondence, which can be found here: https://www.ontariohealth.ca/about-us/our-people.
 
-clean_data<-clean_data%>%
-  mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
-                            city!="Toronto"~ "All other cities"
-                          )))%>%
-  mutate(location=as.factor(location))
-
-ggplot(clean_data %>% count(location) %>%    # then count number in each group
-         mutate(pct=n/sum(n),               # Calculate percent within each group
-                ypos = 200),  # Calculate label positions
-       aes(location, n, fill=location)) +
-  geom_bar(stat="identity") +
-  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
-  theme_classic()+
-  thm1
-
-```
-
-The proportion of responses from the city of Toronto is 31.3% of the total in the clean dataset.
-From the (Census data)[https://www.toronto.ca/wp-content/uploads/2022/02/92e3-City-Planning-2021-Census-Backgrounder-Population-Dwellings-Backgrounder.pdf], the city of Toronto as of 2021 has a population of 2,794,356, corresponding to 19.6% of the total population of the Province.
-
-# Health Regions approach
-
-Following on a suggestion by BN on Jan 09, the idea is to use broader regions for geographical analysis.
-The Health Regions of Ontario will be used.
-However, these Health Regions do not match the census divisions.
-I have not been able to find a dataset from Health Ontario that lists each municipality and its corresponding Health Region.
-I will use a multi-stage approach to incorporate the information into the dataset:
-
-- Get dataset from Paul Allen regarding long-term care homes in Ontario (https://paulallen.ca/consolidated-dataset-of-ltc-homes-in-ontario/).
-This dataset has communities and the Local Health Integration Network (LHINs) where long-term care homes were located.
-I will keep the community and LHIN information.
-
-- Add column for Health Region based on LHIN: The Health Regions substituted LHINs, a list of their correspondence can be found here: https://www.ontariohealth.ca/about-us/our-people.
-I will add this information.
-
-- Merge the dataset and check if there are missing municipalities. Manually add missing municipalities.
-
-## Steps
+- Third, after merging the dataset, we manually added LHINs to those municipalities that did not have an entry at this stage.
 
 The dataset from Paul Allen was downloaded and saved as `Consolidated_LTC_dataset.csv` in the `data` directory.
-One thing to note is that there is a missing observation (coded as "Not provided") for one of the entries (city of Napanee, located in the Lennox and Addington County).
+One thing to note is that there was a missing observation (coded as "Not provided") for one of the entries (city of Napanee, located in the Lennox and Addington County).
 Under the LHIN divisions, Napanee was in the South East LHIN.
-Also, there is one entry that says "244 Main Street East" as the community entry but it should be "Stayner" (the address provided belongs to Stayner.) The information is fixed before adding the Health Region.
+Also, there is one entry that says "244 Main Street East" as the community entry but it should be "Stayner" (the address provided belongs to Stayner).
+The information was fixed before adding the Health Region.
 
@@
 
-Now, combine datasets and write csv file for those cities that were not assigned a Health Region.
+Next, we combined the datasets and wrote csv file with those cities that were not assigned a Health Region.
+The next code chunk does these steps (note that the line for writing the csv file has ben commented as it was run only once).
 
@@
 
-To complete the missing LHIN information, the correspoding LHINs were obtained from the LHIN websites:
+To obtain the missing LHINswe, obtained information from the LHIN websites, which listed all the municipalities within each LHIN. The websites were:
 
-https://www.southeasthealthline.ca/aboutourregion.aspx?region=RuralFrontenacLennoxAddington
+- South East [link](https://www.southeasthealthline.ca/aboutourregion.aspx?region=RuralFrontenacLennoxAddington)
 
-https://www.nsmhealthline.ca/aboutourregion.aspx
+- North Simcoe Muskoka [link](https://www.nsmhealthline.ca/aboutourregion.aspx)
 
-https://www.champlainhealthline.ca/aboutourregion.aspx?region=StormontDundasGlen
+- Champlain [link](https://www.champlainhealthline.ca/aboutourregion.aspx?region=StormontDundasGlen)
 
-https://www.wwhealthline.ca/aboutourregion.aspx?region=KitchenerWaterlooWellesleyWilmotWoolwich
+- Waterloo Wellington [link](https://www.wwhealthline.ca/aboutourregion.aspx?region=KitchenerWaterlooWellesleyWilmotWoolwich)
+
+- North West [link](https://www.northwesthealthline.ca/aboutourregion.aspx)
+
+- North East [link](https://www.northeasthealthline.ca/aboutourregion.aspx)
+
+- Erie St. Clair [link](https://www.eriestclairhealthline.ca/aboutourregion.aspx)
+
+- South West [link](https://www.southwesthealthline.ca/aboutourregion.aspx)
+
+- Hamilton Niagara Haldimand Brant [link](https://www.hnhbhealthline.ca/aboutourregion.aspx)
+
+- Central West [link](https://www.centralwesthealthline.ca/aboutourregion.aspx)
+
+- Central East [link](https://www.centraleasthealthline.ca/aboutourregion.aspx)
+
+- Mississauga Halton [link](https://www.mississaugahaltonhealthline.ca/aboutourregion.aspx?region=HaltonHills)
+
+- Toronto Central [link](https://www.torontocentralhealthline.ca/aboutourregion.aspx)
 
@@
 For example, Etobicoke was divided between the Central, Central East, and Toronto Central LHINs.
-We chose in these cases the LHIN that covered the larger geographical region of each city (Toronto Central LHIN in the case of Etobicoke).
-We now assign LHINs to the dataset and Health Regions.
-The dataset with the information of the missing LHINs is called `missing_health_regions_updated.csv`
+We chose in these case the LHIN that covered the larger geographical region of each city (Toronto Central LHIN in the case of Etobicoke).
+We next assigned LHINs to the entries were they were missing using the information from the webistes, and created a csv with the updated information (the file is called `missing_health_regions_updated.csv`).
+The next code chunk loads this file, assigns LHINs and creates a new column in the dataset for Health Region.
 
@@
 
-At this point, the data has been cleaned and prepared for formal analysis.
+At this point, the data was ready for formal analysis.
 
-The clean dataset will be saved as a *.csv file.
+The clean dataset and completed dataset was saved as a *.csv file called `clean_dataset`.
 
@@
 
-write.csv(clean_data,here("data","clean_dataset.csv"),row.names = FALSE)
+#write.csv(clean_data,here("data","clean_dataset.csv"),row.names = FALSE)
 ```

### 2023-02-07 manuscript/02-methods.qmd
created file with details of data cleaning

@@
 |:------------------|-----------------------------------------------------|
-| Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
+| Age group            | 16-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
 | Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
@@
 In the case of municipalities that did not appear in the long-term dataset, we manually searched each city in the LHINs websites in order to provide geographical information.
-Details of the complete data cleaning process are in the Appendix.
+Details of the data cleaning process are in the github repository: https://github.com/aimundo/Fields_COVID-19/.
 

### 2023-02-07 manuscript/03-results.qmd
created file with details of data cleaning

@@
 The reference groups in each case were set as follows: 16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
-Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.44), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.4) and Other Race/Ethnicity (a group that included minorities not found elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000-39,999 (OR=0.23).
+Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.44), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.4) and Other Race/Ethnicity (a group those that identified as Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000-39,999 (OR=0.23).
 

### 2023-02-07 manuscript/appendix.qmd
created file with details of data cleaning

--- 2023-01-11 data_cleaning/Data_cleaning_and_covariates_Jan_09.qmd
@@
 ---
-title: "Fields COVID-19 dataset exploration"
-author: "Ariel Mundo"
+title: '**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**'
+subtitle: Appendix
+author: Ariel Mundo Ortiz
 format: pdf
-toc: true
-toc-depth: 2
-editor: source
-urlcolor: blue
 #bibliography: refs.bib
+#csl: jama.csl
+#link-citations: yes #adds links to the citations
 ---
@@
 
-This dataset is about vaccination and other COVID measures in different cities in Ontario. The covariates available are:
+The original dataset of the *Survey of COVID-19 related Behaviours and Attitudes* included different covariates (age, age-group, employment, remote work within the last month, etc).
+However, we are interested in the following covariates:
 
--   Age (only above 16, if age above 75 then it appears as 98)
--   Age-group (generated from age)
--   Employment status
--   Remote work within the last month
--   If person receives paid sick leave
--   Number of people in household
--   Number of people from household that attend school
--   Chronic illnesses within the household (*I will not analyze this one as the question is too open and asks about age and disease at the same time*)
+-   Age-group
 -   Race
--   Three first digits of postal code
--   Day, month and year the survey was accessed
-
-There are many responses from the survey on the dataset, but there are some that I would be interesed in analyzing:
-
+-   Location
+-   Income
 -   "Have you received the first dose of the COVID vaccine?" (y/n)
 
--   (If answered "yes" above) "Have you received the second dose of the COVID vaccine?" (y/n)
+# Data Cleaning
 
--   (If answer was "no" to the first question) "If a vaccine was made available to you you would:"
-
-          - definitely get
-          - definitely not get
-          - probably get
-          - probably not get
-
-**Note: As of Dec 30, 2022 I no longer intend to analyze this last question as after cleaning the dataset it has a missing rate of about 76% and therefore it has too few observations compared to the first two questions**.
-
-# Preliminary Analyses
-
-This document focuses on some preliminary analyses of the data I undertook to assess its representativeness. I performed five major tasks for this preliminary analysis:
-
--   Choosing covariates to assess their missigness rates
--   Cleaning the data in order to keep all the covariates with a 0% missigness rate
--   Race and ethnicity representativeness of the clean data
--   Trends in the survey answers by geographical location
--   Trends in survey answers by group age
-
-Overall the goal of this analysis was to see what adjustments would need to be made to formally analyze the data.
-
-As way of comparison, I used the 2016 census data for Ontario.
+The data was cleaned in order to keep the covariates, determine their differences to the 2016 Census Data, and identify and remove outliers.
+All analyses were carried in RStudio (2022.12.0 Build 353) and R version 4.2.2.
 
@@
 
-#the dataframe has missing answers as blankspaces, need to change them all to NA
+#the dataframe has missing answers as blank spaces, need to change the blank spaces all to NA
 
 data<-tibble(read.csv(here("data","Fields_data.csv"),na.strings = c("")))
-
-#set the time and date fields in the format that is required by lubridate
-
-data$time_date_code<-ymd(data$time_date_code)
-```
-
-```{r, rename-columns,echo=FALSE,warning=FALSE}
 
@@
 #get the variables of interest, convert to factors those required
+
 d1<-data%>%select(
-  time_date_code,
   city,
-  age,
   age_group,
   q02_first_dose,
-  q02a_second_dose,
-  q11_employed,
-  q13_hh_size,
   q09_hh_income,
-  q02b_vaccine,
-  q16_race,
-  q11a_work_remotely,
-  q11b_sick_leave,
-  q14_hh_attending_school,
-  q24a_postal_first,
-  q24b_postal_second,
-  q24c_postal_third)%>%
-  rename(date=time_date_code,
-         first_dose=q02_first_dose,
-         second_dose=q02a_second_dose,
-         get_vaccine=q02b_vaccine,
-         h_size=q13_hh_size,
+  q16_race)%>%
+  rename(first_dose=q02_first_dose,
          income=q09_hh_income,
-         employed=q11_employed,
-         race=q16_race,
-         remote_work=q11a_work_remotely,
-         sick_leave=q11b_sick_leave,
-         school=q14_hh_attending_school,
-         pc_1=q24a_postal_first,
-         pc_2=q24b_postal_second,
-         pc_3=q24c_postal_third
-         )%>%
+         race=q16_race)%>%
   mutate(across(c("age_group",
             "city",
-            "employed",
-            "h_size",
             "income",
-            "race",
-            "remote_work",
-            "sick_leave",
-            "school"),as.factor))
+            "race"),
+            as.factor))
 
@@
 
-## Choosing covariates
+## Covariate Missigness
 
-The next step is to see from the covariates what are the ones with the highest number of missing observations, in order to decide what can we include in the model.
+Next, we determined the number of missing observations in the covariates.
 
@@
   select(age_group,
-         employed,
-         h_size,
-         income,race,
-         remote_work,
-         sick_leave,
-         school,
-         pc_1,
-         pc_2,
-         pc_3)%>%
- summarise_all(funs(sum(!is.na(.)/length(.))))%>%#count percentage of NAs
+         income,
+         race,
+         first_dose)%>%
+ summarise_all(funs(sum(!is.na(.)/length(.))))%>% #count percentage of NAs
   pivot_longer(everything())%>%
@@
   select(age_group,
-         employed,
-         h_size,
-         income,race,
-         remote_work,
-         sick_leave,
-         school,
-         pc_1,
-         pc_2,
-         pc_3)%>%
+         income,
+         first_dose,
+         race)%>%
  summarise_all(funs(sum(is.na(.)/length(.))))%>%#count percentage of NAs
@@
   select(age_group,
-         employed,
-         age,
          income,
@@
          city,
-         h_size,
-         date,
-         first_dose,
-         second_dose,
-         get_vaccine)%>%
-  drop_na(employed)
-
-### histogram plots to see issues about income, age and household #####
-
-# clean_data%>%
-#   ggplot(aes(x=age))+
-#   geom_histogram(binwidth=10)+
-#   facet_wrap(~income)
-
-a<-d1%>%
-  mutate(income_ord = fct_relevel(income,
-                                   "under_15000",
-                                   "15000_24999",
-                                   "25000_39999",
-                                   "40000_59999",
-                                   "60000_89999",
-                                   "90000_109999",
-                                  "over_110000"
-                                   ))%>%
- #drop_na(age,income,h_size)%>%
-  ggplot(aes(x=age))+
-  geom_histogram(binwidth=10)+
-  facet_wrap(~income_ord)+
-  ylim(c(0,1200))+
-  theme(strip.text.x = element_text(size = 8))+
-  theme_classic()+
-  thm1
-
-b<-d1%>%
-  mutate(h_size_ord = fct_relevel(h_size,
-                                   "1",
-                                   "2",
-                                   "3",
-                                   "4",
-                                   "5",
-                                   "6",
-                                   "7",
-                                   "8",
-                                   "9",
-                                   "10_or_more"
-                                   ))%>%
-  #drop_na(age,income,h_size_ord)%>%
-  ggplot(aes(x=age))+
-  geom_histogram(binwidth=10)+
-  facet_wrap(~h_size_ord)+
-  ylim(c(0,200))+
-  theme(strip.text.x = element_text(size = 8))+
-  theme_classic()+
-  thm1
-
-c<-clean_data%>%
-   mutate(income_ord = fct_relevel(income,
-                                   "under_15000",
-                                   "15000_24999",
-                                   "25000_39999",
-                                   "40000_59999",
-                                   "60000_89999",
-                                   "90000_109999",
-                                  "over_110000"
-                                   ))%>%
-  ggplot(aes(x=age))+
-  geom_histogram(binwidth=10)+
-  facet_wrap(~income_ord)+
-  theme(strip.text.x = element_text(size = 5))+
-  theme_classic()+
-  thm1
-
-d<-clean_data%>%
-   mutate(h_size_ord = fct_relevel(h_size,
-                                   "1",
-                                   "2",
-                                   "3",
-                                   "4",
-                                   "5",
-                                   "6",
-                                   "7",
-                                   "8",
-                                   "9",
-                                   "10_or_more"
-                                   ))%>%
-  #drop_na(age,income,h_size_ord)%>%
-  ggplot(aes(x=age))+
-  geom_histogram(binwidth=10)+
-  facet_wrap(~h_size_ord)+
-  ylim(c(0,200))+
-  theme_classic()+
-  theme(strip.text.x = element_text(size = 8))+
-  thm1
-
-a+b+plot_annotation(title="Original dataset")
-
-c+d+plot_annotation(title="Clean dataset")
-
-e<-clean_data%>%
-  #drop_na(age,income,h_size_ord)%>%
-  ggplot(aes(x=age,fill=employed))+
-  geom_histogram(stat="count",binwidth=10)+
-  #ylim(c(0,200))+
-facet_wrap(~income)+
-  theme_classic()+
-  ggtitle("Clean dataset")+
-  #theme(strip.text.x = element_text(size = 8))+
-  thm1
-
-f<-d1%>%
-  #drop_na(age,income,h_size_ord)%>%
-  ggplot(aes(x=age,fill=employed))+
-  geom_histogram(stat="count",binwidth=10)+
-  #ylim(c(0,200))+
-facet_wrap(~income)+
-  theme_classic()+
-  ylim(c(0,100))+
-  ggtitle("Original dataset")+
-  #theme(strip.text.x = element_text(size = 8))+
-  thm1
-
-f+e
-
-t4<-clean_data%>%
- summarise_all(funs(sum(is.na(.)/length(.))))%>% #count percentage of NAs
-  pivot_longer(everything())%>%
-  arrange(desc(value))%>%
-  rename(missigness=value,
-         variable=name)%>%
-  gt()%>%
-  fmt_percent(
-   columns=missigness,
-    decimals=1)
-t4
+         first_dose,)%>%
+  drop_na(race)
 

### 2023-02-07 manuscript/main.qmd
created file with details of data cleaning

@@
 ---
-title: '**Socio-economic and geographic determinants of Covid-19 vaccination**'
+title: '**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**'
 author: Ariel Mundo Ortiz
@@
 
-<!-- \FloatBarrier -->
-
 ```{r} 

### 2023-02-07 protocol/Background_and_methods_Jan_24.qmd
created file with details of data cleaning

@@
 | Variable             | Values                                                                                                                          |
-|:-------------------|----------------------------------------------------|
+|:---------------------|---------------------------------------------------------------------------------------------------------------------------------|
 | Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |

### 2023-02-08 data_cleaning/Data_cleaning_final.qmd
Fixed appendix, cleaning folder, latex preamble

@@
 
+# Purpose
+
+This document focuses on the cleaning of the data from the Survey of COVID-19 related Behaviours and Attitudes which was used as the data source for the analysis in the main paper.
+The cleaning process encompassed exploratory analyses, covariate selection, removal of outliers, and adding geographical information to the data for analysis (Health Regions in Ontario).
+At the end of the data cleaning process, a dataset ready for for formal analyses was produced.
+This document can be rendered to re-create all the steps used in the cleaning process and to generate the same dataset that was used for analysis.
+
+Note, however, that code chunks where intermediate csv files are generated (such as those for missing municipalities, or missing Health Regions) are commented in the current document as they were executed once to create the files.
+Also, to reduce the lenght of the document, the code in the code chunks does not appear in the rendered file.
+
 # Background
@@
           - probably not get
-
-# Preliminary Analyses
-
-This document focuses on exploratory analyses, covariate selection, and outlier identification and removal. The goal of the exploratory analyses was to clean the dataset for formal analyses.
-
-As way of comparison, I used the 2016 census data for Ontario.
 
@@
 
-The next step consisted in identify the missing rates of the covariates, and determinewhich of those could be included in the model.
+The next step consisted in identify the missing rates of the covariates, and determine which of those could be included in the model.
 The following code chunk creates a table with the number of missing observations and percentages.
@@
 
-At this stage in the analysis, we identified that outliers were present in the data.
-Specifically, individuals that were under 25 years of age and that reported having an income \>110k while living in a household of 1.
-We therefore, created some diagnostic plots to explore the reported household composition of those \<25 with income\>110k, and we removed the outliers from the dataset.
+At this stage in the analysis, we examined the data for outliers.
+We identified certain entries in the survey that corresponded to individuals that were under 25 years of age and that reported having an income \>110k while living in a household 1.
+The next code chunk creates a plot of the number of observations that have these characteristics, which were around 20.
+Next, these outliers were removed from the dataset.
 
-```{r,outliers}
+```{r,outliers, echo=FALSE, warning=FALSE,message=FALSE}
 outliers<-clean_data %>%
@@
 
-## Race and Ethnicity
-
-Next, we explored how the race and ethnicity information from the dataset compared to the data from the Census. The following chunk creates a summary table for the Race variable.
-
-```{r,ethnic information}
-#| echo: false
-#| warning: false
-#| tbl-cap: Ethnic information from the clean dataset
-  t6<-clean_data%>%
-    select(race)%>%
-    group_by(race)%>%
-    summarise(observations = n()) %>%
-  mutate(percentage = observations / sum(observations))%>%
-    gt()%>%
-      fmt_percent(
-   columns=percentage,
-    decimals=1)
-
-t6
-```
-
-It is important to mention that the categories for race/ethnicity provided in the survey did not match the categories used in the Census.
-Therefore, we used a combination of sources to obtain the information presented in @tbl-races, where the categories used could be matched to the Census Information.
-The data sources for the table were:
-
-- Fact sheet from the Provice of Ontario for Visible Minorities [link](https://www.ontario.ca/document/2016-census-highlights/fact-sheet-9-ethnic-origin-and-visible-minorities): Used to obtain percentages for Arab, Black, East Asian/Pacific Islander (adding Chinese, Korean, and Japanese percentages), Latin American, Mixed (using the percentage for "multiple visible minority"), Other (obtained by adding the Southeast Asian, Filipino, West Asian, and Minority not identified elsewhere percentages), South Asian.
-_Accessed on January 05, 2022_
-
-- Census Profile for Ontario [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Aboriginal%20peoples&TABID=1&type=1): Used to obtain percentage of Aboriginal population.
-_Accessed on January 05, 2022_
-
-- Wikipedia entry for Ontario demographics [link](https://en.wikipedia.org/wiki/Demographics_of_Ontario): To corroborate that the percentage of population reported as "European" in this website matched the percentage obtained for "White Caucasian" that was independently obtained by obtaining the difference in population proportion after subtracting the sum of Visible Minorities and Aboriginal Population percentages.
-
-| Ethnicity/Race | percentage |
-|----------------|:-----------|
-| Arab           | 1.6%       |
-| Black          | 4.7%       |
-| East Asian/Pacific Islander     | 6.6%       |
-| Indigenous     | 2.8%       |
-| Latin American | 1.5%       |
-| Mixed          | 1.0%       |
-| Other          | 5.2%       |
-| South Asian    | 8.7%       |
-| White Caucasian| 67.8%      |
-| **Total**      | 99.9%      |
-: Reference Data from Race/Ethnicity {#tbl-races}
-
-Visible Minorities: 29.3%
-Not a Visible Minority: 70.7% (Aboriginal 2.8%, White Caucasian 67.8%)
-
-## Age groups
-
-According to the Census, the distribution of the different the age groups for the province of Ontario are as follows:
-
-| Group age   | Percentage of population |
-|-------------|:-------------------------|
-| 15-24       | 12.7%                    |
-| 25-34       | 12.9%                    |
-| 35-44       | 12.8%                    |
-| 45-54       | 14.9%                    |
-| 55-64       | 13.7%                    |
-| 65 and over | 16.7                     |
-
-: Age distributions from the 2016 Census for Ontario
-
-To compare the survey data to the Census, the next chunk creates a barplot of the age groups in the survey.
-Here, it can be noticed that the age-group distribution from the survey data is different from the Census data.
-
-```{r,trends-by-age}
-#| echo: false
-#| warning: false
-#| fig-height: 4
-#| fig-width: 8
-#| fig-cap: Age group distributions from the dataset
-
-p2<-clean_data%>%
-  select(age_group)%>%
-   mutate(age_group_ord = fct_relevel(age_group,
-                                      "16_24",
-                                      "25_34",
-                                      "35_44",
-                                      "45_54",
-                                      "55_64",
-                                  "65_and_over"
-                                          ))
-
-  ggplot(p2%>%count(age_group_ord) %>%    # G count number in each group
-         mutate(pct=n/sum(n),               # Calculate percent within each group
-                ypos = 200),  # Calculate label positions
-       aes(age_group_ord, n, fill=age_group_ord)) +
-  geom_bar(stat="identity") +
-  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
-  theme_classic()+
-  thm1
-
-```
-
-## Income
-
-Survey respondents answered the question "What is your household annual income?".
-To compare the distribution of responses to this question from the survey and the Census data, we used the "Household total income groups in 2015 for private households" from the Census data (available in the Census Data for Ontario website [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Income&TABID=1&type=1)).
-
-| Household income range (CAD) | Percentage |
-|------------------------------|:-----------|
-| \< 15,000                    | 5.7%       |
-| 15,000 - 24,999              | 7.5%       |
-| 25,000 - 39,999              | 11.6%      |
-| 40,000- 59,999               | 15.4%      |
-| 60,000 - 89,999              | 19.5%      |
-| \>90,000                     | 40.3%      |
-
-: Income percentages from the 2016 Census for Ontario
-
-One difference that is noticeable is that the brackets for income in the census data are different than the brackets used in the survey.
-The census does CAD 4,999 brackets (e.g., CAD 5,000- CAD 9,9999) up to CAD 49,999, followed by CAD 9,999 brackets up to CAD 99,999.
-After that, the brackets increase to CAD 24,999, and therefore, it is not possible to obtain percentages for the 90,000-109,999 and \>110,000 brackets from the survey.
-
-Therefore, in the following code chunk, we created an additional category for income in the dataset that matched the information from the Census, and barplots to visualize the proportion of each income bracket in the dataset.
-
-```{r,income}
-#| echo: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 4
-
-clean_data<-clean_data%>%
-    mutate(income_ord = case_when(
-                                  income %in% "90000_109999" ~ "over_90000",
-                                  income %in%"over_110000" ~ "over_90000",
-                                  TRUE ~ as.character(income)
-                                ))%>%
-  mutate(income_ord=as.factor(income_ord))
-
-p3<-clean_data%>%
-  mutate(income_ord = fct_relevel(income_ord,
-                                  "under_15000",
-                                  "15000_24999",
-                                  "25000_39999",
-                                  "40000_59999",
-                                  "60000_89999",
-                                  "over_90000"
-                                ))
-
-ggplot(p3%>% count(income_ord)%>%  # count number in each group
-         mutate(pct=n/sum(n),               # Calculate percent within each group
-                ypos = 200),  # Calculate label positions
-       aes(income_ord, n, fill=income_ord)) +
-  geom_bar(stat="identity") +
-  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
-  theme_classic()+
-  thm2
-
-```
-
-We identified differences again between the survey data and the Census data.
-For example, the \<15,000 bracket accounts for about 25% of the responses, a much higher rate than the proportion in the Census.
-The other brackets were different as well from the Census data.
-
 ## Geographical Information
@@
 
-```{r,join-datasets}
+```{r,join-datasets, echo=FALSE}
 
@@
 
-```{r, missing-geog-area}
+```{r, missing-geog-area, echo= FALSE}
 
@@
 
-```{r,merge-datasets-2}
+```{r,merge-datasets-2, echo=FALSE}
 #load missing municipalities dataset
@@
 
-```{r,get-health-data}
+```{r,get-health-data, echo=FALSE}
 
@@
 
-```{r,combine}
+```{r,combine, echo=FALSE}
 
@@
 
-```{r,merge-health-regions}
+```{r,merge-health-regions, echo=FALSE}
 

### 2023-02-08 manuscript/02-methods.qmd
Fixed appendix, cleaning folder, latex preamble

@@
 Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000.
-After cleaning, the dataset contained 3,709 unique entries.
+The next step consisted in processing the geographical information in the survey (city where the survey was responded) in order to match each city to its correspondent Health Region.
+Details of this process are given below.
 
@@
 In the case of municipalities that did not appear in the long-term dataset, we manually searched each city in the LHINs websites in order to provide geographical information.
-Details of the data cleaning process are in the github repository: https://github.com/aimundo/Fields_COVID-19/.
+The raw dataset, clean dataset, details of the data cleaning process, and the addition of Health Region and LHIN information are described in detail in the github repository: https://github.com/aimundo/Fields_COVID-19/.
 

### 2023-02-08 manuscript/03-results.qmd
Fixed appendix, cleaning folder, latex preamble

@@
 #| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"
-source(here::here("code","raking_logistic_regression.R"))
 
+#logistic regression
+source(here::here("code","regression.R"))
+
+# create table with the model output
 m1 %>%

### 2023-02-08 manuscript/appendix.qmd
Fixed appendix, cleaning folder, latex preamble

@@
 author: Ariel Mundo Ortiz
-format: pdf
+format: 
+  pdf:
+    include-before-body: latex/appendix_preamble.sty
 #bibliography: refs.bib
@@
 thm3<-scale_fill_scico_d(palette="fes",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")
+
+clean_data <- read.csv(here("data","clean_dataset.csv"))
+
+# remove rows with missing observations for race to obtain accurate numbers
+
+clean_data<-clean_data %>% drop_na(race)
 ```
 
-# Background
+This document provides details on assessing the socio-economic information from the clean survey dataset, and how it compared to the Census Data for ontario.
+At the end, it also provides information about the corrections implemented into the data.
+The variables examined were: race/ethnicity, age groups, income, and population by Health Region.
+The computations are performed in the `clean_dataset.csv` file that was created at the end of the data cleaning process.
+Population-wide totals were obtained from the 2016, available in the [Statistics Canada Website](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/Page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=&Code2=&Data=Count&SearchText=Ontario&Sear).
 
-The original dataset of the *Survey of COVID-19 related Behaviours and Attitudes* included different covariates (age, age-group, employment, remote work within the last month, etc).
-However, we are interested in the following covariates:
+# Race and Ethnicity
 
--   Age-group
--   Race
--   Location
--   Income
--   "Have you received the first dose of the COVID vaccine?" (y/n)
-
-# Data Cleaning
-
-The data was cleaned in order to keep the covariates, determine their differences to the 2016 Census Data, and identify and remove outliers.
-All analyses were carried in RStudio (2022.12.0 Build 353) and R version 4.2.2.
-
-```{r,load-data,warning=FALSE,echo=FALSE}
-
-#the dataframe has missing answers as blank spaces, need to change the blank spaces all to NA
-
-data<-tibble(read.csv(here("data","Fields_data.csv"),na.strings = c("")))
-
-#get a new dataframe and extract variables of interest and make them factors
-
-#get the variables of interest, convert to factors those required
-
-d1<-data%>%select(
-  city,
-  age_group,
-  q02_first_dose,
-  q09_hh_income,
-  q16_race)%>%
-  rename(first_dose=q02_first_dose,
-         income=q09_hh_income,
-         race=q16_race)%>%
-  mutate(across(c("age_group",
-            "city",
-            "income",
-            "race"),
-            as.factor))
-
-```
-
-## Covariate Missigness
-
-Next, we determined the number of missing observations in the covariates.
-
-```{r}
-#| echo: false
-#| warning: false
-#| tbl-cap: Percentage of missing observations \n all covariates
-
-t1<-d1%>%
-  select(age_group,
-         income,
-         race,
-         first_dose)%>%
- summarise_all(funs(sum(!is.na(.)/length(.))))%>% #count percentage of NAs
-  pivot_longer(everything())%>%
-  arrange(desc(value))%>%
-  rename(observations=value,
-         variable=name)
-
-#get actual observations
-t2<-d1%>%
-  select(age_group,
-         income,
-         first_dose,
-         race)%>%
- summarise_all(funs(sum(is.na(.)/length(.))))%>%#count percentage of NAs
-  pivot_longer(everything())%>%
-  arrange(desc(value))%>%
-  rename(missigness=value,
-         variable=name)
-
-#joining both to make table
-
-t3<-left_join(t1,t2,by="variable")
-
-#make table
-
-t3%>%
-  gt()%>%
-  fmt_percent(
-   columns=missigness,
-    decimals=1)
-
-```
-
-From the table, it can be seen that the covariate with the least amount of observations is "sick_leave".
-
-However, there are a couple of things we want to consider: The answer to "sick_leave" was answered only by those that reported to be employed (the survey design made this response conditional).
-Therefore, those unemployed would be excluded in an analysis that considers this variable.
-
-At the very least, it would be interesting to analyze how the reported status of vaccination changed over geographical location and time by:
-
--   age group
--   income
--   race
--   employment status
-
-From these covariates, the one with the highest missing rate of observations is the employment status.
-I next cleaned the data in order to have complete observations about employment, and to see how the missing rates looked for the other covariates.
-
-```{r}
-#| echo: false
-#| warning: false
-#| tbl-cap: Missing observations and histograms of the data
-#| fig-width: 10
-
-clean_data<-d1%>%
-  select(age_group,
-         income,
-         race,
-         city,
-         first_dose,)%>%
-  drop_na(race)
-
-```
-
-There seem to be outliers where people \<25 y of age report having an income \>110k while living in a household of 1. Explore the reported household composition of those \<25 with income\>110k.
-
-```{r,outliers}
-outliers<-clean_data %>%
-  filter(age_group=="16_24" &
-           h_size==1 &
-           income=="over_110000"
-
-  )
-
-outliers%>%
-  ggplot(aes(x=h_size))+
-  geom_histogram(stat="count")
-
-#From the plot it is about 21 entries that are outliers. Now, remove these outliers
-
-clean_data<-anti_join(clean_data,outliers)
-
-```
-
-This table shows something interesting due to the nature of the dataset: the missing observations across variables are different, which is why now, although in the original dataset there were more observations of "race" than "employment", when we select complete observations for the latter we lose some observations for race.
-The second step of cleaning would be to remove the missing observations from race as well, and see what the proportions of covariates and responses are when compared to the census data.
-
-```{r,complete-obs}
-#| echo: false
-#| warning: false
-#| tbl-cap: Clean dataset missigness
-
-t7<-clean_data%>%
-  drop_na(race)%>%
- summarise_all(funs(sum(is.na(.)/length(.))))%>% #count percentage of NAs
-  pivot_longer(everything())%>%
-  arrange(desc(value))%>%
-  rename(missigness=value,
-         variable=name)%>%
-  gt()%>%
-  fmt_percent(
-   columns=missigness,
-    decimals=1)
-t7
-```
-
-It can be seen that after removing all missing observations in the covariates, the only response that still has missing observations is the answers for the second dose of the vaccine, with 27% of missing observations.
-
-## Race and Ethnicity
-
-Once the data was clean, I explored how the race and ethnicity information from the dataset compared to the data from the Census.
+First, we explored how the race and ethnicity information from the dataset and how it compared to the data from the Census for Ontario.
+The following chunk creates a summary table for the Race variable.
 
@@
 
-Because the categories for race/ethnicity in the survey did not match the categories from the Canadian Census, we used a combination of sources to obtain the corresponding percentages to match the survey categories.
-Sources are indicated below:
+It is important to mention that the categories for race/ethnicity provided in the survey did not match the categories used in the Census.
+Therefore, we used a combination of sources to obtain estimates of racial/ethnic distribution that matched the categories from the survey.The data sources were:
 
@@
 - Wikipedia entry for Ontario demographics [link](https://en.wikipedia.org/wiki/Demographics_of_Ontario): To corroborate that the percentage of population reported as "European" in this website matched the percentage obtained for "White Caucasian" that was independently obtained by obtaining the difference in population proportion after subtracting the sum of Visible Minorities and Aboriginal Population percentages.
+The aggregated information can be found in in @tbl-races, where the totals per race/ethnic category are presented.
 
-| Ethnicity/Race | percentage |
-|----------------|:-----------|
-| Arab           | 1.6%       |
-| Black          | 4.7%       |
-| East Asian/Pacific Islander     | 6.6%       |
-| Indigenous     | 2.8%       |
-| Latin American | 1.5%       |
-| Mixed          | 1.0%       |
-| Other          | 5.2%       |
-| South Asian    | 8.7%       |
-| White Caucasian| 67.8%      |
-| **Total**      | 99.9%      |
-: Reference Data from Race/Ethnicity
+| Ethnicity/Race | Percentage | Population totals |
+|----------------|:-----------| :-----------------|
+| Arab           | 1.6%       | 212782    |
+| Black          | 4.7%       | 638346|
+| East Asian/Pacific Islander     | 6.6%       | 886592 |
+| Indigenous     | 2.8%       | 376558 |
+| Latin American | 1.5%       | 197020 |
+| Mixed          | 1.0%       | 130033 |
+| Other          | 5.2%       | 705333 |
+| South Asian    | 8.7%       | 1166361 |
+| White Caucasian| 67.8%      | 9118079 |
+| **Total**      | 99.9%      | 13431105 |
+: Reference Data from the 2016 Census from Race/Ethnicity in Ontario {#tbl-races}
 
@@
 
-## Age groups
+It can be seen that there are differences in the distribution of each ethnicity between the Census data and the survey.
+For example, Arabs/Middle Easteners correspond to 5.9% of the survey responses, but population-wise individuals that identify in this ethnic group correspond to 1.6% of the Ontario population.
 
-Another important aspect to consider is how the trends in the responses look over time by age group. This is because from the census data, the age groups for the province are as follows:
+# Age Groups
 
-| Group age   | Percentage of population |
-|-------------|:-------------------------|
-| 15-24       | 12.7%                    |
-| 25-34       | 12.9%                    |
-| 35-44       | 12.8%                    |
-| 45-54       | 14.9%                    |
-| 55-64       | 13.7%                    |
-| 65 and over | 16.7                     |
+According to the Census, the distribution of the different the age groups for the province of Ontario are as follows:
 
+| Group age   | Percentage               | Population Totals |
+|-------------|:-------------------------|:------------------|
+| 15-24       | 12.7%                    | 1707959           |
+| 25-34       | 12.9%                    | 1734856           |
+| 35-44       | 12.8%                    | 1721407           |
+| 45-54       | 14.9%                    | 2003826           |
+| 55-64       | 13.7%                    | 1842444           |
+| 65 and over | 16.7%                    |2245898            |
 : Age distributions from the 2016 Census for Ontario
 
-However, when plotted the proportion of answers at each time point among groups, the distribution looks rather different.
+To compare the survey data to the Census, the next chunk creates a barplot of the age groups in the survey.
+Here, it can be noticed that the age-group distribution from the survey data is different from the Census data.
 
@@
 
-It can be seen from the graph that the proportion of answers by group age do not follow the overall distribution from the province.
+# Income
 
-## Income
+Survey respondents answered the question "What is your household annual income?".
+To compare the distribution of responses to this question from the survey and the Census data, we used the "Household total income groups in 2015 for private households" from the Census data (available in the Census Data for Ontario website [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Income&TABID=1&type=1)).
 
-The question from the survey was "What is your household annual income?".
-To compare the results of this answer, we used the "Household total income groups in 2015 for private households" from the Census data (available in the Census Data for Ontario website [link](https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E&Geo1=PR&Code1=35&Geo2=PR&Code2=01&SearchText=Ontario&SearchType=Begins&SearchPR=01&B1=Income&TABID=1&type=1)).
-
-| Household income range (CAD) | Percentage |
-|------------------------------|:-----------|
-| \< 15,000                    | 5.7%       |
-| 15,000 - 24,999              | 7.5%       |
-| 25,000 - 39,999              | 11.6%      |
-| 40,000- 59,999               | 15.4%      |
-| 60,000 - 89,999              | 19.5%      |
-| \>90,000                     | 40.3%      |
+| Household income range (CAD) | Percentage | Population Totals |
+|------------------------------|:-----------|:------------------|
+| \< 15,000                    | 5.7%       | 294643            |
+| 15,000 - 24,999              | 7.5%       | 387688            |
+| 25,000 - 39,999              | 11.6%      | 599624            |
+| 40,000- 59,999               | 15.4%      | 796052            |
+| 60,000 - 89,999              | 19.5%      | 1007988           |
+| \>90,000                     | 40.3%      | 2083176           |
 
@@
 #| warning: false
+#| fig-height: 4
 #| fig-width: 8
-#| fig-height: 4
+#| fig-cap: Income bracket distribution in the dataset
 
@@
   geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
+  xlab("Income")+
+  theme(legend.title = element_blank())+
   theme_classic()+
@@
 
-Again, there are differences here in the patterns of response.
-The \<15,000 bracket accounts for about 20% of the responses, a much higher rate than the percentage they represent from the census, and there is variation within the other brackets as well.
+We identified differences again between the survey data and the Census data.
+For example, the \<15,000 bracket accounts for about 24% of the responses, a much higher rate than the proportion from the Census (5.7\%).
+The other brackets were different as well from the Census data.
 
-## Geographical location
+# Health Regions Population
 
-Next, we are interested in analyzing the geographical distribution of the answers the participants within the geographical areas defined by the Association of Municipalities from Ontario.
+Finally, we obtained population totals for each of the Health Regions included in the dataset from the Ontario Health Business Plan for 2022-2023 (available [here](https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf)).
+The population totals are presented in @tbl-health-regions.
 
-**Update: the populations by census division/city were found on Jan 8, 2023 (from https://www12.statcan.gc.ca/census-recensement/2021/dp-pd/prof/index.cfm?Lang=E).
-The correction will be done using these numbers instead.
-The numbers can be found in `census_population_by_division.csv` in `data` directory.**
+| Health Region | Population Totals |
+|---------------|:------------------|
+| North East    | 232299            |
+| North West    |557000             |
+| West          |4095589            |
+| East          |3742520            |
+| Central       |5032410            | 
+| Toronto       |1440644            |
 
-There are two parts to the geographical analysis:
+: Population Totals for the Health Regions in Ontario {#tbl-health-regions}
 
-1) Add to the dataset the data for the municipalities so each city has the corresponding region it belongs to.
-1) Determine percentage of answers that correspond to Toronto.
-This is important because although the Census regions do not match the exact geographical boundaries of the municipalities and thus a direct correction for this cannot be easily obtained, Toronto is the largest urban area in the province and it is important to identify if corrections are needed with regard to the proportions from this place in the survey.
+Note that as indicated in the paper, we excluded observations from the North East and North West Health Regions, and therefore, the population totals for these areas were not used in the final analysis.
 
-### Joining the municipalities data
+# Corrections
 
-We have obtained and cleaned the data of the municipalities from the province of Ontario [website](https://www.ontario.ca/page/list-ontario-municipalities#section-3), and cleaned the dataset to obtain the city names and geographical locations.
-Further details can be found in `municipalities.qmd` in the `data_cleaning` directory.
-We will join and match the geographical location from the municipalities dataset to the clean dataset we have obtained in this document so far after removing the entries that do not have a corresponding geographical location.
-
-```{r,join-datasets}
-
-clean_data<-clean_data%>%
-  filter(city!="None")
-
-municipalities<-read.csv(here("data","municipalities_clean.csv"))
-
-municipalities$Municipal.status<-as.factor(municipalities$Municipal.status)
-
-municipalities$Geographic.area<-as.factor(municipalities$Geographic.area)
-
-municipalities$city<-as.factor(municipalities$city)
-
-clean_data<-left_join(clean_data,municipalities,by="city")
-
-```
-
-Need to explore which entries were left without a geographical region:
-
-```{r, missing-geog-area}
-
-test<-clean_data %>%
-  filter(is.na(Geographic.area))%>%
-  distinct(city)
-
-```
-
-There are 2744 observations that did not get a geographical region, and they correspond to 187 unique cities.
-These will be exported to a csv file in order to manually write their geographical areas following the Association of Municipalities of Ontario divisions, and using Wikipedia to check the status of each municipality.
-
-```{r,write-file}
-#| echo: false
-
-#write.csv(test,here("data","missing_municipalities.csv"),row.names = FALSE)
-
-```
-
-The code chunk above was ran once to get the names of the municipalities.
-After manually entering the geographic areas, the file was saved as `missing_municipalities_updated.csv`, and this will be the file to be used for the next steps.
-
-Note: there is one city "Kinburn", but there are two communities with such name, one in Huron County and one in Carelton County, for the time being Huron County is assigned, will check later if is better to remove it.
-
-Also one occurence of "Sydenham", which can be a ward in Kingston or a community in Frotenac. Will assign Frotenac for the time being.
-
-### Merging missing municipalities and geographic area names
-
-After manually assigning the geographical regions to the municipalities that were missing (which can be found in `missing_municipalities.csv` in the `data` directory), these will be merged as the dataset `geographic_areas.csv` (also in `data`), which contains the titles for each region (e.g., "County", "Region").
-
-```{r,merge-datasets-2}
-#load missing municipalities dataset
-
-missing_municipalities<-read.csv(here("data","missing_municipalities_updated.csv"))
-
-#combining geographical regions
-
-clean_data<-clean_data %>%
-  left_join(missing_municipalities, by = c("city")) %>%
-  mutate(Geographic.area = coalesce(Geographic.area.x,Geographic.area.y)) %>%
-  select(-c(Geographic.area.x,Geographic.area.y))
-
-clean_data$Geographic.area<-as.factor(clean_data$Geographic.area)
-
-# load the the titles for each region
-
-geographic_areas<-read.csv(here("data","geographic_areas.csv"))
-
-geographic_areas$Geographic_area<-as.factor(geographic_areas$Geographic_area)
-
-geographic_areas$Geographic_area_title <-as.factor(geographic_areas$Geographic_area_title)
-
-#merge the datasets
-
-clean_data <-left_join(clean_data,geographic_areas,by=c("Geographic.area"="Geographic_area")) %>%
-   subset(select=-c(Full_title.y))
-
-```
-
-## Proportion of answers within Toronto
-
-**This section needs to be modified as the correction has been done by geographical area on Jan 8, 2023**
-
-The final step is of the geographical analysis is to determine the proportion of answers from Toronto.
-
-```{r,plot-location}
-#| echo: false
-#| warning: false
-#| fig-width: 8
-#| fig-height: 4
-#| fig-cap: Percentage of survey answers by city
-
-clean_data<-clean_data%>%
-  mutate(location=as.character(case_when(city=="Toronto"~"Toronto",
-                            city!="Toronto"~ "All other cities"
-                          )))%>%
-  mutate(location=as.factor(location))
-
-ggplot(clean_data %>% count(location) %>%    # then count number in each group
-         mutate(pct=n/sum(n),               # Calculate percent within each group
-                ypos = 200),  # Calculate label positions
-       aes(location, n, fill=location)) +
-  geom_bar(stat="identity") +
-  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
-  theme_classic()+
-  thm1
-
-```
-
-The proportion of responses from the city of Toronto is 31.3% of the total in the clean dataset.
-From the (Census data)[https://www.toronto.ca/wp-content/uploads/2022/02/92e3-City-Planning-2021-Census-Backgrounder-Population-Dwellings-Backgrounder.pdf], the city of Toronto as of 2021 has a population of 2,794,356, corresponding to 19.6% of the total population of the Province.
-
-# Health Regions approach
-
-Following on a suggestion by BN on Jan 09, the idea is to use broader regions for geographical analysis.
-The Health Regions of Ontario will be used.
-However, these Health Regions do not match the census divisions.
-I have not been able to find a dataset from Health Ontario that lists each municipality and its corresponding Health Region.
-I will use a multi-stage approach to incorporate the information into the dataset:
-
-- Get dataset from Paul Allen regarding long-term care homes in Ontario (https://paulallen.ca/consolidated-dataset-of-ltc-homes-in-ontario/).
-This dataset has communities and the Local Health Integration Network (LHINs) where long-term care homes were located.
-I will keep the community and LHIN information.
-
-- Add column for Health Region based on LHIN: The Health Regions substituted LHINs, a list of their correspondence can be found here: https://www.ontariohealth.ca/about-us/our-people.
-I will add this information.
-
-- Merge the dataset and check if there are missing municipalities. Manually add missing municipalities.
-
-## Steps
-
-The dataset from Paul Allen was downloaded and saved as `Consolidated_LTC_dataset.csv` in the `data` directory.
-One thing to note is that there is a missing observation (coded as "Not provided") for one of the entries (city of Napanee, located in the Lennox and Addington County).
-Under the LHIN divisions, Napanee was in the South East LHIN.
-Also, there is one entry that says "244 Main Street East" as the community entry but it should be "Stayner" (the address provided belongs to Stayner.) The information is fixed before adding the Health Region.
-
-```{r,get-health-data}
-
-#load data
-ltc_data <- read.csv(here("data","Consolidated_LTC_dataset.csv"))
-
-## checking missing observations
-
-sum(ltc_data$X_LHIN=="Not provided")
-
-ltc_data <- ltc_data %>%
-  mutate(X_LHIN=replace(X_LHIN,X_LHIN=="Not provided","South East"))
-
-ltc_data <- ltc_data %>%
-  mutate(COMMUNITY=replace(COMMUNITY,COMMUNITY==" 244 Main Street East","Stayner"))
-
-#keep relevant columns and add Health region information
-ltc_data <-ltc_data %>%
- select(c("COMMUNITY","X_LHIN"))%>%
-  rename(city=COMMUNITY,
-         LHIN=X_LHIN)%>%
-  mutate(Health_Region=
-    case_when(LHIN=="Central"~ "Central",
-              LHIN=="Central West"~"Central",
-              LHIN=="Mississauga Halton"~"Central",
-              LHIN=="North Simcoe Muskoka"~"Central",
-              LHIN=="Central East"~"East",
-              LHIN=="South East"~"East",
-              LHIN=="Champlain"~"East",
-              LHIN=="North East"~"North East",
-              LHIN=="North West"~"North West",
-              LHIN=="Toronto Central"~"Toronto",
-              LHIN=="South West"~"West",
-              LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)"~"West",
-              LHIN=="Waterloo Wellington"~"West",
-              LHIN=="Erie St. Clair"~"West"
-    )
-  )%>%
-  distinct(city,.keep_all = TRUE)
-
-```
-
-Now, combine datasets and write csv file for those cities that were not assigned a Health Region.
-
-```{r,combine}
-
-clean_data<-left_join(clean_data,ltc_data,by="city")
-
- missing_health_regions<-clean_data %>% filter(is.na(LHIN)) %>%
-   distinct(city,.keep_all = TRUE)%>%
-   select(city,Geographic.area,Geographic_area_title,LHIN,Health_Region)
-
-#write.csv(missing_health_regions,here("data","missing_health_regions.csv"),row.names = FALSE)
-
-```
-
-To complete the missing LHIN information, the correspoding LHINs were obtained from the LHIN websites:
-
-https://www.southeasthealthline.ca/aboutourregion.aspx?region=RuralFrontenacLennoxAddington
-
-https://www.nsmhealthline.ca/aboutourregion.aspx
-
-https://www.champlainhealthline.ca/aboutourregion.aspx?region=StormontDundasGlen
-
-https://www.wwhealthline.ca/aboutourregion.aspx?region=KitchenerWaterlooWellesleyWilmotWoolwich
-
-Some cities did not belong entirely to a LHIN.
-For example, Etobicoke was divided between the Central, Central East, and Toronto Central LHINs.
-We chose in these cases the LHIN that covered the larger geographical region of each city (Toronto Central LHIN in the case of Etobicoke).
-We now assign LHINs to the dataset and Health Regions.
-The dataset with the information of the missing LHINs is called `missing_health_regions_updated.csv`
-
-```{r,merge-health-regions}
-
-#load health regions with missing entries fixed
-
-mhru<-read.csv(here("data","missing_health_regions_updated.csv"))
-
-#rename columns
-mhru <- mhru %>%
-  rename(Geographic_area_title=Geographic.area.title)
-
-clean_data<-clean_data %>%
-  left_join(mhru, by = c("city")) %>%
-   mutate(LHIN = coalesce(LHIN.x,LHIN.y)) %>%
-  select(-c(LHIN.x,LHIN.y,Full_title.x,Geographic.area.y,Geographic_area_title.y))%>%
-  mutate(Health_Region=
-    case_when(LHIN=="Central"~ "Central",
-              LHIN=="Central West"~"Central",
-              LHIN=="Mississauga Halton"~"Central",
-              LHIN=="North Simcoe Muskoka"~"Central",
-              LHIN=="Central East"~"East",
-              LHIN=="South East"~"East",
-              LHIN=="Champlain"~"East",
-              LHIN=="North East"~"North East",
-              LHIN=="North West"~"North West",
-              LHIN=="Toronto Central"~"Toronto",
-              LHIN=="South West"~"West",
-              LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)"~"West",
-              LHIN=="Waterloo Wellington"~"West",
-              LHIN=="Erie St. Clair"~"West"
-    ))%>%
-  rename(Geographic_area=Geographic.area.x,Geographic_area_title=Geographic_area_title.x)
-
-```
-
-At this point, the data has been cleaned and prepared for formal analysis.
-
-The clean dataset will be saved as a *.csv file.
-
-```{r,save-clean-dataset}
-#| echo: false
-
-write.csv(clean_data,here("data","clean_dataset.csv"),row.names = FALSE)
-```
+We used the iterative proportional fitting procedure (_raking_) in order to account for the differences in income, age groups, race/ethnicity and the population from the Health Regions.
+Details of the implementation can be found in the R Script `raking_logistic_regression.R` in the `code` directory.

### 2023-02-08 protocol/Background_and_methods_Jan_24.qmd
Fixed appendix, cleaning folder, latex preamble

--- 2023-02-06 protocol/Background_and_methods_Jan_24.qmd
@@
 | Variable             | Values                                                                                                                          |
-|:-------------------|----------------------------------------------------|
+|:--------------------|---------------------------------------------------|
 | Age group            | 15-24,25-34,35-44,45-54,55-64, 65+                                                                                              |

### 2023-02-08 code/data_preparation.R

@@ -1,5 +1,6 @@
-## this script takes the clean data set and does some preparatory steps such as recoding
-## variables to create a table summary
+# this script takes the clean dataset generated from the data cleaning process and does some preparatory steps such as 
+# variable recoding, creating a variable for the dependent variable in the regression, and creating 
+# a summary table of the dataset
 
 data<-read.csv(here("data","clean_dataset.csv"))
 

### 2023-02-08 code/raking.R

@@ -1,0 +1,97 @@
+#Each of these dataframes provides the population totals for the different variables and their categories
+
+
+Ages<-data.frame(Age_group=c("16_24",
+                             "25_34",
+                             "35_44",
+                             "45_54",
+                             "55_64",
+                             "65_and_over"),
+                 Freq=c(1707959,
+                        1734856,
+                        1721407,
+                        2003826,
+                        1842444,
+                        2245898))
+
+
+
+Races<-data.frame(Race=c("arab_middle_eastern",
+                         "black",
+                         "east_asian_pacific_islander",
+                         "indigenous",
+                         "latin_american",
+                         "mixed",
+                         "other",
+                         "south_asian",
+                         "white_caucasian"),
+                  Freq=c(212782,
+                         638346,
+                         886592,
+                         376558,
+                         197020,
+                         130033,
+                         705333,
+                         1166361,
+                         9118079 ))
+
+Incomes<-data.frame(Income=c("under_15000",
+                             "15000_24999",
+                             "25000_39999",
+                             "40000_59999",
+                             "60000_89999",
+                             "over_90000"),
+                    Freq=c(294643,
+                           387688,
+                           599624,
+                           796052,
+                           1007988,
+                           2083176
+                           
+                    ))
+
+
+# corrections for Health Regions, without the North East and North West regions
+
+Health_Regions<-data.frame(Health_Region=c(
+  #"North West",
+  #"North East",
+  "West",
+  "East",
+  "Central",
+  "Toronto"),
+  Freq=c(#232299,
+    #557000,
+    4095589,
+    3742520,
+    5032410,
+    1440644
+  )
+)
+
+
+
+### Raking ####
+
+## First, provide a survey design object. This is done using the dataset. The syntax
+## means that there is no stratification (id=~1). When the line below is run it will throw
+## a warning because we do not have "design" weights in the dataset (other datasets do have this).
+
+
+a1<-svydesign(id=~1,data=clean_data)
+
+## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage. 
+
+
+## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
+## from the data frames from above.
+
+a1_rake<-rake(a1,
+              sample=list(~Age_group,
+                          ~Race,
+                          ~Income,
+                          ~Health_Region),
+              population=list(Ages,
+                              Races,
+                              Incomes,
+                              Health_Regions))

### 2023-02-08 code/regression.R

@@ -1,0 +1,13 @@
+## This script uses the object created in the raking.R script to fit a fixed-effects logistic regression
+
+
+
+#call the script for raking
+source(here::here("code","raking.R"))
+
+# logistic regression model
+m1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income+Race*Income,
+           design=a1_rake,
+           family = quasibinomial(), 
+           control= list(maxit=25))
+

### 2023-02-10 manuscript/04-discussion.qmd
Working on Discussion

# Discussion

The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022]. The availability of vaccines at the end of 2020 made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021]. However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of Covid-19 have faced multiple challenges that have complicated the achievement of global immunity.

Among the different challenges faced by Covid-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and vaccine inequity  between low and high income countries[@tamey2022]. However, it is also well established that even in the case of high income countries that have had ample access to vaccines since 2020, such as the US, the UK, and Canada, there have been challenges in vaccination efforts due to differences in vaccine uptake among different segments of the population. More specifically, lower vaccine uptake has been associated with socio-economic factors such as race (i.e., identifying as Black, Asian, Indigenous) and household income (typically within lower income brackets) [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021]. Reasons given for this association have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020].

- Influence of geography

- ontario and Health Regions

- Results

- Limitations

### 2023-02-14 manuscript/04-discussion.qmd
working on discussion

@@
 
-Among the different challenges faced by Covid-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and vaccine inequity  between low and high income countries[@tamey2022].
+Among the different challenges faced by Covid-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and inequity in vaccine access between low and high income countries [@tamey2022].
 However, it is also well established that even in the case of high income countries that have had ample access to vaccines since 2020, such as the US, the UK, and Canada, there have been challenges in vaccination efforts due to differences in vaccine uptake among different segments of the population.
@@
 Reasons given for this association have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020].
+
+In addition, vaccine uptake is influenced by geography as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021].
+However, the majority of studies that have analyzed spatial differences in vaccination have been carried outside of Canada, where spatial analyses have been focused at the country level [@lavoie2022].
+
+In this regard, we explored spatial and socio-economic determinants of vaccination status in the province of Ontario.
+This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Networks (LHINs) and their incorporation into larger Health Regions [@dong2022].
+The idea behind the change was to address the inequalities that were identified under the LHIN approach [@tsasis2012], and in the case of Covid-19 vaccination.
+
+Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of a Covid-19 vaccine (@tbl-descriptive-stats).
+However, the likelihood of vaccination significant differed between age, race, and income (@tbl-model).
+These results are consistent with other studies that have shown significant lower vaccination in individuals within low income brackets [citations here].
+In our case, individuals with a household income below CAD 39,999 were around 60% less likely of having received the vaccine when compared to individuals in high income households (with an income above 90,000 CAD).
 

### 2023-02-14 manuscript/main.qmd
Working on Discussion
working on discussion

--- 2023-02-06 manuscript/main.qmd
@@
 ---
-title: '**Socio-economic and geographic determinants of Covid-19 vaccination**'
+title: '**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**'
 author: Ariel Mundo Ortiz
 format: pdf
+linestretch: 2
 bibliography: refs.bib
@@
 
+```{r}
+#| child: 04-discussion.qmd
+```
+
 # References

### 2023-02-14 manuscript/01-background.qmd
updated discussion

--- 2023-02-06 01-background.qmd
@@
 Indeed, multiple obstacles that have complicated vaccination efforts against COVID-19 have been identified: inequality in vaccine access, vaccine hesitancy, and differences in vaccination rates across different segments of the population [@gerretsen2021; @nafilyan2021; @malik2020].
-In the case of Canada, lower vaccine uptake has been associated with socio-economic factors such as younger age, educational level, presence of children in the household, lack of a regular healthcare provider, ethnic origin, and financial instability [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
+In the case of Canada, lower vaccine uptake has been associated with socio-economic factors such as younger age, educational level, presence of children in the household, lack of a regular healthcare provider, ethnic origin, and financial instability [@guay2022;@muhajarine2021;@hussain2022].
 

### 2023-02-14 manuscript/03-results.qmd
updated discussion

@@
   modify_header(label = "**Variable**") %>%
-  as_kable_extra(booktabs = TRUE) %>%
+  as_kable_extra(booktabs = TRUE,linesep="") %>%
   # reduce font size to make table fit. 
@@
 
+\footnotesize
+
+\renewcommand{\arraystretch}{0.5}
 ```{r,model-table}
@@
     bold_labels()%>%
-  as_kable_extra(booktabs = TRUE, longtable=T) %>%
+  as_kable_extra(booktabs = TRUE, longtable=T, linesep="") %>%
   # reduce font size to make table fit. 

### 2023-02-14 manuscript/04-discussion.qmd
updated discussion

--- 2023-02-10 manuscript/04-discussion.qmd
@@
 
+In addition, vaccine uptake is influenced by geography as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021].
+However, in the case of Canada, most studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022].
+Therefore, we explored spatial and socio-economic determinants of vaccination status in the province of Ontario.
+This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Networks (LHINs) and their incorporation into larger Health Regions [@dong2022].
+The idea behind the change aimed at reducing the inequalities in healthcare that were identified under the LHIN approach [@tsasis2012], and thus, examining if there are significant differences in vaccination between the Health Regions could be important to consider by decision-makers in the development of public health policies that aim to address intra-provincial health disparities.
+
+Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the Covid-19 vaccine (@tbl-descriptive-stats).
+This is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022].
+However, we also identified differences in vaccination likelihood based on different socio-economic factors.
+More specifically, our results show that those that were above 65 years old were only 44% likely to have received a vaccine when compared to those between 16 and 24 years old, that individuals with a household income below CAD 39,999 were around 60% less likely of having received the vaccine when compared to individuals in high income households (with an income above 90,000 CAD), and that individuals who identified as Arab and Middle Easterns, Black, Indigenous, Latin American, South Asian, Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere had significantly lower probability of vaccination  (ranging between 15% and 40%) when compared to White/Caucasian individuals (@tbl-model).
+
+These results are consistent with other studies that have shown an association between lower vaccination rates in racial minorities and lower household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
+However, it is interesting to note reduced vaccination likelihood in those aged 65 and over, as other studies have reported higher vaccination acceptance in this age group [@macdonald2021;@guay2022].
+
+Although there was no significant difference between vaccination likelihood among the Health Regions, there was a significant lower vaccination likelihood in the Central Health Region in individuals with a household income between CAD 15,000-24,999 (@tbl-model).
+This result would be consistent in the context of to Ontario Health, The Central Health Region comprises the former Central, Central West, Mississauga Halton, and North Simcoe Muskoka LHINs with 48.3% immigrant population and 47% of the population identifying as visible minority, the highest percentage among all regions [@ontariohealth].
+
 - Influence of geography
 
-- ontario and Health Regions
+- ontario and Health Regions: find information about poverty and racial groups.
 

### 2023-02-21 data_cleaning/Data_cleaning_Feb_17.qmd
re-grouped data and ran model, fixed code

--- 2023-02-08 data_cleaning/Data_cleaning_final.qmd
@@
 library(lubridate)
+library(gtsummary)
 library(patchwork)
@@
   select(age_group,
-         employed,
-         h_size,
-         income,race,
+         income,
+         race,
          remote_work,
@@
          city,
-         h_size,
-         date,
-         first_dose,
-         second_dose,
-         get_vaccine)%>%
-  drop_na(employed)
+         first_dose)%>%
+  drop_na(age_group,
+          income,
+          race,
+          city,
+          first_dose)
+
+data_summary<-clean_data %>%
+  tbl_summary(percent="row",
+              by= first_dose,
+              include=c(income,
+                        age_group,
+                        race)
+  ) %>% bold_labels()
+
+## original clean dataset
+
+data1<-read.csv(here("data","clean_dataset.csv"))
+
+data_summary2<-data1 %>%
+  drop_na(race)%>%
+  tbl_summary(percent="row",
+              by= first_dose,
+              include=c(income,
+                        age_group,
+                        race)
+  ) %>% bold_labels()
+
+tbl_merge(tbls=list(data_summary2,data_summary),tab_spanner = c("Original Clean Dataset","Dataset choosing covariates first"))
 

### 2023-02-21 data_cleaning/data_grouping_Feb_20.qmd
re-grouped data and ran model, fixed code

---
title: "**Data cleaning and model testing with grouped categories**"
subtitle: Re-working groups
author: "Ariel Mundo"
format: pdf
toc: true
toc-depth: 2
editor: source
urlcolor: blue
#bibliography: refs.bib
---

```{r, setup, echo=FALSE}
#| warning: false
library(tidyverse)
library(here)
library(lubridate)
library(gtsummary)
library(patchwork)
library(visdat)
library(survey)
library(gt)
library(scales)
library(scico)

thm1<-scale_fill_scico_d(palette="buda",begin=0.3, end=0.8, direction = -1, aesthetics = "fill")

thm2<-scale_fill_scico_d(palette="oleron",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")

thm3<-scale_fill_scico_d(palette="fes",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")
```

## Process

Merge income as: below 24k, 25-60k, 60k and above
Age: 16-34, 35-54, 55 and over

Interactions: Race * Income, Race * Health Region

```{r, prepare-data}

data1<-read.csv(here("data","clean_dataset.csv"))

data1 <- data1 %>% 
  drop_na(race)

data_summary2<-data1 %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income,
                        age_group,
                        race,
                        Health_Region)) %>% 
  bold_labels()

data_summary2

## Recode income

data1<-data1%>%
    mutate(income_ord = case_when(
                                  income %in% "under_15000" ~ "under_25000",
                                  income %in% "15000_24999" ~ "under_25000",
                                  income %in% "25000_39999" ~ "25000_59999",
                                  income %in% "40000_59999" ~ "25000_59999",
                                  income %in% "60000_89999" ~ "60000_and_above",
                                  income %in% "90000_109999" ~ "60000_and_above",
                                  income %in%"over_110000" ~ "60000_and_above",
                                ))%>%
  mutate(income_ord=as.factor(income_ord))

data1<-data1%>%
    mutate(age_group_ord = case_when(
                                  age_group %in% "16_24" ~ "16_34",
                                  age_group %in% "25_34" ~ "16_34",
                                  age_group %in% "35_44" ~ "35_54",
                                  age_group %in% "45_54" ~ "35_54",
                                  age_group %in% "55_64" ~ "55_and_over",
                                  age_group %in% "65_and_over" ~ "55_and_over",
                                ))%>%
  mutate(age_group_ord=as.factor(age_group_ord))

```

```{r,raking}
#remove north health regions (low numbers)

#get health regions to eliminate

data1<-data1 %>%
  select(-Municipal.status) %>%
  rename(Race=race,Age_group_ord=age_group_ord,Income=income_ord)

data1<-data1 %<>%
  mutate_at(c("Income","Race","Age_group_ord","Health_Region"),factor)

data1 <- data1 %>%
  filter(Health_Region!="North East" , Health_Region!="North West")

#eliminate the levels of the removed health regions
data1 <- droplevels(data1)

```

# Raking

```{r,raking}
Ages<-data.frame(Age_group_ord=c("16_34",
                             "35_54",
                             "55_and_over"),
                 Freq=c(3442815,
                        3725233,
                        4088342))

Races<-data.frame(Race=c("arab_middle_eastern",
                         "black",
                         "east_asian_pacific_islander",
                         "indigenous",
                         "latin_american",
                         "mixed",
                         "other",
                         "south_asian",
                         "white_caucasian"),
                  Freq=c(212782,
                         638346,
                         886592,
                         376558,
                         197020,
                         130033,
                         705333,
                         1166361,
                         9118079 ))

Incomes<-data.frame(Income=c("under_25000",
                             "25000_59999",
                             "60000_and_above"),
                    Freq=c(682331,
                           1395676,
                           3091164
                    ))

# corrections for Health Regions, without the North East and North West regions

Health_Regions<-data.frame(Health_Region=c(
  #"North West",
  #"North East",
  "West",
  "East",
  "Central",
  "Toronto"),
  Freq=c(#232299,
    #557000,
    4095589,
    3742520,
    5032410,
    1440644
  )
)

data1<-data1 %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))

data1<- within(data1, Age_group_ord <- relevel(Age_group_ord, ref = "16_34"))
data1<- within(data1, Race <- relevel(Race, ref = "white_caucasian"))
data1<- within(data1, Health_Region <- relevel(Health_Region, ref = "Toronto"))
data1<- within(data1, Income <- relevel(Income, ref = "60000_and_above"))

a1<-svydesign(id=~1,data=data1)

## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage.

## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
## from the data frames from above.

a1_rake<-rake(a1,
              sample=list(~Age_group_ord,
                          ~Race,
                          ~Income,
                          ~Health_Region),
              population=list(Ages,
                              Races,
                              Incomes,
                              Health_Regions))

#model

m1<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

m2<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region+Race*Income,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

m3<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region+Race*Health_Region,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

m4<-svyglm(first_dose_m~Age_group_ord+Income+Race+Health_Region+Race*Income+Race*Health_Region,
           design=a1_rake,
           family = quasibinomial(), 
           control= list(maxit=25))

t1<-m1 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

t2<-m2 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

t3<-m3 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

t4<-m4 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()

tm<-tbl_merge(tbls=list(t1,t2,t3,t4),tab_spanner = c("No interactions","Race x Income","Race x Health Region", "Race x Income, Race x Health Region "))

tm

t5<-data1 %>% 
  tbl_cross(row = Race, col = Income)

t6<- data1 %>%
  tbl_cross (row=Race, col= Health_Region)

tbl_stack(tbls=list(t4,t5),group_header = c("Race and Income", "Race and Health Region"))

```

### 2023-02-21 manuscript/02-methods.qmd
re-grouped data and ran model, fixed code

@@
 
-Additionally, users who indicated they were under the age of 16 were exited from the survey, no record was created, and these users were unable to navigate back to the "age select" screen.
-The personal identifier information from each user was automatically scrubbed and replaced by a unique ID.
+Additionally, users who indicated they were under the age of 16 were exited from the survey without creating a record, furthermore, these users were unable to navigate back to the "age select" screen.
+The personal identifier information from each user that participated in the survey was automatically scrubbed and replaced by a unique ID.
 Respondents were drawn exclusively from the province of Ontario, as per their devices meta-data.
 
-## Survey responses
+## Survey responses {#sec-socio-demographic-factors}
 
-### Socio-demographic factors {#sec-socio-demographic-factors}
+Socio-economic information selected from the survey answers included age group, income bracket, and race/ethnicity.
+The levels of each socio-economic factor used for analysis appear in @tbl-covariates.
 
-Socio-economic information selected for analysis from the survey answers included age group, income bracket, and race/ethnicity.
-Information about these socio-economic factors in the survey is provided in @tbl-covariates.
-
-| Variable             | Values                                                                                                                          |
-|:------------------|-----------------------------------------------------|
-| Age group            | 16-24,25-34,35-44,45-54,55-64, 65+                                                                                              |
-| Income bracket (CAD) | \<15,000, 15,000-24,999, 25,000-39,999, 40,000-59,999, 60,000-89,999, \>90,000                                                  |
+| Variable             | Levels                                            |            
+|:---------------------|---------------------------------------------------|
+| Age group            | 16-34,35-54,55 and over                           |
+| Income bracket (CAD) | under 25,000, 25,000-59,999, 60,000 and above     |
 | Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, Other |
-|                      |                                                                                                                                 |
 
@@
 
-We used a fixed-effects logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors and the Health Regions from @sec-geographical-location, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
+We used a logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @sec-socio-demographic-factors, the Health Regions from @sec-geographical-location, and the interactions between Race and Health Region, and Race and income as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
 The model appears in @eq-model3,
@@
 \begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Income)} + \beta_6 \textrm{ (Age group} \times \textrm{Income)} + \beta_7 \textrm{ (Race} \times \textrm{Income)}
+\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Race)} + \beta_6 \textrm{ (Income} \times \textrm{Race)}
 \end{aligned}
@@
 
-Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, $\beta_1...\beta_7$ indicate the coefficients for each of the fixed and interaction effects of age group, race, Health Region, and income.
+Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term.
 The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.

### 2023-02-21 manuscript/03-results.qmd
re-grouped data and ran model, fixed code

@@
 @tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors.
-The reference groups in each case were set as follows: 16 to 24 years (age group), White Caucasian (Race), Toronto (Health Region), Over CAD 90,000 (Income).
-Socio-economic factors with statistically significant odds ratios were the age group of 65 years and over (OR=0.44), persons who identified as Arab/Middle Eastern (OR=0.15), Black (OR=0.17), Indigenous (OR=0.27), Latin American (OR=0.24), South Asian (OR=0.4) and Other Race/Ethnicity (a group those that identified as Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere, OR=0.17), and those that lived in a household with an income bracket between CAD 25,000-39,999 (OR=0.23).
+The reference groups in each case were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income).
+Statistically significant odds ratios were identified for individuals with a household income under CAD 25,000 (OR=0.37) and between CAD 25,000-59,999 (OR=0.59), persons who identified as Arab/Middle Eastern (OR=0.31), Black (OR=0.32), Latin American (OR=0.28), Other Race/Ethnicity (a group those that identified as Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere, OR=0.22), and inviduals living in the West Health Region (OR=1.55).
 
-Regarding the interaction effects, the interaction of the Central Health Region and income bracket between 15000 to 24999 CAD was statistically significant (OR=0.4).
-In age group and income, significant effects were estimated for those 65 years and older with a household income between CAD 25000-39999 (OR=7.21), and with an income between CAD 40000-59999 (OR=3.12).
-In the case of race and income, significant effects were estimated for individuals that identified as Arab/Middle Easterners with a family income between 15000 and 24999 (OR=5.68) or income under CAD 15000 (OR=4.21), Black individuals with a family income under CAD 15000 (OR=4.26) or between CAD 15000-24999 (OR=3.35); additionally, there were statistically significant interaction effects in the case of those that identified as belonging to a race not identified elsewhere ("Other") within multiple income brackets: under CAD 15000 (OR=3.73), between 15000 and 24999 (OR 11.2), between 25000 and 39999 CAD (OR=4.67), and with income CAD 40000-59999 (OR=9.58).
-Finally, there was a significant interaction effect for South Asian individuals with an income under CAD 15000 (OR=3.10).
+Moreover, statistically-significant higher odd ratios were determined in the case of income and race; specifically, for individuals with a household income below CAD 25,000 that identified as Arab/Middle Eastern (OR=3.05), Black (3.19), Latin American (OR=2.80), and Other (4.59); and for individuals with a household income between CAD 25,000 and 59,999 that identified as Other racial minority (OR=6.93).
+
+For Health Region and race, significant lower vaccination odds were identified for Black individuals in the Central Health Region (OR=0.44), and Other racial minorities and South Asian individuals that live in the West Health Region (ORs=0.41).
 
@@
                  label=list(Health_Region ~ "Health Region",
-                 Age_group ~ "Age Group")) %>%
+                 Age_group_ord ~ "Age Group")) %>%
     bold_labels()%>%

### 2023-02-21 manuscript/04-discussion.qmd
re-grouped data and ran model, fixed code

@@
 
-Although there was no significant difference between vaccination likelihood among the Health Regions, there was a significant lower vaccination likelihood in the Central Health Region in individuals with a household income between CAD 15,000-24,999 (@tbl-model).
-This result would be consistent in the context of to Ontario Health, The Central Health Region comprises the former Central, Central West, Mississauga Halton, and North Simcoe Muskoka LHINs with 48.3% immigrant population and 47% of the population identifying as visible minority, the highest percentage among all regions [@ontariohealth].
+Although there was no significant difference between vaccination odds among the Health Regions, there was a significant lower vaccination odds in the Central Health Region in individuals with a household income between CAD 15,000-24,999 (OR=0.4, @tbl-model).
+The Central Health Region comprises the former Central, Central West, Mississauga Halton, and North Simcoe Muskoka LHINs with 48.3% of immigrant population and 47% of the population identifying as visible minority, the highest percentage among all Health Regions [@ontariohealth]; moreover, the fact that the Central West LHIN, now within the Central Health Region, had the highest proportion of low-income families in 2018 (12.9\% of males and 16.7\% of females among the lowest 20% of Canadian family incomes [@buajitti2018])
 

### 2023-02-21 manuscript/appendix.qmd
re-grouped data and ran model, fixed code

@@
 It can be seen that there are differences in the distribution of each ethnicity between the Census data and the survey.
-For example, Arabs/Middle Easteners correspond to 5.9% of the survey responses, but population-wise individuals that identify in this ethnic group correspond to 1.6% of the Ontario population.
+For example, Arabs/Middle Easterner correspond to 5.9% of the survey responses, but population-wise individuals that identify in this ethnic group correspond to 1.6% of the Ontario population.
 

### 2023-02-21 code/data_preparation.R

@@ -33,20 +33,52 @@ clean_data <-clean_data %>%
 #eliminate the levels of the removed health regions
 clean_data <- droplevels(clean_data)
 
+
+## Recode income
+
+
+clean_data<-clean_data%>%
+  mutate(income_ord = case_when(
+    income %in% "under_15000" ~ "under_25000",
+    income %in% "15000_24999" ~ "under_25000",
+    income %in% "25000_39999" ~ "25000_59999",
+    income %in% "40000_59999" ~ "25000_59999",
+    income %in% "60000_89999" ~ "60000_and_above",
+    income %in% "90000_109999" ~ "60000_and_above",
+    income %in%"over_110000" ~ "60000_and_above",
+  ))%>%
+  mutate(income_ord=as.factor(income_ord))
+
+#re-group age groups
+
+clean_data<-clean_data%>%
+  mutate(Age_group_ord = case_when(
+    Age_group %in% "16_24" ~ "16_34",
+    Age_group %in% "25_34" ~ "16_34",
+    Age_group %in% "35_44" ~ "35_54",
+    Age_group %in% "45_54" ~ "35_54",
+    Age_group %in% "55_64" ~ "55_and_over",
+    Age_group %in% "65_and_over" ~ "55_and_over",
+  ))%>%
+  mutate(Age_group_ord=as.factor(Age_group_ord))
+
+
 ## relevel for reference groups in the model
 
-clean_data<- within(clean_data, Age_group <- relevel(Age_group, ref = "16_24"))
+clean_data<- within(clean_data, Age_group_ord <- relevel(Age_group_ord, ref = "16_34"))
 clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
-clean_data<- within(clean_data, Income <- relevel(Income, ref = "over_90000"))
+clean_data<- within(clean_data, income_ord <- relevel(income_ord, ref = "60000_and_above"))
+
 
 #create summary table
 
 data_summary<-clean_data %>%
   tbl_summary(percent="row",
               by= first_dose,
-              include=c(Income,
-                        Age_group,
+              include=c(income_ord,
+                        Age_group_ord,
                         Health_Region,
                         Race)
   ) %>% bold_labels()
+

### 2023-02-21 code/raking.R

@@ -1,18 +1,12 @@
 #Each of these dataframes provides the population totals for the different variables and their categories
 
 
-Ages<-data.frame(Age_group=c("16_24",
-                             "25_34",
-                             "35_44",
-                             "45_54",
-                             "55_64",
-                             "65_and_over"),
-                 Freq=c(1707959,
-                        1734856,
-                        1721407,
-                        2003826,
-                        1842444,
-                        2245898))
+Ages<-data.frame(Age_group_ord=c("16_34",
+                                 "35_54",
+                                 "55_and_over"),
+                 Freq=c(3442815,
+                        3725233,
+                        4088342))
 
 
 
@@ -35,19 +29,12 @@ Races<-data.frame(Race=c("arab_middle_eastern",
                          1166361,
                          9118079 ))
 
-Incomes<-data.frame(Income=c("under_15000",
-                             "15000_24999",
-                             "25000_39999",
-                             "40000_59999",
-                             "60000_89999",
-                             "over_90000"),
-                    Freq=c(294643,
-                           387688,
-                           599624,
-                           796052,
-                           1007988,
-                           2083176
-                           
+Incomes<-data.frame(income_ord=c("under_25000",
+                             "25000_59999",
+                             "60000_and_above"),
+                    Freq=c(682331,
+                           1395676,
+                           3091164
                     ))
 
 
@@ -87,9 +74,9 @@ a1<-svydesign(id=~1,data=clean_data)
 ## from the data frames from above.
 
 a1_rake<-rake(a1,
-              sample=list(~Age_group,
+              sample=list(~Age_group_ord,
                           ~Race,
-                          ~Income,
+                          ~income_ord,
                           ~Health_Region),
               population=list(Ages,
                               Races,

### 2023-02-21 code/regression.R

@@ -6,8 +6,10 @@
 source(here::here("code","raking.R"))
 
 # logistic regression model
-m1<-svyglm(first_dose_m~Age_group+Race+Health_Region*Income+Health_Region+Age_group*Income+Race*Income,
+
+m1<-svyglm(first_dose_m~Age_group_ord+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
            design=a1_rake,
            family = quasibinomial(), 
            control= list(maxit=25))
+
 

### 2023-02-24 manuscript/01-background.qmd
working on discussion

@@
 For example, in the US and Canada vaccine campaigns began in mid-December of 2020 [@tanne2020; @bogoch2022].
-Although it has been estimated that vaccines against COVID-19 have prevented around 14 millions of deaths worldwide [@watson2022], the rollout of COVID-19 vaccines has faced multiple challenges since its inception.
+Although it has been estimated that COVID-19 vaccines have prevented around 14 millions of deaths worldwide [@watson2022], the rollout of COVID-19 vaccines has faced multiple challenges since its inception.
 

### 2023-02-24 manuscript/03-results.qmd
working on discussion

@@
 Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it.
-Regarding socio-economic factors, respondents who reported living in a household with an income under CAD 15,000 or above 90,000 represented approximately 50\% of the total number of entries, with 71\% and 75\% of the respondents in each income bracket indicating to have received the first dose of the vaccine, respectively.
-The age group between 16-24 years had the highest representation in the survey responses (837, 23.6\% of all responses), followed by the age group of those between 65 and over (692, 19.5\% of the total); 73\% of those in the 16 to 24 years age group reported to have received the vaccine, whereas 68\% of those above 65 years indicated the same.
-The Health Region with the highest number of participants in the survey was Toronto, accounting for 1,342 entries (37.7\%) and a 72\% vaccination rate.
-Regarding race, individuals that identified as White/Caucasian represented 1313 (37\%) of all entries and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine.
-On the other hand, those that identified as Indigenous had the lowest vaccine uptake with 60\% indicating to have received the vaccine.
+Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\% of the total number of entries, those within the CAD 25,000-59,999 income bracket represented 25% of the total sample, and those with an income above CAD 60,000 represented 38 \% of the sample; across all income brackets, the percentage of individuals that reported having received a first dose of the vaccine was consistent, above 69\%.
+
+Within the age groups of survey respondents, the age group between 16-34 years had the highest representation in the survey responses (1,521, 42.8\% of all responses).
+Within this age bracket, 73\% of respondents indicated having received the vaccine, whereas the lowest vaccination rate was in the bracket of those 55 years of age and above, with a total of 72\%.
+The Health Region with highest representation in the survey was Toronto, accounting for 1,324 entries (37.2\%), with a vaccination rate of 72\%.
+Regarding race/ethnicity, individuals that identified as White/Caucasian represented 1313 (37\%) of all entries and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine.
+On the other hand, the ethnic group with the lowest number of entries in the survey was Latin American, with a total of 180, or 5\% of all entries.
+Vaccination rates across all minority groups were below the value reported by White/Caucasians, with the lowest vaccination rate (60\%) being reported by individuals that identified as Indigenous.
 
@@
 
-@tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors.
-The reference groups in each case were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income).
-Statistically significant odds ratios were identified for individuals with a household income under CAD 25,000 (OR=0.37) and between CAD 25,000-59,999 (OR=0.59), persons who identified as Arab/Middle Eastern (OR=0.31), Black (OR=0.32), Latin American (OR=0.28), Other Race/Ethnicity (a group those that identified as Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere, OR=0.22), and inviduals living in the West Health Region (OR=1.55).
+@tbl-model shows the results of the logistic regression on vaccination status using socio-economic factors (age group, income, race), geographical areas (Health Regions) and the interactions between income and race and Health Region and race.
+The reference groups were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income).
+There were no statistically significant differences in vaccination rates within the age groups, but significant odds ratios were estimated for individuals within the rest of the covariates.
+Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, respectively).
+Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.28, respectively); additionally, those individuals in the Other Race/Ethnicity group (a group that included Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere) had even lower odds of vaccination than the other minority groups, with an OR of 0.22.
+Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.55).
 
-Moreover, statistically-significant higher odd ratios were determined in the case of income and race; specifically, for individuals with a household income below CAD 25,000 that identified as Arab/Middle Eastern (OR=3.05), Black (3.19), Latin American (OR=2.80), and Other (4.59); and for individuals with a household income between CAD 25,000 and 59,999 that identified as Other racial minority (OR=6.93).
+Moreover, statistically-significant higher odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05), Black (OR=3.19), Latin American (OR=2.80), or that belonged to other minority groups  (OR=4.59).
+Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.93).
 
-For Health Region and race, significant lower vaccination odds were identified for Black individuals in the Central Health Region (OR=0.44), and Other racial minorities and South Asian individuals that live in the West Health Region (ORs=0.41).
+For the interaction of Health Region and race, significant lower vaccination odds were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44).
+Finally, individuals that identified as part of other racial minorities or South Asian iand that lived in the West Health Region had significantly lower odds of vaccination  (ORs=0.41).
 

### 2023-02-24 manuscript/04-discussion.qmd
working on discussion

@@
 
-The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022].
-The availability of vaccines at the end of 2020 made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021].
+The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022], and their availability towards the end of 2020 in certain countries made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021].
 However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of Covid-19 have faced multiple challenges that have complicated the achievement of global immunity.
@@
 This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Networks (LHINs) and their incorporation into larger Health Regions [@dong2022].
-The idea behind the change aimed at reducing the inequalities in healthcare that were identified under the LHIN approach [@tsasis2012], and thus, examining if there are significant differences in vaccination between the Health Regions could be important to consider by decision-makers in the development of public health policies that aim to address intra-provincial health disparities.
+The idea behind the change aimed at reducing the inequalities in healthcare that were identified under the LHIN approach [@tsasis2012], and thus, examining differences in vaccination between the Health Regions can provide decision makers with information regarding intra-provincial health disparities that may need to be addressed in the future.
 
@@
 This is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022].
-However, we also identified differences in vaccination likelihood based on different socio-economic factors.
-More specifically, our results show that those that were above 65 years old were only 44% likely to have received a vaccine when compared to those between 16 and 24 years old, that individuals with a household income below CAD 39,999 were around 60% less likely of having received the vaccine when compared to individuals in high income households (with an income above 90,000 CAD), and that individuals who identified as Arab and Middle Easterns, Black, Indigenous, Latin American, South Asian, Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere had significantly lower probability of vaccination  (ranging between 15% and 40%) when compared to White/Caucasian individuals (@tbl-model).
+However, we identified differences in vaccination based on socio-economic and geographical factors.
+First, our results show significant differences in vaccination odds in individuals with a household income below CAD 60,000, and in individuals belonging to minority racial or ethnic groups.
+Those who identified as Black, Latin American, or  belonging to different minorities not included as categories in the survey (Southeast Asian, Filipino, West Asian, and Minority not identified elsewhere) had vaccination odds below 33% when compared to individuals that identified as White/Caucasian (@tbl-model).
+Geographically, significantly higher vaccination odds were identified in the West Health Region, which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
 
-These results are consistent with other studies that have shown an association between lower vaccination rates in racial minorities and lower household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
-However, it is interesting to note reduced vaccination likelihood in those aged 65 and over, as other studies have reported higher vaccination acceptance in this age group [@macdonald2021;@guay2022].
+The lack of difference in vaccination uptake across different age groups is consistent with other studies that have shown similar results [@macdonald2021;@guay2022].
+In the case of Ontario, where there was around 90% of vaccination uptake in the age group above 65 years old, and xxx percentages in xxx group and xxx group, these results show an overall high vaccination uptake at the provincial level.
+These results are consistent with other studies that have shown an association between lower vaccination uptake, identifying as part of a racial minority, or having a low household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
 

### 2023-03-01 manuscript/02-methods.qmd
Working on discussion, updated appendix

@@
 
-Socio-economic information selected from the survey answers included age group, income bracket, and race/ethnicity.
+Socio-economic information selected from the survey answers included age group and income brackets, and race/ethnicity.
 The levels of each socio-economic factor used for analysis appear in @tbl-covariates.
@@
 Following a preliminary analysis, the dataset was cleaned in order to only contain the socio-economic information provided in @tbl-covariates and vaccination status.
-The cleaning process also included removing outliers that were identified during the preliminary analyses.
-Specifically, we removed those respondents that indicated to be below 25 years of age, living in a household of size 1, and that reported an income above CAD 110,000.
-The next step consisted in processing the geographical information in the survey (city where the survey was responded) in order to match each city to its correspondent Health Region.
-Details of this process are given below.
+The cleaning process also included removing outliers that were identified during the preliminary analyses, and processing the geographical information in the survey (city where the survey was responded) in order to match each city to its correspondent Health Region.
 

### 2023-03-01 manuscript/03-results.qmd
Working on discussion, updated appendix

@@
 #| label: tbl-descriptive-stats
-#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey"
+#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey (by Vaccination Status)"
 #| message: false
@@
 source(here::here("code","data_preparation.R"))
+
+#create summary table
+
+data_summary<-clean_data %>%
+  tbl_summary(percent="row",
+              by= first_dose,
+              include=c(income_ord,
+                        Age_group_ord,
+                        Health_Region,
+                        Race)
+  ) %>% bold_labels()
 
@@
 The reference groups were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income).
-There were no statistically significant differences in vaccination rates within the age groups, but significant odds ratios were estimated for individuals within the rest of the covariates.
+There were no statistically significant differences in vaccination rates within the age groups from the survey, but significant odds ratios were estimated for other covariates.
 Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, respectively).
@@
 
-Moreover, statistically-significant higher odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05), Black (OR=3.19), Latin American (OR=2.80), or that belonged to other minority groups  (OR=4.59).
+Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05), Black (OR=3.19), Latin American (OR=2.80), or that belonged to other minority groups  (OR=4.59).
 Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.93).
 
-For the interaction of Health Region and race, significant lower vaccination odds were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44).
-Finally, individuals that identified as part of other racial minorities or South Asian iand that lived in the West Health Region had significantly lower odds of vaccination  (ORs=0.41).
+For the interaction of Health Region and race, significant odds of vaccination were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41).
 
 \footnotesize
-
 \renewcommand{\arraystretch}{0.5}
@@
 ```
+
+\normalsize

### 2023-03-01 manuscript/04-discussion.qmd
Working on discussion, updated appendix

@@
 In addition, vaccine uptake is influenced by geography as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021].
-However, in the case of Canada, most studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022].
+However, in the case of Canada, studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022].
 Therefore, we explored spatial and socio-economic determinants of vaccination status in the province of Ontario.
-This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Networks (LHINs) and their incorporation into larger Health Regions [@dong2022].
-The idea behind the change aimed at reducing the inequalities in healthcare that were identified under the LHIN approach [@tsasis2012], and thus, examining differences in vaccination between the Health Regions can provide decision makers with information regarding intra-provincial health disparities that may need to be addressed in the future.
+This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Network (LHIN) system and the incorporation of regions covered by LHINs into larger Health Regions [@dong2022].
+Because the idea behind the change aimed to reduce the inequalities in healthcare that were identified under the LHIN model [@tsasis2012], examining differences in vaccination between the Health Regions can provide decision-makers with insight regarding intra-provincial health disparities that may need to be addressed in future vaccination or public health campaigns.
 
-Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the Covid-19 vaccine (@tbl-descriptive-stats).
-This is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022].
-However, we identified differences in vaccination based on socio-economic and geographical factors.
-First, our results show significant differences in vaccination odds in individuals with a household income below CAD 60,000, and in individuals belonging to minority racial or ethnic groups.
-Those who identified as Black, Latin American, or  belonging to different minorities not included as categories in the survey (Southeast Asian, Filipino, West Asian, and Minority not identified elsewhere) had vaccination odds below 33% when compared to individuals that identified as White/Caucasian (@tbl-model).
-Geographically, significantly higher vaccination odds were identified in the West Health Region, which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the Covid-19 vaccine (@tbl-descriptive-stats), and that there were no significant differences in vaccination odds among the age groups considered in the survey.
+This result is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022], and with the vaccination information provided by Public Health Ontario,  which shows that for the period where the Fields survey ran (Sept 30, 2021-Jan 17, 2022) there was a minimum of  80% of first dose vaccination coverage among all the age groups considered in the survey [@ontario-covid].
 
-The lack of difference in vaccination uptake across different age groups is consistent with other studies that have shown similar results [@macdonald2021;@guay2022].
-In the case of Ontario, where there was around 90% of vaccination uptake in the age group above 65 years old, and xxx percentages in xxx group and xxx group, these results show an overall high vaccination uptake at the provincial level.
-These results are consistent with other studies that have shown an association between lower vaccination uptake, identifying as part of a racial minority, or having a low household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
+However, we identified intra-provincial differences in vaccination based on socio-economic and geographical factors.
+First, our results show significant differences in vaccination odds in individuals with a household income below CAD 60,000 and in individuals belonging to visible minority groups.
+Those who identified as Black, Latin American, or  belonging to a minority group not included in the survey (Southeast Asian, Filipino, West Asian, and Minority not identified elsewhere) had vaccination odds below 33% when compared to individuals that identified as White/Caucasian (@tbl-model).
+These results are consistent with other studies that have shown lower vaccination rates in individuals that identify as part of a racial minority, or that have a low household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
+
+In this study, we also decided to explore the interactions between income and race and race and Health Region, as it is known that many individuals within racial minority groups perform tend to occupy certain types of occupations that fall within income brackets that have been shown to be associated with differences in vaccination uptake.
+In other words, we decided to explore if there were differences in vaccination within racial groups in certain income brackets and in certain the Health Regions.
+In this regard, it is interesting to note that although overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals (@tbl-model), the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, for which vaccination odds decreased in lower income brackets (when compared to the CAD 60,000 and over bracket, Supplementary Figure A-3).
+More specifically, the change in odds of vaccination increased in individuals who identified as Arab/Middle Eastern, Black, Latin American, or belonging to other minority groups with a household income below CAD 25,000, which was also true for individuals in other racial minority groups with an income between CAD 25,000-59,999 (@tbl-model, Supplementary Figure A-3).
+
+This result is likely due to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include occupations such as grocery store workers, gas station workers, warehouse and distribution workers, and manufacturing workers, all being occupations for which a low income is to be expected.
+In the case of Ontario, essential workers had priority for Covid-19 vaccination [@mishra2021], which would explain the higher odds of vaccination for these individuals in certain income brackets, in contrast to the lower odds of vaccination for the same type of individuals in higher income brackets.
+In other words, it is possible that the type of occupation played an important role in the odds of vaccination in these racial minority groups.
+
+Additionally, significant higher vaccination odds were identified in the West Health Region when compared to the Health Region of Toronto (@tbl-model).
+The West Health Region comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+In this case, overall higher odds of vaccination can be explained by the fact that in the survey, about 47% of the entries for this Health Region corresponded to White/Caucasian individuals, who reported an overall 83% vaccination rate (Supplementary Table A-6).
+However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@tbl-model].
+In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-4).
+
+The West Health Region, according to Ontario Health, has 13.2% of its population identifying as visible minority, and 2.5% of it identifying as Indigenous [@ontariohealth].
+In the case of this analysis, it is possible that these significant interactions are due to the fact that 50% of the answers from this region came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
+Clair, which were previously reported to have more than 20% of their population in the lowest income quintile [@buajitti2018] (Supplementary Table A-7) while about 55% of the answers from areas previously under the aforementioned LHINs also corresponded to individuals that identified in the minority groups.
+This would mean that although all minority groups had lower odds of vaccination in the rest of the Health Regions, the combination of higher representation of individuals from low-income areas that identified as minorities resulted in the overall significant lower odds of vaccination.
 

### 2023-03-01 manuscript/appendix.qmd
Working on discussion, updated appendix

@@
 library(tidyverse)
+library(magrittr)
 library(here)
-library(lubridate)
 library(patchwork)
-library(visdat)
 library(gt)
-library(scales)
+library(gtsummary)
+library(survey)
+library(emmeans)
 library(scico)
@@
 thm3<-scale_fill_scico_d(palette="fes",begin=0.0, end=1.0, direction = -1, aesthetics = "fill")
+
+thm4<-scale_fill_scico_d(palette="tokyo",begin=0.3, end=0.8, direction = -1, aesthetics = c("colour","fill"))
+
+thm5<-scale_fill_scico_d(palette="imola",begin=0.3, end=0.8, direction = -1, aesthetics = c("colour","fill"))
 
@@
 Details of the implementation can be found in the R Script `raking_logistic_regression.R` in the `code` directory.
+
+# Interactions from the Regression Model
+
+The plots below show the interactions for Race and Income and Race and Health Region based on the regression model presented in the main paper.
+
+```{r,interaction-race-income}
+#| echo: false
+#| warning: false
+#| fig-height: 4
+#| fig-width: 6
+#| fig-cap: Interaction plot between Race and Income
+
+#call scripts for data preparation and regression
+source(here::here("code","data_preparation.R"))
+source(here::here("code","regression.R"))
+
+race_income<-emmip(m1, Race ~ income_ord, CIs=TRUE,plotit = FALSE)
+
+ggplot(data=race_income,aes(x=income_ord,group=Race,color=Race,fill=Race))+
+  geom_ribbon(aes(ymin=LCL,ymax=UCL),alpha=0.5,show.legend = FALSE)+
+  geom_line(aes(group=Race,y=yvar),show.legend = FALSE)+
+  facet_wrap(~Race)+
+  ylab("Linear predictor")+
+  theme_classic()+
+  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
+  thm5
+```
+
+```{r,interaction-race-hr}
+#| echo: false
+#| warning: false
+#| fig-height: 4
+#| fig-width: 6
+#| fig-cap: Interaction plot between Race and Health Region
+
+race_hr<-emmip(m1, Race ~ Health_Region, CIs=TRUE,plotit = FALSE)
+
+ggplot(data=race_hr,aes(x=Health_Region,group=Race,color=Race,fill=Race))+
+  geom_ribbon(aes(ymin=LCL,ymax=UCL),alpha=0.5,show.legend = FALSE)+
+  geom_line(aes(group=Race,y=yvar),show.legend = FALSE)+
+  facet_wrap(~Race)+
+  ylab("Linear predictor")+
+  theme_classic()+thm4
+
+```
+
+# Socio-economic Information from the Survey Data
+
+The tables below present summaries of information from the Fields survey (from the clean dataset analyzed in the main paper), showing responses by White/Caucasian individuals by Health Region and proportion of answers by LHIN in the West Health Region.
+
+```{r,health-region}
+#| label: tbl-hr-percentages
+#| tbl-cap: "Reported Vaccination Status from the Survey for White/Caucasian Individuals by Health Region"
+#| message: false
+#| echo: false
+
+clean_data %>% subset (Race=="white_caucasian")%>%
+  tbl_summary(
+    percent="row",
+    by= first_dose,
+    include=c(Health_Region),
+    sort = all_categorical() ~ "frequency"
+  ) %>% bold_labels()
+```
+
+```{r,West-HR-LHIN}
+#| label: tbl-west-hr-lhin
+#| tbl-cap: "Proportion of answers by LHIN (West Health Region)"
+#| message: false
+#| echo: false
+
+clean_data %>% subset (Health_Region=="West")%>%
+tbl_summary(include=c(LHIN),
+            sort = all_categorical() ~ "frequency") %>% 
+  bold_labels()
+
+clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
+tbl_summary(include=c(Race),
+            sort = all_categorical() ~ "frequency") %>% 
+  bold_labels()
+
+```

### 2023-03-01 code/data_preparation.R

@@ -69,16 +69,3 @@ clean_data<- within(clean_data, Age_group_ord <- relevel(Age_group_ord, ref = "1
 clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
 clean_data<- within(clean_data, income_ord <- relevel(income_ord, ref = "60000_and_above"))
-
-
-#create summary table
-
-data_summary<-clean_data %>%
-  tbl_summary(percent="row",
-              by= first_dose,
-              include=c(income_ord,
-                        Age_group_ord,
-                        Health_Region,
-                        Race)
-  ) %>% bold_labels()
-

### 2023-03-02 manuscript/01-background.qmd
Abstract, discussion, conclusion

@@
 
-In this study we examined self-reported COVID-19 vaccination status within the province of Ontario using a combination of socio-economic factors(such as ethnic origin, age, and income) and geographical analysis (at the level of the Health Regions)  in order to determine intra-provincial disparities in vaccine uptake and address the ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].
+In this study, we examined self-reported COVID-19 vaccination status within the province of Ontario using a combination of socio-economic factors(such as ethnic origin, age, and income) and geographical analysis (at the level of the Health Regions) in order to determine intra-provincial differences in vaccine uptake and address the ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].

### 2023-03-02 manuscript/02-methods.qmd
Abstract, discussion, conclusion

@@
 In the case of municipalities that did not appear in the long-term dataset, we manually searched each city in the LHINs websites in order to provide geographical information.
-The raw dataset, clean dataset, details of the data cleaning process, and the addition of Health Region and LHIN information are described in detail in the github repository: https://github.com/aimundo/Fields_COVID-19/.
+The raw dataset, clean dataset, details of the data cleaning process, and the addition of Health Region and LHIN information are described in detail in the GitHub repository for this paper, which can be found at [https://github.com/aimundo/Fields_COVID-19/](https://github.com/aimundo/Fields_COVID-19/).
 
@@
 
-Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term.
+Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a COVID-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term.
 The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.

### 2023-03-02 manuscript/03-results.qmd
Abstract, discussion, conclusion

@@
 
-@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed.
+@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed.
 The total number of entries from the survey in the dataset after cleaning was 3,551.
@@
 #| label: tbl-descriptive-stats
-#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey (by Vaccination Status)"
+#| tbl-cap: "Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination Status)"
 #| message: false
@@
                         Health_Region,
-                        Race)
+                        Race),
+              label=list(Health_Region ~ "Health Region",
+                 Age_group_ord ~ "Age Group",
+                 income_ord="Income")
   ) %>% bold_labels()
@@
 Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, respectively).
-Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.28, respectively); additionally, those individuals in the Other Race/Ethnicity group (a group that included Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere) had even lower odds of vaccination than the other minority groups, with an OR of 0.22.
-Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.55).
+Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.28, and _p_=0.004,<0.001 and 0.004, respectively); additionally, those individuals in the Other Race/Ethnicity group (a group that included Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere) had even lower odds of vaccination than the other minority groups (OR=0.22, _p_<0.001).
+Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.55, _p_=0.029).
 
-Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05), Black (OR=3.19), Latin American (OR=2.80), or that belonged to other minority groups  (OR=4.59).
-Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.93).
+Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001).
+Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.93, _p_<0.001).
 
-For the interaction of Health Region and race, significant odds of vaccination were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41).
+For the interaction of Health Region and race, significant odds of vaccination were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44, _p_=0.046), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41, _p_=0.032 and _p_=0.037, respectively).
 
@@
                  label=list(Health_Region ~ "Health Region",
-                 Age_group_ord ~ "Age Group")) %>%
+                 Age_group_ord ~ "Age Group",
+                 income_ord="Income")) %>%
     bold_labels()%>%

### 2023-03-02 manuscript/04-discussion.qmd
Abstract, discussion, conclusion

@@
 
-The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022], and their availability towards the end of 2020 in certain countries made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021].
-However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of Covid-19 have faced multiple challenges that have complicated the achievement of global immunity.
+The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022].
+Vaccine availability towards the end of 2020 in certain countries made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021].
+However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of COVID-19 have faced multiple challenges that have complicated the achievement of global immunity.
 
-Among the different challenges faced by Covid-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and inequity in vaccine access between low and high income countries [@tamey2022].
+Among the different challenges faced by COVID-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and inequity in vaccine access between low and high income countries [@tamey2022].
 However, it is also well established that even in the case of high income countries that have had ample access to vaccines since 2020, such as the US, the UK, and Canada, there have been challenges in vaccination efforts due to differences in vaccine uptake among different segments of the population.
@@
 
-In addition, vaccine uptake is influenced by geography as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021].
+In addition, vaccine uptake is influenced by geography, as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021].
 However, in the case of Canada, studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022].
@@
 
-Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the Covid-19 vaccine (@tbl-descriptive-stats), and that there were no significant differences in vaccination odds among the age groups considered in the survey.
+Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats), and that there were no significant differences in vaccination odds among the age groups considered in the survey.
 This result is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022], and with the vaccination information provided by Public Health Ontario,  which shows that for the period where the Fields survey ran (Sept 30, 2021-Jan 17, 2022) there was a minimum of  80% of first dose vaccination coverage among all the age groups considered in the survey [@ontario-covid].
@@
 
-This result is likely due to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include occupations such as grocery store workers, gas station workers, warehouse and distribution workers, and manufacturing workers, all being occupations for which a low income is to be expected.
-In the case of Ontario, essential workers had priority for Covid-19 vaccination [@mishra2021], which would explain the higher odds of vaccination for these individuals in certain income brackets, in contrast to the lower odds of vaccination for the same type of individuals in higher income brackets.
-In other words, it is possible that the type of occupation played an important role in the odds of vaccination in these racial minority groups.
+This result is likely due to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include occupations such as grocery store workers, gas station workers, warehouse and distribution workers, and manufacturing workers, all being occupations for which an income within the significant brackets is to be expected.
+In the case of Ontario, essential workers had priority for COVID-19 vaccination [@mishra2021], which would explain the higher odds of vaccination for these individuals in certain income brackets, in contrast to the lower odds of vaccination for the same type of individuals with higher household income.
+In other words, it is possible that the type of occupation played an important role in increasing the odds of vaccination in these racial minority groups.
 
@@
 The West Health Region comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
-In this case, overall higher odds of vaccination can be explained by the fact that in the survey, about 47% of the entries for this Health Region corresponded to White/Caucasian individuals, who reported an overall 83% vaccination rate (Supplementary Table A-6).
+In this case, a possible rationale for the results is the fact that in the survey, about 47% of the entries for this Health Region corresponded to White/Caucasian individuals, who reported an overall 83% vaccination rate (Supplementary Table A-6).
 However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@tbl-model].
@@
 
-The West Health Region, according to Ontario Health, has 13.2% of its population identifying as visible minority, and 2.5% of it identifying as Indigenous [@ontariohealth].
-In the case of this analysis, it is possible that these significant interactions are due to the fact that 50% of the answers from this region came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
-Clair, which were previously reported to have more than 20% of their population in the lowest income quintile [@buajitti2018] (Supplementary Table A-7) while about 55% of the answers from areas previously under the aforementioned LHINs also corresponded to individuals that identified in the minority groups.
-This would mean that although all minority groups had lower odds of vaccination in the rest of the Health Regions, the combination of higher representation of individuals from low-income areas that identified as minorities resulted in the overall significant lower odds of vaccination.
+According to Ontario Health, 13.2% of the population in the West Health Region identifies as a visible minority, whereas 2.5% identifies as Indigenous [@ontariohealth].
+In the case of this analysis, the estimated lower odds are likely to be explained from a socio-economic perspective.
+In fact, 50% of the answers from this region in the survey came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
+Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-7).
+Therefore, this result partly reinforces the well-known existing association between low vaccination rates and income, but it additionally indicates that there were intra-regional differences in vaccination.
+Interestingly, a disproportionate number of COVID-19 cases and low vaccination rate (under 50%) have been previously reported in the South Asian community of Ontario [@anand2022]; in this regard, our result provides additional context  by showing that within the South Asian community, there were differences in vaccination uptake across Ontario.
+Moreover, because significant lower odds of vaccination were also identified  other minority groups, this provides a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.
 
-Although there was no significant difference between vaccination odds among the Health Regions, there was a significant lower vaccination odds in the Central Health Region in individuals with a household income between CAD 15,000-24,999 (OR=0.4, @tbl-model).
-The Central Health Region comprises the former Central, Central West, Mississauga Halton, and North Simcoe Muskoka LHINs with 48.3% of immigrant population and 47% of the population identifying as visible minority, the highest percentage among all Health Regions [@ontariohealth]; moreover, the fact that the Central West LHIN, now within the Central Health Region, had the highest proportion of low-income families in 2018 (12.9\% of males and 16.7\% of females among the lowest 20% of Canadian family incomes [@buajitti2018])
+There are some limitations to the present study.
+First, the data collection design, which allowed respondents to withdraw from the survey at any point, resulted in a high number of unique entries in the survey with multiple missing answers.
+Because we focused on entries that had complete observations in the covariates of interest for our analysis, it is possible that some information was not considered by excluding observations that had information in other variables (such as work from home, or number of persons in the household).
+However, we attempted to minimize this possibility by correcting the dataset using information from the Census.
+More granular corrections, which for example could be based on demographic information by municipality, could be used in the future to obtain a more accurate approximation to the population totals of the province.
+Additionally, the results in this study are based on self-reported data, where the risk of bias exist.
+Despite this, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], the effect of self-reported bias is likely to not be significant in this case.
 
-- Influence of geography
-
-- ontario and Health Regions: find information about poverty and racial groups.
-
-- Results
-
-- Limitations
+Finally, it is likely that there have been differences in vaccination across the province as more doses of the vaccine were administered and as successive variants emerged.
+Because this study focused only on vaccination status regarding the first dose of the vaccine within a relatively short time window, it can only provide a snapshot of the societal dynamics behind the pandemic.
+Nonetheless, the results presented here can serve as a starting point to motivate future longitudinal research that aims to quantify geographical differences within vulnerable segments of the population, and that can be used to inform the development of public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.

### 2023-03-02 manuscript/appendix.qmd
Abstract, discussion, conclusion

@@
 #| fig-width: 6
-#| fig-cap: Interaction plot between Race and Income
+#| fig-cap: Estimated marginal means for the combination between Race and Income in the regression model.
 
@@
 #| fig-width: 6
-#| fig-cap: Interaction plot between Race and Health Region
+#| fig-cap: Estimated marginal means for the combination between Race and Health Region in the regression model.
 
@@
 
-clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
-tbl_summary(include=c(Race),
-            sort = all_categorical() ~ "frequency") %>% 
-  bold_labels()
-
+# clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
+# tbl_summary(include=c(Race),
+#             sort = all_categorical() ~ "frequency") %>% 
+#   bold_labels()
+# 
+# 
+# clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
+# tbl_summary(by=first_dose,
+#             percent="row",
+#   include=c(Race),
+#             sort = all_categorical() ~ "frequency") %>% 
+#   bold_labels()
+# 
+# clean_data %>% subset (LHIN=="Erie St. Clair")%>%
+# tbl_summary(by=first_dose,
+#             percent="row",
+#   include=c(Race),
+#             sort = all_categorical() ~ "frequency") %>% 
+#   bold_labels()
 ```

### 2023-03-02 manuscript/main.qmd
Abstract, discussion, conclusion

@@
 
+The COVID-19 pandemic continues to be a worldwide public health concern.
+Although vaccines against this disease were rapidly developed, vaccination uptake has not ben equal across all the segments of the population.
+In particular, it has been shown that there have been differences in vaccine uptake across different segments of the population.
+However, there are also differences in vaccination across geographical areas, which might be important to consider in the development of future public health policies against COVID-19.
+In this study, we examined the relationship between vaccination status (having received the first dose of a COVID-19 vaccine), and different socio-economic and geographical factors.
+Our results show differences in vaccination due to race/ethnicity, income, Health Regions (geographical areas used for health service access in Ontario), and their interactions.
+In particular, we show
+that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32, 0.28, and _p_=0.004,_p_<0.001 and _p_=0.004, respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets.
+Finally, we also identified lower odds of vaccination within different minority groups in West Health Region.
+This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population that have been largely impacted by the pandemic.
+
 # Keywords {-}
+
+Covid-19, vaccination, survey, socio-economic factors, visible minorities.
 
@@
 
+```{r}
+#| child: 05-conclusion.qmd
+```
+
 # References

### 2023-03-03 manuscript/appendix.qmd
updated author info, font

@@
 ---
-title: '**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**'
+title: '**COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata**'
 subtitle: Appendix
-author: Ariel Mundo Ortiz
+authors:
+  - name: Ariel Mundo Ortiz
+    affiliations:
+      - ref: crm
+    email: ariel.mundo.ortiz@umontreal.ca
+    orcid: 0000-0002-6014-4538
+  - name: Bouchra Nasri
+    affiliations:
+      - ref: espum
+    email: bouchra.nasri@umontreal.ca
+    corresponding: true
+affiliations:
+  - id: crm
+    name: Centre de Recherches Mathématiques, University of Montreal, Montréal, Canada
+  - id: espum
+    name: Department of Social and Preventive Medicine, École de Santé Publique, University of Montreal, Montréal, Canada 
+filters:
+  - authors-block
 format: 
   pdf:
+    mainfont: "Palatino Linotype"
     include-before-body: latex/appendix_preamble.sty
-#bibliography: refs.bib
-#csl: jama.csl
-#link-citations: yes #adds links to the citations
+linestretch: 2
 ---

### 2023-03-03 manuscript/main.qmd
updated author info, font

@@
 ---
-title: '**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**'
-author: Ariel Mundo Ortiz
-format: pdf
+title: '**COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata**'
+authors:
+  - name: Ariel Mundo Ortiz
+    affiliations:
+      - ref: crm
+    email: ariel.mundo.ortiz@umontreal.ca
+    orcid: 0000-0002-6014-4538
+  - name: Bouchra Nasri
+    affiliations:
+      - ref: espum
+    email: bouchra.nasri@umontreal.ca
+    corresponding: true
+affiliations:
+  - id: crm
+    name: Centre de Recherches Mathématiques, University of Montreal, Montréal, Canada
+  - id: espum
+    name: Department of Social and Preventive Medicine, École de Santé Publique, University of Montreal, Montréal, Canada 
+filters:
+  - authors-block
+format: 
+  pdf:
+    include-in-header: latex/preamble.sty
+    mainfont: "Palatino Linotype"
+    keep-tex: true
 linestretch: 2
-bibliography: refs.bib
-csl: jama.csl
+bibliography: references/refs.bib
+csl: references/jama.csl
 link-citations: yes #adds links to the citations

### 2023-03-03 manuscript/main.tex
updated author info, font

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmainfont[]{Palatino Linotype}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

%line numbers
%\usepackage{mathpazo}
\usepackage{lineno}
\linenumbers

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata},
  pdfauthor={Ariel Mundo Ortiz1; Bouchra Nasri2,},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{COVID-19 vaccination in Ontario: Exploring
intra-provincial variations within Health Regions and socio-economic
strata}}
\author{Ariel Mundo Ortiz\textsuperscript{1} \and Bouchra
Nasri\textsuperscript{2,*}}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

\setstretch{2}
\textsuperscript{1} Centre de Recherches Mathématiques, University of
Montreal, Montréal, Canada\\
\textsuperscript{2} Department of Social and Preventive Medicine, École
de Santé Publique, University of Montreal, Montréal, Canada

\textsuperscript{*} Correspondence:
\href{mailto:bouchra.nasri@umontreal.ca}{Bouchra Nasri
\textless{}bouchra.nasri@umontreal.ca\textgreater{}}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

The COVID-19 pandemic continues to be a worldwide public health concern.
Although vaccines against this disease were rapidly developed,
vaccination uptake has not ben equal across all the segments of the
population. In particular, it has been shown that there have been
differences in vaccine uptake across different segments of the
population. However, there are also differences in vaccination across
geographical areas, which might be important to consider in the
development of future public health policies against COVID-19. In this
study, we examined the relationship between vaccination status (having
received the first dose of a COVID-19 vaccine), and different
socio-economic and geographical factors. Our results show differences in
vaccination due to race/ethnicity, income, Health Regions (geographical
areas used for health service access in Ontario), and their
interactions. In particular, we show that individuals who identified as
Arab/Middle Eastern, Black, or Latin American, had significantly lower
odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32,
0.28, and \emph{p}=0.004,\emph{p}\textless0.001 and \emph{p}=0.004,
respectively), and that individuals with a household income below CAD
25,000 who identified as Arab/Middle Eastern (OR=3.05, \emph{p}=0.013),
Black (OR=3.19, \emph{p}=0.004), Latin American (OR=2.80,
\emph{p}=0.041), or that belonged to other minority groups (OR=4.59,
\emph{p}\textless0.001) had higher odds of vaccination than individuals
from the same racial/ethnic group in higher income brackets. Finally, we
also identified lower odds of vaccination within different minority
groups in West Health Region. This study shows that there is an ongoing
need to better understand and address differences in vaccination uptake
across diverse segments of the population that have been largely
impacted by the pandemic.

\hypertarget{keywords}{%
\section*{Keywords}\label{keywords}}
\addcontentsline{toc}{section}{Keywords}

Covid-19, vaccination, survey, socio-economic factors, visible
minorities.

\hypertarget{background}{%
\section{Background}\label{background}}

The COVID-19 pandemic continues around the world with more than 600
million confirmed cases as of November of
2022\textsuperscript{\protect\hyperlink{ref-WHO-Covid}{1}}. During the
first months of the pandemic in early 2020, non-pharmaceutical
interventions such as masking and social distancing were the only
methods available to manage the spread of the disease; however, the
rapid development of vaccines permitted their approval and use in some
countries towards the last month part of 2020. For example, in the US
and Canada vaccine campaigns began in mid-December of
2020\textsuperscript{\protect\hyperlink{ref-tanne2020}{2},\protect\hyperlink{ref-bogoch2022}{3}}.
Although it has been estimated that COVID-19 vaccines have prevented
around 14 millions of deaths
worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{4}}, the
rollout of COVID-19 vaccines has faced multiple challenges since its
inception.

Indeed, multiple obstacles that have complicated vaccination efforts
against COVID-19 have been identified: inequality in vaccine access,
vaccine hesitancy, and differences in vaccination rates across different
segments of the
population\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{5}--\protect\hyperlink{ref-malik2020}{7}}.
In the case of Canada, lower vaccine uptake has been associated with
socio-economic factors such as younger age, educational level, presence
of children in the household, lack of a regular healthcare provider,
ethnic origin, and financial
instability\textsuperscript{\protect\hyperlink{ref-guay2022}{8}--\protect\hyperlink{ref-hussain2022}{10}}.

Additionally, vaccination is influenced by changes in geography. In this
regard, it has been shown that there have been spatial differences in
COVID-19 vaccination rates due to regional differences in attitudes
towards
vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{7}},
geographical differences in vaccine access and supply, vaccination
location availability, and lack of prioritization of vulnerable
groups\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3},\protect\hyperlink{ref-nguyen2021}{11}}.

Studies that analyze geographical variations in vaccine uptake are
important as they can help inform public health decision-makers to
design policies to that consider spatial changes to address vaccination
disparities. In this regard, previous geographical (spatial) analyses of
vaccination rates have shown that variations in vaccine uptake can occur
within small governmental administrative units (e.g., counties in the
case of the
US)\textsuperscript{\protect\hyperlink{ref-mollalo2021}{12}--\protect\hyperlink{ref-bhuiyan2022}{15}},
and that geographical analyses can be predictive of booster uptake
patterns\textsuperscript{\protect\hyperlink{ref-wood2022}{16}}.

In Canada, studies that have used a spatial approach to analyze vaccine
uptake have shown disparities in vaccination rates across low and high
income neighborhoods in the city of
Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{17}}, among
adolescents from deprived neighborhoods in the city of
Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{18}}, and
have also highlighted disparities in vaccination status depending on
age, income, and ethnic origin at the provincial
level\textsuperscript{\protect\hyperlink{ref-guay2022}{8}}. However,
there is limited information on differences in vaccination status inside
the provinces. Such analysis is important as it can help identify
inequalities that may exist within these geographical areas while
providing a granular view of intra-provincial differences that can help
understand the barriers for vaccine uptake.

In this study, we examined self-reported COVID-19 vaccination status
within the province of Ontario using a combination of socio-economic
factors(such as ethnic origin, age, and income) and geographical
analysis (at the level of the Health Regions) in order to determine
intra-provincial differences in vaccine uptake and address the ongoing
need of socio-economic information that can provide a rationale for the
disparities in vaccination observed within some racial
groups\textsuperscript{\protect\hyperlink{ref-cnat2022b}{19}}.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-source-survey-overview}{%
\subsection{Data source: survey
overview}\label{data-source-survey-overview}}

We used data from the \emph{Survey of COVID-19 related Behaviours and
Attitudes}, a repeated cross sectional survey focused on the Canadian
province of Ontario which ran from Sept 30, 2021 until January 17, 2022
and that was comissioned by the Fields Institute for Research in
Mathematical Sciences (henceforth Fields) and the Mathematical Modelling
of COVID-19 Task Force. The survey was conducted by a third-party
service provider (RIWI Corp.), under ethical guidance from the
University of Toronto.

Briefly, the survey was deployed used random domain intercept
technology, where if users clicked on a registered but commercially
inactive web link or typed in a web address for a site that was dormant,
they had a random chance of that link being temporarily managed by the
company that administered the survey and instead of coming across a
notification about the status of the site(``this page does not exist''),
the survey was deployed to the
user\textsuperscript{\protect\hyperlink{ref-sargent2022}{20}}. Users
then decided whether to anonymously participate, and those that
participated were able to exit the survey at any time. After the survey
closed, regardless if it was complete or incomplete, access was denied
to any further users with the same internet protocol (IP) address and
the domain entry point rotated such that if a user were to attempt to
access the survey again, share the link, or enter via the same address
using an alternative IP address, the survey would not deploy. This
effectively meant that a user could participate only once in the survey.

Additionally, users who indicated they were under the age of 16 were
exited from the survey without creating a record, furthermore, these
users were unable to navigate back to the ``age select'' screen. The
personal identifier information from each user that participated in the
survey was automatically scrubbed and replaced by a unique ID.
Respondents were drawn exclusively from the province of Ontario, as per
their devices meta-data.

\hypertarget{sec-socio-demographic-factors}{%
\subsection{Survey responses}\label{sec-socio-demographic-factors}}

Socio-economic information selected from the survey answers included age
group and income brackets, and race/ethnicity. The levels of each
socio-economic factor used for analysis appear in
Table~\ref{tbl-covariates}.

\hypertarget{tbl-covariates}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3014}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6986}}@{}}
\caption{\label{tbl-covariates}Selected socio-economic factors from the
survey}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Levels
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Levels
\end{minipage} \\
\midrule()
\endhead
Age group & 16-34,35-54,55 and over \\
Income bracket (CAD) & under 25,000, 25,000-59,999, 60,000 and above \\
Race/ethnicity & Arab/Middle Eastern, Black, East Asian/Pacific
Islander, Indigenous, Latin American, Mixed, South Asian, White
Caucasian, Other \\
\bottomrule()
\end{longtable}

Furthermore, the information on vaccination status was provided by
survey participants who answered the question ``Have you received the
first dose of the COVID vaccine?'', with possible answers ``yes'' and
``no''.

\hypertarget{data-cleaning}{%
\subsection{Data cleaning}\label{data-cleaning}}

The original dataset contained 39,029 entries (where each entry
corresponded to a set of answers provided by a unique respondent).
Following a preliminary analysis, the dataset was cleaned in order to
only contain the socio-economic information provided in
Table~\ref{tbl-covariates} and vaccination status. The cleaning process
also included removing outliers that were identified during the
preliminary analyses, and processing the geographical information in the
survey (city where the survey was responded) in order to match each city
to its correspondent Health Region.

\hypertarget{sec-geographical-location}{%
\subsubsection{Geographical location}\label{sec-geographical-location}}

For each survey participant certain data was automatically captured,
including the nearest municipality (city). This resulted in a total of
578 different municipalities within the clean dataset. Due to our
interest in analyzing the differences between Health Regions, we
assigned the city of each entry to its correspondent Health Region
following a multi-step process. Briefly, we used Local Health Integrated
Networks (LHINs) to assign a Health Region to each entry in the survey.
LHINs were the geographical divisions for health used by Ontario before
the adoption of the Health Regions; because of the lack of a publicly
available list of all municipalities within each Health Region, we used
a dataset of long-term care homes and LHINs to match each city to LHIN,
followed by matching each LHIN to a Health Region following the
information provided on the Ontario Health Website, where the list of
LHINs and corresponding Health Regions is available. In the case of
municipalities that did not appear in the long-term dataset, we manually
searched each city in the LHINs websites in order to provide
geographical information. The raw dataset, clean dataset, details of the
data cleaning process, and the addition of Health Region and LHIN
information are described in detail in the GitHub repository for this
paper, which can be found at
\url{https://github.com/aimundo/Fields_COVID-19/}.

Following an assessment of the number of entries corresponding to each
Health Region in the final dataset, only 107 observations (4.3\% of the
total) corresponded to cities located in the North West and North East
Health Regions. Due to the low number of entries, we omitted these
Health Regions from further analyses. Therefore, the total number of
unique entries used for analysis was 3,551 which included the East,
Central, Toronto, and West Health Regions.

\hypertarget{corrections}{%
\subsection{Corrections}\label{corrections}}

We identified differences between the proportions of all the
socio-economic factors included in the analysis
(Table~\ref{tbl-covariates}) and the 2016 Canada Census data for
Ontario. Additionally, because the Census divisions do not match the
exact boundaries of the Health Regions, we also obtained population
estimates for each Health Region from the Ontario Health website in
order to correct for the population totals. We used an iterative
proportional fitting procedure
(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{21}}
to correct for socio-economic factors and Health Region populations
using the R \texttt{survey} package. Details about the correction can be
found in the Appendix.

\hypertarget{statistical-analyses}{%
\subsection{Statistical analyses}\label{statistical-analyses}}

We used a logistic regression model to estimate the probability of
vaccination depending on the socio-economic factors described in
Section~\ref{sec-socio-demographic-factors}, the Health Regions from
Section~\ref{sec-geographical-location}, and the interactions between
Race and Health Region, and Race and income as previous studies have
shown that socio-economic factors and their interactions are significant
predictors of intent of vaccination and vaccination
status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{22}--\protect\hyperlink{ref-cnat2022a}{24}}.
The model appears in Equation~\ref{eq-model3},

\begin{equation}\protect\hypertarget{eq-model3}{}{
\begin{aligned}
\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Race)} + \beta_6 \textrm{ (Income} \times \textrm{Race)}
\end{aligned}
}\label{eq-model3}\end{equation}

Where \(p\textrm{(vac)}\) indicates the probability of having received
the first dose of a COVID-19 vaccine, \(\beta_0\) indicates the
population intercept, and \(\beta_1...\beta_6\) indicate the
coefficients for each term. The the model was fitted using the function
\texttt{svyglm} from the \texttt{survey} R package in order to
incorporate the correction in sampling probability obtained from raking.

All analyses were conducted in RStudio (2022.12.0 Build 353), using R
4.2.2 and the packages
\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{25}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{26}},
and \texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{27}}.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{survey-results}{%
\subsection{Survey Results}\label{survey-results}}

Table~\ref{tbl-descriptive-stats} shows the descriptive statistics
(uncorrected) from the Fields COVID-19 survey data for vaccination
status and each of the covariates analyzed. The total number of entries
from the survey in the dataset after cleaning was 3,551. Overall, 26.9\%
of survey respondents (958) reported not having received the first dose
of the vaccine, whereas 73.1\% (2,593) reported having received it.
Within each socio-economic factor, respondents who reported living in a
household with an income under CAD 25,000 represented 37\% of the total
number of entries, those within the CAD 25,000-59,999 income bracket
represented 25\% of the total sample, and those with an income above CAD
60,000 represented 38 \% of the sample; across all income brackets, the
percentage of individuals that reported having received a first dose of
the vaccine was consistent, above 69\%.

Within the age groups of survey respondents, the age group between 16-34
years had the highest representation in the survey responses (1,521,
42.8\% of all responses). Within this age bracket, 73\% of respondents
indicated having received the vaccine, whereas the lowest vaccination
rate was in the bracket of those 55 years of age and above, with a total
of 72\%. The Health Region with highest representation in the survey was
Toronto, accounting for 1,324 entries (37.2\%), with a vaccination rate
of 72\%. Regarding race/ethnicity, individuals that identified as
White/Caucasian represented 1313 (37\%) of all entries and had the
highest vaccination uptake with 82\% of them indicating to have received
the COVID-19 vaccine. On the other hand, the ethnic group with the
lowest number of entries in the survey was Latin American, with a total
of 180, or 5\% of all entries. Vaccination rates across all minority
groups were below the value reported by White/Caucasians, with the
lowest vaccination rate (60\%) being reported by individuals that
identified as Indigenous.

\hypertarget{tbl-descriptive-stats}{}
\begin{table}
\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination
Status) }\tabularnewline

\centering\begingroup\fontsize{9}{11}\selectfont

\begin{tabular}{lcc}
\toprule
\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,593\\
\midrule
\textbf{Income} &  & \\
\hspace{1em}60000\_and\_above & 305 (23\%) & 1,049 (77\%)\\
\hspace{1em}25000\_59999 & 253 (28\%) & 636 (72\%)\\
\hspace{1em}under\_25000 & 400 (31\%) & 908 (69\%)\\
\textbf{Age Group} &  & \\
\hspace{1em}16\_34 & 409 (27\%) & 1,112 (73\%)\\
\hspace{1em}35\_54 & 252 (26\%) & 712 (74\%)\\
\hspace{1em}55\_and\_over & 297 (28\%) & 769 (72\%)\\
\textbf{Health Region} &  & \\
\hspace{1em}Toronto & 371 (28\%) & 953 (72\%)\\
\hspace{1em}Central & 224 (28\%) & 581 (72\%)\\
\hspace{1em}East & 135 (23\%) & 448 (77\%)\\
\hspace{1em}West & 228 (27\%) & 611 (73\%)\\
\textbf{Race} &  & \\
\hspace{1em}white\_caucasian & 233 (18\%) & 1,080 (82\%)\\
\hspace{1em}arab\_middle\_eastern & 76 (36\%) & 138 (64\%)\\
\hspace{1em}black & 114 (38\%) & 184 (62\%)\\
\hspace{1em}east\_asian\_pacific\_islander & 69 (23\%) & 234 (77\%)\\
\hspace{1em}indigenous & 76 (40\%) & 115 (60\%)\\
\hspace{1em}latin\_american & 69 (38\%) & 111 (62\%)\\
\hspace{1em}mixed & 105 (34\%) & 205 (66\%)\\
\hspace{1em}other & 128 (35\%) & 239 (65\%)\\
\hspace{1em}south\_asian & 88 (23\%) & 287 (77\%)\\
\bottomrule
\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
\end{tabular}
\endgroup{}
\end{table}

\hypertarget{multivariate-regression}{%
\subsection{Multivariate Regression}\label{multivariate-regression}}

Table~\ref{tbl-model} shows the results of the logistic regression on
vaccination status using socio-economic factors (age group, income,
race), geographical areas (Health Regions) and the interactions between
income and race and Health Region and race. The reference groups were
set as follows: 16 to 34 years (age group), White Caucasian (Race),
Toronto (Health Region), CAD 60,000 and over (Income). There were no
statistically significant differences in vaccination rates within the
age groups from the survey, but significant odds ratios were estimated
for other covariates. Within household income brackets, individuals with
an income under CAD 25,000 or between CAD 25,000-59,999 had
significantly lower odds of vaccination than those with an income above
CAD 60,000 (ORs=0.37 and 0.59, respectively). Within Race/Ethnicity,
individuals who identified as Arab/Middle Eastern, Black, or Latin
American, had significantly lower odds of vaccination than those in the
White/Caucasian group (ORs=0.31, 0.32, 0.28, and
\emph{p}=0.004,\textless0.001 and 0.004, respectively); additionally,
those individuals in the Other Race/Ethnicity group (a group that
included Southeast Asian, Filipino, West Asian, and Minorities Not
Identified Elsewhere) had even lower odds of vaccination than the other
minority groups (OR=0.22, \emph{p}\textless0.001). Regarding Health
Regions, individuals that reported living in the West Health Region
(which comprises the regions of Waterloo and Niagara, the counties of
Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
Brant, and Chatham-Kent) had significantly higher odds of vaccination
than those in the Health Region of Toronto (OR=1.55, \emph{p}=0.029).

Moreover, statistically-significant odd ratios were determined in the
case of the interaction of income and race; specifically, for
individuals with a household income below CAD 25,000 who identified as
Arab/Middle Eastern (OR=3.05, \emph{p}=0.013), Black (OR=3.19,
\emph{p}=0.004), Latin American (OR=2.80, \emph{p}=0.041), or that
belonged to other minority groups (OR=4.59, \emph{p}\textless0.001).
Within the CAD 25,000-59,999 income bracket, individuals who identified
as belonging to other racial minority groups had significantly higher
odds of vaccination (OR=6.93, \emph{p}\textless0.001).

For the interaction of Health Region and race, significant odds of
vaccination were identified for Black individuals in the Central Health
Region, which comprises the region of York, counties of Dufferin and
Simcoe and the district of Muskoka (OR=0.44, \emph{p}=0.046), and in
individuals that identified as part of other racial minorities or South
Asian that lived in the West Health Region (ORs=0.41, \emph{p}=0.032 and
\emph{p}=0.037, respectively).

\footnotesize
\renewcommand{\arraystretch}{0.5}

\hypertarget{tbl-model}{}
\begin{longtable}{lccc}
\caption{\label{tbl-model}Multiple Regression Analysis-Predictors of Vaccination Status }\tabularnewline

\toprule
\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\textbf{Age Group} &  &  & \\
\hspace{1em}16\_34 & — & — & \\
\hspace{1em}35\_54 & 0.90 & 0.67, 1.21 & 0.5\\
\hspace{1em}55\_and\_over & 0.99 & 0.74, 1.32 & >0.9\\
\textbf{Income} &  &  & \\
\hspace{1em}60000\_and\_above & — & — & \\
\hspace{1em}25000\_59999 & 0.59 & 0.39, 0.89 & 0.011\\
\hspace{1em}under\_25000 & 0.37 & 0.25, 0.56 & <0.001\\
\textbf{Race} &  &  & \\
\hspace{1em}white\_caucasian & — & — & \\
\hspace{1em}arab\_middle\_eastern & 0.31 & 0.14, 0.69 & 0.004\\
\hspace{1em}black & 0.32 & 0.17, 0.60 & <0.001\\
\hspace{1em}east\_asian\_pacific\_islander & 1.15 & 0.50, 2.66 & 0.7\\
\hspace{1em}indigenous & 0.44 & 0.19, 1.02 & 0.056\\
\hspace{1em}latin\_american & 0.28 & 0.11, 0.67 & 0.004\\
\hspace{1em}mixed & 0.64 & 0.25, 1.65 & 0.4\\
\hspace{1em}other & 0.22 & 0.12, 0.41 & <0.001\\
\hspace{1em}south\_asian & 0.91 & 0.49, 1.69 & 0.8\\
\textbf{Health Region} &  &  & \\
\hspace{1em}Toronto & — & — & \\
\hspace{1em}Central & 1.47 & 0.92, 2.35 & 0.11\\
\hspace{1em}East & 1.42 & 0.90, 2.23 & 0.13\\
\hspace{1em}West & 1.55 & 1.05, 2.30 & 0.029\\
\textbf{Income * Race} &  &  & \\
\hspace{1em}25000\_59999 * arab\_middle\_eastern & 1.79 & 0.67, 4.83 & 0.2\\
\hspace{1em}under\_25000 * arab\_middle\_eastern & 3.05 & 1.26, 7.39 & 0.013\\
\hspace{1em}25000\_59999 * black & 1.34 & 0.59, 3.05 & 0.5\\
\hspace{1em}under\_25000 * black & 3.19 & 1.45, 6.99 & 0.004\\
\hspace{1em}25000\_59999 * east\_asian\_pacific\_islander & 0.42 & 0.17, 1.05 & 0.062\\
\hspace{1em}under\_25000 * east\_asian\_pacific\_islander & 1.16 & 0.47, 2.86 & 0.8\\
\hspace{1em}25000\_59999 * indigenous & 1.36 & 0.48, 3.89 & 0.6\\
\hspace{1em}under\_25000 * indigenous & 1.45 & 0.55, 3.80 & 0.5\\
\hspace{1em}25000\_59999 * latin\_american & 1.24 & 0.45, 3.43 & 0.7\\
\hspace{1em}under\_25000 * latin\_american & 2.80 & 1.04, 7.51 & 0.041\\
\hspace{1em}25000\_59999 * mixed & 0.85 & 0.32, 2.26 & 0.7\\
\hspace{1em}under\_25000 * mixed & 1.10 & 0.37, 3.27 & 0.9\\
\hspace{1em}25000\_59999 * other & 6.93 & 2.65, 18.1 & <0.001\\
\hspace{1em}under\_25000 * other & 4.59 & 2.33, 9.05 & <0.001\\
\hspace{1em}25000\_59999 * south\_asian & 1.20 & 0.51, 2.85 & 0.7\\
\hspace{1em}under\_25000 * south\_asian & 2.00 & 0.93, 4.30 & 0.077\\
\textbf{Race * Health Region} &  &  & \\
\hspace{1em}arab\_middle\_eastern * Central & 0.66 & 0.26, 1.70 & 0.4\\
\hspace{1em}black * Central & 0.44 & 0.19, 0.98 & 0.046\\
\hspace{1em}east\_asian\_pacific\_islander * Central & 0.98 & 0.38, 2.53 & >0.9\\
\hspace{1em}indigenous * Central & 0.63 & 0.22, 1.79 & 0.4\\
\hspace{1em}latin\_american * Central & 0.67 & 0.23, 1.96 & 0.5\\
\hspace{1em}mixed * Central & 0.73 & 0.24, 2.22 & 0.6\\
\hspace{1em}other * Central & 0.80 & 0.36, 1.78 & 0.6\\
\hspace{1em}south\_asian * Central & 0.54 & 0.25, 1.20 & 0.13\\
\hspace{1em}arab\_middle\_eastern * East & 0.43 & 0.13, 1.45 & 0.2\\
\hspace{1em}black * East & 0.83 & 0.34, 2.04 & 0.7\\
\hspace{1em}east\_asian\_pacific\_islander * East & 0.86 & 0.29, 2.56 & 0.8\\
\hspace{1em}indigenous * East & 0.69 & 0.23, 2.08 & 0.5\\
\hspace{1em}latin\_american * East & 1.03 & 0.32, 3.34 & >0.9\\
\hspace{1em}mixed * East & 0.91 & 0.28, 3.03 & 0.9\\
\hspace{1em}other * East & 1.05 & 0.39, 2.83 & >0.9\\
\hspace{1em}south\_asian * East & 0.52 & 0.19, 1.45 & 0.2\\
\hspace{1em}arab\_middle\_eastern * West & 1.00 & 0.37, 2.73 & >0.9\\
\hspace{1em}black * West & 0.76 & 0.32, 1.80 & 0.5\\
\hspace{1em}east\_asian\_pacific\_islander * West & 0.52 & 0.20, 1.34 & 0.2\\
\hspace{1em}indigenous * West & 0.39 & 0.14, 1.09 & 0.073\\
\hspace{1em}latin\_american * West & 0.94 & 0.32, 2.72 & >0.9\\
\hspace{1em}mixed * West & 0.37 & 0.12, 1.16 & 0.089\\
\hspace{1em}other * West & 0.41 & 0.18, 0.93 & 0.032\\
\hspace{1em}south\_asian * West & 0.41 & 0.18, 0.95 & 0.037\\*
\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\
\end{longtable}

\normalsize

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The rapid development of COVID-19 vaccines has been considered as a
major achievement of modern
medicine\textsuperscript{\protect\hyperlink{ref-davis2022}{28}}. Vaccine
availability towards the end of 2020 in certain countries made some
believe that they would be a determinant factor in a rapid ending of the
pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{29}}.
However, despite previous successful vaccination campaigns that were
crucial to control diseases such as smallpox and
polio\textsuperscript{\protect\hyperlink{ref-kayser2021}{30}},
vaccination efforts in the case of COVID-19 have faced multiple
challenges that have complicated the achievement of global immunity.

Among the different challenges faced by COVID-19 vaccination efforts are
the development of new variants due to inadequate public health
measures\textsuperscript{\protect\hyperlink{ref-li2021}{31}} and
inequity in vaccine access between low and high income
countries\textsuperscript{\protect\hyperlink{ref-tamey2022}{32}}.
However, it is also well established that even in the case of high
income countries that have had ample access to vaccines since 2020, such
as the US, the UK, and Canada, there have been challenges in vaccination
efforts due to differences in vaccine uptake among different segments of
the population. More specifically, lower vaccine uptake has been
associated with socio-economic factors such as race (i.e., identifying
as Black, Asian, Indigenous) and household income (typically within
lower income
brackets)\textsuperscript{\protect\hyperlink{ref-willis2021}{33}--\protect\hyperlink{ref-khubchandani2021}{36}}.
Reasons given for this association have included medical mistrust due to
systemic medical racism, mistrust in vaccines, and the influence of
conspiracy
theories\textsuperscript{\protect\hyperlink{ref-willis2021}{33},\protect\hyperlink{ref-stoler2021}{35},\protect\hyperlink{ref-bogart2021}{37}--\protect\hyperlink{ref-freeman2020}{39}}.

In addition, vaccine uptake is influenced by geography, as shown by
different studies that have identified intra-regional differences in
vaccine
uptake\textsuperscript{\protect\hyperlink{ref-mollalo2021}{12},\protect\hyperlink{ref-pallathadka2022}{40},\protect\hyperlink{ref-huang2022}{41}}.
However, in the case of Canada, studies that have analyzed spatial
differences in vaccination have been focused in country-wide or
province-wide
estimates\textsuperscript{\protect\hyperlink{ref-guay2022}{8},\protect\hyperlink{ref-lavoie2022}{42}}.
Therefore, we explored spatial and socio-economic determinants of
vaccination status in the province of Ontario. This province is of
particular interest as it has seen recently major structural health
changes with the dissolution of the Local Health Integrated Network
(LHIN) system and the incorporation of regions covered by LHINs into
larger Health
Regions\textsuperscript{\protect\hyperlink{ref-dong2022}{43}}. Because
the idea behind the change aimed to reduce the inequalities in
healthcare that were identified under the LHIN
model\textsuperscript{\protect\hyperlink{ref-tsasis2012}{44}}, examining
differences in vaccination between the Health Regions can provide
decision-makers with insight regarding intra-provincial health
disparities that may need to be addressed in future vaccination or
public health campaigns.

Our results indicate that across the most densely populated Health
Regions of Ontario, almost three quarters of the surveyed individuals
reported having received the first dose of the COVID-19 vaccine
(Table~\ref{tbl-descriptive-stats}), and that there were no significant
differences in vaccination odds among the age groups considered in the
survey. This result is consistent with overall vaccination rates
reported for Canada, which have been relatively higher when compared to
other high income
countries\textsuperscript{\protect\hyperlink{ref-dube2022}{45}}, with
vaccination uptake rates across different age groups presented in other
studies\textsuperscript{\protect\hyperlink{ref-guay2022}{8},\protect\hyperlink{ref-macdonald2021}{46}},
and with the vaccination information provided by Public Health Ontario,
which shows that for the period where the Fields survey ran (Sept 30,
2021-Jan 17, 2022) there was a minimum of 80\% of first dose vaccination
coverage among all the age groups considered in the
survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{47}}.

However, we identified intra-provincial differences in vaccination based
on socio-economic and geographical factors. First, our results show
significant differences in vaccination odds in individuals with a
household income below CAD 60,000 and in individuals belonging to
visible minority groups. Those who identified as Black, Latin American,
or belonging to a minority group not included in the survey (Southeast
Asian, Filipino, West Asian, and Minority not identified elsewhere) had
vaccination odds below 33\% when compared to individuals that identified
as White/Caucasian (Table~\ref{tbl-model}). These results are consistent
with other studies that have shown lower vaccination rates in
individuals that identify as part of a racial minority, or that have a
low household
income\textsuperscript{\protect\hyperlink{ref-guay2022}{8}--\protect\hyperlink{ref-hussain2022}{10},\protect\hyperlink{ref-carter2022}{48}}.

In this study, we also decided to explore the interactions between
income and race and race and Health Region, as it is known that many
individuals within racial minority groups perform tend to occupy certain
types of occupations that fall within income brackets that have been
shown to be associated with differences in vaccination uptake. In other
words, we decided to explore if there were differences in vaccination
within racial groups in certain income brackets and in certain the
Health Regions. In this regard, it is interesting to note that although
overall self-reported vaccination rates were found to be statistically
significantly lower in various racial minority groups when compared to
White/Caucasian individuals (Table~\ref{tbl-model}), the change in odds
of vaccination within certain racial groups and income strata was
actually positive, in contrast to the White/Caucasian group, for which
vaccination odds decreased in lower income brackets (when compared to
the CAD 60,000 and over bracket, Supplementary Figure A-3). More
specifically, the change in odds of vaccination increased in individuals
who identified as Arab/Middle Eastern, Black, Latin American, or
belonging to other minority groups with a household income below CAD
25,000, which was also true for individuals in other racial minority
groups with an income between CAD 25,000-59,999 (Table~\ref{tbl-model},
Supplementary Figure A-3).

This result is likely due to the fact that individuals that belong
racial minority groups tend to perform occupations that have been deemed
as ``essential'' in the context of the
pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{49},\protect\hyperlink{ref-ct2021}{50}},
which include occupations such as grocery store workers, gas station
workers, warehouse and distribution workers, and manufacturing workers,
all being occupations for which an income within the significant
brackets is to be expected. In the case of Ontario, essential workers
had priority for COVID-19
vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{51}},
which would explain the higher odds of vaccination for these individuals
in certain income brackets, in contrast to the lower odds of vaccination
for the same type of individuals with higher household income. In other
words, it is possible that the type of occupation played an important
role in increasing the odds of vaccination in these racial minority
groups.

Additionally, significant higher vaccination odds were identified in the
West Health Region when compared to the Health Region of Toronto
(Table~\ref{tbl-model}). The West Health Region comprises the regions of
Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and
the cities of Hamilton, Haldimand, Brant, and Chatham-Kent. In this
case, a possible rationale for the results is the fact that in the
survey, about 47\% of the entries for this Health Region corresponded to
White/Caucasian individuals, who reported an overall 83\% vaccination
rate (Supplementary Table A-6). However, the interaction effect of
Health Region and race was also significant in the case of individuals
identifying as South Asian or other minorities not included in the
survey Table~\ref{tbl-model}. In this case, the results of the
interaction term in the model indicate that the odds of vaccination for
those within the South Asian and Other minority groups in the West
Region decreased when compared to the other Health Regions
(Supplementary Figure A-4).

According to Ontario Health, 13.2\% of the population in the West Health
Region identifies as a visible minority, whereas 2.5\% identifies as
Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{52}}.
In the case of this analysis, the estimated lower odds are likely to be
explained from a socio-economic perspective. In fact, 50\% of the
answers from this region in the survey came from the former LHINs of
Hamilton Niagara Haldimand Brant, and Erie St.~Clair, both which are
among the regions of Ontario with the highest proportion of their
population (more than 20\%) in the lowest income
quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{53}}
(Supplementary Table A-7). Therefore, this result partly reinforces the
well-known existing association between low vaccination rates and
income, but it additionally indicates that there were intra-regional
differences in vaccination. Interestingly, a disproportionate number of
COVID-19 cases and low vaccination rate (under 50\%) have been
previously reported in the South Asian community of
Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{54}}; in this
regard, our result provides additional context by showing that within
the South Asian community, there were differences in vaccination uptake
across Ontario. Moreover, because significant lower odds of vaccination
were also identified other minority groups, this provides a rationale
for future studies that explore how vaccination uptake varies across
different minority groups within Ontario and other Canadian provinces.

There are some limitations to the present study. First, the data
collection design, which allowed respondents to withdraw from the survey
at any point, resulted in a high number of unique entries in the survey
with multiple missing answers. Because we focused on entries that had
complete observations in the covariates of interest for our analysis, it
is possible that some information was not considered by excluding
observations that had information in other variables (such as work from
home, or number of persons in the household). However, we attempted to
minimize this possibility by correcting the dataset using information
from the Census. More granular corrections, which for example could be
based on demographic information by municipality, could be used in the
future to obtain a more accurate approximation to the population totals
of the province. Additionally, the results in this study are based on
self-reported data, where the risk of bias exist. Despite this, because
in the context of COVID-19 it has been shown that good agreement exists
between self-reported and documented vaccination
status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{55}}, the
effect of self-reported bias is likely to not be significant in this
case.

Finally, it is likely that there have been differences in vaccination
across the province as more doses of the vaccine were administered and
as successive variants emerged. Because this study focused only on
vaccination status regarding the first dose of the vaccine within a
relatively short time window, it can only provide a snapshot of the
societal dynamics behind the pandemic. Nonetheless, the results
presented here can serve as a starting point to motivate future
longitudinal research that aims to quantify geographical differences
within vulnerable segments of the population, and that can be used to
inform the development of public health policies within the province of
Ontario or across other provinces that aim to minimize disparities in
health access.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

This study explored differences in COVID-19 vaccination across the
province of Ontario between late 2021 and early 2022 by taking into
consideration socio-economic factors, such as income and race, their
interactions, and the Health Regions within the province. Our results
show that, during the period analyzed, significant differences in
vaccination existed across different visible minority groups, income
brackets, and Health Regions, showing intra-provincial disparities in
vaccine uptake. As the COVID-19 continues around the world, it important
that future public policies take into consideration how to adequately
reach individuals within minority groups that live across geographical
areas where less probabilities of being vaccinated are likely. At the
moment, this is an ongoing issue that needs to be addressed to ensure a
more homogeneous outcome from the pandemic.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-WHO-Covid}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{{World Health Organization Coronavirus (COVID-19)
Dashboard}. Accessed November 27, 2022. \url{https://covid19.who.int/}}

\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
2020:m4799.
doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}

\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
social and policy view of vaccination efforts in ontario, canada.
\emph{Cambridge Journal of Regions, Economy and Society}. Published
online November 2022.
doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}

\leavevmode\vadjust pre{\hypertarget{ref-watson2022}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Watson OJ, Barnsley G, Toor J, Hogan AB, Winskill P,
Ghani AC. Global impact of the first year of {COVID}-19 vaccination: A
mathematical modelling study. \emph{The Lancet Infectious Diseases}.
2022;22(9):1293-1302.
doi:\href{https://doi.org/10.1016/s1473-3099(22)00320-6}{10.1016/s1473-3099(22)00320-6}}

\leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
determinants of {COVID}-19 vaccine hesitancy. Inbaraj LR, ed.
\emph{{PLOS} {ONE}}. 2021;16(11):e0258462.
doi:\href{https://doi.org/10.1371/journal.pone.0258462}{10.1371/journal.pone.0258462}}

\leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
inequality in {COVID}-19 vaccination coverage among elderly adults in
england: A national linked data study. \emph{{BMJ} Open}.
2021;11(7):e053402.
doi:\href{https://doi.org/10.1136/bmjopen-2021-053402}{10.1136/bmjopen-2021-053402}}

\leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
of {COVID}-19 vaccine acceptance in the {US}.
\emph{{EClinicalMedicine}}. 2020;26:100495.
doi:\href{https://doi.org/10.1016/j.eclinm.2020.100495}{10.1016/j.eclinm.2020.100495}}

\leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
inequalities in {COVID}-19 vaccination uptake and intent: Results from
the canadian community health survey 2021. \emph{{BMC} Public Health}.
2022;22(1).
doi:\href{https://doi.org/10.1186/s12889-022-14090-z}{10.1186/s12889-022-14090-z}}

\leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
Fahlman M, Kallio N. {COVID}-19 vaccine hesitancy and refusal and
associated factors in an adult population in saskatchewan, canada:
Evidence from predictive modelling. Gesser-Edelsburg A, ed. \emph{{PLOS}
{ONE}}. 2021;16(11):e0259513.
doi:\href{https://doi.org/10.1371/journal.pone.0259513}{10.1371/journal.pone.0259513}}

\leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
Overcoming {COVID}-19 vaccine hesitancy among ethnic minorities: A
systematic review of {UK} studies. \emph{Vaccine}.
2022;40(25):3413-3432.
doi:\href{https://doi.org/10.1016/j.vaccine.2022.04.030}{10.1016/j.vaccine.2022.04.030}}

\leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
Changes in {COVID}-19 vaccination receipt and intention to vaccinate by
socioeconomic characteristics and geographic area, united states,
january 6 {\textendash} march 29, 2021. \emph{Annals of Medicine}.
2021;53(1):1419-1428.
doi:\href{https://doi.org/10.1080/07853890.2021.1957998}{10.1080/07853890.2021.1957998}}

\leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
vaccine hesitancy in the united states. \emph{International Journal of
Environmental Research and Public Health}. 2021;18(18):9488.
doi:\href{https://doi.org/10.3390/ijerph18189488}{10.3390/ijerph18189488}}

\leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
spatial process: {COVID}-19 fully vaccinated rates in u.s. counties.
\emph{American Journal of Preventive Medicine}. 2022;63(6):954-961.
doi:\href{https://doi.org/10.1016/j.amepre.2022.06.006}{10.1016/j.amepre.2022.06.006}}

\leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
the spatiotemporal heterogeneity of the {COVID}-19 vaccination
landscape. \emph{American Journal of Epidemiology}.
2022;191(10):1792-1802.
doi:\href{https://doi.org/10.1093/aje/kwac080}{10.1093/aje/kwac080}}

\leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
social vulnerability index to assess {COVID}-19 vaccine uptake in
louisiana. \emph{{GeoJournal}}. Published online December 2022.
doi:\href{https://doi.org/10.1007/s10708-022-10802-5}{10.1007/s10708-022-10802-5}}

\leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
future spatial patterns in {COVID}-19 booster vaccine uptake. Published
online September 2022.
doi:\href{https://doi.org/10.1101/2022.08.30.22279415}{10.1101/2022.08.30.22279415}}

\leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
trajectories. \emph{Socius: Sociological Research for a Dynamic World}.
2021;7:237802312110529.
doi:\href{https://doi.org/10.1177/23780231211052946}{10.1177/23780231211052946}}

\leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
Social inequalities in {COVID}-19 vaccine acceptance and uptake for
children and adolescents in montreal, canada. \emph{Vaccine}.
2021;39(49):7140-7145.
doi:\href{https://doi.org/10.1016/j.vaccine.2021.10.077}{10.1016/j.vaccine.2021.10.077}}

\leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
review on vaccine hesitancy in black communities in canada: Critical
issues and research failures. \emph{Vaccines}. 2022;10(11):1937.
doi:\href{https://doi.org/10.3390/vaccines10111937}{10.3390/vaccines10111937}}

\leavevmode\vadjust pre{\hypertarget{ref-sargent2022}{}}%
\CSLLeftMargin{20. }%
\CSLRightInline{Sargent RH, Laurie S, Weakland LF, et al. Use of random
domain intercept technology to track {COVID}-19 vaccination rates in
real time across the united states: Survey study. \emph{Journal of
Medical Internet Research}. 2022;24(7):e37920.
doi:\href{https://doi.org/10.2196/37920}{10.2196/37920}}

\leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
\CSLLeftMargin{21. }%
\CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
a sampled frequency table when the expected marginal totals are known.
\emph{The Annals of Mathematical Statistics}. 1940;11(4):427-444.
doi:\href{https://doi.org/10.1214/aoms/1177731829}{10.1214/aoms/1177731829}}

\leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
\CSLLeftMargin{22. }%
\CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
Corlin L. Disparities in national and state estimates of {COVID}-19
vaccination receipt and intent to vaccinate by race/ethnicity, income,
and age group among adults~\(\geq\)~18~years, united states.
\emph{Vaccine}. 2022;40(1):107-113.
doi:\href{https://doi.org/10.1016/j.vaccine.2021.11.040}{10.1016/j.vaccine.2021.11.040}}

\leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
\CSLLeftMargin{23. }%
\CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
Zikmund-Fisher BJ. Vaccine hesitancy and rejection of a vaccine for the
novel coronavirus in the united states. \emph{Frontiers in Immunology}.
2021;12.
doi:\href{https://doi.org/10.3389/fimmu.2021.558270}{10.3389/fimmu.2021.558270}}

\leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
\CSLLeftMargin{24. }%
\CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
and factors related to {COVID}-19 vaccine hesitancy and unwillingness in
canada: A systematic review and meta-analysis. \emph{Journal of Medical
Virology}. 2022;95(1).
doi:\href{https://doi.org/10.1002/jmv.28156}{10.1002/jmv.28156}}

\leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
\CSLLeftMargin{25. }%
\CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
2011.}

\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
\CSLLeftMargin{26. }%
\CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
{tidyverse}. \emph{Journal of Open Source Software}. 2019;4(43):1686.
doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}}

\leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
\CSLLeftMargin{27. }%
\CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
Markdown Publishing System}.; 2022.
\url{https://CRAN.R-project.org/package=quarto}}

\leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
\CSLLeftMargin{28. }%
\CSLRightInline{Davis CJ, Golding M, McKay R. Efficacy information
influences intention to take COVID-19 vaccine. \emph{British Journal of
Health Psychology}. 2022;27(2):300-319.
doi:\url{https://doi.org/10.1111/bjhp.12546}}

\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
\CSLLeftMargin{29. }%
\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}

\leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
\CSLLeftMargin{30. }%
\CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
and emerging issues. \emph{Human Vaccines {\&}amp\(\mathsemicolon\)
Immunotherapeutics}. 2021;17(12):5255-5268.
doi:\href{https://doi.org/10.1080/21645515.2021.1977057}{10.1080/21645515.2021.1977057}}

\leavevmode\vadjust pre{\hypertarget{ref-li2021}{}}%
\CSLLeftMargin{31. }%
\CSLRightInline{Li Q, Wang J, Tang Y, Lu H. Next-generation COVID-19
vaccines: Opportunities for vaccine development and challenges in
tackling COVID-19. \emph{Drug Discoveries \& Therapeutics}.
2021;15(3):118-123.
doi:\href{https://doi.org/10.5582/ddt.2021.0105}{10.5582/ddt.2021.0105}}

\leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
\CSLLeftMargin{32. }%
\CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
to achieve global covid-19 vaccine equity. \emph{{BMJ}}. Published
online March 2022:e070650.
doi:\href{https://doi.org/10.1136/bmj-2022-070650}{10.1136/bmj-2022-070650}}

\leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
\CSLLeftMargin{33. }%
\CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
{COVID}-19 vaccine hesitancy: Race/ethnicity, trust, and fear.
\emph{Clinical and Translational Science}. 2021;14(6):2200-2207.
doi:\href{https://doi.org/10.1111/cts.13077}{10.1111/cts.13077}}

\leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
\CSLLeftMargin{34. }%
\CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
accepting {COVID}-19 vaccination during and after pregnancy, and for
their babies: A multi-methods study in the {UK}. \emph{{BMC} Pregnancy
and Childbirth}. 2022;22(1).
doi:\href{https://doi.org/10.1186/s12884-021-04321-3}{10.1186/s12884-021-04321-3}}

\leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
\CSLLeftMargin{35. }%
\CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
limits of medical trust in mitigating {COVID}-19 vaccine hesitancy among
black americans. \emph{Journal of General Internal Medicine}.
2021;36(11):3629-3631.
doi:\href{https://doi.org/10.1007/s11606-021-06743-3}{10.1007/s11606-021-06743-3}}

\leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
Sharma M, Webb FJ. {COVID}-19 vaccination hesitancy in the united
states: A rapid national assessment. \emph{Journal of Community Health}.
2021;46(2):270-277.
doi:\href{https://doi.org/10.1007/s10900-020-00958-x}{10.1007/s10900-020-00958-x}}

\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
\CSLLeftMargin{37. }%
\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
related medical mistrust, health impacts, and potential vaccine
hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}

\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
\CSLLeftMargin{38. }%
\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
canada. \emph{Canadian Medical Association Journal}.
2021;193(11):E381-E383.
doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}

\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
\CSLLeftMargin{39. }%
\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
2020;52(14):3127-3141.
doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}

\leavevmode\vadjust pre{\hypertarget{ref-pallathadka2022}{}}%
\CSLLeftMargin{40. }%
\CSLRightInline{Pallathadka A, Chang H, Han D. What explains spatial
variations of {COVID}-19 vaccine hesitancy?: A
social-ecological-technological systems approach. \emph{Environmental
Research: Health}. 2022;1(1):011001.
doi:\href{https://doi.org/10.1088/2752-5309/ac8ac2}{10.1088/2752-5309/ac8ac2}}

\leavevmode\vadjust pre{\hypertarget{ref-huang2022}{}}%
\CSLLeftMargin{41. }%
\CSLRightInline{Huang Q, Cutter SL. Spatial-temporal differences of
{COVID}-19 vaccinations in the u.s. \emph{Urban Informatics}. 2022;1(1).
doi:\href{https://doi.org/10.1007/s44212-022-00019-9}{10.1007/s44212-022-00019-9}}

\leavevmode\vadjust pre{\hypertarget{ref-lavoie2022}{}}%
\CSLLeftMargin{42. }%
\CSLRightInline{Lavoie K, Gosselin-Boucher V, Stojanovic J, et al.
Understanding national trends in {COVID}-19 vaccine hesitancy in canada:
Results from five sequential cross-sectional representative surveys
spanning april 2020{\textendash}march 2021. \emph{{BMJ} Open}.
2022;12(4):e059411.
doi:\href{https://doi.org/10.1136/bmjopen-2021-059411}{10.1136/bmjopen-2021-059411}}

\leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
\CSLLeftMargin{43. }%
\CSLRightInline{Dong L, Sahu R, Black R. Governance in the
transformational journey toward integrated healthcare: The case of
ontario. \emph{Journal of Information Technology Teaching Cases}.
Published online December 2022:204388692211473.
doi:\href{https://doi.org/10.1177/20438869221147313}{10.1177/20438869221147313}}

\leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
\CSLLeftMargin{44. }%
\CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
integrated care: A complex-adaptive systems perspective.
\emph{International Journal of Integrated Care}. 2012;12(5).
doi:\href{https://doi.org/10.5334/ijic.843}{10.5334/ijic.843}}

\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
\CSLLeftMargin{45. }%
\CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
compulsion: The case of {COVID}-19 vaccination in canada.
\emph{Vaccine}. 2022;40(29):3923-3926.
doi:\href{https://doi.org/10.1016/j.vaccine.2022.05.053}{10.1016/j.vaccine.2022.05.053}}

\leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
\CSLLeftMargin{46. }%
\CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
of canada {COVID}-19 report: Enhancing {COVID}-19 vaccine acceptance in
canada. Blais JM, ed. \emph{{FACETS}}. 2021;6:1184-1246.
doi:\href{https://doi.org/10.1139/facets-2021-0037}{10.1139/facets-2021-0037}}

\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
\CSLLeftMargin{47. }%
\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
2023.
\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}

\leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
\CSLLeftMargin{48. }%
\CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
{COVID}-19 vaccine uptake in southeastern ontario, canada: Monitoring
and addressing health inequities. \emph{Journal of Public Health
Management and Practice}. 2022;28(6):615-623.
doi:\href{https://doi.org/10.1097/phh.0000000000001565}{10.1097/phh.0000000000001565}}

\leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
\CSLLeftMargin{49. }%
\CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
and other infection exposure according to race and ethnicity.
\emph{American Journal of Industrial Medicine}. 2020;63(9):817-820.
doi:\href{https://doi.org/10.1002/ajim.23145}{10.1002/ajim.23145}}

\leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
\CSLLeftMargin{50. }%
\CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
review of {COVID}-19 and vulnerable workers: Intersecting occupational
and public health issues. \emph{American Journal of Industrial
Medicine}. 2021;64(7):551-566.
doi:\href{https://doi.org/10.1002/ajim.23256}{10.1002/ajim.23256}}

\leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
\CSLLeftMargin{51. }%
\CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
Strategy for Ontario {COVID}-19 Hotspots and Essential Workers}. Ontario
{COVID}-19 Science Advisory Table; 2021.
doi:\href{https://doi.org/10.47326/ocsat.2021.02.26.1.0}{10.47326/ocsat.2021.02.26.1.0}}

\leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
\CSLLeftMargin{52. }%
\CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
Health;
\url{https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf};
2022.}

\leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
\CSLLeftMargin{53. }%
\CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
Rosella LC. Ontario atlas of adult mortality, 1992-2015: Trends in
{L}ocal {H}ealth {I}ntegration {N}etworks. Published online 2018.
\url{https://tspace.library.utoronto.ca/handle/1807/82836}}

\leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
\CSLLeftMargin{54. }%
\CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
and risk factors for {SARS}-{CoV}-2 infection in a south asian community
in ontario: A cross-sectional analysis of a prospective cohort study.
\emph{{CMAJ} Open}. 2022;10(3):E599-E609.
doi:\href{https://doi.org/10.9778/cmajo.20220031}{10.9778/cmajo.20220031}}

\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
\CSLLeftMargin{55. }%
\CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
vaccination status by self-report versus source documentation: Impact on
measuring {COVID}-19 vaccine effectiveness. \emph{Influenza and Other
Respiratory Viruses}. 2022;16(6):1101-1111.
doi:\href{https://doi.org/10.1111/irv.13023}{10.1111/irv.13023}}

\end{CSLReferences}

\end{document}

### 2023-03-03 manuscript/main.qmd
updated

@@
 In particular, we show
-that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32, 0.28, and _p_=0.004,_p_<0.001 and _p_=0.004, respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets.
+that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32, 0.28, and _p_=0.004, _p_<0.001 and _p_=0.004, respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets.
 Finally, we also identified lower odds of vaccination within different minority groups in West Health Region.

### 2023-03-03 manuscript/main.tex
updated

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32,
-0.28, and \emph{p}=0.004,\emph{p}\textless0.001 and \emph{p}=0.004,
+0.28, and \emph{p}=0.004, \emph{p}\textless0.001 and \emph{p}=0.004,
 respectively), and that individuals with a household income below CAD

### 2023-03-03 manuscript/main.tex
minor changes

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, interior hidden, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 

### 2023-03-03 code/raking_and_model.R

@@ -1,461 +1,0 @@
-#### this script uses the clean dataset to: 1) correct the observations in the survey
-#### response using raking and 2) compute a model of logistic regression
-
-#load packages
-
-library(weights)
-library(here)
-library(survey)
-library(magrittr)
-library(tidyverse)
-library(lme4)
-library(gtsummary)
-
-data<-read.csv(here("data","clean_dataset.csv"))
-
-# remove rows with missing observations for race
-
-clean_data<-data%>%drop_na(race)
-
-#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
-
-clean_data<-clean_data%>%
-  select(-Municipal.status)
-
-#make variables of interest factors
-clean_data<-clean_data %<>%
-  mutate_at(c("income_ord","race","age_group","location","Geographic_area","Geographic_area_title"),factor)
-# recode the response of interest for categorical analysis
-clean_data<-clean_data %>%
-  mutate(first_dose_m=case_when(
-    first_dose=="yes"~1,
-    first_dose=="no"~0
-  ))
-
-
-# get percentages to see how different the data is from the Census (details of the Census values are in
-# Data_cleaning_and_covariates_Jan_05.qmd in directory "data cleaning")
-
-wpct(clean_data$income_ord)
-wpct(clean_data$age_group)
-wpct(clean_data$race)
-wpct(clean_data$location)
-
-##### Summary of the dataset####
-
-
-data_summary<-clean_data %>%
-  tbl_summary(percent="row",
-    by= first_dose,
-    include=c(income_ord,
-    age_group,
-    Health_Region,
-    race)
-  )
-
-################ RAKING ###################################
-## I had originally used the percentages from the census, but the "population values"
-## that need to be provided in the examples of the survey package are all based on total counts.
-## This would make sense as to estimate the sampling probability we input from the dataset are counts as well.
-## I transformed all the percentages using the total population for Ontario (from the 2016 census) and 
-## household counts (also from the 2016 Census).
-
-## Each of these dataframes provide the population totals for the different variables and their categories
-
-ages<-data.frame(age_group=c("16_24",
-                             "25_34",
-                             "35_44",
-                             "45_54",
-                             "55_64",
-                             "65_and_over"),
-                 Freq=c(1707959,
-                        1734856,
-                        1721407,
-                        2003826,
-                        1842444,
-                        2245898))
-
-
-
-races<-data.frame(race=c("arab_middle_eastern",
-                         "black",
-                         "east_asian_pacific_islander",
[…]

### 2023-03-03 code/raking_logistic_regression.R

@@ -1,111 +1,0 @@
-#Each of these dataframes provides the population totals for the different variables and their categories
-
-# clean_data %>%
-#   tbl_summary(
-#     include=c(
-#     Health_Region,)
-#   ) %>% bold_labels()
-
-
-Ages<-data.frame(Age_group=c("16_24",
-                             "25_34",
-                             "35_44",
-                             "45_54",
-                             "55_64",
-                             "65_and_over"),
-                 Freq=c(1707959,
-                        1734856,
-                        1721407,
-                        2003826,
-                        1842444,
-                        2245898))
-
-
-
-Races<-data.frame(Race=c("arab_middle_eastern",
-                         "black",
-                         "east_asian_pacific_islander",
-                         "indigenous",
-                         "latin_american",
-                         "mixed",
-                         "other",
-                         "south_asian",
-                         "white_caucasian"),
-                  Freq=c(212782,
-                         638346,
-                         886592,
-                         376558,
-                         197020,
-                         130033,
-                         705333,
-                         1166361,
-                         9118079 ))
-
-Incomes<-data.frame(Income=c("under_15000",
-                             "15000_24999",
-                             "25000_39999",
-                             "40000_59999",
-                             "60000_89999",
-                             "over_90000"),
-                    Freq=c(294643,
-                           387688,
-                           599624,
-                           796052,
-                           1007988,
-                           2083176
-                           
-                    ))
-
-
-# corrections for Health Regions, without the North East and North West regions
-
-Health_Regions<-data.frame(Health_Region=c(
-  #"North West",
-  #"North East",
-  "West",
-  "East",
-  "Central",
-  "Toronto"),
-  Freq=c(#232299,
-    #557000,
-    4095589,
-    3742520,
-    5032410,
-    1440644
-  )
-)
-
-
-
-### Raking ####
-
-## First, provide a survey design object. This is done using the dataset. The syntax
-## means that there is no stratification (id=~1). When the line below is run it will throw
-## a warning because we do not have "design" weights in the dataset (other datasets do have this).
-
-
-a1<-svydesign(id=~1,data=clean_data)
-
-## because there are no sampling weights, each observation as a probability of 1 of being sampled at this stage. 
-
-
-## The next line uses the survey design object a1, and the variables we will adjust for, with the population corrections
-## from the data frames from above.
-
-a1_rake<-rake(a1,
-              sample=list(~Age_group,
-                          ~Race,
[…]

### 2023-03-22 manuscript/appendix.tex
updated analyses, removed shp files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata},
  pdfauthor={Ariel Mundo Ortiz1; Bouchra Nasri2,},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{COVID-19 vaccination in Ontario: Exploring
intra-provincial variations within Health Regions and socio-economic
strata}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Appendix}
\author{Ariel Mundo Ortiz\textsuperscript{1} \and Bouchra
Nasri\textsuperscript{2,*}}
\date{}

\begin{document}
\maketitle
%start sections and page numbers with A
\setcounter{page}{1}
\renewcommand\thesection{A}
\renewcommand\thesubsection{\thesection.\arabic{subsection}}
\renewcommand{\thepage}{A-\arabic{page}}
\renewcommand{\thetable}{A-\arabic{table}}
\renewcommand\thefigure{\thesection-\arabic{figure}}

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, frame hidden]}{\end{tcolorbox}}\fi

\setstretch{2}
\textsuperscript{1} Centre de Recherches Mathématiques, University of
Montreal, Montréal, Canada\\
\textsuperscript{2} Department of Social and Preventive Medicine, École
de Santé Publique, University of Montreal, Montréal, Canada

\textsuperscript{*} Correspondence:
\href{mailto:bouchra.nasri@umontreal.ca}{Bouchra Nasri
\textless{}bouchra.nasri@umontreal.ca\textgreater{}}

This document provides details on assessing the socio-economic
information from the clean survey dataset, and how it compared to the
Census Data for ontario. At the end, it also provides information about
the corrections implemented into the data. The variables examined were:
race/ethnicity, age groups, income, and population by Health Region. The
computations are performed in the \texttt{clean\_dataset.csv} file that
was created at the end of the data cleaning process. Population-wide
totals were obtained from the 2016, available in the
\href{https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/Page.cfm?Lang=E\&Geo1=PR\&Code1=35\&Geo2=\&Code2=\&Data=Count\&SearchText=Ontario\&Sear}{Statistics
Canada Website}.

\hypertarget{race-and-ethnicity}{%
\section{Race and Ethnicity}\label{race-and-ethnicity}}

First, we explored how the race and ethnicity information from the
dataset and how it compared to the data from the Census for Ontario. The
following chunk creates a summary table for the Race variable.

\begin{longtable}{lrr}
\caption{Ethnic information from the clean dataset}\tabularnewline

\toprule
race & observations & percentage \\ 
\midrule
arab\_middle\_eastern & 218 & $5.9\%$ \\ 
black & 306 & $8.3\%$ \\ 
east\_asian\_pacific\_islander & 308 & $8.3\%$ \\ 
indigenous & 216 & $5.8\%$ \\ 
latin\_american & 182 & $4.9\%$ \\ 
mixed & 324 & $8.7\%$ \\ 
other & 389 & $10.5\%$ \\ 
south\_asian & 383 & $10.3\%$ \\ 
white\_caucasian & 1383 & $37.3\%$ \\ 
\bottomrule
\end{longtable}

It is important to mention that the categories for race/ethnicity
provided in the survey did not match the categories used in the Census.
Therefore, we used a combination of sources to obtain estimates of
racial/ethnic distribution that matched the categories from the
survey.The data sources were:

\begin{itemize}
\item
  Fact sheet from the Provice of Ontario for Visible Minorities
  \href{https://www.ontario.ca/document/2016-census-highlights/fact-sheet-9-ethnic-origin-and-visible-minorities}{link}:
  Used to obtain percentages for Arab, Black, East Asian/Pacific
  Islander (adding Chinese, Korean, and Japanese percentages), Latin
  American, Mixed (using the percentage for ``multiple visible
  minority''), Other (obtained by adding the Southeast Asian, Filipino,
  West Asian, and Minority not identified elsewhere percentages), South
  Asian. \emph{Accessed on January 05, 2022}
\item
  Census Profile for Ontario
  \href{https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E\&Geo1=PR\&Code1=35\&Geo2=PR\&Code2=01\&SearchText=Ontario\&SearchType=Begins\&SearchPR=01\&B1=Aboriginal\%20peoples\&TABID=1\&type=1}{link}:
  Used to obtain percentage of Aboriginal population. \emph{Accessed on
  January 05, 2022}
\item
  Wikipedia entry for Ontario demographics
  \href{https://en.wikipedia.org/wiki/Demographics_of_Ontario}{link}: To
  corroborate that the percentage of population reported as ``European''
  in this website matched the percentage obtained for ``White
  Caucasian'' that was independently obtained by obtaining the
  difference in population proportion after subtracting the sum of
  Visible Minorities and Aboriginal Population percentages. The
  aggregated information can be found in in Table~\ref{tbl-races}, where
  the totals per race/ethnic category are presented.
\end{itemize}

\hypertarget{tbl-races}{}
\begin{longtable}[]{@{}lll@{}}
\caption{\label{tbl-races}Reference Data from the 2016 Census from
Race/Ethnicity in Ontario}\tabularnewline
\toprule()
Ethnicity/Race & Percentage & Population totals \\
\midrule()
\endfirsthead
\toprule()
Ethnicity/Race & Percentage & Population totals \\
\midrule()
\endhead
Arab & 1.6\% & 212782 \\
Black & 4.7\% & 638346 \\
East Asian/Pacific Islander & 6.6\% & 886592 \\
Indigenous & 2.8\% & 376558 \\
Latin American & 1.5\% & 197020 \\
Mixed & 1.0\% & 130033 \\
Other & 5.2\% & 705333 \\
South Asian & 8.7\% & 1166361 \\
White Caucasian & 67.8\% & 9118079 \\
\textbf{Total} & 99.9\% & 13431105 \\
\bottomrule()
\end{longtable}

Visible Minorities: 29.3\% Not a Visible Minority: 70.7\% (Aboriginal
2.8\%, White Caucasian 67.8\%)

It can be seen that there are differences in the distribution of each
ethnicity between the Census data and the survey. For example,
Arabs/Middle Easterner correspond to 5.9\% of the survey responses, but
population-wise individuals that identify in this ethnic group
correspond to 1.6\% of the Ontario population.

\hypertarget{age-groups}{%
\section{Age Groups}\label{age-groups}}

According to the Census, the distribution of the different the age
groups for the province of Ontario are as follows:

\begin{longtable}[]{@{}lll@{}}
\caption{Age distributions from the 2016 Census for
Ontario}\tabularnewline
\toprule()
Group age & Percentage & Population Totals \\
\midrule()
\endfirsthead
\toprule()
Group age & Percentage & Population Totals \\
\midrule()
\endhead
15-24 & 12.7\% & 1707959 \\
25-34 & 12.9\% & 1734856 \\
35-44 & 12.8\% & 1721407 \\
45-54 & 14.9\% & 2003826 \\
55-64 & 13.7\% & 1842444 \\
65 and over & 16.7\% & 2245898 \\
\bottomrule()
\end{longtable}

To compare the survey data to the Census, the next chunk creates a
barplot of the age groups in the survey. Here, it can be noticed that
the age-group distribution from the survey data is different from the
Census data.

\begin{figure}

{\centering \includegraphics{appendix_files/figure-pdf/trends-by-age-1.pdf}

}

\caption{Age group distributions from the dataset}

\end{figure}

\hypertarget{income}{%
\section{Income}\label{income}}

Survey respondents answered the question ``What is your household annual
income?''. To compare the distribution of responses to this question
from the survey and the Census data, we used the ``Household total
income groups in 2015 for private households'' from the Census data
(available in the Census Data for Ontario website
\href{https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/page.cfm?Lang=E\&Geo1=PR\&Code1=35\&Geo2=PR\&Code2=01\&SearchText=Ontario\&SearchType=Begins\&SearchPR=01\&B1=Income\&TABID=1\&type=1}{link}).

\begin{longtable}[]{@{}lll@{}}
\caption{Income percentages from the 2016 Census for
Ontario}\tabularnewline
\toprule()
Household income range (CAD) & Percentage & Population Totals \\
\midrule()
\endfirsthead
\toprule()
Household income range (CAD) & Percentage & Population Totals \\
\midrule()
\endhead
\textless{} 15,000 & 5.7\% & 294643 \\
15,000 - 24,999 & 7.5\% & 387688 \\
25,000 - 39,999 & 11.6\% & 599624 \\
40,000- 59,999 & 15.4\% & 796052 \\
60,000 - 89,999 & 19.5\% & 1007988 \\
\textgreater90,000 & 40.3\% & 2083176 \\
\bottomrule()
\end{longtable}

One difference that is noticeable is that the brackets for income in the
census data are different than the brackets used in the survey. The
census does CAD 4,999 brackets (e.g., CAD 5,000- CAD 9,9999) up to CAD
49,999, followed by CAD 9,999 brackets up to CAD 99,999. After that, the
brackets increase to CAD 24,999, and therefore, it is not possible to
obtain percentages for the 90,000-109,999 and \textgreater110,000
brackets from the survey.

Therefore, in the following code chunk, we created an additional
category for income in the dataset that matched the information from the
Census, and barplots to visualize the proportion of each income bracket
in the dataset.

\begin{figure}

{\centering \includegraphics{appendix_files/figure-pdf/income-1.pdf}

}

\caption{Income bracket distribution in the dataset}

\end{figure}

We identified differences again between the survey data and the Census
data. For example, the \textless15,000 bracket accounts for about 24\%
of the responses, a much higher rate than the proportion from the Census
(5.7\%). The other brackets were different as well from the Census data.

\hypertarget{health-regions-population}{%
\section{Health Regions Population}\label{health-regions-population}}

Finally, we obtained population totals for each of the Health Regions
included in the dataset from the Ontario Health Business Plan for
2022-2023 (available
\href{https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf}{here}).
The population totals are presented in Table~\ref{tbl-health-regions}.

\hypertarget{tbl-health-regions}{}
\begin{longtable}[]{@{}ll@{}}
\caption{\label{tbl-health-regions}Population Totals for the Health
Regions in Ontario}\tabularnewline
\toprule()
Health Region & Population Totals \\
\midrule()
\endfirsthead
\toprule()
Health Region & Population Totals \\
\midrule()
\endhead
North East & 232299 \\
North West & 557000 \\
West & 4095589 \\
East & 3742520 \\
Central & 5032410 \\
Toronto & 1440644 \\
\bottomrule()
\end{longtable}

Note that as indicated in the paper, we excluded observations from the
North East and North West Health Regions, and therefore, the population
totals for these areas were not used in the final analysis.

\hypertarget{corrections}{%
\section{Corrections}\label{corrections}}

We used the iterative proportional fitting procedure (\emph{raking}) in
order to account for the differences in income, age groups,
race/ethnicity and the population from the Health Regions. Details of
the implementation can be found in the R Script
\texttt{raking\_logistic\_regression.R} in the \texttt{code} directory.

\hypertarget{interactions-from-the-regression-model}{%
\section{Interactions from the Regression
Model}\label{interactions-from-the-regression-model}}

The plots below show the interactions for Race and Income and Race and
Health Region based on the regression model presented in the main paper.

\begin{figure}

{\centering \includegraphics{appendix_files/figure-pdf/interaction-race-income-1.pdf}

}

\caption{Estimated marginal means for the combination between Race and
Income in the regression model.}

\end{figure}

\begin{figure}

{\centering \includegraphics{appendix_files/figure-pdf/interaction-race-hr-1.pdf}

}

\caption{Estimated marginal means for the combination between Race and
Health Region in the regression model.}

\end{figure}

\hypertarget{socio-economic-information-from-the-survey-data}{%
\section{Socio-economic Information from the Survey
Data}\label{socio-economic-information-from-the-survey-data}}

The tables below present summaries of information from the Fields survey
(from the clean dataset analyzed in the main paper), showing responses
by White/Caucasian individuals by Health Region and proportion of
answers by LHIN in the West Health Region.

\hypertarget{tbl-hr-percentages}{}
\begin{longtable}[]{@{}lcc@{}}
\caption{\label{tbl-hr-percentages}Reported Vaccination Status from the
Survey for White/Caucasian Individuals by Health Region}\tabularnewline
\toprule()
\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N =
1,080 \\
\midrule()
\endfirsthead
\toprule()
\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N =
1,080 \\
\midrule()
\endhead
\textbf{Health\_Region} & & \\
Toronto & 85 (21\%) & 321 (79\%) \\
West & 69 (17\%) & 329 (83\%) \\
East & 41 (15\%) & 236 (85\%) \\
Central & 38 (16\%) & 194 (84\%) \\
\bottomrule()
\end{longtable}

\hypertarget{tbl-west-hr-lhin}{}
\begin{longtable}[]{@{}lc@{}}
\caption{\label{tbl-west-hr-lhin}Proportion of answers by LHIN (West
Health Region)}\tabularnewline
\toprule()
\textbf{Characteristic} & \textbf{N = 839} \\
\midrule()
\endfirsthead
\toprule()
\textbf{Characteristic} & \textbf{N = 839} \\
\midrule()
\endhead
\textbf{LHIN} & \\
Hamilton Niagara Haldimand Brant (Hnhb) & 271 (32\%) \\
Waterloo Wellington & 227 (27\%) \\
South West & 187 (22\%) \\
Erie St.~Clair & 154 (18\%) \\
\bottomrule()
\end{longtable}

\end{document}

### 2023-03-22 manuscript_v2/01-background.qmd
updated analyses, removed shp files

# Background

The vaccines against COVID-19 have been considered a major achievement of modern medicine as their rapid development[@davis2022] allowed the start of broad vaccination campaigns towards the end of 2020 in certain countries such as the US[@tanne2020] and Canada[@bogoch2022]. This made some believe that vaccines were destined to be a determinant factor in a rapid ending of the pandemic [@thelancet2021]. However, although it has been estimated that COVID-19 vaccines have prevented around 14 million of deaths worldwide [@watson2022], their implementation has been far from being equal to that of the vaccines of smallpox and polio, which were implemented on a global scale and that were indeed crucial to control these diseases [@kayser2021]. In fact, the rollout of COVID-19 vaccines has faced multiple challenges since its inception which ultimately have hampered the goal of global immunity.

This ongoing lack of a successful global rollout of the COVID-19 vaccines is a multifaceted issue resulting from, as among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020]. In particular, it is well established that differences in vaccination uptake have been present even in the case of countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within racial minorities (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021]. Reasons given for lower vaccine uptake in these cases have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020]. More specifically, in the case of Canada, lower vaccine uptake has been observed in young individuals, those with a low educational level, households with children, those without a regular healthcare provider, individuals that identify as part of a visible minorities or Indigenous, and those that are financially unstable [@guay2022;@muhajarine2021;@hussain2022].

However, it is important to consider that vaccination uptake can also be influenced by geographical (spatial) factors. 
In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], spatial differences in vaccine access and supply, vaccination location availability, and lack of prioritization of areas where vulnerable groups reside [@bogoch2022; @nguyen2021]. Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that and that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022]. Overall, the existing body of  research on geographical/spatial differences in vaccination has provided valuable information that can be used by decision-makers in the development of public health policies that are focused on improving vaccination within specific administrative or geographical areas. However, spatial analyses of vaccination have been very limited in Canada, as existing studies have been focused only at certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022]. However, there is a lack of studies that explore spatial heterogeneities in vaccination at the intra-provincial level.

Ontario is the most populated province in Canada, and has recently seen major changes on its healthcare structure as between 2006 and 2019, it provided healthcare access using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs). However, this approach was complex and caused systemic inequalities [@tsasis2012]. In late 2019, the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East) in an effort to improve the healthcare system of the province[@dong2022]. Because the adoption of the Health Regions has been recent, there is an ongoing need to analyze the impact of their adoption and intra-regional differences that might exist, which is specially important in the context of the COVID-19 pandemic.

To address this need, in this study we analyzed differences in self-reported COVID-19 vaccination status using socio-economic (e.g., income, racial/ethnic identification), and spatial  information at the level of the Health Regions, to identify the existence of intra-provincial differences that might need to be addressed to ensure that the healthcare system of Ontario is more inclusive and that responds to the needs of its most vulnerable population.

### 2023-03-22 manuscript_v2/02-methods.qmd
updated analyses, removed shp files

# Methods

## Data {#sec-data}

We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto. The survey collected socio-economic information from participants (@tbl-covariates), recorded their location (using the nearest municipality), and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no". The original dataset contained 39,029 entries (where each entry corresponded to a unique respondent).

This dataset was cleaned to remove outliers that were identified during preliminary analyses, and processing the geographical information in the survey (city where the survey was responded) in order to match each city to its correspondent Health Region.

| Variable             | Levels                                            |            
|:---------------------|---------------------------------------------------|
| Age group            | 16-34,35-54,55 and over                           |
| Income bracket (CAD) | under 25,000, 25,000-59,999, 60,000 and above     |
| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, Other |

: Socio-economic factors from the Fields COVID-19 survey {#tbl-covariates}

The clean dataset contained responses from more than 200 different municipalities within Ontario (@fig-map). Because of the lack of a publicly available list of all municipalities within each Health Region, we used a dataset of long-term care homes and LHINs to match each city to LHIN, followed by matching each LHIN to a Health Region following the information provided on the Ontario Health Website, where the list of LHINs and corresponding Health Regions is available. In the case of municipalities that did not appear in the long-term care home dataset, we manually searched each city in the LHINs websites in order to provide geographical information. The original dataset, clean dataset, and details on the data cleaning process are described in detail in the GitHub repository for this paper, which can be found at [https://github.com/aimundo/Fields_COVID-19/](https://github.com/aimundo/Fields_COVID-19/).

Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions. The low representation of these Health Regions in the dataset is noticeable in @fig-map, where the number of cities in each Health Region is relatively low compared to the most populated Health Regions, such as the Toronto or Central Regions. We omitted the North East and North East Health Regions from further analyses due to the low number of entries. Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions.

```{r,map}
#| label: fig-map
#| fig-cap: "Geographic representation of the data collected by the _Survey of COVID-19 related Behaviours and Attitudes_, collected by the Fields Institute in Ontario. The municipalities (cities) from where survey participants provided answers (in the clean dataset) appear as points. The Health six Regions are color-coded. Internal boundaries within certain Health Regions indicate areas that belonged to the Local Integrated Health Networks (LHINs), the geographic areas for healthcare in Ontario before the adoption of the Health Regions."
#| message: false
#| echo: false
#| fig-height: 3
#| figh-width: 2

knitr::include_graphics(here("data","map_data","map.pdf"))

```

## Statistical analyses

We used a logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @tbl-covariates, the Health Regions from Ontario indicated in @sec-data, and the interactions between Race and Health Region, and Race and income, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].

<!-- $$ -->
<!-- \begin{aligned} -->
<!-- \log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Race)} + \beta_6 \textrm{ (Income} \times \textrm{Race)} -->
<!-- \end{aligned} -->
<!-- $$ {#eq-model3} -->

<!-- Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a COVID-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term.  -->

The model was fitted first to the clean dataset to obtain uncorrected estimates. Additionally, because we identified differences between the proportions of all the socio-economic factors included in the analysis (@tbl-covariates) and the Census data for Ontario, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using Census socio-economic data and Health Region population totals, in order to obtain corrected estimates from the model. Details regarding the correction can be found in the Appendix. 
All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], and `quarto` [@quarto].

### 2023-03-22 manuscript_v2/03-results.qmd
updated analyses, removed shp files

--- 2023-03-02 manuscript/03-results.qmd
@@
 
-@tbl-model shows the results of the logistic regression on vaccination status using socio-economic factors (age group, income, race), geographical areas (Health Regions) and the interactions between income and race and Health Region and race.
-The reference groups were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income).
+@tbl-model shows the results of the logistic regression models (for the uncorrected and corrected data) on vaccination status using socio-economic factors (age group, income, race), geographical areas (Health Regions) and the interactions between income and race and Health Region and race.
 There were no statistically significant differences in vaccination rates within the age groups from the survey, but significant odds ratios were estimated for other covariates.
-Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, respectively).
+Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, _p_=0.011 and <0.001, respectively).
 Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.28, and _p_=0.004,<0.001 and 0.004, respectively); additionally, those individuals in the Other Race/Ethnicity group (a group that included Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere) had even lower odds of vaccination than the other minority groups (OR=0.22, _p_<0.001).
@@
 
+#call the script for raking
+source(here::here("code","raking.R"))
+
 #logistic regression
@@
 
-# create table with the model output
-m1 %>%
+# create tables with the model outputs
+
+t0<-m0 %>%
   tbl_regression(exponentiate = TRUE,
@@
                             income_ord="Income")) %>%
+  bold_labels()
+
+t1<-m1 %>%
+  tbl_regression(exponentiate = TRUE,
+                 label=list(Health_Region ~ "Health Region",
+                            Age_group_ord ~ "Age Group",
+                            income_ord="Income")) %>%
+  bold_labels()
+
+tbl_merge <-
+  tbl_merge(
+    tbls = list(t0, t1),
+    tab_spanner = c("**Uncorrected**", "**Corrected**")
+  )%>%
   bold_labels()%>%
@@
   kableExtra::kable_styling(latex_options = "repeat_header")
+
+tbl_merge
+
+# m1 %>%
+#   tbl_regression(exponentiate = TRUE,
+#                  label=list(Health_Region ~ "Health Region",
+#                  Age_group_ord ~ "Age Group",
+#                  income_ord="Income")) %>%
+#     bold_labels()%>%
+#   as_kable_extra(booktabs = TRUE, longtable=T, linesep="") %>%
+#   # reduce font size to make table fit. 
+#   # you may also use the `latex_options = "scale_down" argument here.
+#   kableExtra::kable_styling(latex_options = "repeat_header")
 ```

### 2023-03-22 manuscript_v2/04-discussion.qmd
updated analyses, removed shp files

--- 2023-03-02 manuscript/04-discussion.qmd
@@
 
-The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022].
-Vaccine availability towards the end of 2020 in certain countries made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021].
-However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of COVID-19 have faced multiple challenges that have complicated the achievement of global immunity.
-
-Among the different challenges faced by COVID-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and inequity in vaccine access between low and high income countries [@tamey2022].
-However, it is also well established that even in the case of high income countries that have had ample access to vaccines since 2020, such as the US, the UK, and Canada, there have been challenges in vaccination efforts due to differences in vaccine uptake among different segments of the population.
-More specifically, lower vaccine uptake has been associated with socio-economic factors such as race (i.e., identifying as Black, Asian, Indigenous) and household income (typically within lower income brackets) [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021].
-Reasons given for this association have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020].
-
-In addition, vaccine uptake is influenced by geography, as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021].
-However, in the case of Canada, studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022].
-Therefore, we explored spatial and socio-economic determinants of vaccination status in the province of Ontario.
-This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Network (LHIN) system and the incorporation of regions covered by LHINs into larger Health Regions [@dong2022].
-Because the idea behind the change aimed to reduce the inequalities in healthcare that were identified under the LHIN model [@tsasis2012], examining differences in vaccination between the Health Regions can provide decision-makers with insight regarding intra-provincial health disparities that may need to be addressed in future vaccination or public health campaigns.
+The change from LHINs to Health Regions for healthcare in Ontario was aimed at reducing healthcare disparities [@tsasis2012].
+The relatively recent nature of the change, which occurred in late 2019, highlights the importance of studies that can determine if there are intra-regional differences in healthcare that need to be addressed, particularly in the context of the COVID-19 pandemic.
 
@@
 More granular corrections, which for example could be based on demographic information by municipality, could be used in the future to obtain a more accurate approximation to the population totals of the province.
-Additionally, the results in this study are based on self-reported data, where the risk of bias exist.
+Moreover, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map).
+Although low representation from these areas is based on the fact that these regions only account for 5\% of the total population of Ontario, these regions are the home to more than 100,000 individuals that identify as Indigenous [@ontariohealth], a minority group that has historically suffered from reduced access to health care and discrimination [@mosby2021].
+Therefore, there is a need for additional studies that focus on these low-populated Health Regions in Ontario where disparities in vaccination might be significant and understudied.
+
+The results in this study are based on self-reported data, where the risk of bias exist.
 Despite this, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], the effect of self-reported bias is likely to not be significant in this case.
-
 Finally, it is likely that there have been differences in vaccination across the province as more doses of the vaccine were administered and as successive variants emerged.

### 2023-03-22 manuscript_v2/appendix.qmd
updated analyses, removed shp files

--- 2023-03-02 manuscript/appendix.qmd
@@
 ---
-title: '**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**'
+title: '**COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata**'
 subtitle: Appendix
-author: Ariel Mundo Ortiz
+authors:
+  - name: Ariel Mundo Ortiz
+    affiliations:
+      - ref: crm
+    email: ariel.mundo.ortiz@umontreal.ca
+    orcid: 0000-0002-6014-4538
+  - name: Bouchra Nasri
+    affiliations:
+      - ref: espum
+    email: bouchra.nasri@umontreal.ca
+    corresponding: true
+affiliations:
+  - id: crm
+    name: Centre de Recherches Mathématiques, University of Montreal, Montréal, Canada
+  - id: espum
+    name: Department of Social and Preventive Medicine, École de Santé Publique, University of Montreal, Montréal, Canada 
+filters:
+  - authors-block
 format: 
   pdf:
+    #mainfont: "Palatino Linotype"
     include-before-body: latex/appendix_preamble.sty
-#bibliography: refs.bib
-#csl: jama.csl
-#link-citations: yes #adds links to the citations
+linestretch: 2
+keep-tex: true
 ---

### 2023-03-22 manuscript_v2/main.qmd
updated analyses, removed shp files

--- 2023-03-03 manuscript/main.qmd
@@
     include-in-header: latex/preamble.sty
-    mainfont: "Palatino Linotype"
+    #mainfont: "Palatino Linotype"
     keep-tex: true
@@
 that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32, 0.28, and _p_=0.004, _p_<0.001 and _p_=0.004, respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets.
-Finally, we also identified lower odds of vaccination within different minority groups in West Health Region.
+Finally, we also identified lower odds of vaccination within certain minority groups in the West Health Region, which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
 This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population that have been largely impacted by the pandemic.

### 2023-03-22 manuscript_v2/main.tex
updated analyses, removed shp files

--- 2023-03-03 manuscript/main.tex
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-  \setmainfont[]{Palatino Linotype}
 \fi
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, interior hidden, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 from the same racial/ethnic group in higher income brackets. Finally, we
-also identified lower odds of vaccination within different minority
-groups in West Health Region. This study shows that there is an ongoing
-need to better understand and address differences in vaccination uptake
-across diverse segments of the population that have been largely
-impacted by the pandemic.
+also identified lower odds of vaccination within certain minority groups
+in the West Health Region, which comprises the regions of Waterloo and
+Niagara, the counties of Wellington, Essex and Lambton, and the cities
+of Hamilton, Haldimand, Brant, and Chatham-Kent. This study shows that
+there is an ongoing need to better understand and address differences in
+vaccination uptake across diverse segments of the population that have
+been largely impacted by the pandemic.
 
@@
 
-The COVID-19 pandemic continues around the world with more than 600
-million confirmed cases as of November of
-2022\textsuperscript{\protect\hyperlink{ref-WHO-Covid}{1}}. During the
-first months of the pandemic in early 2020, non-pharmaceutical
-interventions such as masking and social distancing were the only
-methods available to manage the spread of the disease; however, the
-rapid development of vaccines permitted their approval and use in some
-countries towards the last month part of 2020. For example, in the US
-and Canada vaccine campaigns began in mid-December of
-2020\textsuperscript{\protect\hyperlink{ref-tanne2020}{2},\protect\hyperlink{ref-bogoch2022}{3}}.
-Although it has been estimated that COVID-19 vaccines have prevented
-around 14 millions of deaths
-worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{4}}, the
-rollout of COVID-19 vaccines has faced multiple challenges since its
-inception.
-
-Indeed, multiple obstacles that have complicated vaccination efforts
-against COVID-19 have been identified: inequality in vaccine access,
-vaccine hesitancy, and differences in vaccination rates across different
-segments of the
-population\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{5}--\protect\hyperlink{ref-malik2020}{7}}.
-In the case of Canada, lower vaccine uptake has been associated with
-socio-economic factors such as younger age, educational level, presence
-of children in the household, lack of a regular healthcare provider,
-ethnic origin, and financial
-instability\textsuperscript{\protect\hyperlink{ref-guay2022}{8}--\protect\hyperlink{ref-hussain2022}{10}}.
-
-Additionally, vaccination is influenced by changes in geography. In this
-regard, it has been shown that there have been spatial differences in
-COVID-19 vaccination rates due to regional differences in attitudes
-towards
-vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{7}},
-geographical differences in vaccine access and supply, vaccination
-location availability, and lack of prioritization of vulnerable
-groups\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3},\protect\hyperlink{ref-nguyen2021}{11}}.
-
-Studies that analyze geographical variations in vaccine uptake are
-important as they can help inform public health decision-makers to
-design policies to that consider spatial changes to address vaccination
-disparities. In this regard, previous geographical (spatial) analyses of
-vaccination rates have shown that variations in vaccine uptake can occur
-within small governmental administrative units (e.g., counties in the
-case of the
-US)\textsuperscript{\protect\hyperlink{ref-mollalo2021}{12}--\protect\hyperlink{ref-bhuiyan2022}{15}},
-and that geographical analyses can be predictive of booster uptake
-patterns\textsuperscript{\protect\hyperlink{ref-wood2022}{16}}.
-
-In Canada, studies that have used a spatial approach to analyze vaccine
-uptake have shown disparities in vaccination rates across low and high
-income neighborhoods in the city of
-Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{17}}, among
-adolescents from deprived neighborhoods in the city of
-Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{18}}, and
-have also highlighted disparities in vaccination status depending on
-age, income, and ethnic origin at the provincial
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{8}}. However,
-there is limited information on differences in vaccination status inside
-the provinces. Such analysis is important as it can help identify
-inequalities that may exist within these geographical areas while
-providing a granular view of intra-provincial differences that can help
-understand the barriers for vaccine uptake.
-
-In this study, we examined self-reported COVID-19 vaccination status
-within the province of Ontario using a combination of socio-economic
-factors(such as ethnic origin, age, and income) and geographical
-analysis (at the level of the Health Regions) in order to determine
-intra-provincial differences in vaccine uptake and address the ongoing
-need of socio-economic information that can provide a rationale for the
-disparities in vaccination observed within some racial
-groups\textsuperscript{\protect\hyperlink{ref-cnat2022b}{19}}.
+The vaccines against COVID-19 have been considered a major achievement
+of modern medicine as their rapid
+development\textsuperscript{\protect\hyperlink{ref-davis2022}{1}}
+allowed the start of broad vaccination campaigns towards the end of 2020
+in certain countries such as the
+US\textsuperscript{\protect\hyperlink{ref-tanne2020}{2}} and
+Canada\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3}}. This made
+some believe that vaccines were destined to be a determinant factor in a
+rapid ending of the
+pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{4}}.
+However, although it has been estimated that COVID-19 vaccines have
+prevented around 14 million of deaths
+worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{5}}, their
+implementation has been far from being equal to that of the vaccines of
+smallpox and polio, which were implemented on a global scale and that
+were indeed crucial to control these
+diseases\textsuperscript{\protect\hyperlink{ref-kayser2021}{6}}. In
+fact, the rollout of COVID-19 vaccines has faced multiple challenges
+since its inception which ultimately have hampered the goal of global
+immunity.
+
+This ongoing lack of a successful global rollout of the COVID-19
+vaccines is a multifaceted issue resulting from, as among other things,
+the development of new variants due to inadequate public health
+measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
+in vaccine access between high- and low-income
+countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{8},\protect\hyperlink{ref-tamey2022}{9}},
+vaccine
+hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{10}}, and
+differences in vaccination uptake across different segments of the
+population\textsuperscript{\protect\hyperlink{ref-malik2020}{11}}. In
+particular, it is well established that differences in vaccination
+uptake have been present even in the case of countries that have had
+ample access to vaccines since 2020 (such as the US, the UK, and
+Canada), where lower vaccine uptake has been observed within racial
+minorities (i.e., individuals that identify as Black, Asian, or
+Indigenous), and in individuals within low income
+brackets\textsuperscript{\protect\hyperlink{ref-willis2021}{12}--\protect\hyperlink{ref-khubchandani2021}{15}}.
+Reasons given for lower vaccine uptake in these cases have included
+medical mistrust due to systemic medical racism, mistrust in vaccines,
+and the influence of conspiracy
+theories\textsuperscript{\protect\hyperlink{ref-willis2021}{12},\protect\hyperlink{ref-stoler2021}{14},\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
+More specifically, in the case of Canada, lower vaccine uptake has been
+observed in young individuals, those with a low educational level,
+households with children, those without a regular healthcare provider,
+individuals that identify as part of a visible minorities or Indigenous,
+and those that are financially
+unstable\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.
+
+However, it is important to consider that vaccination uptake can also be
+influenced by geographical (spatial) factors. In this regard,
+differences in COVID-19 vaccination rates have been associated with
+varied regional attitudes towards
+vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{11}},
+spatial differences in vaccine access and supply, vaccination location
+availability, and lack of prioritization of areas where vulnerable
+groups
+reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3},\protect\hyperlink{ref-nguyen2021}{22}}.
+Other studies have also shown heterogeneity in vaccine uptake within
+small governmental administrative units such as
+counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{23}--\protect\hyperlink{ref-bhuiyan2022}{26}},
+and that and that accounting for geographical differences in vaccination
+can help predict patterns of booster
+uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{27}}. Overall,
+the existing body of research on geographical/spatial differences in
+vaccination has provided valuable information that can be used by
+decision-makers in the development of public health policies that are
+focused on improving vaccination within specific administrative or
+geographical areas. However, spatial analyses of vaccination have been
+very limited in Canada, as existing studies have been focused only at
+certain cities (such as
+Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}}, or
+Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}), or
+have explored differences at a province-wide
+level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. However,
+there is a lack of studies that explore spatial heterogeneities in
+vaccination at the intra-provincial level.
+
+Ontario is the most populated province in Canada, and has recently seen
+major changes on its healthcare structure as between 2006 and 2019, it
+provided healthcare access using 14 intra-provincial divisions called
+the Local Health Integrated Networks (LHINs). However, this approach was
+complex and caused systemic
+inequalities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. In
+late 2019, the 14 LHINs were phased out and the areas they covered were
+incorporated into 6 Health Regions (North East, North West, Central,
+Toronto, West, and East) in an effort to improve the healthcare system
+of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
+Because the adoption of the Health Regions has been recent, there is an
+ongoing need to analyze the impact of their adoption and intra-regional
+differences that might exist, which is specially important in the
+context of the COVID-19 pandemic.
+
+To address this need, in this study we analyzed differences in
+self-reported COVID-19 vaccination status using socio-economic (e.g.,
+income, racial/ethnic identification), and spatial information at the
+level of the Health Regions, to identify the existence of
+intra-provincial differences that might need to be addressed to ensure
+that the healthcare system of Ontario is more inclusive and that
+responds to the needs of its most vulnerable population.
 
@@
 
-\hypertarget{data-source-survey-overview}{%
-\subsection{Data source: survey
-overview}\label{data-source-survey-overview}}
+\hypertarget{sec-data}{%
+\subsection{Data}\label{sec-data}}
 
@@
 Attitudes}, a repeated cross sectional survey focused on the Canadian
-province of Ontario which ran from Sept 30, 2021 until January 17, 2022
-and that was comissioned by the Fields Institute for Research in
-Mathematical Sciences (henceforth Fields) and the Mathematical Modelling
-of COVID-19 Task Force. The survey was conducted by a third-party
-service provider (RIWI Corp.), under ethical guidance from the
-University of Toronto.
-
-Briefly, the survey was deployed used random domain intercept
-technology, where if users clicked on a registered but commercially
-inactive web link or typed in a web address for a site that was dormant,
-they had a random chance of that link being temporarily managed by the
-company that administered the survey and instead of coming across a
-notification about the status of the site(``this page does not exist''),
-the survey was deployed to the
-user\textsuperscript{\protect\hyperlink{ref-sargent2022}{20}}. Users
-then decided whether to anonymously participate, and those that
-participated were able to exit the survey at any time. After the survey
-closed, regardless if it was complete or incomplete, access was denied
-to any further users with the same internet protocol (IP) address and
-the domain entry point rotated such that if a user were to attempt to
-access the survey again, share the link, or enter via the same address
-using an alternative IP address, the survey would not deploy. This
-effectively meant that a user could participate only once in the survey.
-
-Additionally, users who indicated they were under the age of 16 were
-exited from the survey without creating a record, furthermore, these
-users were unable to navigate back to the ``age select'' screen. The
-personal identifier information from each user that participated in the
-survey was automatically scrubbed and replaced by a unique ID.
-Respondents were drawn exclusively from the province of Ontario, as per
-their devices meta-data.
-
-\hypertarget{sec-socio-demographic-factors}{%
-\subsection{Survey responses}\label{sec-socio-demographic-factors}}
-
-Socio-economic information selected from the survey answers included age
-group and income brackets, and race/ethnicity. The levels of each
-socio-economic factor used for analysis appear in
-Table~\ref{tbl-covariates}.
+province of Ontario that was commissioned by the Fields Institute for
+Research in Mathematical Sciences (henceforth Fields) and the
+Mathematical Modelling of COVID-19 Task Force under ethical guidance
+from the University of Toronto. The survey collected socio-economic
+information from participants (Table~\ref{tbl-covariates}), recorded
+their location (using the nearest municipality), and asked information
+on vaccination status by using the question ``Have you received the
+first dose of the COVID vaccine?'', with possible answers ``yes'' and
+``no''. The original dataset contained 39,029 entries (where each entry
+corresponded to a unique respondent).
+
+This dataset was cleaned to remove outliers that were identified during
+preliminary analyses, and processing the geographical information in the
+survey (city where the survey was responded) in order to match each city
+to its correspondent Health Region.
 
@@
   >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6986}}@{}}
-\caption{\label{tbl-covariates}Selected socio-economic factors from the
-survey}\tabularnewline
+\caption{\label{tbl-covariates}Socio-economic factors from the Fields
+COVID-19 survey}\tabularnewline
 \toprule()
@@
 
-Furthermore, the information on vaccination status was provided by
-survey participants who answered the question ``Have you received the
-first dose of the COVID vaccine?'', with possible answers ``yes'' and
-``no''.
-
-\hypertarget{data-cleaning}{%
-\subsection{Data cleaning}\label{data-cleaning}}
-
-The original dataset contained 39,029 entries (where each entry
-corresponded to a set of answers provided by a unique respondent).
-Following a preliminary analysis, the dataset was cleaned in order to
-only contain the socio-economic information provided in
-Table~\ref{tbl-covariates} and vaccination status. The cleaning process
-also included removing outliers that were identified during the
-preliminary analyses, and processing the geographical information in the
-survey (city where the survey was responded) in order to match each city
-to its correspondent Health Region.
-
-\hypertarget{sec-geographical-location}{%
-\subsubsection{Geographical location}\label{sec-geographical-location}}
-
-For each survey participant certain data was automatically captured,
-including the nearest municipality (city). This resulted in a total of
-578 different municipalities within the clean dataset. Due to our
-interest in analyzing the differences between Health Regions, we
-assigned the city of each entry to its correspondent Health Region
-following a multi-step process. Briefly, we used Local Health Integrated
-Networks (LHINs) to assign a Health Region to each entry in the survey.
-LHINs were the geographical divisions for health used by Ontario before
-the adoption of the Health Regions; because of the lack of a publicly
-available list of all municipalities within each Health Region, we used
-a dataset of long-term care homes and LHINs to match each city to LHIN,
-followed by matching each LHIN to a Health Region following the
-information provided on the Ontario Health Website, where the list of
-LHINs and corresponding Health Regions is available. In the case of
-municipalities that did not appear in the long-term dataset, we manually
-searched each city in the LHINs websites in order to provide
-geographical information. The raw dataset, clean dataset, details of the
-data cleaning process, and the addition of Health Region and LHIN
-information are described in detail in the GitHub repository for this
-paper, which can be found at
+The clean dataset contained responses from more than 200 different
+municipalities within Ontario (Figure~\ref{fig-map}). Because of the
+lack of a publicly available list of all municipalities within each
+Health Region, we used a dataset of long-term care homes and LHINs to
+match each city to LHIN, followed by matching each LHIN to a Health
+Region following the information provided on the Ontario Health Website,
+where the list of LHINs and corresponding Health Regions is available.
+In the case of municipalities that did not appear in the long-term care
+home dataset, we manually searched each city in the LHINs websites in
+order to provide geographical information. The original dataset, clean
+dataset, and details on the data cleaning process are described in
+detail in the GitHub repository for this paper, which can be found at
 \url{https://github.com/aimundo/Fields_COVID-19/}.
@@
 total) corresponded to cities located in the North West and North East
-Health Regions. Due to the low number of entries, we omitted these
-Health Regions from further analyses. Therefore, the total number of
-unique entries used for analysis was 3,551 which included the East,
+Health Regions. The low representation of these Health Regions in the
+dataset is noticeable in Figure~\ref{fig-map}, where the number of
+cities in each Health Region is relatively low compared to the most
+populated Health Regions, such as the Toronto or Central Regions. We
+omitted the North East and North East Health Regions from further
+analyses due to the low number of entries. Therefore, the total number
+of unique entries used for analysis was 3,551 which included the East,
 Central, Toronto, and West Health Regions.
 
-\hypertarget{corrections}{%
-\subsection{Corrections}\label{corrections}}
-
-We identified differences between the proportions of all the
-socio-economic factors included in the analysis
-(Table~\ref{tbl-covariates}) and the 2016 Canada Census data for
-Ontario. Additionally, because the Census divisions do not match the
-exact boundaries of the Health Regions, we also obtained population
-estimates for each Health Region from the Ontario Health website in
-order to correct for the population totals. We used an iterative
-proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{21}}
-to correct for socio-economic factors and Health Region populations
-using the R \texttt{survey} package. Details about the correction can be
-found in the Appendix.
+\begin{figure}
+
+{\centering \includegraphics{../data/map_data/map.pdf}
+
+}
+
+\caption{\label{fig-map}Geographic representation of the data collected
+by the \emph{Survey of COVID-19 related Behaviours and Attitudes},
+collected by the Fields Institute in Ontario. The municipalities
+(cities) from where survey participants provided answers (in the clean
+dataset) appear as points. The Health six Regions are color-coded.
+Internal boundaries within certain Health Regions indicate areas that
+belonged to the Local Integrated Health Networks (LHINs), the geographic
+areas for healthcare in Ontario before the adoption of the Health
+Regions.}
+
+\end{figure}
 
@@
 vaccination depending on the socio-economic factors described in
-Section~\ref{sec-socio-demographic-factors}, the Health Regions from
-Section~\ref{sec-geographical-location}, and the interactions between
-Race and Health Region, and Race and income as previous studies have
-shown that socio-economic factors and their interactions are significant
-predictors of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{22}--\protect\hyperlink{ref-cnat2022a}{24}}.
-The model appears in Equation~\ref{eq-model3},
-
-\begin{equation}\protect\hypertarget{eq-model3}{}{
-\begin{aligned}
-\log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Race)} + \beta_6 \textrm{ (Income} \times \textrm{Race)}
-\end{aligned}
-}\label{eq-model3}\end{equation}
-
-Where \(p\textrm{(vac)}\) indicates the probability of having received
-the first dose of a COVID-19 vaccine, \(\beta_0\) indicates the
-population intercept, and \(\beta_1...\beta_6\) indicate the
-coefficients for each term. The the model was fitted using the function
-\texttt{svyglm} from the \texttt{survey} R package in order to
-incorporate the correction in sampling probability obtained from raking.
-
-All analyses were conducted in RStudio (2022.12.0 Build 353), using R
-4.2.2 and the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{25}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{26}},
-and \texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{27}}.
+Table~\ref{tbl-covariates}, the Health Regions from Ontario indicated in
+Section~\ref{sec-data}, and the interactions between Race and Health
+Region, and Race and income, as previous studies have shown that
+socio-economic factors and their interactions are significant predictors
+of intent of vaccination and vaccination
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
+
+The model was fitted first to the clean dataset to obtain uncorrected
+estimates. Additionally, because we identified differences between the
+proportions of all the socio-economic factors included in the analysis
+(Table~\ref{tbl-covariates}) and the Census data for Ontario, we used an
+iterative proportional fitting procedure
+(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{35}}
+to correct the data using Census socio-economic data and Health Region
+population totals, in order to obtain corrected estimates from the
+model. Details regarding the correction can be found in the Appendix.
+All analyses were conducted in R 4.2.2 using the packages
+\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
+and \texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}}.
 
@@
 
-Table~\ref{tbl-model} shows the results of the logistic regression on
-vaccination status using socio-economic factors (age group, income,
-race), geographical areas (Health Regions) and the interactions between
-income and race and Health Region and race. The reference groups were
-set as follows: 16 to 34 years (age group), White Caucasian (Race),
-Toronto (Health Region), CAD 60,000 and over (Income). There were no
-statistically significant differences in vaccination rates within the
-age groups from the survey, but significant odds ratios were estimated
-for other covariates. Within household income brackets, individuals with
-an income under CAD 25,000 or between CAD 25,000-59,999 had
-significantly lower odds of vaccination than those with an income above
-CAD 60,000 (ORs=0.37 and 0.59, respectively). Within Race/Ethnicity,
+Table~\ref{tbl-model} shows the results of the logistic regression
+models (for the uncorrected and corrected data) on vaccination status
+using socio-economic factors (age group, income, race), geographical
+areas (Health Regions) and the interactions between income and race and
+Health Region and race. There were no statistically significant
+differences in vaccination rates within the age groups from the survey,
+but significant odds ratios were estimated for other covariates. Within
+household income brackets, individuals with an income under CAD 25,000
+or between CAD 25,000-59,999 had significantly lower odds of vaccination
+than those with an income above CAD 60,000 (ORs=0.37 and 0.59,
+\emph{p}=0.011 and \textless0.001, respectively). Within Race/Ethnicity,
 individuals who identified as Arab/Middle Eastern, Black, or Latin
@@
 \hypertarget{tbl-model}{}
-\begin{longtable}{lccc}
+\begin{longtable}{lcccccc}
 \caption{\label{tbl-model}Multiple Regression Analysis-Predictors of Vaccination Status }\tabularnewline
@@
 \toprule
-\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{\textbf{Uncorrected}} & \multicolumn{3}{c}{\textbf{Corrected}} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
 \midrule
 \endfirsthead
-\multicolumn{4}{@{}l}{\textit{(continued)}}\\
+\multicolumn{7}{@{}l}{\textit{(continued)}}\\
 \toprule
-\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{\textbf{Uncorrected}} & \multicolumn{3}{c}{\textbf{Corrected}} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
+\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
 \midrule
@@
 \endlastfoot
-\textbf{Age Group} &  &  & \\
-\hspace{1em}16\_34 & — & — & \\
-\hspace{1em}35\_54 & 0.90 & 0.67, 1.21 & 0.5\\
-\hspace{1em}55\_and\_over & 0.99 & 0.74, 1.32 & >0.9\\
-\textbf{Income} &  &  & \\
-\hspace{1em}60000\_and\_above & — & — & \\
-\hspace{1em}25000\_59999 & 0.59 & 0.39, 0.89 & 0.011\\
-\hspace{1em}under\_25000 & 0.37 & 0.25, 0.56 & <0.001\\
-\textbf{Race} &  &  & \\
-\hspace{1em}white\_caucasian & — & — & \\
-\hspace{1em}arab\_middle\_eastern & 0.31 & 0.14, 0.69 & 0.004\\
-\hspace{1em}black & 0.32 & 0.17, 0.60 & <0.001\\
-\hspace{1em}east\_asian\_pacific\_islander & 1.15 & 0.50, 2.66 & 0.7\\
-\hspace{1em}indigenous & 0.44 & 0.19, 1.02 & 0.056\\
-\hspace{1em}latin\_american & 0.28 & 0.11, 0.67 & 0.004\\
-\hspace{1em}mixed & 0.64 & 0.25, 1.65 & 0.4\\
-\hspace{1em}other & 0.22 & 0.12, 0.41 & <0.001\\
-\hspace{1em}south\_asian & 0.91 & 0.49, 1.69 & 0.8\\
-\textbf{Health Region} &  &  & \\
-\hspace{1em}Toronto & — & — & \\
-\hspace{1em}Central & 1.47 & 0.92, 2.35 & 0.11\\
-\hspace{1em}East & 1.42 & 0.90, 2.23 & 0.13\\
-\hspace{1em}West & 1.55 & 1.05, 2.30 & 0.029\\
-\textbf{Income * Race} &  &  & \\
-\hspace{1em}25000\_59999 * arab\_middle\_eastern & 1.79 & 0.67, 4.83 & 0.2\\
-\hspace{1em}under\_25000 * arab\_middle\_eastern & 3.05 & 1.26, 7.39 & 0.013\\
-\hspace{1em}25000\_59999 * black & 1.34 & 0.59, 3.05 & 0.5\\
-\hspace{1em}under\_25000 * black & 3.19 & 1.45, 6.99 & 0.004\\
-\hspace{1em}25000\_59999 * east\_asian\_pacific\_islander & 0.42 & 0.17, 1.05 & 0.062\\
-\hspace{1em}under\_25000 * east\_asian\_pacific\_islander & 1.16 & 0.47, 2.86 & 0.8\\
-\hspace{1em}25000\_59999 * indigenous & 1.36 & 0.48, 3.89 & 0.6\\
-\hspace{1em}under\_25000 * indigenous & 1.45 & 0.55, 3.80 & 0.5\\
-\hspace{1em}25000\_59999 * latin\_american & 1.24 & 0.45, 3.43 & 0.7\\
-\hspace{1em}under\_25000 * latin\_american & 2.80 & 1.04, 7.51 & 0.041\\
-\hspace{1em}25000\_59999 * mixed & 0.85 & 0.32, 2.26 & 0.7\\
-\hspace{1em}under\_25000 * mixed & 1.10 & 0.37, 3.27 & 0.9\\
-\hspace{1em}25000\_59999 * other & 6.93 & 2.65, 18.1 & <0.001\\
-\hspace{1em}under\_25000 * other & 4.59 & 2.33, 9.05 & <0.001\\
-\hspace{1em}25000\_59999 * south\_asian & 1.20 & 0.51, 2.85 & 0.7\\
-\hspace{1em}under\_25000 * south\_asian & 2.00 & 0.93, 4.30 & 0.077\\
-\textbf{Race * Health Region} &  &  & \\
-\hspace{1em}arab\_middle\_eastern * Central & 0.66 & 0.26, 1.70 & 0.4\\
-\hspace{1em}black * Central & 0.44 & 0.19, 0.98 & 0.046\\
-\hspace{1em}east\_asian\_pacific\_islander * Central & 0.98 & 0.38, 2.53 & >0.9\\
-\hspace{1em}indigenous * Central & 0.63 & 0.22, 1.79 & 0.4\\
-\hspace{1em}latin\_american * Central & 0.67 & 0.23, 1.96 & 0.5\\
-\hspace{1em}mixed * Central & 0.73 & 0.24, 2.22 & 0.6\\
-\hspace{1em}other * Central & 0.80 & 0.36, 1.78 & 0.6\\
-\hspace{1em}south\_asian * Central & 0.54 & 0.25, 1.20 & 0.13\\
-\hspace{1em}arab\_middle\_eastern * East & 0.43 & 0.13, 1.45 & 0.2\\
-\hspace{1em}black * East & 0.83 & 0.34, 2.04 & 0.7\\
-\hspace{1em}east\_asian\_pacific\_islander * East & 0.86 & 0.29, 2.56 & 0.8\\
-\hspace{1em}indigenous * East & 0.69 & 0.23, 2.08 & 0.5\\
-\hspace{1em}latin\_american * East & 1.03 & 0.32, 3.34 & >0.9\\
-\hspace{1em}mixed * East & 0.91 & 0.28, 3.03 & 0.9\\
-\hspace{1em}other * East & 1.05 & 0.39, 2.83 & >0.9\\
-\hspace{1em}south\_asian * East & 0.52 & 0.19, 1.45 & 0.2\\
-\hspace{1em}arab\_middle\_eastern * West & 1.00 & 0.37, 2.73 & >0.9\\
-\hspace{1em}black * West & 0.76 & 0.32, 1.80 & 0.5\\
-\hspace{1em}east\_asian\_pacific\_islander * West & 0.52 & 0.20, 1.34 & 0.2\\
-\hspace{1em}indigenous * West & 0.39 & 0.14, 1.09 & 0.073\\
-\hspace{1em}latin\_american * West & 0.94 & 0.32, 2.72 & >0.9\\
-\hspace{1em}mixed * West & 0.37 & 0.12, 1.16 & 0.089\\
-\hspace{1em}other * West & 0.41 & 0.18, 0.93 & 0.032\\
-\hspace{1em}south\_asian * West & 0.41 & 0.18, 0.95 & 0.037\\*
-\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\
+\textbf{Age Group} &  &  &  &  &  & \\
+\hspace{1em}16\_34 & — & — &  & — & — & \\
+\hspace{1em}35\_54 & 0.93 & 0.77, 1.13 & 0.5 & 0.90 & 0.67, 1.21 & 0.5\\
+\hspace{1em}55\_and\_over & 0.74 & 0.61, 0.89 & 0.002 & 0.99 & 0.74, 1.32 & >0.9\\
+\textbf{Income} &  &  &  &  &  & \\
+\hspace{1em}60000\_and\_above & — & — &  & — & — & \\
+\hspace{1em}25000\_59999 & 0.59 & 0.41, 0.84 & 0.004 & 0.59 & 0.39, 0.89 & 0.011\\
+\hspace{1em}under\_25000 & 0.41 & 0.29, 0.58 & <0.001 & 0.37 & 0.25, 0.56 & <0.001\\
+\textbf{Race} &  &  &  &  &  & \\
+\hspace{1em}white\_caucasian & — & — &  & — & — & \\
+\hspace{1em}arab\_middle\_eastern & 0.24 & 0.12, 0.50 & <0.001 & 0.31 & 0.14, 0.69 & 0.004\\
+\hspace{1em}black & 0.30 & 0.17, 0.54 & <0.001 & 0.32 & 0.17, 0.60 & <0.001\\
+\hspace{1em}east\_asian\_pacific\_islander & 0.74 & 0.36, 1.52 & 0.4 & 1.15 & 0.50, 2.66 & 0.7\\
+\hspace{1em}indigenous & 0.37 & 0.17, 0.81 & 0.013 & 0.44 & 0.19, 1.02 & 0.056\\
+\hspace{1em}latin\_american & 0.28 & 0.13, 0.59 & <0.001 & 0.28 & 0.11, 0.67 & 0.004\\
+\hspace{1em}mixed & 0.59 & 0.31, 1.12 & 0.11 & 0.64 & 0.25, 1.65 & 0.4\\
+\hspace{1em}other & 0.20 & 0.11, 0.35 & <0.001 & 0.22 & 0.12, 0.41 & <0.001\\
+\hspace{1em}south\_asian & 0.80 & 0.44, 1.45 & 0.5 & 0.91 & 0.49, 1.69 & 0.8\\
+\textbf{Health Region} &  &  &  &  &  & \\
+\hspace{1em}Toronto & — & — &  & — & — & \\
+\hspace{1em}Central & 1.30 & 0.85, 2.00 & 0.2 & 1.47 & 0.92, 2.35 & 0.11\\
+\hspace{1em}East & 1.54 & 1.01, 2.34 & 0.044 & 1.42 & 0.90, 2.23 & 0.13\\
+\hspace{1em}West & 1.35 & 0.95, 1.94 & 0.10 & 1.55 & 1.05, 2.30 & 0.029\\
+\textbf{Income * Race} &  &  &  &  &  & \\
+\hspace{1em}25000\_59999 * arab\_middle\_eastern & 2.16 & 0.93, 4.99 & 0.072 & 1.79 & 0.67, 4.83 & 0.2\\
+\hspace{1em}under\_25000 * arab\_middle\_eastern & 2.96 & 1.39, 6.26 & 0.005 & 3.05 & 1.26, 7.39 & 0.013\\
+\hspace{1em}25000\_59999 * black & 1.19 & 0.60, 2.39 & 0.6 & 1.34 & 0.59, 3.05 & 0.5\\
+\hspace{1em}under\_25000 * black & 2.88 & 1.48, 5.59 & 0.002 & 3.19 & 1.45, 6.99 & 0.004\\
+\hspace{1em}25000\_59999 * east\_asian\_pacific\_islander & 0.95 & 0.44, 2.07 & >0.9 & 0.42 & 0.17, 1.05 & 0.062\\
+\hspace{1em}under\_25000 * east\_asian\_pacific\_islander & 1.91 & 0.90, 4.04 & 0.090 & 1.16 & 0.47, 2.86 & 0.8\\
+\hspace{1em}25000\_59999 * indigenous & 1.81 & 0.74, 4.43 & 0.2 & 1.36 & 0.48, 3.89 & 0.6\\
+\hspace{1em}under\_25000 * indigenous & 1.64 & 0.73, 3.69 & 0.2 & 1.45 & 0.55, 3.80 & 0.5\\
+\hspace{1em}25000\_59999 * latin\_american & 0.89 & 0.38, 2.10 & 0.8 & 1.24 & 0.45, 3.43 & 0.7\\
+\hspace{1em}under\_25000 * latin\_american & 3.09 & 1.33, 7.16 & 0.009 & 2.80 & 1.04, 7.51 & 0.041\\
+\hspace{1em}25000\_59999 * mixed & 0.86 & 0.39, 1.93 & 0.7 & 0.85 & 0.32, 2.26 & 0.7\\
+\hspace{1em}under\_25000 * mixed & 1.26 & 0.64, 2.47 & 0.5 & 1.10 & 0.37, 3.27 & 0.9\\
+\hspace{1em}25000\_59999 * other & 5.46 & 2.41, 12.3 & <0.001 & 6.93 & 2.65, 18.1 & <0.001\\
+\hspace{1em}under\_25000 * other & 4.06 & 2.25, 7.31 & <0.001 & 4.59 & 2.33, 9.05 & <0.001\\
+\hspace{1em}25000\_59999 * south\_asian & 1.13 & 0.54, 2.36 & 0.7 & 1.20 & 0.51, 2.85 & 0.7\\
+\hspace{1em}under\_25000 * south\_asian & 1.59 & 0.83, 3.06 & 0.2 & 2.00 & 0.93, 4.30 & 0.077\\
+\textbf{Race * Health Region} &  &  &  &  &  & \\
+\hspace{1em}arab\_middle\_eastern * Central & 0.75 & 0.33, 1.71 & 0.5 & 0.66 & 0.26, 1.70 & 0.4\\
+\hspace{1em}black * Central & 0.48 & 0.23, 1.01 & 0.055 & 0.44 & 0.19, 0.98 & 0.046\\
+\hspace{1em}east\_asian\_pacific\_islander * Central & 0.84 & 0.37, 1.88 & 0.7 & 0.98 & 0.38, 2.53 & >0.9\\
+\hspace{1em}indigenous * Central & 0.60 & 0.24, 1.51 & 0.3 & 0.63 & 0.22, 1.79 & 0.4\\
+\hspace{1em}latin\_american * Central & 0.69 & 0.28, 1.72 & 0.4 & 0.67 & 0.23, 1.96 & 0.5\\
+\hspace{1em}mixed * Central & 0.63 & 0.29, 1.35 & 0.2 & 0.73 & 0.24, 2.22 & 0.6\\
+\hspace{1em}other * Central & 0.96 & 0.48, 1.92 & >0.9 & 0.80 & 0.36, 1.78 & 0.6\\
+\hspace{1em}south\_asian * Central & 0.70 & 0.34, 1.45 & 0.3 & 0.54 & 0.25, 1.20 & 0.13\\
+\hspace{1em}arab\_middle\_eastern * East & 0.58 & 0.20, 1.65 & 0.3 & 0.43 & 0.13, 1.45 & 0.2\\
+\hspace{1em}black * East & 0.82 & 0.36, 1.86 & 0.6 & 0.83 & 0.34, 2.04 & 0.7\\
+\hspace{1em}east\_asian\_pacific\_islander * East & 0.82 & 0.30, 2.24 & 0.7 & 0.86 & 0.29, 2.56 & 0.8\\
+\hspace{1em}indigenous * East & 0.55 & 0.21, 1.44 & 0.2 & 0.69 & 0.23, 2.08 & 0.5\\
+\hspace{1em}latin\_american * East & 0.79 & 0.28, 2.23 & 0.7 & 1.03 & 0.32, 3.34 & >0.9\\
+\hspace{1em}mixed * East & 0.71 & 0.31, 1.60 & 0.4 & 0.91 & 0.28, 3.03 & 0.9\\
+\hspace{1em}other * East & 0.86 & 0.37, 1.99 & 0.7 & 1.05 & 0.39, 2.83 & >0.9\\
+\hspace{1em}south\_asian * East & 0.50 & 0.20, 1.24 & 0.14 & 0.52 & 0.19, 1.45 & 0.2\\
+\hspace{1em}arab\_middle\_eastern * West & 1.16 & 0.48, 2.77 & 0.7 & 1.00 & 0.37, 2.73 & >0.9\\
+\hspace{1em}black * West & 0.77 & 0.36, 1.65 & 0.5 & 0.76 & 0.32, 1.80 & 0.5\\
+\hspace{1em}east\_asian\_pacific\_islander * West & 0.54 & 0.24, 1.20 & 0.13 & 0.52 & 0.20, 1.34 & 0.2\\
+\hspace{1em}indigenous * West & 0.44 & 0.19, 1.02 & 0.056 & 0.39 & 0.14, 1.09 & 0.073\\
+\hspace{1em}latin\_american * West & 0.97 & 0.39, 2.43 & >0.9 & 0.94 & 0.32, 2.72 & >0.9\\
+\hspace{1em}mixed * West & 0.53 & 0.25, 1.11 & 0.092 & 0.37 & 0.12, 1.16 & 0.089\\
+\hspace{1em}other * West & 0.55 & 0.27, 1.12 & 0.10 & 0.41 & 0.18, 0.93 & 0.032\\
+\hspace{1em}south\_asian * West & 0.50 & 0.23, 1.07 & 0.075 & 0.41 & 0.18, 0.95 & 0.037\\*
+\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\
 \end{longtable}
@@
 
-The rapid development of COVID-19 vaccines has been considered as a
-major achievement of modern
-medicine\textsuperscript{\protect\hyperlink{ref-davis2022}{28}}. Vaccine
-availability towards the end of 2020 in certain countries made some
-believe that they would be a determinant factor in a rapid ending of the
-pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{29}}.
-However, despite previous successful vaccination campaigns that were
-crucial to control diseases such as smallpox and
-polio\textsuperscript{\protect\hyperlink{ref-kayser2021}{30}},
-vaccination efforts in the case of COVID-19 have faced multiple
-challenges that have complicated the achievement of global immunity.
-
-Among the different challenges faced by COVID-19 vaccination efforts are
-the development of new variants due to inadequate public health
-measures\textsuperscript{\protect\hyperlink{ref-li2021}{31}} and
-inequity in vaccine access between low and high income
-countries\textsuperscript{\protect\hyperlink{ref-tamey2022}{32}}.
-However, it is also well established that even in the case of high
-income countries that have had ample access to vaccines since 2020, such
-as the US, the UK, and Canada, there have been challenges in vaccination
-efforts due to differences in vaccine uptake among different segments of
-the population. More specifically, lower vaccine uptake has been
-associated with socio-economic factors such as race (i.e., identifying
-as Black, Asian, Indigenous) and household income (typically within
-lower income
-brackets)\textsuperscript{\protect\hyperlink{ref-willis2021}{33}--\protect\hyperlink{ref-khubchandani2021}{36}}.
-Reasons given for this association have included medical mistrust due to
-systemic medical racism, mistrust in vaccines, and the influence of
-conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-willis2021}{33},\protect\hyperlink{ref-stoler2021}{35},\protect\hyperlink{ref-bogart2021}{37}--\protect\hyperlink{ref-freeman2020}{39}}.
-
-In addition, vaccine uptake is influenced by geography, as shown by
-different studies that have identified intra-regional differences in
-vaccine
-uptake\textsuperscript{\protect\hyperlink{ref-mollalo2021}{12},\protect\hyperlink{ref-pallathadka2022}{40},\protect\hyperlink{ref-huang2022}{41}}.
-However, in the case of Canada, studies that have analyzed spatial
-differences in vaccination have been focused in country-wide or
-province-wide
-estimates\textsuperscript{\protect\hyperlink{ref-guay2022}{8},\protect\hyperlink{ref-lavoie2022}{42}}.
-Therefore, we explored spatial and socio-economic determinants of
-vaccination status in the province of Ontario. This province is of
-particular interest as it has seen recently major structural health
-changes with the dissolution of the Local Health Integrated Network
-(LHIN) system and the incorporation of regions covered by LHINs into
-larger Health
-Regions\textsuperscript{\protect\hyperlink{ref-dong2022}{43}}. Because
-the idea behind the change aimed to reduce the inequalities in
-healthcare that were identified under the LHIN
-model\textsuperscript{\protect\hyperlink{ref-tsasis2012}{44}}, examining
-differences in vaccination between the Health Regions can provide
-decision-makers with insight regarding intra-provincial health
-disparities that may need to be addressed in future vaccination or
-public health campaigns.
+The change from LHINs to Health Regions for healthcare in Ontario was
+aimed at reducing healthcare
+disparities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. The
+relatively recent nature of the change, which occurred in late 2019,
+highlights the importance of studies that can determine if there are
+intra-regional differences in healthcare that need to be addressed,
+particularly in the context of the COVID-19 pandemic.
 
@@
 other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{45}}, with
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{39}}, with
 vaccination uptake rates across different age groups presented in other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{8},\protect\hyperlink{ref-macdonald2021}{46}},
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{40}},
 and with the vaccination information provided by Public Health Ontario,
@@
 coverage among all the age groups considered in the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{47}}.
+survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}.
 
@@
 low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{8}--\protect\hyperlink{ref-hussain2022}{10},\protect\hyperlink{ref-carter2022}{48}}.
+income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{42}}.
 
@@
 as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{49},\protect\hyperlink{ref-ct2021}{50}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{43},\protect\hyperlink{ref-ct2021}{44}},
 which include occupations such as grocery store workers, gas station
@@
 had priority for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{51}},
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{45}},
 which would explain the higher odds of vaccination for these individuals
@@
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{52}}.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}.
 In the case of this analysis, the estimated lower odds are likely to be
@@
 population (more than 20\%) in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{53}}
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{47}}
 (Supplementary Table A-7). Therefore, this result partly reinforces the
@@
 previously reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{54}}; in this
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{48}}; in this
 regard, our result provides additional context by showing that within
@@
 future to obtain a more accurate approximation to the population totals
-of the province. Additionally, the results in this study are based on
-self-reported data, where the risk of bias exist. Despite this, because
-in the context of COVID-19 it has been shown that good agreement exists
-between self-reported and documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{55}}, the
+of the province. Moreover, our analysis did not consider the North West
+and North East Health Regions, due to the low number of entries from
+these areas in the survey (Figure~\ref{fig-map}). Although low
+representation from these areas is based on the fact that these regions
+only account for 5\% of the total population of Ontario, these regions
+are the home to more than 100,000 individuals that identify as
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}, a
+minority group that has historically suffered from reduced access to
+health care and
+discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{17}}.
+Therefore, there is a need for additional studies that focus on these
+low-populated Health Regions in Ontario where disparities in vaccination
+might be significant and understudied.
+
+The results in this study are based on self-reported data, where the
+risk of bias exist. Despite this, because in the context of COVID-19 it
+has been shown that good agreement exists between self-reported and
+documented vaccination
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{49}}, the
 effect of self-reported bias is likely to not be significant in this
-case.
-
-Finally, it is likely that there have been differences in vaccination
-across the province as more doses of the vaccine were administered and
-as successive variants emerged. Because this study focused only on
-vaccination status regarding the first dose of the vaccine within a
-relatively short time window, it can only provide a snapshot of the
-societal dynamics behind the pandemic. Nonetheless, the results
-presented here can serve as a starting point to motivate future
+case. Finally, it is likely that there have been differences in
+vaccination across the province as more doses of the vaccine were
+administered and as successive variants emerged. Because this study
+focused only on vaccination status regarding the first dose of the
+vaccine within a relatively short time window, it can only provide a
+snapshot of the societal dynamics behind the pandemic. Nonetheless, the
+results presented here can serve as a starting point to motivate future
 longitudinal research that aims to quantify geographical differences
@@
 \begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-WHO-Covid}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
 \CSLLeftMargin{1. }%
-\CSLRightInline{{World Health Organization Coronavirus (COVID-19)
-Dashboard}. Accessed November 27, 2022. \url{https://covid19.who.int/}}
+\CSLRightInline{Davis CJ, Golding M, McKay R. Efficacy information
+influences intention to take COVID-19 vaccine. \emph{British Journal of
+Health Psychology}. 2022;27(2):300-319.
+doi:\url{https://doi.org/10.1111/bjhp.12546}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
+end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
+doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-watson2022}{}}%
-\CSLLeftMargin{4. }%
+\CSLLeftMargin{5. }%
 \CSLRightInline{Watson OJ, Barnsley G, Toor J, Hogan AB, Winskill P,
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
+\CSLLeftMargin{6. }%
+\CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
+and emerging issues. \emph{Human Vaccines {\&}amp\(\mathsemicolon\)
+Immunotherapeutics}. 2021;17(12):5255-5268.
+doi:\href{https://doi.org/10.1080/21645515.2021.1977057}{10.1080/21645515.2021.1977057}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-li2021}{}}%
+\CSLLeftMargin{7. }%
+\CSLRightInline{Li Q, Wang J, Tang Y, Lu H. Next-generation COVID-19
+vaccines: Opportunities for vaccine development and challenges in
+tackling COVID-19. \emph{Drug Discoveries \& Therapeutics}.
+2021;15(3):118-123.
+doi:\href{https://doi.org/10.5582/ddt.2021.0105}{10.5582/ddt.2021.0105}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
-\CSLLeftMargin{5. }%
+\CSLLeftMargin{8. }%
 \CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
+\CSLLeftMargin{9. }%
+\CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
+to achieve global covid-19 vaccine equity. \emph{{BMJ}}. Published
+online March 2022:e070650.
+doi:\href{https://doi.org/10.1136/bmj-2022-070650}{10.1136/bmj-2022-070650}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
-\CSLLeftMargin{6. }%
+\CSLLeftMargin{10. }%
 \CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
-\CSLLeftMargin{7. }%
+\CSLLeftMargin{11. }%
 \CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
+\CSLLeftMargin{12. }%
+\CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
+{COVID}-19 vaccine hesitancy: Race/ethnicity, trust, and fear.
+\emph{Clinical and Translational Science}. 2021;14(6):2200-2207.
+doi:\href{https://doi.org/10.1111/cts.13077}{10.1111/cts.13077}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
+\CSLLeftMargin{13. }%
+\CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
+accepting {COVID}-19 vaccination during and after pregnancy, and for
+their babies: A multi-methods study in the {UK}. \emph{{BMC} Pregnancy
+and Childbirth}. 2022;22(1).
+doi:\href{https://doi.org/10.1186/s12884-021-04321-3}{10.1186/s12884-021-04321-3}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
+\CSLLeftMargin{14. }%
+\CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
+limits of medical trust in mitigating {COVID}-19 vaccine hesitancy among
+black americans. \emph{Journal of General Internal Medicine}.
+2021;36(11):3629-3631.
+doi:\href{https://doi.org/10.1007/s11606-021-06743-3}{10.1007/s11606-021-06743-3}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
+\CSLLeftMargin{15. }%
+\CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
+Sharma M, Webb FJ. {COVID}-19 vaccination hesitancy in the united
+states: A rapid national assessment. \emph{Journal of Community Health}.
+2021;46(2):270-277.
+doi:\href{https://doi.org/10.1007/s10900-020-00958-x}{10.1007/s10900-020-00958-x}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
+\CSLLeftMargin{16. }%
+\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
+related medical mistrust, health impacts, and potential vaccine
+hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
+of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
+doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
+\CSLLeftMargin{17. }%
+\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
+roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
+canada. \emph{Canadian Medical Association Journal}.
+2021;193(11):E381-E383.
+doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
+\CSLLeftMargin{18. }%
+\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
+hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
+and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
+2020;52(14):3127-3141.
+doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
-\CSLLeftMargin{8. }%
+\CSLLeftMargin{19. }%
 \CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
-\CSLLeftMargin{9. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
-\CSLLeftMargin{10. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
-\CSLLeftMargin{11. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
-\CSLLeftMargin{16. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
@@
 \leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
-\CSLLeftMargin{19. }%
-\CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
-review on vaccine hesitancy in black communities in canada: Critical
-issues and research failures. \emph{Vaccines}. 2022;10(11):1937.
-doi:\href{https://doi.org/10.3390/vaccines10111937}{10.3390/vaccines10111937}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sargent2022}{}}%
-\CSLLeftMargin{20. }%
-\CSLRightInline{Sargent RH, Laurie S, Weakland LF, et al. Use of random
-domain intercept technology to track {COVID}-19 vaccination rates in
-real time across the united states: Survey study. \emph{Journal of
-Medical Internet Research}. 2022;24(7):e37920.
-doi:\href{https://doi.org/10.2196/37920}{10.2196/37920}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{21. }%
-\CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
-a sampled frequency table when the expected marginal totals are known.
-\emph{The Annals of Mathematical Statistics}. 1940;11(4):427-444.
-doi:\href{https://doi.org/10.1214/aoms/1177731829}{10.1214/aoms/1177731829}}
+\leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
+\CSLLeftMargin{30. }%
+\CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
+integrated care: A complex-adaptive systems perspective.
+\emph{International Journal of Integrated Care}. 2012;12(5).
+doi:\href{https://doi.org/10.5334/ijic.843}{10.5334/ijic.843}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
+\CSLLeftMargin{31. }%
+\CSLRightInline{Dong L, Sahu R, Black R. Governance in the
+transformational journey toward integrated healthcare: The case of
+ontario. \emph{Journal of Information Technology Teaching Cases}.
+Published online December 2022:204388692211473.
+doi:\href{https://doi.org/10.1177/20438869221147313}{10.1177/20438869221147313}}
 
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
+\CSLLeftMargin{35. }%
+\CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
+a sampled frequency table when the expected marginal totals are known.
+\emph{The Annals of Mathematical Statistics}. 1940;11(4):427-444.
+doi:\href{https://doi.org/10.1214/aoms/1177731829}{10.1214/aoms/1177731829}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
-\CSLLeftMargin{28. }%
-\CSLRightInline{Davis CJ, Golding M, McKay R. Efficacy information
-influences intention to take COVID-19 vaccine. \emph{British Journal of
-Health Psychology}. 2022;27(2):300-319.
-doi:\url{https://doi.org/10.1111/bjhp.12546}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
-\CSLLeftMargin{29. }%
-\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
-end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
-doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
-\CSLLeftMargin{30. }%
-\CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
-and emerging issues. \emph{Human Vaccines {\&}amp\(\mathsemicolon\)
-Immunotherapeutics}. 2021;17(12):5255-5268.
-doi:\href{https://doi.org/10.1080/21645515.2021.1977057}{10.1080/21645515.2021.1977057}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-li2021}{}}%
-\CSLLeftMargin{31. }%
-\CSLRightInline{Li Q, Wang J, Tang Y, Lu H. Next-generation COVID-19
-vaccines: Opportunities for vaccine development and challenges in
-tackling COVID-19. \emph{Drug Discoveries \& Therapeutics}.
-2021;15(3):118-123.
-doi:\href{https://doi.org/10.5582/ddt.2021.0105}{10.5582/ddt.2021.0105}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
-\CSLLeftMargin{32. }%
-\CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
-to achieve global covid-19 vaccine equity. \emph{{BMJ}}. Published
-online March 2022:e070650.
-doi:\href{https://doi.org/10.1136/bmj-2022-070650}{10.1136/bmj-2022-070650}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
-\CSLLeftMargin{33. }%
-\CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
-{COVID}-19 vaccine hesitancy: Race/ethnicity, trust, and fear.
-\emph{Clinical and Translational Science}. 2021;14(6):2200-2207.
-doi:\href{https://doi.org/10.1111/cts.13077}{10.1111/cts.13077}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
-\CSLLeftMargin{34. }%
-\CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
-accepting {COVID}-19 vaccination during and after pregnancy, and for
-their babies: A multi-methods study in the {UK}. \emph{{BMC} Pregnancy
-and Childbirth}. 2022;22(1).
-doi:\href{https://doi.org/10.1186/s12884-021-04321-3}{10.1186/s12884-021-04321-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
-\CSLLeftMargin{35. }%
-\CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
-limits of medical trust in mitigating {COVID}-19 vaccine hesitancy among
-black americans. \emph{Journal of General Internal Medicine}.
-2021;36(11):3629-3631.
-doi:\href{https://doi.org/10.1007/s11606-021-06743-3}{10.1007/s11606-021-06743-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
-\CSLLeftMargin{36. }%
-\CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
-Sharma M, Webb FJ. {COVID}-19 vaccination hesitancy in the united
-states: A rapid national assessment. \emph{Journal of Community Health}.
-2021;46(2):270-277.
-doi:\href{https://doi.org/10.1007/s10900-020-00958-x}{10.1007/s10900-020-00958-x}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
-\CSLLeftMargin{37. }%
-\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
-related medical mistrust, health impacts, and potential vaccine
-hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
-of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
-doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
-\CSLLeftMargin{38. }%
-\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
-roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
-canada. \emph{Canadian Medical Association Journal}.
-2021;193(11):E381-E383.
-doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
 \CSLLeftMargin{39. }%
-\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
-hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
-and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
-2020;52(14):3127-3141.
-doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-pallathadka2022}{}}%
-\CSLLeftMargin{40. }%
-\CSLRightInline{Pallathadka A, Chang H, Han D. What explains spatial
-variations of {COVID}-19 vaccine hesitancy?: A
-social-ecological-technological systems approach. \emph{Environmental
-Research: Health}. 2022;1(1):011001.
-doi:\href{https://doi.org/10.1088/2752-5309/ac8ac2}{10.1088/2752-5309/ac8ac2}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-huang2022}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{Huang Q, Cutter SL. Spatial-temporal differences of
-{COVID}-19 vaccinations in the u.s. \emph{Urban Informatics}. 2022;1(1).
-doi:\href{https://doi.org/10.1007/s44212-022-00019-9}{10.1007/s44212-022-00019-9}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lavoie2022}{}}%
-\CSLLeftMargin{42. }%
-\CSLRightInline{Lavoie K, Gosselin-Boucher V, Stojanovic J, et al.
-Understanding national trends in {COVID}-19 vaccine hesitancy in canada:
-Results from five sequential cross-sectional representative surveys
-spanning april 2020{\textendash}march 2021. \emph{{BMJ} Open}.
-2022;12(4):e059411.
-doi:\href{https://doi.org/10.1136/bmjopen-2021-059411}{10.1136/bmjopen-2021-059411}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
-\CSLLeftMargin{43. }%
-\CSLRightInline{Dong L, Sahu R, Black R. Governance in the
-transformational journey toward integrated healthcare: The case of
-ontario. \emph{Journal of Information Technology Teaching Cases}.
-Published online December 2022:204388692211473.
-doi:\href{https://doi.org/10.1177/20438869221147313}{10.1177/20438869221147313}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
-\CSLLeftMargin{44. }%
-\CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
-integrated care: A complex-adaptive systems perspective.
-\emph{International Journal of Integrated Care}. 2012;12(5).
-doi:\href{https://doi.org/10.5334/ijic.843}{10.5334/ijic.843}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{45. }%
 \CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-03-22 code/age_analysis.R

@@ -1,0 +1,32 @@
+clean_data %>%
+  tbl_summary(percent="row",
+              by= first_dose,
+              include=c(income_ord,
+                        Age_group,
+                        Health_Region,
+                        Race),
+              label=list(Health_Region ~ "Health Region",
+                         Age_group ~ "Age Group",
+                         income_ord="Income")
+  ) %>% bold_labels()
+
+clean_data$test<-as.factor(clean_data$age)
+
+clean_data %>%
+  tbl_summary(percent="row",
+              by= first_dose,
+              include=test)%>% 
+  bold_labels()
+
+ggplot(data=clean_data,aes(x=age,y=first_dose_m,group=age))+geom_bar(stat="identity")
+
+
+test2 <- clean_data %>% 
+  group_by(age) %>% 
+  count(first_dose_m) %>% 
+  mutate(pct = n / sum(n)) 
+
+# Plot 
+ggplot(test2, aes(x = age, y = pct, fill = age)) +
+  geom_col(position = "dodge") +
+  facet_wrap(~ age) 

### 2023-03-22 code/data_preparation_map.R

@@ -1,0 +1,29 @@
+## data preparation for map, including the North West and North East Health Regions
+
+# this script takes the clean dataset generated from the data cleaning process and does some preparatory steps such as 
+# variable recoding, creating a variable for the dependent variable in the regression, and creating 
+# a summary table of the dataset
+
+data<-read.csv(here("data","clean_dataset.csv"))
+
+# remove rows with missing observations for race
+
+clean_data<-data%>%drop_na(race)
+
+#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
+
+clean_data<-clean_data %>%
+  select(-Municipal.status) %>%
+  rename(Race=race,Age_group=age_group,Income=income_ord)
+
+#make variables of interest factors
+clean_data<-clean_data %<>%
+  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)
+
+# recode the response of interest for categorical analysis
+clean_data<-clean_data %>%
+  mutate(first_dose_m=case_when(
+    first_dose=="yes"~1,
+    first_dose=="no"~0
+  ))
+

### 2023-03-22 code/map.R

@@ -1,0 +1,176 @@
+#stand-alone script to generate a map of Ontario with the cities included in the survey
+
+###########################################################################
+# IMPORTANT: The shapefiles used to create the map are too large to the uploaded to GitHub. Therefore, information
+# on where the data can be obtained is provided below. The user needs to download the files
+# from the links provided, unzip them, and place them in a directory where they can be read.
+# In this case, the files were placed in the "map_data" folder within the "data" directory.
+
+# Shapefile for LHINs
+  ## Description for the dataset from the Ontario website:
+  ## https://data.ontario.ca/dataset/ontario-s-health-region-geographic-data/resource/9fe4e91a-adff-4a34-bbe6-d43d596e4867
+  ## the data can be downloaded from:
+  # https://www150.statcan.gc.ca/n1/en/pub/82-402-x/2018001/data-donnees/boundary-limites/arcinfo/HRL_035b18a-eng.zip?st=QxbZnLnU
+
+
+# Shapefile for cities 
+  ## Description for the dataset:
+  ## https://natural-resources.canada.ca/earth-sciences/geography/download-geographical-names-data/9245
+  ##The data can be downloaded from:
+  # https://ftp.cartes.canada.ca/pub/nrcan_rncan/vector/geobase_cgn_toponyme/prov_shp_eng/cgn_on_shp_eng.zip
+
+library(rgdal)
+library(sf)
+library(viridis)
+library(here)
+library(tidyverse)
+library(ggspatial)
+library(magrittr)
+library(scico)
+
+thm1<-scale_fill_viridis_d(option="mako",begin=0,end=1,direction = 1)
+
+thm2<-scale_fill_scico_d(palette = "nuuk",begin=0,end=1,direction = -1)
+
+mytheme<-theme(axis.line=element_blank(),
+               axis.text.x=element_blank(),
+               axis.text.y=element_blank(),
+               axis.ticks=element_blank(),
+               axis.title.x=element_blank(),
+               axis.title.y=element_blank(),
+               legend.position="right",
+               panel.background=element_blank(),
+               panel.border=element_blank(),
+               panel.grid.major=element_blank(),
+               panel.grid.minor=element_blank(),
+               plot.background=element_blank())
+
+
+
+# use sf to read lhin data
[…]

### 2023-03-22 code/regression.R

@@ -1,11 +1,13 @@
 ## This script uses the object created in the raking.R script to fit a fixed-effects logistic regression
 
+#model for uncorrected data
+m0<-svyglm(first_dose_m~Age_group_ord+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
+           design=a1,
+           family = quasibinomial(), 
+           control= list(maxit=25))
 
 
-#call the script for raking
-source(here::here("code","raking.R"))
-
-# logistic regression model
+# model for corrected data
 
 m1<-svyglm(first_dose_m~Age_group_ord+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
            design=a1_rake,

### 2023-03-23 manuscript_v2/01-background.qmd
code for age analysis

@@
 
-The vaccines against COVID-19 have been considered a major achievement of modern medicine as their rapid development[@davis2022] allowed the start of broad vaccination campaigns towards the end of 2020 in certain countries such as the US[@tanne2020] and Canada[@bogoch2022].
+The vaccines against COVID-19 have been considered a major achievement of modern medicine as their rapid development allowed the start of broad vaccination campaigns towards the end of 2020 in certain countries, such as the US and Canada [@davis2022;@bogoch2022;@tanne2020].
 This made some believe that vaccines were destined to be a determinant factor in a rapid ending of the pandemic [@thelancet2021].
 However, although it has been estimated that COVID-19 vaccines have prevented around 14 million of deaths worldwide [@watson2022], their implementation has been far from being equal to that of the vaccines of smallpox and polio, which were implemented on a global scale and that were indeed crucial to control these diseases [@kayser2021].
-In fact, the rollout of COVID-19 vaccines has faced multiple challenges since its inception which ultimately have hampered the goal of global immunity.
+In fact, the rollout of COVID-19 vaccines has faced multiple challenges since its inception which ultimately have hampered their use to achieve the ultimate goal of global immunity.
 
-This ongoing lack of a successful global rollout of the COVID-19 vaccines is a multifaceted issue resulting from, as among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020].
-In particular, it is well established that differences in vaccination uptake have been present even in the case of countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within racial minorities (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021].
-Reasons given for lower vaccine uptake in these cases have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020].
-More specifically, in the case of Canada, lower vaccine uptake has been observed in young individuals, those with a low educational level, households with children, those without a regular healthcare provider, individuals that identify as part of a visible minorities or Indigenous, and those that are financially unstable [@guay2022;@muhajarine2021;@hussain2022].
+This problematic of COVID-19 vaccines rollout is a multifaceted issue resulting from, among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020].
+In particular, it is well established that differences in vaccination uptake have been present even in countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within racial minorities (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021].
+Reasons given for lower vaccine uptake in these cases have included medical mistrust due to systemic medical racism [@stoler2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
+Moreover, in the case of Canada, lower vaccine uptake has been observed in young individuals, those with a low educational level, households with children, those without a regular healthcare provider, individuals that identify as part of a visible minorities or Indigenous, and those that are financially unstable [@guay2022;@muhajarine2021;@hussain2022].
 
@@
 Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that and that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022].
-Overall, the existing body of  research on geographical/spatial differences in vaccination has provided valuable information that can be used by decision-makers in the development of public health policies that are focused on improving vaccination within specific administrative or geographical areas.
-However, spatial analyses of vaccination have been very limited in Canada, as existing studies have been focused only at certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022].
-However, there is a lack of studies that explore spatial heterogeneities in vaccination at the intra-provincial level.
+Overall, the evidence provided by the literature demonstrates the existence of spatially-driven heterogeneities in vaccine uptake that be used by decision-makers in the development of public health policies that are focused on addressing these disparities within specific administrative or geographical areas.
 
-Ontario is the most populated province in Canada, and has recently seen major changes on its healthcare structure as between 2006 and 2019, it provided healthcare access using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs).
-However, this approach was complex and caused systemic inequalities [@tsasis2012].
-In late 2019, the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East) in an effort to improve the healthcare system of the province[@dong2022].
-Because the adoption of the Health Regions has been recent, there is an ongoing need to analyze the impact of their adoption and intra-regional differences that might exist, which is specially important in the context of the COVID-19 pandemic.
+However, such analyses have been carried mostly in territories outside of Canada, where available studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022].
+Thus, there is a need for studies that explore spatial differences in vaccination within the Canadian territory and that consequently, can help identify disparities that need to be addressed within specific areas in each province.
 
-To address this need, in this study we analyzed differences in self-reported COVID-19 vaccination status using socio-economic (e.g., income, racial/ethnic identification), and spatial  information at the level of the Health Regions, to identify the existence of intra-provincial differences that might need to be addressed to ensure that the healthcare system of Ontario is more inclusive and that responds to the needs of its most vulnerable population.
+This need is particularly important in the case of Ontario, the most populated province in Canada.
+Between 2006 and 2019, Ontario provided healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs).
+However, this approach was complex, bureaucratic, and led to systemic inequalities [@tsasis2012].
+In late 2019, the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East) in an effort to improve the healthcare system of the province [@dong2022].
+Because the adoption of the Health Regions occurred at a relatively recent time, there is an ongoing need to analyze the impact of this measure and identify the existence of intra-regional differences that might exist, and which could be specially important in the context of the COVID-19 pandemic.
+
+In this study, we analyzed differences in self-reported COVID-19 vaccination status in Ontario using socio-economic (e.g., income, racial/ethnic identification), and spatial  information at the level of the Health Regions to identify the existence of differences that might need to be addressed to ensure that the healthcare system of Ontario is more inclusive and that responds to the needs of its most vulnerable population.

### 2023-03-23 manuscript_v2/02-methods.qmd
code for age analysis

@@
 Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions.
-The low representation of these Health Regions in the dataset is noticeable in @fig-map, where the number of cities in each Health Region is relatively low compared to the most populated Health Regions, such as the Toronto or Central Regions.
+The low representation of these Health Regions in the dataset is noticeable in @fig-map,which shows that responses from these areas came from a relatively low number of cities when compared to the most populated Health Regions, such as the Toronto or Central Regions.
 We omitted the North East and North East Health Regions from further analyses due to the low number of entries.
@@
 
-<!-- $$ -->
-<!-- \begin{aligned} -->
-<!-- \log \left( \frac{p\textrm{(vac)}}{1-p\textrm{(vac)}} \right) = \beta_0+ \beta_{1}\textrm{(Age group)} +\beta_{2} \textrm{ Race} + \beta_3 \textrm{ Health Region} + \beta_4 \textrm{ Income}+\\ \beta_5\textrm{(Health Region} \times \textrm{Race)} + \beta_6 \textrm{ (Income} \times \textrm{Race)} -->
-<!-- \end{aligned} -->
-<!-- $$ {#eq-model3} -->
-
-<!-- Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a COVID-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term.
--->
-
 The model was fitted first to the clean dataset to obtain uncorrected estimates.

### 2023-03-23 manuscript_v2/03-results.qmd
code for age analysis

@@
 
-@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed.
-The total number of entries from the survey in the dataset after cleaning was 3,551.
+@tbl-descriptive-stats shows the descriptive statistics from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed.
+The total number of entries analyzed was 3,551.
 Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it.

### 2023-03-23 manuscript_v2/main.tex
code for age analysis

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 The vaccines against COVID-19 have been considered a major achievement
-of modern medicine as their rapid
-development\textsuperscript{\protect\hyperlink{ref-davis2022}{1}}
-allowed the start of broad vaccination campaigns towards the end of 2020
-in certain countries such as the
-US\textsuperscript{\protect\hyperlink{ref-tanne2020}{2}} and
-Canada\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3}}. This made
-some believe that vaccines were destined to be a determinant factor in a
-rapid ending of the
+of modern medicine as their rapid development allowed the start of broad
+vaccination campaigns towards the end of 2020 in certain countries, such
+as the US and
+Canada\textsuperscript{\protect\hyperlink{ref-davis2022}{1}--\protect\hyperlink{ref-tanne2020}{3}}.
+This made some believe that vaccines were destined to be a determinant
+factor in a rapid ending of the
 pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{4}}.
@@
 fact, the rollout of COVID-19 vaccines has faced multiple challenges
-since its inception which ultimately have hampered the goal of global
-immunity.
-
-This ongoing lack of a successful global rollout of the COVID-19
-vaccines is a multifaceted issue resulting from, as among other things,
-the development of new variants due to inadequate public health
+since its inception which ultimately have hampered their use to achieve
+the ultimate goal of global immunity.
+
+This problematic of COVID-19 vaccines rollout is a multifaceted issue
+resulting from, among other things, the development of new variants due
+to inadequate public health
 measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
@@
 particular, it is well established that differences in vaccination
-uptake have been present even in the case of countries that have had
-ample access to vaccines since 2020 (such as the US, the UK, and
-Canada), where lower vaccine uptake has been observed within racial
-minorities (i.e., individuals that identify as Black, Asian, or
-Indigenous), and in individuals within low income
+uptake have been present even in countries that have had ample access to
+vaccines since 2020 (such as the US, the UK, and Canada), where lower
+vaccine uptake has been observed within racial minorities (i.e.,
+individuals that identify as Black, Asian, or Indigenous), and in
+individuals within low income
 brackets\textsuperscript{\protect\hyperlink{ref-willis2021}{12}--\protect\hyperlink{ref-khubchandani2021}{15}}.
 Reasons given for lower vaccine uptake in these cases have included
-medical mistrust due to systemic medical racism, mistrust in vaccines,
-and the influence of conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-willis2021}{12},\protect\hyperlink{ref-stoler2021}{14},\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
-More specifically, in the case of Canada, lower vaccine uptake has been
-observed in young individuals, those with a low educational level,
-households with children, those without a regular healthcare provider,
-individuals that identify as part of a visible minorities or Indigenous,
-and those that are financially
+medical mistrust due to systemic medical
+racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{14}}, mistrust
+in vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{12}}, and
+the influence of conspiracy
+theories\textsuperscript{\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
+Moreover, in the case of Canada, lower vaccine uptake has been observed
+in young individuals, those with a low educational level, households
+with children, those without a regular healthcare provider, individuals
+that identify as part of a visible minorities or Indigenous, and those
+that are financially
 unstable\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.
@@
 groups
-reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{3},\protect\hyperlink{ref-nguyen2021}{22}}.
+reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2},\protect\hyperlink{ref-nguyen2021}{22}}.
 Other studies have also shown heterogeneity in vaccine uptake within
@@
 uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{27}}. Overall,
-the existing body of research on geographical/spatial differences in
-vaccination has provided valuable information that can be used by
+the evidence provided by the literature demonstrates the existence of
+spatially-driven heterogeneities in vaccine uptake that be used by
 decision-makers in the development of public health policies that are
-focused on improving vaccination within specific administrative or
-geographical areas. However, spatial analyses of vaccination have been
-very limited in Canada, as existing studies have been focused only at
-certain cities (such as
-Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}}, or
-Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}), or
-have explored differences at a province-wide
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. However,
-there is a lack of studies that explore spatial heterogeneities in
-vaccination at the intra-provincial level.
-
-Ontario is the most populated province in Canada, and has recently seen
-major changes on its healthcare structure as between 2006 and 2019, it
-provided healthcare access using 14 intra-provincial divisions called
-the Local Health Integrated Networks (LHINs). However, this approach was
-complex and caused systemic
+focused on addressing these disparities within specific administrative
+or geographical areas.
+
+However, such analyses have been carried mostly in territories outside
+of Canada, where available studies have been focused in certain cities
+(such as Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}},
+or Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}),
+or have explored differences at a province-wide
+level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. Thus, there
+is a need for studies that explore spatial differences in vaccination
+within the Canadian territory and that consequently, can help identify
+disparities that need to be addressed within specific areas in each
+province.
+
+This need is particularly important in the case of Ontario, the most
+populated province in Canada. Between 2006 and 2019, Ontario provided
+healthcare access to its inhabitants using 14 intra-provincial divisions
+called the Local Health Integrated Networks (LHINs). However, this
+approach was complex, bureaucratic, and led to systemic
 inequalities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. In
@@
 of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
-Because the adoption of the Health Regions has been recent, there is an
-ongoing need to analyze the impact of their adoption and intra-regional
-differences that might exist, which is specially important in the
-context of the COVID-19 pandemic.
-
-To address this need, in this study we analyzed differences in
-self-reported COVID-19 vaccination status using socio-economic (e.g.,
-income, racial/ethnic identification), and spatial information at the
-level of the Health Regions, to identify the existence of
-intra-provincial differences that might need to be addressed to ensure
-that the healthcare system of Ontario is more inclusive and that
-responds to the needs of its most vulnerable population.
+Because the adoption of the Health Regions occurred at a relatively
+recent time, there is an ongoing need to analyze the impact of this
+measure and identify the existence of intra-regional differences that
+might exist, and which could be specially important in the context of
+the COVID-19 pandemic.
+
+In this study, we analyzed differences in self-reported COVID-19
+vaccination status in Ontario using socio-economic (e.g., income,
+racial/ethnic identification), and spatial information at the level of
+the Health Regions to identify the existence of differences that might
+need to be addressed to ensure that the healthcare system of Ontario is
+more inclusive and that responds to the needs of its most vulnerable
+population.
 
@@
 Health Regions. The low representation of these Health Regions in the
-dataset is noticeable in Figure~\ref{fig-map}, where the number of
-cities in each Health Region is relatively low compared to the most
-populated Health Regions, such as the Toronto or Central Regions. We
-omitted the North East and North East Health Regions from further
-analyses due to the low number of entries. Therefore, the total number
-of unique entries used for analysis was 3,551 which included the East,
-Central, Toronto, and West Health Regions.
+dataset is noticeable in Figure~\ref{fig-map},which shows that responses
+from these areas came from a relatively low number of cities when
+compared to the most populated Health Regions, such as the Toronto or
+Central Regions. We omitted the North East and North East Health Regions
+from further analyses due to the low number of entries. Therefore, the
+total number of unique entries used for analysis was 3,551 which
+included the East, Central, Toronto, and West Health Regions.
 
@@
 
-Table~\ref{tbl-descriptive-stats} shows the descriptive statistics
-(uncorrected) from the Fields COVID-19 survey data for vaccination
-status and each of the covariates analyzed. The total number of entries
-from the survey in the dataset after cleaning was 3,551. Overall, 26.9\%
-of survey respondents (958) reported not having received the first dose
-of the vaccine, whereas 73.1\% (2,593) reported having received it.
-Within each socio-economic factor, respondents who reported living in a
-household with an income under CAD 25,000 represented 37\% of the total
-number of entries, those within the CAD 25,000-59,999 income bracket
-represented 25\% of the total sample, and those with an income above CAD
-60,000 represented 38 \% of the sample; across all income brackets, the
-percentage of individuals that reported having received a first dose of
-the vaccine was consistent, above 69\%.
+Table~\ref{tbl-descriptive-stats} shows the descriptive statistics from
+the Fields COVID-19 survey data for vaccination status and each of the
+covariates analyzed. The total number of entries analyzed was 3,551.
+Overall, 26.9\% of survey respondents (958) reported not having received
+the first dose of the vaccine, whereas 73.1\% (2,593) reported having
+received it. Within each socio-economic factor, respondents who reported
+living in a household with an income under CAD 25,000 represented 37\%
+of the total number of entries, those within the CAD 25,000-59,999
+income bracket represented 25\% of the total sample, and those with an
+income above CAD 60,000 represented 38 \% of the sample; across all
+income brackets, the percentage of individuals that reported having
+received a first dose of the vaccine was consistent, above 69\%.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
 \CSLLeftMargin{2. }%
-\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
-pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
-2020:m4799.
-doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
-\CSLLeftMargin{3. }%
 \CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
@@
 doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
+pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
+2020:m4799.
+doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}
 

### 2023-03-23 code/age_analysis.R

@@ -1,3 +1,5 @@
+library(patchwork)
+
 clean_data %>%
   tbl_summary(percent="row",
               by= first_dose,
@@ -10,6 +12,7 @@ clean_data %>%
                          income_ord="Income")
   ) %>% bold_labels()
 
+
 clean_data$test<-as.factor(clean_data$age)
 
 clean_data %>%
@@ -18,15 +21,42 @@ clean_data %>%
               include=test)%>% 
   bold_labels()
 
-ggplot(data=clean_data,aes(x=age,y=first_dose_m,group=age))+geom_bar(stat="identity")
+##important: according to the data dictionary, the age of those >75 was
+## coded as "98"
+ggplot(data=clean_data,
+       aes(x=age,y=first_dose_m,group=age))+
+  geom_bar(stat="identity")
 
 
 test2 <- clean_data %>% 
-  group_by(age) %>% 
+  group_by(age,Age_group,Age_group_ord) %>% 
   count(first_dose_m) %>% 
   mutate(pct = n / sum(n)) 
 
 # Plot 
-ggplot(test2, aes(x = age, y = pct, fill = age)) +
-  geom_col(position = "dodge") +
-  facet_wrap(~ age) 
+a<-test2 %>%
+  subset(first_dose_m==1 & Age_group_ord=="16_34")%>%
+ggplot(aes(x = age, y = pct, fill = age)) +
+  geom_smooth(show.legend = FALSE)+
+ geom_point(show.legend = FALSE)+
+  facet_wrap(~Age_group_ord)
+
+b<-test2 %>%
+  subset(first_dose_m==1 & Age_group_ord=="35_54")%>%
+  ggplot(aes(x = age, y = pct, fill = age)) +
+  geom_smooth(show.legend = FALSE)+
+  geom_point(show.legend = FALSE)+
+  facet_wrap(~Age_group_ord)
+
+c<-test2 %>%
+  subset(first_dose_m==1 & Age_group_ord=="55_and_over")%>%
+  ggplot(aes(x = age, y = pct, fill = age)) +
+  geom_smooth(show.legend = FALSE)+
+  geom_point(show.legend = FALSE)+
+  facet_wrap(~Age_group_ord)
+
+a+b+c
+
+
+## try to get percentage of people that said yes per age group and plot it
+

### 2023-03-24 manuscript_v2/02-methods.qmd
updating discussion

@@
 
-We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto.
+We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto, and which ran between September 30th, 2021 and January 17th,2022.
 The survey collected socio-economic information from participants (@tbl-covariates), recorded their location (using the nearest municipality), and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".
@@
 We omitted the North East and North East Health Regions from further analyses due to the low number of entries.
-Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions.
+Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions and that covered the period between September 30th,2021 and December 12th, 2021.
 

### 2023-03-24 manuscript_v2/04-discussion.qmd
updating discussion

@@
 
-The change from LHINs to Health Regions for healthcare in Ontario was aimed at reducing healthcare disparities [@tsasis2012].
-The relatively recent nature of the change, which occurred in late 2019, highlights the importance of studies that can determine if there are intra-regional differences in healthcare that need to be addressed, particularly in the context of the COVID-19 pandemic.
+There existence of healthcare disparities within Ontario is a topic of particular interest in due to the recent change in the healthcare system of this region, which eliminated the LHIN model and adopted larger Health Regions system in late 2019 in an attempt to address the disparities of the former approach [@tsasis2012;@dong2022].
+In this context, analyzing COVID-19 vaccination estimates within the province is important as they can provide an initial assessment of variations that might need to be addressed by decision-makers.
 
 Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats), and that there were no significant differences in vaccination odds among the age groups considered in the survey.
-This result is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022], and with the vaccination information provided by Public Health Ontario,  which shows that for the period where the Fields survey ran (Sept 30, 2021-Jan 17, 2022) there was a minimum of  80% of first dose vaccination coverage among all the age groups considered in the survey [@ontario-covid].
+This result is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022].
+
+It is worth mentioning that the province-wide vaccination rate from the data for those 55 years of age and older, which was 72\%, can appear to be significantly lower from the vaccination rate reported for the same period by the public health agency of Ontario (Public Health Ontario), which reported a rate of 88.4\% at the start of the period covered by the data (September 30th, 2021 until December 12th, 2021) for those 50 years of age and older [@ontario-covid].
+We believe the difference in this case can be explained by the fact that such province-wide estimates mask regional differences, which have been present in vaccination rates in the province since the start.
+This becomes apparent when vaccination rates are disaggregated by the Public Health Units in Ontario.
+For example, the Public Health Unit of Lambton (a region within the West Health Region) and the Public Health Unit of Haliburton, Kawartha, and the Pine Ridge District (an area covered by the East Health Region) both reported vaccination lower rates of 78\% for those 50 years of age and older at the beginning of the period of interest, whereas other regions  [@ontario-covid].
+Moreover, differences in vaccination rates within the province continue as of today (late March of 2023), when, according to Public Health Ontario, some regions have less than 75\% vaccination rates [@ontario-covid-map].
+
+age and older for the period covered by the data used for analysis in this study (late September of 2021 until mid-December of the same year) the vaccination rate provided by Public Health Ontario,  which shows that for the period where the Fields survey ran (Sept 30, 2021-Jan 17, 2022) there was a minimum of  80% of first dose vaccination coverage among all the age groups considered in the survey [@ontario-covid].
 

### 2023-03-24 manuscript_v2/appendix.qmd
updating discussion

@@
 
-clean_data <- read.csv(here("data","clean_dataset.csv"))
-
-# remove rows with missing observations for race to obtain accurate numbers
-
-clean_data<-clean_data %>% drop_na(race)
+#call scripts for data preparation and regression
+source(here::here("code","data_preparation.R"))
+source(here::here("code","raking.R"))
+source(here::here("code","regression.R"))
 ```
@@
   t6<-clean_data%>%
-    select(race)%>%
-    group_by(race)%>%
+    select(Race)%>%
+    group_by(Race)%>%
     summarise(observations = n()) %>%
@@
 p2<-clean_data%>%
-  select(age_group)%>%
-   mutate(age_group_ord = fct_relevel(age_group,
+  select(Age_group)%>%
+   mutate(Age_group_ord = fct_relevel(Age_group,
                                       "16_24",
@@
 
-  ggplot(p2%>%count(age_group_ord) %>%    # G count number in each group
+  ggplot(p2%>%count(Age_group_ord) %>%    # G count number in each group
          mutate(pct=n/sum(n),               # Calculate percent within each group
                 ypos = 200),  # Calculate label positions
-       aes(age_group_ord, n, fill=age_group_ord)) +
+       aes(Age_group_ord, n, fill=Age_group_ord)) +
   geom_bar(stat="identity") +
@@
 
+# Vaccination rates by Age
+
+The table below provides the percentages of vaccination from the survey data.
+
+```{r,vacc-table}
+#| label: tbl-vaccination
+#| tbl-cap: "Vaccination Status by Age from the Survey Data"
+#| message: false
+#| echo: false
+
+clean_data$test<-as.factor(clean_data$age)
+
+tvac<-clean_data %>%
+  tbl_summary(percent="row",
+              by= first_dose,
+              include=test,
+              label=test~"Age")%>%
+  bold_labels()%>%
+as_kable_extra(booktabs = TRUE,linesep="",longtable=T) %>%
+  kableExtra::kable_styling(latex_options = "repeat_header")
+
+tvac
+```
+
 # Interactions from the Regression Model
@@
 #| fig-cap: Estimated marginal means for the combination between Race and Income in the regression model.
-
-#call scripts for data preparation and regression
-source(here::here("code","data_preparation.R"))
-source(here::here("code","regression.R"))
 
@@
     sort = all_categorical() ~ "frequency"
-  ) %>% bold_labels()
+  ) %>% bold_labels()%>%
+  as_kable_extra(booktabs = TRUE,linesep="")
 ```
@@
             sort = all_categorical() ~ "frequency") %>% 
-  bold_labels()
+  bold_labels()%>%
+  as_kable_extra(booktabs = TRUE,linesep="")
 

### 2023-03-24 manuscript_v2/appendix.tex
updating discussion

--- 2023-03-22 manuscript/appendix.tex
@@
 \usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \KOMAoption{captions}{tableheading}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, breakable, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\begin{longtable}{lrr}
+\begin{longtable}{crr}
 \caption{Ethnic information from the clean dataset}\tabularnewline
@@
 \toprule
-race & observations & percentage \\ 
+Race & observations & percentage \\ 
 \midrule
-arab\_middle\_eastern & 218 & $5.9\%$ \\ 
-black & 306 & $8.3\%$ \\ 
-east\_asian\_pacific\_islander & 308 & $8.3\%$ \\ 
-indigenous & 216 & $5.8\%$ \\ 
-latin\_american & 182 & $4.9\%$ \\ 
-mixed & 324 & $8.7\%$ \\ 
-other & 389 & $10.5\%$ \\ 
-south\_asian & 383 & $10.3\%$ \\ 
-white\_caucasian & 1383 & $37.3\%$ \\ 
+white\_caucasian & 1313 & $37.0\%$ \\ 
+arab\_middle\_eastern & 214 & $6.0\%$ \\ 
+black & 298 & $8.4\%$ \\ 
+east\_asian\_pacific\_islander & 303 & $8.5\%$ \\ 
+indigenous & 191 & $5.4\%$ \\ 
+latin\_american & 180 & $5.1\%$ \\ 
+mixed & 310 & $8.7\%$ \\ 
+other & 367 & $10.3\%$ \\ 
+south\_asian & 375 & $10.6\%$ \\ 
 \bottomrule
@@
 
+\hypertarget{vaccination-rates-by-age}{%
+\section{Vaccination rates by Age}\label{vaccination-rates-by-age}}
+
+The table below provides the percentages of vaccination from the survey
+data.
+
+\hypertarget{tbl-vaccination}{}
+\begin{longtable}{lcc}
+\caption{\label{tbl-vaccination}Vaccination Status by Age from the Survey Data }\tabularnewline
+
+\toprule
+\textbf{Characteristic} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,593\\
+\midrule
+\endfirsthead
+\multicolumn{3}{@{}l}{\textit{(continued)}}\\
+\toprule
+\textbf{Characteristic} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,593\\
+\midrule
+\endhead
+
+\endfoot
+\bottomrule
+\endlastfoot
+\textbf{Age} &  & \\
+\hspace{1em}16 & 11 (21\%) & 41 (79\%)\\
+\hspace{1em}17 & 15 (25\%) & 45 (75\%)\\
+\hspace{1em}18 & 53 (36\%) & 93 (64\%)\\
+\hspace{1em}19 & 33 (31\%) & 74 (69\%)\\
+\hspace{1em}20 & 18 (17\%) & 85 (83\%)\\
+\hspace{1em}21 & 30 (25\%) & 92 (75\%)\\
+\hspace{1em}22 & 19 (23\%) & 62 (77\%)\\
+\hspace{1em}23 & 21 (24\%) & 67 (76\%)\\
+\hspace{1em}24 & 23 (29\%) & 55 (71\%)\\
+\hspace{1em}25 & 25 (24\%) & 80 (76\%)\\
+\hspace{1em}26 & 23 (30\%) & 53 (70\%)\\
+\hspace{1em}27 & 16 (25\%) & 49 (75\%)\\
+\hspace{1em}28 & 13 (20\%) & 52 (80\%)\\
+\hspace{1em}29 & 13 (26\%) & 37 (74\%)\\
+\hspace{1em}30 & 12 (21\%) & 44 (79\%)\\
+\hspace{1em}31 & 17 (38\%) & 28 (62\%)\\
+\hspace{1em}32 & 17 (27\%) & 47 (73\%)\\
+\hspace{1em}33 & 17 (26\%) & 48 (74\%)\\
+\hspace{1em}34 & 33 (35\%) & 60 (65\%)\\
+\hspace{1em}35 & 31 (34\%) & 59 (66\%)\\
+\hspace{1em}36 & 16 (25\%) & 49 (75\%)\\
+\hspace{1em}37 & 15 (30\%) & 35 (70\%)\\
+\hspace{1em}38 & 10 (28\%) & 26 (72\%)\\
+\hspace{1em}39 & 5 (16\%) & 26 (84\%)\\
+\hspace{1em}40 & 7 (16\%) & 36 (84\%)\\
+\hspace{1em}41 & 5 (14\%) & 31 (86\%)\\
+\hspace{1em}42 & 7 (19\%) & 30 (81\%)\\
+\hspace{1em}43 & 9 (17\%) & 45 (83\%)\\
+\hspace{1em}44 & 24 (33\%) & 48 (67\%)\\
+\hspace{1em}45 & 23 (30\%) & 53 (70\%)\\
+\hspace{1em}46 & 13 (24\%) & 41 (76\%)\\
+\hspace{1em}47 & 18 (37\%) & 31 (63\%)\\
+\hspace{1em}48 & 10 (33\%) & 20 (67\%)\\
+\hspace{1em}49 & 8 (28\%) & 21 (72\%)\\
+\hspace{1em}50 & 7 (16\%) & 38 (84\%)\\
+\hspace{1em}51 & 7 (23\%) & 23 (77\%)\\
+\hspace{1em}52 & 9 (20\%) & 36 (80\%)\\
+\hspace{1em}53 & 9 (22\%) & 31 (78\%)\\
+\hspace{1em}54 & 19 (37\%) & 33 (63\%)\\
+\hspace{1em}55 & 12 (23\%) & 41 (77\%)\\
+\hspace{1em}56 & 13 (26\%) & 37 (74\%)\\
+\hspace{1em}57 & 12 (22\%) & 43 (78\%)\\
+\hspace{1em}58 & 6 (14\%) & 38 (86\%)\\
+\hspace{1em}59 & 4 (12\%) & 28 (88\%)\\
+\hspace{1em}60 & 3 (11\%) & 25 (89\%)\\
+\hspace{1em}61 & 1 (4.5\%) & 21 (95\%)\\
+\hspace{1em}62 & 8 (26\%) & 23 (74\%)\\
+\hspace{1em}63 & 5 (19\%) & 21 (81\%)\\
+\hspace{1em}64 & 11 (33\%) & 22 (67\%)\\
+\hspace{1em}65 & 11 (20\%) & 45 (80\%)\\
+\hspace{1em}66 & 11 (28\%) & 28 (72\%)\\
+\hspace{1em}67 & 8 (22\%) & 28 (78\%)\\
+\hspace{1em}68 & 5 (19\%) & 21 (81\%)\\
+\hspace{1em}69 & 17 (39\%) & 27 (61\%)\\
+\hspace{1em}70 & 4 (14\%) & 24 (86\%)\\
+\hspace{1em}71 & 1 (3.8\%) & 25 (96\%)\\
+\hspace{1em}72 & 2 (9.1\%) & 20 (91\%)\\
+\hspace{1em}73 & 3 (12\%) & 22 (88\%)\\
+\hspace{1em}74 & 13 (31\%) & 29 (69\%)\\
+\hspace{1em}98 & 147 (42\%) & 201 (58\%)\\*
+\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\end{longtable}
+
 \hypertarget{interactions-from-the-regression-model}{%
@@
 \hypertarget{tbl-hr-percentages}{}
-\begin{longtable}[]{@{}lcc@{}}
-\caption{\label{tbl-hr-percentages}Reported Vaccination Status from the
-Survey for White/Caucasian Individuals by Health Region}\tabularnewline
-\toprule()
-\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N =
-1,080 \\
-\midrule()
-\endfirsthead
-\toprule()
-\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N =
-1,080 \\
-\midrule()
-\endhead
+\begin{table}
+\caption{\label{tbl-hr-percentages}Reported Vaccination Status from the Survey for White/Caucasian
+Individuals by Health Region }\tabularnewline
+
+\centering
+\begin{tabular}{lcc}
+\toprule
+\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N = 1,080\\
+\midrule
 \textbf{Health\_Region} &  & \\
-Toronto & 85 (21\%) & 321 (79\%) \\
-West & 69 (17\%) & 329 (83\%) \\
-East & 41 (15\%) & 236 (85\%) \\
-Central & 38 (16\%) & 194 (84\%) \\
-\bottomrule()
-\end{longtable}
+\hspace{1em}Toronto & 85 (21\%) & 321 (79\%)\\
+\hspace{1em}West & 69 (17\%) & 329 (83\%)\\
+\hspace{1em}East & 41 (15\%) & 236 (85\%)\\
+\hspace{1em}Central & 38 (16\%) & 194 (84\%)\\
+\bottomrule
+\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\end{tabular}
+\end{table}
 
 \hypertarget{tbl-west-hr-lhin}{}
-\begin{longtable}[]{@{}lc@{}}
-\caption{\label{tbl-west-hr-lhin}Proportion of answers by LHIN (West
-Health Region)}\tabularnewline
-\toprule()
-\textbf{Characteristic} & \textbf{N = 839} \\
-\midrule()
-\endfirsthead
-\toprule()
-\textbf{Characteristic} & \textbf{N = 839} \\
-\midrule()
-\endhead
+\begin{table}
+\caption{\label{tbl-west-hr-lhin}Proportion of answers by LHIN (West Health Region) }\tabularnewline
+
+\centering
+\begin{tabular}{lc}
+\toprule
+\textbf{Characteristic} & \textbf{N = 839}\\
+\midrule
 \textbf{LHIN} & \\
-Hamilton Niagara Haldimand Brant (Hnhb) & 271 (32\%) \\
-Waterloo Wellington & 227 (27\%) \\
-South West & 187 (22\%) \\
-Erie St.~Clair & 154 (18\%) \\
-\bottomrule()
-\end{longtable}
+\hspace{1em}Hamilton Niagara Haldimand Brant (Hnhb) & 271 (32\%)\\
+\hspace{1em}Waterloo Wellington & 227 (27\%)\\
+\hspace{1em}South West & 187 (22\%)\\
+\hspace{1em}Erie St. Clair & 154 (18\%)\\
+\bottomrule
+\multicolumn{2}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\end{tabular}
+\end{table}
 

### 2023-03-24 manuscript_v2/main.qmd
updating discussion

@@
 library(patchwork)
-
 ```

### 2023-03-26 manuscript_v2/04-discussion.qmd
updated discussion

--- 2023-03-22 manuscript_v2/04-discussion.qmd
@@
 
-The change from LHINs to Health Regions for healthcare in Ontario was aimed at reducing healthcare disparities [@tsasis2012].
-The relatively recent nature of the change, which occurred in late 2019, highlights the importance of studies that can determine if there are intra-regional differences in healthcare that need to be addressed, particularly in the context of the COVID-19 pandemic.
+There existence of healthcare disparities within Ontario is a topic of particular interest in due to the recent change in the healthcare system of this region, which eliminated the LHIN model and adopted larger Health Regions system in late 2019 in an attempt to address the disparities of the former approach [@tsasis2012;@dong2022].
+In this context, analyzing COVID-19 vaccination estimates within the province is important as they can provide an initial assessment of variations that might need to be addressed by decision-makers.
 
-Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats), and that there were no significant differences in vaccination odds among the age groups considered in the survey.
-This result is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022], and with the vaccination information provided by Public Health Ontario,  which shows that for the period where the Fields survey ran (Sept 30, 2021-Jan 17, 2022) there was a minimum of  80% of first dose vaccination coverage among all the age groups considered in the survey [@ontario-covid].
+Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats).
+It is worth mentioning that province-wide vaccination rates for the period of interest are somewhat different from those of the survey, particularly in the case of those 55 years of age and older, which in the survey had a vaccination rate of 72\%, against a rate of 88.4\% reported for the closest age bracket (50 years of age and older) reported by Public Health Ontario  at the start of the period covered by the data (between September 30th, 2021 and December 12th, 2021) [@ontario-covid].
+In this regard, differences are to be expected because the survey represents a random sample from the population, and therefore, the sampling process is likely to cause variations between the values from the survey and province-wide estimates.
 
-However, we identified intra-provincial differences in vaccination based on socio-economic and geographical factors.
-First, our results show significant differences in vaccination odds in individuals with a household income below CAD 60,000 and in individuals belonging to visible minority groups.
-Those who identified as Black, Latin American, or  belonging to a minority group not included in the survey (Southeast Asian, Filipino, West Asian, and Minority not identified elsewhere) had vaccination odds below 33% when compared to individuals that identified as White/Caucasian (@tbl-model).
+We believe that these differences did not have a significant impact in our analyses, because our results indicate that there were no significant differences in vaccination odds among the age groups considered in the survey, in agreement with the overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], and with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022].
+Moreover, vaccination rates within each age in the dataset are in good agreement with province-wide estimates (e.g., a rate of 95\% for those with 61 years of age, Supplementary Table A-6).
+Additionally, when overall vaccination rates for the province are disaggregated, it can be seen that regional differences were present during the period analyzed, which can be masked by province-wide estimates.
+For example, the Public Health Unit of Lambton (a region within the West Health Region) and the Public Health Unit of Haliburton, Kawartha, and the Pine Ridge District (an area covered by the East Health Region) reported lower vaccination rates (78\%) for those 50 years of age and older at the beginning of the period of interest, whereas other regions had higher vaccination rates [@ontario-covid].
+To this day, differences in vaccination rates within the province are present, because according to Public Health Ontario, as of March of 2023 some regions have less than 75\% vaccination rate [@ontario-covid-map].
+Overall, these reasons indicate that the estimates obtained from the data are in good agreement with the trends from the population.
+
+We identified significant intra-provincial differences in vaccination based on socio-economic and geographical factors.
+First, our results show differences in odds of vaccination in individuals with a household income below CAD 60,000 and in individuals belonging to visible minority groups.
+Those who identified as Arab/Middle Eastern, Black, Latin American, or  that belonged to a minority group not included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were less than a third of individuals that identified as White/Caucasian (@tbl-model).
 These results are consistent with other studies that have shown lower vaccination rates in individuals that identify as part of a racial minority, or that have a low household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
@@
 The West Health Region comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
-In this case, a possible rationale for the results is the fact that in the survey, about 47% of the entries for this Health Region corresponded to White/Caucasian individuals, who reported an overall 83% vaccination rate (Supplementary Table A-6).
+In this case, a possible rationale for the results is the fact that in the survey, about 47% of the entries for this Health Region corresponded to White/Caucasian individuals, who reported an overall 83% vaccination rate (Supplementary Table A-7).
 However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@tbl-model].
@@
 In fact, 50% of the answers from this region in the survey came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
-Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-7).
+Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-8).
 Therefore, this result partly reinforces the well-known existing association between low vaccination rates and income, but it additionally indicates that there were intra-regional differences in vaccination.
@@
 
-The results in this study are based on self-reported data, where the risk of bias exist.
-Despite this, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], the effect of self-reported bias is likely to not be significant in this case.
+The results in this study are based on self-reported data, where there is a risk that biased values are reported.
+Despite this, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], and therefore, the effect of self-reported bias is likely to not be significant in our analyses.
 Finally, it is likely that there have been differences in vaccination across the province as more doses of the vaccine were administered and as successive variants emerged.
 Because this study focused only on vaccination status regarding the first dose of the vaccine within a relatively short time window, it can only provide a snapshot of the societal dynamics behind the pandemic.
-Nonetheless, the results presented here can serve as a starting point to motivate future longitudinal research that aims to quantify geographical differences within vulnerable segments of the population, and that can be used to inform the development of public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.
+Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.

### 2023-03-26 manuscript_v2/main.tex
updated discussion

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 Mathematical Modelling of COVID-19 Task Force under ethical guidance
-from the University of Toronto. The survey collected socio-economic
+from the University of Toronto, and which ran between September 30th,
+2021 and January 17th,2022. The survey collected socio-economic
 information from participants (Table~\ref{tbl-covariates}), recorded
@@
 total number of unique entries used for analysis was 3,551 which
-included the East, Central, Toronto, and West Health Regions.
+included the East, Central, Toronto, and West Health Regions and that
+covered the period between September 30th,2021 and December 12th, 2021.
 
@@
 
-The change from LHINs to Health Regions for healthcare in Ontario was
-aimed at reducing healthcare
-disparities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. The
-relatively recent nature of the change, which occurred in late 2019,
-highlights the importance of studies that can determine if there are
-intra-regional differences in healthcare that need to be addressed,
-particularly in the context of the COVID-19 pandemic.
+There existence of healthcare disparities within Ontario is a topic of
+particular interest in due to the recent change in the healthcare system
+of this region, which eliminated the LHIN model and adopted larger
+Health Regions system in late 2019 in an attempt to address the
+disparities of the former
+approach\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
+In this context, analyzing COVID-19 vaccination estimates within the
+province is important as they can provide an initial assessment of
+variations that might need to be addressed by decision-makers.
 
@@
 reported having received the first dose of the COVID-19 vaccine
-(Table~\ref{tbl-descriptive-stats}), and that there were no significant
-differences in vaccination odds among the age groups considered in the
-survey. This result is consistent with overall vaccination rates
-reported for Canada, which have been relatively higher when compared to
-other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{39}}, with
-vaccination uptake rates across different age groups presented in other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{40}},
-and with the vaccination information provided by Public Health Ontario,
-which shows that for the period where the Fields survey ran (Sept 30,
-2021-Jan 17, 2022) there was a minimum of 80\% of first dose vaccination
-coverage among all the age groups considered in the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}.
-
-However, we identified intra-provincial differences in vaccination based
-on socio-economic and geographical factors. First, our results show
-significant differences in vaccination odds in individuals with a
-household income below CAD 60,000 and in individuals belonging to
-visible minority groups. Those who identified as Black, Latin American,
-or belonging to a minority group not included in the survey (Southeast
-Asian, Filipino, West Asian, and Minority not identified elsewhere) had
-vaccination odds below 33\% when compared to individuals that identified
-as White/Caucasian (Table~\ref{tbl-model}). These results are consistent
-with other studies that have shown lower vaccination rates in
-individuals that identify as part of a racial minority, or that have a
-low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{42}}.
+(Table~\ref{tbl-descriptive-stats}). It is worth mentioning that
+province-wide vaccination rates for the period of interest are somewhat
+different from those of the survey, particularly in the case of those 55
+years of age and older, which in the survey had a vaccination rate of
+72\%, against a rate of 88.4\% reported for the closest age bracket (50
+years of age and older) reported by Public Health Ontario at the start
+of the period covered by the data (between September 30th, 2021 and
+December 12th,
+2021)\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. In
+this regard, differences are to be expected because the survey
+represents a random sample from the population, and therefore, the
+sampling process is likely to cause variations between the values from
+the survey and province-wide estimates.
+
+We believe that these differences did not have a significant impact in
+our analyses, because our results indicate that there were no
+significant differences in vaccination odds among the age groups
+considered in the survey, in agreement with the overall vaccination
+rates reported for Canada, which have been relatively higher when
+compared to other high income
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{40}}, and
+with vaccination uptake rates across different age groups presented in
+other
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{41}}.
+Moreover, vaccination rates within each age in the dataset are in good
+agreement with province-wide estimates (e.g., a rate of 95\% for those
+with 61 years of age, Supplementary Table A-6). Additionally, when
+overall vaccination rates for the province are disaggregated, it can be
+seen that regional differences were present during the period analyzed,
+which can be masked by province-wide estimates. For example, the Public
+Health Unit of Lambton (a region within the West Health Region) and the
+Public Health Unit of Haliburton, Kawartha, and the Pine Ridge District
+(an area covered by the East Health Region) reported lower vaccination
+rates (78\%) for those 50 years of age and older at the beginning of the
+period of interest, whereas other regions had higher vaccination
+rates\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. To
+this day, differences in vaccination rates within the province are
+present, because according to Public Health Ontario, as of March of 2023
+some regions have less than 75\% vaccination
+rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{42}}.
+Overall, these reasons indicate that the estimates obtained from the
+data are in good agreement with the trends from the population.
+
+We identified significant intra-provincial differences in vaccination
+based on socio-economic and geographical factors. First, our results
+show differences in odds of vaccination in individuals with a household
+income below CAD 60,000 and in individuals belonging to visible minority
+groups. Those who identified as Arab/Middle Eastern, Black, Latin
+American, or that belonged to a minority group not included in the
+survey (Southeast Asian, Filipino, West Asian, and minority groups not
+identified elsewhere) had vaccination odds that were less than a third
+of individuals that identified as White/Caucasian
+(Table~\ref{tbl-model}). These results are consistent with other studies
+that have shown lower vaccination rates in individuals that identify as
+part of a racial minority, or that have a low household
+income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{43}}.
 
@@
 as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{43},\protect\hyperlink{ref-ct2021}{44}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{44},\protect\hyperlink{ref-ct2021}{45}},
 which include occupations such as grocery store workers, gas station
@@
 had priority for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{45}},
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{46}},
 which would explain the higher odds of vaccination for these individuals
@@
 White/Caucasian individuals, who reported an overall 83\% vaccination
-rate (Supplementary Table A-6). However, the interaction effect of
+rate (Supplementary Table A-7). However, the interaction effect of
 Health Region and race was also significant in the case of individuals
@@
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{47}}.
 In the case of this analysis, the estimated lower odds are likely to be
@@
 population (more than 20\%) in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{47}}
-(Supplementary Table A-7). Therefore, this result partly reinforces the
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{48}}
+(Supplementary Table A-8). Therefore, this result partly reinforces the
 well-known existing association between low vaccination rates and
@@
 previously reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{48}}; in this
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{49}}; in this
 regard, our result provides additional context by showing that within
@@
 are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}, a
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{47}}, a
 minority group that has historically suffered from reduced access to
@@
 
-The results in this study are based on self-reported data, where the
-risk of bias exist. Despite this, because in the context of COVID-19 it
-has been shown that good agreement exists between self-reported and
-documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{49}}, the
-effect of self-reported bias is likely to not be significant in this
-case. Finally, it is likely that there have been differences in
-vaccination across the province as more doses of the vaccine were
-administered and as successive variants emerged. Because this study
-focused only on vaccination status regarding the first dose of the
-vaccine within a relatively short time window, it can only provide a
-snapshot of the societal dynamics behind the pandemic. Nonetheless, the
-results presented here can serve as a starting point to motivate future
-longitudinal research that aims to quantify geographical differences
-within vulnerable segments of the population, and that can be used to
-inform the development of public health policies within the province of
+The results in this study are based on self-reported data, where there
+is a risk that biased values are reported. Despite this, because in the
+context of COVID-19 it has been shown that good agreement exists between
+self-reported and documented vaccination
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{50}}, and
+therefore, the effect of self-reported bias is likely to not be
+significant in our analyses. Finally, it is likely that there have been
+differences in vaccination across the province as more doses of the
+vaccine were administered and as successive variants emerged. Because
+this study focused only on vaccination status regarding the first dose
+of the vaccine within a relatively short time window, it can only
+provide a snapshot of the societal dynamics behind the pandemic.
+Nonetheless, the results presented here can serve as a starting point to
+motivate the collection of robust longitudinal data that can be used to
+quantify geographical and temporal differences within vulnerable
+segments of the population, and that can be used to inform the
+development of adequate public health policies within the province of
 Ontario or across other provinces that aim to minimize disparities in
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
+2023.
+\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
-2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
+\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
+\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=maps}}
 
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-03-26 manuscript_v2/appendix.qmd
Updated files

@@
 We used the iterative proportional fitting procedure (_raking_) in order to account for the differences in income, age groups, race/ethnicity and the population from the Health Regions.
-Details of the implementation can be found in the R Script `raking_logistic_regression.R` in the `code` directory.
+Details of the implementation can be found in the R Script `raking.R` in the `code` directory.
 
@@
 
-The tables below present summaries of information from the Fields survey (from the clean dataset analyzed in the main paper), showing responses by White/Caucasian individuals by Health Region and proportion of answers by LHIN in the West Health Region.
+Tables A-7 and A-8 present summaries of information from the Fields survey (from the clean dataset analyzed in the main paper), showing responses by White/Caucasian individuals by Health Region and proportion of answers by LHIN in the West Health Region.
 

### 2023-04-10 manuscript_v2/main.qmd
changed interlining

@@
     keep-tex: true
-linestretch: 2
+#linestretch: 1
 bibliography: references/refs.bib

### 2023-04-10 manuscript_v2/main.tex
changed interlining

@@
 \usepackage{lmodern}
-\usepackage{setspace}
 \usepackage{iftex}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
-
-\setstretch{2}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+
 \textsuperscript{1} Centre de Recherches Mathématiques, University of

### 2023-04-14 manuscript_v2/03-results.qmd
Added month to model, need to double-check

@@
 
+t2<-m2 %>%
+  tbl_regression(exponentiate = TRUE,
+                 label=list(Health_Region ~ "Health Region",
+                            Age_group_ord ~ "Age Group",
+                            income_ord="Income")) %>%
+  bold_labels()
+
+# ggplot(data=clean_data,aes(y=first_dose))+
+#   geom_bar()+facet_wrap(~Month)
+
 tbl_merge <-
   tbl_merge(
-    tbls = list(t0, t1),
+    tbls = list(t0, t2),
     tab_spanner = c("**Uncorrected**", "**Corrected**")

### 2023-04-14 manuscript_v2/main.qmd
Added month to model, need to double-check

@@
 library(lme4)
+library(lubridate)
 library(kableExtra)

### 2023-04-14 manuscript_v2/main.tex
Added month to model, need to double-check

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, frame hidden, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
 
@@
 \toprule
-\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,593\\
+\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,591\\
 \midrule
 \textbf{Income} &  & \\
-\hspace{1em}60000\_and\_above & 305 (23\%) & 1,049 (77\%)\\
+\hspace{1em}60000\_and\_above & 305 (23\%) & 1,048 (77\%)\\
 \hspace{1em}25000\_59999 & 253 (28\%) & 636 (72\%)\\
-\hspace{1em}under\_25000 & 400 (31\%) & 908 (69\%)\\
+\hspace{1em}under\_25000 & 400 (31\%) & 907 (69\%)\\
 \textbf{Age Group} &  & \\
-\hspace{1em}16\_34 & 409 (27\%) & 1,112 (73\%)\\
+\hspace{1em}16\_34 & 409 (27\%) & 1,111 (73\%)\\
 \hspace{1em}35\_54 & 252 (26\%) & 712 (74\%)\\
-\hspace{1em}55\_and\_over & 297 (28\%) & 769 (72\%)\\
+\hspace{1em}55\_and\_over & 297 (28\%) & 768 (72\%)\\
 \textbf{Health Region} &  & \\
-\hspace{1em}Toronto & 371 (28\%) & 953 (72\%)\\
-\hspace{1em}Central & 224 (28\%) & 581 (72\%)\\
+\hspace{1em}Toronto & 371 (28\%) & 952 (72\%)\\
+\hspace{1em}Central & 224 (28\%) & 580 (72\%)\\
 \hspace{1em}East & 135 (23\%) & 448 (77\%)\\
@@
 \textbf{Race} &  & \\
-\hspace{1em}white\_caucasian & 233 (18\%) & 1,080 (82\%)\\
+\hspace{1em}white\_caucasian & 233 (18\%) & 1,079 (82\%)\\
 \hspace{1em}arab\_middle\_eastern & 76 (36\%) & 138 (64\%)\\
@@
 \hspace{1em}other & 128 (35\%) & 239 (65\%)\\
-\hspace{1em}south\_asian & 88 (23\%) & 287 (77\%)\\
+\hspace{1em}south\_asian & 88 (24\%) & 286 (76\%)\\
 \bottomrule
@@
 \hspace{1em}16\_34 & — & — &  & — & — & \\
-\hspace{1em}35\_54 & 0.93 & 0.77, 1.13 & 0.5 & 0.90 & 0.67, 1.21 & 0.5\\
+\hspace{1em}35\_54 & 0.93 & 0.77, 1.14 & 0.5 & 0.90 & 0.67, 1.21 & 0.5\\
 \hspace{1em}55\_and\_over & 0.74 & 0.61, 0.89 & 0.002 & 0.99 & 0.74, 1.32 & >0.9\\
+\textbf{Month} &  &  &  &  &  & \\
+\hspace{1em}October & — & — &  & — & — & \\
+\hspace{1em}December & 1.16 & 0.91, 1.49 & 0.2 & 1.16 & 0.78, 1.71 & 0.5\\
+\hspace{1em}November & 0.95 & 0.81, 1.12 & 0.6 & 0.92 & 0.71, 1.18 & 0.5\\
 \textbf{Income} &  &  &  &  &  & \\
 \hspace{1em}60000\_and\_above & — & — &  & — & — & \\
-\hspace{1em}25000\_59999 & 0.59 & 0.41, 0.84 & 0.004 & 0.59 & 0.39, 0.89 & 0.011\\
+\hspace{1em}25000\_59999 & 0.59 & 0.41, 0.84 & 0.003 & 0.59 & 0.39, 0.88 & 0.010\\
 \hspace{1em}under\_25000 & 0.41 & 0.29, 0.58 & <0.001 & 0.37 & 0.25, 0.56 & <0.001\\
@@
 \hspace{1em}white\_caucasian & — & — &  & — & — & \\
-\hspace{1em}arab\_middle\_eastern & 0.24 & 0.12, 0.50 & <0.001 & 0.31 & 0.14, 0.69 & 0.004\\
+\hspace{1em}arab\_middle\_eastern & 0.24 & 0.12, 0.49 & <0.001 & 0.31 & 0.14, 0.68 & 0.003\\
 \hspace{1em}black & 0.30 & 0.17, 0.54 & <0.001 & 0.32 & 0.17, 0.60 & <0.001\\
-\hspace{1em}east\_asian\_pacific\_islander & 0.74 & 0.36, 1.52 & 0.4 & 1.15 & 0.50, 2.66 & 0.7\\
-\hspace{1em}indigenous & 0.37 & 0.17, 0.81 & 0.013 & 0.44 & 0.19, 1.02 & 0.056\\
-\hspace{1em}latin\_american & 0.28 & 0.13, 0.59 & <0.001 & 0.28 & 0.11, 0.67 & 0.004\\
-\hspace{1em}mixed & 0.59 & 0.31, 1.12 & 0.11 & 0.64 & 0.25, 1.65 & 0.4\\
+\hspace{1em}east\_asian\_pacific\_islander & 0.73 & 0.36, 1.51 & 0.4 & 1.15 & 0.50, 2.66 & 0.7\\
+\hspace{1em}indigenous & 0.38 & 0.17, 0.82 & 0.014 & 0.45 & 0.20, 1.04 & 0.061\\
+\hspace{1em}latin\_american & 0.28 & 0.13, 0.59 & <0.001 & 0.27 & 0.11, 0.66 & 0.004\\
+\hspace{1em}mixed & 0.58 & 0.30, 1.11 & 0.10 & 0.63 & 0.24, 1.62 & 0.3\\
 \hspace{1em}other & 0.20 & 0.11, 0.35 & <0.001 & 0.22 & 0.12, 0.41 & <0.001\\
-\hspace{1em}south\_asian & 0.80 & 0.44, 1.45 & 0.5 & 0.91 & 0.49, 1.69 & 0.8\\
+\hspace{1em}south\_asian & 0.79 & 0.44, 1.43 & 0.4 & 0.90 & 0.48, 1.67 & 0.7\\
 \textbf{Health Region} &  &  &  &  &  & \\
 \hspace{1em}Toronto & — & — &  & — & — & \\
-\hspace{1em}Central & 1.30 & 0.85, 2.00 & 0.2 & 1.47 & 0.92, 2.35 & 0.11\\
-\hspace{1em}East & 1.54 & 1.01, 2.34 & 0.044 & 1.42 & 0.90, 2.23 & 0.13\\
-\hspace{1em}West & 1.35 & 0.95, 1.94 & 0.10 & 1.55 & 1.05, 2.30 & 0.029\\
+\hspace{1em}Central & 1.30 & 0.84, 2.00 & 0.2 & 1.47 & 0.92, 2.36 & 0.11\\
+\hspace{1em}East & 1.54 & 1.01, 2.34 & 0.044 & 1.41 & 0.90, 2.22 & 0.14\\
+\hspace{1em}West & 1.35 & 0.94, 1.93 & 0.10 & 1.54 & 1.04, 2.29 & 0.031\\
 \textbf{Income * Race} &  &  &  &  &  & \\
-\hspace{1em}25000\_59999 * arab\_middle\_eastern & 2.16 & 0.93, 4.99 & 0.072 & 1.79 & 0.67, 4.83 & 0.2\\
-\hspace{1em}under\_25000 * arab\_middle\_eastern & 2.96 & 1.39, 6.26 & 0.005 & 3.05 & 1.26, 7.39 & 0.013\\
-\hspace{1em}25000\_59999 * black & 1.19 & 0.60, 2.39 & 0.6 & 1.34 & 0.59, 3.05 & 0.5\\
-\hspace{1em}under\_25000 * black & 2.88 & 1.48, 5.59 & 0.002 & 3.19 & 1.45, 6.99 & 0.004\\
-\hspace{1em}25000\_59999 * east\_asian\_pacific\_islander & 0.95 & 0.44, 2.07 & >0.9 & 0.42 & 0.17, 1.05 & 0.062\\
-\hspace{1em}under\_25000 * east\_asian\_pacific\_islander & 1.91 & 0.90, 4.04 & 0.090 & 1.16 & 0.47, 2.86 & 0.8\\
-\hspace{1em}25000\_59999 * indigenous & 1.81 & 0.74, 4.43 & 0.2 & 1.36 & 0.48, 3.89 & 0.6\\
-\hspace{1em}under\_25000 * indigenous & 1.64 & 0.73, 3.69 & 0.2 & 1.45 & 0.55, 3.80 & 0.5\\
-\hspace{1em}25000\_59999 * latin\_american & 0.89 & 0.38, 2.10 & 0.8 & 1.24 & 0.45, 3.43 & 0.7\\
-\hspace{1em}under\_25000 * latin\_american & 3.09 & 1.33, 7.16 & 0.009 & 2.80 & 1.04, 7.51 & 0.041\\
-\hspace{1em}25000\_59999 * mixed & 0.86 & 0.39, 1.93 & 0.7 & 0.85 & 0.32, 2.26 & 0.7\\
-\hspace{1em}under\_25000 * mixed & 1.26 & 0.64, 2.47 & 0.5 & 1.10 & 0.37, 3.27 & 0.9\\
-\hspace{1em}25000\_59999 * other & 5.46 & 2.41, 12.3 & <0.001 & 6.93 & 2.65, 18.1 & <0.001\\
-\hspace{1em}under\_25000 * other & 4.06 & 2.25, 7.31 & <0.001 & 4.59 & 2.33, 9.05 & <0.001\\
-\hspace{1em}25000\_59999 * south\_asian & 1.13 & 0.54, 2.36 & 0.7 & 1.20 & 0.51, 2.85 & 0.7\\
-\hspace{1em}under\_25000 * south\_asian & 1.59 & 0.83, 3.06 & 0.2 & 2.00 & 0.93, 4.30 & 0.077\\
+\hspace{1em}25000\_59999 * arab\_middle\_eastern & 2.19 & 0.95, 5.05 & 0.067 & 1.85 & 0.68, 4.99 & 0.2\\
+\hspace{1em}under\_25000 * arab\_middle\_eastern & 2.95 & 1.39, 6.24 & 0.005 & 3.08 & 1.27, 7.47 & 0.013\\
+\hspace{1em}25000\_59999 * black & 1.18 & 0.59, 2.36 & 0.6 & 1.31 & 0.58, 2.98 & 0.5\\
+\hspace{1em}under\_25000 * black & 2.84 & 1.46, 5.53 & 0.002 & 3.15 & 1.43, 6.92 & 0.004\\
+\hspace{1em}25000\_59999 * east\_asian\_pacific\_islander & 0.97 & 0.44, 2.10 & >0.9 & 0.42 & 0.17, 1.06 & 0.066\\
+\hspace{1em}under\_25000 * east\_asian\_pacific\_islander & 1.90 & 0.90, 4.01 & 0.094 & 1.15 & 0.47, 2.84 & 0.8\\
+\hspace{1em}25000\_59999 * indigenous & 1.77 & 0.72, 4.32 & 0.2 & 1.33 & 0.46, 3.80 & 0.6\\
+\hspace{1em}under\_25000 * indigenous & 1.60 & 0.71, 3.60 & 0.3 & 1.41 & 0.54, 3.72 & 0.5\\
+\hspace{1em}25000\_59999 * latin\_american & 0.89 & 0.38, 2.10 & 0.8 & 1.25 & 0.45, 3.45 & 0.7\\
+\hspace{1em}under\_25000 * latin\_american & 3.10 & 1.34, 7.21 & 0.009 & 2.81 & 1.04, 7.59 & 0.041\\
+\hspace{1em}25000\_59999 * mixed & 0.87 & 0.39, 1.93 & 0.7 & 0.87 & 0.33, 2.33 & 0.8\\
+\hspace{1em}under\_25000 * mixed & 1.27 & 0.65, 2.49 & 0.5 & 1.12 & 0.38, 3.33 & 0.8\\
+\hspace{1em}25000\_59999 * other & 5.50 & 2.43, 12.5 & <0.001 & 6.96 & 2.67, 18.2 & <0.001\\
+\hspace{1em}under\_25000 * other & 4.09 & 2.27, 7.37 & <0.001 & 4.63 & 2.34, 9.13 & <0.001\\
+\hspace{1em}25000\_59999 * south\_asian & 1.14 & 0.55, 2.38 & 0.7 & 1.22 & 0.52, 2.87 & 0.7\\
+\hspace{1em}under\_25000 * south\_asian & 1.58 & 0.82, 3.05 & 0.2 & 2.01 & 0.93, 4.34 & 0.075\\
 \textbf{Race * Health Region} &  &  &  &  &  & \\
-\hspace{1em}arab\_middle\_eastern * Central & 0.75 & 0.33, 1.71 & 0.5 & 0.66 & 0.26, 1.70 & 0.4\\
-\hspace{1em}black * Central & 0.48 & 0.23, 1.01 & 0.055 & 0.44 & 0.19, 0.98 & 0.046\\
-\hspace{1em}east\_asian\_pacific\_islander * Central & 0.84 & 0.37, 1.88 & 0.7 & 0.98 & 0.38, 2.53 & >0.9\\
-\hspace{1em}indigenous * Central & 0.60 & 0.24, 1.51 & 0.3 & 0.63 & 0.22, 1.79 & 0.4\\
-\hspace{1em}latin\_american * Central & 0.69 & 0.28, 1.72 & 0.4 & 0.67 & 0.23, 1.96 & 0.5\\
-\hspace{1em}mixed * Central & 0.63 & 0.29, 1.35 & 0.2 & 0.73 & 0.24, 2.22 & 0.6\\
-\hspace{1em}other * Central & 0.96 & 0.48, 1.92 & >0.9 & 0.80 & 0.36, 1.78 & 0.6\\
-\hspace{1em}south\_asian * Central & 0.70 & 0.34, 1.45 & 0.3 & 0.54 & 0.25, 1.20 & 0.13\\
-\hspace{1em}arab\_middle\_eastern * East & 0.58 & 0.20, 1.65 & 0.3 & 0.43 & 0.13, 1.45 & 0.2\\
-\hspace{1em}black * East & 0.82 & 0.36, 1.86 & 0.6 & 0.83 & 0.34, 2.04 & 0.7\\
-\hspace{1em}east\_asian\_pacific\_islander * East & 0.82 & 0.30, 2.24 & 0.7 & 0.86 & 0.29, 2.56 & 0.8\\
-\hspace{1em}indigenous * East & 0.55 & 0.21, 1.44 & 0.2 & 0.69 & 0.23, 2.08 & 0.5\\
-\hspace{1em}latin\_american * East & 0.79 & 0.28, 2.23 & 0.7 & 1.03 & 0.32, 3.34 & >0.9\\
-\hspace{1em}mixed * East & 0.71 & 0.31, 1.60 & 0.4 & 0.91 & 0.28, 3.03 & 0.9\\
-\hspace{1em}other * East & 0.86 & 0.37, 1.99 & 0.7 & 1.05 & 0.39, 2.83 & >0.9\\
-\hspace{1em}south\_asian * East & 0.50 & 0.20, 1.24 & 0.14 & 0.52 & 0.19, 1.45 & 0.2\\
-\hspace{1em}arab\_middle\_eastern * West & 1.16 & 0.48, 2.77 & 0.7 & 1.00 & 0.37, 2.73 & >0.9\\
-\hspace{1em}black * West & 0.77 & 0.36, 1.65 & 0.5 & 0.76 & 0.32, 1.80 & 0.5\\
-\hspace{1em}east\_asian\_pacific\_islander * West & 0.54 & 0.24, 1.20 & 0.13 & 0.52 & 0.20, 1.34 & 0.2\\
-\hspace{1em}indigenous * West & 0.44 & 0.19, 1.02 & 0.056 & 0.39 & 0.14, 1.09 & 0.073\\
-\hspace{1em}latin\_american * West & 0.97 & 0.39, 2.43 & >0.9 & 0.94 & 0.32, 2.72 & >0.9\\
-\hspace{1em}mixed * West & 0.53 & 0.25, 1.11 & 0.092 & 0.37 & 0.12, 1.16 & 0.089\\
-\hspace{1em}other * West & 0.55 & 0.27, 1.12 & 0.10 & 0.41 & 0.18, 0.93 & 0.032\\
-\hspace{1em}south\_asian * West & 0.50 & 0.23, 1.07 & 0.075 & 0.41 & 0.18, 0.95 & 0.037\\*
+\hspace{1em}arab\_middle\_eastern * Central & 0.76 & 0.33, 1.71 & 0.5 & 0.66 & 0.25, 1.69 & 0.4\\
+\hspace{1em}black * Central & 0.48 & 0.23, 1.02 & 0.055 & 0.44 & 0.19, 0.99 & 0.046\\
+\hspace{1em}east\_asian\_pacific\_islander * Central & 0.84 & 0.38, 1.89 & 0.7 & 0.98 & 0.38, 2.53 & >0.9\\
+\hspace{1em}indigenous * Central & 0.60 & 0.24, 1.50 & 0.3 & 0.63 & 0.22, 1.78 & 0.4\\
+\hspace{1em}latin\_american * Central & 0.69 & 0.28, 1.71 & 0.4 & 0.66 & 0.22, 1.93 & 0.4\\
+\hspace{1em}mixed * Central & 0.63 & 0.30, 1.35 & 0.2 & 0.73 & 0.24, 2.23 & 0.6\\
+\hspace{1em}other * Central & 0.98 & 0.49, 1.95 & >0.9 & 0.81 & 0.36, 1.81 & 0.6\\
+\hspace{1em}south\_asian * Central & 0.71 & 0.34, 1.47 & 0.4 & 0.55 & 0.25, 1.21 & 0.14\\
+\hspace{1em}arab\_middle\_eastern * East & 0.57 & 0.20, 1.63 & 0.3 & 0.43 & 0.13, 1.43 & 0.2\\
+\hspace{1em}black * East & 0.84 & 0.37, 1.89 & 0.7 & 0.86 & 0.35, 2.11 & 0.7\\
+\hspace{1em}east\_asian\_pacific\_islander * East & 0.82 & 0.30, 2.24 & 0.7 & 0.86 & 0.29, 2.54 & 0.8\\
+\hspace{1em}indigenous * East & 0.56 & 0.21, 1.45 & 0.2 & 0.70 & 0.23, 2.11 & 0.5\\
+\hspace{1em}latin\_american * East & 0.80 & 0.28, 2.25 & 0.7 & 1.03 & 0.31, 3.40 & >0.9\\
+\hspace{1em}mixed * East & 0.70 & 0.31, 1.60 & 0.4 & 0.92 & 0.28, 3.05 & 0.9\\
+\hspace{1em}other * East & 0.87 & 0.38, 2.02 & 0.7 & 1.08 & 0.40, 2.91 & 0.9\\
+\hspace{1em}south\_asian * East & 0.50 & 0.20, 1.24 & 0.14 & 0.51 & 0.19, 1.42 & 0.2\\
+\hspace{1em}arab\_middle\_eastern * West & 1.18 & 0.49, 2.82 & 0.7 & 1.02 & 0.37, 2.77 & >0.9\\
+\hspace{1em}black * West & 0.78 & 0.36, 1.67 & 0.5 & 0.77 & 0.32, 1.82 & 0.5\\
+\hspace{1em}east\_asian\_pacific\_islander * West & 0.54 & 0.24, 1.21 & 0.14 & 0.52 & 0.20, 1.36 & 0.2\\
+\hspace{1em}indigenous * West & 0.45 & 0.19, 1.05 & 0.064 & 0.40 & 0.14, 1.12 & 0.082\\
+\hspace{1em}latin\_american * West & 0.99 & 0.39, 2.48 & >0.9 & 0.96 & 0.33, 2.79 & >0.9\\
+\hspace{1em}mixed * West & 0.53 & 0.25, 1.10 & 0.088 & 0.37 & 0.12, 1.15 & 0.085\\
+\hspace{1em}other * West & 0.55 & 0.27, 1.12 & 0.10 & 0.41 & 0.18, 0.92 & 0.032\\
+\hspace{1em}south\_asian * West & 0.50 & 0.24, 1.08 & 0.078 & 0.41 & 0.18, 0.95 & 0.037\\*
 \multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\

### 2023-04-14 code/data_preparation.R

@@ -63,9 +63,31 @@ clean_data<-clean_data%>%
   mutate(Age_group_ord=as.factor(Age_group_ord))
 
 
+
+## get month from the data
+
+clean_data$date <-as_date(clean_data$date)
+
+
+clean_data<-clean_data %>%
+  mutate(Month=format(date,"%B"))
+
+clean_data<-clean_data %>%
+  mutate(MonthDay=as.numeric(format(date,"%m.%d")))
+
+
+clean_data$Month <- as.factor(clean_data$Month)
+
 ## relevel for reference groups in the model
 
 clean_data<- within(clean_data, Age_group_ord <- relevel(Age_group_ord, ref = "16_34"))
 clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
 clean_data<- within(clean_data, income_ord <- relevel(income_ord, ref = "60000_and_above"))
+clean_data<- within(clean_data, Month <- relevel(Month, ref = "October"))
+
+# remove september, only 2 observations
+
+clean_data <-clean_data %>%
+  subset(Month!="September")
+  

### 2023-04-14 code/regression.R

@@ -1,7 +1,7 @@
 ## This script uses the object created in the raking.R script to fit a fixed-effects logistic regression
 
 #model for uncorrected data
-m0<-svyglm(first_dose_m~Age_group_ord+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
+m0<-svyglm(first_dose_m~Age_group_ord+Month+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
            design=a1,
            family = quasibinomial(), 
            control= list(maxit=25))
@@ -15,3 +15,7 @@ m1<-svyglm(first_dose_m~Age_group_ord+income_ord+Race+Health_Region+Race*income_
            control= list(maxit=25))
 
 
+m2<-svyglm(first_dose_m~Age_group_ord+Month+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
+           design=a1_rake,
+           family = quasibinomial(), 
+           control= list(maxit=25))

### 2023-04-16 manuscript_v3/01-background.qmd
Updated months, fix descriptive table as September still appears

--- 2023-03-23 manuscript_v2/01-background.qmd
@@
 This made some believe that vaccines were destined to be a determinant factor in a rapid ending of the pandemic [@thelancet2021].
-However, although it has been estimated that COVID-19 vaccines have prevented around 14 million of deaths worldwide [@watson2022], their implementation has been far from being equal to that of the vaccines of smallpox and polio, which were implemented on a global scale and that were indeed crucial to control these diseases [@kayser2021].
+However, although it has been estimated that COVID-19 vaccines have prevented around 14 million of deaths worldwide [@watson2022], their implementation has been far from being equal to that of the smallpox and polio vaccines, which were implemented on a global scale and that were crucial to control these diseases [@kayser2021].
 In fact, the rollout of COVID-19 vaccines has faced multiple challenges since its inception which ultimately have hampered their use to achieve the ultimate goal of global immunity.
 
-This problematic of COVID-19 vaccines rollout is a multifaceted issue resulting from, among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020].
+This problematic in the rollout of the COVID-19 vaccines is a multifaceted issue resulting from, among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020].
 In particular, it is well established that differences in vaccination uptake have been present even in countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within racial minorities (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021].
@@
 In late 2019, the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East) in an effort to improve the healthcare system of the province [@dong2022].
-Because the adoption of the Health Regions occurred at a relatively recent time, there is an ongoing need to analyze the impact of this measure and identify the existence of intra-regional differences that might exist, and which could be specially important in the context of the COVID-19 pandemic.
+Because the adoption of the Health Regions is relatively recent there is an ongoing need to analyze the impact of this measure and identify the existence of intra-regional differences that might exist, and which could be specially important in the context of the COVID-19 pandemic.
 
-In this study, we analyzed differences in self-reported COVID-19 vaccination status in Ontario using socio-economic (e.g., income, racial/ethnic identification), and spatial  information at the level of the Health Regions to identify the existence of differences that might need to be addressed to ensure that the healthcare system of Ontario is more inclusive and that responds to the needs of its most vulnerable population.
+In this study, we analyzed differences in self-reported COVID-19 vaccination status in Ontario using socio-economic (e.g., income, racial/ethnic identification), temporal (through a consecutive period of three months in late 2021), and spatial  information (using the Health Regions of Ontario) to identify the existence of differences that might need to be addressed to ensure that the healthcare system of the province is more inclusive and that responds to the needs of its most vulnerable population.

### 2023-04-16 manuscript_v3/03-results.qmd
Updated months, fix descriptive table as September still appears

--- 2023-04-14 manuscript_v2/03-results.qmd
@@
 @tbl-descriptive-stats shows the descriptive statistics from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed.
-The total number of entries analyzed was 3,551.
-Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it.
-Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\% of the total number of entries, those within the CAD 25,000-59,999 income bracket represented 25% of the total sample, and those with an income above CAD 60,000 represented 38 \% of the sample; across all income brackets, the percentage of individuals that reported having received a first dose of the vaccine was consistent, above 69\%.
+The total number of entries analyzed was 3,549.
+Overall, 27\% (958) of survey respondents  reported not having received the first dose of the vaccine, whereas 73\% (2,591) reported having received it.
+Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\%  (1307) of the number of entries, those within the CAD 25,000-59,999 income bracket represented 25% (889) of the total sample, and those with an income above CAD 60,000 represented 38 \% (1353) of the sample.
+Between the three income brackets, the lower vaccination rate corresponded to the under CAD 25,000 group, with 69\% .
 
-Within the age groups of survey respondents, the age group between 16-34 years had the highest representation in the survey responses (1,521, 42.8\% of all responses).
-Within this age bracket, 73\% of respondents indicated having received the vaccine, whereas the lowest vaccination rate was in the bracket of those 55 years of age and above, with a total of 72\%.
-The Health Region with highest representation in the survey was Toronto, accounting for 1,324 entries (37.2\%), with a vaccination rate of 72\%.
-Regarding race/ethnicity, individuals that identified as White/Caucasian represented 1313 (37\%) of all entries and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine.
+Within the age groups of survey respondents, the age group between 16-34 years had the highest representation in the survey responses (1,520 or 42.8\% of all responses).
+Within this age bracket, 73\% of respondents indicated having received the vaccine, whereas the lowest vaccination rate was in the bracket of those 55 years of age and above, with 72\%.
+The Health Region with highest representation in the survey was Toronto, accounting for 1,323 entries (37.3\% of the total), with a vaccination rate of 72\%.
+Regarding race/ethnicity, individuals that identified as White/Caucasian represented 37\% of all entries (1,312) and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine.
 On the other hand, the ethnic group with the lowest number of entries in the survey was Latin American, with a total of 180, or 5\% of all entries.
@@
                         Health_Region,
+                        Month,
                         Race),
@@
 
-@tbl-model shows the results of the logistic regression models (for the uncorrected and corrected data) on vaccination status using socio-economic factors (age group, income, race), geographical areas (Health Regions) and the interactions between income and race and Health Region and race.
-There were no statistically significant differences in vaccination rates within the age groups from the survey, but significant odds ratios were estimated for other covariates.
-Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, _p_=0.011 and <0.001, respectively).
-Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.28, and _p_=0.004,<0.001 and 0.004, respectively); additionally, those individuals in the Other Race/Ethnicity group (a group that included Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere) had even lower odds of vaccination than the other minority groups (OR=0.22, _p_<0.001).
-Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.55, _p_=0.029).
+@tbl-model shows the results of the logistic regression models (for the uncorrected and corrected data) on vaccination status using socio-economic factors (age group, income, race), months covered by the survey (October, November, and December), geographical areas (Health Regions) and the interactions between income and race and Health Region and race.
+The corrected results show that there were no statistically significant differences in vaccination rate between the age groups or across the months covered by the survey as in both cases the odd ratios were similar.
+However, between the different household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, _p_<0.001 and _p_=0.010, respectively).
+Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.32, 0.32, 0.27, and _p_=0.003, _p_<0.001 and _p_=0.004, respectively); additionally, those individuals that reported to belong to the "Other" Race/Ethnicity group (which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census) had even lower odds of vaccination than the other minority groups (OR=0.22, _p_<0.001).
+Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, _p_=0.031).
 
-Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001).
-Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.93, _p_<0.001).
+Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.08, _p_=0.013), Black (OR=3.15, _p_=0.004), Latin American (OR=2.81, _p_=0.041), or that belonged to other minority groups  (OR=4.63, _p_<0.001).
+Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.96, _p_<0.001).
 

### 2023-04-16 manuscript_v3/04-discussion.qmd
Updated months, fix descriptive table as September still appears

--- 2023-03-26 manuscript_v2/04-discussion.qmd
@@
 
-There existence of healthcare disparities within Ontario is a topic of particular interest in due to the recent change in the healthcare system of this region, which eliminated the LHIN model and adopted larger Health Regions system in late 2019 in an attempt to address the disparities of the former approach [@tsasis2012;@dong2022].
+There existence of healthcare disparities within Ontario is a topic of particular interest as it motivated the recent change in the healthcare system of this region, which a system based on Health Regions in late 2019 in an attempt to address the disparities from the LHIN model  [@tsasis2012;@dong2022].
 In this context, analyzing COVID-19 vaccination estimates within the province is important as they can provide an initial assessment of variations that might need to be addressed by decision-makers.
@@
 Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats).
-It is worth mentioning that province-wide vaccination rates for the period of interest are somewhat different from those of the survey, particularly in the case of those 55 years of age and older, which in the survey had a vaccination rate of 72\%, against a rate of 88.4\% reported for the closest age bracket (50 years of age and older) reported by Public Health Ontario  at the start of the period covered by the data (between September 30th, 2021 and December 12th, 2021) [@ontario-covid].
+It is worth mentioning that province-wide vaccination rates for the period of interest are somewhat different from those of the survey, particularly in the case of those 55 years of age and older, which in the survey had a vaccination rate of 72\%, against a rate of 88.4\% reported for the closest age bracket (50 years of age and older) reported by Public Health Ontario at the start of the period covered by the data (between October 1st, 2021 and December 12th, 2021) [@ontario-covid].
 In this regard, differences are to be expected because the survey represents a random sample from the population, and therefore, the sampling process is likely to cause variations between the values from the survey and province-wide estimates.

### 2023-04-16 manuscript_v3/05-conclusion.qmd
Updated months, fix descriptive table as September still appears

--- 2023-04-10 manuscript_v2/main.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
+# Conclusion
 
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-%line numbers
-%\usepackage{mathpazo}
-\usepackage{lineno}
-\linenumbers
-
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-\KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
-\ifdefined\contentsname
-  \renewcommand*\contentsname{Table of contents}
-\else
-  \newcommand\contentsname{Table of contents}
-\fi
-\ifdefined\listfigurename
-  \renewcommand*\listfigurename{List of Figures}
-\else
-  \newcommand\listfigurename{List of Figures}
-\fi
-\ifdefined\listtablename
-  \renewcommand*\listtablename{List of Tables}
-\else
-  \newcommand\listtablename{List of Tables}
-\fi
-\ifdefined\figurename
-  \renewcommand*\figurename{Figure}
-\else
-  \newcommand\figurename{Figure}
-\fi
-\ifdefined\tablename
-  \renewcommand*\tablename{Table}
-\else
-  \newcommand\tablename{Table}
-\fi
-}
-\@ifpackageloaded{float}{}{\usepackage{float}}
-\floatstyle{ruled}
-\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata},
-  pdfauthor={Ariel Mundo Ortiz1; Bouchra Nasri2,},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-
-\title{\textbf{COVID-19 vaccination in Ontario: Exploring
-intra-provincial variations within Health Regions and socio-economic
-strata}}
-\author{Ariel Mundo Ortiz\textsuperscript{1} \and Bouchra
-Nasri\textsuperscript{2,*}}
-\date{}
-
-\begin{document}
-\maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
-
-\textsuperscript{1} Centre de Recherches Mathématiques, University of
-Montreal, Montréal, Canada\\
-\textsuperscript{2} Department of Social and Preventive Medicine, École
-de Santé Publique, University of Montreal, Montréal, Canada
-
-\textsuperscript{*} Correspondence:
-\href{mailto:bouchra.nasri@umontreal.ca}{Bouchra Nasri
-\textless{}bouchra.nasri@umontreal.ca\textgreater{}}
-
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
-
-The COVID-19 pandemic continues to be a worldwide public health concern.
-Although vaccines against this disease were rapidly developed,
-vaccination uptake has not ben equal across all the segments of the
-population. In particular, it has been shown that there have been
-differences in vaccine uptake across different segments of the
-population. However, there are also differences in vaccination across
-geographical areas, which might be important to consider in the
-development of future public health policies against COVID-19. In this
-study, we examined the relationship between vaccination status (having
-received the first dose of a COVID-19 vaccine), and different
-socio-economic and geographical factors. Our results show differences in
-vaccination due to race/ethnicity, income, Health Regions (geographical
-areas used for health service access in Ontario), and their
-interactions. In particular, we show that individuals who identified as
-Arab/Middle Eastern, Black, or Latin American, had significantly lower
-odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32,
-0.28, and \emph{p}=0.004, \emph{p}\textless0.001 and \emph{p}=0.004,
-respectively), and that individuals with a household income below CAD
-25,000 who identified as Arab/Middle Eastern (OR=3.05, \emph{p}=0.013),
-Black (OR=3.19, \emph{p}=0.004), Latin American (OR=2.80,
-\emph{p}=0.041), or that belonged to other minority groups (OR=4.59,
-\emph{p}\textless0.001) had higher odds of vaccination than individuals
-from the same racial/ethnic group in higher income brackets. Finally, we
-also identified lower odds of vaccination within certain minority groups
-in the West Health Region, which comprises the regions of Waterloo and
-Niagara, the counties of Wellington, Essex and Lambton, and the cities
-of Hamilton, Haldimand, Brant, and Chatham-Kent. This study shows that
-there is an ongoing need to better understand and address differences in
-vaccination uptake across diverse segments of the population that have
-been largely impacted by the pandemic.
-
-\hypertarget{keywords}{%
-\section*{Keywords}\label{keywords}}
-\addcontentsline{toc}{section}{Keywords}
-
-Covid-19, vaccination, survey, socio-economic factors, visible
-minorities.
-
-\hypertarget{background}{%
-\section{Background}\label{background}}
-
-The vaccines against COVID-19 have been considered a major achievement
-of modern medicine as their rapid development allowed the start of broad
-vaccination campaigns towards the end of 2020 in certain countries, such
-as the US and
-Canada\textsuperscript{\protect\hyperlink{ref-davis2022}{1}--\protect\hyperlink{ref-tanne2020}{3}}.
-This made some believe that vaccines were destined to be a determinant
-factor in a rapid ending of the
-pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{4}}.
-However, although it has been estimated that COVID-19 vaccines have
-prevented around 14 million of deaths
-worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{5}}, their
-implementation has been far from being equal to that of the vaccines of
-smallpox and polio, which were implemented on a global scale and that
-were indeed crucial to control these
-diseases\textsuperscript{\protect\hyperlink{ref-kayser2021}{6}}. In
-fact, the rollout of COVID-19 vaccines has faced multiple challenges
-since its inception which ultimately have hampered their use to achieve
-the ultimate goal of global immunity.
-
-This problematic of COVID-19 vaccines rollout is a multifaceted issue
-resulting from, among other things, the development of new variants due
-to inadequate public health
-measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
-in vaccine access between high- and low-income
-countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{8},\protect\hyperlink{ref-tamey2022}{9}},
-vaccine
-hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{10}}, and
-differences in vaccination uptake across different segments of the
-population\textsuperscript{\protect\hyperlink{ref-malik2020}{11}}. In
-particular, it is well established that differences in vaccination
-uptake have been present even in countries that have had ample access to
-vaccines since 2020 (such as the US, the UK, and Canada), where lower
-vaccine uptake has been observed within racial minorities (i.e.,
-individuals that identify as Black, Asian, or Indigenous), and in
-individuals within low income
-brackets\textsuperscript{\protect\hyperlink{ref-willis2021}{12}--\protect\hyperlink{ref-khubchandani2021}{15}}.
-Reasons given for lower vaccine uptake in these cases have included
-medical mistrust due to systemic medical
-racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{14}}, mistrust
-in vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{12}}, and
-the influence of conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
-Moreover, in the case of Canada, lower vaccine uptake has been observed
-in young individuals, those with a low educational level, households
-with children, those without a regular healthcare provider, individuals
-that identify as part of a visible minorities or Indigenous, and those
-that are financially
-unstable\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.
-
-However, it is important to consider that vaccination uptake can also be
-influenced by geographical (spatial) factors. In this regard,
-differences in COVID-19 vaccination rates have been associated with
-varied regional attitudes towards
-vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{11}},
-spatial differences in vaccine access and supply, vaccination location
-availability, and lack of prioritization of areas where vulnerable
-groups
-reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2},\protect\hyperlink{ref-nguyen2021}{22}}.
-Other studies have also shown heterogeneity in vaccine uptake within
-small governmental administrative units such as
-counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{23}--\protect\hyperlink{ref-bhuiyan2022}{26}},
-and that and that accounting for geographical differences in vaccination
-can help predict patterns of booster
-uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{27}}. Overall,
-the evidence provided by the literature demonstrates the existence of
-spatially-driven heterogeneities in vaccine uptake that be used by
-decision-makers in the development of public health policies that are
-focused on addressing these disparities within specific administrative
-or geographical areas.
-
-However, such analyses have been carried mostly in territories outside
-of Canada, where available studies have been focused in certain cities
-(such as Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}},
-or Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}),
-or have explored differences at a province-wide
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. Thus, there
-is a need for studies that explore spatial differences in vaccination
-within the Canadian territory and that consequently, can help identify
-disparities that need to be addressed within specific areas in each
-province.
-
-This need is particularly important in the case of Ontario, the most
-populated province in Canada. Between 2006 and 2019, Ontario provided
-healthcare access to its inhabitants using 14 intra-provincial divisions
-called the Local Health Integrated Networks (LHINs). However, this
-approach was complex, bureaucratic, and led to systemic
-inequalities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. In
-late 2019, the 14 LHINs were phased out and the areas they covered were
-incorporated into 6 Health Regions (North East, North West, Central,
-Toronto, West, and East) in an effort to improve the healthcare system
-of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
-Because the adoption of the Health Regions occurred at a relatively
-recent time, there is an ongoing need to analyze the impact of this
-measure and identify the existence of intra-regional differences that
-might exist, and which could be specially important in the context of
-the COVID-19 pandemic.
-
-In this study, we analyzed differences in self-reported COVID-19
-vaccination status in Ontario using socio-economic (e.g., income,
-racial/ethnic identification), and spatial information at the level of
-the Health Regions to identify the existence of differences that might
-need to be addressed to ensure that the healthcare system of Ontario is
-more inclusive and that responds to the needs of its most vulnerable
-population.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{sec-data}{%
-\subsection{Data}\label{sec-data}}
-
-We used data from the \emph{Survey of COVID-19 related Behaviours and
-Attitudes}, a repeated cross sectional survey focused on the Canadian
-province of Ontario that was commissioned by the Fields Institute for
-Research in Mathematical Sciences (henceforth Fields) and the
-Mathematical Modelling of COVID-19 Task Force under ethical guidance
-from the University of Toronto, and which ran between September 30th,
-2021 and January 17th,2022. The survey collected socio-economic
-information from participants (Table~\ref{tbl-covariates}), recorded
-their location (using the nearest municipality), and asked information
-on vaccination status by using the question ``Have you received the
-first dose of the COVID vaccine?'', with possible answers ``yes'' and
-``no''. The original dataset contained 39,029 entries (where each entry
-corresponded to a unique respondent).
-
-This dataset was cleaned to remove outliers that were identified during
-preliminary analyses, and processing the geographical information in the
-survey (city where the survey was responded) in order to match each city
-to its correspondent Health Region.
-
-\hypertarget{tbl-covariates}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3014}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6986}}@{}}
-\caption{\label{tbl-covariates}Socio-economic factors from the Fields
-COVID-19 survey}\tabularnewline
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Variable
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Levels
-\end{minipage} \\
-\midrule()
-\endfirsthead
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Variable
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Levels
-\end{minipage} \\
-\midrule()
-\endhead
-Age group & 16-34,35-54,55 and over \\
-Income bracket (CAD) & under 25,000, 25,000-59,999, 60,000 and above \\
-Race/ethnicity & Arab/Middle Eastern, Black, East Asian/Pacific
-Islander, Indigenous, Latin American, Mixed, South Asian, White
-Caucasian, Other \\
-\bottomrule()
-\end{longtable}
-
-The clean dataset contained responses from more than 200 different
-municipalities within Ontario (Figure~\ref{fig-map}). Because of the
-lack of a publicly available list of all municipalities within each
-Health Region, we used a dataset of long-term care homes and LHINs to
-match each city to LHIN, followed by matching each LHIN to a Health
-Region following the information provided on the Ontario Health Website,
-where the list of LHINs and corresponding Health Regions is available.
-In the case of municipalities that did not appear in the long-term care
-home dataset, we manually searched each city in the LHINs websites in
-order to provide geographical information. The original dataset, clean
-dataset, and details on the data cleaning process are described in
-detail in the GitHub repository for this paper, which can be found at
-\url{https://github.com/aimundo/Fields_COVID-19/}.
-
-Following an assessment of the number of entries corresponding to each
-Health Region in the final dataset, only 107 observations (4.3\% of the
-total) corresponded to cities located in the North West and North East
-Health Regions. The low representation of these Health Regions in the
-dataset is noticeable in Figure~\ref{fig-map},which shows that responses
-from these areas came from a relatively low number of cities when
-compared to the most populated Health Regions, such as the Toronto or
-Central Regions. We omitted the North East and North East Health Regions
-from further analyses due to the low number of entries. Therefore, the
-total number of unique entries used for analysis was 3,551 which
-included the East, Central, Toronto, and West Health Regions and that
-covered the period between September 30th,2021 and December 12th, 2021.
-
-\begin{figure}
-
-{\centering \includegraphics{../data/map_data/map.pdf}
-
-}
-
-\caption{\label{fig-map}Geographic representation of the data collected
-by the \emph{Survey of COVID-19 related Behaviours and Attitudes},
-collected by the Fields Institute in Ontario. The municipalities
-(cities) from where survey participants provided answers (in the clean
-dataset) appear as points. The Health six Regions are color-coded.
-Internal boundaries within certain Health Regions indicate areas that
-belonged to the Local Integrated Health Networks (LHINs), the geographic
-areas for healthcare in Ontario before the adoption of the Health
-Regions.}
-
-\end{figure}
-
-\hypertarget{statistical-analyses}{%
-\subsection{Statistical analyses}\label{statistical-analyses}}
-
-We used a logistic regression model to estimate the probability of
-vaccination depending on the socio-economic factors described in
-Table~\ref{tbl-covariates}, the Health Regions from Ontario indicated in
-Section~\ref{sec-data}, and the interactions between Race and Health
-Region, and Race and income, as previous studies have shown that
-socio-economic factors and their interactions are significant predictors
-of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
-
-The model was fitted first to the clean dataset to obtain uncorrected
-estimates. Additionally, because we identified differences between the
-proportions of all the socio-economic factors included in the analysis
-(Table~\ref{tbl-covariates}) and the Census data for Ontario, we used an
-iterative proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{35}}
-to correct the data using Census socio-economic data and Health Region
-population totals, in order to obtain corrected estimates from the
-model. Details regarding the correction can be found in the Appendix.
-All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
-and \texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}}.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{survey-results}{%
-\subsection{Survey Results}\label{survey-results}}
-
-Table~\ref{tbl-descriptive-stats} shows the descriptive statistics from
-the Fields COVID-19 survey data for vaccination status and each of the
-covariates analyzed. The total number of entries analyzed was 3,551.
-Overall, 26.9\% of survey respondents (958) reported not having received
-the first dose of the vaccine, whereas 73.1\% (2,593) reported having
-received it. Within each socio-economic factor, respondents who reported
-living in a household with an income under CAD 25,000 represented 37\%
-of the total number of entries, those within the CAD 25,000-59,999
-income bracket represented 25\% of the total sample, and those with an
-income above CAD 60,000 represented 38 \% of the sample; across all
-income brackets, the percentage of individuals that reported having
-received a first dose of the vaccine was consistent, above 69\%.
-
-Within the age groups of survey respondents, the age group between 16-34
-years had the highest representation in the survey responses (1,521,
-42.8\% of all responses). Within this age bracket, 73\% of respondents
-indicated having received the vaccine, whereas the lowest vaccination
-rate was in the bracket of those 55 years of age and above, with a total
-of 72\%. The Health Region with highest representation in the survey was
-Toronto, accounting for 1,324 entries (37.2\%), with a vaccination rate
-of 72\%. Regarding race/ethnicity, individuals that identified as
-White/Caucasian represented 1313 (37\%) of all entries and had the
-highest vaccination uptake with 82\% of them indicating to have received
-the COVID-19 vaccine. On the other hand, the ethnic group with the
-lowest number of entries in the survey was Latin American, with a total
-of 180, or 5\% of all entries. Vaccination rates across all minority
-groups were below the value reported by White/Caucasians, with the
-lowest vaccination rate (60\%) being reported by individuals that
-identified as Indigenous.
-
-\hypertarget{tbl-descriptive-stats}{}
-\begin{table}
-\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination
-Status) }\tabularnewline
-
-\centering\begingroup\fontsize{9}{11}\selectfont
-
-\begin{tabular}{lcc}
-\toprule
-\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,593\\
-\midrule
-\textbf{Income} &  & \\
-\hspace{1em}60000\_and\_above & 305 (23\%) & 1,049 (77\%)\\
-\hspace{1em}25000\_59999 & 253 (28\%) & 636 (72\%)\\
-\hspace{1em}under\_25000 & 400 (31\%) & 908 (69\%)\\
-\textbf{Age Group} &  & \\
-\hspace{1em}16\_34 & 409 (27\%) & 1,112 (73\%)\\
-\hspace{1em}35\_54 & 252 (26\%) & 712 (74\%)\\
-\hspace{1em}55\_and\_over & 297 (28\%) & 769 (72\%)\\
-\textbf{Health Region} &  & \\
-\hspace{1em}Toronto & 371 (28\%) & 953 (72\%)\\
-\hspace{1em}Central & 224 (28\%) & 581 (72\%)\\
-\hspace{1em}East & 135 (23\%) & 448 (77\%)\\
-\hspace{1em}West & 228 (27\%) & 611 (73\%)\\
-\textbf{Race} &  & \\
-\hspace{1em}white\_caucasian & 233 (18\%) & 1,080 (82\%)\\
-\hspace{1em}arab\_middle\_eastern & 76 (36\%) & 138 (64\%)\\
-\hspace{1em}black & 114 (38\%) & 184 (62\%)\\
-\hspace{1em}east\_asian\_pacific\_islander & 69 (23\%) & 234 (77\%)\\
-\hspace{1em}indigenous & 76 (40\%) & 115 (60\%)\\
-\hspace{1em}latin\_american & 69 (38\%) & 111 (62\%)\\
-\hspace{1em}mixed & 105 (34\%) & 205 (66\%)\\
-\hspace{1em}other & 128 (35\%) & 239 (65\%)\\
-\hspace{1em}south\_asian & 88 (23\%) & 287 (77\%)\\
-\bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
-\end{tabular}
-\endgroup{}
-\end{table}
-
-\hypertarget{multivariate-regression}{%
-\subsection{Multivariate Regression}\label{multivariate-regression}}
-
-Table~\ref{tbl-model} shows the results of the logistic regression
-models (for the uncorrected and corrected data) on vaccination status
-using socio-economic factors (age group, income, race), geographical
-areas (Health Regions) and the interactions between income and race and
-Health Region and race. There were no statistically significant
-differences in vaccination rates within the age groups from the survey,
-but significant odds ratios were estimated for other covariates. Within
-household income brackets, individuals with an income under CAD 25,000
-or between CAD 25,000-59,999 had significantly lower odds of vaccination
-than those with an income above CAD 60,000 (ORs=0.37 and 0.59,
-\emph{p}=0.011 and \textless0.001, respectively). Within Race/Ethnicity,
-individuals who identified as Arab/Middle Eastern, Black, or Latin
-American, had significantly lower odds of vaccination than those in the
-White/Caucasian group (ORs=0.31, 0.32, 0.28, and
-\emph{p}=0.004,\textless0.001 and 0.004, respectively); additionally,
-those individuals in the Other Race/Ethnicity group (a group that
-included Southeast Asian, Filipino, West Asian, and Minorities Not
-Identified Elsewhere) had even lower odds of vaccination than the other
-minority groups (OR=0.22, \emph{p}\textless0.001). Regarding Health
-Regions, individuals that reported living in the West Health Region
-(which comprises the regions of Waterloo and Niagara, the counties of
-Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
-Brant, and Chatham-Kent) had significantly higher odds of vaccination
-than those in the Health Region of Toronto (OR=1.55, \emph{p}=0.029).
-
-Moreover, statistically-significant odd ratios were determined in the
-case of the interaction of income and race; specifically, for
-individuals with a household income below CAD 25,000 who identified as
-Arab/Middle Eastern (OR=3.05, \emph{p}=0.013), Black (OR=3.19,
-\emph{p}=0.004), Latin American (OR=2.80, \emph{p}=0.041), or that
-belonged to other minority groups (OR=4.59, \emph{p}\textless0.001).
-Within the CAD 25,000-59,999 income bracket, individuals who identified
-as belonging to other racial minority groups had significantly higher
-odds of vaccination (OR=6.93, \emph{p}\textless0.001).
-
-For the interaction of Health Region and race, significant odds of
-vaccination were identified for Black individuals in the Central Health
-Region, which comprises the region of York, counties of Dufferin and
-Simcoe and the district of Muskoka (OR=0.44, \emph{p}=0.046), and in
-individuals that identified as part of other racial minorities or South
-Asian that lived in the West Health Region (ORs=0.41, \emph{p}=0.032 and
-\emph{p}=0.037, respectively).
-
-\footnotesize
-\renewcommand{\arraystretch}{0.5}
-
-\hypertarget{tbl-model}{}
-\begin{longtable}{lcccccc}
-\caption{\label{tbl-model}Multiple Regression Analysis-Predictors of Vaccination Status }\tabularnewline
-
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{\textbf{Uncorrected}} & \multicolumn{3}{c}{\textbf{Corrected}} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
-\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
-\midrule
-\endfirsthead
-\multicolumn{7}{@{}l}{\textit{(continued)}}\\
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{\textbf{Uncorrected}} & \multicolumn{3}{c}{\textbf{Corrected}} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
-\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
-\midrule
-\endhead
-
-\endfoot
-\bottomrule
-\endlastfoot
-\textbf{Age Group} &  &  &  &  &  & \\
-\hspace{1em}16\_34 & — & — &  & — & — & \\
-\hspace{1em}35\_54 & 0.93 & 0.77, 1.13 & 0.5 & 0.90 & 0.67, 1.21 & 0.5\\
-\hspace{1em}55\_and\_over & 0.74 & 0.61, 0.89 & 0.002 & 0.99 & 0.74, 1.32 & >0.9\\
-\textbf{Income} &  &  &  &  &  & \\
-\hspace{1em}60000\_and\_above & — & — &  & — & — & \\
-\hspace{1em}25000\_59999 & 0.59 & 0.41, 0.84 & 0.004 & 0.59 & 0.39, 0.89 & 0.011\\
-\hspace{1em}under\_25000 & 0.41 & 0.29, 0.58 & <0.001 & 0.37 & 0.25, 0.56 & <0.001\\
-\textbf{Race} &  &  &  &  &  & \\
-\hspace{1em}white\_caucasian & — & — &  & — & — & \\
-\hspace{1em}arab\_middle\_eastern & 0.24 & 0.12, 0.50 & <0.001 & 0.31 & 0.14, 0.69 & 0.004\\
-\hspace{1em}black & 0.30 & 0.17, 0.54 & <0.001 & 0.32 & 0.17, 0.60 & <0.001\\
-\hspace{1em}east\_asian\_pacific\_islander & 0.74 & 0.36, 1.52 & 0.4 & 1.15 & 0.50, 2.66 & 0.7\\
-\hspace{1em}indigenous & 0.37 & 0.17, 0.81 & 0.013 & 0.44 & 0.19, 1.02 & 0.056\\
-\hspace{1em}latin\_american & 0.28 & 0.13, 0.59 & <0.001 & 0.28 & 0.11, 0.67 & 0.004\\
-\hspace{1em}mixed & 0.59 & 0.31, 1.12 & 0.11 & 0.64 & 0.25, 1.65 & 0.4\\
-\hspace{1em}other & 0.20 & 0.11, 0.35 & <0.001 & 0.22 & 0.12, 0.41 & <0.001\\
-\hspace{1em}south\_asian & 0.80 & 0.44, 1.45 & 0.5 & 0.91 & 0.49, 1.69 & 0.8\\
-\textbf{Health Region} &  &  &  &  &  & \\
-\hspace{1em}Toronto & — & — &  & — & — & \\
-\hspace{1em}Central & 1.30 & 0.85, 2.00 & 0.2 & 1.47 & 0.92, 2.35 & 0.11\\
-\hspace{1em}East & 1.54 & 1.01, 2.34 & 0.044 & 1.42 & 0.90, 2.23 & 0.13\\
-\hspace{1em}West & 1.35 & 0.95, 1.94 & 0.10 & 1.55 & 1.05, 2.30 & 0.029\\
-\textbf{Income * Race} &  &  &  &  &  & \\
-\hspace{1em}25000\_59999 * arab\_middle\_eastern & 2.16 & 0.93, 4.99 & 0.072 & 1.79 & 0.67, 4.83 & 0.2\\
-\hspace{1em}under\_25000 * arab\_middle\_eastern & 2.96 & 1.39, 6.26 & 0.005 & 3.05 & 1.26, 7.39 & 0.013\\
-\hspace{1em}25000\_59999 * black & 1.19 & 0.60, 2.39 & 0.6 & 1.34 & 0.59, 3.05 & 0.5\\
-\hspace{1em}under\_25000 * black & 2.88 & 1.48, 5.59 & 0.002 & 3.19 & 1.45, 6.99 & 0.004\\
-\hspace{1em}25000\_59999 * east\_asian\_pacific\_islander & 0.95 & 0.44, 2.07 & >0.9 & 0.42 & 0.17, 1.05 & 0.062\\
-\hspace{1em}under\_25000 * east\_asian\_pacific\_islander & 1.91 & 0.90, 4.04 & 0.090 & 1.16 & 0.47, 2.86 & 0.8\\
-\hspace{1em}25000\_59999 * indigenous & 1.81 & 0.74, 4.43 & 0.2 & 1.36 & 0.48, 3.89 & 0.6\\
-\hspace{1em}under\_25000 * indigenous & 1.64 & 0.73, 3.69 & 0.2 & 1.45 & 0.55, 3.80 & 0.5\\
-\hspace{1em}25000\_59999 * latin\_american & 0.89 & 0.38, 2.10 & 0.8 & 1.24 & 0.45, 3.43 & 0.7\\
-\hspace{1em}under\_25000 * latin\_american & 3.09 & 1.33, 7.16 & 0.009 & 2.80 & 1.04, 7.51 & 0.041\\
-\hspace{1em}25000\_59999 * mixed & 0.86 & 0.39, 1.93 & 0.7 & 0.85 & 0.32, 2.26 & 0.7\\
-\hspace{1em}under\_25000 * mixed & 1.26 & 0.64, 2.47 & 0.5 & 1.10 & 0.37, 3.27 & 0.9\\
-\hspace{1em}25000\_59999 * other & 5.46 & 2.41, 12.3 & <0.001 & 6.93 & 2.65, 18.1 & <0.001\\
-\hspace{1em}under\_25000 * other & 4.06 & 2.25, 7.31 & <0.001 & 4.59 & 2.33, 9.05 & <0.001\\
-\hspace{1em}25000\_59999 * south\_asian & 1.13 & 0.54, 2.36 & 0.7 & 1.20 & 0.51, 2.85 & 0.7\\
-\hspace{1em}under\_25000 * south\_asian & 1.59 & 0.83, 3.06 & 0.2 & 2.00 & 0.93, 4.30 & 0.077\\
-\textbf{Race * Health Region} &  &  &  &  &  & \\
-\hspace{1em}arab\_middle\_eastern * Central & 0.75 & 0.33, 1.71 & 0.5 & 0.66 & 0.26, 1.70 & 0.4\\
-\hspace{1em}black * Central & 0.48 & 0.23, 1.01 & 0.055 & 0.44 & 0.19, 0.98 & 0.046\\
-\hspace{1em}east\_asian\_pacific\_islander * Central & 0.84 & 0.37, 1.88 & 0.7 & 0.98 & 0.38, 2.53 & >0.9\\
-\hspace{1em}indigenous * Central & 0.60 & 0.24, 1.51 & 0.3 & 0.63 & 0.22, 1.79 & 0.4\\
-\hspace{1em}latin\_american * Central & 0.69 & 0.28, 1.72 & 0.4 & 0.67 & 0.23, 1.96 & 0.5\\
-\hspace{1em}mixed * Central & 0.63 & 0.29, 1.35 & 0.2 & 0.73 & 0.24, 2.22 & 0.6\\
-\hspace{1em}other * Central & 0.96 & 0.48, 1.92 & >0.9 & 0.80 & 0.36, 1.78 & 0.6\\
-\hspace{1em}south\_asian * Central & 0.70 & 0.34, 1.45 & 0.3 & 0.54 & 0.25, 1.20 & 0.13\\
-\hspace{1em}arab\_middle\_eastern * East & 0.58 & 0.20, 1.65 & 0.3 & 0.43 & 0.13, 1.45 & 0.2\\
-\hspace{1em}black * East & 0.82 & 0.36, 1.86 & 0.6 & 0.83 & 0.34, 2.04 & 0.7\\
-\hspace{1em}east\_asian\_pacific\_islander * East & 0.82 & 0.30, 2.24 & 0.7 & 0.86 & 0.29, 2.56 & 0.8\\
-\hspace{1em}indigenous * East & 0.55 & 0.21, 1.44 & 0.2 & 0.69 & 0.23, 2.08 & 0.5\\
-\hspace{1em}latin\_american * East & 0.79 & 0.28, 2.23 & 0.7 & 1.03 & 0.32, 3.34 & >0.9\\
-\hspace{1em}mixed * East & 0.71 & 0.31, 1.60 & 0.4 & 0.91 & 0.28, 3.03 & 0.9\\
-\hspace{1em}other * East & 0.86 & 0.37, 1.99 & 0.7 & 1.05 & 0.39, 2.83 & >0.9\\
-\hspace{1em}south\_asian * East & 0.50 & 0.20, 1.24 & 0.14 & 0.52 & 0.19, 1.45 & 0.2\\
-\hspace{1em}arab\_middle\_eastern * West & 1.16 & 0.48, 2.77 & 0.7 & 1.00 & 0.37, 2.73 & >0.9\\
-\hspace{1em}black * West & 0.77 & 0.36, 1.65 & 0.5 & 0.76 & 0.32, 1.80 & 0.5\\
-\hspace{1em}east\_asian\_pacific\_islander * West & 0.54 & 0.24, 1.20 & 0.13 & 0.52 & 0.20, 1.34 & 0.2\\
-\hspace{1em}indigenous * West & 0.44 & 0.19, 1.02 & 0.056 & 0.39 & 0.14, 1.09 & 0.073\\
-\hspace{1em}latin\_american * West & 0.97 & 0.39, 2.43 & >0.9 & 0.94 & 0.32, 2.72 & >0.9\\
-\hspace{1em}mixed * West & 0.53 & 0.25, 1.11 & 0.092 & 0.37 & 0.12, 1.16 & 0.089\\
-\hspace{1em}other * West & 0.55 & 0.27, 1.12 & 0.10 & 0.41 & 0.18, 0.93 & 0.032\\
-\hspace{1em}south\_asian * West & 0.50 & 0.23, 1.07 & 0.075 & 0.41 & 0.18, 0.95 & 0.037\\*
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\
-\end{longtable}
-
-\normalsize
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-There existence of healthcare disparities within Ontario is a topic of
-particular interest in due to the recent change in the healthcare system
-of this region, which eliminated the LHIN model and adopted larger
-Health Regions system in late 2019 in an attempt to address the
-disparities of the former
-approach\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
-In this context, analyzing COVID-19 vaccination estimates within the
-province is important as they can provide an initial assessment of
-variations that might need to be addressed by decision-makers.
-
-Our results indicate that across the most densely populated Health
-Regions of Ontario, almost three quarters of the surveyed individuals
-reported having received the first dose of the COVID-19 vaccine
-(Table~\ref{tbl-descriptive-stats}). It is worth mentioning that
-province-wide vaccination rates for the period of interest are somewhat
-different from those of the survey, particularly in the case of those 55
-years of age and older, which in the survey had a vaccination rate of
-72\%, against a rate of 88.4\% reported for the closest age bracket (50
-years of age and older) reported by Public Health Ontario at the start
-of the period covered by the data (between September 30th, 2021 and
-December 12th,
-2021)\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. In
-this regard, differences are to be expected because the survey
-represents a random sample from the population, and therefore, the
-sampling process is likely to cause variations between the values from
-the survey and province-wide estimates.
-
-We believe that these differences did not have a significant impact in
-our analyses, because our results indicate that there were no
-significant differences in vaccination odds among the age groups
-considered in the survey, in agreement with the overall vaccination
-rates reported for Canada, which have been relatively higher when
-compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{40}}, and
-with vaccination uptake rates across different age groups presented in
-other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{41}}.
-Moreover, vaccination rates within each age in the dataset are in good
-agreement with province-wide estimates (e.g., a rate of 95\% for those
-with 61 years of age, Supplementary Table A-6). Additionally, when
-overall vaccination rates for the province are disaggregated, it can be
-seen that regional differences were present during the period analyzed,
-which can be masked by province-wide estimates. For example, the Public
-Health Unit of Lambton (a region within the West Health Region) and the
-Public Health Unit of Haliburton, Kawartha, and the Pine Ridge District
-(an area covered by the East Health Region) reported lower vaccination
-rates (78\%) for those 50 years of age and older at the beginning of the
-period of interest, whereas other regions had higher vaccination
-rates\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. To
-this day, differences in vaccination rates within the province are
-present, because according to Public Health Ontario, as of March of 2023
-some regions have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{42}}.
-Overall, these reasons indicate that the estimates obtained from the
-data are in good agreement with the trends from the population.
-
-We identified significant intra-provincial differences in vaccination
-based on socio-economic and geographical factors. First, our results
-show differences in odds of vaccination in individuals with a household
-income below CAD 60,000 and in individuals belonging to visible minority
-groups. Those who identified as Arab/Middle Eastern, Black, Latin
-American, or that belonged to a minority group not included in the
-survey (Southeast Asian, Filipino, West Asian, and minority groups not
-identified elsewhere) had vaccination odds that were less than a third
-of individuals that identified as White/Caucasian
-(Table~\ref{tbl-model}). These results are consistent with other studies
-that have shown lower vaccination rates in individuals that identify as
-part of a racial minority, or that have a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{43}}.
-
-In this study, we also decided to explore the interactions between
-income and race and race and Health Region, as it is known that many
-individuals within racial minority groups perform tend to occupy certain
-types of occupations that fall within income brackets that have been
-shown to be associated with differences in vaccination uptake. In other
-words, we decided to explore if there were differences in vaccination
-within racial groups in certain income brackets and in certain the
-Health Regions. In this regard, it is interesting to note that although
-overall self-reported vaccination rates were found to be statistically
-significantly lower in various racial minority groups when compared to
-White/Caucasian individuals (Table~\ref{tbl-model}), the change in odds
-of vaccination within certain racial groups and income strata was
-actually positive, in contrast to the White/Caucasian group, for which
-vaccination odds decreased in lower income brackets (when compared to
-the CAD 60,000 and over bracket, Supplementary Figure A-3). More
-specifically, the change in odds of vaccination increased in individuals
-who identified as Arab/Middle Eastern, Black, Latin American, or
-belonging to other minority groups with a household income below CAD
-25,000, which was also true for individuals in other racial minority
-groups with an income between CAD 25,000-59,999 (Table~\ref{tbl-model},
-Supplementary Figure A-3).
-
-This result is likely due to the fact that individuals that belong
-racial minority groups tend to perform occupations that have been deemed
-as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{44},\protect\hyperlink{ref-ct2021}{45}},
-which include occupations such as grocery store workers, gas station
-workers, warehouse and distribution workers, and manufacturing workers,
-all being occupations for which an income within the significant
-brackets is to be expected. In the case of Ontario, essential workers
-had priority for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{46}},
-which would explain the higher odds of vaccination for these individuals
-in certain income brackets, in contrast to the lower odds of vaccination
-for the same type of individuals with higher household income. In other
-words, it is possible that the type of occupation played an important
-role in increasing the odds of vaccination in these racial minority
-groups.
-
-Additionally, significant higher vaccination odds were identified in the
-West Health Region when compared to the Health Region of Toronto
-(Table~\ref{tbl-model}). The West Health Region comprises the regions of
-Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and
-the cities of Hamilton, Haldimand, Brant, and Chatham-Kent. In this
-case, a possible rationale for the results is the fact that in the
-survey, about 47\% of the entries for this Health Region corresponded to
-White/Caucasian individuals, who reported an overall 83\% vaccination
-rate (Supplementary Table A-7). However, the interaction effect of
-Health Region and race was also significant in the case of individuals
-identifying as South Asian or other minorities not included in the
-survey Table~\ref{tbl-model}. In this case, the results of the
-interaction term in the model indicate that the odds of vaccination for
-those within the South Asian and Other minority groups in the West
-Region decreased when compared to the other Health Regions
-(Supplementary Figure A-4).
-
-According to Ontario Health, 13.2\% of the population in the West Health
-Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{47}}.
-In the case of this analysis, the estimated lower odds are likely to be
-explained from a socio-economic perspective. In fact, 50\% of the
-answers from this region in the survey came from the former LHINs of
-Hamilton Niagara Haldimand Brant, and Erie St.~Clair, both which are
-among the regions of Ontario with the highest proportion of their
-population (more than 20\%) in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{48}}
-(Supplementary Table A-8). Therefore, this result partly reinforces the
-well-known existing association between low vaccination rates and
-income, but it additionally indicates that there were intra-regional
-differences in vaccination. Interestingly, a disproportionate number of
-COVID-19 cases and low vaccination rate (under 50\%) have been
-previously reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{49}}; in this
-regard, our result provides additional context by showing that within
-the South Asian community, there were differences in vaccination uptake
-across Ontario. Moreover, because significant lower odds of vaccination
-were also identified other minority groups, this provides a rationale
-for future studies that explore how vaccination uptake varies across
-different minority groups within Ontario and other Canadian provinces.
-
-There are some limitations to the present study. First, the data
-collection design, which allowed respondents to withdraw from the survey
-at any point, resulted in a high number of unique entries in the survey
-with multiple missing answers. Because we focused on entries that had
-complete observations in the covariates of interest for our analysis, it
-is possible that some information was not considered by excluding
-observations that had information in other variables (such as work from
-home, or number of persons in the household). However, we attempted to
-minimize this possibility by correcting the dataset using information
-from the Census. More granular corrections, which for example could be
-based on demographic information by municipality, could be used in the
-future to obtain a more accurate approximation to the population totals
-of the province. Moreover, our analysis did not consider the North West
-and North East Health Regions, due to the low number of entries from
-these areas in the survey (Figure~\ref{fig-map}). Although low
-representation from these areas is based on the fact that these regions
-only account for 5\% of the total population of Ontario, these regions
-are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{47}}, a
-minority group that has historically suffered from reduced access to
-health care and
-discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{17}}.
-Therefore, there is a need for additional studies that focus on these
-low-populated Health Regions in Ontario where disparities in vaccination
-might be significant and understudied.
-
-The results in this study are based on self-reported data, where there
-is a risk that biased values are reported. Despite this, because in the
-context of COVID-19 it has been shown that good agreement exists between
-self-reported and documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{50}}, and
-therefore, the effect of self-reported bias is likely to not be
-significant in our analyses. Finally, it is likely that there have been
-differences in vaccination across the province as more doses of the
-vaccine were administered and as successive variants emerged. Because
-this study focused only on vaccination status regarding the first dose
-of the vaccine within a relatively short time window, it can only
-provide a snapshot of the societal dynamics behind the pandemic.
-Nonetheless, the results presented here can serve as a starting point to
-motivate the collection of robust longitudinal data that can be used to
-quantify geographical and temporal differences within vulnerable
-segments of the population, and that can be used to inform the
-development of adequate public health policies within the province of
-Ontario or across other provinces that aim to minimize disparities in
-health access.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This study explored differences in COVID-19 vaccination across the
-province of Ontario between late 2021 and early 2022 by taking into
-consideration socio-economic factors, such as income and race, their
-interactions, and the Health Regions within the province. Our results
-show that, during the period analyzed, significant differences in
-vaccination existed across different visible minority groups, income
-brackets, and Health Regions, showing intra-provincial disparities in
-vaccine uptake. As the COVID-19 continues around the world, it important
-that future public policies take into consideration how to adequately
-reach individuals within minority groups that live across geographical
-areas where less probabilities of being vaccinated are likely. At the
-moment, this is an ongoing issue that needs to be addressed to ensure a
-more homogeneous outcome from the pandemic.
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
-\CSLLeftMargin{1. }%
-\CSLRightInline{Davis CJ, Golding M, McKay R. Efficacy information
-influences intention to take COVID-19 vaccine. \emph{British Journal of
-Health Psychology}. 2022;27(2):300-319.
-doi:\url{https://doi.org/10.1111/bjhp.12546}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
-\CSLLeftMargin{2. }%
-\CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
-social and policy view of vaccination efforts in ontario, canada.
-\emph{Cambridge Journal of Regions, Economy and Society}. Published
-online November 2022.
-doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
-\CSLLeftMargin{3. }%
-\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
-pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
-2020:m4799.
-doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
-\CSLLeftMargin{4. }%
-\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
-end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
-doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-watson2022}{}}%
-\CSLLeftMargin{5. }%
-\CSLRightInline{Watson OJ, Barnsley G, Toor J, Hogan AB, Winskill P,
-Ghani AC. Global impact of the first year of {COVID}-19 vaccination: A
-mathematical modelling study. \emph{The Lancet Infectious Diseases}.
-2022;22(9):1293-1302.
-doi:\href{https://doi.org/10.1016/s1473-3099(22)00320-6}{10.1016/s1473-3099(22)00320-6}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
-\CSLLeftMargin{6. }%
-\CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
-and emerging issues. \emph{Human Vaccines {\&}amp\(\mathsemicolon\)
-Immunotherapeutics}. 2021;17(12):5255-5268.
-doi:\href{https://doi.org/10.1080/21645515.2021.1977057}{10.1080/21645515.2021.1977057}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-li2021}{}}%
-\CSLLeftMargin{7. }%
-\CSLRightInline{Li Q, Wang J, Tang Y, Lu H. Next-generation COVID-19
-vaccines: Opportunities for vaccine development and challenges in
-tackling COVID-19. \emph{Drug Discoveries \& Therapeutics}.
-2021;15(3):118-123.
-doi:\href{https://doi.org/10.5582/ddt.2021.0105}{10.5582/ddt.2021.0105}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
-\CSLLeftMargin{8. }%
-\CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
-determinants of {COVID}-19 vaccine hesitancy. Inbaraj LR, ed.
-\emph{{PLOS} {ONE}}. 2021;16(11):e0258462.
-doi:\href{https://doi.org/10.1371/journal.pone.0258462}{10.1371/journal.pone.0258462}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
-\CSLLeftMargin{9. }%
-\CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
-to achieve global covid-19 vaccine equity. \emph{{BMJ}}. Published
-online March 2022:e070650.
-doi:\href{https://doi.org/10.1136/bmj-2022-070650}{10.1136/bmj-2022-070650}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
-\CSLLeftMargin{10. }%
-\CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
-inequality in {COVID}-19 vaccination coverage among elderly adults in
-england: A national linked data study. \emph{{BMJ} Open}.
-2021;11(7):e053402.
-doi:\href{https://doi.org/10.1136/bmjopen-2021-053402}{10.1136/bmjopen-2021-053402}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
-\CSLLeftMargin{11. }%
-\CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
-of {COVID}-19 vaccine acceptance in the {US}.
-\emph{{EClinicalMedicine}}. 2020;26:100495.
-doi:\href{https://doi.org/10.1016/j.eclinm.2020.100495}{10.1016/j.eclinm.2020.100495}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
-\CSLLeftMargin{12. }%
-\CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
-{COVID}-19 vaccine hesitancy: Race/ethnicity, trust, and fear.
-\emph{Clinical and Translational Science}. 2021;14(6):2200-2207.
-doi:\href{https://doi.org/10.1111/cts.13077}{10.1111/cts.13077}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
-\CSLLeftMargin{13. }%
-\CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
-accepting {COVID}-19 vaccination during and after pregnancy, and for
-their babies: A multi-methods study in the {UK}. \emph{{BMC} Pregnancy
-and Childbirth}. 2022;22(1).
-doi:\href{https://doi.org/10.1186/s12884-021-04321-3}{10.1186/s12884-021-04321-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
-\CSLLeftMargin{14. }%
-\CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
-limits of medical trust in mitigating {COVID}-19 vaccine hesitancy among
-black americans. \emph{Journal of General Internal Medicine}.
-2021;36(11):3629-3631.
-doi:\href{https://doi.org/10.1007/s11606-021-06743-3}{10.1007/s11606-021-06743-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
-\CSLLeftMargin{15. }%
-\CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
-Sharma M, Webb FJ. {COVID}-19 vaccination hesitancy in the united
-states: A rapid national assessment. \emph{Journal of Community Health}.
-2021;46(2):270-277.
-doi:\href{https://doi.org/10.1007/s10900-020-00958-x}{10.1007/s10900-020-00958-x}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
-\CSLLeftMargin{16. }%
-\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
-related medical mistrust, health impacts, and potential vaccine
-hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
-of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
-doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
-\CSLLeftMargin{17. }%
-\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
-roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
-canada. \emph{Canadian Medical Association Journal}.
-2021;193(11):E381-E383.
-doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
-\CSLLeftMargin{18. }%
-\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
-hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
-and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
-2020;52(14):3127-3141.
-doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
-\CSLLeftMargin{19. }%
-\CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
-inequalities in {COVID}-19 vaccination uptake and intent: Results from
-the canadian community health survey 2021. \emph{{BMC} Public Health}.
-2022;22(1).
-doi:\href{https://doi.org/10.1186/s12889-022-14090-z}{10.1186/s12889-022-14090-z}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
-\CSLLeftMargin{20. }%
-\CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
-Fahlman M, Kallio N. {COVID}-19 vaccine hesitancy and refusal and
-associated factors in an adult population in saskatchewan, canada:
-Evidence from predictive modelling. Gesser-Edelsburg A, ed. \emph{{PLOS}
-{ONE}}. 2021;16(11):e0259513.
-doi:\href{https://doi.org/10.1371/journal.pone.0259513}{10.1371/journal.pone.0259513}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
-\CSLLeftMargin{21. }%
-\CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
-Overcoming {COVID}-19 vaccine hesitancy among ethnic minorities: A
-systematic review of {UK} studies. \emph{Vaccine}.
-2022;40(25):3413-3432.
-doi:\href{https://doi.org/10.1016/j.vaccine.2022.04.030}{10.1016/j.vaccine.2022.04.030}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
-\CSLLeftMargin{22. }%
-\CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
-Changes in {COVID}-19 vaccination receipt and intention to vaccinate by
-socioeconomic characteristics and geographic area, united states,
-january 6 {\textendash} march 29, 2021. \emph{Annals of Medicine}.
-2021;53(1):1419-1428.
-doi:\href{https://doi.org/10.1080/07853890.2021.1957998}{10.1080/07853890.2021.1957998}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
-\CSLLeftMargin{23. }%
-\CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
-vaccine hesitancy in the united states. \emph{International Journal of
-Environmental Research and Public Health}. 2021;18(18):9488.
-doi:\href{https://doi.org/10.3390/ijerph18189488}{10.3390/ijerph18189488}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
-\CSLLeftMargin{24. }%
-\CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
-spatial process: {COVID}-19 fully vaccinated rates in u.s. counties.
-\emph{American Journal of Preventive Medicine}. 2022;63(6):954-961.
-doi:\href{https://doi.org/10.1016/j.amepre.2022.06.006}{10.1016/j.amepre.2022.06.006}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
-\CSLLeftMargin{25. }%
-\CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
-the spatiotemporal heterogeneity of the {COVID}-19 vaccination
-landscape. \emph{American Journal of Epidemiology}.
-2022;191(10):1792-1802.
-doi:\href{https://doi.org/10.1093/aje/kwac080}{10.1093/aje/kwac080}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
-\CSLLeftMargin{26. }%
-\CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
-social vulnerability index to assess {COVID}-19 vaccine uptake in
-louisiana. \emph{{GeoJournal}}. Published online December 2022.
-doi:\href{https://doi.org/10.1007/s10708-022-10802-5}{10.1007/s10708-022-10802-5}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
-\CSLLeftMargin{27. }%
-\CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
-future spatial patterns in {COVID}-19 booster vaccine uptake. Published
-online September 2022.
-doi:\href{https://doi.org/10.1101/2022.08.30.22279415}{10.1101/2022.08.30.22279415}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
-\CSLLeftMargin{28. }%
-\CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
-trajectories. \emph{Socius: Sociological Research for a Dynamic World}.
-2021;7:237802312110529.
-doi:\href{https://doi.org/10.1177/23780231211052946}{10.1177/23780231211052946}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
-\CSLLeftMargin{29. }%
-\CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
-Social inequalities in {COVID}-19 vaccine acceptance and uptake for
-children and adolescents in montreal, canada. \emph{Vaccine}.
-2021;39(49):7140-7145.
-doi:\href{https://doi.org/10.1016/j.vaccine.2021.10.077}{10.1016/j.vaccine.2021.10.077}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
-\CSLLeftMargin{30. }%
-\CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
-integrated care: A complex-adaptive systems perspective.
-\emph{International Journal of Integrated Care}. 2012;12(5).
-doi:\href{https://doi.org/10.5334/ijic.843}{10.5334/ijic.843}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
-\CSLLeftMargin{31. }%
-\CSLRightInline{Dong L, Sahu R, Black R. Governance in the
-transformational journey toward integrated healthcare: The case of
-ontario. \emph{Journal of Information Technology Teaching Cases}.
-Published online December 2022:204388692211473.
-doi:\href{https://doi.org/10.1177/20438869221147313}{10.1177/20438869221147313}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{32. }%
-\CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
-Corlin L. Disparities in national and state estimates of {COVID}-19
-vaccination receipt and intent to vaccinate by race/ethnicity, income,
-and age group among adults~\(\geq\)~18~years, united states.
-\emph{Vaccine}. 2022;40(1):107-113.
-doi:\href{https://doi.org/10.1016/j.vaccine.2021.11.040}{10.1016/j.vaccine.2021.11.040}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{33. }%
-\CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
-Zikmund-Fisher BJ. Vaccine hesitancy and rejection of a vaccine for the
-novel coronavirus in the united states. \emph{Frontiers in Immunology}.
-2021;12.
-doi:\href{https://doi.org/10.3389/fimmu.2021.558270}{10.3389/fimmu.2021.558270}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{34. }%
-\CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
-and factors related to {COVID}-19 vaccine hesitancy and unwillingness in
-canada: A systematic review and meta-analysis. \emph{Journal of Medical
-Virology}. 2022;95(1).
-doi:\href{https://doi.org/10.1002/jmv.28156}{10.1002/jmv.28156}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{35. }%
-\CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
-a sampled frequency table when the expected marginal totals are known.
-\emph{The Annals of Mathematical Statistics}. 1940;11(4):427-444.
-doi:\href{https://doi.org/10.1214/aoms/1177731829}{10.1214/aoms/1177731829}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{36. }%
-\CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
-2011.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{37. }%
-\CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
-{tidyverse}. \emph{Journal of Open Source Software}. 2019;4(43):1686.
-doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{38. }%
-\CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
-Markdown Publishing System}.; 2022.
-\url{https://CRAN.R-project.org/package=quarto}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{39. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
-2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{40. }%
-\CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
-compulsion: The case of {COVID}-19 vaccination in canada.
-\emph{Vaccine}. 2022;40(29):3923-3926.
-doi:\href{https://doi.org/10.1016/j.vaccine.2022.05.053}{10.1016/j.vaccine.2022.05.053}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
-of canada {COVID}-19 report: Enhancing {COVID}-19 vaccine acceptance in
-canada. Blais JM, ed. \emph{{FACETS}}. 2021;6:1184-1246.
-doi:\href{https://doi.org/10.1139/facets-2021-0037}{10.1139/facets-2021-0037}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{42. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=maps}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{43. }%
-\CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
-{COVID}-19 vaccine uptake in southeastern ontario, canada: Monitoring
-and addressing health inequities. \emph{Journal of Public Health
-Management and Practice}. 2022;28(6):615-623.
-doi:\href{https://doi.org/10.1097/phh.0000000000001565}{10.1097/phh.0000000000001565}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{44. }%
-\CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
-and other infection exposure according to race and ethnicity.
-\emph{American Journal of Industrial Medicine}. 2020;63(9):817-820.
-doi:\href{https://doi.org/10.1002/ajim.23145}{10.1002/ajim.23145}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{45. }%
-\CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
-review of {COVID}-19 and vulnerable workers: Intersecting occupational
-and public health issues. \emph{American Journal of Industrial
-Medicine}. 2021;64(7):551-566.
-doi:\href{https://doi.org/10.1002/ajim.23256}{10.1002/ajim.23256}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{46. }%
-\CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
-Strategy for Ontario {COVID}-19 Hotspots and Essential Workers}. Ontario
-{COVID}-19 Science Advisory Table; 2021.
-doi:\href{https://doi.org/10.47326/ocsat.2021.02.26.1.0}{10.47326/ocsat.2021.02.26.1.0}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{47. }%
-\CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
-Health;
-\url{https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf};
-2022.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{48. }%
-\CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
-Rosella LC. Ontario atlas of adult mortality, 1992-2015: Trends in
-{L}ocal {H}ealth {I}ntegration {N}etworks. Published online 2018.
-\url{https://tspace.library.utoronto.ca/handle/1807/82836}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{49. }%
-\CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
-and risk factors for {SARS}-{CoV}-2 infection in a south asian community
-in ontario: A cross-sectional analysis of a prospective cohort study.
-\emph{{CMAJ} Open}. 2022;10(3):E599-E609.
-doi:\href{https://doi.org/10.9778/cmajo.20220031}{10.9778/cmajo.20220031}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{50. }%
-\CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
-vaccination status by self-report versus source documentation: Impact on
-measuring {COVID}-19 vaccine effectiveness. \emph{Influenza and Other
-Respiratory Viruses}. 2022;16(6):1101-1111.
-doi:\href{https://doi.org/10.1111/irv.13023}{10.1111/irv.13023}}
-
-\end{CSLReferences}
-
-\end{document}
+This study explored differences in COVID-19 vaccination across the province of Ontario between late 2021 and early 2022 by taking into consideration socio-economic factors, such as income and race, their interactions, and the Health Regions within the province.
+Our results show that, during the period analyzed, significant differences in vaccination existed across different visible minority groups, income brackets, and Health Regions, showing  intra-provincial disparities in vaccine uptake.
+As the COVID-19 continues around the world, it important that future public policies take into consideration how to adequately reach individuals within minority groups that live across geographical areas where less probabilities of being vaccinated are likely.
+At the moment, this is an ongoing issue that needs to be addressed to ensure a more homogeneous outcome from the pandemic.

### 2023-04-16 manuscript_v3/appendix.tex
Updated months, fix descriptive table as September still appears

--- 2023-03-24 manuscript_v2/appendix.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, breakable, enhanced, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, breakable, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
 
@@
 race/ethnicity and the population from the Health Regions. Details of
-the implementation can be found in the R Script
-\texttt{raking\_logistic\_regression.R} in the \texttt{code} directory.
+the implementation can be found in the R Script \texttt{raking.R} in the
+\texttt{code} directory.
 
@@
 
-The tables below present summaries of information from the Fields survey
-(from the clean dataset analyzed in the main paper), showing responses
-by White/Caucasian individuals by Health Region and proportion of
-answers by LHIN in the West Health Region.
+Tables A-7 and A-8 present summaries of information from the Fields
+survey (from the clean dataset analyzed in the main paper), showing
+responses by White/Caucasian individuals by Health Region and proportion
+of answers by LHIN in the West Health Region.
 

### 2023-04-16 manuscript_v3/main.qmd
Updated months, fix descriptive table as September still appears

--- 2023-04-14 manuscript_v2/main.qmd
@@
 ---
-title: '**COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata**'
+title: '**Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata**'
 authors:
@@
       - ref: crm
+      - ref: espum
     email: ariel.mundo.ortiz@umontreal.ca
@@
     affiliations:
+      - ref: crm
       - ref: espum

### 2023-04-16 manuscript_v3/main.tex
Updated months, fix descriptive table as September still appears

--- 2023-04-14 manuscript_v2/main.tex
@@
 \hypersetup{
-  pdftitle={COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata},
-  pdfauthor={Ariel Mundo Ortiz1; Bouchra Nasri2,},
+  pdftitle={Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata},
+  pdfauthor={Ariel Mundo Ortiz1,2; Bouchra Nasri1,2,},
   colorlinks=true,
@@
 
-\title{\textbf{COVID-19 vaccination in Ontario: Exploring
-intra-provincial variations within Health Regions and socio-economic
-strata}}
-\author{Ariel Mundo Ortiz\textsuperscript{1} \and Bouchra
-Nasri\textsuperscript{2,*}}
+\title{\textbf{Differences in COVID-19 vaccination in the province of
+Ontario across Health Regions and socio-economic strata}}
+\author{Ariel Mundo Ortiz\textsuperscript{1,2} \and Bouchra
+Nasri\textsuperscript{1,2,*}}
 \date{}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, frame hidden, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{5}}, their
-implementation has been far from being equal to that of the vaccines of
-smallpox and polio, which were implemented on a global scale and that
-were indeed crucial to control these
+implementation has been far from being equal to that of the smallpox and
+polio vaccines, which were implemented on a global scale and that were
+crucial to control these
 diseases\textsuperscript{\protect\hyperlink{ref-kayser2021}{6}}. In
@@
 
-This problematic of COVID-19 vaccines rollout is a multifaceted issue
-resulting from, among other things, the development of new variants due
-to inadequate public health
+This problematic in the rollout of the COVID-19 vaccines is a
+multifaceted issue resulting from, among other things, the development
+of new variants due to inadequate public health
 measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
@@
 of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
-Because the adoption of the Health Regions occurred at a relatively
-recent time, there is an ongoing need to analyze the impact of this
-measure and identify the existence of intra-regional differences that
-might exist, and which could be specially important in the context of
-the COVID-19 pandemic.
+Because the adoption of the Health Regions is relatively recent there is
+an ongoing need to analyze the impact of this measure and identify the
+existence of intra-regional differences that might exist, and which
+could be specially important in the context of the COVID-19 pandemic.
 
@@
 vaccination status in Ontario using socio-economic (e.g., income,
-racial/ethnic identification), and spatial information at the level of
-the Health Regions to identify the existence of differences that might
-need to be addressed to ensure that the healthcare system of Ontario is
-more inclusive and that responds to the needs of its most vulnerable
-population.
+racial/ethnic identification), temporal (through a consecutive period of
+three months in late 2021), and spatial information (using the Health
+Regions of Ontario) to identify the existence of differences that might
+need to be addressed to ensure that the healthcare system of the
+province is more inclusive and that responds to the needs of its most
+vulnerable population.
 
@@
 information from participants (Table~\ref{tbl-covariates}), recorded
-their location (using the nearest municipality), and asked information
-on vaccination status by using the question ``Have you received the
-first dose of the COVID vaccine?'', with possible answers ``yes'' and
-``no''. The original dataset contained 39,029 entries (where each entry
-corresponded to a unique respondent).
+their location (using the nearest municipality), the date of access to
+the survey, and asked information on vaccination status by using the
+question ``Have you received the first dose of the COVID vaccine?'',
+with possible answers ``yes'' and ``no''. The original dataset contained
+39,029 entries (where each entry corresponded to a unique respondent).
 
 This dataset was cleaned to remove outliers that were identified during
-preliminary analyses, and processing the geographical information in the
-survey (city where the survey was responded) in order to match each city
-to its correspondent Health Region.
+preliminary analyses, and the geographical information in the survey
+(city where the survey was responded) was used to match each city to its
+correspondent Health Region.
 
@@
 Health Regions. The low representation of these Health Regions in the
-dataset is noticeable in Figure~\ref{fig-map},which shows that responses
-from these areas came from a relatively low number of cities when
-compared to the most populated Health Regions, such as the Toronto or
-Central Regions. We omitted the North East and North East Health Regions
-from further analyses due to the low number of entries. Therefore, the
-total number of unique entries used for analysis was 3,551 which
-included the East, Central, Toronto, and West Health Regions and that
-covered the period between September 30th,2021 and December 12th, 2021.
+dataset is noticeable in Figure~\ref{fig-map}, which shows that
+responses from these areas came from a relatively low number of cities
+when compared to the most populated Health Regions, such as the Toronto
+or Central Regions. We omitted the North East and North East Health
+Regions from further analyses due to the low number of entries.
+Therefore, the total number of unique entries used for analysis was
+3,549 which included the East, Central, Toronto, and West Health Regions
+covering the period between October 1st,2021 and December 12th, 2021.
 
@@
 vaccination depending on the socio-economic factors described in
-Table~\ref{tbl-covariates}, the Health Regions from Ontario indicated in
-Section~\ref{sec-data}, and the interactions between Race and Health
-Region, and Race and income, as previous studies have shown that
-socio-economic factors and their interactions are significant predictors
-of intent of vaccination and vaccination
+Table~\ref{tbl-covariates}, the month when the survey was answered, the
+Health Regions from Ontario indicated in Section~\ref{sec-data}, and the
+interactions between Race and Health Region, and Race and income, as
+previous studies have shown that socio-economic factors and their
+interactions are significant predictors of intent of vaccination and
+vaccination
 status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
@@
 The model was fitted first to the clean dataset to obtain uncorrected
-estimates. Additionally, because we identified differences between the
-proportions of all the socio-economic factors included in the analysis
+estimates. Because we identified differences between the proportions of
+all the socio-economic factors included in the analysis
 (Table~\ref{tbl-covariates}) and the Census data for Ontario, we used an
@@
 the Fields COVID-19 survey data for vaccination status and each of the
-covariates analyzed. The total number of entries analyzed was 3,551.
-Overall, 26.9\% of survey respondents (958) reported not having received
-the first dose of the vaccine, whereas 73.1\% (2,593) reported having
+covariates analyzed. The total number of entries analyzed was 3,549.
+Overall, 27\% (958) of survey respondents reported not having received
+the first dose of the vaccine, whereas 73\% (2,591) reported having
 received it. Within each socio-economic factor, respondents who reported
 living in a household with an income under CAD 25,000 represented 37\%
-of the total number of entries, those within the CAD 25,000-59,999
-income bracket represented 25\% of the total sample, and those with an
-income above CAD 60,000 represented 38 \% of the sample; across all
-income brackets, the percentage of individuals that reported having
-received a first dose of the vaccine was consistent, above 69\%.
+(1307) of the number of entries, those within the CAD 25,000-59,999
+income bracket represented 25\% (889) of the total sample, and those
+with an income above CAD 60,000 represented 38 \% (1353) of the sample.
+Between the three income brackets, the lower vaccination rate
+corresponded to the under CAD 25,000 group, with 69\% .
 
 Within the age groups of survey respondents, the age group between 16-34
-years had the highest representation in the survey responses (1,521,
+years had the highest representation in the survey responses (1,520 or
 42.8\% of all responses). Within this age bracket, 73\% of respondents
 indicated having received the vaccine, whereas the lowest vaccination
-rate was in the bracket of those 55 years of age and above, with a total
-of 72\%. The Health Region with highest representation in the survey was
-Toronto, accounting for 1,324 entries (37.2\%), with a vaccination rate
-of 72\%. Regarding race/ethnicity, individuals that identified as
-White/Caucasian represented 1313 (37\%) of all entries and had the
+rate was in the bracket of those 55 years of age and above, with 72\%.
+The Health Region with highest representation in the survey was Toronto,
+accounting for 1,323 entries (37.3\% of the total), with a vaccination
+rate of 72\%. Regarding race/ethnicity, individuals that identified as
+White/Caucasian represented 37\% of all entries (1,312) and had the
 highest vaccination uptake with 82\% of them indicating to have received
@@
 identified as Indigenous.
+
+\begin{verbatim}
+Warning: Outer names are only allowed for unnamed scalar atomic inputs
+\end{verbatim}
 
@@
 \hspace{1em}West & 228 (27\%) & 611 (73\%)\\
+\textbf{Month} &  & \\
+\hspace{1em}October & 469 (27\%) & 1,263 (73\%)\\
+\hspace{1em}November & 376 (28\%) & 980 (72\%)\\
+\hspace{1em}December & 113 (25\%) & 348 (75\%)\\
+\hspace{1em}September & 0 (NA\%) & 0 (NA\%)\\
 \textbf{Race} &  & \\
@@
 models (for the uncorrected and corrected data) on vaccination status
-using socio-economic factors (age group, income, race), geographical
-areas (Health Regions) and the interactions between income and race and
-Health Region and race. There were no statistically significant
-differences in vaccination rates within the age groups from the survey,
-but significant odds ratios were estimated for other covariates. Within
-household income brackets, individuals with an income under CAD 25,000
-or between CAD 25,000-59,999 had significantly lower odds of vaccination
-than those with an income above CAD 60,000 (ORs=0.37 and 0.59,
-\emph{p}=0.011 and \textless0.001, respectively). Within Race/Ethnicity,
-individuals who identified as Arab/Middle Eastern, Black, or Latin
-American, had significantly lower odds of vaccination than those in the
-White/Caucasian group (ORs=0.31, 0.32, 0.28, and
-\emph{p}=0.004,\textless0.001 and 0.004, respectively); additionally,
-those individuals in the Other Race/Ethnicity group (a group that
-included Southeast Asian, Filipino, West Asian, and Minorities Not
-Identified Elsewhere) had even lower odds of vaccination than the other
-minority groups (OR=0.22, \emph{p}\textless0.001). Regarding Health
-Regions, individuals that reported living in the West Health Region
-(which comprises the regions of Waterloo and Niagara, the counties of
-Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
-Brant, and Chatham-Kent) had significantly higher odds of vaccination
-than those in the Health Region of Toronto (OR=1.55, \emph{p}=0.029).
+using socio-economic factors (age group, income, race), months covered
+by the survey (October, November, and December), geographical areas
+(Health Regions) and the interactions between income and race and Health
+Region and race. The corrected results show that there were no
+statistically significant differences in vaccination rate between the
+age groups or across the months covered by the survey as in both cases
+the odd ratios were similar. However, between the different household
+income brackets, individuals with an income under CAD 25,000 or between
+CAD 25,000-59,999 had significantly lower odds of vaccination than those
+with an income above CAD 60,000 (ORs=0.37 and 0.59,
+\emph{p}\textless0.001 and \emph{p}=0.010, respectively). Within
+Race/Ethnicity, individuals who identified as Arab/Middle Eastern,
+Black, or Latin American, had significantly lower odds of vaccination
+than those in the White/Caucasian group (ORs=0.32, 0.32, 0.27, and
+\emph{p}=0.003, \emph{p}\textless0.001 and \emph{p}=0.004,
+respectively); additionally, those individuals that reported to belong
+to the ``Other'' Race/Ethnicity group (which included the Southeast
+Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere
+groups according to the Census) had even lower odds of vaccination than
+the other minority groups (OR=0.22, \emph{p}\textless0.001). Regarding
+Health Regions, individuals that reported living in the West Health
+Region (which comprises the regions of Waterloo and Niagara, the
+counties of Wellington, Essex, and Lambton, and the cities of Hamilton,
+Haldimand, Brant, and Chatham-Kent) had significantly higher odds of
+vaccination than those in the Health Region of Toronto (OR=1.54,
+\emph{p}=0.031).
 
@@
 individuals with a household income below CAD 25,000 who identified as
-Arab/Middle Eastern (OR=3.05, \emph{p}=0.013), Black (OR=3.19,
-\emph{p}=0.004), Latin American (OR=2.80, \emph{p}=0.041), or that
-belonged to other minority groups (OR=4.59, \emph{p}\textless0.001).
+Arab/Middle Eastern (OR=3.08, \emph{p}=0.013), Black (OR=3.15,
+\emph{p}=0.004), Latin American (OR=2.81, \emph{p}=0.041), or that
+belonged to other minority groups (OR=4.63, \emph{p}\textless0.001).
 Within the CAD 25,000-59,999 income bracket, individuals who identified
 as belonging to other racial minority groups had significantly higher
-odds of vaccination (OR=6.93, \emph{p}\textless0.001).
+odds of vaccination (OR=6.96, \emph{p}\textless0.001).
 
@@
 \hspace{1em}October & — & — &  & — & — & \\
+\hspace{1em}November & 0.95 & 0.81, 1.12 & 0.6 & 0.92 & 0.71, 1.18 & 0.5\\
 \hspace{1em}December & 1.16 & 0.91, 1.49 & 0.2 & 1.16 & 0.78, 1.71 & 0.5\\
-\hspace{1em}November & 0.95 & 0.81, 1.12 & 0.6 & 0.92 & 0.71, 1.18 & 0.5\\
 \textbf{Income} &  &  &  &  &  & \\
@@
 There existence of healthcare disparities within Ontario is a topic of
-particular interest in due to the recent change in the healthcare system
-of this region, which eliminated the LHIN model and adopted larger
-Health Regions system in late 2019 in an attempt to address the
-disparities of the former
-approach\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
+particular interest as it motivated the recent change in the healthcare
+system of this region, which a system based on Health Regions in late
+2019 in an attempt to address the disparities from the LHIN
+model\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
 In this context, analyzing COVID-19 vaccination estimates within the
@@
 years of age and older) reported by Public Health Ontario at the start
-of the period covered by the data (between September 30th, 2021 and
+of the period covered by the data (between October 1st, 2021 and
 December 12th,

### 2023-04-16 code/data_preparation.R

@@ -77,6 +77,7 @@ clean_data<-clean_data %>%
 
 
 clean_data$Month <- as.factor(clean_data$Month)
+clean_data$Month <- fct_relevel(clean_data$Month,levels=c("October","November","December"))
 
 ## relevel for reference groups in the model
 
@@ -84,9 +85,9 @@ clean_data<- within(clean_data, Age_group_ord <- relevel(Age_group_ord, ref = "1
 clean_data<- within(clean_data, Race <- relevel(Race, ref = "white_caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
 clean_data<- within(clean_data, income_ord <- relevel(income_ord, ref = "60000_and_above"))
-clean_data<- within(clean_data, Month <- relevel(Month, ref = "October"))
+#clean_data<- within(clean_data, Month <- relevel(Month, ref = "October"))
 
-# remove september, only 2 observations
+# remove september, as it only has 2 observations
 
 clean_data <-clean_data %>%
   subset(Month!="September")

### 2023-04-17 manuscript_v3/04-discussion.qmd
Fixed model, discussion, appendix

@@
 
-There existence of healthcare disparities within Ontario is a topic of particular interest as it motivated the recent change in the healthcare system of this region, which a system based on Health Regions in late 2019 in an attempt to address the disparities from the LHIN model  [@tsasis2012;@dong2022].
-In this context, analyzing COVID-19 vaccination estimates within the province is important as they can provide an initial assessment of variations that might need to be addressed by decision-makers.
+The existence of healthcare disparities in Ontario motivated the recent change in the healthcare system of the province, which changed the LHIN model for a Health Region model in late 2019  [@tsasis2012;@dong2022].
+In this context, analyzing COVID-19 vaccination estimates between the Health Regions can serve as an indicator of ongoing intra-provincial disparities that may need to be addressed to ensure that the Health Region model is able to improve health access for the inhabitants of Ontario, which faces unique challenges due to its condition as the most populated and the most ethnically diverse province of Canada.
 
-Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats).
+Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of surveyed individuals reported to have received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats).
 It is worth mentioning that province-wide vaccination rates for the period of interest are somewhat different from those of the survey, particularly in the case of those 55 years of age and older, which in the survey had a vaccination rate of 72\%, against a rate of 88.4\% reported for the closest age bracket (50 years of age and older) reported by Public Health Ontario at the start of the period covered by the data (between October 1st, 2021 and December 12th, 2021) [@ontario-covid].
-In this regard, differences are to be expected because the survey represents a random sample from the population, and therefore, the sampling process is likely to cause variations between the values from the survey and province-wide estimates.
+In this case, differences between the survey and province-wide estimates are to be expected as the data from survey represents a random sample from the overall population.
 
-We believe that these differences did not have a significant impact in our analyses, because our results indicate that there were no significant differences in vaccination odds among the age groups considered in the survey, in agreement with the overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], and with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022].
-Moreover, vaccination rates within each age in the dataset are in good agreement with province-wide estimates (e.g., a rate of 95\% for those with 61 years of age, Supplementary Table A-6).
-Additionally, when overall vaccination rates for the province are disaggregated, it can be seen that regional differences were present during the period analyzed, which can be masked by province-wide estimates.
-For example, the Public Health Unit of Lambton (a region within the West Health Region) and the Public Health Unit of Haliburton, Kawartha, and the Pine Ridge District (an area covered by the East Health Region) reported lower vaccination rates (78\%) for those 50 years of age and older at the beginning of the period of interest, whereas other regions had higher vaccination rates [@ontario-covid].
-To this day, differences in vaccination rates within the province are present, because according to Public Health Ontario, as of March of 2023 some regions have less than 75\% vaccination rate [@ontario-covid-map].
-Overall, these reasons indicate that the estimates obtained from the data are in good agreement with the trends from the population.
+However, this variation did not lead to discrepant results, as the estimates from the model indicate that no significant differences in vaccination odds were identified among the age groups analyzed, in agreement with both overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], and with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022].
+In other words, although vaccination rates obtained from the survey were slightly lower than the provincial estimates, these values still represented a valid approximation; this notion is reinforced by the consistency in the proportion of vaccination rates (@tbl-descriptive-stats) and vaccination odds (@tbl-model) across the period covered by the survey, which follow the vaccination rates from Public Health Ontario and which indicate that due to the relatively high coverage achieved in the population at that point, there were no abrupt shift in the trends, which increased by around 3% across all age groups during the three months (October, November, and December of 2021 [@ontario-covid]).
+Moreover, there was good agreement between vaccination rates within each age in the raw dataset and province-wide estimates (e.g., a rate of 95\% for those with 61 years of age, Supplementary Table A-6, which is similar to the value reported by Public Health Ontario).
+It is also important to mention that regional differences can be masked by overall estimates, as when overall vaccination rates for the province (from Public Health Ontario) are disaggregated, it can be seen that regional differences during the period analyzed existed.
+For example, the Public Health Unit of Lambton (a region within the West Health Region) and the Public Health Unit of Haliburton, Kawartha, and the Pine Ridge District (an area covered by the East Health Region) reported lower vaccination rates (78\%) for those 50 years of age and older at the beginning of the period of interest, in contrast with other regions that had vaccination rates above 80\% [@ontario-covid].
+To this day, differences in vaccination rates within the province continue, as according to Public Health Ontario, as of March of 2023 some regions still have less than 75\% vaccination rate [@ontario-covid-map].
 
@@
 Interestingly, a disproportionate number of COVID-19 cases and low vaccination rate (under 50%) have been previously reported in the South Asian community of Ontario [@anand2022]; in this regard, our result provides additional context  by showing that within the South Asian community, there were differences in vaccination uptake across Ontario.
-Moreover, because significant lower odds of vaccination were also identified  other minority groups, this provides a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.
+Moreover, because significant lower odds of vaccination were also identified in other minority groups, this provides a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.
 

### 2023-04-17 manuscript_v3/appendix.qmd
Fixed model, discussion, appendix

--- 2023-03-26 manuscript_v2/appendix.qmd
@@
 ---
-title: '**COVID-19 vaccination in Ontario: Exploring intra-provincial variations within Health Regions and socio-economic strata**'
+title: '**Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata**'
 subtitle: Appendix
@@
 library(gt)
+library(lubridate)
 library(gtsummary)
@@
   geom_bar(stat="identity") +
+    xlab("Age group")+
   geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"), y=ypos))+
@@
 #| fig-width: 6
-#| fig-cap: Estimated marginal means for the combination between Race and Income in the regression model.
+#| fig-cap: Estimated marginal means for the interaction between Race and Income in the regression model.
 
-race_income<-emmip(m1, Race ~ income_ord, CIs=TRUE,plotit = FALSE)
+race_income<-emmip(m2, Race ~ income_ord, CIs=TRUE,plotit = FALSE)
 
@@
 #| fig-width: 6
-#| fig-cap: Estimated marginal means for the combination between Race and Health Region in the regression model.
+#| fig-cap: Estimated marginal means for the interaction between Race and Health Region in the regression model.
 
-race_hr<-emmip(m1, Race ~ Health_Region, CIs=TRUE,plotit = FALSE)
+race_hr<-emmip(m2, Race ~ Health_Region, CIs=TRUE,plotit = FALSE)
 
@@
 
-clean_data %>% subset (Race=="white_caucasian")%>%
+clean_data %>% subset (Race=="White/Caucasian")%>%
   tbl_summary(
@@
 
-# clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
-# tbl_summary(include=c(Race),
-#             sort = all_categorical() ~ "frequency") %>% 
-#   bold_labels()
-# 
-# 
-# clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
-# tbl_summary(by=first_dose,
-#             percent="row",
-#   include=c(Race),
-#             sort = all_categorical() ~ "frequency") %>% 
-#   bold_labels()
-# 
-# clean_data %>% subset (LHIN=="Erie St. Clair")%>%
-# tbl_summary(by=first_dose,
-#             percent="row",
-#   include=c(Race),
-#             sort = all_categorical() ~ "frequency") %>% 
-#   bold_labels()
 ```

### 2023-04-17 manuscript_v3/appendix.tex
Fixed model, discussion, appendix

@@
 
-\title{\textbf{COVID-19 vaccination in Ontario: Exploring
-intra-provincial variations within Health Regions and socio-economic
-strata}}
+\title{\textbf{Differences in COVID-19 vaccination in the province of
+Ontario across Health Regions and socio-economic strata}}
 \usepackage{etoolbox}
@@
 \subtitle{Appendix}
-\author{Ariel Mundo Ortiz\textsuperscript{1} \and Bouchra
-Nasri\textsuperscript{2,*}}
+\author{Ariel Mundo Ortiz\textsuperscript{1,2} \and Bouchra
+Nasri\textsuperscript{1,2,*}}
 \date{}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, breakable, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 \midrule
-white\_caucasian & 1313 & $37.0\%$ \\ 
-arab\_middle\_eastern & 214 & $6.0\%$ \\ 
-black & 298 & $8.4\%$ \\ 
-east\_asian\_pacific\_islander & 303 & $8.5\%$ \\ 
-indigenous & 191 & $5.4\%$ \\ 
-latin\_american & 180 & $5.1\%$ \\ 
-mixed & 310 & $8.7\%$ \\ 
-other & 367 & $10.3\%$ \\ 
-south\_asian & 375 & $10.6\%$ \\ 
+White/Caucasian & 1312 & $37.0\%$ \\ 
+Arab/Middle Eastern & 214 & $6.0\%$ \\ 
+Black & 298 & $8.4\%$ \\ 
+East Asian/Pacific Islander & 303 & $8.5\%$ \\ 
+Indigenous & 191 & $5.4\%$ \\ 
+Latin American & 180 & $5.1\%$ \\ 
+Mixed & 310 & $8.7\%$ \\ 
+Other & 367 & $10.3\%$ \\ 
+South Asian & 374 & $10.5\%$ \\ 
 \bottomrule
@@
 \toprule
-\textbf{Characteristic} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,593\\
+\textbf{Characteristic} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,591\\
 \midrule
@@
 \hspace{1em}25 & 25 (24\%) & 80 (76\%)\\
-\hspace{1em}26 & 23 (30\%) & 53 (70\%)\\
+\hspace{1em}26 & 23 (31\%) & 52 (69\%)\\
 \hspace{1em}27 & 16 (25\%) & 49 (75\%)\\
@@
 \hspace{1em}57 & 12 (22\%) & 43 (78\%)\\
-\hspace{1em}58 & 6 (14\%) & 38 (86\%)\\
+\hspace{1em}58 & 6 (14\%) & 37 (86\%)\\
 \hspace{1em}59 & 4 (12\%) & 28 (88\%)\\
@@
 
-\caption{Estimated marginal means for the combination between Race and
+\caption{Estimated marginal means for the interaction between Race and
 Income in the regression model.}
@@
 
-\caption{Estimated marginal means for the combination between Race and
+\caption{Estimated marginal means for the interaction between Race and
 Health Region in the regression model.}
@@
 \toprule
-\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N = 1,080\\
+\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N = 1,079\\
 \midrule
@@
 \hspace{1em}East & 41 (15\%) & 236 (85\%)\\
-\hspace{1em}Central & 38 (16\%) & 194 (84\%)\\
+\hspace{1em}Central & 38 (16\%) & 193 (84\%)\\
 \bottomrule

### 2023-04-17 manuscript_v3/main.qmd
Fixed model, discussion, appendix

@@
 Our results show differences in vaccination due to race/ethnicity, income, Health Regions (geographical areas used for health service access in Ontario), and their interactions.
-In particular, we show
-that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32, 0.28, and _p_=0.004, _p_<0.001 and _p_=0.004, respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets.
-Finally, we also identified lower odds of vaccination within certain minority groups in the West Health Region, which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+In particular, we show that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs and 95% CIs=0.31[0.14,0.68], 0.32[0.17,0.60], 0.27[0.11,0.66], respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.08, 95% CI=[1.27,7.47]), Black (OR=3.15,95% CI=[1.43,6.92]), Latin American (OR=2.81, 95% CI=[1.
+04,7.59]), or that belonged to other minority groups  (OR=4.63,95% CI=[2.34,9.13]) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets.
+Finally, we also identified lower odds of vaccination within certain minority groups in the West Health Region.
 This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population that have been largely impacted by the pandemic.

### 2023-04-17 manuscript_v3/main.tex
Fixed model, discussion, appendix

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
 
@@
 Arab/Middle Eastern, Black, or Latin American, had significantly lower
-odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32,
-0.28, and \emph{p}=0.004, \emph{p}\textless0.001 and \emph{p}=0.004,
+odds of vaccination than White/Caucasian individuals (ORs and 95\%
+CIs=0.31{[}0.14,0.68{]}, 0.32{[}0.17,0.60{]}, 0.27{[}0.11,0.66{]},
 respectively), and that individuals with a household income below CAD
-25,000 who identified as Arab/Middle Eastern (OR=3.05, \emph{p}=0.013),
-Black (OR=3.19, \emph{p}=0.004), Latin American (OR=2.80,
-\emph{p}=0.041), or that belonged to other minority groups (OR=4.59,
-\emph{p}\textless0.001) had higher odds of vaccination than individuals
-from the same racial/ethnic group in higher income brackets. Finally, we
-also identified lower odds of vaccination within certain minority groups
-in the West Health Region, which comprises the regions of Waterloo and
-Niagara, the counties of Wellington, Essex and Lambton, and the cities
-of Hamilton, Haldimand, Brant, and Chatham-Kent. This study shows that
-there is an ongoing need to better understand and address differences in
-vaccination uptake across diverse segments of the population that have
-been largely impacted by the pandemic.
+25,000 who identified as Arab/Middle Eastern (OR=3.08, 95\%
+CI={[}1.27,7.47{]}), Black (OR=3.15,95\% CI={[}1.43,6.92{]}), Latin
+American (OR=2.81, 95\% CI={[}1. 04,7.59{]}), or that belonged to other
+minority groups (OR=4.63,95\% CI={[}2.34,9.13{]}) had higher odds of
+vaccination than individuals from the same racial/ethnic group in higher
+income brackets. Finally, we also identified lower odds of vaccination
+within certain minority groups in the West Health Region. This study
+shows that there is an ongoing need to better understand and address
+differences in vaccination uptake across diverse segments of the
+population that have been largely impacted by the pandemic.
 
@@
 accounting for 1,323 entries (37.3\% of the total), with a vaccination
-rate of 72\%. Regarding race/ethnicity, individuals that identified as
-White/Caucasian represented 37\% of all entries (1,312) and had the
-highest vaccination uptake with 82\% of them indicating to have received
-the COVID-19 vaccine. On the other hand, the ethnic group with the
-lowest number of entries in the survey was Latin American, with a total
-of 180, or 5\% of all entries. Vaccination rates across all minority
-groups were below the value reported by White/Caucasians, with the
-lowest vaccination rate (60\%) being reported by individuals that
-identified as Indigenous.
-
-\begin{verbatim}
-Warning: Outer names are only allowed for unnamed scalar atomic inputs
-\end{verbatim}
+rate of 72\%. Across the three months covered by the survey, the highest
+number of entries was in October of 2021, with a total of 1,732
+respondents, which corresponded to 49\% of all the survey entries.
+Additionally, the proportion of respondents who reported being
+vaccinated was similar in all months, ranging between 72\% and 75\%,
+with the highest value corresponding to December of 2021.
+
+Regarding race/ethnicity, individuals that identified as White/Caucasian
+represented 37\% of all entries (1,312) and had the highest vaccination
+uptake with 82\% of them indicating to have received the COVID-19
+vaccine. On the other hand, the ethnic group with the lowest number of
+entries in the survey was Latin American, with a total of 180, or 5\% of
+all entries. Vaccination rates across all minority groups were below the
+value reported by White/Caucasians, with the lowest vaccination rate
+(60\%) being reported by individuals that identified as Indigenous.
 
@@
 \midrule
-\textbf{Income} &  & \\
-\hspace{1em}60000\_and\_above & 305 (23\%) & 1,048 (77\%)\\
-\hspace{1em}25000\_59999 & 253 (28\%) & 636 (72\%)\\
-\hspace{1em}under\_25000 & 400 (31\%) & 907 (69\%)\\
+\textbf{Income (CAD)} &  & \\
+\hspace{1em}60000 and above & 305 (23\%) & 1,048 (77\%)\\
+\hspace{1em}25000-59999 & 253 (28\%) & 636 (72\%)\\
+\hspace{1em}under 25000 & 400 (31\%) & 907 (69\%)\\
 \textbf{Age Group} &  & \\
-\hspace{1em}16\_34 & 409 (27\%) & 1,111 (73\%)\\
-\hspace{1em}35\_54 & 252 (26\%) & 712 (74\%)\\
-\hspace{1em}55\_and\_over & 297 (28\%) & 768 (72\%)\\
+\hspace{1em}16-34 & 409 (27\%) & 1,111 (73\%)\\
+\hspace{1em}35-54 & 252 (26\%) & 712 (74\%)\\
+\hspace{1em}55 and over & 297 (28\%) & 768 (72\%)\\
 \textbf{Health Region} &  & \\
@@
 \hspace{1em}December & 113 (25\%) & 348 (75\%)\\
-\hspace{1em}September & 0 (NA\%) & 0 (NA\%)\\
 \textbf{Race} &  & \\
-\hspace{1em}white\_caucasian & 233 (18\%) & 1,079 (82\%)\\
-\hspace{1em}arab\_middle\_eastern & 76 (36\%) & 138 (64\%)\\
-\hspace{1em}black & 114 (38\%) & 184 (62\%)\\
-\hspace{1em}east\_asian\_pacific\_islander & 69 (23\%) & 234 (77\%)\\
-\hspace{1em}indigenous & 76 (40\%) & 115 (60\%)\\
-\hspace{1em}latin\_american & 69 (38\%) & 111 (62\%)\\
-\hspace{1em}mixed & 105 (34\%) & 205 (66\%)\\
-\hspace{1em}other & 128 (35\%) & 239 (65\%)\\
-\hspace{1em}south\_asian & 88 (24\%) & 286 (76\%)\\
+\hspace{1em}White/Caucasian & 233 (18\%) & 1,079 (82\%)\\
+\hspace{1em}Arab/Middle Eastern & 76 (36\%) & 138 (64\%)\\
+\hspace{1em}Black & 114 (38\%) & 184 (62\%)\\
+\hspace{1em}East Asian/Pacific Islander & 69 (23\%) & 234 (77\%)\\
+\hspace{1em}Indigenous & 76 (40\%) & 115 (60\%)\\
+\hspace{1em}Latin American & 69 (38\%) & 111 (62\%)\\
+\hspace{1em}Mixed & 105 (34\%) & 205 (66\%)\\
+\hspace{1em}Other & 128 (35\%) & 239 (65\%)\\
+\hspace{1em}South Asian & 88 (24\%) & 286 (76\%)\\
 \bottomrule
@@
 \textbf{Age Group} &  &  &  &  &  & \\
-\hspace{1em}16\_34 & — & — &  & — & — & \\
-\hspace{1em}35\_54 & 0.93 & 0.77, 1.14 & 0.5 & 0.90 & 0.67, 1.21 & 0.5\\
-\hspace{1em}55\_and\_over & 0.74 & 0.61, 0.89 & 0.002 & 0.99 & 0.74, 1.32 & >0.9\\
+\hspace{1em}16-34 & — & — &  & — & — & \\
+\hspace{1em}35-54 & 0.93 & 0.77, 1.14 & 0.5 & 0.90 & 0.67, 1.21 & 0.5\\
+\hspace{1em}55 and over & 0.74 & 0.61, 0.89 & 0.002 & 0.99 & 0.74, 1.32 & >0.9\\
 \textbf{Month} &  &  &  &  &  & \\
@@
 \hspace{1em}December & 1.16 & 0.91, 1.49 & 0.2 & 1.16 & 0.78, 1.71 & 0.5\\
-\textbf{Income} &  &  &  &  &  & \\
-\hspace{1em}60000\_and\_above & — & — &  & — & — & \\
-\hspace{1em}25000\_59999 & 0.59 & 0.41, 0.84 & 0.003 & 0.59 & 0.39, 0.88 & 0.010\\
-\hspace{1em}under\_25000 & 0.41 & 0.29, 0.58 & <0.001 & 0.37 & 0.25, 0.56 & <0.001\\
+\textbf{Income (CAD)} &  &  &  &  &  & \\
+\hspace{1em}60000 and above & — & — &  & — & — & \\
+\hspace{1em}25000-59999 & 0.59 & 0.41, 0.84 & 0.003 & 0.59 & 0.39, 0.88 & 0.010\\
+\hspace{1em}under 25000 & 0.41 & 0.29, 0.58 & <0.001 & 0.37 & 0.25, 0.56 & <0.001\\
 \textbf{Race} &  &  &  &  &  & \\
-\hspace{1em}white\_caucasian & — & — &  & — & — & \\
-\hspace{1em}arab\_middle\_eastern & 0.24 & 0.12, 0.49 & <0.001 & 0.31 & 0.14, 0.68 & 0.003\\
-\hspace{1em}black & 0.30 & 0.17, 0.54 & <0.001 & 0.32 & 0.17, 0.60 & <0.001\\
-\hspace{1em}east\_asian\_pacific\_islander & 0.73 & 0.36, 1.51 & 0.4 & 1.15 & 0.50, 2.66 & 0.7\\
-\hspace{1em}indigenous & 0.38 & 0.17, 0.82 & 0.014 & 0.45 & 0.20, 1.04 & 0.061\\
-\hspace{1em}latin\_american & 0.28 & 0.13, 0.59 & <0.001 & 0.27 & 0.11, 0.66 & 0.004\\
-\hspace{1em}mixed & 0.58 & 0.30, 1.11 & 0.10 & 0.63 & 0.24, 1.62 & 0.3\\
-\hspace{1em}other & 0.20 & 0.11, 0.35 & <0.001 & 0.22 & 0.12, 0.41 & <0.001\\
-\hspace{1em}south\_asian & 0.79 & 0.44, 1.43 & 0.4 & 0.90 & 0.48, 1.67 & 0.7\\
+\hspace{1em}White/Caucasian & — & — &  & — & — & \\
+\hspace{1em}Arab/Middle Eastern & 0.24 & 0.12, 0.49 & <0.001 & 0.31 & 0.14, 0.68 & 0.003\\
+\hspace{1em}Black & 0.30 & 0.17, 0.54 & <0.001 & 0.32 & 0.17, 0.60 & <0.001\\
+\hspace{1em}East Asian/Pacific Islander & 0.73 & 0.36, 1.51 & 0.4 & 1.15 & 0.50, 2.66 & 0.7\\
+\hspace{1em}Indigenous & 0.38 & 0.17, 0.82 & 0.014 & 0.45 & 0.20, 1.04 & 0.061\\
+\hspace{1em}Latin American & 0.28 & 0.13, 0.59 & <0.001 & 0.27 & 0.11, 0.66 & 0.004\\
+\hspace{1em}Mixed & 0.58 & 0.30, 1.11 & 0.10 & 0.63 & 0.24, 1.62 & 0.3\\
+\hspace{1em}Other & 0.20 & 0.11, 0.35 & <0.001 & 0.22 & 0.12, 0.41 & <0.001\\
+\hspace{1em}South Asian & 0.79 & 0.44, 1.43 & 0.4 & 0.90 & 0.48, 1.67 & 0.7\\
 \textbf{Health Region} &  &  &  &  &  & \\
@@
 \hspace{1em}West & 1.35 & 0.94, 1.93 & 0.10 & 1.54 & 1.04, 2.29 & 0.031\\
-\textbf{Income * Race} &  &  &  &  &  & \\
-\hspace{1em}25000\_59999 * arab\_middle\_eastern & 2.19 & 0.95, 5.05 & 0.067 & 1.85 & 0.68, 4.99 & 0.2\\
-\hspace{1em}under\_25000 * arab\_middle\_eastern & 2.95 & 1.39, 6.24 & 0.005 & 3.08 & 1.27, 7.47 & 0.013\\
-\hspace{1em}25000\_59999 * black & 1.18 & 0.59, 2.36 & 0.6 & 1.31 & 0.58, 2.98 & 0.5\\
-\hspace{1em}under\_25000 * black & 2.84 & 1.46, 5.53 & 0.002 & 3.15 & 1.43, 6.92 & 0.004\\
-\hspace{1em}25000\_59999 * east\_asian\_pacific\_islander & 0.97 & 0.44, 2.10 & >0.9 & 0.42 & 0.17, 1.06 & 0.066\\
-\hspace{1em}under\_25000 * east\_asian\_pacific\_islander & 1.90 & 0.90, 4.01 & 0.094 & 1.15 & 0.47, 2.84 & 0.8\\
-\hspace{1em}25000\_59999 * indigenous & 1.77 & 0.72, 4.32 & 0.2 & 1.33 & 0.46, 3.80 & 0.6\\
-\hspace{1em}under\_25000 * indigenous & 1.60 & 0.71, 3.60 & 0.3 & 1.41 & 0.54, 3.72 & 0.5\\
-\hspace{1em}25000\_59999 * latin\_american & 0.89 & 0.38, 2.10 & 0.8 & 1.25 & 0.45, 3.45 & 0.7\\
-\hspace{1em}under\_25000 * latin\_american & 3.10 & 1.34, 7.21 & 0.009 & 2.81 & 1.04, 7.59 & 0.041\\
-\hspace{1em}25000\_59999 * mixed & 0.87 & 0.39, 1.93 & 0.7 & 0.87 & 0.33, 2.33 & 0.8\\
-\hspace{1em}under\_25000 * mixed & 1.27 & 0.65, 2.49 & 0.5 & 1.12 & 0.38, 3.33 & 0.8\\
-\hspace{1em}25000\_59999 * other & 5.50 & 2.43, 12.5 & <0.001 & 6.96 & 2.67, 18.2 & <0.001\\
-\hspace{1em}under\_25000 * other & 4.09 & 2.27, 7.37 & <0.001 & 4.63 & 2.34, 9.13 & <0.001\\
-\hspace{1em}25000\_59999 * south\_asian & 1.14 & 0.55, 2.38 & 0.7 & 1.22 & 0.52, 2.87 & 0.7\\
-\hspace{1em}under\_25000 * south\_asian & 1.58 & 0.82, 3.05 & 0.2 & 2.01 & 0.93, 4.34 & 0.075\\
+\textbf{Income (CAD) * Race} &  &  &  &  &  & \\
+\hspace{1em}25000-59999 * Arab/Middle Eastern & 2.19 & 0.95, 5.05 & 0.067 & 1.85 & 0.68, 4.99 & 0.2\\
+\hspace{1em}under 25000 * Arab/Middle Eastern & 2.95 & 1.39, 6.24 & 0.005 & 3.08 & 1.27, 7.47 & 0.013\\
+\hspace{1em}25000-59999 * Black & 1.18 & 0.59, 2.36 & 0.6 & 1.31 & 0.58, 2.98 & 0.5\\
+\hspace{1em}under 25000 * Black & 2.84 & 1.46, 5.53 & 0.002 & 3.15 & 1.43, 6.92 & 0.004\\
+\hspace{1em}25000-59999 * East Asian/Pacific Islander & 0.97 & 0.44, 2.10 & >0.9 & 0.42 & 0.17, 1.06 & 0.066\\
+\hspace{1em}under 25000 * East Asian/Pacific Islander & 1.90 & 0.90, 4.01 & 0.094 & 1.15 & 0.47, 2.84 & 0.8\\
+\hspace{1em}25000-59999 * Indigenous & 1.77 & 0.72, 4.32 & 0.2 & 1.33 & 0.46, 3.80 & 0.6\\
+\hspace{1em}under 25000 * Indigenous & 1.60 & 0.71, 3.60 & 0.3 & 1.41 & 0.54, 3.72 & 0.5\\
+\hspace{1em}25000-59999 * Latin American & 0.89 & 0.38, 2.10 & 0.8 & 1.25 & 0.45, 3.45 & 0.7\\
+\hspace{1em}under 25000 * Latin American & 3.10 & 1.34, 7.21 & 0.009 & 2.81 & 1.04, 7.59 & 0.041\\
+\hspace{1em}25000-59999 * Mixed & 0.87 & 0.39, 1.93 & 0.7 & 0.87 & 0.33, 2.33 & 0.8\\
+\hspace{1em}under 25000 * Mixed & 1.27 & 0.65, 2.49 & 0.5 & 1.12 & 0.38, 3.33 & 0.8\\
+\hspace{1em}25000-59999 * Other & 5.50 & 2.43, 12.5 & <0.001 & 6.96 & 2.67, 18.2 & <0.001\\
+\hspace{1em}under 25000 * Other & 4.09 & 2.27, 7.37 & <0.001 & 4.63 & 2.34, 9.13 & <0.001\\
+\hspace{1em}25000-59999 * South Asian & 1.14 & 0.55, 2.38 & 0.7 & 1.22 & 0.52, 2.87 & 0.7\\
+\hspace{1em}under 25000 * South Asian & 1.58 & 0.82, 3.05 & 0.2 & 2.01 & 0.93, 4.34 & 0.075\\
 \textbf{Race * Health Region} &  &  &  &  &  & \\
-\hspace{1em}arab\_middle\_eastern * Central & 0.76 & 0.33, 1.71 & 0.5 & 0.66 & 0.25, 1.69 & 0.4\\
-\hspace{1em}black * Central & 0.48 & 0.23, 1.02 & 0.055 & 0.44 & 0.19, 0.99 & 0.046\\
-\hspace{1em}east\_asian\_pacific\_islander * Central & 0.84 & 0.38, 1.89 & 0.7 & 0.98 & 0.38, 2.53 & >0.9\\
-\hspace{1em}indigenous * Central & 0.60 & 0.24, 1.50 & 0.3 & 0.63 & 0.22, 1.78 & 0.4\\
-\hspace{1em}latin\_american * Central & 0.69 & 0.28, 1.71 & 0.4 & 0.66 & 0.22, 1.93 & 0.4\\
-\hspace{1em}mixed * Central & 0.63 & 0.30, 1.35 & 0.2 & 0.73 & 0.24, 2.23 & 0.6\\
-\hspace{1em}other * Central & 0.98 & 0.49, 1.95 & >0.9 & 0.81 & 0.36, 1.81 & 0.6\\
-\hspace{1em}south\_asian * Central & 0.71 & 0.34, 1.47 & 0.4 & 0.55 & 0.25, 1.21 & 0.14\\
-\hspace{1em}arab\_middle\_eastern * East & 0.57 & 0.20, 1.63 & 0.3 & 0.43 & 0.13, 1.43 & 0.2\\
-\hspace{1em}black * East & 0.84 & 0.37, 1.89 & 0.7 & 0.86 & 0.35, 2.11 & 0.7\\
-\hspace{1em}east\_asian\_pacific\_islander * East & 0.82 & 0.30, 2.24 & 0.7 & 0.86 & 0.29, 2.54 & 0.8\\
-\hspace{1em}indigenous * East & 0.56 & 0.21, 1.45 & 0.2 & 0.70 & 0.23, 2.11 & 0.5\\
-\hspace{1em}latin\_american * East & 0.80 & 0.28, 2.25 & 0.7 & 1.03 & 0.31, 3.40 & >0.9\\
-\hspace{1em}mixed * East & 0.70 & 0.31, 1.60 & 0.4 & 0.92 & 0.28, 3.05 & 0.9\\
-\hspace{1em}other * East & 0.87 & 0.38, 2.02 & 0.7 & 1.08 & 0.40, 2.91 & 0.9\\
-\hspace{1em}south\_asian * East & 0.50 & 0.20, 1.24 & 0.14 & 0.51 & 0.19, 1.42 & 0.2\\
-\hspace{1em}arab\_middle\_eastern * West & 1.18 & 0.49, 2.82 & 0.7 & 1.02 & 0.37, 2.77 & >0.9\\
-\hspace{1em}black * West & 0.78 & 0.36, 1.67 & 0.5 & 0.77 & 0.32, 1.82 & 0.5\\
-\hspace{1em}east\_asian\_pacific\_islander * West & 0.54 & 0.24, 1.21 & 0.14 & 0.52 & 0.20, 1.36 & 0.2\\
-\hspace{1em}indigenous * West & 0.45 & 0.19, 1.05 & 0.064 & 0.40 & 0.14, 1.12 & 0.082\\
-\hspace{1em}latin\_american * West & 0.99 & 0.39, 2.48 & >0.9 & 0.96 & 0.33, 2.79 & >0.9\\
-\hspace{1em}mixed * West & 0.53 & 0.25, 1.10 & 0.088 & 0.37 & 0.12, 1.15 & 0.085\\
-\hspace{1em}other * West & 0.55 & 0.27, 1.12 & 0.10 & 0.41 & 0.18, 0.92 & 0.032\\
-\hspace{1em}south\_asian * West & 0.50 & 0.24, 1.08 & 0.078 & 0.41 & 0.18, 0.95 & 0.037\\*
+\hspace{1em}Arab/Middle Eastern * Central & 0.76 & 0.33, 1.71 & 0.5 & 0.66 & 0.25, 1.69 & 0.4\\
+\hspace{1em}Black * Central & 0.48 & 0.23, 1.02 & 0.055 & 0.44 & 0.19, 0.99 & 0.046\\
+\hspace{1em}East Asian/Pacific Islander * Central & 0.84 & 0.38, 1.89 & 0.7 & 0.98 & 0.38, 2.53 & >0.9\\
+\hspace{1em}Indigenous * Central & 0.60 & 0.24, 1.50 & 0.3 & 0.63 & 0.22, 1.78 & 0.4\\
+\hspace{1em}Latin American * Central & 0.69 & 0.28, 1.71 & 0.4 & 0.66 & 0.22, 1.93 & 0.4\\
+\hspace{1em}Mixed * Central & 0.63 & 0.30, 1.35 & 0.2 & 0.73 & 0.24, 2.23 & 0.6\\
+\hspace{1em}Other * Central & 0.98 & 0.49, 1.95 & >0.9 & 0.81 & 0.36, 1.81 & 0.6\\
+\hspace{1em}South Asian * Central & 0.71 & 0.34, 1.47 & 0.4 & 0.55 & 0.25, 1.21 & 0.14\\
+\hspace{1em}Arab/Middle Eastern * East & 0.57 & 0.20, 1.63 & 0.3 & 0.43 & 0.13, 1.43 & 0.2\\
+\hspace{1em}Black * East & 0.84 & 0.37, 1.89 & 0.7 & 0.86 & 0.35, 2.11 & 0.7\\
+\hspace{1em}East Asian/Pacific Islander * East & 0.82 & 0.30, 2.24 & 0.7 & 0.86 & 0.29, 2.54 & 0.8\\
+\hspace{1em}Indigenous * East & 0.56 & 0.21, 1.45 & 0.2 & 0.70 & 0.23, 2.11 & 0.5\\
+\hspace{1em}Latin American * East & 0.80 & 0.28, 2.25 & 0.7 & 1.03 & 0.31, 3.40 & >0.9\\
+\hspace{1em}Mixed * East & 0.70 & 0.31, 1.60 & 0.4 & 0.92 & 0.28, 3.05 & 0.9\\
+\hspace{1em}Other * East & 0.87 & 0.38, 2.02 & 0.7 & 1.08 & 0.40, 2.91 & 0.9\\
+\hspace{1em}South Asian * East & 0.50 & 0.20, 1.24 & 0.14 & 0.51 & 0.19, 1.42 & 0.2\\
+\hspace{1em}Arab/Middle Eastern * West & 1.18 & 0.49, 2.82 & 0.7 & 1.02 & 0.37, 2.77 & >0.9\\
+\hspace{1em}Black * West & 0.78 & 0.36, 1.67 & 0.5 & 0.77 & 0.32, 1.82 & 0.5\\
+\hspace{1em}East Asian/Pacific Islander * West & 0.54 & 0.24, 1.21 & 0.14 & 0.52 & 0.20, 1.36 & 0.2\\
+\hspace{1em}Indigenous * West & 0.45 & 0.19, 1.05 & 0.064 & 0.40 & 0.14, 1.12 & 0.082\\
+\hspace{1em}Latin American * West & 0.99 & 0.39, 2.48 & >0.9 & 0.96 & 0.33, 2.79 & >0.9\\
+\hspace{1em}Mixed * West & 0.53 & 0.25, 1.10 & 0.088 & 0.37 & 0.12, 1.15 & 0.085\\
+\hspace{1em}Other * West & 0.55 & 0.27, 1.12 & 0.10 & 0.41 & 0.18, 0.92 & 0.032\\
+\hspace{1em}South Asian * West & 0.50 & 0.24, 1.08 & 0.078 & 0.41 & 0.18, 0.95 & 0.037\\*
 \multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\
@@
 
-There existence of healthcare disparities within Ontario is a topic of
-particular interest as it motivated the recent change in the healthcare
-system of this region, which a system based on Health Regions in late
-2019 in an attempt to address the disparities from the LHIN
-model\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
-In this context, analyzing COVID-19 vaccination estimates within the
-province is important as they can provide an initial assessment of
-variations that might need to be addressed by decision-makers.
+The existence of healthcare disparities in Ontario motivated the recent
+change in the healthcare system of the province, which changed the LHIN
+model for a Health Region model in late
+2019\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
+In this context, analyzing COVID-19 vaccination estimates between the
+Health Regions can serve as an indicator of ongoing intra-provincial
+disparities that may need to be addressed to ensure that the Health
+Region model is able to improve health access for the inhabitants of
+Ontario, which faces unique challenges due to its condition as the most
+populated and the most ethnically diverse province of Canada.
 
 Our results indicate that across the most densely populated Health
-Regions of Ontario, almost three quarters of the surveyed individuals
-reported having received the first dose of the COVID-19 vaccine
+Regions of Ontario, almost three quarters of surveyed individuals
+reported to have received the first dose of the COVID-19 vaccine
 (Table~\ref{tbl-descriptive-stats}). It is worth mentioning that
@@
 2021)\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. In
-this regard, differences are to be expected because the survey
-represents a random sample from the population, and therefore, the
-sampling process is likely to cause variations between the values from
-the survey and province-wide estimates.
-
-We believe that these differences did not have a significant impact in
-our analyses, because our results indicate that there were no
-significant differences in vaccination odds among the age groups
-considered in the survey, in agreement with the overall vaccination
-rates reported for Canada, which have been relatively higher when
-compared to other high income
+this case, differences between the survey and province-wide estimates
+are to be expected as the data from survey represents a random sample
+from the overall population.
+
+However, this variation did not lead to discrepant results, as the
+estimates from the model indicate that no significant differences in
+vaccination odds were identified among the age groups analyzed, in
+agreement with both overall vaccination rates reported for Canada, which
+have been relatively higher when compared to other high income
 countries\textsuperscript{\protect\hyperlink{ref-dube2022}{40}}, and
@@
 studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{41}}.
-Moreover, vaccination rates within each age in the dataset are in good
-agreement with province-wide estimates (e.g., a rate of 95\% for those
-with 61 years of age, Supplementary Table A-6). Additionally, when
-overall vaccination rates for the province are disaggregated, it can be
-seen that regional differences were present during the period analyzed,
-which can be masked by province-wide estimates. For example, the Public
+In other words, although vaccination rates obtained from the survey were
+slightly lower than the provincial estimates, these values still
+represented a valid approximation; this notion is reinforced by the
+consistency in the proportion of vaccination rates
+(Table~\ref{tbl-descriptive-stats}) and vaccination odds
+(Table~\ref{tbl-model}) across the period covered by the survey, which
+follow the vaccination rates from Public Health Ontario and which
+indicate that due to the relatively high coverage achieved in the
+population at that point, there were no abrupt shift in the trends,
+which increased by around 3\% across all age groups during the three
+months (October, November, and December of
+2021\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}).
+Moreover, there was good agreement between vaccination rates within each
+age in the raw dataset and province-wide estimates (e.g., a rate of 95\%
+for those with 61 years of age, Supplementary Table A-6, which is
+similar to the value reported by Public Health Ontario). It is also
+important to mention that regional differences can be masked by overall
+estimates, as when overall vaccination rates for the province (from
+Public Health Ontario) are disaggregated, it can be seen that regional
+differences during the period analyzed existed. For example, the Public
 Health Unit of Lambton (a region within the West Health Region) and the
@@
 rates (78\%) for those 50 years of age and older at the beginning of the
-period of interest, whereas other regions had higher vaccination
-rates\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. To
-this day, differences in vaccination rates within the province are
-present, because according to Public Health Ontario, as of March of 2023
-some regions have less than 75\% vaccination
+period of interest, in contrast with other regions that had vaccination
+rates above
+80\%\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. To this
+day, differences in vaccination rates within the province continue, as
+according to Public Health Ontario, as of March of 2023 some regions
+still have less than 75\% vaccination
 rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{42}}.
-Overall, these reasons indicate that the estimates obtained from the
-data are in good agreement with the trends from the population.
 
@@
 across Ontario. Moreover, because significant lower odds of vaccination
-were also identified other minority groups, this provides a rationale
+were also identified in other minority groups, this provides a rationale
 for future studies that explore how vaccination uptake varies across

### 2023-04-17 code/data_preparation.R

@@ -14,10 +14,6 @@ clean_data<-clean_data %>%
   select(-Municipal.status) %>%
   rename(Race=race,Age_group=age_group,Income=income_ord)
 
-#make variables of interest factors
-clean_data<-clean_data %<>%
-  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)
-
 # recode the response of interest for categorical analysis
 clean_data<-clean_data %>%
   mutate(first_dose_m=case_when(
@@ -29,23 +25,17 @@ clean_data<-clean_data %>%
 clean_data <-clean_data %>%
   filter(Health_Region!="North East" , Health_Region!="North West")
 
-
-#eliminate the levels of the removed health regions
-clean_data <- droplevels(clean_data)
-
-
 ## Recode income
-
 
 clean_data<-clean_data%>%
   mutate(income_ord = case_when(
-    income %in% "under_15000" ~ "under_25000",
-    income %in% "15000_24999" ~ "under_25000",
-    income %in% "25000_39999" ~ "25000_59999",
-    income %in% "40000_59999" ~ "25000_59999",
-    income %in% "60000_89999" ~ "60000_and_above",
-    income %in% "90000_109999" ~ "60000_and_above",
-    income %in%"over_110000" ~ "60000_and_above",
+    income %in% "under_15000" ~ "under 25000",
+    income %in% "15000_24999" ~ "under 25000",
+    income %in% "25000_39999" ~ "25000-59999",
+    income %in% "40000_59999" ~ "25000-59999",
+    income %in% "60000_89999" ~ "60000 and above",
+    income %in% "90000_109999" ~ "60000 and above",
+    income %in%"over_110000" ~ "60000 and above",
   ))%>%
   mutate(income_ord=as.factor(income_ord))
 
@@ -53,16 +43,29 @@ clean_data<-clean_data%>%
 
 clean_data<-clean_data%>%
   mutate(Age_group_ord = case_when(
-    Age_group %in% "16_24" ~ "16_34",
-    Age_group %in% "25_34" ~ "16_34",
-    Age_group %in% "35_44" ~ "35_54",
-    Age_group %in% "45_54" ~ "35_54",
-    Age_group %in% "55_64" ~ "55_and_over",
-    Age_group %in% "65_and_over" ~ "55_and_over",
+    Age_group %in% "16_24" ~ "16-34",
+    Age_group %in% "25_34" ~ "16-34",
+    Age_group %in% "35_44" ~ "35-54",
[…]

### 2023-04-17 code/raking.R

@@ -1,24 +1,24 @@
 #Each of these dataframes provides the population totals for the different variables and their categories
 
 
-Ages<-data.frame(Age_group_ord=c("16_34",
-                                 "35_54",
-                                 "55_and_over"),
+Ages<-data.frame(Age_group_ord=c("16-34",
+                                 "35-54",
+                                 "55 and over"),
                  Freq=c(3442815,
                         3725233,
                         4088342))
 
 
 
-Races<-data.frame(Race=c("arab_middle_eastern",
-                         "black",
-                         "east_asian_pacific_islander",
-                         "indigenous",
-                         "latin_american",
-                         "mixed",
-                         "other",
-                         "south_asian",
-                         "white_caucasian"),
+Races<-data.frame(Race=c("Arab/Middle Eastern",
+                         "Black",
+                         "East Asian/Pacific Islander",
+                         "Indigenous",
+                         "Latin American",
+                         "Mixed",
+                         "Other",
+                         "South Asian",
+                         "White/Caucasian"),
                   Freq=c(212782,
                          638346,
                          886592,
@@ -29,9 +29,9 @@ Races<-data.frame(Race=c("arab_middle_eastern",
                          1166361,
                          9118079 ))
 
-Incomes<-data.frame(income_ord=c("under_25000",
-                             "25000_59999",
-                             "60000_and_above"),
+Incomes<-data.frame(income_ord=c("under 25000",
+                             "25000-59999",
+                             "60000 and above"),
                     Freq=c(682331,
                            1395676,
                            3091164

### 2023-04-17 code/regression.R

@@ -9,12 +9,6 @@ m0<-svyglm(first_dose_m~Age_group_ord+Month+income_ord+Race+Health_Region+Race*i
 
 # model for corrected data
 
-m1<-svyglm(first_dose_m~Age_group_ord+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
-           design=a1_rake,
-           family = quasibinomial(), 
-           control= list(maxit=25))
-
-
 m2<-svyglm(first_dose_m~Age_group_ord+Month+income_ord+Race+Health_Region+Race*income_ord+Race*Health_Region,
            design=a1_rake,
            family = quasibinomial(), 

### 2023-04-23 manuscript_v4/02-methods.qmd
Started v4 to implement final changes before submission

--- 2023-03-24 manuscript_v2/02-methods.qmd
@@
 We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto, and which ran between September 30th, 2021 and January 17th,2022.
-The survey collected socio-economic information from participants (@tbl-covariates), recorded their location (using the nearest municipality), and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".
+The survey collected socio-economic information from participants (@tbl-covariates), recorded their location (using the nearest municipality), the date of access to the survey, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".
 The original dataset contained 39,029 entries (where each entry corresponded to a unique respondent).
 
-This dataset was cleaned to remove outliers that were identified during preliminary analyses, and processing the geographical information in the survey (city where the survey was responded) in order to match each city to its correspondent Health Region.
+This dataset was cleaned to remove outliers that were identified during preliminary analyses, and the geographical information in the survey (city where the survey was responded) was used to match each city to its correspondent Health Region.
 
@@
 Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions.
-The low representation of these Health Regions in the dataset is noticeable in @fig-map,which shows that responses from these areas came from a relatively low number of cities when compared to the most populated Health Regions, such as the Toronto or Central Regions.
+The low representation of these Health Regions in the dataset is noticeable in @fig-map, which shows that responses from these areas came from a relatively low number of cities when compared to the most populated Health Regions, such as the Toronto or Central Regions.
 We omitted the North East and North East Health Regions from further analyses due to the low number of entries.
-Therefore, the total number of unique entries used for analysis was 3,551 which included the East, Central, Toronto, and West Health Regions and that covered the period between September 30th,2021 and December 12th, 2021.
+Therefore, the total number of unique entries used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021.
 
@@
 
-We used a logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @tbl-covariates, the Health Regions from Ontario indicated in @sec-data, and the interactions between Race and Health Region, and Race and income, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
+We used a logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @tbl-covariates, the month when the survey was answered, the Health Regions from Ontario indicated in @sec-data, and the interactions between Race and Health Region, and Race and income, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
 
 The model was fitted first to the clean dataset to obtain uncorrected estimates.
-Additionally, because we identified differences between the proportions of all the socio-economic factors included in the analysis (@tbl-covariates) and the Census data for Ontario, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using Census socio-economic data and Health Region population totals, in order to obtain corrected estimates from the model.
+Because we identified differences between the proportions of all the socio-economic factors included in the analysis (@tbl-covariates) and the Census data for Ontario, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using Census socio-economic data and Health Region population totals, in order to obtain corrected estimates from the model.
 Details regarding the correction can be found in the Appendix.

### 2023-04-23 manuscript_v4/03-results.qmd
Started v4 to implement final changes before submission

--- 2023-04-16 manuscript_v3/03-results.qmd
@@
 The Health Region with highest representation in the survey was Toronto, accounting for 1,323 entries (37.3\% of the total), with a vaccination rate of 72\%.
+Across the three months covered by the survey, the highest number of entries was in October of 2021, with a total of 1,732 respondents, which corresponded to 49\% of all the survey entries.
+Additionally, the proportion of respondents who reported being vaccinated was similar in all months, ranging between 72\% and 75\%, with the highest value corresponding to December of 2021.
+
 Regarding race/ethnicity, individuals that identified as White/Caucasian represented 37\% of all entries (1,312) and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine.
@@
                  Age_group_ord ~ "Age Group",
-                 income_ord="Income")
+                 income_ord="Income (CAD)")
   ) %>% bold_labels()
@@
                             Age_group_ord ~ "Age Group",
-                            income_ord="Income")) %>%
+                            income_ord="Income (CAD)")) %>%
   bold_labels()
 
-t1<-m1 %>%
-  tbl_regression(exponentiate = TRUE,
-                 label=list(Health_Region ~ "Health Region",
-                            Age_group_ord ~ "Age Group",
-                            income_ord="Income")) %>%
-  bold_labels()
+# t1<-m1 %>%
+#   tbl_regression(exponentiate = TRUE,
+#                  label=list(Health_Region ~ "Health Region",
+#                             Age_group_ord ~ "Age Group",
+#                             income_ord="Income")) %>%
+#   bold_labels()
 
@@
                             Age_group_ord ~ "Age Group",
-                            income_ord="Income")) %>%
+                            income_ord="Income (CAD)")) %>%
   bold_labels()

### 2023-04-23 manuscript_v4/03-results.qmd
need to fix modelplot figure

@@
                  income_ord="Income (CAD)")
-  ) %>% bold_labels()
+  ) %>% add_p("prop.test")%>%
+  bold_labels()
 
@@
 
-t0<-m0 %>%
-  tbl_regression(exponentiate = TRUE,
-                 label=list(Health_Region ~ "Health Region",
-                            Age_group_ord ~ "Age Group",
-                            income_ord="Income (CAD)")) %>%
-  bold_labels()
+# t0<-m0 %>%
+#   tbl_regression(exponentiate = TRUE,
+#                  label=list(Health_Region ~ "Health Region",
+#                             Age_group_ord ~ "Age Group",
+#                             income_ord="Income (CAD)")) %>%
+#   bold_labels()
+
+modelplot(m0)+
+  aes(color=ifelse(p.value<0.05, "Significant","Non-significant"))+
+  scale_color_manual(values=c("black","red"))
+
+modelplot(m2)+
+  aes(color=ifelse(p.value<0.05, "Significant","Non-significant"))+
+  scale_color_manual(values=c("black","red"))
 

### 2023-04-24 manuscript_v4/01-background.qmd
updating v4

--- 2023-04-16 manuscript_v3/01-background.qmd
@@
 This problematic in the rollout of the COVID-19 vaccines is a multifaceted issue resulting from, among other things, the development of new variants due to inadequate public health measures [@li2021],  inequality in vaccine access between high- and low-income countries[@gerretsen2021; @tamey2022], vaccine hesitancy[@nafilyan2021], and differences in vaccination uptake across different segments of the population [@malik2020].
-In particular, it is well established that differences in vaccination uptake have been present even in countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within racial minorities (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021].
+In particular, it is well established that differences in vaccination uptake have been present even in countries that have had ample access to vaccines since 2020 (such as the US, the UK, and Canada),  where lower vaccine uptake has been observed  within certain racial groups (i.e., individuals that identify as Black, Asian, or Indigenous), and in individuals within low income brackets [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021].
 Reasons given for lower vaccine uptake in these cases have included medical mistrust due to systemic medical racism [@stoler2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
-Moreover, in the case of Canada, lower vaccine uptake has been observed in young individuals, those with a low educational level, households with children, those without a regular healthcare provider, individuals that identify as part of a visible minorities or Indigenous, and those that are financially unstable [@guay2022;@muhajarine2021;@hussain2022].
+Moreover, in the case of Canada, lower vaccine uptake has been observed in young individuals, those with a low educational level, households with children, those without a regular healthcare provider, individuals that identify as part of certain equity-deserving groups, and those with a low household income [@guay2022;@muhajarine2021;@hussain2022].
 
@@
 In late 2019, the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East) in an effort to improve the healthcare system of the province [@dong2022].
-Because the adoption of the Health Regions is relatively recent there is an ongoing need to analyze the impact of this measure and identify the existence of intra-regional differences that might exist, and which could be specially important in the context of the COVID-19 pandemic.
+Because the adoption of the Health Regions is relatively recent, there is an ongoing need to analyze the impact of this measure and identify disparities in health access that might exist across the Health Regions, which can be specially important in the context of the COVID-19 pandemic.
 
-In this study, we analyzed differences in self-reported COVID-19 vaccination status in Ontario using socio-economic (e.g., income, racial/ethnic identification), temporal (through a consecutive period of three months in late 2021), and spatial  information (using the Health Regions of Ontario) to identify the existence of differences that might need to be addressed to ensure that the healthcare system of the province is more inclusive and that responds to the needs of its most vulnerable population.
+Therefore, in this study we hypothesized that there were differences in vaccination uptake between the Health Regions of Ontario during the last quarter of 2021.To better contextualize potential differences, we also included socio-economic factors in our analysis in order to identify in which demographic groups these differences were significant.

### 2023-04-24 manuscript_v4/02-methods.qmd
updating v4

@@
 This dataset was cleaned to remove outliers that were identified during preliminary analyses, and the geographical information in the survey (city where the survey was responded) was used to match each city to its correspondent Health Region.
-
-| Variable             | Levels                                            |            
-|:---------------------|---------------------------------------------------|
-| Age group            | 16-34,35-54,55 and over                           |
-| Income bracket (CAD) | under 25,000, 25,000-59,999, 60,000 and above     |
-| Race/ethnicity       | Arab/Middle Eastern, Black, East Asian/Pacific Islander, Indigenous, Latin American, Mixed, South Asian, White Caucasian, Other |
-
-: Socio-economic factors from the Fields COVID-19 survey {#tbl-covariates}
 

### 2023-04-24 manuscript_v4/03-results.qmd
updating v4

--- 2023-04-23 manuscript_v4/03-results.qmd
@@
                  income_ord="Income (CAD)")
-  ) %>% bold_labels()
+  ) %>% add_p()%>%
+  bold_labels()
 
@@
 
-\footnotesize
-\renewcommand{\arraystretch}{0.5}
-```{r,model-table}
+```{r,model-uncorr-fig}
 #| message: false
@@
 #| echo: false
-#| label: tbl-model
-#| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"
+#| fig-align: left
+#| fig-width: 7
+#| fig-height: 7
+#| label: fig-model-uncorr
+#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. For the interaction terms, only those with a statistically significant values are shown."
 
@@
 
-# create tables with the model outputs
+#create dataframe from modelplot, remove names of factors, drop the intercept
 
-t0<-m0 %>%
-  tbl_regression(exponentiate = TRUE,
-                 label=list(Health_Region ~ "Health Region",
-                            Age_group_ord ~ "Age Group",
-                            income_ord="Income (CAD)")) %>%
-  bold_labels()
+mplot1<-modelplot(m0,draw=FALSE,coef_omit = "Intercept") %>%
+  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
 
-# t1<-m1 %>%
-#   tbl_regression(exponentiate = TRUE,
-#                  label=list(Health_Region ~ "Health Region",
-#                             Age_group_ord ~ "Age Group",
-#                             income_ord="Income")) %>%
-#   bold_labels()
+#add column by factor 
+mplot1$factor<-c(rep("Age",2),
+                 rep("Month",2),
+                 rep("Income",2),
+                 rep("Race/Ethnicity",8),
+                 rep("Health Region",3),
+                 rep("Income and\nRace/Ethnicity",16),
+                 rep("Race/Ethnicity\nand\nHealth Region",24)
+                 )
 
-t2<-m2 %>%
-  tbl_regression(exponentiate = TRUE,
-                 label=list(Health_Region ~ "Health Region",
-                            Age_group_ord ~ "Age Group",
-                            income_ord="Income (CAD)")) %>%
-  bold_labels()
+#order factors for plot
+#mplot1$term_ord <- factor(mplot1$term,ordered=TRUE,levels=mplot1$term)
 
-# ggplot(data=clean_data,aes(y=first_dose))+
-#   geom_bar()+facet_wrap(~Month)
+mplot1$factor <- factor(mplot1$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
 
-tbl_merge <-
-  tbl_merge(
-    tbls = list(t0, t2),
-    tab_spanner = c("**Uncorrected**", "**Corrected**")
-  )%>%
-  bold_labels()%>%
-  as_kable_extra(booktabs = TRUE, longtable=T, linesep="") %>%
-  # reduce font size to make table fit. 
-  # you may also use the `latex_options = "scale_down" argument here.
-  kableExtra::kable_styling(latex_options = "repeat_header")
+# remove non-significant interaction factors
 
-tbl_merge
+test<-mplot1%>%
+  filter(str_detect(term,':') & p.value>0.05)
 
-# m1 %>%
-#   tbl_regression(exponentiate = TRUE,
-#                  label=list(Health_Region ~ "Health Region",
-#                  Age_group_ord ~ "Age Group",
-#                  income_ord="Income")) %>%
-#     bold_labels()%>%
-#   as_kable_extra(booktabs = TRUE, longtable=T, linesep="") %>%
-#   # reduce font size to make table fit. 
-#   # you may also use the `latex_options = "scale_down" argument here.
-#   kableExtra::kable_styling(latex_options = "repeat_header")
+#remove non-significant interactions from dataframe
+mplot1<-anti_join(mplot1,test)
+
+mplot1<-mplot1%>%
+mutate(term=str_replace_all(term,':'," and "))
+
+#make plot
+mplot1 %>% 
+ggplot(aes(y=term))+
+  geom_vline(xintercept = 0, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate))+
+  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
+  scale_y_discrete(limits=rev)+
+  xlab("Coefficient estimates and 95% confidence intervals")+
+  ylab("")+
+  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y",labeller = label_wrap_gen(multi_line = TRUE))+
+  theme_classic()
 ```
 
-\normalsize
+```{r,model-corr-fig}
+#| message: false
+#| warning: false
+#| echo: false
+#| fig-align: left
+#| fig-width: 7
+#| fig-height: 7
+#| label: fig-model-corr
+#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. For the interaction terms, only those with a statistically significant values are shown."
+
+#create dataframe from modelplot, remove names of factors, drop the intercept
+
+mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept") %>%
+  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
+
+#add column by factor 
+mplot2$factor<-c(rep("Age",2),
+                 rep("Month",2),
+                 rep("Income",2),
+                 rep("Race/Ethnicity",8),
+                 rep("Health Region",3),
+                 rep("Income and\nRace/Ethnicity",16),
+                 rep("Race/Ethnicity\nand\nHealth Region",24)
+                 )
+
+#order factors for plot
+#mplot2$term_ord <- factor(mplot2$term,ordered=TRUE,levels=mplot2$term)
+
+mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
+
+# remove non-significant interaction factors
+
+test<-mplot2%>%
+  filter(str_detect(term,':') & p.value>0.05)
+
+mplot2<-anti_join(mplot2,test)
+
+mplot2<-mplot2%>%
+mutate(term=str_replace_all(term,':'," and "))
+
+mplot2 %>% 
+ggplot(aes(y=term))+
+  geom_vline(xintercept = 0, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate))+
+  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
+  scale_y_discrete(limits=rev)+
+  xlab("Coefficient estimates and 95% confidence intervals")+
+  ylab("")+
+  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
+  theme_classic()
+```

### 2023-04-24 manuscript_v4/main.qmd
updating v4

--- 2023-04-17 manuscript_v3/main.qmd
@@
 In this study, we examined the relationship between vaccination status (having received the first dose of a COVID-19 vaccine), and different socio-economic and geographical factors.
-Our results show differences in vaccination due to race/ethnicity, income, Health Regions (geographical areas used for health service access in Ontario), and their interactions.
-In particular, we show that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs and 95% CIs=0.31[0.14,0.68], 0.32[0.17,0.60], 0.27[0.11,0.66], respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.08, 95% CI=[1.27,7.47]), Black (OR=3.15,95% CI=[1.43,6.92]), Latin American (OR=2.81, 95% CI=[1.
-04,7.59]), or that belonged to other minority groups  (OR=4.63,95% CI=[2.34,9.13]) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets.
-Finally, we also identified lower odds of vaccination within certain minority groups in the West Health Region.
-This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population that have been largely impacted by the pandemic.
+Our results show that during the last three months of 2021, individuals in certain equity-deserving groups (visible minorities) were three times less likely to be vaccinated than White/Caucasian individuals across the province and that in some cases, within these groups individuals in low income brackets had significantly higher odds of vaccination when compared to their peers in high income brackets.
+Finally, we identified significantly lower odds of vaccination in the West Health Region of Ontario within certain equity-deserving groups.
+This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population of Ontario that have been largely impacted by the pandemic.
 

### 2023-04-24 manuscript_v4/main.tex
updating v4

--- 2023-04-17 manuscript_v3/main.tex
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, enhanced, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 received the first dose of a COVID-19 vaccine), and different
-socio-economic and geographical factors. Our results show differences in
-vaccination due to race/ethnicity, income, Health Regions (geographical
-areas used for health service access in Ontario), and their
-interactions. In particular, we show that individuals who identified as
-Arab/Middle Eastern, Black, or Latin American, had significantly lower
-odds of vaccination than White/Caucasian individuals (ORs and 95\%
-CIs=0.31{[}0.14,0.68{]}, 0.32{[}0.17,0.60{]}, 0.27{[}0.11,0.66{]},
-respectively), and that individuals with a household income below CAD
-25,000 who identified as Arab/Middle Eastern (OR=3.08, 95\%
-CI={[}1.27,7.47{]}), Black (OR=3.15,95\% CI={[}1.43,6.92{]}), Latin
-American (OR=2.81, 95\% CI={[}1. 04,7.59{]}), or that belonged to other
-minority groups (OR=4.63,95\% CI={[}2.34,9.13{]}) had higher odds of
-vaccination than individuals from the same racial/ethnic group in higher
-income brackets. Finally, we also identified lower odds of vaccination
-within certain minority groups in the West Health Region. This study
-shows that there is an ongoing need to better understand and address
-differences in vaccination uptake across diverse segments of the
-population that have been largely impacted by the pandemic.
+socio-economic and geographical factors. Our results show that during
+the last three months of 2021, individuals in certain equity-deserving
+groups (visible minorities) were three times less likely to be
+vaccinated than White/Caucasian individuals across the province and that
+in some cases, within these groups individuals in low income brackets
+had significantly higher odds of vaccination when compared to their
+peers in high income brackets. Finally, we identified significantly
+lower odds of vaccination in the West Health Region of Ontario within
+certain equity-deserving groups. This study shows that there is an
+ongoing need to better understand and address differences in vaccination
+uptake across diverse segments of the population of Ontario that have
+been largely impacted by the pandemic.
 
@@
 vaccines since 2020 (such as the US, the UK, and Canada), where lower
-vaccine uptake has been observed within racial minorities (i.e.,
+vaccine uptake has been observed within certain racial groups (i.e.,
 individuals that identify as Black, Asian, or Indigenous), and in
@@
 with children, those without a regular healthcare provider, individuals
-that identify as part of a visible minorities or Indigenous, and those
-that are financially
-unstable\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.
+that identify as part of certain equity-deserving groups, and those with
+a low household
+income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.
 
@@
 of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
-Because the adoption of the Health Regions is relatively recent there is
-an ongoing need to analyze the impact of this measure and identify the
-existence of intra-regional differences that might exist, and which
-could be specially important in the context of the COVID-19 pandemic.
-
-In this study, we analyzed differences in self-reported COVID-19
-vaccination status in Ontario using socio-economic (e.g., income,
-racial/ethnic identification), temporal (through a consecutive period of
-three months in late 2021), and spatial information (using the Health
-Regions of Ontario) to identify the existence of differences that might
-need to be addressed to ensure that the healthcare system of the
-province is more inclusive and that responds to the needs of its most
-vulnerable population.
+Because the adoption of the Health Regions is relatively recent, there
+is an ongoing need to analyze the impact of this measure and identify
+disparities in health access that might exist across the Health Regions,
+which can be specially important in the context of the COVID-19
+pandemic.
+
+Therefore, in this study we hypothesized that there were differences in
+vaccination uptake between the Health Regions of Ontario during the last
+quarter of 2021.To better contextualize potential differences, we also
+included socio-economic factors in our analysis in order to identify in
+which demographic groups these differences were significant.
 
@@
 2021 and January 17th,2022. The survey collected socio-economic
-information from participants (Table~\ref{tbl-covariates}), recorded
+information from participants (\textbf{?@tbl-covariates}), recorded
 their location (using the nearest municipality), the date of access to
@@
 correspondent Health Region.
-
-\hypertarget{tbl-covariates}{}
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.3014}}
-  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.6986}}@{}}
-\caption{\label{tbl-covariates}Socio-economic factors from the Fields
-COVID-19 survey}\tabularnewline
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Variable
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Levels
-\end{minipage} \\
-\midrule()
-\endfirsthead
-\toprule()
-\begin{minipage}[b]{\linewidth}\raggedright
-Variable
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-Levels
-\end{minipage} \\
-\midrule()
-\endhead
-Age group & 16-34,35-54,55 and over \\
-Income bracket (CAD) & under 25,000, 25,000-59,999, 60,000 and above \\
-Race/ethnicity & Arab/Middle Eastern, Black, East Asian/Pacific
-Islander, Indigenous, Latin American, Mixed, South Asian, White
-Caucasian, Other \\
-\bottomrule()
-\end{longtable}
 
@@
 vaccination depending on the socio-economic factors described in
-Table~\ref{tbl-covariates}, the month when the survey was answered, the
+\textbf{?@tbl-covariates}, the month when the survey was answered, the
 Health Regions from Ontario indicated in Section~\ref{sec-data}, and the
@@
 all the socio-economic factors included in the analysis
-(Table~\ref{tbl-covariates}) and the Census data for Ontario, we used an
+(\textbf{?@tbl-covariates}) and the Census data for Ontario, we used an
 iterative proportional fitting procedure
@@
 
-\begin{tabular}{lcc}
+\begin{tabular}{lccc}
 \toprule
-\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,591\\
+\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,591 & \textbf{p-value}\\
 \midrule
-\textbf{Income (CAD)} &  & \\
-\hspace{1em}60000 and above & 305 (23\%) & 1,048 (77\%)\\
-\hspace{1em}25000-59999 & 253 (28\%) & 636 (72\%)\\
-\hspace{1em}under 25000 & 400 (31\%) & 907 (69\%)\\
-\textbf{Age Group} &  & \\
-\hspace{1em}16-34 & 409 (27\%) & 1,111 (73\%)\\
-\hspace{1em}35-54 & 252 (26\%) & 712 (74\%)\\
-\hspace{1em}55 and over & 297 (28\%) & 768 (72\%)\\
-\textbf{Health Region} &  & \\
-\hspace{1em}Toronto & 371 (28\%) & 952 (72\%)\\
-\hspace{1em}Central & 224 (28\%) & 580 (72\%)\\
-\hspace{1em}East & 135 (23\%) & 448 (77\%)\\
-\hspace{1em}West & 228 (27\%) & 611 (73\%)\\
-\textbf{Month} &  & \\
-\hspace{1em}October & 469 (27\%) & 1,263 (73\%)\\
-\hspace{1em}November & 376 (28\%) & 980 (72\%)\\
-\hspace{1em}December & 113 (25\%) & 348 (75\%)\\
-\textbf{Race} &  & \\
-\hspace{1em}White/Caucasian & 233 (18\%) & 1,079 (82\%)\\
-\hspace{1em}Arab/Middle Eastern & 76 (36\%) & 138 (64\%)\\
-\hspace{1em}Black & 114 (38\%) & 184 (62\%)\\
-\hspace{1em}East Asian/Pacific Islander & 69 (23\%) & 234 (77\%)\\
-\hspace{1em}Indigenous & 76 (40\%) & 115 (60\%)\\
-\hspace{1em}Latin American & 69 (38\%) & 111 (62\%)\\
-\hspace{1em}Mixed & 105 (34\%) & 205 (66\%)\\
-\hspace{1em}Other & 128 (35\%) & 239 (65\%)\\
-\hspace{1em}South Asian & 88 (24\%) & 286 (76\%)\\
+\textbf{Income (CAD)} &  &  & <0.001\\
+\hspace{1em}60000 and above & 305 (23\%) & 1,048 (77\%) & \\
+\hspace{1em}25000-59999 & 253 (28\%) & 636 (72\%) & \\
+\hspace{1em}under 25000 & 400 (31\%) & 907 (69\%) & \\
+\textbf{Age Group} &  &  & 0.7\\
+\hspace{1em}16-34 & 409 (27\%) & 1,111 (73\%) & \\
+\hspace{1em}35-54 & 252 (26\%) & 712 (74\%) & \\
+\hspace{1em}55 and over & 297 (28\%) & 768 (72\%) & \\
+\textbf{Health Region} &  &  & 0.14\\
+\hspace{1em}Toronto & 371 (28\%) & 952 (72\%) & \\
+\hspace{1em}Central & 224 (28\%) & 580 (72\%) & \\
+\hspace{1em}East & 135 (23\%) & 448 (77\%) & \\
+\hspace{1em}West & 228 (27\%) & 611 (73\%) & \\
+\textbf{Month} &  &  & 0.4\\
+\hspace{1em}October & 469 (27\%) & 1,263 (73\%) & \\
+\hspace{1em}November & 376 (28\%) & 980 (72\%) & \\
+\hspace{1em}December & 113 (25\%) & 348 (75\%) & \\
+\textbf{Race} &  &  & <0.001\\
+\hspace{1em}White/Caucasian & 233 (18\%) & 1,079 (82\%) & \\
+\hspace{1em}Arab/Middle Eastern & 76 (36\%) & 138 (64\%) & \\
+\hspace{1em}Black & 114 (38\%) & 184 (62\%) & \\
+\hspace{1em}East Asian/Pacific Islander & 69 (23\%) & 234 (77\%) & \\
+\hspace{1em}Indigenous & 76 (40\%) & 115 (60\%) & \\
+\hspace{1em}Latin American & 69 (38\%) & 111 (62\%) & \\
+\hspace{1em}Mixed & 105 (34\%) & 205 (66\%) & \\
+\hspace{1em}Other & 128 (35\%) & 239 (65\%) & \\
+\hspace{1em}South Asian & 88 (24\%) & 286 (76\%) & \\
 \bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{2} Pearson's Chi-squared test}\\
 \end{tabular}
@@
 
-Table~\ref{tbl-model} shows the results of the logistic regression
-models (for the uncorrected and corrected data) on vaccination status
-using socio-economic factors (age group, income, race), months covered
-by the survey (October, November, and December), geographical areas
-(Health Regions) and the interactions between income and race and Health
-Region and race. The corrected results show that there were no
-statistically significant differences in vaccination rate between the
-age groups or across the months covered by the survey as in both cases
-the odd ratios were similar. However, between the different household
-income brackets, individuals with an income under CAD 25,000 or between
-CAD 25,000-59,999 had significantly lower odds of vaccination than those
-with an income above CAD 60,000 (ORs=0.37 and 0.59,
-\emph{p}\textless0.001 and \emph{p}=0.010, respectively). Within
-Race/Ethnicity, individuals who identified as Arab/Middle Eastern,
-Black, or Latin American, had significantly lower odds of vaccination
-than those in the White/Caucasian group (ORs=0.32, 0.32, 0.27, and
-\emph{p}=0.003, \emph{p}\textless0.001 and \emph{p}=0.004,
-respectively); additionally, those individuals that reported to belong
-to the ``Other'' Race/Ethnicity group (which included the Southeast
-Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere
-groups according to the Census) had even lower odds of vaccination than
-the other minority groups (OR=0.22, \emph{p}\textless0.001). Regarding
-Health Regions, individuals that reported living in the West Health
-Region (which comprises the regions of Waterloo and Niagara, the
-counties of Wellington, Essex, and Lambton, and the cities of Hamilton,
-Haldimand, Brant, and Chatham-Kent) had significantly higher odds of
-vaccination than those in the Health Region of Toronto (OR=1.54,
-\emph{p}=0.031).
+\textbf{?@tbl-model} shows the results of the logistic regression models
+(for the uncorrected and corrected data) on vaccination status using
+socio-economic factors (age group, income, race), months covered by the
+survey (October, November, and December), geographical areas (Health
+Regions) and the interactions between income and race and Health Region
+and race. The corrected results show that there were no statistically
+significant differences in vaccination rate between the age groups or
+across the months covered by the survey as in both cases the odd ratios
+were similar. However, between the different household income brackets,
+individuals with an income under CAD 25,000 or between CAD 25,000-59,999
+had significantly lower odds of vaccination than those with an income
+above CAD 60,000 (ORs=0.37 and 0.59, \emph{p}\textless0.001 and
+\emph{p}=0.010, respectively). Within Race/Ethnicity, individuals who
+identified as Arab/Middle Eastern, Black, or Latin American, had
+significantly lower odds of vaccination than those in the
+White/Caucasian group (ORs=0.32, 0.32, 0.27, and \emph{p}=0.003,
+\emph{p}\textless0.001 and \emph{p}=0.004, respectively); additionally,
+those individuals that reported to belong to the ``Other''
+Race/Ethnicity group (which included the Southeast Asian, Filipino, West
+Asian, and Minorities Not Identified Elsewhere groups according to the
+Census) had even lower odds of vaccination than the other minority
+groups (OR=0.22, \emph{p}\textless0.001). Regarding Health Regions,
+individuals that reported living in the West Health Region (which
+comprises the regions of Waterloo and Niagara, the counties of
+Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
+Brant, and Chatham-Kent) had significantly higher odds of vaccination
+than those in the Health Region of Toronto (OR=1.54, \emph{p}=0.031).
 
@@
 
-\footnotesize
-\renewcommand{\arraystretch}{0.5}
-
-\hypertarget{tbl-model}{}
-\begin{longtable}{lcccccc}
-\caption{\label{tbl-model}Multiple Regression Analysis-Predictors of Vaccination Status }\tabularnewline
-
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{\textbf{Uncorrected}} & \multicolumn{3}{c}{\textbf{Corrected}} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
-\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
-\midrule
-\endfirsthead
-\multicolumn{7}{@{}l}{\textit{(continued)}}\\
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{\textbf{Uncorrected}} & \multicolumn{3}{c}{\textbf{Corrected}} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
-\textbf{Characteristic} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value} & \textbf{OR} & \textbf{95\% CI} & \textbf{p-value}\\
-\midrule
-\endhead
-
-\endfoot
-\bottomrule
-\endlastfoot
-\textbf{Age Group} &  &  &  &  &  & \\
-\hspace{1em}16-34 & — & — &  & — & — & \\
-\hspace{1em}35-54 & 0.93 & 0.77, 1.14 & 0.5 & 0.90 & 0.67, 1.21 & 0.5\\
-\hspace{1em}55 and over & 0.74 & 0.61, 0.89 & 0.002 & 0.99 & 0.74, 1.32 & >0.9\\
-\textbf{Month} &  &  &  &  &  & \\
-\hspace{1em}October & — & — &  & — & — & \\
-\hspace{1em}November & 0.95 & 0.81, 1.12 & 0.6 & 0.92 & 0.71, 1.18 & 0.5\\
-\hspace{1em}December & 1.16 & 0.91, 1.49 & 0.2 & 1.16 & 0.78, 1.71 & 0.5\\
-\textbf{Income (CAD)} &  &  &  &  &  & \\
-\hspace{1em}60000 and above & — & — &  & — & — & \\
-\hspace{1em}25000-59999 & 0.59 & 0.41, 0.84 & 0.003 & 0.59 & 0.39, 0.88 & 0.010\\
-\hspace{1em}under 25000 & 0.41 & 0.29, 0.58 & <0.001 & 0.37 & 0.25, 0.56 & <0.001\\
-\textbf{Race} &  &  &  &  &  & \\
-\hspace{1em}White/Caucasian & — & — &  & — & — & \\
-\hspace{1em}Arab/Middle Eastern & 0.24 & 0.12, 0.49 & <0.001 & 0.31 & 0.14, 0.68 & 0.003\\
-\hspace{1em}Black & 0.30 & 0.17, 0.54 & <0.001 & 0.32 & 0.17, 0.60 & <0.001\\
-\hspace{1em}East Asian/Pacific Islander & 0.73 & 0.36, 1.51 & 0.4 & 1.15 & 0.50, 2.66 & 0.7\\
-\hspace{1em}Indigenous & 0.38 & 0.17, 0.82 & 0.014 & 0.45 & 0.20, 1.04 & 0.061\\
-\hspace{1em}Latin American & 0.28 & 0.13, 0.59 & <0.001 & 0.27 & 0.11, 0.66 & 0.004\\
-\hspace{1em}Mixed & 0.58 & 0.30, 1.11 & 0.10 & 0.63 & 0.24, 1.62 & 0.3\\
-\hspace{1em}Other & 0.20 & 0.11, 0.35 & <0.001 & 0.22 & 0.12, 0.41 & <0.001\\
-\hspace{1em}South Asian & 0.79 & 0.44, 1.43 & 0.4 & 0.90 & 0.48, 1.67 & 0.7\\
-\textbf{Health Region} &  &  &  &  &  & \\
-\hspace{1em}Toronto & — & — &  & — & — & \\
-\hspace{1em}Central & 1.30 & 0.84, 2.00 & 0.2 & 1.47 & 0.92, 2.36 & 0.11\\
-\hspace{1em}East & 1.54 & 1.01, 2.34 & 0.044 & 1.41 & 0.90, 2.22 & 0.14\\
-\hspace{1em}West & 1.35 & 0.94, 1.93 & 0.10 & 1.54 & 1.04, 2.29 & 0.031\\
-\textbf{Income (CAD) * Race} &  &  &  &  &  & \\
-\hspace{1em}25000-59999 * Arab/Middle Eastern & 2.19 & 0.95, 5.05 & 0.067 & 1.85 & 0.68, 4.99 & 0.2\\
-\hspace{1em}under 25000 * Arab/Middle Eastern & 2.95 & 1.39, 6.24 & 0.005 & 3.08 & 1.27, 7.47 & 0.013\\
-\hspace{1em}25000-59999 * Black & 1.18 & 0.59, 2.36 & 0.6 & 1.31 & 0.58, 2.98 & 0.5\\
-\hspace{1em}under 25000 * Black & 2.84 & 1.46, 5.53 & 0.002 & 3.15 & 1.43, 6.92 & 0.004\\
-\hspace{1em}25000-59999 * East Asian/Pacific Islander & 0.97 & 0.44, 2.10 & >0.9 & 0.42 & 0.17, 1.06 & 0.066\\
-\hspace{1em}under 25000 * East Asian/Pacific Islander & 1.90 & 0.90, 4.01 & 0.094 & 1.15 & 0.47, 2.84 & 0.8\\
-\hspace{1em}25000-59999 * Indigenous & 1.77 & 0.72, 4.32 & 0.2 & 1.33 & 0.46, 3.80 & 0.6\\
-\hspace{1em}under 25000 * Indigenous & 1.60 & 0.71, 3.60 & 0.3 & 1.41 & 0.54, 3.72 & 0.5\\
-\hspace{1em}25000-59999 * Latin American & 0.89 & 0.38, 2.10 & 0.8 & 1.25 & 0.45, 3.45 & 0.7\\
-\hspace{1em}under 25000 * Latin American & 3.10 & 1.34, 7.21 & 0.009 & 2.81 & 1.04, 7.59 & 0.041\\
-\hspace{1em}25000-59999 * Mixed & 0.87 & 0.39, 1.93 & 0.7 & 0.87 & 0.33, 2.33 & 0.8\\
-\hspace{1em}under 25000 * Mixed & 1.27 & 0.65, 2.49 & 0.5 & 1.12 & 0.38, 3.33 & 0.8\\
-\hspace{1em}25000-59999 * Other & 5.50 & 2.43, 12.5 & <0.001 & 6.96 & 2.67, 18.2 & <0.001\\
-\hspace{1em}under 25000 * Other & 4.09 & 2.27, 7.37 & <0.001 & 4.63 & 2.34, 9.13 & <0.001\\
-\hspace{1em}25000-59999 * South Asian & 1.14 & 0.55, 2.38 & 0.7 & 1.22 & 0.52, 2.87 & 0.7\\
-\hspace{1em}under 25000 * South Asian & 1.58 & 0.82, 3.05 & 0.2 & 2.01 & 0.93, 4.34 & 0.075\\
-\textbf{Race * Health Region} &  &  &  &  &  & \\
-\hspace{1em}Arab/Middle Eastern * Central & 0.76 & 0.33, 1.71 & 0.5 & 0.66 & 0.25, 1.69 & 0.4\\
-\hspace{1em}Black * Central & 0.48 & 0.23, 1.02 & 0.055 & 0.44 & 0.19, 0.99 & 0.046\\
-\hspace{1em}East Asian/Pacific Islander * Central & 0.84 & 0.38, 1.89 & 0.7 & 0.98 & 0.38, 2.53 & >0.9\\
-\hspace{1em}Indigenous * Central & 0.60 & 0.24, 1.50 & 0.3 & 0.63 & 0.22, 1.78 & 0.4\\
-\hspace{1em}Latin American * Central & 0.69 & 0.28, 1.71 & 0.4 & 0.66 & 0.22, 1.93 & 0.4\\
-\hspace{1em}Mixed * Central & 0.63 & 0.30, 1.35 & 0.2 & 0.73 & 0.24, 2.23 & 0.6\\
-\hspace{1em}Other * Central & 0.98 & 0.49, 1.95 & >0.9 & 0.81 & 0.36, 1.81 & 0.6\\
-\hspace{1em}South Asian * Central & 0.71 & 0.34, 1.47 & 0.4 & 0.55 & 0.25, 1.21 & 0.14\\
-\hspace{1em}Arab/Middle Eastern * East & 0.57 & 0.20, 1.63 & 0.3 & 0.43 & 0.13, 1.43 & 0.2\\
-\hspace{1em}Black * East & 0.84 & 0.37, 1.89 & 0.7 & 0.86 & 0.35, 2.11 & 0.7\\
-\hspace{1em}East Asian/Pacific Islander * East & 0.82 & 0.30, 2.24 & 0.7 & 0.86 & 0.29, 2.54 & 0.8\\
-\hspace{1em}Indigenous * East & 0.56 & 0.21, 1.45 & 0.2 & 0.70 & 0.23, 2.11 & 0.5\\
-\hspace{1em}Latin American * East & 0.80 & 0.28, 2.25 & 0.7 & 1.03 & 0.31, 3.40 & >0.9\\
-\hspace{1em}Mixed * East & 0.70 & 0.31, 1.60 & 0.4 & 0.92 & 0.28, 3.05 & 0.9\\
-\hspace{1em}Other * East & 0.87 & 0.38, 2.02 & 0.7 & 1.08 & 0.40, 2.91 & 0.9\\
-\hspace{1em}South Asian * East & 0.50 & 0.20, 1.24 & 0.14 & 0.51 & 0.19, 1.42 & 0.2\\
-\hspace{1em}Arab/Middle Eastern * West & 1.18 & 0.49, 2.82 & 0.7 & 1.02 & 0.37, 2.77 & >0.9\\
-\hspace{1em}Black * West & 0.78 & 0.36, 1.67 & 0.5 & 0.77 & 0.32, 1.82 & 0.5\\
-\hspace{1em}East Asian/Pacific Islander * West & 0.54 & 0.24, 1.21 & 0.14 & 0.52 & 0.20, 1.36 & 0.2\\
-\hspace{1em}Indigenous * West & 0.45 & 0.19, 1.05 & 0.064 & 0.40 & 0.14, 1.12 & 0.082\\
-\hspace{1em}Latin American * West & 0.99 & 0.39, 2.48 & >0.9 & 0.96 & 0.33, 2.79 & >0.9\\
-\hspace{1em}Mixed * West & 0.53 & 0.25, 1.10 & 0.088 & 0.37 & 0.12, 1.15 & 0.085\\
-\hspace{1em}Other * West & 0.55 & 0.27, 1.12 & 0.10 & 0.41 & 0.18, 0.92 & 0.032\\
-\hspace{1em}South Asian * West & 0.50 & 0.24, 1.08 & 0.078 & 0.41 & 0.18, 0.95 & 0.037\\*
-\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{1} OR = Odds Ratio, CI = Confidence Interval}\\
-\end{longtable}
-
-\normalsize
+\begin{figure}
+
+\includegraphics{main_files/figure-pdf/fig-model-uncorr-1.pdf} \hfill{}
+
+\caption{\label{fig-model-uncorr}Coefficient estimates and confidence
+intervals for the uncorrected model. For the interaction terms, only
+those with a statistically significant values are shown.}
+
+\end{figure}
+
+\begin{figure}
+
+\includegraphics{main_files/figure-pdf/fig-model-corr-1.pdf} \hfill{}
+
+\caption{\label{fig-model-corr}Coefficient estimates and confidence
+intervals for the corrected model. For the interaction terms, only those
+with a statistically significant values are shown.}
+
+\end{figure}
 
@@
 (Table~\ref{tbl-descriptive-stats}) and vaccination odds
-(Table~\ref{tbl-model}) across the period covered by the survey, which
+(\textbf{?@tbl-model}) across the period covered by the survey, which
 follow the vaccination rates from Public Health Ontario and which
@@
 of individuals that identified as White/Caucasian
-(Table~\ref{tbl-model}). These results are consistent with other studies
+(\textbf{?@tbl-model}). These results are consistent with other studies
 that have shown lower vaccination rates in individuals that identify as
@@
 significantly lower in various racial minority groups when compared to
-White/Caucasian individuals (Table~\ref{tbl-model}), the change in odds
+White/Caucasian individuals (\textbf{?@tbl-model}), the change in odds
 of vaccination within certain racial groups and income strata was
@@
 25,000, which was also true for individuals in other racial minority
-groups with an income between CAD 25,000-59,999 (Table~\ref{tbl-model},
+groups with an income between CAD 25,000-59,999 (\textbf{?@tbl-model},
 Supplementary Figure A-3).
@@
 West Health Region when compared to the Health Region of Toronto
-(Table~\ref{tbl-model}). The West Health Region comprises the regions of
+(\textbf{?@tbl-model}). The West Health Region comprises the regions of
 Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and
@@
 identifying as South Asian or other minorities not included in the
-survey Table~\ref{tbl-model}. In this case, the results of the
+survey \textbf{?@tbl-model}. In this case, the results of the
 interaction term in the model indicate that the odds of vaccination for

### 2023-04-26 manuscript_v4/01-background.qmd
updating v4

@@
 
-Therefore, in this study we hypothesized that there were differences in vaccination uptake between the Health Regions of Ontario during the last quarter of 2021.To better contextualize potential differences, we also included socio-economic factors in our analysis in order to identify in which demographic groups these differences were significant.
+Therefore, in this study we hypothesized that there were differences in vaccination uptake between the different Health Regions of Ontario during the last quarter of 2021.
+By including socio-economic factors in our analysis, we aimed at identifying in which groups these differences were significant in order to provide an assessment of the current state of healthcare access in Ontario.

### 2023-04-26 manuscript_v4/02-methods.qmd
updating v4

@@
 
-## Data {#sec-data}
+## Data and Methods {#sec-data}
 
-We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences (henceforth Fields) and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto, and which ran between September 30th, 2021 and January 17th,2022.
-The survey collected socio-economic information from participants (@tbl-covariates), recorded their location (using the nearest municipality), the date of access to the survey, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".
-The original dataset contained 39,029 entries (where each entry corresponded to a unique respondent).
+We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto, and which ran between September 30th, 2021 and January 17th,2022.
+The survey collected socio-economic information from participants (@tbl-descriptive-stats), their location (nearest municipality,  as shown in @fig-map), the date of access to the survey, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".
+The original dataset contained 39,029 observations (where each observation corresponded to a unique respondent).
 
-This dataset was cleaned to remove outliers that were identified during preliminary analyses, and the geographical information in the survey (city where the survey was responded) was used to match each city to its correspondent Health Region.
-
-The clean dataset contained responses from more than 200 different municipalities within Ontario (@fig-map).
-Because of the lack of a publicly available list of all municipalities within each Health Region, we used a dataset of long-term care homes and LHINs to match each city to LHIN, followed by matching each LHIN to a Health Region following the information provided on the Ontario Health Website, where the list of LHINs and corresponding Health Regions is available.
-In the case of municipalities that did not appear in the long-term care home dataset, we manually searched each city in the LHINs websites in order to provide geographical information.
-The original dataset, clean dataset, and details on the data cleaning process are described in detail in the GitHub repository for this paper, which can be found at [https://github.com/aimundo/Fields_COVID-19/](https://github.com/aimundo/Fields_COVID-19/).
-
-Following an assessment of the number of entries corresponding to each Health Region in the final dataset, only 107 observations (4.3% of the total) corresponded to cities located in the North West and North East Health Regions.
-The low representation of these Health Regions in the dataset is noticeable in @fig-map, which shows that responses from these areas came from a relatively low number of cities when compared to the most populated Health Regions, such as the Toronto or Central Regions.
-We omitted the North East and North East Health Regions from further analyses due to the low number of entries.
-Therefore, the total number of unique entries used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021.
+Preliminary analyses of the data included the removal of outliers, of observations where respondents did not provide answers in all the covariates of interest, matching the city of each observations with its corresponding LHIN and Health Region, and removing observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions).
+After all the preliminary analyses indicated above, the total number of observations used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021.
+The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.
 
@@
 
-We used a logistic regression model to estimate the probability of vaccination depending on the socio-economic factors described in @tbl-covariates, the month when the survey was answered, the Health Regions from Ontario indicated in @sec-data, and the interactions between Race and Health Region, and Race and income, as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
-
-The model was fitted first to the clean dataset to obtain uncorrected estimates.
-Because we identified differences between the proportions of all the socio-economic factors included in the analysis (@tbl-covariates) and the Census data for Ontario, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using Census socio-economic data and Health Region population totals, in order to obtain corrected estimates from the model.
+We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
+Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals;and fitted the regression model to the uncorrected and corrected data.
 Details regarding the correction can be found in the Appendix.
-All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], and `quarto` [@quarto].
+All analyses were conducted in R 4.2.2 using the packages `survey` [@lumley2011],`tidyverse` [@wickham2019], `quarto` [@quarto], `modelsummary` [@modelsummary], and `gtsummary` [@gtsummary].

### 2023-04-26 manuscript_v4/main.tex
updating v4

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, enhanced, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 Therefore, in this study we hypothesized that there were differences in
-vaccination uptake between the Health Regions of Ontario during the last
-quarter of 2021.To better contextualize potential differences, we also
-included socio-economic factors in our analysis in order to identify in
-which demographic groups these differences were significant.
+vaccination uptake between the different Health Regions of Ontario
+during the last quarter of 2021. By including socio-economic factors in
+our analysis, we aimed at identifying in which groups these differences
+were significant in order to provide an assessment of the current state
+of healthcare access in Ontario.
 
@@
 \hypertarget{sec-data}{%
-\subsection{Data}\label{sec-data}}
+\subsection{Data and Methods}\label{sec-data}}
 
@@
 province of Ontario that was commissioned by the Fields Institute for
-Research in Mathematical Sciences (henceforth Fields) and the
-Mathematical Modelling of COVID-19 Task Force under ethical guidance
-from the University of Toronto, and which ran between September 30th,
-2021 and January 17th,2022. The survey collected socio-economic
-information from participants (\textbf{?@tbl-covariates}), recorded
-their location (using the nearest municipality), the date of access to
-the survey, and asked information on vaccination status by using the
-question ``Have you received the first dose of the COVID vaccine?'',
-with possible answers ``yes'' and ``no''. The original dataset contained
-39,029 entries (where each entry corresponded to a unique respondent).
-
-This dataset was cleaned to remove outliers that were identified during
-preliminary analyses, and the geographical information in the survey
-(city where the survey was responded) was used to match each city to its
-correspondent Health Region.
-
-The clean dataset contained responses from more than 200 different
-municipalities within Ontario (Figure~\ref{fig-map}). Because of the
-lack of a publicly available list of all municipalities within each
-Health Region, we used a dataset of long-term care homes and LHINs to
-match each city to LHIN, followed by matching each LHIN to a Health
-Region following the information provided on the Ontario Health Website,
-where the list of LHINs and corresponding Health Regions is available.
-In the case of municipalities that did not appear in the long-term care
-home dataset, we manually searched each city in the LHINs websites in
-order to provide geographical information. The original dataset, clean
-dataset, and details on the data cleaning process are described in
-detail in the GitHub repository for this paper, which can be found at
-\url{https://github.com/aimundo/Fields_COVID-19/}.
-
-Following an assessment of the number of entries corresponding to each
-Health Region in the final dataset, only 107 observations (4.3\% of the
-total) corresponded to cities located in the North West and North East
-Health Regions. The low representation of these Health Regions in the
-dataset is noticeable in Figure~\ref{fig-map}, which shows that
-responses from these areas came from a relatively low number of cities
-when compared to the most populated Health Regions, such as the Toronto
-or Central Regions. We omitted the North East and North East Health
-Regions from further analyses due to the low number of entries.
-Therefore, the total number of unique entries used for analysis was
-3,549 which included the East, Central, Toronto, and West Health Regions
-covering the period between October 1st,2021 and December 12th, 2021.
+Research in Mathematical Sciences and the Mathematical Modelling of
+COVID-19 Task Force under ethical guidance from the University of
+Toronto, and which ran between September 30th, 2021 and January
+17th,2022. The survey collected socio-economic information from
+participants (Table~\ref{tbl-descriptive-stats}), their location
+(nearest municipality, as shown in Figure~\ref{fig-map}), the date of
+access to the survey, and asked information on vaccination status by
+using the question ``Have you received the first dose of the COVID
+vaccine?'', with possible answers ``yes'' and ``no''. The original
+dataset contained 39,029 observations (where each observation
+corresponded to a unique respondent).
+
+Preliminary analyses of the data included the removal of outliers, of
+observations where respondents did not provide answers in all the
+covariates of interest, matching the city of each observations with its
+corresponding LHIN and Health Region, and removing observations from
+areas with low representation (107 observations corresponding to the
+North West and North East Health Regions). After all the preliminary
+analyses indicated above, the total number of observations used for
+analysis was 3,549 which included the East, Central, Toronto, and West
+Health Regions covering the period between October 1st,2021 and December
+12th, 2021. The original dataset, clean dataset, and details on the data
+cleaning process are described in detail in the
+\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
+for this paper.
 
@@
 
-We used a logistic regression model to estimate the probability of
-vaccination depending on the socio-economic factors described in
-\textbf{?@tbl-covariates}, the month when the survey was answered, the
-Health Regions from Ontario indicated in Section~\ref{sec-data}, and the
-interactions between Race and Health Region, and Race and income, as
-previous studies have shown that socio-economic factors and their
-interactions are significant predictors of intent of vaccination and
-vaccination
+We used a logistic regression model to examine the impact of the Health
+Regions in vaccination rates while considering the socio-economic
+factors and and months covered by the survey
+(Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
+Health Region and Race and income), as previous studies have shown that
+socio-economic factors and their interactions are significant predictors
+of intent of vaccination and vaccination
 status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
-
-The model was fitted first to the clean dataset to obtain uncorrected
-estimates. Because we identified differences between the proportions of
-all the socio-economic factors included in the analysis
-(\textbf{?@tbl-covariates}) and the Census data for Ontario, we used an
-iterative proportional fitting procedure
+Because we identified differences in representativity between the survey
+data and the estimates from the Census, we used an iterative
+proportional fitting procedure
 (\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{35}}
-to correct the data using Census socio-economic data and Health Region
-population totals, in order to obtain corrected estimates from the
-model. Details regarding the correction can be found in the Appendix.
-All analyses were conducted in R 4.2.2 using the packages
+to correct the data using data from the Census and Health Region
+population totals;and fitted the regression model to the uncorrected and
+corrected data. Details regarding the correction can be found in the
+Appendix. All analyses were conducted in R 4.2.2 using the packages
 \texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
-and \texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}}.
+\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}},
+\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{39}},
+and
+\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{40}}.
 
@@
 December 12th,
-2021)\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. In
+2021)\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}. In
 this case, differences between the survey and province-wide estimates
@@
 have been relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{40}}, and
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{42}}, and
 with vaccination uptake rates across different age groups presented in
 other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{41}}.
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{43}}.
 In other words, although vaccination rates obtained from the survey were
@@
 months (October, November, and December of
-2021\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}).
+2021\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}).
 Moreover, there was good agreement between vaccination rates within each
@@
 rates above
-80\%\textsuperscript{\protect\hyperlink{ref-ontario-covid}{39}}. To this
+80\%\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}. To this
 day, differences in vaccination rates within the province continue, as
@@
 still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{42}}.
+rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{44}}.
 
@@
 part of a racial minority, or that have a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{43}}.
+income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{45}}.
 
@@
 as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{44},\protect\hyperlink{ref-ct2021}{45}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{46},\protect\hyperlink{ref-ct2021}{47}},
 which include occupations such as grocery store workers, gas station
@@
 had priority for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{46}},
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{48}},
 which would explain the higher odds of vaccination for these individuals
@@
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{47}}.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{49}}.
 In the case of this analysis, the estimated lower odds are likely to be
@@
 population (more than 20\%) in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{48}}
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{50}}
 (Supplementary Table A-8). Therefore, this result partly reinforces the
@@
 previously reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{49}}; in this
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{51}}; in this
 regard, our result provides additional context by showing that within
@@
 are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{47}}, a
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{49}}, a
 minority group that has historically suffered from reduced access to
@@
 self-reported and documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{50}}, and
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{52}}, and
 therefore, the effect of self-reported bias is likely to not be
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
+in {R}. \emph{Journal of Statistical Software}. 2022;103(1):1-23.
+doi:\href{https://doi.org/10.18637/jss.v103.i01}{10.18637/jss.v103.i01}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
+\CSLLeftMargin{40. }%
+\CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
+Reproducible summary tables with the gtsummary package. \emph{{The R
+Journal}}. 2021;13:570-580.
+doi:\href{https://doi.org/10.32614/RJ-2021-053}{10.32614/RJ-2021-053}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-04-28 manuscript_v4/02-methods.qmd
need to fix figure to exponentiated

@@
 
+Comparisons between the variables of the survey were performed using the $\chi^2$ test.
 We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
-Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals;and fitted the regression model to the uncorrected and corrected data.
+Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data.
+The reference groups in the model were: 16-34 years (age group), October(Month), CAD 60,000 and above (income), White/Caucasian (Race), Toronto (Health Region).
 Details regarding the correction can be found in the Appendix.

### 2023-04-28 manuscript_v4/03-results.qmd
need to fix figure to exponentiated

@@
 
-## Survey Results
+## Sample Characteristics
 
-@tbl-descriptive-stats shows the descriptive statistics from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed.
-The total number of entries analyzed was 3,549.
-Overall, 27\% (958) of survey respondents  reported not having received the first dose of the vaccine, whereas 73\% (2,591) reported having received it.
-Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\%  (1307) of the number of entries, those within the CAD 25,000-59,999 income bracket represented 25% (889) of the total sample, and those with an income above CAD 60,000 represented 38 \% (1353) of the sample.
-Between the three income brackets, the lower vaccination rate corresponded to the under CAD 25,000 group, with 69\% .
-
-Within the age groups of survey respondents, the age group between 16-34 years had the highest representation in the survey responses (1,520 or 42.8\% of all responses).
-Within this age bracket, 73\% of respondents indicated having received the vaccine, whereas the lowest vaccination rate was in the bracket of those 55 years of age and above, with 72\%.
-The Health Region with highest representation in the survey was Toronto, accounting for 1,323 entries (37.3\% of the total), with a vaccination rate of 72\%.
-Across the three months covered by the survey, the highest number of entries was in October of 2021, with a total of 1,732 respondents, which corresponded to 49\% of all the survey entries.
-Additionally, the proportion of respondents who reported being vaccinated was similar in all months, ranging between 72\% and 75\%, with the highest value corresponding to December of 2021.
-
-Regarding race/ethnicity, individuals that identified as White/Caucasian represented 37\% of all entries (1,312) and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine.
-On the other hand, the ethnic group with the lowest number of entries in the survey was Latin American, with a total of 180, or 5\% of all entries.
-Vaccination rates across all minority groups were below the value reported by White/Caucasians, with the lowest vaccination rate (60\%) being reported by individuals that identified as Indigenous.
+@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status.
+A total of 3,549 observations were used for analysis, from which 27\% (958) corresponded to individuals that reported not having received the first dose of the vaccine.
+Respondents between 16 and 34 years of age, and those that lived within the Health Region of Toronto had the highest number of observations in the data (42.8\% and 37.3\% of the total, respectively).
+Moreover, individuals who identified as White or Caucasian reported the highest vaccination uptake (82\%), in contrast to respondents who identified as which had had the lowest vaccination uptake rate (60\%) across all groups.
+From all the covariates, significant associations were identified between income, race, and vaccination rate.
 
@@
                  income_ord="Income (CAD)")
-  ) %>% add_p()%>%
+  ) %>% add_p(test = everything () ~ "chisq.test")%>%
   bold_labels()
@@
 
+# test<-apply(xtabs(~first_dose_m + Race,clean_data), 2, chisq.test)
+# 
+# 
+# clean_data %>%
+#   filter(income_ord %in% c("60000 and above"))%>%
+#   tbl_summary(percent="row",
+#               by= first_dose_m,
+#               missing="no",
+#               include=Health_Region,
+#               label=list(
+#                  income_ord="Income (CAD)")
+#   ) %>% add_p()%>%
+#   bold_labels()
+
+# clean_data %>%
+#   filter(income_ord %in% c("25000-59999"))%>%
+#   tbl_summary(percent="row",
+#               by= income_ord,
+#               missing="no",
+#               include=(income_ord),
+#               label=list(
+#                  income_ord="Income (CAD)")
+#   ) %>% add_p()%>%
+#   bold_labels()
+
 ```
@@
 
-@tbl-model shows the results of the logistic regression models (for the uncorrected and corrected data) on vaccination status using socio-economic factors (age group, income, race), months covered by the survey (October, November, and December), geographical areas (Health Regions) and the interactions between income and race and Health Region and race.
-The corrected results show that there were no statistically significant differences in vaccination rate between the age groups or across the months covered by the survey as in both cases the odd ratios were similar.
-However, between the different household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, _p_<0.001 and _p_=0.010, respectively).
-Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.32, 0.32, 0.27, and _p_=0.003, _p_<0.001 and _p_=0.004, respectively); additionally, those individuals that reported to belong to the "Other" Race/Ethnicity group (which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census) had even lower odds of vaccination than the other minority groups (OR=0.22, _p_<0.001).
-Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, _p_=0.031).
+@fig-model-uncorr shows the estimates from the logistic regression model of vaccination status for the uncorrected data, whereas the estimated obtained from the corrected data appear in @fig-model-corr.
+The results show significantly higher odds of vaccination in individuals with a high household income, and those that identified as White or Caucasian when compared with individuals in lower income brackets or that identified as part of equity-deserving groups.
 
-Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.08, _p_=0.013), Black (OR=3.15, _p_=0.004), Latin American (OR=2.81, _p_=0.041), or that belonged to other minority groups  (OR=4.63, _p_<0.001).
+Specifically, significantly lower odds of vaccination were identified for those with a household income under CAD 25,000 (OR=0.374, CI=[0.251,0.559]) and those with an income between CAD 25,000 and 59,999 (OR=0.586, CI=[0.390,0.882]).
+Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, or that belonged to "Other" racial groups, which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and CIs=[0.14,0.68], [0.17,0.60], [0.11,0.66], and [0.12,0.41], respectively).
+Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, CI=[1.04,2.29]).
+
+Interestingly, higher odds of vaccination were also estimated in the case of individuals in equity-deserving groups with a household income below CAD 25,000 when compared to those with a household income above CAD 60,000.
+This held true in the case of Arab/Middle Eastern (OR=3.08, _p_=0.013), Black (OR=3.15, _p_=0.004), and Latin American (OR=2.81, _p_=0.041) individuals, as well as respondents who belonged to "Other" minority groups  (OR=4.63, _p_<0.001).
 Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.96, _p_<0.001).
@@
 #| label: fig-model-uncorr
-#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. For the interaction terms, only those with a statistically significant values are shown."
+#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."
 
@@
 
-mplot1<-modelplot(m0,draw=FALSE,coef_omit = "Intercept") %>%
+mplot1<-modelplot(m0,draw=FALSE,
+                  coef_omit = "Intercept", exponentiate = TRUE) %>%
   mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
@@
 #| label: fig-model-corr
-#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. For the interaction terms, only those with a statistically significant values are shown."
+#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."
 
@@
 
-mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept") %>%
+mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept",exponentiate = TRUE) %>%
   mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
@@
 ggplot(aes(y=term))+
-  geom_vline(xintercept = 0, linetype="dashed",color="red") +
+  geom_vline(xintercept = 5.73, linetype="dashed",color="red") +
   geom_point(aes(x=estimate))+

### 2023-04-28 manuscript_v4/main.tex
need to fix figure to exponentiated

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-We used a logistic regression model to examine the impact of the Health
-Regions in vaccination rates while considering the socio-economic
-factors and and months covered by the survey
+Comparisons between the variables of the survey were performed using the
+\(\chi^2\) test. We used a logistic regression model to examine the
+impact of the Health Regions in vaccination rates while considering the
+socio-economic factors and and months covered by the survey
 (Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
@@
 to correct the data using data from the Census and Health Region
-population totals;and fitted the regression model to the uncorrected and
-corrected data. Details regarding the correction can be found in the
-Appendix. All analyses were conducted in R 4.2.2 using the packages
+population totals; and fitted the regression model to the uncorrected
+and corrected data. The reference groups in the model were: 16-34 years
+(age group), October(Month), CAD 60,000 and above (income),
+White/Caucasian (Race), Toronto (Health Region). Details regarding the
+correction can be found in the Appendix. All analyses were conducted in
+R 4.2.2 using the packages
 \texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
@@
 
-\hypertarget{survey-results}{%
-\subsection{Survey Results}\label{survey-results}}
-
-Table~\ref{tbl-descriptive-stats} shows the descriptive statistics from
-the Fields COVID-19 survey data for vaccination status and each of the
-covariates analyzed. The total number of entries analyzed was 3,549.
-Overall, 27\% (958) of survey respondents reported not having received
-the first dose of the vaccine, whereas 73\% (2,591) reported having
-received it. Within each socio-economic factor, respondents who reported
-living in a household with an income under CAD 25,000 represented 37\%
-(1307) of the number of entries, those within the CAD 25,000-59,999
-income bracket represented 25\% (889) of the total sample, and those
-with an income above CAD 60,000 represented 38 \% (1353) of the sample.
-Between the three income brackets, the lower vaccination rate
-corresponded to the under CAD 25,000 group, with 69\% .
-
-Within the age groups of survey respondents, the age group between 16-34
-years had the highest representation in the survey responses (1,520 or
-42.8\% of all responses). Within this age bracket, 73\% of respondents
-indicated having received the vaccine, whereas the lowest vaccination
-rate was in the bracket of those 55 years of age and above, with 72\%.
-The Health Region with highest representation in the survey was Toronto,
-accounting for 1,323 entries (37.3\% of the total), with a vaccination
-rate of 72\%. Across the three months covered by the survey, the highest
-number of entries was in October of 2021, with a total of 1,732
-respondents, which corresponded to 49\% of all the survey entries.
-Additionally, the proportion of respondents who reported being
-vaccinated was similar in all months, ranging between 72\% and 75\%,
-with the highest value corresponding to December of 2021.
-
-Regarding race/ethnicity, individuals that identified as White/Caucasian
-represented 37\% of all entries (1,312) and had the highest vaccination
-uptake with 82\% of them indicating to have received the COVID-19
-vaccine. On the other hand, the ethnic group with the lowest number of
-entries in the survey was Latin American, with a total of 180, or 5\% of
-all entries. Vaccination rates across all minority groups were below the
-value reported by White/Caucasians, with the lowest vaccination rate
-(60\%) being reported by individuals that identified as Indigenous.
+\hypertarget{sample-characteristics}{%
+\subsection{Sample Characteristics}\label{sample-characteristics}}
+
+Table~\ref{tbl-descriptive-stats} shows the characteristics of the data
+from the Fields COVID-19 survey used for analysis with regard to
+vaccination status. A total of 3,549 observations were used for
+analysis, from which 27\% (958) corresponded to individuals that
+reported not having received the first dose of the vaccine. Respondents
+between 16 and 34 years of age, and those that lived within the Health
+Region of Toronto had the highest number of observations in the data
+(42.8\% and 37.3\% of the total, respectively). Moreover, individuals
+who identified as White or Caucasian reported the highest vaccination
+uptake (82\%), in contrast to respondents who identified as which had
+had the lowest vaccination uptake rate (60\%) across all groups. From
+all the covariates, significant associations were identified between
+income, race, and vaccination rate.
 
@@
 
-\textbf{?@tbl-model} shows the results of the logistic regression models
-(for the uncorrected and corrected data) on vaccination status using
-socio-economic factors (age group, income, race), months covered by the
-survey (October, November, and December), geographical areas (Health
-Regions) and the interactions between income and race and Health Region
-and race. The corrected results show that there were no statistically
-significant differences in vaccination rate between the age groups or
-across the months covered by the survey as in both cases the odd ratios
-were similar. However, between the different household income brackets,
-individuals with an income under CAD 25,000 or between CAD 25,000-59,999
-had significantly lower odds of vaccination than those with an income
-above CAD 60,000 (ORs=0.37 and 0.59, \emph{p}\textless0.001 and
-\emph{p}=0.010, respectively). Within Race/Ethnicity, individuals who
+Figure~\ref{fig-model-uncorr} shows the estimates from the logistic
+regression model of vaccination status for the uncorrected data, whereas
+the estimated obtained from the corrected data appear in
+Figure~\ref{fig-model-corr}. The results show significantly higher odds
+of vaccination in individuals with a high household income, and those
+that identified as White or Caucasian when compared with individuals in
+lower income brackets or that identified as part of equity-deserving
+groups.
+
+Specifically, significantly lower odds of vaccination were identified
+for those with a household income under CAD 25,000 (OR=0.374,
+CI={[}0.251,0.559{]}) and those with an income between CAD 25,000 and
+59,999 (OR=0.586, CI={[}0.390,0.882{]}). Additionally, individuals who
 identified as Arab/Middle Eastern, Black, or Latin American, had
@@
 \caption{\label{fig-model-uncorr}Coefficient estimates and confidence
-intervals for the uncorrected model. For the interaction terms, only
-those with a statistically significant values are shown.}
+intervals for the uncorrected model. Only statistically significant
+interaction terms are shown. Full estimates can be found in the
+Appendix.}
 
@@
 \caption{\label{fig-model-corr}Coefficient estimates and confidence
-intervals for the corrected model. For the interaction terms, only those
-with a statistically significant values are shown.}
+intervals for the corrected model. Only statistically significant
+interaction terms are shown. Full estimates can be found in the
+Appendix.}
 

### 2023-04-28 manuscript_v4/03-results.qmd
Updated files

@@
   ) %>% add_p(test = everything () ~ "chisq.test")%>%
-  bold_labels()
+  bold_labels() %>%
+  modify_table_styling(
+    columns=label,
+    rows=label=="Other",
+    footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
+  )
+
+# data_summary %>%
+#   modify_header(label = "**Variable**") %>%
+#   as_kable_extra(booktabs = TRUE,linesep="") %>%
+#   # reduce font size to make table fit. 
+#   # you may also use the `latex_options = "scale_down"` argument here.
+#   kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)
 
@@
   modify_header(label = "**Variable**") %>%
-  as_kable_extra(booktabs = TRUE,linesep="") %>%
-  # reduce font size to make table fit. 
-  # you may also use the `latex_options = "scale_down"` argument here.
-  kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)
+  as_gt()
 
@@
 @fig-model-uncorr shows the estimates from the logistic regression model of vaccination status for the uncorrected data, whereas the estimated obtained from the corrected data appear in @fig-model-corr.
-The results show significantly higher odds of vaccination in individuals with a high household income, and those that identified as White or Caucasian when compared with individuals in lower income brackets or that identified as part of equity-deserving groups.
-
-Specifically, significantly lower odds of vaccination were identified for those with a household income under CAD 25,000 (OR=0.374, CI=[0.251,0.559]) and those with an income between CAD 25,000 and 59,999 (OR=0.586, CI=[0.390,0.882]).
-Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, or that belonged to "Other" racial groups, which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and CIs=[0.14,0.68], [0.17,0.60], [0.11,0.66], and [0.12,0.41], respectively).
-Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, CI=[1.04,2.29]).
-
-Interestingly, higher odds of vaccination were also estimated in the case of individuals in equity-deserving groups with a household income below CAD 25,000 when compared to those with a household income above CAD 60,000.
-This held true in the case of Arab/Middle Eastern (OR=3.08, _p_=0.013), Black (OR=3.15, _p_=0.004), and Latin American (OR=2.81, _p_=0.041) individuals, as well as respondents who belonged to "Other" minority groups  (OR=4.63, _p_<0.001).
-Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.96, _p_<0.001).
-
-For the interaction of Health Region and race, significant odds of vaccination were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44, _p_=0.046), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41, _p_=0.032 and _p_=0.037, respectively).
+The results show significantly lower odds of vaccination in individuals with a low household income and those that identified as part of equity-deserving groups when compared to individuals in high income brackets or that identified as White or Caucasian.
 
@@
 #| label: fig-model-uncorr
-#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."
+#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model.
+Only statistically significant interaction terms are shown.
+Full interaction terms can be found in Supplementary Figure A-3."
 
@@
 ggplot(aes(y=term))+
-  geom_vline(xintercept = 0, linetype="dashed",color="red") +
-  geom_point(aes(x=estimate))+
+  geom_vline(xintercept = 1, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate),size=2)+
   geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
   scale_y_discrete(limits=rev)+
-  xlab("Coefficient estimates and 95% confidence intervals")+
+  xlab("")+
   ylab("")+
@@
 ```
+
+Specifically, significantly lower odds of vaccination were identified for those with a household income under CAD 25,000 (OR=0.37, CI=[0.25,0.56]) and those with an income between CAD 25,000 and 59,999 (OR=0.59, CI=[0.39,0.88]).
+Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, or that belonged to "Other" racial groups, which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and CIs=[0.14,0.68], [0.17,0.60], [0.11,0.66], and [0.12,0.41], respectively).
+Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, CI=[1.04,2.29]).
+
+Interestingly, individuals in equity-deserving groups with a household income below CAD 25,000 had higher odds of vaccination (when compared to those with a household income above CAD 60,000).
+This held true in the case of Arab/Middle Eastern (OR=3.08, CI=[1.27,7.47]), Black (OR=3.15, CI=[1.43,6.92]), and Latin American (OR=2.81, CI=[1.04,7.59]) individuals, as well as respondents who belonged to "Other" minority groups  (OR=4.63, CI=[2.34,9.13]), who also had higher odds of vaccination in the CAD 25,000-59,999 income bracket (OR=6.96, CI=[2.67,18.16]).
+Finally, significantly lower odds of vaccination were identified (when compared to the Toronto Health Region) for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44, CI=[0.19,0.99]), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41, and CIs=[0.18,0.92] and [0.18,0.95], respectively).
 
@@
 #| label: fig-model-corr
-#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model. Only statistically significant interaction terms are shown. Full estimates can be found in the Appendix."
+#| fig-cap: "Coefficient estimates and confidence intervals for the corrected model.
+Only statistically significant interaction terms are shown.
+Full interaction terms can be found in Supplementary Figure A-4."
 
@@
 ggplot(aes(y=term))+
-  geom_vline(xintercept = 5.73, linetype="dashed",color="red") +
-  geom_point(aes(x=estimate))+
+  geom_vline(xintercept = 1, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate),size=2)+
   geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
   scale_y_discrete(limits=rev)+
-  xlab("Coefficient estimates and 95% confidence intervals")+
+  xlab("")+
   ylab("")+

### 2023-04-28 manuscript_v4/05-conclusion.qmd
Updated files

# Conclusion

This study explored differences in COVID-19 vaccination across the province of Ontario during the last quarter of 2021 taking into consideration socio-economic factors, such as income and race, their interactions, and the Health Regions within the province. Our results show that during the period analyzed, differences in vaccination uptake existed across multiple equity-deserving groups in the province, and that these differences were significant in two of the Health Regions analyzed. It is important that future public health policies in Ontario take into consideration how to adequately reach individuals from equity-deserving groups that might live in areas of the province where access to healthcare might be difficult. Only in this way the goal of the Health Region model, which aims at reducing disparities, will become successful.

### 2023-04-28 manuscript_v4/appendix.qmd
Updated files

--- 2023-04-17 manuscript_v3/appendix.qmd
@@
   pdf:
-    #mainfont: "Palatino Linotype"
-    include-before-body: latex/appendix_preamble.sty
+    include-in-header:
+      - text: |
+         \usepackage{placeins}
+    include-before-body: 
+      - file: latex/appendix_preamble.sty
 linestretch: 2
@@
 library(gtsummary)
+library(modelsummary)
 library(survey)
@@
 
-# Vaccination rates by Age
+# Complete Model Estimates
 
-The table below provides the percentages of vaccination from the survey data.
+The figures below present the complete estimates from the regression models.
 
-```{r,vacc-table}
-#| label: tbl-vaccination
-#| tbl-cap: "Vaccination Status by Age from the Survey Data"
+```{r,mod-uncorr-appendix}
 #| message: false
+#| warning: false
 #| echo: false
+#| fig-align: left
+#| fig-width: 7
+#| fig-height: 9
+#| label: fig-model-uncorr-appendix
+#| fig-cap: "Full coefficient estimates and confidence intervals for the uncorrected model."
 
-clean_data$test<-as.factor(clean_data$age)
+#create dataframe from modelplot, remove names of factors, drop the intercept
 
-tvac<-clean_data %>%
-  tbl_summary(percent="row",
-              by= first_dose,
-              include=test,
-              label=test~"Age")%>%
-  bold_labels()%>%
-as_kable_extra(booktabs = TRUE,linesep="",longtable=T) %>%
-  kableExtra::kable_styling(latex_options = "repeat_header")
+mplot1<-modelplot(m0,draw=FALSE,
+                  coef_omit = "Intercept", exponentiate = TRUE) %>%
+  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
 
-tvac
+#add column by factor 
+mplot1$factor<-c(rep("Age",2),
+                 rep("Month",2),
+                 rep("Income",2),
+                 rep("Race/Ethnicity",8),
+                 rep("Health Region",3),
+                 rep("Income and\nRace/Ethnicity",16),
+                 rep("Race/Ethnicity\nand\nHealth Region",24)
+                 )
+
+#order factors for plot
+#mplot1$term_ord <- factor(mplot1$term,ordered=TRUE,levels=mplot1$term)
+
+mplot1$factor <- factor(mplot1$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
+
+#remove non-interaction terms
+
+mplot1<-mplot1%>%
+ subset(str_detect(term,':'))
+
+#remove non-significant interactions from dataframe
+#mplot1<-anti_join(mplot1,test)
+
+mplot1<-mplot1%>%
+mutate(term=str_replace_all(term,':'," and "))
+
+#make plot
+mplot1 %>% 
+ggplot(aes(y=term))+
+  geom_vline(xintercept = 1, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate),size=2)+
+  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
+  scale_y_discrete(limits=rev)+
+  xlab("")+
+  ylab("")+
+  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y",labeller = label_wrap_gen(multi_line = TRUE))+
+  theme_classic()
 ```
+
+\FloatBarrier
+
+```{r,model-corr-appendix}
+#| message: false
+#| warning: false
+#| echo: false
+#| fig-align: left
+#| fig-width: 7
+#| fig-height: 9
+#| label: fig-model-corr-appendix
+#| fig-cap: "Full coefficient estimates and confidence intervals for the corrected model."
+
+#create dataframe from modelplot, remove names of factors, drop the intercept
+
+mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept",exponentiate = TRUE) %>%
+  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
+
+#add column by factor 
+mplot2$factor<-c(rep("Age",2),
+                 rep("Month",2),
+                 rep("Income",2),
+                 rep("Race/Ethnicity",8),
+                 rep("Health Region",3),
+                 rep("Income and\nRace/Ethnicity",16),
+                 rep("Race/Ethnicity\nand\nHealth Region",24)
+                 )
+
+#order factors for plot
+#mplot2$term_ord <- factor(mplot2$term,ordered=TRUE,levels=mplot2$term)
+
+mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
+
+#remove non-interaction terms
+
+mplot2<-mplot2%>%
+ subset(str_detect(term,':'))
+
+mplot2<-mplot2%>%
+mutate(term=str_replace_all(term,':'," and "))
+
+mplot2 %>% 
+ggplot(aes(y=term))+
+  geom_vline(xintercept = 1, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate),size=2)+
+  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
+  scale_y_discrete(limits=rev)+
+  xlab("")+
+  ylab("")+
+  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
+  theme_classic()
+```
+
+\FloatBarrier
 
@@
 
-The plots below show the interactions for Race and Income and Race and Health Region based on the regression model presented in the main paper.
+The plots below show the interactions for Race and Income and Race and Health Region based on the corrected regression model presented in the main paper.
 
@@
   ylab("Linear predictor")+
+  xlab("Income")+
   theme_classic()+
@@
   ylab("Linear predictor")+
-  theme_classic()+thm4
+  xlab("Health Region")+
+  theme_classic()+
+  thm4
 

### 2023-04-28 manuscript_v4/main.tex
Updated files

@@
 \usepackage{xcolor}
+\usepackage{amsmath}
+\usepackage{caption}
 \KOMAoption{captions}{tableheading}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, breakable, sharp corners, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, enhanced, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 \hypertarget{tbl-descriptive-stats}{}
-\begin{table}
+\setlength{\LTpost}{0mm}
+\begin{longtable}{lccc}
 \caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination
@@
 
-\centering\begingroup\fontsize{9}{11}\selectfont
-
-\begin{tabular}{lccc}
 \toprule
-\textbf{Variable} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,591 & \textbf{p-value}\\
+\textbf{Variable} & \textbf{no}, N = 958\textsuperscript{1} & \textbf{yes}, N = 2,591\textsuperscript{1} & \textbf{p-value}\textsuperscript{2} \\ 
 \midrule
-\textbf{Income (CAD)} &  &  & <0.001\\
-\hspace{1em}60000 and above & 305 (23\%) & 1,048 (77\%) & \\
-\hspace{1em}25000-59999 & 253 (28\%) & 636 (72\%) & \\
-\hspace{1em}under 25000 & 400 (31\%) & 907 (69\%) & \\
-\textbf{Age Group} &  &  & 0.7\\
-\hspace{1em}16-34 & 409 (27\%) & 1,111 (73\%) & \\
-\hspace{1em}35-54 & 252 (26\%) & 712 (74\%) & \\
-\hspace{1em}55 and over & 297 (28\%) & 768 (72\%) & \\
-\textbf{Health Region} &  &  & 0.14\\
-\hspace{1em}Toronto & 371 (28\%) & 952 (72\%) & \\
-\hspace{1em}Central & 224 (28\%) & 580 (72\%) & \\
-\hspace{1em}East & 135 (23\%) & 448 (77\%) & \\
-\hspace{1em}West & 228 (27\%) & 611 (73\%) & \\
-\textbf{Month} &  &  & 0.4\\
-\hspace{1em}October & 469 (27\%) & 1,263 (73\%) & \\
-\hspace{1em}November & 376 (28\%) & 980 (72\%) & \\
-\hspace{1em}December & 113 (25\%) & 348 (75\%) & \\
-\textbf{Race} &  &  & <0.001\\
-\hspace{1em}White/Caucasian & 233 (18\%) & 1,079 (82\%) & \\
-\hspace{1em}Arab/Middle Eastern & 76 (36\%) & 138 (64\%) & \\
-\hspace{1em}Black & 114 (38\%) & 184 (62\%) & \\
-\hspace{1em}East Asian/Pacific Islander & 69 (23\%) & 234 (77\%) & \\
-\hspace{1em}Indigenous & 76 (40\%) & 115 (60\%) & \\
-\hspace{1em}Latin American & 69 (38\%) & 111 (62\%) & \\
-\hspace{1em}Mixed & 105 (34\%) & 205 (66\%) & \\
-\hspace{1em}Other & 128 (35\%) & 239 (65\%) & \\
-\hspace{1em}South Asian & 88 (24\%) & 286 (76\%) & \\
+Income (CAD) &  &  & <0.001 \\ 
+    60000 and above & 305 (23\%) & 1,048 (77\%) &  \\ 
+    25000-59999 & 253 (28\%) & 636 (72\%) &  \\ 
+    under 25000 & 400 (31\%) & 907 (69\%) &  \\ 
+Age Group &  &  & 0.7 \\ 
+    16-34 & 409 (27\%) & 1,111 (73\%) &  \\ 
+    35-54 & 252 (26\%) & 712 (74\%) &  \\ 
+    55 and over & 297 (28\%) & 768 (72\%) &  \\ 
+Health Region &  &  & 0.14 \\ 
+    Toronto & 371 (28\%) & 952 (72\%) &  \\ 
+    Central & 224 (28\%) & 580 (72\%) &  \\ 
+    East & 135 (23\%) & 448 (77\%) &  \\ 
+    West & 228 (27\%) & 611 (73\%) &  \\ 
+Month &  &  & 0.4 \\ 
+    October & 469 (27\%) & 1,263 (73\%) &  \\ 
+    November & 376 (28\%) & 980 (72\%) &  \\ 
+    December & 113 (25\%) & 348 (75\%) &  \\ 
+Race &  &  & <0.001 \\ 
+    White/Caucasian & 233 (18\%) & 1,079 (82\%) &  \\ 
+    Arab/Middle Eastern & 76 (36\%) & 138 (64\%) &  \\ 
+    Black & 114 (38\%) & 184 (62\%) &  \\ 
+    East Asian/Pacific Islander & 69 (23\%) & 234 (77\%) &  \\ 
+    Indigenous & 76 (40\%) & 115 (60\%) &  \\ 
+    Latin American & 69 (38\%) & 111 (62\%) &  \\ 
+    Mixed & 105 (34\%) & 205 (66\%) &  \\ 
+    Other\textsuperscript{3} & 128 (35\%) & 239 (65\%) &  \\ 
+    South Asian & 88 (24\%) & 286 (76\%) &  \\ 
 \bottomrule
-\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
-\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{2} Pearson's Chi-squared test}\\
-\end{tabular}
-\endgroup{}
-\end{table}
+\end{longtable}
+\begin{minipage}{\linewidth}
+\textsuperscript{1}n (\%)\\
+\textsuperscript{2}Pearson\textquotesingle{}s Chi-squared test\\
+\textsuperscript{3}Southeast Asian, Filipino, West Asian,
+and minorities not identified elsewhere according to the Census.\\
+\end{minipage}
 
@@
 the estimated obtained from the corrected data appear in
-Figure~\ref{fig-model-corr}. The results show significantly higher odds
-of vaccination in individuals with a high household income, and those
-that identified as White or Caucasian when compared with individuals in
-lower income brackets or that identified as part of equity-deserving
-groups.
-
-Specifically, significantly lower odds of vaccination were identified
-for those with a household income under CAD 25,000 (OR=0.374,
-CI={[}0.251,0.559{]}) and those with an income between CAD 25,000 and
-59,999 (OR=0.586, CI={[}0.390,0.882{]}). Additionally, individuals who
-identified as Arab/Middle Eastern, Black, or Latin American, had
-significantly lower odds of vaccination than those in the
-White/Caucasian group (ORs=0.32, 0.32, 0.27, and \emph{p}=0.003,
-\emph{p}\textless0.001 and \emph{p}=0.004, respectively); additionally,
-those individuals that reported to belong to the ``Other''
-Race/Ethnicity group (which included the Southeast Asian, Filipino, West
-Asian, and Minorities Not Identified Elsewhere groups according to the
-Census) had even lower odds of vaccination than the other minority
-groups (OR=0.22, \emph{p}\textless0.001). Regarding Health Regions,
-individuals that reported living in the West Health Region (which
-comprises the regions of Waterloo and Niagara, the counties of
-Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
-Brant, and Chatham-Kent) had significantly higher odds of vaccination
-than those in the Health Region of Toronto (OR=1.54, \emph{p}=0.031).
-
-Moreover, statistically-significant odd ratios were determined in the
-case of the interaction of income and race; specifically, for
-individuals with a household income below CAD 25,000 who identified as
-Arab/Middle Eastern (OR=3.08, \emph{p}=0.013), Black (OR=3.15,
-\emph{p}=0.004), Latin American (OR=2.81, \emph{p}=0.041), or that
-belonged to other minority groups (OR=4.63, \emph{p}\textless0.001).
-Within the CAD 25,000-59,999 income bracket, individuals who identified
-as belonging to other racial minority groups had significantly higher
-odds of vaccination (OR=6.96, \emph{p}\textless0.001).
-
-For the interaction of Health Region and race, significant odds of
-vaccination were identified for Black individuals in the Central Health
-Region, which comprises the region of York, counties of Dufferin and
-Simcoe and the district of Muskoka (OR=0.44, \emph{p}=0.046), and in
-individuals that identified as part of other racial minorities or South
-Asian that lived in the West Health Region (ORs=0.41, \emph{p}=0.032 and
-\emph{p}=0.037, respectively).
+Figure~\ref{fig-model-corr}. The results show significantly lower odds
+of vaccination in individuals with a low household income and those that
+identified as part of equity-deserving groups when compared to
+individuals in high income brackets or that identified as White or
+Caucasian.
 
@@
 intervals for the uncorrected model. Only statistically significant
-interaction terms are shown. Full estimates can be found in the
-Appendix.}
+interaction terms are shown. Full interaction terms can be found in
+Supplementary Figure A-3.}
 
 \end{figure}
+
+Specifically, significantly lower odds of vaccination were identified
+for those with a household income under CAD 25,000 (OR=0.37,
+CI={[}0.25,0.56{]}) and those with an income between CAD 25,000 and
+59,999 (OR=0.59, CI={[}0.39,0.88{]}). Additionally, individuals who
+identified as Arab/Middle Eastern, Black, or Latin American, or that
+belonged to ``Other'' racial groups, which included the Southeast Asian,
+Filipino, West Asian, and Minorities Not Identified Elsewhere groups
+according to the Census, had significantly lower odds of vaccination
+than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and
+CIs={[}0.14,0.68{]}, {[}0.17,0.60{]}, {[}0.11,0.66{]}, and
+{[}0.12,0.41{]}, respectively). Regarding Health Regions, individuals
+that reported living in the West Health Region (which includes the
+regions of Waterloo and Niagara, the counties of Wellington, Essex, and
+Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent)
+had significantly higher odds of vaccination than those in the Health
+Region of Toronto (OR=1.54, CI={[}1.04,2.29{]}).
+
+Interestingly, individuals in equity-deserving groups with a household
+income below CAD 25,000 had higher odds of vaccination (when compared to
+those with a household income above CAD 60,000). This held true in the
+case of Arab/Middle Eastern (OR=3.08, CI={[}1.27,7.47{]}), Black
+(OR=3.15, CI={[}1.43,6.92{]}), and Latin American (OR=2.81,
+CI={[}1.04,7.59{]}) individuals, as well as respondents who belonged to
+``Other'' minority groups (OR=4.63, CI={[}2.34,9.13{]}), who also had
+higher odds of vaccination in the CAD 25,000-59,999 income bracket
+(OR=6.96, CI={[}2.67,18.16{]}). Finally, significantly lower odds of
+vaccination were identified (when compared to the Toronto Health Region)
+for Black individuals in the Central Health Region, which comprises the
+region of York, counties of Dufferin and Simcoe and the district of
+Muskoka (OR=0.44, CI={[}0.19,0.99{]}), and in individuals that
+identified as part of other racial minorities or South Asian that lived
+in the West Health Region (ORs=0.41, and CIs={[}0.18,0.92{]} and
+{[}0.18,0.95{]}, respectively).
 
@@
 intervals for the corrected model. Only statistically significant
-interaction terms are shown. Full estimates can be found in the
-Appendix.}
+interaction terms are shown. Full interaction terms can be found in
+Supplementary Figure A-4.}
 
@@
 The existence of healthcare disparities in Ontario motivated the recent
-change in the healthcare system of the province, which changed the LHIN
-model for a Health Region model in late
+change in the healthcare system of the province, which switched from the
+LHIN model to the Health Region model in late
 2019\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
-In this context, analyzing COVID-19 vaccination estimates between the
-Health Regions can serve as an indicator of ongoing intra-provincial
-disparities that may need to be addressed to ensure that the Health
-Region model is able to improve health access for the inhabitants of
-Ontario, which faces unique challenges due to its condition as the most
-populated and the most ethnically diverse province of Canada.
+This change is relatively recent, and because of this, is likely that
+there are ongoing differences in healthcare access across the province.
+In this study we hypothesized that differences in COVID-19 vaccination
+uptake were present between the Health Regions, aiming at determining
+which socio-demographic groups could be impacted by these disparities
+and to provide decision-makers with information that could be used to
+develop policies focused on reducing or eliminating these differences
+and ensure that the Health Region model is able to improve health access
+for the inhabitants of Ontario, a province that faces unique challenges
+due to its condition as the most populated and the most ethnically
+diverse province of Canada.
 
@@
 reported to have received the first dose of the COVID-19 vaccine
-(Table~\ref{tbl-descriptive-stats}). It is worth mentioning that
-province-wide vaccination rates for the period of interest are somewhat
-different from those of the survey, particularly in the case of those 55
-years of age and older, which in the survey had a vaccination rate of
-72\%, against a rate of 88.4\% reported for the closest age bracket (50
-years of age and older) reported by Public Health Ontario at the start
-of the period covered by the data (between October 1st, 2021 and
-December 12th,
-2021)\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}. In
-this case, differences between the survey and province-wide estimates
-are to be expected as the data from survey represents a random sample
-from the overall population.
-
-However, this variation did not lead to discrepant results, as the
-estimates from the model indicate that no significant differences in
-vaccination odds were identified among the age groups analyzed, in
-agreement with both overall vaccination rates reported for Canada, which
-have been relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{42}}, and
-with vaccination uptake rates across different age groups presented in
-other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{43}}.
-In other words, although vaccination rates obtained from the survey were
-slightly lower than the provincial estimates, these values still
-represented a valid approximation; this notion is reinforced by the
-consistency in the proportion of vaccination rates
-(Table~\ref{tbl-descriptive-stats}) and vaccination odds
-(\textbf{?@tbl-model}) across the period covered by the survey, which
-follow the vaccination rates from Public Health Ontario and which
-indicate that due to the relatively high coverage achieved in the
-population at that point, there were no abrupt shift in the trends,
-which increased by around 3\% across all age groups during the three
-months (October, November, and December of
-2021\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}).
-Moreover, there was good agreement between vaccination rates within each
-age in the raw dataset and province-wide estimates (e.g., a rate of 95\%
-for those with 61 years of age, Supplementary Table A-6, which is
-similar to the value reported by Public Health Ontario). It is also
-important to mention that regional differences can be masked by overall
-estimates, as when overall vaccination rates for the province (from
-Public Health Ontario) are disaggregated, it can be seen that regional
-differences during the period analyzed existed. For example, the Public
-Health Unit of Lambton (a region within the West Health Region) and the
-Public Health Unit of Haliburton, Kawartha, and the Pine Ridge District
-(an area covered by the East Health Region) reported lower vaccination
-rates (78\%) for those 50 years of age and older at the beginning of the
-period of interest, in contrast with other regions that had vaccination
-rates above
-80\%\textsuperscript{\protect\hyperlink{ref-ontario-covid}{41}}. To this
-day, differences in vaccination rates within the province continue, as
-according to Public Health Ontario, as of March of 2023 some regions
-still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{44}}.
-
-We identified significant intra-provincial differences in vaccination
-based on socio-economic and geographical factors. First, our results
-show differences in odds of vaccination in individuals with a household
-income below CAD 60,000 and in individuals belonging to visible minority
-groups. Those who identified as Arab/Middle Eastern, Black, Latin
-American, or that belonged to a minority group not included in the
-survey (Southeast Asian, Filipino, West Asian, and minority groups not
-identified elsewhere) had vaccination odds that were less than a third
-of individuals that identified as White/Caucasian
-(\textbf{?@tbl-model}). These results are consistent with other studies
-that have shown lower vaccination rates in individuals that identify as
-part of a racial minority, or that have a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{45}}.
-
-In this study, we also decided to explore the interactions between
-income and race and race and Health Region, as it is known that many
-individuals within racial minority groups perform tend to occupy certain
-types of occupations that fall within income brackets that have been
-shown to be associated with differences in vaccination uptake. In other
-words, we decided to explore if there were differences in vaccination
-within racial groups in certain income brackets and in certain the
-Health Regions. In this regard, it is interesting to note that although
-overall self-reported vaccination rates were found to be statistically
-significantly lower in various racial minority groups when compared to
-White/Caucasian individuals (\textbf{?@tbl-model}), the change in odds
-of vaccination within certain racial groups and income strata was
-actually positive, in contrast to the White/Caucasian group, for which
-vaccination odds decreased in lower income brackets (when compared to
-the CAD 60,000 and over bracket, Supplementary Figure A-3). More
-specifically, the change in odds of vaccination increased in individuals
-who identified as Arab/Middle Eastern, Black, Latin American, or
-belonging to other minority groups with a household income below CAD
-25,000, which was also true for individuals in other racial minority
-groups with an income between CAD 25,000-59,999 (\textbf{?@tbl-model},
-Supplementary Figure A-3).
-
-This result is likely due to the fact that individuals that belong
-racial minority groups tend to perform occupations that have been deemed
-as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{46},\protect\hyperlink{ref-ct2021}{47}},
-which include occupations such as grocery store workers, gas station
-workers, warehouse and distribution workers, and manufacturing workers,
-all being occupations for which an income within the significant
-brackets is to be expected. In the case of Ontario, essential workers
-had priority for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{48}},
-which would explain the higher odds of vaccination for these individuals
-in certain income brackets, in contrast to the lower odds of vaccination
-for the same type of individuals with higher household income. In other
-words, it is possible that the type of occupation played an important
-role in increasing the odds of vaccination in these racial minority
-groups.
+(Table~\ref{tbl-descriptive-stats}). We identified significant
+intra-provincial differences in vaccination based on socio-economic and
+geographical factors. First, our results show differences in odds of
+vaccination in individuals with a household income below CAD 60,000 and
+in individuals belonging to visible minority groups. Those who
+identified as Arab/Middle Eastern, Black, Latin American, or that
+belonged to a minority group not included in the survey (Southeast
+Asian, Filipino, West Asian, and minority groups not identified
+elsewhere) had vaccination odds that were less than a third of
+individuals that identified as White/Caucasian
+(Figure~\ref{fig-model-corr}). These results are consistent with other
+studies that have shown lower vaccination rates in individuals that
+identify as part of a racial minority, or that have a low household
+income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{41}}.
+In the case of Ontario, a possible rationale for this difference is
+vaccine access, which has been identified as an important contributing
+factor of disparities in vaccination for equity-deserving
+groups\textsuperscript{\protect\hyperlink{ref-njoku2021}{42}}, and that
+occurred in Toronto earlier in 2021, where it has been shown that
+vaccination venues were scarce in low socio-economic areas that had the
+highest burden of COVID-19 at the
+time\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2}}.
+
+It is interesting to note that although overall self-reported
+vaccination rates were found to be statistically significantly lower in
+various racial minority groups when compared to White/Caucasian
+individuals, the change in odds of vaccination within certain racial
+groups and income strata was actually positive, in contrast to the
+White/Caucasian group, where vaccination odds decreased in lower income
+brackets (when compared to the CAD 60,000 and over bracket,
+Supplementary Figure A-5). More specifically, the change in odds of
+vaccination increased in individuals who identified as Arab/Middle
+Eastern, Black, Latin American, or belonging to other minority groups
+with a household income below CAD 25,000, which was also true for
+individuals in other racial minority groups with an income between CAD
+25,000-59,999.
+
+This result is likely due in part to the fact that individuals that
+belong racial minority groups tend to perform occupations that have been
+deemed as ``essential'' in the context of the
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{43},\protect\hyperlink{ref-ct2021}{44}},
+which include grocery store workers, gas station, warehouse,
+distribution, and manufacturing workers, all being occupations for which
+an income within the significant brackets identified in the analysis is
+to be expected. In the case of Ontario, essential workers had priority
+for COVID-19
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{45}},
+which would explain the higher odds of vaccination for these
+individuals. In other words, it is possible that the type of occupation
+from individuals in equity-deserving groups played an important role in
+increasing the odds of vaccination.
 
@@
 West Health Region when compared to the Health Region of Toronto
-(\textbf{?@tbl-model}). The West Health Region comprises the regions of
-Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and
-the cities of Hamilton, Haldimand, Brant, and Chatham-Kent. In this
-case, a possible rationale for the results is the fact that in the
-survey, about 47\% of the entries for this Health Region corresponded to
-White/Caucasian individuals, who reported an overall 83\% vaccination
-rate (Supplementary Table A-7). However, the interaction effect of
-Health Region and race was also significant in the case of individuals
+(Figure~\ref{fig-model-corr}). The West Health Region comprises the
+regions of Waterloo and Niagara, the counties of Wellington, Essex and
+Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+In this case, these results could reflect the fact that in the survey,
+about half of the entries for this Health Region corresponded to
+White/Caucasian individuals, who reported a vaccination rate of 83\%
+(Supplementary Table A-6). However, the interaction effect of Health
+Region and race was also significant in the case of individuals
 identifying as South Asian or other minorities not included in the
-survey \textbf{?@tbl-model}. In this case, the results of the
+survey Figure~\ref{fig-model-corr}. In this case, the results of the
 interaction term in the model indicate that the odds of vaccination for
@@
 Region decreased when compared to the other Health Regions
-(Supplementary Figure A-4).
+(Supplementary Figure A-6).
 
@@
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{49}}.
-In the case of this analysis, the estimated lower odds are likely to be
-explained from a socio-economic perspective. In fact, 50\% of the
-answers from this region in the survey came from the former LHINs of
-Hamilton Niagara Haldimand Brant, and Erie St.~Clair, both which are
-among the regions of Ontario with the highest proportion of their
-population (more than 20\%) in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{50}}
-(Supplementary Table A-8). Therefore, this result partly reinforces the
-well-known existing association between low vaccination rates and
-income, but it additionally indicates that there were intra-regional
-differences in vaccination. Interestingly, a disproportionate number of
-COVID-19 cases and low vaccination rate (under 50\%) have been
-previously reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{51}}; in this
-regard, our result provides additional context by showing that within
-the South Asian community, there were differences in vaccination uptake
-across Ontario. Moreover, because significant lower odds of vaccination
-were also identified in other minority groups, this provides a rationale
-for future studies that explore how vaccination uptake varies across
-different minority groups within Ontario and other Canadian provinces.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}.
+Therefore, the estimated lower odds are likely to be explained from a
+socio-economic perspective. In fact, 50\% of the answers from this
+region in the survey came from the former LHINs of Hamilton Niagara
+Haldimand Brant, and Erie St.~Clair, both which are among the regions of
+Ontario with the highest proportion of their population (more than 20\%)
+in the lowest income
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{47}}
+(Supplementary Table A-7). Interestingly, a disproportionate number of
+COVID-19 cases and low vaccination rate (under 50\%) have been also
+reported in the South Asian community of
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{48}}. These
+results provide context to the observed differences in vaccination among
+the Health Regions of the province, showing which equity-deserving and
+socio-demographic groups in certain regions need to be prioritized in
+future vaccination strategies. Moreover, these results provide a
+rationale for future studies that explore how vaccination uptake varies
+across different minority groups within Ontario and other Canadian
+provinces.
 
@@
 are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{49}}, a
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}, a
 minority group that has historically suffered from reduced access to
@@
 
-The results in this study are based on self-reported data, where there
-is a risk that biased values are reported. Despite this, because in the
-context of COVID-19 it has been shown that good agreement exists between
-self-reported and documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{52}}, and
-therefore, the effect of self-reported bias is likely to not be
-significant in our analyses. Finally, it is likely that there have been
-differences in vaccination across the province as more doses of the
-vaccine were administered and as successive variants emerged. Because
-this study focused only on vaccination status regarding the first dose
-of the vaccine within a relatively short time window, it can only
-provide a snapshot of the societal dynamics behind the pandemic.
-Nonetheless, the results presented here can serve as a starting point to
-motivate the collection of robust longitudinal data that can be used to
-quantify geographical and temporal differences within vulnerable
-segments of the population, and that can be used to inform the
-development of adequate public health policies within the province of
-Ontario or across other provinces that aim to minimize disparities in
-health access.
+It is also worth mentioning that province-wide vaccination rates for the
+period of interest are somewhat different from those of the survey,
+particularly in the case of those 55 years of age and older, which in
+the survey had a vaccination rate of 72\%, against a rate of 88.4\%
+reported for the closest age bracket (50 years of age and older)
+reported by Public Health Ontario at the start of the period covered by
+the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{49}}.
+However, we found good agreement between the estimates from the model
+and overall vaccination rates reported for Canada, which have been
+relatively higher when compared to other high income
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{50}}, and
+with vaccination uptake rates across different age groups presented in
+other
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{51}}.
+
+In other words, although vaccination rates obtained from the survey were
+slightly lower than the provincial estimates, these values still
+represented a valid approximation to overall trends; this notion is
+reinforced by the consistency in the proportion of vaccination rates
+(Table~\ref{tbl-descriptive-stats}) and vaccination odds
+(Figure~\ref{fig-model-corr}) across the period covered by the survey,
+which closely match the vaccination rates from Public Health Ontario and
+which indicate that due to the relatively high coverage achieved in the
+population at that point, vaccination rates increased by around 3\%
+across all age groups during the three months covered by the
+survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{49}}. It is
+also important to mention that to this day, differences in vaccination
+rates within the province continue. As of March of 2023 some regions
+still have less than 75\% vaccination
+rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{52}}, and
+although data for the period analyzed in this study is not
+publicly-available, it is likely that differences in vaccination rates
+were higher at the time, being partially captured by the survey.
+
+The results in this study are based on self-reported data, where bias
+might be present. However, because in the context of COVID-19 it has
+been shown that good agreement exists between self-reported and
+documented vaccination
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{53}}, we
+believe that our data was able to provide a valid assessment of
+vaccination in the province. Finally, this study focused on first-dose
+vaccination status within a relatively short time window, and therefore
+can only provide a snapshot of the societal dynamics behind the
+pandemic. Nonetheless, the results presented here can serve as a
+starting point to motivate the collection of robust longitudinal data
+that can be used to quantify geographical and temporal differences
+within vulnerable segments of the population, and that can be used to
+inform the development of adequate public health policies within the
+province of Ontario or across other provinces that aim to minimize
+disparities in health access.
 
@@
 This study explored differences in COVID-19 vaccination across the
-province of Ontario between late 2021 and early 2022 by taking into
+province of Ontario during the last quarter of 2021 taking into
 consideration socio-economic factors, such as income and race, their
 interactions, and the Health Regions within the province. Our results
-show that, during the period analyzed, significant differences in
-vaccination existed across different visible minority groups, income
-brackets, and Health Regions, showing intra-provincial disparities in
-vaccine uptake. As the COVID-19 continues around the world, it important
-that future public policies take into consideration how to adequately
-reach individuals within minority groups that live across geographical
-areas where less probabilities of being vaccinated are likely. At the
-moment, this is an ongoing issue that needs to be addressed to ensure a
-more homogeneous outcome from the pandemic.
+show that during the period analyzed, differences in vaccination uptake
+existed across multiple equity-deserving groups in the province, and
+that these differences were significant in two of the Health Regions
+analyzed. It is important that future public health policies in Ontario
+take into consideration how to adequately reach individuals from
+equity-deserving groups that might live in areas of the province where
+access to healthcare might be difficult. Only in this way the goal of
+the Health Region model, which aims at reducing disparities, will become
+successful.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
 \CSLLeftMargin{41. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
-2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{42. }%
-\CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
-compulsion: The case of {COVID}-19 vaccination in canada.
-\emph{Vaccine}. 2022;40(29):3923-3926.
-doi:\href{https://doi.org/10.1016/j.vaccine.2022.05.053}{10.1016/j.vaccine.2022.05.053}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{43. }%
-\CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
-of canada {COVID}-19 report: Enhancing {COVID}-19 vaccine acceptance in
-canada. Blais JM, ed. \emph{{FACETS}}. 2021;6:1184-1246.
-doi:\href{https://doi.org/10.1139/facets-2021-0037}{10.1139/facets-2021-0037}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{44. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=maps}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{45. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
+Structural barriers and racial and ethnic inequities in {COVID}-19
+vaccination. \emph{International Journal of Environmental Research and
+Public Health}. 2021;18(18):9904.
+doi:\href{https://doi.org/10.3390/ijerph18189904}{10.3390/ijerph18189904}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
+\CSLLeftMargin{49. }%
+\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
+2023.
+\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
+\CSLLeftMargin{50. }%
+\CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
+compulsion: The case of {COVID}-19 vaccination in canada.
+\emph{Vaccine}. 2022;40(29):3923-3926.
+doi:\href{https://doi.org/10.1016/j.vaccine.2022.05.053}{10.1016/j.vaccine.2022.05.053}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
+\CSLLeftMargin{51. }%
+\CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
+of canada {COVID}-19 report: Enhancing {COVID}-19 vaccine acceptance in
+canada. Blais JM, ed. \emph{{FACETS}}. 2021;6:1184-1246.
+doi:\href{https://doi.org/10.1139/facets-2021-0037}{10.1139/facets-2021-0037}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
+\CSLLeftMargin{52. }%
+\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
+\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=maps}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-05-04 code/complete_case_test.R
Working with "complete" cases

@@ -1,0 +1,387 @@
+library(here)
+library(tidyverse)
+library(naniar)
+library(gtsummary)
+library(lubridate)
+
+rawdata<-read.csv(here("data","Fields_data.csv"),na.strings = c(""))
+clean_data <-read.csv(here("data","clean_dataset.csv"))
+
+## select the variables used in the analysis
+
+rawdata <- rawdata %>%
+  select(age,age_group,q09_hh_income,q16_race,city,q02_first_dose,complete,time_date_code)
+
+# The dataset had a column called "complete". The data dictionary was not
+# clear about what this meant, it seems that observations that are "complete" 
+# have complete observations for vaccination status, and all other covariates, except race, with a 4%
+# of missing obs.
+
+test<-rawdata %>%
+  subset(complete=="complete") 
+
+test <- test %>%
+  rename(date=time_date_code,
+    first_dose=q02_first_dose,
+         income=q09_hh_income,
+         race=q16_race,
+  )%>%
+  mutate(across(c("age_group",
+                  "city",
+                  "income",
+                  "race"),as.factor))
+
+
+## Recode income
+
+test<-test%>%
+  mutate(income_ord = case_when(
+    income %in% "under_15000" ~ "under 25000",
+    income %in% "15000_24999" ~ "under 25000",
+    income %in% "25000_39999" ~ "25000-59999",
+    income %in% "40000_59999" ~ "25000-59999",
+    income %in% "60000_89999" ~ "60000 and above",
+    income %in% "90000_109999" ~ "60000 and above",
+    income %in%"over_110000" ~ "60000 and above",
+  ))%>%
+  mutate(income_ord=as.factor(income_ord))
+
+#re-group age groups
+
+test<-test%>%
+  mutate(Age_group_ord = case_when(
+    age_group %in% "16_24" ~ "16-34",
+    age_group %in% "25_34" ~ "16-34",
+    age_group %in% "35_44" ~ "35-54",
+    age_group %in% "45_54" ~ "35-54",
+    age_group %in% "55_64" ~ "55 and over",
+    age_group %in% "65_and_over" ~ "55 and over",
+  ))%>%
+  mutate(Age_group_ord=as.factor(Age_group_ord))
+
+#fix Race/Ethnicity levels
+
+test<- test %>%
+  mutate(Race= case_when(
+    race %in% "white_caucasian" ~ "White/Caucasian",
[…]

### 2023-05-04 code/missing_data_analysis.R

@@ -1,0 +1,90 @@
+## script to analyze missing data pattern in the original dataset and compare with
+## the clean dataset
+library(here)
+library(tidyverse)
+library(naniar)
+library(gtsummary)
+
+rawdata<-read.csv(here("data","Fields_data.csv"),na.strings = c(""))
+clean_data <-read.csv(here("data","clean_dataset.csv"))
+
+## select the variables used in the analysis
+
+rawdata <- rawdata %>%
+ select(age,age_group,q09_hh_income,q16_race,city,q02_first_dose,complete)
+
+
+# recode variables to match the column names from the clean dataset , and re-group age and income
+
+
+
+rawdata <- rawdata %>%
+  rename(first_dose=q02_first_dose,
+         income=q09_hh_income,
+         race=q16_race,
+  )%>%
+  mutate(across(c("age_group",
+                  "city",
+                  "income",
+                  "race"),as.factor))
+
+
+## Recode income
+
+rawdata<-rawdata%>%
+  mutate(income_ord = case_when(
+    income %in% "under_15000" ~ "under 25000",
+    income %in% "15000_24999" ~ "under 25000",
+    income %in% "25000_39999" ~ "25000-59999",
+    income %in% "40000_59999" ~ "25000-59999",
+    income %in% "60000_89999" ~ "60000 and above",
+    income %in% "90000_109999" ~ "60000 and above",
+    income %in%"over_110000" ~ "60000 and above",
+  ))%>%
+  mutate(income_ord=as.factor(income_ord))
+
+#re-group age groups
+
+rawdata<-rawdata%>%
+  mutate(Age_group_ord = case_when(
+    age_group %in% "16_24" ~ "16-34",
+    age_group %in% "25_34" ~ "16-34",
+    age_group %in% "35_44" ~ "35-54",
+    age_group %in% "45_54" ~ "35-54",
+    age_group %in% "55_64" ~ "55 and over",
+    age_group %in% "65_and_over" ~ "55 and over",
+  ))%>%
+  mutate(Age_group_ord=as.factor(Age_group_ord))
+
+#fix Race/Ethnicity levels
+
+rawdata <- rawdata %>%
+  mutate(Race= case_when(
+    race %in% "white_caucasian" ~ "White/Caucasian",
+    race %in% "arab_middle_eastern" ~ "Arab/Middle Eastern",
+    race %in% "black" ~ "Black",
+    race %in% "east_asian_pacific_islander" ~ "East Asian/Pacific Islander",
[…]

### 2023-05-05 code/complete_case_test.R
Added observations, added Geographical regions and updated LHINs

@@ -1,8 +1,13 @@
 library(here)
 library(tidyverse)
 library(naniar)
+library(survey)
 library(gtsummary)
 library(lubridate)
+library(modelsummary)
+
+# for reference, the dataset used in the last version the manuscript had 3,549 obs.
+
 
 rawdata<-read.csv(here("data","Fields_data.csv"),na.strings = c(""))
 clean_data <-read.csv(here("data","clean_dataset.csv"))
@@ -19,6 +24,13 @@ rawdata <- rawdata %>%
 
 test<-rawdata %>%
   subset(complete=="complete") 
+
+# remove cities entered as "None"
+
+test<-test %>%
+  subset(city!="None")
+
+#recode variable names
 
 test <- test %>%
   rename(date=time_date_code,
@@ -88,12 +100,8 @@ municipalities$city<-as.factor(municipalities$city)
 
 test<-left_join(test,municipalities,by="city")
 
-test %>%
-  filter(is.na(Geographic.area))%>%
-  distinct(city)
 
-
-missing_municipalities<-read.csv(here("data","missing_municipalities_updated.csv"))
+missing_municipalities<-read.csv(here("data","missing_municipalities_updated_May_05_2023.csv"))
 
 
 #combining geographical regions
@@ -104,6 +112,13 @@ test<-test %>%
   select(-c(Geographic.area.x,Geographic.area.y))
 
 test$Geographic.area<-as.factor(test$Geographic.area)
+
+## what cities are missing geographical region
+
+t1<-test %>%
+  filter(is.na(Geographic.area))
+
+## no areas missing, good!
 
 # load the the titles for each region
 
@@ -172,9 +187,9 @@ missing_health_regions<-test %>% filter(is.na(LHIN)) %>%
   distinct(city,.keep_all = TRUE)%>%
   select(city,Geographic.area,Geographic_area_title,LHIN,Health_Region)
 
+## keep in mind that Mitchell was coded as Mitchell/Ontario in the dataset, might need to be chagned
 
-
-mhru<-read.csv(here("data","missing_health_regions_updated.csv"))
+mhru<-read.csv(here("data","missing_health_regions_updated_May_05_2023.csv"))
 
 #rename columns
 mhru <- mhru %>%
@@ -202,6 +217,14 @@ test<-test %>%
                      LHIN=="Erie St. Clair"~"West"
            ))%>%
[…]

### 2023-05-08 manuscript_v5/02-methods.qmd
new directory for version, tidy code

--- 2023-04-28 manuscript_v4/02-methods.qmd
@@
 
-Preliminary analyses of the data included the removal of outliers, of observations where respondents did not provide answers in all the covariates of interest, matching the city of each observations with its corresponding LHIN and Health Region, and removing observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions).
+Preliminary analyses of the data included the removal of outliers (**should we still do this?
+it's only 19 observations with income >110k and household of 1, but we are not even using such income bracket in the analysis because we re-grouped the data, and the household size variable has 90% missing rate**), of observations where respondents did not provide answers in all the covariates of interest, matching the city of each observations with its corresponding LHIN and Health Region, and removing observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions).
 After all the preliminary analyses indicated above, the total number of observations used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021.
@@
 
-Comparisons between the variables of the survey were performed using the $\chi^2$ test.
 We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
 Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data.
-The reference groups in the model were: 16-34 years (age group), October(Month), CAD 60,000 and above (income), White/Caucasian (Race), Toronto (Health Region).
 Details regarding the correction can be found in the Appendix.

### 2023-05-08 manuscript_v5/03-results.qmd
new directory for version, tidy code

--- 2023-04-28 manuscript_v4/03-results.qmd
@@
 @tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status.
-A total of 3,549 observations were used for analysis, from which 27\% (958) corresponded to individuals that reported not having received the first dose of the vaccine.
-Respondents between 16 and 34 years of age, and those that lived within the Health Region of Toronto had the highest number of observations in the data (42.8\% and 37.3\% of the total, respectively).
-Moreover, individuals who identified as White or Caucasian reported the highest vaccination uptake (82\%), in contrast to respondents who identified as which had had the lowest vaccination uptake rate (60\%) across all groups.
-From all the covariates, significant associations were identified between income, race, and vaccination rate.
+A total of **6,236** observations were used for analysis, from which 24\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
+Across the socio-demographic characteristics collected by the survey, higher vaccination uptake was reported by individuals in the highest income bracket considered, those between 16 and 34 years of age, individuals that lived in the East Health Region, and individuals that identified as White or Caucasian.
+During the months covered by the survey, the highest vaccination uptake was observed in January of 2022.
 
@@
 
-source(here::here("code","data_preparation.R"))
+source(here::here("code","data_preparation_May_08.R"))
 
@@
     footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
-  )
-
-# data_summary %>%
-#   modify_header(label = "**Variable**") %>%
-#   as_kable_extra(booktabs = TRUE,linesep="") %>%
-#   # reduce font size to make table fit. 
-#   # you may also use the `latex_options = "scale_down"` argument here.
-#   kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)
-
-data_summary %>%
+  ) %>%
   modify_header(label = "**Variable**") %>%
@@
 
-# test<-apply(xtabs(~first_dose_m + Race,clean_data), 2, chisq.test)
-# 
-# 
-# clean_data %>%
-#   filter(income_ord %in% c("60000 and above"))%>%
-#   tbl_summary(percent="row",
-#               by= first_dose_m,
-#               missing="no",
-#               include=Health_Region,
-#               label=list(
-#                  income_ord="Income (CAD)")
-#   ) %>% add_p()%>%
-#   bold_labels()
-
-# clean_data %>%
-#   filter(income_ord %in% c("25000-59999"))%>%
-#   tbl_summary(percent="row",
-#               by= income_ord,
-#               missing="no",
-#               include=(income_ord),
-#               label=list(
-#                  income_ord="Income (CAD)")
-#   ) %>% add_p()%>%
-#   bold_labels()
+data_summary
 
@@
 #call the script for raking
-source(here::here("code","raking.R"))
+source(here("code","raking.R"))
 
 #logistic regression
-source(here::here("code","regression.R"))
+source(here("code","regression.R"))
 
-#create dataframe from modelplot, remove names of factors, drop the intercept
+# call script that creates figure for model
 
-mplot1<-modelplot(m0,draw=FALSE,
-                  coef_omit = "Intercept", exponentiate = TRUE) %>%
-  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
+source(here("code","uncorr_model.R"))
 
-#add column by factor 
-mplot1$factor<-c(rep("Age",2),
-                 rep("Month",2),
-                 rep("Income",2),
-                 rep("Race/Ethnicity",8),
-                 rep("Health Region",3),
-                 rep("Income and\nRace/Ethnicity",16),
-                 rep("Race/Ethnicity\nand\nHealth Region",24)
-                 )
+#call figure created in the script above
+uncorr_mod
 
-#order factors for plot
-#mplot1$term_ord <- factor(mplot1$term,ordered=TRUE,levels=mplot1$term)
-
-mplot1$factor <- factor(mplot1$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
-
-# remove non-significant interaction factors
-
-test<-mplot1%>%
-  filter(str_detect(term,':') & p.value>0.05)
-
-#remove non-significant interactions from dataframe
-mplot1<-anti_join(mplot1,test)
-
-mplot1<-mplot1%>%
-mutate(term=str_replace_all(term,':'," and "))
-
-#make plot
-mplot1 %>% 
-ggplot(aes(y=term))+
-  geom_vline(xintercept = 1, linetype="dashed",color="red") +
-  geom_point(aes(x=estimate),size=2)+
-  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
-  scale_y_discrete(limits=rev)+
-  xlab("")+
-  ylab("")+
-  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y",labeller = label_wrap_gen(multi_line = TRUE))+
-  theme_classic()
 ```
@@
 
-#create dataframe from modelplot, remove names of factors, drop the intercept
+#call script with code for figure
+source(here("code","corr_model.R"))
 
-mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept",exponentiate = TRUE) %>%
-  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
-
-#add column by factor 
-mplot2$factor<-c(rep("Age",2),
-                 rep("Month",2),
-                 rep("Income",2),
-                 rep("Race/Ethnicity",8),
-                 rep("Health Region",3),
-                 rep("Income and\nRace/Ethnicity",16),
-                 rep("Race/Ethnicity\nand\nHealth Region",24)
-                 )
-
-#order factors for plot
-#mplot2$term_ord <- factor(mplot2$term,ordered=TRUE,levels=mplot2$term)
-
-mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
-
-# remove non-significant interaction factors
-
-test<-mplot2%>%
-  filter(str_detect(term,':') & p.value>0.05)
-
-mplot2<-anti_join(mplot2,test)
-
-mplot2<-mplot2%>%
-mutate(term=str_replace_all(term,':'," and "))
-
-mplot2 %>% 
-ggplot(aes(y=term))+
-  geom_vline(xintercept = 1, linetype="dashed",color="red") +
-  geom_point(aes(x=estimate),size=2)+
-  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
-  scale_y_discrete(limits=rev)+
-  xlab("")+
-  ylab("")+
-  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
-  theme_classic()
+#call figure
+corr_mod
 ```

### 2023-05-08 manuscript_v5/04-discussion.qmd
new directory for version, tidy code

--- 2023-04-28 manuscript_v4/main.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
+# Discussion
 
-\usepackage{amsmath,amssymb}
-\usepackage{lmodern}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+The existence of healthcare disparities in Ontario motivated the recent change in the healthcare system of the province, which switched from the LHIN model to the Health Region model in late 2019  [@tsasis2012;@dong2022].
+This change is relatively recent, and because of this, is likely that there are ongoing differences in healthcare access across the province.
+In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions, aiming at determining which socio-demographic groups could be impacted by these disparities and to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to improve health access for the inhabitants of Ontario, a province that faces unique challenges due to its condition as the most populated and the most ethnically diverse province of Canada.
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of surveyed individuals reported to have received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats).
+We identified significant intra-provincial differences in vaccination based on socio-economic and geographical factors.
+First, our results show differences in odds of vaccination in individuals with a household income below CAD 60,000 and in individuals belonging to visible minority groups.
+Those who identified as Arab/Middle Eastern, Black, Latin American, or  that belonged to a minority group not included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were less than a third of individuals that identified as White/Caucasian (@fig-model-corr).
+These results are consistent with other studies that have shown lower vaccination rates in individuals that identify as part of a racial minority, or that have a low household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
+In the case of Ontario, a possible rationale for this difference is vaccine access, which has been identified as an important contributing factor of disparities in vaccination for equity-deserving groups [@njoku2021], and that occurred in Toronto earlier in 2021, where it has been shown that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 at the time [@bogoch2022].
 
-%line numbers
-%\usepackage{mathpazo}
-\usepackage{lineno}
-\linenumbers
+It is interesting to note that although overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets (when compared to the CAD 60,000 and over bracket, Supplementary Figure A-5).
+More specifically, the change in odds of vaccination increased in individuals who identified as Arab/Middle Eastern, Black, Latin American, or belonging to other minority groups with a household income below CAD 25,000, which was also true for individuals in other racial minority groups with an income between CAD 25,000-59,999.
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-\usepackage{amsmath}
-\usepackage{caption}
-\KOMAoption{captions}{tableheading}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
-\ifdefined\contentsname
-  \renewcommand*\contentsname{Table of contents}
-\else
-  \newcommand\contentsname{Table of contents}
-\fi
-\ifdefined\listfigurename
-  \renewcommand*\listfigurename{List of Figures}
-\else
-  \newcommand\listfigurename{List of Figures}
-\fi
-\ifdefined\listtablename
-  \renewcommand*\listtablename{List of Tables}
-\else
-  \newcommand\listtablename{List of Tables}
-\fi
-\ifdefined\figurename
-  \renewcommand*\figurename{Figure}
-\else
-  \newcommand\figurename{Figure}
-\fi
-\ifdefined\tablename
-  \renewcommand*\tablename{Table}
-\else
-  \newcommand\tablename{Table}
-\fi
-}
-\@ifpackageloaded{float}{}{\usepackage{float}}
-\floatstyle{ruled}
-\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata},
-  pdfauthor={Ariel Mundo Ortiz1,2; Bouchra Nasri1,2,},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
+This result is likely due in part to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include grocery store workers, gas station, warehouse, distribution, and manufacturing workers, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
+In the case of Ontario, essential workers had priority for COVID-19 vaccination [@mishra2021], which would explain the higher odds of vaccination for these individuals.
+In other words, it is possible that the type of occupation from individuals in equity-deserving groups played an important role in increasing the odds of vaccination.
 
-\title{\textbf{Differences in COVID-19 vaccination in the province of
-Ontario across Health Regions and socio-economic strata}}
-\author{Ariel Mundo Ortiz\textsuperscript{1,2} \and Bouchra
-Nasri\textsuperscript{1,2,*}}
-\date{}
+Additionally, significant higher vaccination odds were identified in the West Health Region when compared to the Health Region of Toronto (@fig-model-corr).
+The West Health Region comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+In this case, these results could reflect the fact that in the survey, about half of the entries for this Health Region corresponded to White/Caucasian individuals, who reported a vaccination rate of 83\% (Supplementary Table A-6).
+However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@fig-model-corr].
+In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-6).
 
-\begin{document}
-\maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, enhanced, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
+According to Ontario Health, 13.2% of the population in the West Health Region identifies as a visible minority, whereas 2.5% identifies as Indigenous [@ontariohealth].
+Therefore, the estimated lower odds are likely to be explained from a socio-economic perspective.
+In fact, 50% of the answers from this region in the survey came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
+Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-7).
+Interestingly, a disproportionate number of COVID-19 cases and low vaccination rate (under 50%) have been also reported in the South Asian community of Ontario [@anand2022].
+These results provide context to the observed differences in vaccination among the Health Regions of the province, showing which equity-deserving and socio-demographic groups in certain regions need to be prioritized in future vaccination strategies.
+Moreover, these results provide a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.
 
-\textsuperscript{1} Centre de Recherches Mathématiques, University of
-Montreal, Montréal, Canada\\
-\textsuperscript{2} Department of Social and Preventive Medicine, École
-de Santé Publique, University of Montreal, Montréal, Canada
+There are some limitations to the present study.
+First, the data collection design, which allowed respondents to withdraw from the survey at any point, resulted in a high number of unique entries in the survey with multiple missing answers.
+Because we focused on entries that had complete observations in the covariates of interest for our analysis, it is possible that some information was not considered by excluding observations that had information in other variables (such as work from home, or number of persons in the household).
+However, we attempted to minimize this possibility by correcting the dataset using information from the Census.
+More granular corrections, which for example could be based on demographic information by municipality, could be used in the future to obtain a more accurate approximation to the population totals of the province.
+Moreover, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map).
+Although low representation from these areas is based on the fact that these regions only account for 5\% of the total population of Ontario, these regions are the home to more than 100,000 individuals that identify as Indigenous [@ontariohealth], a minority group that has historically suffered from reduced access to health care and discrimination [@mosby2021].
+Therefore, there is a need for additional studies that focus on these low-populated Health Regions in Ontario where disparities in vaccination might be significant and understudied.
 
-\textsuperscript{*} Correspondence:
-\href{mailto:bouchra.nasri@umontreal.ca}{Bouchra Nasri
-\textless{}bouchra.nasri@umontreal.ca\textgreater{}}
+It is also worth mentioning that province-wide vaccination rates for the period of interest are somewhat different from those of the survey, particularly in the case of those 55 years of age and older, which in the survey had a vaccination rate of 72\%, against a rate of 88.4\% reported for the closest age bracket (50 years of age and older) reported by Public Health Ontario at the start of the period covered by the data [@ontario-covid].
+However, we found good agreement between the estimates from the model and overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], and with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022].
 
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
+In other words, although vaccination rates obtained from the survey were slightly lower than the provincial estimates, these values still represented a valid approximation to overall trends; this notion is reinforced by the consistency in the proportion of vaccination rates (@tbl-descriptive-stats) and vaccination odds (@fig-model-corr) across the period covered by the survey, which closely match the vaccination rates from Public Health Ontario and which indicate that due to the relatively high coverage achieved in the population at that point, vaccination rates increased by around 3% across all age groups during the three months covered by the survey [@ontario-covid].
+It is also important to mention that to this day, differences in vaccination rates within the province continue.
+As of March of 2023 some regions still have less than 75\% vaccination rate [@ontario-covid-map], and although data for the period analyzed in this study is not publicly-available, it is likely that differences in vaccination rates were higher at the time, being partially captured by the survey.
 
-The COVID-19 pandemic continues to be a worldwide public health concern.
-Although vaccines against this disease were rapidly developed,
-vaccination uptake has not ben equal across all the segments of the
-population. In particular, it has been shown that there have been
-differences in vaccine uptake across different segments of the
-population. However, there are also differences in vaccination across
-geographical areas, which might be important to consider in the
-development of future public health policies against COVID-19. In this
-study, we examined the relationship between vaccination status (having
-received the first dose of a COVID-19 vaccine), and different
-socio-economic and geographical factors. Our results show that during
-the last three months of 2021, individuals in certain equity-deserving
-groups (visible minorities) were three times less likely to be
-vaccinated than White/Caucasian individuals across the province and that
-in some cases, within these groups individuals in low income brackets
-had significantly higher odds of vaccination when compared to their
-peers in high income brackets. Finally, we identified significantly
-lower odds of vaccination in the West Health Region of Ontario within
-certain equity-deserving groups. This study shows that there is an
-ongoing need to better understand and address differences in vaccination
-uptake across diverse segments of the population of Ontario that have
-been largely impacted by the pandemic.
-
-\hypertarget{keywords}{%
-\section*{Keywords}\label{keywords}}
-\addcontentsline{toc}{section}{Keywords}
-
-Covid-19, vaccination, survey, socio-economic factors, visible
-minorities.
-
-\hypertarget{background}{%
-\section{Background}\label{background}}
-
-The vaccines against COVID-19 have been considered a major achievement
-of modern medicine as their rapid development allowed the start of broad
-vaccination campaigns towards the end of 2020 in certain countries, such
-as the US and
-Canada\textsuperscript{\protect\hyperlink{ref-davis2022}{1}--\protect\hyperlink{ref-tanne2020}{3}}.
-This made some believe that vaccines were destined to be a determinant
-factor in a rapid ending of the
-pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{4}}.
-However, although it has been estimated that COVID-19 vaccines have
-prevented around 14 million of deaths
-worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{5}}, their
-implementation has been far from being equal to that of the smallpox and
-polio vaccines, which were implemented on a global scale and that were
-crucial to control these
-diseases\textsuperscript{\protect\hyperlink{ref-kayser2021}{6}}. In
-fact, the rollout of COVID-19 vaccines has faced multiple challenges
-since its inception which ultimately have hampered their use to achieve
-the ultimate goal of global immunity.
-
-This problematic in the rollout of the COVID-19 vaccines is a
-multifaceted issue resulting from, among other things, the development
-of new variants due to inadequate public health
-measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
-in vaccine access between high- and low-income
-countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{8},\protect\hyperlink{ref-tamey2022}{9}},
-vaccine
-hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{10}}, and
-differences in vaccination uptake across different segments of the
-population\textsuperscript{\protect\hyperlink{ref-malik2020}{11}}. In
-particular, it is well established that differences in vaccination
-uptake have been present even in countries that have had ample access to
-vaccines since 2020 (such as the US, the UK, and Canada), where lower
-vaccine uptake has been observed within certain racial groups (i.e.,
-individuals that identify as Black, Asian, or Indigenous), and in
-individuals within low income
-brackets\textsuperscript{\protect\hyperlink{ref-willis2021}{12}--\protect\hyperlink{ref-khubchandani2021}{15}}.
-Reasons given for lower vaccine uptake in these cases have included
-medical mistrust due to systemic medical
-racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{14}}, mistrust
-in vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{12}}, and
-the influence of conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
-Moreover, in the case of Canada, lower vaccine uptake has been observed
-in young individuals, those with a low educational level, households
-with children, those without a regular healthcare provider, individuals
-that identify as part of certain equity-deserving groups, and those with
-a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.
-
-However, it is important to consider that vaccination uptake can also be
-influenced by geographical (spatial) factors. In this regard,
-differences in COVID-19 vaccination rates have been associated with
-varied regional attitudes towards
-vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{11}},
-spatial differences in vaccine access and supply, vaccination location
-availability, and lack of prioritization of areas where vulnerable
-groups
-reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2},\protect\hyperlink{ref-nguyen2021}{22}}.
-Other studies have also shown heterogeneity in vaccine uptake within
-small governmental administrative units such as
-counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{23}--\protect\hyperlink{ref-bhuiyan2022}{26}},
-and that and that accounting for geographical differences in vaccination
-can help predict patterns of booster
-uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{27}}. Overall,
-the evidence provided by the literature demonstrates the existence of
-spatially-driven heterogeneities in vaccine uptake that be used by
-decision-makers in the development of public health policies that are
-focused on addressing these disparities within specific administrative
-or geographical areas.
-
-However, such analyses have been carried mostly in territories outside
-of Canada, where available studies have been focused in certain cities
-(such as Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}},
-or Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}),
-or have explored differences at a province-wide
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. Thus, there
-is a need for studies that explore spatial differences in vaccination
-within the Canadian territory and that consequently, can help identify
-disparities that need to be addressed within specific areas in each
-province.
-
-This need is particularly important in the case of Ontario, the most
-populated province in Canada. Between 2006 and 2019, Ontario provided
-healthcare access to its inhabitants using 14 intra-provincial divisions
-called the Local Health Integrated Networks (LHINs). However, this
-approach was complex, bureaucratic, and led to systemic
-inequalities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. In
-late 2019, the 14 LHINs were phased out and the areas they covered were
-incorporated into 6 Health Regions (North East, North West, Central,
-Toronto, West, and East) in an effort to improve the healthcare system
-of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
-Because the adoption of the Health Regions is relatively recent, there
-is an ongoing need to analyze the impact of this measure and identify
-disparities in health access that might exist across the Health Regions,
-which can be specially important in the context of the COVID-19
-pandemic.
-
-Therefore, in this study we hypothesized that there were differences in
-vaccination uptake between the different Health Regions of Ontario
-during the last quarter of 2021. By including socio-economic factors in
-our analysis, we aimed at identifying in which groups these differences
-were significant in order to provide an assessment of the current state
-of healthcare access in Ontario.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{sec-data}{%
-\subsection{Data and Methods}\label{sec-data}}
-
-We used data from the \emph{Survey of COVID-19 related Behaviours and
-Attitudes}, a repeated cross sectional survey focused on the Canadian
-province of Ontario that was commissioned by the Fields Institute for
-Research in Mathematical Sciences and the Mathematical Modelling of
-COVID-19 Task Force under ethical guidance from the University of
-Toronto, and which ran between September 30th, 2021 and January
-17th,2022. The survey collected socio-economic information from
-participants (Table~\ref{tbl-descriptive-stats}), their location
-(nearest municipality, as shown in Figure~\ref{fig-map}), the date of
-access to the survey, and asked information on vaccination status by
-using the question ``Have you received the first dose of the COVID
-vaccine?'', with possible answers ``yes'' and ``no''. The original
-dataset contained 39,029 observations (where each observation
-corresponded to a unique respondent).
-
-Preliminary analyses of the data included the removal of outliers, of
-observations where respondents did not provide answers in all the
-covariates of interest, matching the city of each observations with its
-corresponding LHIN and Health Region, and removing observations from
-areas with low representation (107 observations corresponding to the
-North West and North East Health Regions). After all the preliminary
-analyses indicated above, the total number of observations used for
-analysis was 3,549 which included the East, Central, Toronto, and West
-Health Regions covering the period between October 1st,2021 and December
-12th, 2021. The original dataset, clean dataset, and details on the data
-cleaning process are described in detail in the
-\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
-for this paper.
-
-\begin{figure}
-
-{\centering \includegraphics{../data/map_data/map.pdf}
-
-}
-
-\caption{\label{fig-map}Geographic representation of the data collected
-by the \emph{Survey of COVID-19 related Behaviours and Attitudes},
-collected by the Fields Institute in Ontario. The municipalities
-(cities) from where survey participants provided answers (in the clean
-dataset) appear as points. The Health six Regions are color-coded.
-Internal boundaries within certain Health Regions indicate areas that
-belonged to the Local Integrated Health Networks (LHINs), the geographic
-areas for healthcare in Ontario before the adoption of the Health
-Regions.}
-
-\end{figure}
-
-\hypertarget{statistical-analyses}{%
-\subsection{Statistical analyses}\label{statistical-analyses}}
-
-Comparisons between the variables of the survey were performed using the
-\(\chi^2\) test. We used a logistic regression model to examine the
-impact of the Health Regions in vaccination rates while considering the
-socio-economic factors and and months covered by the survey
-(Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
-Health Region and Race and income), as previous studies have shown that
-socio-economic factors and their interactions are significant predictors
-of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
-Because we identified differences in representativity between the survey
-data and the estimates from the Census, we used an iterative
-proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{35}}
-to correct the data using data from the Census and Health Region
-population totals; and fitted the regression model to the uncorrected
-and corrected data. The reference groups in the model were: 16-34 years
-(age group), October(Month), CAD 60,000 and above (income),
-White/Caucasian (Race), Toronto (Health Region). Details regarding the
-correction can be found in the Appendix. All analyses were conducted in
-R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{39}},
-and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{40}}.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{sample-characteristics}{%
-\subsection{Sample Characteristics}\label{sample-characteristics}}
-
-Table~\ref{tbl-descriptive-stats} shows the characteristics of the data
-from the Fields COVID-19 survey used for analysis with regard to
-vaccination status. A total of 3,549 observations were used for
-analysis, from which 27\% (958) corresponded to individuals that
-reported not having received the first dose of the vaccine. Respondents
-between 16 and 34 years of age, and those that lived within the Health
-Region of Toronto had the highest number of observations in the data
-(42.8\% and 37.3\% of the total, respectively). Moreover, individuals
-who identified as White or Caucasian reported the highest vaccination
-uptake (82\%), in contrast to respondents who identified as which had
-had the lowest vaccination uptake rate (60\%) across all groups. From
-all the covariates, significant associations were identified between
-income, race, and vaccination rate.
-
-\hypertarget{tbl-descriptive-stats}{}
-\setlength{\LTpost}{0mm}
-\begin{longtable}{lccc}
-\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination
-Status) }\tabularnewline
-
-\toprule
-\textbf{Variable} & \textbf{no}, N = 958\textsuperscript{1} & \textbf{yes}, N = 2,591\textsuperscript{1} & \textbf{p-value}\textsuperscript{2} \\ 
-\midrule
-Income (CAD) &  &  & <0.001 \\ 
-    60000 and above & 305 (23\%) & 1,048 (77\%) &  \\ 
-    25000-59999 & 253 (28\%) & 636 (72\%) &  \\ 
-    under 25000 & 400 (31\%) & 907 (69\%) &  \\ 
-Age Group &  &  & 0.7 \\ 
-    16-34 & 409 (27\%) & 1,111 (73\%) &  \\ 
-    35-54 & 252 (26\%) & 712 (74\%) &  \\ 
-    55 and over & 297 (28\%) & 768 (72\%) &  \\ 
-Health Region &  &  & 0.14 \\ 
-    Toronto & 371 (28\%) & 952 (72\%) &  \\ 
-    Central & 224 (28\%) & 580 (72\%) &  \\ 
-    East & 135 (23\%) & 448 (77\%) &  \\ 
-    West & 228 (27\%) & 611 (73\%) &  \\ 
-Month &  &  & 0.4 \\ 
-    October & 469 (27\%) & 1,263 (73\%) &  \\ 
-    November & 376 (28\%) & 980 (72\%) &  \\ 
-    December & 113 (25\%) & 348 (75\%) &  \\ 
-Race &  &  & <0.001 \\ 
-    White/Caucasian & 233 (18\%) & 1,079 (82\%) &  \\ 
-    Arab/Middle Eastern & 76 (36\%) & 138 (64\%) &  \\ 
-    Black & 114 (38\%) & 184 (62\%) &  \\ 
-    East Asian/Pacific Islander & 69 (23\%) & 234 (77\%) &  \\ 
-    Indigenous & 76 (40\%) & 115 (60\%) &  \\ 
-    Latin American & 69 (38\%) & 111 (62\%) &  \\ 
-    Mixed & 105 (34\%) & 205 (66\%) &  \\ 
-    Other\textsuperscript{3} & 128 (35\%) & 239 (65\%) &  \\ 
-    South Asian & 88 (24\%) & 286 (76\%) &  \\ 
-\bottomrule
-\end{longtable}
-\begin{minipage}{\linewidth}
-\textsuperscript{1}n (\%)\\
-\textsuperscript{2}Pearson\textquotesingle{}s Chi-squared test\\
-\textsuperscript{3}Southeast Asian, Filipino, West Asian,
-and minorities not identified elsewhere according to the Census.\\
-\end{minipage}
-
-\hypertarget{multivariate-regression}{%
-\subsection{Multivariate Regression}\label{multivariate-regression}}
-
-Figure~\ref{fig-model-uncorr} shows the estimates from the logistic
-regression model of vaccination status for the uncorrected data, whereas
-the estimated obtained from the corrected data appear in
-Figure~\ref{fig-model-corr}. The results show significantly lower odds
-of vaccination in individuals with a low household income and those that
-identified as part of equity-deserving groups when compared to
-individuals in high income brackets or that identified as White or
-Caucasian.
-
-\begin{figure}
-
-\includegraphics{main_files/figure-pdf/fig-model-uncorr-1.pdf} \hfill{}
-
-\caption{\label{fig-model-uncorr}Coefficient estimates and confidence
-intervals for the uncorrected model. Only statistically significant
-interaction terms are shown. Full interaction terms can be found in
-Supplementary Figure A-3.}
-
-\end{figure}
-
-Specifically, significantly lower odds of vaccination were identified
-for those with a household income under CAD 25,000 (OR=0.37,
-CI={[}0.25,0.56{]}) and those with an income between CAD 25,000 and
-59,999 (OR=0.59, CI={[}0.39,0.88{]}). Additionally, individuals who
-identified as Arab/Middle Eastern, Black, or Latin American, or that
-belonged to ``Other'' racial groups, which included the Southeast Asian,
-Filipino, West Asian, and Minorities Not Identified Elsewhere groups
-according to the Census, had significantly lower odds of vaccination
-than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and
-CIs={[}0.14,0.68{]}, {[}0.17,0.60{]}, {[}0.11,0.66{]}, and
-{[}0.12,0.41{]}, respectively). Regarding Health Regions, individuals
-that reported living in the West Health Region (which includes the
-regions of Waterloo and Niagara, the counties of Wellington, Essex, and
-Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent)
-had significantly higher odds of vaccination than those in the Health
-Region of Toronto (OR=1.54, CI={[}1.04,2.29{]}).
-
-Interestingly, individuals in equity-deserving groups with a household
-income below CAD 25,000 had higher odds of vaccination (when compared to
-those with a household income above CAD 60,000). This held true in the
-case of Arab/Middle Eastern (OR=3.08, CI={[}1.27,7.47{]}), Black
-(OR=3.15, CI={[}1.43,6.92{]}), and Latin American (OR=2.81,
-CI={[}1.04,7.59{]}) individuals, as well as respondents who belonged to
-``Other'' minority groups (OR=4.63, CI={[}2.34,9.13{]}), who also had
-higher odds of vaccination in the CAD 25,000-59,999 income bracket
-(OR=6.96, CI={[}2.67,18.16{]}). Finally, significantly lower odds of
-vaccination were identified (when compared to the Toronto Health Region)
-for Black individuals in the Central Health Region, which comprises the
-region of York, counties of Dufferin and Simcoe and the district of
-Muskoka (OR=0.44, CI={[}0.19,0.99{]}), and in individuals that
-identified as part of other racial minorities or South Asian that lived
-in the West Health Region (ORs=0.41, and CIs={[}0.18,0.92{]} and
-{[}0.18,0.95{]}, respectively).
-
-\begin{figure}
-
-\includegraphics{main_files/figure-pdf/fig-model-corr-1.pdf} \hfill{}
-
-\caption{\label{fig-model-corr}Coefficient estimates and confidence
-intervals for the corrected model. Only statistically significant
-interaction terms are shown. Full interaction terms can be found in
-Supplementary Figure A-4.}
-
-\end{figure}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-The existence of healthcare disparities in Ontario motivated the recent
-change in the healthcare system of the province, which switched from the
-LHIN model to the Health Region model in late
-2019\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
-This change is relatively recent, and because of this, is likely that
-there are ongoing differences in healthcare access across the province.
-In this study we hypothesized that differences in COVID-19 vaccination
-uptake were present between the Health Regions, aiming at determining
-which socio-demographic groups could be impacted by these disparities
-and to provide decision-makers with information that could be used to
-develop policies focused on reducing or eliminating these differences
-and ensure that the Health Region model is able to improve health access
-for the inhabitants of Ontario, a province that faces unique challenges
-due to its condition as the most populated and the most ethnically
-diverse province of Canada.
-
-Our results indicate that across the most densely populated Health
-Regions of Ontario, almost three quarters of surveyed individuals
-reported to have received the first dose of the COVID-19 vaccine
-(Table~\ref{tbl-descriptive-stats}). We identified significant
-intra-provincial differences in vaccination based on socio-economic and
-geographical factors. First, our results show differences in odds of
-vaccination in individuals with a household income below CAD 60,000 and
-in individuals belonging to visible minority groups. Those who
-identified as Arab/Middle Eastern, Black, Latin American, or that
-belonged to a minority group not included in the survey (Southeast
-Asian, Filipino, West Asian, and minority groups not identified
-elsewhere) had vaccination odds that were less than a third of
-individuals that identified as White/Caucasian
-(Figure~\ref{fig-model-corr}). These results are consistent with other
-studies that have shown lower vaccination rates in individuals that
-identify as part of a racial minority, or that have a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{41}}.
-In the case of Ontario, a possible rationale for this difference is
-vaccine access, which has been identified as an important contributing
-factor of disparities in vaccination for equity-deserving
-groups\textsuperscript{\protect\hyperlink{ref-njoku2021}{42}}, and that
-occurred in Toronto earlier in 2021, where it has been shown that
-vaccination venues were scarce in low socio-economic areas that had the
-highest burden of COVID-19 at the
-time\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2}}.
-
-It is interesting to note that although overall self-reported
-vaccination rates were found to be statistically significantly lower in
-various racial minority groups when compared to White/Caucasian
-individuals, the change in odds of vaccination within certain racial
-groups and income strata was actually positive, in contrast to the
-White/Caucasian group, where vaccination odds decreased in lower income
-brackets (when compared to the CAD 60,000 and over bracket,
-Supplementary Figure A-5). More specifically, the change in odds of
-vaccination increased in individuals who identified as Arab/Middle
-Eastern, Black, Latin American, or belonging to other minority groups
-with a household income below CAD 25,000, which was also true for
-individuals in other racial minority groups with an income between CAD
-25,000-59,999.
-
-This result is likely due in part to the fact that individuals that
-belong racial minority groups tend to perform occupations that have been
-deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{43},\protect\hyperlink{ref-ct2021}{44}},
-which include grocery store workers, gas station, warehouse,
-distribution, and manufacturing workers, all being occupations for which
-an income within the significant brackets identified in the analysis is
-to be expected. In the case of Ontario, essential workers had priority
-for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{45}},
-which would explain the higher odds of vaccination for these
-individuals. In other words, it is possible that the type of occupation
-from individuals in equity-deserving groups played an important role in
-increasing the odds of vaccination.
-
-Additionally, significant higher vaccination odds were identified in the
-West Health Region when compared to the Health Region of Toronto
-(Figure~\ref{fig-model-corr}). The West Health Region comprises the
-regions of Waterloo and Niagara, the counties of Wellington, Essex and
-Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
-In this case, these results could reflect the fact that in the survey,
-about half of the entries for this Health Region corresponded to
-White/Caucasian individuals, who reported a vaccination rate of 83\%
-(Supplementary Table A-6). However, the interaction effect of Health
-Region and race was also significant in the case of individuals
-identifying as South Asian or other minorities not included in the
-survey Figure~\ref{fig-model-corr}. In this case, the results of the
-interaction term in the model indicate that the odds of vaccination for
-those within the South Asian and Other minority groups in the West
-Region decreased when compared to the other Health Regions
-(Supplementary Figure A-6).
-
-According to Ontario Health, 13.2\% of the population in the West Health
-Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}.
-Therefore, the estimated lower odds are likely to be explained from a
-socio-economic perspective. In fact, 50\% of the answers from this
-region in the survey came from the former LHINs of Hamilton Niagara
-Haldimand Brant, and Erie St.~Clair, both which are among the regions of
-Ontario with the highest proportion of their population (more than 20\%)
-in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{47}}
-(Supplementary Table A-7). Interestingly, a disproportionate number of
-COVID-19 cases and low vaccination rate (under 50\%) have been also
-reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{48}}. These
-results provide context to the observed differences in vaccination among
-the Health Regions of the province, showing which equity-deserving and
-socio-demographic groups in certain regions need to be prioritized in
-future vaccination strategies. Moreover, these results provide a
-rationale for future studies that explore how vaccination uptake varies
-across different minority groups within Ontario and other Canadian
-provinces.
-
-There are some limitations to the present study. First, the data
-collection design, which allowed respondents to withdraw from the survey
-at any point, resulted in a high number of unique entries in the survey
-with multiple missing answers. Because we focused on entries that had
-complete observations in the covariates of interest for our analysis, it
-is possible that some information was not considered by excluding
-observations that had information in other variables (such as work from
-home, or number of persons in the household). However, we attempted to
-minimize this possibility by correcting the dataset using information
-from the Census. More granular corrections, which for example could be
-based on demographic information by municipality, could be used in the
-future to obtain a more accurate approximation to the population totals
-of the province. Moreover, our analysis did not consider the North West
-and North East Health Regions, due to the low number of entries from
-these areas in the survey (Figure~\ref{fig-map}). Although low
-representation from these areas is based on the fact that these regions
-only account for 5\% of the total population of Ontario, these regions
-are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}, a
-minority group that has historically suffered from reduced access to
-health care and
-discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{17}}.
-Therefore, there is a need for additional studies that focus on these
-low-populated Health Regions in Ontario where disparities in vaccination
-might be significant and understudied.
-
-It is also worth mentioning that province-wide vaccination rates for the
-period of interest are somewhat different from those of the survey,
-particularly in the case of those 55 years of age and older, which in
-the survey had a vaccination rate of 72\%, against a rate of 88.4\%
-reported for the closest age bracket (50 years of age and older)
-reported by Public Health Ontario at the start of the period covered by
-the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{49}}.
-However, we found good agreement between the estimates from the model
-and overall vaccination rates reported for Canada, which have been
-relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{50}}, and
-with vaccination uptake rates across different age groups presented in
-other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{51}}.
-
-In other words, although vaccination rates obtained from the survey were
-slightly lower than the provincial estimates, these values still
-represented a valid approximation to overall trends; this notion is
-reinforced by the consistency in the proportion of vaccination rates
-(Table~\ref{tbl-descriptive-stats}) and vaccination odds
-(Figure~\ref{fig-model-corr}) across the period covered by the survey,
-which closely match the vaccination rates from Public Health Ontario and
-which indicate that due to the relatively high coverage achieved in the
-population at that point, vaccination rates increased by around 3\%
-across all age groups during the three months covered by the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{49}}. It is
-also important to mention that to this day, differences in vaccination
-rates within the province continue. As of March of 2023 some regions
-still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{52}}, and
-although data for the period analyzed in this study is not
-publicly-available, it is likely that differences in vaccination rates
-were higher at the time, being partially captured by the survey.
-
-The results in this study are based on self-reported data, where bias
-might be present. However, because in the context of COVID-19 it has
-been shown that good agreement exists between self-reported and
-documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{53}}, we
-believe that our data was able to provide a valid assessment of
-vaccination in the province. Finally, this study focused on first-dose
-vaccination status within a relatively short time window, and therefore
-can only provide a snapshot of the societal dynamics behind the
-pandemic. Nonetheless, the results presented here can serve as a
-starting point to motivate the collection of robust longitudinal data
-that can be used to quantify geographical and temporal differences
-within vulnerable segments of the population, and that can be used to
-inform the development of adequate public health policies within the
-province of Ontario or across other provinces that aim to minimize
-disparities in health access.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This study explored differences in COVID-19 vaccination across the
-province of Ontario during the last quarter of 2021 taking into
-consideration socio-economic factors, such as income and race, their
-interactions, and the Health Regions within the province. Our results
-show that during the period analyzed, differences in vaccination uptake
-existed across multiple equity-deserving groups in the province, and
-that these differences were significant in two of the Health Regions
-analyzed. It is important that future public health policies in Ontario
-take into consideration how to adequately reach individuals from
-equity-deserving groups that might live in areas of the province where
-access to healthcare might be difficult. Only in this way the goal of
-the Health Region model, which aims at reducing disparities, will become
-successful.
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
-\CSLLeftMargin{1. }%
-\CSLRightInline{Davis CJ, Golding M, McKay R. Efficacy information
-influences intention to take COVID-19 vaccine. \emph{British Journal of
-Health Psychology}. 2022;27(2):300-319.
-doi:\url{https://doi.org/10.1111/bjhp.12546}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
-\CSLLeftMargin{2. }%
-\CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
-social and policy view of vaccination efforts in ontario, canada.
-\emph{Cambridge Journal of Regions, Economy and Society}. Published
-online November 2022.
-doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
-\CSLLeftMargin{3. }%
-\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
-pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
-2020:m4799.
-doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
-\CSLLeftMargin{4. }%
-\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
-end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
-doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-watson2022}{}}%
-\CSLLeftMargin{5. }%
-\CSLRightInline{Watson OJ, Barnsley G, Toor J, Hogan AB, Winskill P,
-Ghani AC. Global impact of the first year of {COVID}-19 vaccination: A
-mathematical modelling study. \emph{The Lancet Infectious Diseases}.
-2022;22(9):1293-1302.
-doi:\href{https://doi.org/10.1016/s1473-3099(22)00320-6}{10.1016/s1473-3099(22)00320-6}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
-\CSLLeftMargin{6. }%
-\CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
-and emerging issues. \emph{Human Vaccines {\&}amp\(\mathsemicolon\)
-Immunotherapeutics}. 2021;17(12):5255-5268.
-doi:\href{https://doi.org/10.1080/21645515.2021.1977057}{10.1080/21645515.2021.1977057}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-li2021}{}}%
-\CSLLeftMargin{7. }%
-\CSLRightInline{Li Q, Wang J, Tang Y, Lu H. Next-generation COVID-19
-vaccines: Opportunities for vaccine development and challenges in
-tackling COVID-19. \emph{Drug Discoveries \& Therapeutics}.
-2021;15(3):118-123.
-doi:\href{https://doi.org/10.5582/ddt.2021.0105}{10.5582/ddt.2021.0105}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
-\CSLLeftMargin{8. }%
-\CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
-determinants of {COVID}-19 vaccine hesitancy. Inbaraj LR, ed.
-\emph{{PLOS} {ONE}}. 2021;16(11):e0258462.
-doi:\href{https://doi.org/10.1371/journal.pone.0258462}{10.1371/journal.pone.0258462}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
-\CSLLeftMargin{9. }%
-\CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
-to achieve global covid-19 vaccine equity. \emph{{BMJ}}. Published
-online March 2022:e070650.
-doi:\href{https://doi.org/10.1136/bmj-2022-070650}{10.1136/bmj-2022-070650}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
-\CSLLeftMargin{10. }%
-\CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
-inequality in {COVID}-19 vaccination coverage among elderly adults in
-england: A national linked data study. \emph{{BMJ} Open}.
-2021;11(7):e053402.
-doi:\href{https://doi.org/10.1136/bmjopen-2021-053402}{10.1136/bmjopen-2021-053402}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
-\CSLLeftMargin{11. }%
-\CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
-of {COVID}-19 vaccine acceptance in the {US}.
-\emph{{EClinicalMedicine}}. 2020;26:100495.
-doi:\href{https://doi.org/10.1016/j.eclinm.2020.100495}{10.1016/j.eclinm.2020.100495}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
-\CSLLeftMargin{12. }%
-\CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
-{COVID}-19 vaccine hesitancy: Race/ethnicity, trust, and fear.
-\emph{Clinical and Translational Science}. 2021;14(6):2200-2207.
-doi:\href{https://doi.org/10.1111/cts.13077}{10.1111/cts.13077}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
-\CSLLeftMargin{13. }%
-\CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
-accepting {COVID}-19 vaccination during and after pregnancy, and for
-their babies: A multi-methods study in the {UK}. \emph{{BMC} Pregnancy
-and Childbirth}. 2022;22(1).
-doi:\href{https://doi.org/10.1186/s12884-021-04321-3}{10.1186/s12884-021-04321-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
-\CSLLeftMargin{14. }%
-\CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
-limits of medical trust in mitigating {COVID}-19 vaccine hesitancy among
-black americans. \emph{Journal of General Internal Medicine}.
-2021;36(11):3629-3631.
-doi:\href{https://doi.org/10.1007/s11606-021-06743-3}{10.1007/s11606-021-06743-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
-\CSLLeftMargin{15. }%
-\CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
-Sharma M, Webb FJ. {COVID}-19 vaccination hesitancy in the united
-states: A rapid national assessment. \emph{Journal of Community Health}.
-2021;46(2):270-277.
-doi:\href{https://doi.org/10.1007/s10900-020-00958-x}{10.1007/s10900-020-00958-x}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
-\CSLLeftMargin{16. }%
-\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
-related medical mistrust, health impacts, and potential vaccine
-hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
-of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
-doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
-\CSLLeftMargin{17. }%
-\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
-roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
-canada. \emph{Canadian Medical Association Journal}.
-2021;193(11):E381-E383.
-doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
-\CSLLeftMargin{18. }%
-\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
-hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
-and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
-2020;52(14):3127-3141.
-doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
-\CSLLeftMargin{19. }%
-\CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
-inequalities in {COVID}-19 vaccination uptake and intent: Results from
-the canadian community health survey 2021. \emph{{BMC} Public Health}.
-2022;22(1).
-doi:\href{https://doi.org/10.1186/s12889-022-14090-z}{10.1186/s12889-022-14090-z}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
-\CSLLeftMargin{20. }%
-\CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
-Fahlman M, Kallio N. {COVID}-19 vaccine hesitancy and refusal and
-associated factors in an adult population in saskatchewan, canada:
-Evidence from predictive modelling. Gesser-Edelsburg A, ed. \emph{{PLOS}
-{ONE}}. 2021;16(11):e0259513.
-doi:\href{https://doi.org/10.1371/journal.pone.0259513}{10.1371/journal.pone.0259513}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
-\CSLLeftMargin{21. }%
-\CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
-Overcoming {COVID}-19 vaccine hesitancy among ethnic minorities: A
-systematic review of {UK} studies. \emph{Vaccine}.
-2022;40(25):3413-3432.
-doi:\href{https://doi.org/10.1016/j.vaccine.2022.04.030}{10.1016/j.vaccine.2022.04.030}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
-\CSLLeftMargin{22. }%
-\CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
-Changes in {COVID}-19 vaccination receipt and intention to vaccinate by
-socioeconomic characteristics and geographic area, united states,
-january 6 {\textendash} march 29, 2021. \emph{Annals of Medicine}.
-2021;53(1):1419-1428.
-doi:\href{https://doi.org/10.1080/07853890.2021.1957998}{10.1080/07853890.2021.1957998}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
-\CSLLeftMargin{23. }%
-\CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
-vaccine hesitancy in the united states. \emph{International Journal of
-Environmental Research and Public Health}. 2021;18(18):9488.
-doi:\href{https://doi.org/10.3390/ijerph18189488}{10.3390/ijerph18189488}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
-\CSLLeftMargin{24. }%
-\CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
-spatial process: {COVID}-19 fully vaccinated rates in u.s. counties.
-\emph{American Journal of Preventive Medicine}. 2022;63(6):954-961.
-doi:\href{https://doi.org/10.1016/j.amepre.2022.06.006}{10.1016/j.amepre.2022.06.006}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
-\CSLLeftMargin{25. }%
-\CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
-the spatiotemporal heterogeneity of the {COVID}-19 vaccination
-landscape. \emph{American Journal of Epidemiology}.
-2022;191(10):1792-1802.
-doi:\href{https://doi.org/10.1093/aje/kwac080}{10.1093/aje/kwac080}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
-\CSLLeftMargin{26. }%
-\CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
-social vulnerability index to assess {COVID}-19 vaccine uptake in
-louisiana. \emph{{GeoJournal}}. Published online December 2022.
-doi:\href{https://doi.org/10.1007/s10708-022-10802-5}{10.1007/s10708-022-10802-5}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
-\CSLLeftMargin{27. }%
-\CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
-future spatial patterns in {COVID}-19 booster vaccine uptake. Published
-online September 2022.
-doi:\href{https://doi.org/10.1101/2022.08.30.22279415}{10.1101/2022.08.30.22279415}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
-\CSLLeftMargin{28. }%
-\CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
-trajectories. \emph{Socius: Sociological Research for a Dynamic World}.
-2021;7:237802312110529.
-doi:\href{https://doi.org/10.1177/23780231211052946}{10.1177/23780231211052946}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
-\CSLLeftMargin{29. }%
-\CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
-Social inequalities in {COVID}-19 vaccine acceptance and uptake for
-children and adolescents in montreal, canada. \emph{Vaccine}.
-2021;39(49):7140-7145.
-doi:\href{https://doi.org/10.1016/j.vaccine.2021.10.077}{10.1016/j.vaccine.2021.10.077}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
-\CSLLeftMargin{30. }%
-\CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
-integrated care: A complex-adaptive systems perspective.
-\emph{International Journal of Integrated Care}. 2012;12(5).
-doi:\href{https://doi.org/10.5334/ijic.843}{10.5334/ijic.843}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
-\CSLLeftMargin{31. }%
-\CSLRightInline{Dong L, Sahu R, Black R. Governance in the
-transformational journey toward integrated healthcare: The case of
-ontario. \emph{Journal of Information Technology Teaching Cases}.
-Published online December 2022:204388692211473.
-doi:\href{https://doi.org/10.1177/20438869221147313}{10.1177/20438869221147313}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{32. }%
-\CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
-Corlin L. Disparities in national and state estimates of {COVID}-19
-vaccination receipt and intent to vaccinate by race/ethnicity, income,
-and age group among adults~\(\geq\)~18~years, united states.
-\emph{Vaccine}. 2022;40(1):107-113.
-doi:\href{https://doi.org/10.1016/j.vaccine.2021.11.040}{10.1016/j.vaccine.2021.11.040}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{33. }%
-\CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
-Zikmund-Fisher BJ. Vaccine hesitancy and rejection of a vaccine for the
-novel coronavirus in the united states. \emph{Frontiers in Immunology}.
-2021;12.
-doi:\href{https://doi.org/10.3389/fimmu.2021.558270}{10.3389/fimmu.2021.558270}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{34. }%
-\CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
-and factors related to {COVID}-19 vaccine hesitancy and unwillingness in
-canada: A systematic review and meta-analysis. \emph{Journal of Medical
-Virology}. 2022;95(1).
-doi:\href{https://doi.org/10.1002/jmv.28156}{10.1002/jmv.28156}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{35. }%
-\CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
-a sampled frequency table when the expected marginal totals are known.
-\emph{The Annals of Mathematical Statistics}. 1940;11(4):427-444.
-doi:\href{https://doi.org/10.1214/aoms/1177731829}{10.1214/aoms/1177731829}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{36. }%
-\CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
-2011.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{37. }%
-\CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
-{tidyverse}. \emph{Journal of Open Source Software}. 2019;4(43):1686.
-doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{38. }%
-\CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
-Markdown Publishing System}.; 2022.
-\url{https://CRAN.R-project.org/package=quarto}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{39. }%
-\CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
-in {R}. \emph{Journal of Statistical Software}. 2022;103(1):1-23.
-doi:\href{https://doi.org/10.18637/jss.v103.i01}{10.18637/jss.v103.i01}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{40. }%
-\CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
-Reproducible summary tables with the gtsummary package. \emph{{The R
-Journal}}. 2021;13:570-580.
-doi:\href{https://doi.org/10.32614/RJ-2021-053}{10.32614/RJ-2021-053}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
-{COVID}-19 vaccine uptake in southeastern ontario, canada: Monitoring
-and addressing health inequities. \emph{Journal of Public Health
-Management and Practice}. 2022;28(6):615-623.
-doi:\href{https://doi.org/10.1097/phh.0000000000001565}{10.1097/phh.0000000000001565}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{42. }%
-\CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
-Structural barriers and racial and ethnic inequities in {COVID}-19
-vaccination. \emph{International Journal of Environmental Research and
-Public Health}. 2021;18(18):9904.
-doi:\href{https://doi.org/10.3390/ijerph18189904}{10.3390/ijerph18189904}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{43. }%
-\CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
-and other infection exposure according to race and ethnicity.
-\emph{American Journal of Industrial Medicine}. 2020;63(9):817-820.
-doi:\href{https://doi.org/10.1002/ajim.23145}{10.1002/ajim.23145}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{44. }%
-\CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
-review of {COVID}-19 and vulnerable workers: Intersecting occupational
-and public health issues. \emph{American Journal of Industrial
-Medicine}. 2021;64(7):551-566.
-doi:\href{https://doi.org/10.1002/ajim.23256}{10.1002/ajim.23256}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{45. }%
-\CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
-Strategy for Ontario {COVID}-19 Hotspots and Essential Workers}. Ontario
-{COVID}-19 Science Advisory Table; 2021.
-doi:\href{https://doi.org/10.47326/ocsat.2021.02.26.1.0}{10.47326/ocsat.2021.02.26.1.0}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{46. }%
-\CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
-Health;
-\url{https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf};
-2022.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{47. }%
-\CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
-Rosella LC. Ontario atlas of adult mortality, 1992-2015: Trends in
-{L}ocal {H}ealth {I}ntegration {N}etworks. Published online 2018.
-\url{https://tspace.library.utoronto.ca/handle/1807/82836}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{48. }%
-\CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
-and risk factors for {SARS}-{CoV}-2 infection in a south asian community
-in ontario: A cross-sectional analysis of a prospective cohort study.
-\emph{{CMAJ} Open}. 2022;10(3):E599-E609.
-doi:\href{https://doi.org/10.9778/cmajo.20220031}{10.9778/cmajo.20220031}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{49. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
-2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{50. }%
-\CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
-compulsion: The case of {COVID}-19 vaccination in canada.
-\emph{Vaccine}. 2022;40(29):3923-3926.
-doi:\href{https://doi.org/10.1016/j.vaccine.2022.05.053}{10.1016/j.vaccine.2022.05.053}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{51. }%
-\CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
-of canada {COVID}-19 report: Enhancing {COVID}-19 vaccine acceptance in
-canada. Blais JM, ed. \emph{{FACETS}}. 2021;6:1184-1246.
-doi:\href{https://doi.org/10.1139/facets-2021-0037}{10.1139/facets-2021-0037}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{52. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=maps}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{53. }%
-\CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
-vaccination status by self-report versus source documentation: Impact on
-measuring {COVID}-19 vaccine effectiveness. \emph{Influenza and Other
-Respiratory Viruses}. 2022;16(6):1101-1111.
-doi:\href{https://doi.org/10.1111/irv.13023}{10.1111/irv.13023}}
-
-\end{CSLReferences}
-
-\end{document}
+The results in this study are based on self-reported data, where bias might be present.
+However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid assessment of vaccination in the province.
+Finally, this study focused on first-dose vaccination status  within a relatively short time window, and therefore can only provide a snapshot of the societal dynamics behind the pandemic.
+Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.

### 2023-05-08 manuscript_v5/appendix.tex
new directory for version, tidy code

--- 2023-04-17 manuscript_v3/appendix.tex
@@
 \usepackage{xcolor}
+\usepackage{placeins}
 \KOMAoption{captions}{tableheading}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, breakable, frame hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{vaccination-rates-by-age}{%
-\section{Vaccination rates by Age}\label{vaccination-rates-by-age}}
+\hypertarget{complete-model-estimates}{%
+\section{Complete Model Estimates}\label{complete-model-estimates}}
 
-The table below provides the percentages of vaccination from the survey
-data.
+The figures below present the complete estimates from the regression
+models.
 
-\hypertarget{tbl-vaccination}{}
-\begin{longtable}{lcc}
-\caption{\label{tbl-vaccination}Vaccination Status by Age from the Survey Data }\tabularnewline
+\begin{figure}
 
-\toprule
-\textbf{Characteristic} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,591\\
-\midrule
-\endfirsthead
-\multicolumn{3}{@{}l}{\textit{(continued)}}\\
-\toprule
-\textbf{Characteristic} & \textbf{no}, N = 958 & \textbf{yes}, N = 2,591\\
-\midrule
-\endhead
+\includegraphics{appendix_files/figure-pdf/fig-model-uncorr-appendix-1.pdf} \hfill{}
 
-\endfoot
-\bottomrule
-\endlastfoot
-\textbf{Age} &  & \\
-\hspace{1em}16 & 11 (21\%) & 41 (79\%)\\
-\hspace{1em}17 & 15 (25\%) & 45 (75\%)\\
-\hspace{1em}18 & 53 (36\%) & 93 (64\%)\\
-\hspace{1em}19 & 33 (31\%) & 74 (69\%)\\
-\hspace{1em}20 & 18 (17\%) & 85 (83\%)\\
-\hspace{1em}21 & 30 (25\%) & 92 (75\%)\\
-\hspace{1em}22 & 19 (23\%) & 62 (77\%)\\
-\hspace{1em}23 & 21 (24\%) & 67 (76\%)\\
-\hspace{1em}24 & 23 (29\%) & 55 (71\%)\\
-\hspace{1em}25 & 25 (24\%) & 80 (76\%)\\
-\hspace{1em}26 & 23 (31\%) & 52 (69\%)\\
-\hspace{1em}27 & 16 (25\%) & 49 (75\%)\\
-\hspace{1em}28 & 13 (20\%) & 52 (80\%)\\
-\hspace{1em}29 & 13 (26\%) & 37 (74\%)\\
-\hspace{1em}30 & 12 (21\%) & 44 (79\%)\\
-\hspace{1em}31 & 17 (38\%) & 28 (62\%)\\
-\hspace{1em}32 & 17 (27\%) & 47 (73\%)\\
-\hspace{1em}33 & 17 (26\%) & 48 (74\%)\\
-\hspace{1em}34 & 33 (35\%) & 60 (65\%)\\
-\hspace{1em}35 & 31 (34\%) & 59 (66\%)\\
-\hspace{1em}36 & 16 (25\%) & 49 (75\%)\\
-\hspace{1em}37 & 15 (30\%) & 35 (70\%)\\
-\hspace{1em}38 & 10 (28\%) & 26 (72\%)\\
-\hspace{1em}39 & 5 (16\%) & 26 (84\%)\\
-\hspace{1em}40 & 7 (16\%) & 36 (84\%)\\
-\hspace{1em}41 & 5 (14\%) & 31 (86\%)\\
-\hspace{1em}42 & 7 (19\%) & 30 (81\%)\\
-\hspace{1em}43 & 9 (17\%) & 45 (83\%)\\
-\hspace{1em}44 & 24 (33\%) & 48 (67\%)\\
-\hspace{1em}45 & 23 (30\%) & 53 (70\%)\\
-\hspace{1em}46 & 13 (24\%) & 41 (76\%)\\
-\hspace{1em}47 & 18 (37\%) & 31 (63\%)\\
-\hspace{1em}48 & 10 (33\%) & 20 (67\%)\\
-\hspace{1em}49 & 8 (28\%) & 21 (72\%)\\
-\hspace{1em}50 & 7 (16\%) & 38 (84\%)\\
-\hspace{1em}51 & 7 (23\%) & 23 (77\%)\\
-\hspace{1em}52 & 9 (20\%) & 36 (80\%)\\
-\hspace{1em}53 & 9 (22\%) & 31 (78\%)\\
-\hspace{1em}54 & 19 (37\%) & 33 (63\%)\\
-\hspace{1em}55 & 12 (23\%) & 41 (77\%)\\
-\hspace{1em}56 & 13 (26\%) & 37 (74\%)\\
-\hspace{1em}57 & 12 (22\%) & 43 (78\%)\\
-\hspace{1em}58 & 6 (14\%) & 37 (86\%)\\
-\hspace{1em}59 & 4 (12\%) & 28 (88\%)\\
-\hspace{1em}60 & 3 (11\%) & 25 (89\%)\\
-\hspace{1em}61 & 1 (4.5\%) & 21 (95\%)\\
-\hspace{1em}62 & 8 (26\%) & 23 (74\%)\\
-\hspace{1em}63 & 5 (19\%) & 21 (81\%)\\
-\hspace{1em}64 & 11 (33\%) & 22 (67\%)\\
-\hspace{1em}65 & 11 (20\%) & 45 (80\%)\\
-\hspace{1em}66 & 11 (28\%) & 28 (72\%)\\
-\hspace{1em}67 & 8 (22\%) & 28 (78\%)\\
-\hspace{1em}68 & 5 (19\%) & 21 (81\%)\\
-\hspace{1em}69 & 17 (39\%) & 27 (61\%)\\
-\hspace{1em}70 & 4 (14\%) & 24 (86\%)\\
-\hspace{1em}71 & 1 (3.8\%) & 25 (96\%)\\
-\hspace{1em}72 & 2 (9.1\%) & 20 (91\%)\\
-\hspace{1em}73 & 3 (12\%) & 22 (88\%)\\
-\hspace{1em}74 & 13 (31\%) & 29 (69\%)\\
-\hspace{1em}98 & 147 (42\%) & 201 (58\%)\\*
-\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
-\end{longtable}
+\caption{\label{fig-model-uncorr-appendix}Full coefficient estimates and
+confidence intervals for the uncorrected model.}
+
+\end{figure}
+
+\FloatBarrier
+
+\begin{figure}
+
+\includegraphics{appendix_files/figure-pdf/fig-model-corr-appendix-1.pdf} \hfill{}
+
+\caption{\label{fig-model-corr-appendix}Full coefficient estimates and
+confidence intervals for the corrected model.}
+
+\end{figure}
+
+\FloatBarrier
 
@@
 The plots below show the interactions for Race and Income and Race and
-Health Region based on the regression model presented in the main paper.
+Health Region based on the corrected regression model presented in the
+main paper.
 

### 2023-05-08 manuscript_v5/main.qmd
new directory for version, tidy code

--- 2023-04-24 manuscript_v4/main.qmd
@@
 library(tidyverse)
+library(naniar)
 library(lme4)

### 2023-05-08 manuscript_v5/main.tex
new directory for version, tidy code

--- 2023-04-28 manuscript_v4/main.tex
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, enhanced, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-Preliminary analyses of the data included the removal of outliers, of
-observations where respondents did not provide answers in all the
-covariates of interest, matching the city of each observations with its
-corresponding LHIN and Health Region, and removing observations from
-areas with low representation (107 observations corresponding to the
-North West and North East Health Regions). After all the preliminary
-analyses indicated above, the total number of observations used for
-analysis was 3,549 which included the East, Central, Toronto, and West
-Health Regions covering the period between October 1st,2021 and December
-12th, 2021. The original dataset, clean dataset, and details on the data
-cleaning process are described in detail in the
+Preliminary analyses of the data included the removal of outliers
+(\textbf{should we still do this? it's only 19 observations with income
+\textgreater110k and household of 1, but we are not even using such
+income bracket in the analysis because we re-grouped the data, and the
+household size variable has 90\% missing rate}), of observations where
+respondents did not provide answers in all the covariates of interest,
+matching the city of each observations with its corresponding LHIN and
+Health Region, and removing observations from areas with low
+representation (107 observations corresponding to the North West and
+North East Health Regions). After all the preliminary analyses indicated
+above, the total number of observations used for analysis was 3,549
+which included the East, Central, Toronto, and West Health Regions
+covering the period between October 1st,2021 and December 12th, 2021.
+The original dataset, clean dataset, and details on the data cleaning
+process are described in detail in the
 \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
@@
 
-Comparisons between the variables of the survey were performed using the
-\(\chi^2\) test. We used a logistic regression model to examine the
-impact of the Health Regions in vaccination rates while considering the
-socio-economic factors and and months covered by the survey
+We used a logistic regression model to examine the impact of the Health
+Regions in vaccination rates while considering the socio-economic
+factors and and months covered by the survey
 (Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
@@
 population totals; and fitted the regression model to the uncorrected
-and corrected data. The reference groups in the model were: 16-34 years
-(age group), October(Month), CAD 60,000 and above (income),
-White/Caucasian (Race), Toronto (Health Region). Details regarding the
-correction can be found in the Appendix. All analyses were conducted in
-R 4.2.2 using the packages
+and corrected data. Details regarding the correction can be found in the
+Appendix. All analyses were conducted in R 4.2.2 using the packages
 \texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
@@
 from the Fields COVID-19 survey used for analysis with regard to
-vaccination status. A total of 3,549 observations were used for
-analysis, from which 27\% (958) corresponded to individuals that
-reported not having received the first dose of the vaccine. Respondents
-between 16 and 34 years of age, and those that lived within the Health
-Region of Toronto had the highest number of observations in the data
-(42.8\% and 37.3\% of the total, respectively). Moreover, individuals
-who identified as White or Caucasian reported the highest vaccination
-uptake (82\%), in contrast to respondents who identified as which had
-had the lowest vaccination uptake rate (60\%) across all groups. From
-all the covariates, significant associations were identified between
-income, race, and vaccination rate.
+vaccination status. A total of \textbf{6,236} observations were used for
+analysis, from which 24\% (1,547) corresponded to individuals that
+reported not having received the first dose of the vaccine. Across the
+socio-demographic characteristics collected by the survey, higher
+vaccination uptake was reported by individuals in the highest income
+bracket considered, those between 16 and 34 years of age, individuals
+that lived in the East Health Region, and individuals that identified as
+White or Caucasian. During the months covered by the survey, the highest
+vaccination uptake was observed in January of 2022.
 
@@
 \toprule
-\textbf{Variable} & \textbf{no}, N = 958\textsuperscript{1} & \textbf{yes}, N = 2,591\textsuperscript{1} & \textbf{p-value}\textsuperscript{2} \\ 
+\textbf{Variable} & \textbf{no}, N = 1,547\textsuperscript{1} & \textbf{yes}, N = 4,689\textsuperscript{1} & \textbf{p-value}\textsuperscript{2} \\ 
 \midrule
 Income (CAD) &  &  & <0.001 \\ 
-    60000 and above & 305 (23\%) & 1,048 (77\%) &  \\ 
-    25000-59999 & 253 (28\%) & 636 (72\%) &  \\ 
-    under 25000 & 400 (31\%) & 907 (69\%) &  \\ 
-Age Group &  &  & 0.7 \\ 
-    16-34 & 409 (27\%) & 1,111 (73\%) &  \\ 
-    35-54 & 252 (26\%) & 712 (74\%) &  \\ 
-    55 and over & 297 (28\%) & 768 (72\%) &  \\ 
-Health Region &  &  & 0.14 \\ 
-    Toronto & 371 (28\%) & 952 (72\%) &  \\ 
-    Central & 224 (28\%) & 580 (72\%) &  \\ 
-    East & 135 (23\%) & 448 (77\%) &  \\ 
-    West & 228 (27\%) & 611 (73\%) &  \\ 
-Month &  &  & 0.4 \\ 
+    60000 and above & 542 (21\%) & 1,996 (79\%) &  \\ 
+    25000-59999 & 347 (25\%) & 1,046 (75\%) &  \\ 
+    under 25000 & 658 (29\%) & 1,647 (71\%) &  \\ 
+Age Group &  &  & 0.002 \\ 
+    16-34 & 645 (23\%) & 2,117 (77\%) &  \\ 
+    35-54 & 411 (24\%) & 1,305 (76\%) &  \\ 
+    55 and over & 491 (28\%) & 1,267 (72\%) &  \\ 
+Health Region &  &  & 0.3 \\ 
+    Toronto & 593 (26\%) & 1,709 (74\%) &  \\ 
+    Central & 372 (26\%) & 1,083 (74\%) &  \\ 
+    East & 236 (23\%) & 783 (77\%) &  \\ 
+    West & 346 (24\%) & 1,114 (76\%) &  \\ 
+Month &  &  & <0.001 \\ 
     October & 469 (27\%) & 1,263 (73\%) &  \\ 
     November & 376 (28\%) & 980 (72\%) &  \\ 
-    December & 113 (25\%) & 348 (75\%) &  \\ 
+    December & 181 (24\%) & 565 (76\%) &  \\ 
+    January & 521 (22\%) & 1,881 (78\%) &  \\ 
 Race &  &  & <0.001 \\ 
-    White/Caucasian & 233 (18\%) & 1,079 (82\%) &  \\ 
-    Arab/Middle Eastern & 76 (36\%) & 138 (64\%) &  \\ 
-    Black & 114 (38\%) & 184 (62\%) &  \\ 
-    East Asian/Pacific Islander & 69 (23\%) & 234 (77\%) &  \\ 
-    Indigenous & 76 (40\%) & 115 (60\%) &  \\ 
-    Latin American & 69 (38\%) & 111 (62\%) &  \\ 
-    Mixed & 105 (34\%) & 205 (66\%) &  \\ 
-    Other\textsuperscript{3} & 128 (35\%) & 239 (65\%) &  \\ 
-    South Asian & 88 (24\%) & 286 (76\%) &  \\ 
+    White/Caucasian & 354 (16\%) & 1,871 (84\%) &  \\ 
+    Arab/Middle Eastern & 111 (34\%) & 220 (66\%) &  \\ 
+    Black & 159 (34\%) & 303 (66\%) &  \\ 
+    East Asian/Pacific Islander & 94 (19\%) & 404 (81\%) &  \\ 
+    Indigenous & 112 (37\%) & 194 (63\%) &  \\ 
+    Latin American & 99 (34\%) & 195 (66\%) &  \\ 
+    Mixed & 177 (30\%) & 411 (70\%) &  \\ 
+    Other\textsuperscript{3} & 315 (34\%) & 606 (66\%) &  \\ 
+    South Asian & 126 (21\%) & 485 (79\%) &  \\ 
 \bottomrule

### 2023-05-08 code/corr_model.R

@@ -1,0 +1,41 @@
+## Script to create plot for corrected model
+
+mplot2<-modelplot(m2,draw=FALSE,coef_omit = "Intercept",exponentiate = TRUE) %>%
+  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
+
+#add column by factor 
+mplot2$factor<-c(rep("Age",2),
+                 rep("Month",3),
+                 rep("Income",2),
+                 rep("Race/Ethnicity",8),
+                 rep("Health Region",3),
+                 rep("Income and\nRace/Ethnicity",16),
+                 rep("Race/Ethnicity\nand\nHealth Region",24)
+)
+
+#order factors for plot
+#mplot2$term_ord <- factor(mplot2$term,ordered=TRUE,levels=mplot2$term)
+
+mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
+
+# remove non-significant interaction factors
+
+test<-mplot2%>%
+  filter(str_detect(term,':') & p.value>0.05)
+
+mplot2<-anti_join(mplot2,test)
+
+mplot2<-mplot2%>%
+  mutate(term=str_replace_all(term,':'," and "))
+
+
+corr_mod<-mplot2 %>% 
+  ggplot(aes(y=term))+
+  geom_vline(xintercept = 1, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate),size=2)+
+  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
+  scale_y_discrete(limits=rev)+
+  xlab("")+
+  ylab("")+
+  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y")+
+  theme_classic()

### 2023-05-08 code/data_preparation_May_08.R

@@ -1,13 +1,5 @@
-library(here)
-library(tidyverse)
-library(naniar)
-library(survey)
-library(gtsummary)
-library(lubridate)
-library(modelsummary)
-
-# for reference, the dataset used in the last version the manuscript had 3,549 obs.
-
+## Script to clean and prepare the raw dataset for analysis
+## Upated on May 8, 2023
 
 rawdata<-read.csv(here("data","Fields_data.csv"),na.strings = c(""))
 #clean_data <-read.csv(here("data","clean_dataset.csv"))
@@ -45,7 +37,7 @@ clean_data<-rawdata %>%
 miss_var_summary(clean_data,show_pct=TRUE) %>%
   kableExtra::kbl() %>%
   kableExtra::kable_styling()
-  
+
 # remove cities entered as "None"
 
 clean_data<-clean_data %>%
@@ -55,7 +47,7 @@ clean_data<-clean_data %>%
 
 clean_data <- clean_data %>%
   rename(date=time_date_code,
-    first_dose=q02_first_dose,
+         first_dose=q02_first_dose,
          income=q09_hh_income,
          race=q16_race,
   )%>%
@@ -314,136 +306,3 @@ clean_data<- within(clean_data, Race <- relevel(Race, ref = "White/Caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
 clean_data<- within(clean_data, income_ord <- relevel(income_ord, ref = "60000 and above"))
 clean_data<- within(clean_data, Month <- relevel(Month, ref = "October"))
-
-clean_data %>%
-  tbl_summary(percent="row",
-              by= first_dose,
-              include=c(income_ord,
-                        Age_group_ord,
-                        Health_Region,
-                        Month,
-                        Race),
-              label=list(Health_Region ~ "Health Region",
-                         Age_group_ord ~ "Age Group",
-                         income_ord~"Income (CAD)")
-  ) %>% add_p(test = everything () ~ "chisq.test")%>%
-  bold_labels() %>%
-  modify_table_styling(
-    columns=label,
-    rows=label=="Other",
-    footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
-  )
-
-
-# raking
-#Each of these dataframes provides the population totals for the different variables and their categories
-
-
-Ages<-data.frame(Age_group_ord=c("16-34",
-                                 "35-54",
[…]

### 2023-05-08 code/uncorr_model.R

@@ -1,0 +1,51 @@
+## Script to create plot from model fitted to uncorrected data
+
+mplot1<-modelplot(m0,draw=FALSE,
+                  coef_omit = "Intercept", exponentiate = TRUE) %>%
+  mutate(term=str_remove_all(term,"Race|Age_group_ord|Month|Health_Region|income_ord"))
+
+#add column by factor 
+mplot1$factor<-c(rep("Age",2),
+                 rep("Month",3),
+                 rep("Income",2),
+                 rep("Race/Ethnicity",8),
+                 rep("Health Region",3),
+                 rep("Income and\nRace/Ethnicity",16),
+                 rep("Race/Ethnicity\nand\nHealth Region",24)
+)
+
+#order factors for plot
+#mplot1$term_ord <- factor(mplot1$term,ordered=TRUE,levels=mplot1$term)
+
+mplot1$factor <- factor(mplot1$factor,ordered=TRUE,levels=c("Age",
+                                                            "Month",
+                                                            "Income",
+                                                            "Race/Ethnicity",
+                                                            "Health Region",
+                                                            "Income and\nRace/Ethnicity",
+                                                            "Race/Ethnicity\nand\nHealth Region"))
+
+# identify non-significant interaction factors
+
+test<-mplot1%>%
+  filter(str_detect(term,':') & p.value>0.05)
+
+#remove non-significant interactions from dataframe
+mplot1<-anti_join(mplot1,test)
+
+# fix legend for plotting
+mplot1<-mplot1%>%
+  mutate(term=str_replace_all(term,':'," and "))
+
+#make plot
+uncorr_mod<-mplot1 %>% 
+  ggplot(aes(y=term))+
+  geom_vline(xintercept = 1, linetype="dashed",color="red") +
+  geom_point(aes(x=estimate),size=2)+
+  geom_linerange(aes(xmin=conf.low,xmax=conf.high))+
+  scale_y_discrete(limits=rev)+
+  xlab("")+
+  ylab("")+
+  facet_wrap(~factor,ncol=1,strip.position = "right",scales = "free_y",labeller = label_wrap_gen(multi_line = TRUE))+
+  theme_classic()
+

### 2023-05-08 data_cleaning/data_cleaning_May_08.R

@@ -1,11 +1,3 @@
-library(here)
-library(tidyverse)
-library(naniar)
-library(survey)
-library(gtsummary)
-library(lubridate)
-library(modelsummary)
-
 # for reference, the dataset used in the last version the manuscript had 3,549 obs.
 
 
@@ -45,7 +37,7 @@ clean_data<-rawdata %>%
 miss_var_summary(clean_data,show_pct=TRUE) %>%
   kableExtra::kbl() %>%
   kableExtra::kable_styling()
-  
+
 # remove cities entered as "None"
 
 clean_data<-clean_data %>%
@@ -55,7 +47,7 @@ clean_data<-clean_data %>%
 
 clean_data <- clean_data %>%
   rename(date=time_date_code,
-    first_dose=q02_first_dose,
+         first_dose=q02_first_dose,
          income=q09_hh_income,
          race=q16_race,
   )%>%
@@ -275,8 +267,6 @@ clean_data$Month <- as.factor(clean_data$Month)
 clean_data <- clean_data%>%
   subset(Month!="September")
 
-clean_data$Month <- fct_relevel(clean_data$Month,levels=c("October","November","December","January"))
-
 
 # Remove certain Health REgions due to low number of obs: only 254 in North East
 # and North West Regions at this poing, in contrast with about a thousand from each 
@@ -313,137 +303,4 @@ clean_data<- within(clean_data, Age_group_ord <- relevel(Age_group_ord, ref = "1
 clean_data<- within(clean_data, Race <- relevel(Race, ref = "White/Caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
 clean_data<- within(clean_data, income_ord <- relevel(income_ord, ref = "60000 and above"))
-clean_data<- within(clean_data, Month <- relevel(Month, ref = "October"))
-
-clean_data %>%
-  tbl_summary(percent="row",
-              by= first_dose,
-              include=c(income_ord,
-                        Age_group_ord,
-                        Health_Region,
-                        Month,
-                        Race),
-              label=list(Health_Region ~ "Health Region",
-                         Age_group_ord ~ "Age Group",
-                         income_ord~"Income (CAD)")
-  ) %>% add_p(test = everything () ~ "chisq.test")%>%
-  bold_labels() %>%
-  modify_table_styling(
-    columns=label,
-    rows=label=="Other",
[…]

### 2023-05-09 manuscript_v5/03-results.qmd
updated

@@
 @tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status.
-A total of **6,236** observations were used for analysis, from which 24\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
-Across the socio-demographic characteristics collected by the survey, higher vaccination uptake was reported by individuals in the highest income bracket considered, those between 16 and 34 years of age, individuals that lived in the East Health Region, and individuals that identified as White or Caucasian.
-During the months covered by the survey, the highest vaccination uptake was observed in January of 2022.
+A total of **6,236** observations were used for analysis, from which 24.8\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
+Vaccination rates ranged between 71-79\% across household income brackets, age groups, Health Regions, and the months considered in the survey.
+However, the highest vaccination rates in each category were reported by individuals in the highest income bracket (79\%), those between 16 and 34 years of age (77\%), individuals that lived in the East Health Region (77\%), and  during January of 2022 (78\%).
+Differences were higher between racial/ethnic groups, where the higher vaccination rate was reported by White/Caucasian individuals (84\%), against vaccination rates between 63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and those that reported belonging to "Other" racial groups, which included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere.
 
@@
 
-```{r,model-uncorr-fig}
+```{r,models}
 #| message: false
@@
 #| fig-align: left
-#| fig-width: 7
+#| fig-width: 12
 #| fig-height: 7
@@
 
-# call script that creates figure for model
+# call script that creates figure for uncorrected model
 
@@
 
-#call figure created in the script above
-uncorr_mod
+#call script that creates figure for corrected model
+source(here("code","corr_model.R"))
+
+fig_mod<-uncorr_mod+corr_mod+plot_layout(guides = "collect")+plot_annotation(tag_levels = "A")
+
+#print(fig_mod, vp=viewport(angle=90,height = 0.5,clip="off"))
+fig_mod
 

### 2023-05-09 manuscript_v5/main.qmd
updated

--- 2023-04-24 manuscript_v4/main.qmd
@@
   pdf:
-    include-in-header: latex/preamble.sty
+    include-in-header:
+      - text: |
+          \usepackage{pdflscape}
+      - file: latex/preamble.sty
     #mainfont: "Palatino Linotype"
@@
 library(tidyverse)
+library(naniar)
 library(lme4)
 library(lubridate)
+library(grid)
 library(kableExtra)

### 2023-05-09 manuscript_v5/main.tex
updated

@@
 
-%line numbers
-%\usepackage{mathpazo}
-\usepackage{lineno}
-\linenumbers
-
 \usepackage{booktabs}
@@
 \usepackage{caption}
+%line numbers
+%\usepackage{mathpazo}
+%\usepackage{rotating}
+\usepackage{lineno}
+\linenumbers
+
+\usepackage{pdflscape}
 \KOMAoption{captions}{tableheading}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi
 
@@
 vaccination status. A total of \textbf{6,236} observations were used for
-analysis, from which 24\% (1,547) corresponded to individuals that
-reported not having received the first dose of the vaccine. Across the
-socio-demographic characteristics collected by the survey, higher
-vaccination uptake was reported by individuals in the highest income
-bracket considered, those between 16 and 34 years of age, individuals
-that lived in the East Health Region, and individuals that identified as
-White or Caucasian. During the months covered by the survey, the highest
-vaccination uptake was observed in January of 2022.
+analysis, from which 24.8\% (1,547) corresponded to individuals that
+reported not having received the first dose of the vaccine. Vaccination
+rates ranged between 71-79\% across household income brackets, age
+groups, Health Regions, and the months considered in the survey.
+However, the highest vaccination rates in each category were reported by
+individuals in the highest income bracket (79\%), those between 16 and
+34 years of age (77\%), individuals that lived in the East Health Region
+(77\%), and during January of 2022 (78\%). Differences were higher
+between racial/ethnic groups, where the higher vaccination rate was
+reported by White/Caucasian individuals (84\%), against vaccination
+rates between 63-66\% reported in the case of Arab/Middle Eastern,
+Black, Indigenous, Latin American individuals, and those that reported
+belonging to ``Other'' racial groups, which included Southeast Asian,
+Filipino, West Asian, and minorities not identified elsewhere.
 

### 2023-05-10 manuscript_v5/03-results.qmd
updating results

@@
 
-@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis with regard to vaccination status.
-A total of **6,236** observations were used for analysis, from which 24.8\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
+@tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis.
+The sample contained **6,236** observations, from which 24.8\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
 Vaccination rates ranged between 71-79\% across household income brackets, age groups, Health Regions, and the months considered in the survey.
@@
 
-@fig-model-uncorr shows the estimates from the logistic regression model of vaccination status for the uncorrected data, whereas the estimated obtained from the corrected data appear in @fig-model-corr.
-The results show significantly lower odds of vaccination in individuals with a low household income and those that identified as part of equity-deserving groups when compared to individuals in high income brackets or that identified as White or Caucasian.
+@fig-model-uncorr presents the estimates (as odd ratios) from the logistic regression models for vaccination status using the socio-demographic factors collected by the survey, and their interactions.
+Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of equity-deserving groups.
+However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, certain estimates were not deemed statistically-significant after the correction, in contrast with the estimates obtained from the uncorrected model.
+Specifically, the corrected model showed no significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region.
+<!-- \newpage -->
+
+<!-- \begin{sidewaysfigure} -->
 
@@
 #| fig-align: left
-#| fig-width: 12
-#| fig-height: 7
+#| fig-width: 15
+#| fig-height: 12
 #| label: fig-model-uncorr
@@
 # call script that creates figure for uncorrected model
-
 source(here("code","uncorr_model.R"))
@@
 
-fig_mod<-uncorr_mod+corr_mod+plot_layout(guides = "collect")+plot_annotation(tag_levels = "A")
+fig_mod<-uncorr_mod+corr_mod+plot_layout(guides = "collect")+plot_annotation(tag_levels = "A")&
+    theme(plot.tag = element_text(size=18,face='bold'),
+          #axis.title = element_text(size=11),
+          axis.text = element_text(size=12),
+          strip.text=element_text(size=12))
 
@@
 
-Specifically, significantly lower odds of vaccination were identified for those with a household income under CAD 25,000 (OR=0.37, CI=[0.25,0.56]) and those with an income between CAD 25,000 and 59,999 (OR=0.59, CI=[0.39,0.88]).
-Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, or that belonged to "Other" racial groups, which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and CIs=[0.14,0.68], [0.17,0.60], [0.11,0.66], and [0.12,0.41], respectively).
-Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.54, CI=[1.04,2.29]).
+<!-- \end{sidewaysfigure} -->
+
+However, significantly lower odds of vaccination were identified in the corrected model for those with a household income under CAD 25,000 (OR=0.37, CI=[0.27,0.51]) and those with an income between CAD 25,000 and 59,999 (OR=0.58, CI=[0.42,0.81]).
+Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, Mixed, or that belonged to other racial groups (which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census), had significantly lower odds of vaccination than those in the White/Caucasian group (ORs and CIs=0.28 [0.16,0.51], 0.27 [0.16,0.45], 0.40 [0.21,0.76], 0.53 [0.30,0.92], 0.23 [0.15,0.36]).
+Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent), and in the Central Health Region had significantly higher odds of vaccination than those in the Health Region of Toronto (ORs and CIs=1.59 [1.16,2.19], 1.61 [1.10,2.34]).
 

### 2023-05-10 manuscript_v5/main.qmd
updating results

@@
       - text: |
-          \usepackage{pdflscape}
+          \usepackage{rotfloat}
       - file: latex/preamble.sty
@@
     keep-tex: true
+    link-citations: true #adds links to the citations
 #linestretch: 1
@@
 csl: references/jama.csl
-link-citations: yes #adds links to the citations
+
 ---

### 2023-05-10 manuscript_v5/main.tex
updating results

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
 \fi
@@
 
-\usepackage{pdflscape}
+\usepackage{rotfloat}
 \KOMAoption{captions}{tableheading}
@@
 \makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
 \makeatother
@@
 \@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
 \makeatother
@@
 Ontario across Health Regions and socio-economic strata}}
-\author{Ariel Mundo Ortiz\textsuperscript{1,2} \and Bouchra
-Nasri\textsuperscript{1,2,*}}
+\author{}
 \date{}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, interior hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 Table~\ref{tbl-descriptive-stats} shows the characteristics of the data
-from the Fields COVID-19 survey used for analysis with regard to
-vaccination status. A total of \textbf{6,236} observations were used for
-analysis, from which 24.8\% (1,547) corresponded to individuals that
-reported not having received the first dose of the vaccine. Vaccination
-rates ranged between 71-79\% across household income brackets, age
-groups, Health Regions, and the months considered in the survey.
-However, the highest vaccination rates in each category were reported by
-individuals in the highest income bracket (79\%), those between 16 and
-34 years of age (77\%), individuals that lived in the East Health Region
-(77\%), and during January of 2022 (78\%). Differences were higher
-between racial/ethnic groups, where the higher vaccination rate was
-reported by White/Caucasian individuals (84\%), against vaccination
-rates between 63-66\% reported in the case of Arab/Middle Eastern,
-Black, Indigenous, Latin American individuals, and those that reported
-belonging to ``Other'' racial groups, which included Southeast Asian,
-Filipino, West Asian, and minorities not identified elsewhere.
+from the Fields COVID-19 survey used for analysis. The sample contained
+\textbf{6,236} observations, from which 24.8\% (1,547) corresponded to
+individuals that reported not having received the first dose of the
+vaccine. Vaccination rates ranged between 71-79\% across household
+income brackets, age groups, Health Regions, and the months considered
+in the survey. However, the highest vaccination rates in each category
+were reported by individuals in the highest income bracket (79\%), those
+between 16 and 34 years of age (77\%), individuals that lived in the
+East Health Region (77\%), and during January of 2022 (78\%).
+Differences were higher between racial/ethnic groups, where the higher
+vaccination rate was reported by White/Caucasian individuals (84\%),
+against vaccination rates between 63-66\% reported in the case of
+Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and
+those that reported belonging to ``Other'' racial groups, which included
+Southeast Asian, Filipino, West Asian, and minorities not identified
+elsewhere.
 

### 2023-05-11 manuscript_v5/01-background.qmd
background updated

# Background

As of May of 2023 there have been 765 million confirmed cases of COVID-19 around the world, including 6.8 million deaths [@WHO-Covid]. Although this disease is no longer categorized as a global health emergency by the World Health Organization (WHO) [@rigby2023], there is ongoing concern due to continued transmission, surges in cases and deaths due to new variants [@un2023], and weaknesses in health systems around the world that could be exploited by a novel virus or another public health emergency in the future [@mackey2021].

In particular, a major weakness that has received attention during the pandemic has been related to vaccine uptake. The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic due to the start of vaccination campaigns in certain parts of the world toward the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020]) but inequalities in vaccine uptake made these pharmaceutical interventions ultimately unable to replicate the experience of smallpox, where vaccination on a global scale and was crucial to control this disease [@kayser2021].

Although the problematic on vaccine uptake is a multifaceted issue resulting from a combination of factors, among which are failed public health measures [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021; @tamey2022], and vaccine hesitancy [@nafilyan2021], it is well established that individuals in certain equity-deserving groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages have been disproportionately impacted by it, as indicated by the low vaccine uptake observed in these groups [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].

Reasons given for this problematic have included medical mistrust due to systemic medical racism [@stoler2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].  However, it is important to  also consider that vaccination uptake can be influenced by geographical (spatial) factors. In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], spatial differences in vaccine access and supply, vaccination location availability, and lack of prioritization of areas where vulnerable groups reside [@bogoch2022; @nguyen2021]. Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022].

However, such analyses have been carried mostly in territories outside of Canada, where available studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022]. Therefore, there is a need for studies that explore spatial differences in vaccination within the Canadian territory and that consequently, can help identify disparities that need to be addressed within specific areas in each province.

This is particularly important in the case of Ontario, the most populated province in Canada. Between 2006 and 2019, Ontario provided healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs). However, due to is complexity, bureaucratic burden, and lack of collaborative work this approach resulted in disparities such as excessive expenditures and differences in mortality rates [@tsasis2012;@muratov2018]. In an effort to improve the healthcare system of the province, in late 2019 the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East)  [@dong2022]. Due to the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat. In this regard, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.

Therefore, in this study we hypothesized that there were differences in vaccination uptake between the different Health Regions of Ontario between October of 2021 and January of 2022. By including socio-economic factors in our analysis, we aimed at identifying in which groups these differences were significant in order to provide an assessment of the current state of healthcare access in Ontario.

### 2023-05-11 manuscript_v5/03-results.qmd
background updated

@@
 Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of equity-deserving groups.
-However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, certain estimates were not deemed statistically-significant after the correction, in contrast with the estimates obtained from the uncorrected model.
-Specifically, the corrected model showed no significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region.
+However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, certain estimates were not deemed to be statistically-significant after the correction, in contrast to the estimates from the uncorrected model.
+Specifically, the corrected model showed no significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-model-uncorr,B).
 <!-- \newpage -->
@@
 However, significantly lower odds of vaccination were identified in the corrected model for those with a household income under CAD 25,000 (OR=0.37, CI=[0.27,0.51]) and those with an income between CAD 25,000 and 59,999 (OR=0.58, CI=[0.42,0.81]).
-Additionally, individuals who identified as Arab/Middle Eastern, Black, or Latin American, Mixed, or that belonged to other racial groups (which included the Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere groups according to the Census), had significantly lower odds of vaccination than those in the White/Caucasian group (ORs and CIs=0.28 [0.16,0.51], 0.27 [0.16,0.45], 0.40 [0.21,0.76], 0.53 [0.30,0.92], 0.23 [0.15,0.36]).
-Regarding Health Regions, individuals that reported living in the West Health Region (which includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent), and in the Central Health Region had significantly higher odds of vaccination than those in the Health Region of Toronto (ORs and CIs=1.59 [1.16,2.19], 1.61 [1.10,2.34]).
+Additionally, individuals who identified as Arab/Middle Eastern, Black, Latin American, of mixed background, or that belonged to other racial groups (a category that included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere), had significantly lower odds of vaccination than those in the White/Caucasian group (ORs and CIs=0.28 [0.16,0.51], 0.27 [0.16,0.45], 0.40 [0.21,0.76], 0.53 [0.30,0.92], 0.23 [0.15,0.36]).
+Additionally, individuals that reported living in the Central and West Health Regions had higher odds of vaccination than those in the Health Region of Toronto (ORs and CIs=1.61 [1.10,2.34], and 1.59 [1.16,2.19], respectively).
 
 Interestingly, individuals in equity-deserving groups with a household income below CAD 25,000 had higher odds of vaccination (when compared to those with a household income above CAD 60,000).
-This held true in the case of Arab/Middle Eastern (OR=3.08, CI=[1.27,7.47]), Black (OR=3.15, CI=[1.43,6.92]), and Latin American (OR=2.81, CI=[1.04,7.59]) individuals, as well as respondents who belonged to "Other" minority groups  (OR=4.63, CI=[2.34,9.13]), who also had higher odds of vaccination in the CAD 25,000-59,999 income bracket (OR=6.96, CI=[2.67,18.16]).
-Finally, significantly lower odds of vaccination were identified (when compared to the Toronto Health Region) for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44, CI=[0.19,0.99]), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41, and CIs=[0.18,0.92] and [0.18,0.95], respectively).
+This held true in the case of Arab/Middle Eastern (OR=34, CI=[1.70,6.79]), Black individuals (OR=3.81, CI=[2.05, 7.09]), or those in other racial or ethnic groups (OR=3.19, CI=[2.00,5.09]).
+Additionally, individuals with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern and other racial ethnic groups had higher odds of vaccination (ORs and CIs=6.96 [2.67,18.16], and 3.5 [1.85,6.62]).
+
+Finally, significantly lower odds of vaccination were identified (when compared to the Toronto Health Region) for Black individuals in the Central Health Region (OR=0.39, CI=[0.2,0.75]), Arab/Middle Eastern individuals in the East Health Region (OR=0.41 [0.17, 0.98]), and in the Indigenous and mixed groups in the West Health Region (ORs and CIs=[0.31 [0.14, 0.7] and 0.38 [0.19, 0.76], respectively).
 

### 2023-05-11 manuscript_v5/04-discussion.qmd
background updated

@@
 # Discussion
+
+West Health Region includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+The Central Health Region includes the regions of Peel and Halton, the regional municipality of York, the district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of Etobicoke.
+
+had significantly higher odds of vaccination than those in the Health Region of Toronto
 

### 2023-05-11 manuscript_v5/appendix.qmd
background updated

--- 2023-04-28 manuscript_v4/appendix.qmd
@@
 library(lubridate)
+library(naniar)
 library(gtsummary)
@@
 #call scripts for data preparation and regression
-source(here::here("code","data_preparation.R"))
+source(here::here("code","data_preparation_May_08.R"))
 source(here::here("code","raking.R"))
@@
 p2<-clean_data%>%
-  select(Age_group)%>%
-   mutate(Age_group_ord = fct_relevel(Age_group,
+  select(age_group)%>%
+   mutate(age_group_ord = fct_relevel(age_group,
                                       "16_24",
@@
 
-  ggplot(p2%>%count(Age_group_ord) %>%    # G count number in each group
+  ggplot(p2%>%count(age_group_ord) %>%    # G count number in each group
          mutate(pct=n/sum(n),               # Calculate percent within each group
                 ypos = 200),  # Calculate label positions
-       aes(Age_group_ord, n, fill=Age_group_ord)) +
+       aes(age_group_ord, n, fill=age_group_ord)) +
   geom_bar(stat="identity") +
@@
 clean_data<-clean_data%>%
-    mutate(income_ord = case_when(
+    mutate(income_ord_a = case_when(
                                   income %in% "90000_109999" ~ "over_90000",
@@
                                 ))%>%
-  mutate(income_ord=as.factor(income_ord))
+  mutate(income_ord_a=as.factor(income_ord_a))
 
 p3<-clean_data%>%
-  mutate(income_ord = fct_relevel(income_ord,
+  mutate(income_ord_a = fct_relevel(income_ord_a,
                                   "under_15000",
@@
 
-ggplot(p3%>% count(income_ord)%>%  # count number in each group
+ggplot(p3%>% count(income_ord_a)%>%  # count number in each group
          mutate(pct=n/sum(n),               # Calculate percent within each group
                 ypos = 200),  # Calculate label positions
-       aes(income_ord, n, fill=income_ord)) +
+       aes(income_ord_a, n, fill=income_ord_a)) +
   geom_bar(stat="identity") +
@@
 mplot1$factor<-c(rep("Age",2),
-                 rep("Month",2),
+                 rep("Month",3),
                  rep("Income",2),
@@
 mplot2$factor<-c(rep("Age",2),
-                 rep("Month",2),
+                 rep("Month",3),
                  rep("Income",2),
@@
 
-race_income<-emmip(m2, Race ~ income_ord, CIs=TRUE,plotit = FALSE)
+race_income<-emmip(m2, Race ~ income_ord, CIs=TRUE,plotit = FALSE, type="response")
 
@@
 
-race_hr<-emmip(m2, Race ~ Health_Region, CIs=TRUE,plotit = FALSE)
+race_hr<-emmip(m2, Race ~ Health_Region, CIs=TRUE,plotit = FALSE, type="response")
 
@@
   facet_wrap(~Race)+
-  ylab("Linear predictor")+
+  #ylab("Linear predictor")+
   xlab("Health Region")+

### 2023-05-11 manuscript_v5/appendix.tex
background updated

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{setspace}
@@
 
-\usepackage{amsmath}
 \usepackage{booktabs}
-\usepackage{caption}
 \usepackage{longtable}
@@
 \usepackage{xcolor}
+\usepackage{amsmath}
+\usepackage{caption}
 \usepackage{placeins}
@@
 \subtitle{Appendix}
-\author{Ariel Mundo Ortiz\textsuperscript{1,2} \and Bouchra
-Nasri\textsuperscript{1,2,*}}
+\author{}
 \date{}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 \midrule
-White/Caucasian & 1312 & $37.0\%$ \\ 
-Arab/Middle Eastern & 214 & $6.0\%$ \\ 
-Black & 298 & $8.4\%$ \\ 
-East Asian/Pacific Islander & 303 & $8.5\%$ \\ 
-Indigenous & 191 & $5.4\%$ \\ 
-Latin American & 180 & $5.1\%$ \\ 
-Mixed & 310 & $8.7\%$ \\ 
-Other & 367 & $10.3\%$ \\ 
-South Asian & 374 & $10.5\%$ \\ 
+White/Caucasian & 2225 & $35.7\%$ \\ 
+Arab/Middle Eastern & 331 & $5.3\%$ \\ 
+Black & 462 & $7.4\%$ \\ 
+East Asian/Pacific Islander & 498 & $8.0\%$ \\ 
+Indigenous & 306 & $4.9\%$ \\ 
+Latin American & 294 & $4.7\%$ \\ 
+Mixed & 588 & $9.4\%$ \\ 
+Other & 921 & $14.8\%$ \\ 
+South Asian & 611 & $9.8\%$ \\ 
 \bottomrule
@@
 Race/Ethnicity in Ontario}\tabularnewline
-\toprule()
+\toprule\noalign{}
 Ethnicity/Race & Percentage & Population totals \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 Ethnicity/Race & Percentage & Population totals \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 Arab & 1.6\% & 212782 \\
@@
 \textbf{Total} & 99.9\% & 13431105 \\
-\bottomrule()
 \end{longtable}
@@
 Ontario}\tabularnewline
-\toprule()
+\toprule\noalign{}
 Group age & Percentage & Population Totals \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 Group age & Percentage & Population Totals \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 15-24 & 12.7\% & 1707959 \\
@@
 65 and over & 16.7\% & 2245898 \\
-\bottomrule()
 \end{longtable}
@@
 Ontario}\tabularnewline
-\toprule()
+\toprule\noalign{}
 Household income range (CAD) & Percentage & Population Totals \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 Household income range (CAD) & Percentage & Population Totals \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 \textless{} 15,000 & 5.7\% & 294643 \\
@@
 \textgreater90,000 & 40.3\% & 2083176 \\
-\bottomrule()
 \end{longtable}
@@
 Regions in Ontario}\tabularnewline
-\toprule()
+\toprule\noalign{}
 Health Region & Population Totals \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 Health Region & Population Totals \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 North East & 232299 \\
@@
 Toronto & 1440644 \\
-\bottomrule()
 \end{longtable}
@@
 \toprule
-\textbf{Characteristic} & \textbf{no}, N = 233 & \textbf{yes}, N = 1,079\\
+\textbf{Characteristic} & \textbf{no}, N = 354 & \textbf{yes}, N = 1,871\\
 \midrule
 \textbf{Health\_Region} &  & \\
-\hspace{1em}Toronto & 85 (21\%) & 321 (79\%)\\
-\hspace{1em}West & 69 (17\%) & 329 (83\%)\\
-\hspace{1em}East & 41 (15\%) & 236 (85\%)\\
-\hspace{1em}Central & 38 (16\%) & 193 (84\%)\\
+\hspace{1em}West & 105 (15\%) & 582 (85\%)\\
+\hspace{1em}Toronto & 127 (19\%) & 533 (81\%)\\
+\hspace{1em}East & 66 (14\%) & 396 (86\%)\\
+\hspace{1em}Central & 56 (13\%) & 360 (87\%)\\
 \bottomrule
@@
 \toprule
-\textbf{Characteristic} & \textbf{N = 839}\\
+\textbf{Characteristic} & \textbf{N = 1,460}\\
 \midrule
 \textbf{LHIN} & \\
-\hspace{1em}Hamilton Niagara Haldimand Brant (Hnhb) & 271 (32\%)\\
-\hspace{1em}Waterloo Wellington & 227 (27\%)\\
-\hspace{1em}South West & 187 (22\%)\\
-\hspace{1em}Erie St. Clair & 154 (18\%)\\
+\hspace{1em}Hamilton Niagara Haldimand Brant (Hnhb) & 470 (32\%)\\
+\hspace{1em}Waterloo Wellington & 400 (27\%)\\
+\hspace{1em}South West & 345 (24\%)\\
+\hspace{1em}Erie St. Clair & 245 (17\%)\\
 \bottomrule

### 2023-05-11 manuscript_v5/main.tex
background updated

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, interior hidden, breakable, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, enhanced, boxrule=0pt, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-The vaccines against COVID-19 have been considered a major achievement
-of modern medicine as their rapid development allowed the start of broad
-vaccination campaigns towards the end of 2020 in certain countries, such
-as the US and
-Canada\textsuperscript{\protect\hyperlink{ref-davis2022}{1}--\protect\hyperlink{ref-tanne2020}{3}}.
-This made some believe that vaccines were destined to be a determinant
-factor in a rapid ending of the
-pandemic\textsuperscript{\protect\hyperlink{ref-thelancet2021}{4}}.
-However, although it has been estimated that COVID-19 vaccines have
-prevented around 14 million of deaths
-worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{5}}, their
-implementation has been far from being equal to that of the smallpox and
-polio vaccines, which were implemented on a global scale and that were
-crucial to control these
-diseases\textsuperscript{\protect\hyperlink{ref-kayser2021}{6}}. In
-fact, the rollout of COVID-19 vaccines has faced multiple challenges
-since its inception which ultimately have hampered their use to achieve
-the ultimate goal of global immunity.
-
-This problematic in the rollout of the COVID-19 vaccines is a
-multifaceted issue resulting from, among other things, the development
-of new variants due to inadequate public health
-measures\textsuperscript{\protect\hyperlink{ref-li2021}{7}}, inequality
-in vaccine access between high- and low-income
-countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{8},\protect\hyperlink{ref-tamey2022}{9}},
-vaccine
-hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{10}}, and
-differences in vaccination uptake across different segments of the
-population\textsuperscript{\protect\hyperlink{ref-malik2020}{11}}. In
-particular, it is well established that differences in vaccination
-uptake have been present even in countries that have had ample access to
-vaccines since 2020 (such as the US, the UK, and Canada), where lower
-vaccine uptake has been observed within certain racial groups (i.e.,
-individuals that identify as Black, Asian, or Indigenous), and in
-individuals within low income
-brackets\textsuperscript{\protect\hyperlink{ref-willis2021}{12}--\protect\hyperlink{ref-khubchandani2021}{15}}.
-Reasons given for lower vaccine uptake in these cases have included
-medical mistrust due to systemic medical
-racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{14}}, mistrust
-in vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{12}}, and
+As of May of 2023 there have been 765 million confirmed cases of
+COVID-19 around the world, including 6.8 million
+deaths\textsuperscript{\protect\hyperlink{ref-WHO-Covid}{1}}. Although
+this disease is no longer categorized as a global health emergency by
+the World Health Organization
+(WHO)\textsuperscript{\protect\hyperlink{ref-rigby2023}{2}}, there is
+ongoing concern due to continued transmission, surges in cases and
+deaths due to new
+variants\textsuperscript{\protect\hyperlink{ref-un2023}{3}}, and
+weaknesses in health systems around the world that could be exploited by
+a novel virus or another public health emergency in the
+future\textsuperscript{\protect\hyperlink{ref-mackey2021}{4}}.
+
+In particular, a major weakness that has received attention during the
+pandemic has been related to vaccine uptake. The rapid development of
+vaccines against COVID-19 initially brought the hope of a rapid end to
+the pandemic due to the start of vaccination campaigns in certain parts
+of the world toward the end of
+2020\textsuperscript{\protect\hyperlink{ref-thelancet2021}{5}--\protect\hyperlink{ref-tanne2020}{8}})
+but inequalities in vaccine uptake made these pharmaceutical
+interventions ultimately unable to replicate the experience of smallpox,
+where vaccination on a global scale and was crucial to control this
+disease\textsuperscript{\protect\hyperlink{ref-kayser2021}{9}}.
+
+Although the problematic on vaccine uptake is a multifaceted issue
+resulting from a combination of factors, among which are failed public
+health measures\textsuperscript{\protect\hyperlink{ref-li2021}{10}},
+inequality in vaccine access between high- and low-income
+countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{11},\protect\hyperlink{ref-tamey2022}{12}},
+and vaccine
+hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{13}}, it
+is well established that individuals in certain equity-deserving groups
+(e.g., Black, Asian, or Indigenous) as well as individuals with
+socio-economic disadvantages have been disproportionately impacted by
+it, as indicated by the low vaccine uptake observed in these
+groups\textsuperscript{\protect\hyperlink{ref-willis2021}{14}--\protect\hyperlink{ref-hussain2022}{20}}.
+
+Reasons given for this problematic have included medical mistrust due to
+systemic medical
+racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{16}}, mistrust
+in vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{14}}, and
 the influence of conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-bogart2021}{16}--\protect\hyperlink{ref-freeman2020}{18}}.
-Moreover, in the case of Canada, lower vaccine uptake has been observed
-in young individuals, those with a low educational level, households
-with children, those without a regular healthcare provider, individuals
-that identify as part of certain equity-deserving groups, and those with
-a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21}}.
-
-However, it is important to consider that vaccination uptake can also be
+theories\textsuperscript{\protect\hyperlink{ref-bogart2021}{21}--\protect\hyperlink{ref-freeman2020}{23}}.
+However, it is important to also consider that vaccination uptake can be
 influenced by geographical (spatial) factors. In this regard,
@@
 varied regional attitudes towards
-vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{11}},
+vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{24}},
 spatial differences in vaccine access and supply, vaccination location
@@
 groups
-reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2},\protect\hyperlink{ref-nguyen2021}{22}}.
+reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-nguyen2021}{25}}.
 Other studies have also shown heterogeneity in vaccine uptake within
 small governmental administrative units such as
-counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{23}--\protect\hyperlink{ref-bhuiyan2022}{26}},
-and that and that accounting for geographical differences in vaccination
-can help predict patterns of booster
-uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{27}}. Overall,
-the evidence provided by the literature demonstrates the existence of
-spatially-driven heterogeneities in vaccine uptake that be used by
-decision-makers in the development of public health policies that are
-focused on addressing these disparities within specific administrative
-or geographical areas.
+counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{26}--\protect\hyperlink{ref-bhuiyan2022}{29}},
+and that accounting for geographical differences in vaccination can help
+predict patterns of booster
+uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{30}}.
 
@@
 of Canada, where available studies have been focused in certain cities
-(such as Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{28}},
-or Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{29}}),
+(such as Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{31}},
+or Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{32}}),
 or have explored differences at a province-wide
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. Thus, there
-is a need for studies that explore spatial differences in vaccination
-within the Canadian territory and that consequently, can help identify
-disparities that need to be addressed within specific areas in each
-province.
-
-This need is particularly important in the case of Ontario, the most
+level\textsuperscript{\protect\hyperlink{ref-guay2022}{18}}. Therefore,
+there is a need for studies that explore spatial differences in
+vaccination within the Canadian territory and that consequently, can
+help identify disparities that need to be addressed within specific
+areas in each province.
+
+This is particularly important in the case of Ontario, the most
 populated province in Canada. Between 2006 and 2019, Ontario provided
 healthcare access to its inhabitants using 14 intra-provincial divisions
-called the Local Health Integrated Networks (LHINs). However, this
-approach was complex, bureaucratic, and led to systemic
-inequalities\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30}}. In
-late 2019, the 14 LHINs were phased out and the areas they covered were
+called the Local Health Integrated Networks (LHINs). However, due to is
+complexity, bureaucratic burden, and lack of collaborative work this
+approach resulted in disparities such as excessive expenditures and
+differences in mortality
+rates\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33},\protect\hyperlink{ref-muratov2018}{34}}.
+In an effort to improve the healthcare system of the province, in late
+2019 the 14 LHINs were phased out and the areas they covered were
 incorporated into 6 Health Regions (North East, North West, Central,
-Toronto, West, and East) in an effort to improve the healthcare system
-of the province\textsuperscript{\protect\hyperlink{ref-dong2022}{31}}.
-Because the adoption of the Health Regions is relatively recent, there
-is an ongoing need to analyze the impact of this measure and identify
-disparities in health access that might exist across the Health Regions,
-which can be specially important in the context of the COVID-19
-pandemic.
+Toronto, West, and
+East)\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}. Due to the
+relatively recent adoption of the Health Region model and its alignment
+with the onset of the COVID-19 pandemic, there is a need to analyze if
+there are ongoing disparities in health access under this approach that
+need to be addressed before they are exploited by a new disease or
+public health threat. In this regard, analyzing differences in
+vaccination uptake within the Health Regions and can help identify which
+socio-demographic groups are the most vulnerable and what areas of the
+province deserve special attention by decision-makers.
 
@@
 vaccination uptake between the different Health Regions of Ontario
-during the last quarter of 2021. By including socio-economic factors in
-our analysis, we aimed at identifying in which groups these differences
-were significant in order to provide an assessment of the current state
-of healthcare access in Ontario.
+between October of 2021 and January of 2022. By including socio-economic
+factors in our analysis, we aimed at identifying in which groups these
+differences were significant in order to provide an assessment of the
+current state of healthcare access in Ontario.
 
@@
 of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{32}--\protect\hyperlink{ref-cnat2022a}{34}}.
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{36}--\protect\hyperlink{ref-cnat2022a}{38}}.
 Because we identified differences in representativity between the survey
@@
 proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{35}}
+(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{39}}
 to correct the data using data from the Census and Health Region
@@
 Appendix. All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{36}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{37}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{38}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{39}},
+\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{40}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{41}},
+\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{42}},
+\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{43}},
 and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{40}}.
+\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{44}}.
 
@@
 
-Figure~\ref{fig-model-uncorr} shows the estimates from the logistic
-regression model of vaccination status for the uncorrected data, whereas
-the estimated obtained from the corrected data appear in
-Figure~\ref{fig-model-corr}. The results show significantly lower odds
-of vaccination in individuals with a low household income and those that
-identified as part of equity-deserving groups when compared to
-individuals in high income brackets or that identified as White or
-Caucasian.
+Figure~\ref{fig-model-uncorr} presents the estimates (as odd ratios)
+from the logistic regression models for vaccination status using the
+socio-demographic factors collected by the survey, and their
+interactions. Generally speaking, lower odds of vaccination were
+identified in both cases in individuals characterized by a low household
+income, or that identified as part of equity-deserving groups. However,
+the magnitude of the estimates differed between the uncorrected and
+corrected models and more importantly, certain estimates were not deemed
+to be statistically-significant after the correction, in contrast to the
+estimates from the uncorrected model. Specifically, the corrected model
+showed no significant differences in vaccination odds between the age
+groups considered, the East Health Region, Latin American individuals
+with a household income under CAD 25,000, and Indigenous individuals
+living in the Central Health Region (Figure~\ref{fig-model-uncorr},B).
 
@@
 
-Specifically, significantly lower odds of vaccination were identified
-for those with a household income under CAD 25,000 (OR=0.37,
-CI={[}0.25,0.56{]}) and those with an income between CAD 25,000 and
-59,999 (OR=0.59, CI={[}0.39,0.88{]}). Additionally, individuals who
-identified as Arab/Middle Eastern, Black, or Latin American, or that
-belonged to ``Other'' racial groups, which included the Southeast Asian,
-Filipino, West Asian, and Minorities Not Identified Elsewhere groups
-according to the Census, had significantly lower odds of vaccination
-than those in the White/Caucasian group (ORs=0.31, 0.32, 0.27, 0.22, and
-CIs={[}0.14,0.68{]}, {[}0.17,0.60{]}, {[}0.11,0.66{]}, and
-{[}0.12,0.41{]}, respectively). Regarding Health Regions, individuals
-that reported living in the West Health Region (which includes the
-regions of Waterloo and Niagara, the counties of Wellington, Essex, and
-Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent)
-had significantly higher odds of vaccination than those in the Health
-Region of Toronto (OR=1.54, CI={[}1.04,2.29{]}).
+However, significantly lower odds of vaccination were identified in the
+corrected model for those with a household income under CAD 25,000
+(OR=0.37, CI={[}0.27,0.51{]}) and those with an income between CAD
+25,000 and 59,999 (OR=0.58, CI={[}0.42,0.81{]}). Additionally,
+individuals who identified as Arab/Middle Eastern, Black, Latin
+American, of mixed background, or that belonged to other racial groups
+(a category that included Southeast Asian, Filipino, West Asian, and
+minorities not identified elsewhere), had significantly lower odds of
+vaccination than those in the White/Caucasian group (ORs and CIs=0.28
+{[}0.16,0.51{]}, 0.27 {[}0.16,0.45{]}, 0.40 {[}0.21,0.76{]}, 0.53
+{[}0.30,0.92{]}, 0.23 {[}0.15,0.36{]}). Additionally, individuals that
+reported living in the Central and West Health Regions had higher odds
+of vaccination than those in the Health Region of Toronto (ORs and
+CIs=1.61 {[}1.10,2.34{]}, and 1.59 {[}1.16,2.19{]}, respectively).
 
@@
 those with a household income above CAD 60,000). This held true in the
-case of Arab/Middle Eastern (OR=3.08, CI={[}1.27,7.47{]}), Black
-(OR=3.15, CI={[}1.43,6.92{]}), and Latin American (OR=2.81,
-CI={[}1.04,7.59{]}) individuals, as well as respondents who belonged to
-``Other'' minority groups (OR=4.63, CI={[}2.34,9.13{]}), who also had
-higher odds of vaccination in the CAD 25,000-59,999 income bracket
-(OR=6.96, CI={[}2.67,18.16{]}). Finally, significantly lower odds of
-vaccination were identified (when compared to the Toronto Health Region)
-for Black individuals in the Central Health Region, which comprises the
-region of York, counties of Dufferin and Simcoe and the district of
-Muskoka (OR=0.44, CI={[}0.19,0.99{]}), and in individuals that
-identified as part of other racial minorities or South Asian that lived
-in the West Health Region (ORs=0.41, and CIs={[}0.18,0.92{]} and
-{[}0.18,0.95{]}, respectively).
+case of Arab/Middle Eastern (OR=34, CI={[}1.70,6.79{]}), Black
+individuals (OR=3.81, CI={[}2.05, 7.09{]}), or those in other racial or
+ethnic groups (OR=3.19, CI={[}2.00,5.09{]}). Additionally, individuals
+with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern
+and other racial ethnic groups had higher odds of vaccination (ORs and
+CIs=6.96 {[}2.67,18.16{]}, and 3.5 {[}1.85,6.62{]}).
+
+Finally, significantly lower odds of vaccination were identified (when
+compared to the Toronto Health Region) for Black individuals in the
+Central Health Region (OR=0.39, CI={[}0.2,0.75{]}), Arab/Middle Eastern
+individuals in the East Health Region (OR=0.41 {[}0.17, 0.98{]}), and in
+the Indigenous and mixed groups in the West Health Region (ORs and
+CIs={[}0.31 {[}0.14, 0.7{]} and 0.38 {[}0.19, 0.76{]}, respectively).
 
@@
 
+West Health Region includes the regions of Waterloo and Niagara, the
+counties of Wellington, Essex, and Lambton, and the cities of Hamilton,
+Haldimand, Brant, and Chatham-Kent. The Central Health Region includes
+the regions of Peel and Halton, the regional municipality of York, the
+district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the
+city of Etobicoke.
+
+had significantly higher odds of vaccination than those in the Health
+Region of Toronto
+
 The existence of healthcare disparities in Ontario motivated the recent
@@
 LHIN model to the Health Region model in late
-2019\textsuperscript{\protect\hyperlink{ref-tsasis2012}{30},\protect\hyperlink{ref-dong2022}{31}}.
+2019\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33},\protect\hyperlink{ref-dong2022}{35}}.
 This change is relatively recent, and because of this, is likely that
@@
 identify as part of a racial minority, or that have a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{41}}.
+income\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{45}}.
 In the case of Ontario, a possible rationale for this difference is
@@
 factor of disparities in vaccination for equity-deserving
-groups\textsuperscript{\protect\hyperlink{ref-njoku2021}{42}}, and that
+groups\textsuperscript{\protect\hyperlink{ref-njoku2021}{46}}, and that
 occurred in Toronto earlier in 2021, where it has been shown that
@@
 highest burden of COVID-19 at the
-time\textsuperscript{\protect\hyperlink{ref-bogoch2022}{2}}.
+time\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}.
 
@@
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{43},\protect\hyperlink{ref-ct2021}{44}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{47},\protect\hyperlink{ref-ct2021}{48}},
 which include grocery store workers, gas station, warehouse,
@@
 for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{45}},
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{49}},
 which would explain the higher odds of vaccination for these
@@
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{50}}.
 Therefore, the estimated lower odds are likely to be explained from a
@@
 in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{47}}
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{51}}
 (Supplementary Table A-7). Interestingly, a disproportionate number of
@@
 reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{48}}. These
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{52}}. These
 results provide context to the observed differences in vaccination among
@@
 are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{46}}, a
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{50}}, a
 minority group that has historically suffered from reduced access to
 health care and
-discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{17}}.
+discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{22}}.
 Therefore, there is a need for additional studies that focus on these
@@
 reported by Public Health Ontario at the start of the period covered by
-the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{49}}.
+the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{53}}.
 However, we found good agreement between the estimates from the model
@@
 relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{50}}, and
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{54}}, and
 with vaccination uptake rates across different age groups presented in
 other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{19},\protect\hyperlink{ref-macdonald2021}{51}}.
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{55}}.
 
@@
 across all age groups during the three months covered by the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{49}}. It is
+survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{53}}. It is
 also important to mention that to this day, differences in vaccination
@@
 still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{52}}, and
+rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{56}}, and
 although data for the period analyzed in this study is not
@@
 documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{53}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{57}}, we
 believe that our data was able to provide a valid assessment of
@@
 \begin{CSLReferences}{0}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-WHO-Covid}{}}%
+\CSLLeftMargin{1. }%
+\CSLRightInline{{World Health Organization Coronavirus (COVID-19)
+Dashboard}. Accessed May 11, 2023. \url{https://covid19.who.int/}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-rigby2023}{}}%
+\CSLLeftMargin{2. }%
+\CSLRightInline{Rigby J, Satija B. {WHO} declares end to COVID global
+health emergency. \emph{Reuters}. Published online May 8, 2023. Accessed
+May 11, 2022.
+\url{https://www.reuters.com/business/healthcare-pharmaceuticals/covid-is-no-longer-global-health-emergency-who-2023-05-05/}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-un2023}{}}%
+\CSLLeftMargin{3. }%
+\CSLRightInline{Nations U. {WHO} chief declares end to COVID-19 as a
+global health emergency. \emph{UN News}. Published online May 5, 2023.
+Accessed May 11, 2022.
+\url{https://news.un.org/en/story/2023/05/1136367}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mackey2021}{}}%
+\CSLLeftMargin{4. }%
+\CSLRightInline{Mackey K, Ayers CK, Kondo KK, et al. Racial and ethnic
+disparities in {COVID}-19{\textendash}related infections,
+hospitalizations, and deaths. \emph{Annals of Internal Medicine}.
+2021;174(3):362-373.
+doi:\href{https://doi.org/10.7326/m20-6306}{10.7326/m20-6306}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
+\CSLLeftMargin{5. }%
+\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
+end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
+doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
-\CSLLeftMargin{1. }%
+\CSLLeftMargin{6. }%
 \CSLRightInline{Davis CJ, Golding M, McKay R. Efficacy information
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
-\CSLLeftMargin{2. }%
+\CSLLeftMargin{7. }%
 \CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
-\CSLLeftMargin{3. }%
+\CSLLeftMargin{8. }%
 \CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
-\CSLLeftMargin{4. }%
-\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
-end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
-doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-watson2022}{}}%
-\CSLLeftMargin{5. }%
-\CSLRightInline{Watson OJ, Barnsley G, Toor J, Hogan AB, Winskill P,
-Ghani AC. Global impact of the first year of {COVID}-19 vaccination: A
-mathematical modelling study. \emph{The Lancet Infectious Diseases}.
-2022;22(9):1293-1302.
-doi:\href{https://doi.org/10.1016/s1473-3099(22)00320-6}{10.1016/s1473-3099(22)00320-6}}
-
 \leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
-\CSLLeftMargin{6. }%
+\CSLLeftMargin{9. }%
 \CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
@@
 \leavevmode\vadjust pre{\hypertarget{ref-li2021}{}}%
-\CSLLeftMargin{7. }%
+\CSLLeftMargin{10. }%
 \CSLRightInline{Li Q, Wang J, Tang Y, Lu H. Next-generation COVID-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
-\CSLLeftMargin{8. }%
+\CSLLeftMargin{11. }%
 \CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
-\CSLLeftMargin{9. }%
+\CSLLeftMargin{12. }%
 \CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
-\CSLLeftMargin{10. }%
+\CSLLeftMargin{13. }%
 \CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
-\CSLLeftMargin{11. }%
-\CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
-of {COVID}-19 vaccine acceptance in the {US}.
-\emph{{EClinicalMedicine}}. 2020;26:100495.
-doi:\href{https://doi.org/10.1016/j.eclinm.2020.100495}{10.1016/j.eclinm.2020.100495}}
-
 \leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{14. }%
 \CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{15. }%
 \CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{16. }%
 \CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
@@
 \leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{17. }%
 \CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
-\CSLLeftMargin{16. }%
-\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
-related medical mistrust, health impacts, and potential vaccine
-hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
-of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
-doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
-\CSLLeftMargin{17. }%
-\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
-roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
-canada. \emph{Canadian Medical Association Journal}.
-2021;193(11):E381-E383.
-doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
 \CSLLeftMargin{18. }%
-\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
-hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
-and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
-2020;52(14):3127-3141.
-doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
-\CSLLeftMargin{19. }%
 \CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{19. }%
 \CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
+\CSLLeftMargin{21. }%
+\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
+related medical mistrust, health impacts, and potential vaccine
+hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
+of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
+doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
+\CSLLeftMargin{22. }%
+\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
+roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
+canada. \emph{Canadian Medical Association Journal}.
+2021;193(11):E381-E383.
+doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
+\CSLLeftMargin{23. }%
+\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
+hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
+and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
+2020;52(14):3127-3141.
+doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
+\CSLLeftMargin{24. }%
+\CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
+of {COVID}-19 vaccine acceptance in the {US}.
+\emph{{EClinicalMedicine}}. 2020;26:100495.
+doi:\href{https://doi.org/10.1016/j.eclinm.2020.100495}{10.1016/j.eclinm.2020.100495}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
@@
 \leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-muratov2018}{}}%
+\CSLLeftMargin{34. }%
+\CSLRightInline{Muratov S, Lee J, Holbrook A, et al. Regional variation
+in healthcare spending and mortality among senior high-cost healthcare
+users in ontario, canada: A retrospective matched cohort study.
+\emph{{BMC} Geriatrics}. 2018;18(1).
+doi:\href{https://doi.org/10.1186/s12877-018-0952-7}{10.1186/s12877-018-0952-7}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Dong L, Sahu R, Black R. Governance in the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
@@
 \leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
@@
 \leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-05-12 manuscript_v5/01-background.qmd
updated, working on discussion

@@
 
-This is particularly important in the case of Ontario, the most populated province in Canada.
-Between 2006 and 2019, Ontario provided healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs).
-However, due to is complexity, bureaucratic burden, and lack of collaborative work this approach resulted in disparities such as excessive expenditures and differences in mortality rates [@tsasis2012;@muratov2018].
-In an effort to improve the healthcare system of the province, in late 2019 the 14 LHINs were phased out and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East)  [@dong2022].
+Between 2007 and 2019, Ontario provided healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs).
+However, this approach was complex and bureaucratic, resulting in excessive expenditures, disparities in mortality rates, and the deterioration of certain performance indicators such as wait times and readmissions [@tsasis2012;@muratov2018;@dong2022].
+In an effort to improve the healthcare system of the province, in late 2019 the government dissolved the 14 LHINs  and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East)  [@dong2022].
 Due to the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
-In this regard, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
+In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
+Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
 

### 2023-05-12 manuscript_v5/03-results.qmd
updated, working on discussion

@@
 
+# test<-a1_rake$variables
+# 
+# a1_rake %>%
+# tbl_svysummary(percent="row",
+#               by= first_dose,
+#               include=c(income_ord,
+#                         Age_group_ord,
+#                         Health_Region,
+#                         Month,
+#                         Race),
+#               label=list(Health_Region ~ "Health Region",
+#                  Age_group_ord ~ "Age Group",
+#                  income_ord="Income (CAD)")
+#   ) %>% add_p(test = everything () ~ "chisq.test")%>%
+#   bold_labels() %>%
+#   modify_table_styling(
+#     columns=label,
+#     rows=label=="Other",
+#     footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
+#   ) %>%
+#   modify_header(label = "**Variable**") %>%
+#   as_gt()
+#
+
 ```
@@
 
-@fig-model-uncorr presents the estimates (as odd ratios) from the logistic regression models for vaccination status using the socio-demographic factors collected by the survey, and their interactions.
+@fig-models presents the estimates (as odd ratios) from the logistic regression models for vaccination status using the socio-demographic factors collected by the survey, and their interactions.
 Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of equity-deserving groups.
 However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, certain estimates were not deemed to be statistically-significant after the correction, in contrast to the estimates from the uncorrected model.
-Specifically, the corrected model showed no significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-model-uncorr,B).
+Specifically, the corrected model showed no significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-models,B).
 <!-- \newpage -->
@@
 #| fig-height: 12
-#| label: fig-model-uncorr
+#| label: fig-models
 #| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model.
@@
 
-#print(fig_mod, vp=viewport(angle=90,height = 0.5,clip="off"))
 fig_mod
 
+#table for comparison
+#tbl_merge(tbls = list(tbl_regression(m0,exponentiate=TRUE),tbl_regression(m2,exponentiate = TRUE)))
 ```

### 2023-05-12 manuscript_v5/04-discussion.qmd
updated, working on discussion

@@
 
-The existence of healthcare disparities in Ontario motivated the recent change in the healthcare system of the province, which switched from the LHIN model to the Health Region model in late 2019  [@tsasis2012;@dong2022].
-This change is relatively recent, and because of this, is likely that there are ongoing differences in healthcare access across the province.
-In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions, aiming at determining which socio-demographic groups could be impacted by these disparities and to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to improve health access for the inhabitants of Ontario, a province that faces unique challenges due to its condition as the most populated and the most ethnically diverse province of Canada.
+Between 2007 and 2019, healthcare in Ontario was managed by 14 different LHINs, which aimed to provide an integrated health system for the province.
+However, this approach faced chronic problems such as fragmented electronic health systems [@lysyk2016], the decline of performance indicators (e.g., increased unplanned emergency visits, number of patient readmissions), and inequities in health services access [@lysyk2015].
+Based on these problematics and with the intent of better organizing and delivering care, the provincial government moved to the Health Region model in late 2019 [@dong2022].
 
-Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of surveyed individuals reported to have received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats).
-We identified significant intra-provincial differences in vaccination based on socio-economic and geographical factors.
-First, our results show differences in odds of vaccination in individuals with a household income below CAD 60,000 and in individuals belonging to visible minority groups.
-Those who identified as Arab/Middle Eastern, Black, Latin American, or  that belonged to a minority group not included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were less than a third of individuals that identified as White/Caucasian (@fig-model-corr).
-These results are consistent with other studies that have shown lower vaccination rates in individuals that identify as part of a racial minority, or that have a low household income [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
-In the case of Ontario, a possible rationale for this difference is vaccine access, which has been identified as an important contributing factor of disparities in vaccination for equity-deserving groups [@njoku2021], and that occurred in Toronto earlier in 2021, where it has been shown that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 at the time [@bogoch2022].
+In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its missing of improving health access for the inhabitants of Ontario.
+
+Our results show differences in odds of vaccination in individuals across certain socio-demographic groups.
+Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models).
+These results are consistent with previous studies that have shown lower vaccination rates in individuals with the same socio-demographic characteristics [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
+
+Lower vaccine uptake in the socio-demographic groups indicated above may be influenced in part, by vaccine hesitancy and refusal [@cnat2022a;@cnat2022b], which have been associated in Canadian individuals to concerns on vaccine safety and effectiveness [@basta2022].
+However, it is also likely that accessibility played an important role in vaccine uptake, as it has been shown that this issue plays an important role in the case of equity-deserving groups [@njoku2021].
+Although a spatial analysis of vaccine accessibility in Ontario is beyond the scope of this study, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 at the time in Toronto and other regions of Ontario [@bogoch2022;@iveniuk2021].
 

### 2023-05-12 manuscript_v5/appendix.qmd
updated, working on discussion

@@
 
-Tables A-7 and A-8 present summaries of information from the Fields survey (from the clean dataset analyzed in the main paper), showing responses by White/Caucasian individuals by Health Region and proportion of answers by LHIN in the West Health Region.
+Tables A-7 and A-8 present summaries of information from the Fields survey (from the dataset analyzed in the main paper), showing responses by White/Caucasian individuals by Health Region and proportion of answers by LHIN in the West Health Region.
 
@@
   as_kable_extra(booktabs = TRUE,linesep="")
+
+  clean_data %>%
+  select(Race,Health_Region,first_dose) %>%
+  tbl_strata(
+    strata = Health_Region,
+    .tbl_fun =
+      ~ .x %>%
+        tbl_summary(by = first_dose, missing = "no",percent="row") %>%
+        add_n(),
+    .header = "**{strata}**, N = {n}"
+  )
 ```
@@
 ```
+
+```{r,Central-HR-LHIN}
+#| label: tbl-central-hr-lhin
+#| tbl-cap: "Proportion of answers by LHIN (Central Health Region)"
+#| message: false
+#| echo: false
+
+clean_data %>% subset (Health_Region=="Central")%>%
+tbl_summary(include=c(LHIN),
+            sort = all_categorical() ~ "frequency") %>% 
+  bold_labels()%>%
+  as_kable_extra(booktabs = TRUE,linesep="")
+
+```

### 2023-05-12 manuscript_v5/appendix.tex
updated, working on discussion

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 Tables A-7 and A-8 present summaries of information from the Fields
-survey (from the clean dataset analyzed in the main paper), showing
-responses by White/Caucasian individuals by Health Region and proportion
-of answers by LHIN in the West Health Region.
+survey (from the dataset analyzed in the main paper), showing responses
+by White/Caucasian individuals by Health Region and proportion of
+answers by LHIN in the West Health Region.
 
@@
 
+\hypertarget{tbl-central-hr-lhin}{}
+\begin{table}
+\caption{\label{tbl-central-hr-lhin}Proportion of answers by LHIN (Central Health Region) }\tabularnewline
+
+\centering
+\begin{tabular}{lc}
+\toprule
+\textbf{Characteristic} & \textbf{N = 1,455}\\
+\midrule
+\textbf{LHIN} & \\
+\hspace{1em}Mississauga Halton & 495 (34\%)\\
+\hspace{1em}Central & 428 (29\%)\\
+\hspace{1em}Central West & 362 (25\%)\\
+\hspace{1em}North Simcoe Muskoka & 170 (12\%)\\
+\bottomrule
+\multicolumn{2}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\end{tabular}
+\end{table}
+
 \end{document}

### 2023-05-12 manuscript_v5/main.tex
updated, working on discussion

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, enhanced, boxrule=0pt, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, sharp corners, interior hidden, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-This is particularly important in the case of Ontario, the most
-populated province in Canada. Between 2006 and 2019, Ontario provided
-healthcare access to its inhabitants using 14 intra-provincial divisions
-called the Local Health Integrated Networks (LHINs). However, due to is
-complexity, bureaucratic burden, and lack of collaborative work this
-approach resulted in disparities such as excessive expenditures and
-differences in mortality
-rates\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33},\protect\hyperlink{ref-muratov2018}{34}}.
+Between 2007 and 2019, Ontario provided healthcare access to its
+inhabitants using 14 intra-provincial divisions called the Local Health
+Integrated Networks (LHINs). However, this approach was complex and
+bureaucratic, resulting in excessive expenditures, disparities in
+mortality rates, and the deterioration of certain performance indicators
+such as wait times and
+readmissions\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33}--\protect\hyperlink{ref-dong2022}{35}}.
 In an effort to improve the healthcare system of the province, in late
-2019 the 14 LHINs were phased out and the areas they covered were
-incorporated into 6 Health Regions (North East, North West, Central,
-Toronto, West, and
+2019 the government dissolved the 14 LHINs and the areas they covered
+were incorporated into 6 Health Regions (North East, North West,
+Central, Toronto, West, and
 East)\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}. Due to the
@@
 need to be addressed before they are exploited by a new disease or
-public health threat. In this regard, analyzing differences in
-vaccination uptake within the Health Regions and can help identify which
-socio-demographic groups are the most vulnerable and what areas of the
-province deserve special attention by decision-makers.
+public health threat. In this regard, previous research has highlighted
+disparities in the level of activity of each Health
+Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{36}}.
+Therefore, analyzing differences in vaccination uptake within the Health
+Regions and can help identify which socio-demographic groups are the
+most vulnerable and what areas of the province deserve special attention
+by decision-makers.
 
@@
 of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{36}--\protect\hyperlink{ref-cnat2022a}{38}}.
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{37}--\protect\hyperlink{ref-cnat2022a}{39}}.
 Because we identified differences in representativity between the survey
@@
 proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{39}}
+(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{40}}
 to correct the data using data from the Census and Health Region
@@
 Appendix. All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{40}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{41}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{42}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{43}},
+\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{41}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{42}},
+\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{43}},
+\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{44}},
 and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{44}}.
+\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{45}}.
 
@@
 
-The existence of healthcare disparities in Ontario motivated the recent
-change in the healthcare system of the province, which switched from the
-LHIN model to the Health Region model in late
-2019\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33},\protect\hyperlink{ref-dong2022}{35}}.
-This change is relatively recent, and because of this, is likely that
-there are ongoing differences in healthcare access across the province.
+Between 2007 and 2019, healthcare in Ontario was managed by 14 different
+LHINs, which aimed to provide an integrated health system for the
+province. However, this approach faced chronic problems such as
+fragmented electronic health
+systems\textsuperscript{\protect\hyperlink{ref-lysyk2016}{46}}, the
+decline of performance indicators (e.g., increased unplanned emergency
+visits, number of patient readmissions), and inequities in health
+services access\textsuperscript{\protect\hyperlink{ref-lysyk2015}{47}}.
+Based on these problematics and with the intent of better organizing and
+delivering care, the provincial government moved to the Health Region
+model in late
+2019\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}.
+
 In this study we hypothesized that differences in COVID-19 vaccination
 uptake were present between the Health Regions, aiming at determining
-which socio-demographic groups could be impacted by these disparities
-and to provide decision-makers with information that could be used to
+which socio-demographic groups could be impacted by these disparities in
+order to provide decision-makers with information that could be used to
 develop policies focused on reducing or eliminating these differences
-and ensure that the Health Region model is able to improve health access
-for the inhabitants of Ontario, a province that faces unique challenges
-due to its condition as the most populated and the most ethnically
-diverse province of Canada.
+and ensure that the Health Region model is able to fulfill its missing
+of improving health access for the inhabitants of Ontario.
 
@@
 identify as part of a racial minority, or that have a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{45}}.
+income\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
 In the case of Ontario, a possible rationale for this difference is
@@
 factor of disparities in vaccination for equity-deserving
-groups\textsuperscript{\protect\hyperlink{ref-njoku2021}{46}}, and that
+groups\textsuperscript{\protect\hyperlink{ref-njoku2021}{49}}, and that
 occurred in Toronto earlier in 2021, where it has been shown that
@@
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{47},\protect\hyperlink{ref-ct2021}{48}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{50},\protect\hyperlink{ref-ct2021}{51}},
 which include grocery store workers, gas station, warehouse,
@@
 for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{49}},
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{52}},
 which would explain the higher odds of vaccination for these
@@
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{50}}.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{53}}.
 Therefore, the estimated lower odds are likely to be explained from a
@@
 in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{51}}
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{54}}
 (Supplementary Table A-7). Interestingly, a disproportionate number of
@@
 reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{52}}. These
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{55}}. These
 results provide context to the observed differences in vaccination among
@@
 are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{50}}, a
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{53}}, a
 minority group that has historically suffered from reduced access to
@@
 reported by Public Health Ontario at the start of the period covered by
-the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{53}}.
+the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{56}}.
 However, we found good agreement between the estimates from the model
@@
 relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{54}}, and
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{57}}, and
 with vaccination uptake rates across different age groups presented in
 other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{55}}.
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{58}}.
 
@@
 across all age groups during the three months covered by the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{53}}. It is
+survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{56}}. It is
 also important to mention that to this day, differences in vaccination
@@
 still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{56}}, and
+rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{59}}, and
 although data for the period analyzed in this study is not
@@
 documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{57}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{60}}, we
 believe that our data was able to provide a valid assessment of
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-sethuram2023}{}}%
+\CSLLeftMargin{36. }%
+\CSLRightInline{Sethuram C, McCutcheon T, Liddy C. An environmental scan
+of ontario health teams: A descriptive study. \emph{{BMC} Health
+Services Research}. 2023;23(1).
+doi:\href{https://doi.org/10.1186/s12913-023-09102-6}{10.1186/s12913-023-09102-6}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
@@
 \leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
@@
 \leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lysyk2016}{}}%
+\CSLLeftMargin{46. }%
+\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
+2016. In: \emph{Section 3.03: Electronic Health Records' Implementation
+Status}. Queen's Printer for Ontario; 2016. Accessed May 12, 2023.
+\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en16/v1_303en16.pdf}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lysyk2015}{}}%
+\CSLLeftMargin{47. }%
+\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
+2015. In: \emph{Section 3.08: LHINs - Local Health Integration
+Networks}. Queen's Printer for Ontario; 2015. Accessed May 12, 2023.
+\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en15/3.08en15.pdf}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-05-14 manuscript_v5/01-background.qmd
discussion updated

@@
 
-In particular, a major weakness that has received attention during the pandemic has been related to vaccine uptake.
+In particular, a major weakness that has received attention during the pandemic has been related to inequalities in vaccine uptake.
 The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic due to the start of vaccination campaigns in certain parts of the world toward the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020]) but inequalities in vaccine uptake made these pharmaceutical interventions ultimately unable to replicate the experience of smallpox, where vaccination on a global scale and was crucial to control this disease [@kayser2021].
 
-Although the problematic on vaccine uptake is a multifaceted issue resulting from a combination of factors, among which are failed public health measures [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021; @tamey2022], and vaccine hesitancy [@nafilyan2021], it is well established that individuals in certain equity-deserving groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages have been disproportionately impacted by it, as indicated by the low vaccine uptake observed in these groups [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
+This problematic is a multifaceted issue resulting from a combination of factors, among which are failed public health measures [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021; @tamey2022], and vaccine hesitancy [@nafilyan2021].
+Furthermore, it is well established that this issue has affected in particular individuals in certain equity-deserving groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
 
-Reasons given for this problematic have included medical mistrust due to systemic medical racism [@stoler2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
+Reasons given for this inequality have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
 However, it is important to  also consider that vaccination uptake can be influenced by geographical (spatial) factors.

### 2023-05-14 manuscript_v5/03-results.qmd
discussion updated

@@
 #
-# a1_rake %>%
+
+# ds2<- a1_rake %>%
 #   tbl_svysummary(percent="row",
@@
 #   ) %>%
-#   modify_header(label = "**Variable**") %>%
+#   modify_header(label = "**Variable**")
+# 
+# ds1<- a1 %>%
+#   tbl_svysummary(percent="row",
+#               by= first_dose,
+#               include=c(income_ord,
+#                         Age_group_ord,
+#                         Health_Region,
+#                         Month,
+#                         Race),
+#               label=list(Health_Region ~ "Health Region",
+#                  Age_group_ord ~ "Age Group",
+#                  income_ord="Income (CAD)")
+#   ) %>% add_p(test = everything () ~ "chisq.test")%>%
+#   bold_labels() %>%
+#   modify_table_styling(
+#     columns=label,
+#     rows=label=="Other",
+#     footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
+#   ) %>%
+#   modify_header(label = "**Variable**") 
+# 
+# 
+# tbl_merge(tbls=list(ds1,ds2)) %>%
 #   as_gt()
-#
 
@@
 #| fig-width: 15
-#| fig-height: 12
+#| fig-height: 14
 #| label: fig-models
@@
 fig_mod<-uncorr_mod+corr_mod+plot_layout(guides = "collect")+plot_annotation(tag_levels = "A")&
-    theme(plot.tag = element_text(size=18,face='bold'),
+    theme(plot.tag = element_text(size=22,face='bold'),
           #axis.title = element_text(size=11),
-          axis.text = element_text(size=12),
-          strip.text=element_text(size=12))
+          axis.text = element_text(size=18),
+          strip.text=element_text(size=18))
 
@@
 #call script with code for figure
-source(here("code","corr_model.R"))
+#source(here("code","corr_model.R"))
 
 #call figure
-corr_mod
+#corr_mod
 ```

### 2023-05-14 manuscript_v5/04-discussion.qmd
discussion updated

@@
 
-Lower vaccine uptake in the socio-demographic groups indicated above may be influenced in part, by vaccine hesitancy and refusal [@cnat2022a;@cnat2022b], which have been associated in Canadian individuals to concerns on vaccine safety and effectiveness [@basta2022].
-However, it is also likely that accessibility played an important role in vaccine uptake, as it has been shown that this issue plays an important role in the case of equity-deserving groups [@njoku2021].
-Although a spatial analysis of vaccine accessibility in Ontario is beyond the scope of this study, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 at the time in Toronto and other regions of Ontario [@bogoch2022;@iveniuk2021].
+Lower vaccine uptake in the socio-demographic groups indicated above may be influenced in part, by vaccine hesitancy and refusal, which have been associated in equity-deserving Canadian individuals to concerns on vaccine safety, effectiveness, and experiences of racial discrimination in health settings [@basta2022;@cnat2022a;@cnat2022b;@cnat2023].
+However, it has been shown that structural barriers also play an important role in vaccination uptake.
+In the case of equity-deserving individuals, such barriers include complex scheduling systems, language barriers, lack of adequate public transportation, and lack of accessible vaccination sites [@njoku2021].
+In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with demand [@gill2022] .
+However, to the best of our knowledge, there seems to be a very limited amount of studies in this area, indicating an ongoing need of future studies that examine the longitudinal impact of vaccine accessibility and structural barriers that affected equity-deserving groups within the province.
 
-It is interesting to note that although overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets (when compared to the CAD 60,000 and over bracket, Supplementary Figure A-5).
-More specifically, the change in odds of vaccination increased in individuals who identified as Arab/Middle Eastern, Black, Latin American, or belonging to other minority groups with a household income below CAD 25,000, which was also true for individuals in other racial minority groups with an income between CAD 25,000-59,999.
+Although overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets (when compared to the CAD 60,000 and over bracket, Supplementary Figure A-5).
+Specifically, individuals in low income brackets that belonged to Arab/Middle Eastern, Black, or other minority groups had higher odds of vaccination that their peers with an income above 60,000 CAD.
 
 This result is likely due in part to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include grocery store workers, gas station, warehouse, distribution, and manufacturing workers, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
-In the case of Ontario, essential workers had priority for COVID-19 vaccination [@mishra2021], which would explain the higher odds of vaccination for these individuals.
-In other words, it is possible that the type of occupation from individuals in equity-deserving groups played an important role in increasing the odds of vaccination.
+In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and interestingly, there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022].
+These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation from individuals in equity-deserving groups played an important role in increasing the odds of vaccination in the province.
+
+However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@fig-model-corr].
+In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-6).
+
+According to Ontario Health, 13.2% of the population in the West Health Region identifies as a visible minority, whereas 2.5% identifies as Indigenous [@ontariohealth].
+Therefore, the estimated lower odds are likely to be explained from a socio-economic perspective.
+In fact, 50% of the answers from this region in the survey came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
+Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-7).
+Interestingly, a disproportionate number of COVID-19 cases and low vaccination rate (under 50%) have been also reported in the South Asian community of Ontario [@anand2022].
+These results provide context to the observed differences in vaccination among the Health Regions of the province, showing which equity-deserving and socio-demographic groups in certain regions need to be prioritized in future vaccination strategies.
+Moreover, these results provide a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.
 

### 2023-05-14 manuscript_v5/main.tex
discussion updated

@@
 \usepackage{xcolor}
-\usepackage{amsmath}
 \usepackage{caption}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, sharp corners, interior hidden, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 In particular, a major weakness that has received attention during the
-pandemic has been related to vaccine uptake. The rapid development of
-vaccines against COVID-19 initially brought the hope of a rapid end to
-the pandemic due to the start of vaccination campaigns in certain parts
-of the world toward the end of
+pandemic has been related to inequalities in vaccine uptake. The rapid
+development of vaccines against COVID-19 initially brought the hope of a
+rapid end to the pandemic due to the start of vaccination campaigns in
+certain parts of the world toward the end of
 2020\textsuperscript{\protect\hyperlink{ref-thelancet2021}{5}--\protect\hyperlink{ref-tanne2020}{8}})
@@
 
-Although the problematic on vaccine uptake is a multifaceted issue
-resulting from a combination of factors, among which are failed public
-health measures\textsuperscript{\protect\hyperlink{ref-li2021}{10}},
-inequality in vaccine access between high- and low-income
+This problematic is a multifaceted issue resulting from a combination of
+factors, among which are failed public health
+measures\textsuperscript{\protect\hyperlink{ref-li2021}{10}}, inequality
+in vaccine access between high- and low-income
 countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{11},\protect\hyperlink{ref-tamey2022}{12}},
 and vaccine
-hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{13}}, it
-is well established that individuals in certain equity-deserving groups
-(e.g., Black, Asian, or Indigenous) as well as individuals with
-socio-economic disadvantages have been disproportionately impacted by
-it, as indicated by the low vaccine uptake observed in these
-groups\textsuperscript{\protect\hyperlink{ref-willis2021}{14}--\protect\hyperlink{ref-hussain2022}{20}}.
-
-Reasons given for this problematic have included medical mistrust due to
+hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{13}}.
+Furthermore, it is well established that this issue has affected in
+particular individuals in certain equity-deserving groups (e.g., Black,
+Asian, or Indigenous) as well as individuals with socio-economic
+disadvantages\textsuperscript{\protect\hyperlink{ref-willis2021}{14}--\protect\hyperlink{ref-hussain2022}{20}}.
+
+Reasons given for this inequality have included medical mistrust due to
 systemic medical
-racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{16}}, mistrust
-in vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{14}}, and
+racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{16},\protect\hyperlink{ref-mosby2021}{21}},
+mistrust in
+vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{14}}, and
 the influence of conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-bogart2021}{21}--\protect\hyperlink{ref-freeman2020}{23}}.
+theories\textsuperscript{\protect\hyperlink{ref-mosby2021}{21}--\protect\hyperlink{ref-freeman2020}{23}}.
 However, it is important to also consider that vaccination uptake can be
@@
 \toprule
-\textbf{Variable} & \textbf{no}, N = 1,547\textsuperscript{1} & \textbf{yes}, N = 4,689\textsuperscript{1} & \textbf{p-value}\textsuperscript{2} \\ 
+\textbf{Variable} & \textbf{no}, N = 1,547\textsuperscript{\textit{1}} & \textbf{yes}, N = 4,689\textsuperscript{\textit{1}} & \textbf{p-value}\textsuperscript{\textit{2}} \\ 
 \midrule
@@
     Mixed & 177 (30\%) & 411 (70\%) &  \\ 
-    Other\textsuperscript{3} & 315 (34\%) & 606 (66\%) &  \\ 
+    Other\textsuperscript{\textit{3}} & 315 (34\%) & 606 (66\%) &  \\ 
     South Asian & 126 (21\%) & 485 (79\%) &  \\ 
@@
 \begin{minipage}{\linewidth}
-\textsuperscript{1}n (\%)\\
-\textsuperscript{2}Pearson\textquotesingle{}s Chi-squared test\\
-\textsuperscript{3}Southeast Asian, Filipino, West Asian,
+\textsuperscript{\textit{1}}n (\%)\\
+\textsuperscript{\textit{2}}Pearson's Chi-squared test\\
+\textsuperscript{\textit{3}}Southeast Asian, Filipino, West Asian,
 and minorities not identified elsewhere according to the Census.\\
@@
 
-Figure~\ref{fig-model-uncorr} presents the estimates (as odd ratios)
-from the logistic regression models for vaccination status using the
+Figure~\ref{fig-models} presents the estimates (as odd ratios) from the
+logistic regression models for vaccination status using the
 socio-demographic factors collected by the survey, and their
@@
 with a household income under CAD 25,000, and Indigenous individuals
-living in the Central Health Region (Figure~\ref{fig-model-uncorr},B).
+living in the Central Health Region (Figure~\ref{fig-models},B).
 
@@
 
-\includegraphics{main_files/figure-pdf/fig-model-uncorr-1.pdf} \hfill{}
-
-\caption{\label{fig-model-uncorr}Coefficient estimates and confidence
+\includegraphics{main_files/figure-pdf/fig-models-1.pdf} \hfill{}
+
+\caption{\label{fig-models}Coefficient estimates and confidence
 intervals for the uncorrected model. Only statistically significant
@@
 
-\begin{figure}
-
-\includegraphics{main_files/figure-pdf/fig-model-corr-1.pdf} \hfill{}
-
-\caption{\label{fig-model-corr}Coefficient estimates and confidence
-intervals for the corrected model. Only statistically significant
-interaction terms are shown. Full interaction terms can be found in
-Supplementary Figure A-4.}
-
-\end{figure}
-
 \hypertarget{discussion}{%
@@
 
-Our results indicate that across the most densely populated Health
-Regions of Ontario, almost three quarters of surveyed individuals
-reported to have received the first dose of the COVID-19 vaccine
-(Table~\ref{tbl-descriptive-stats}). We identified significant
-intra-provincial differences in vaccination based on socio-economic and
-geographical factors. First, our results show differences in odds of
-vaccination in individuals with a household income below CAD 60,000 and
-in individuals belonging to visible minority groups. Those who
-identified as Arab/Middle Eastern, Black, Latin American, or that
-belonged to a minority group not included in the survey (Southeast
-Asian, Filipino, West Asian, and minority groups not identified
-elsewhere) had vaccination odds that were less than a third of
-individuals that identified as White/Caucasian
-(Figure~\ref{fig-model-corr}). These results are consistent with other
-studies that have shown lower vaccination rates in individuals that
-identify as part of a racial minority, or that have a low household
-income\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
-In the case of Ontario, a possible rationale for this difference is
-vaccine access, which has been identified as an important contributing
-factor of disparities in vaccination for equity-deserving
-groups\textsuperscript{\protect\hyperlink{ref-njoku2021}{49}}, and that
-occurred in Toronto earlier in 2021, where it has been shown that
-vaccination venues were scarce in low socio-economic areas that had the
-highest burden of COVID-19 at the
-time\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}.
-
-It is interesting to note that although overall self-reported
-vaccination rates were found to be statistically significantly lower in
-various racial minority groups when compared to White/Caucasian
-individuals, the change in odds of vaccination within certain racial
-groups and income strata was actually positive, in contrast to the
-White/Caucasian group, where vaccination odds decreased in lower income
-brackets (when compared to the CAD 60,000 and over bracket,
-Supplementary Figure A-5). More specifically, the change in odds of
-vaccination increased in individuals who identified as Arab/Middle
-Eastern, Black, Latin American, or belonging to other minority groups
-with a household income below CAD 25,000, which was also true for
-individuals in other racial minority groups with an income between CAD
-25,000-59,999.
+Our results show differences in odds of vaccination in individuals
+across certain socio-demographic groups. Specifically, those who
+identified as Arab/Middle Eastern, Black, Latin American, having mixed
+racial or ethnic background, or that belonged to other groups not
+explicitly included in the survey (Southeast Asian, Filipino, West
+Asian, and minority groups not identified elsewhere) had vaccination
+odds that were between a third and a half of that of individuals that
+identified as White/Caucasian (Figure~\ref{fig-models}). These results
+are consistent with previous studies that have shown lower vaccination
+rates in individuals with the same socio-demographic
+characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
+
+Lower vaccine uptake in the socio-demographic groups indicated above may
+be influenced in part, by vaccine hesitancy and refusal, which have been
+associated in equity-deserving Canadian individuals to concerns on
+vaccine safety, effectiveness, and experiences of racial discrimination
+in health
+settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{39},\protect\hyperlink{ref-basta2022}{49}--\protect\hyperlink{ref-cnat2023}{51}}.
+However, it has been shown that structural barriers also play an
+important role in vaccination uptake. In the case of equity-deserving
+individuals, such barriers include complex scheduling systems, language
+barriers, lack of adequate public transportation, and lack of accessible
+vaccination
+sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{52}}. In this
+regard, it is interesting to note that vaccination venues were scarce in
+low socio-economic areas that had the highest burden of COVID-19 in
+Toronto and other regions of Ontario around the time covered by the
+survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{53}},
+and that pharmacies in the Peel region (an area identified as a
+``hotspot'' with high numbers of essential workers and multigenerational
+households) could not keep up with
+demand\textsuperscript{\protect\hyperlink{ref-gill2022}{54}} . However,
+to the best of our knowledge, there seems to be a very limited amount of
+studies in this area, indicating an ongoing need of future studies that
+examine the longitudinal impact of vaccine accessibility and structural
+barriers that affected equity-deserving groups within the province.
+
+Although overall self-reported vaccination rates were found to be
+statistically significantly lower in various racial minority groups when
+compared to White/Caucasian individuals, the change in odds of
+vaccination within certain racial groups and income strata was actually
+positive, in contrast to the White/Caucasian group, where vaccination
+odds decreased in lower income brackets (when compared to the CAD 60,000
+and over bracket, Supplementary Figure A-5). Specifically, individuals
+in low income brackets that belonged to Arab/Middle Eastern, Black, or
+other minority groups had higher odds of vaccination that their peers
+with an income above 60,000 CAD.
 
@@
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{50},\protect\hyperlink{ref-ct2021}{51}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{55},\protect\hyperlink{ref-ct2021}{56}},
 which include grocery store workers, gas station, warehouse,
@@
 an income within the significant brackets identified in the analysis is
-to be expected. In the case of Ontario, essential workers had priority
-for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{52}},
-which would explain the higher odds of vaccination for these
-individuals. In other words, it is possible that the type of occupation
-from individuals in equity-deserving groups played an important role in
-increasing the odds of vaccination.
+to be expected. In Ontario, these workers had priority for COVID-19
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{57}}; and
+interestingly, there is evidence of interventions by vaccination staff
+in certain parts of the province to encourage vaccination uptake by
+these individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}.
+These facts, combined with evidence of increased trends in vaccination
+in this group
+elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{58}},
+suggest that the type of occupation from individuals in equity-deserving
+groups played an important role in increasing the odds of vaccination in
+the province.
 
@@
 West Health Region when compared to the Health Region of Toronto
-(Figure~\ref{fig-model-corr}). The West Health Region comprises the
+(\textbf{?@fig-model-corr}). The West Health Region comprises the
 regions of Waterloo and Niagara, the counties of Wellington, Essex and
@@
 identifying as South Asian or other minorities not included in the
-survey Figure~\ref{fig-model-corr}. In this case, the results of the
+survey \textbf{?@fig-model-corr}. In this case, the results of the
 interaction term in the model indicate that the odds of vaccination for
@@
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{53}}.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{59}}.
 Therefore, the estimated lower odds are likely to be explained from a
@@
 in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{54}}
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{60}}
 (Supplementary Table A-7). Interestingly, a disproportionate number of
@@
 reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{55}}. These
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{61}}. These
 results provide context to the observed differences in vaccination among
@@
 are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{53}}, a
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{59}}, a
 minority group that has historically suffered from reduced access to
 health care and
-discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{22}}.
+discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{21}}.
 Therefore, there is a need for additional studies that focus on these
@@
 reported by Public Health Ontario at the start of the period covered by
-the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{56}}.
+the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{62}}.
 However, we found good agreement between the estimates from the model
@@
 relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{57}}, and
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{63}}, and
 with vaccination uptake rates across different age groups presented in
 other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{58}}.
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{64}}.
 
@@
 (Table~\ref{tbl-descriptive-stats}) and vaccination odds
-(Figure~\ref{fig-model-corr}) across the period covered by the survey,
+(\textbf{?@fig-model-corr}) across the period covered by the survey,
 which closely match the vaccination rates from Public Health Ontario and
@@
 across all age groups during the three months covered by the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{56}}. It is
+survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{62}}. It is
 also important to mention that to this day, differences in vaccination
@@
 still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{59}}, and
+rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{65}}, and
 although data for the period analyzed in this study is not
@@
 documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{60}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{66}}, we
 believe that our data was able to provide a valid assessment of
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
+\CSLLeftMargin{21. }%
+\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
+roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
+canada. \emph{Canadian Medical Association Journal}.
+2021;193(11):E381-E383.
+doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
@@
 doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
-\CSLLeftMargin{22. }%
-\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
-roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
-canada. \emph{Canadian Medical Association Journal}.
-2021;193(11):E381-E383.
-doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-basta2022}{}}%
+\CSLLeftMargin{49. }%
+\CSLRightInline{Basta NE, Sohel N, Sulis G, et al. Factors associated
+with willingness to receive a {COVID}-19 vaccine among 23, 819 adults
+aged 50 years or older: An analysis of the canadian longitudinal study
+on aging. \emph{American Journal of Epidemiology}. 2022;191(6):987-998.
+doi:\href{https://doi.org/10.1093/aje/kwac029}{10.1093/aje/kwac029}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
+\CSLLeftMargin{50. }%
+\CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
+review on vaccine hesitancy in black communities in canada: Critical
+issues and research failures. \emph{Vaccines}. 2022;10(11):1937.
+doi:\href{https://doi.org/10.3390/vaccines10111937}{10.3390/vaccines10111937}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cnat2023}{}}%
+\CSLLeftMargin{51. }%
+\CSLRightInline{Cénat JM, Farahi SMMM, Bakombo SM, et al. Vaccine
+mistrust among black individuals in canada: The major role of health
+literacy, conspiracy theories, and racial discrimination in the
+healthcare system. \emph{Journal of Medical Virology}. 2023;95(4).
+doi:\href{https://doi.org/10.1002/jmv.28738}{10.1002/jmv.28738}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-iveniuk2021}{}}%
+\CSLLeftMargin{53. }%
+\CSLRightInline{Iveniuk J, Leon S. \emph{Uneven Recovery: Measuring
+Covid-19 Vaccine Equity in Ontario}. Wellesley Institute; 2021. Accessed
+May 12, 2023.
+\url{https://www.wellesleyinstitute.com/wp-content/uploads/2021/04/An-uneven-recovery-Measuring-COVID-19-vaccine-equity-in-Ontario.pdf}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gill2022}{}}%
+\CSLLeftMargin{54. }%
+\CSLRightInline{Gill M, Datta D, Gregory P, Austin Z. {COVID}-19
+vaccination in high-risk communities: Case study of brampton, ontario.
+\emph{Canadian Pharmacists Journal / Revue des Pharmaciens du Canada}.
+2022;155(6):345-351.
+doi:\href{https://doi.org/10.1177/17151635221123042}{10.1177/17151635221123042}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-nguyen2021b}{}}%
+\CSLLeftMargin{58. }%
+\CSLRightInline{Nguyen KH, Yankey D, Coy KC, et al. {COVID}-19
+vaccination coverage, intent, knowledge, attitudes, and beliefs among
+essential workers, united states. \emph{Emerging Infectious Diseases}.
+2021;27(11):2908-2913.
+doi:\href{https://doi.org/10.3201/eid2711.211557}{10.3201/eid2711.211557}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-05-14 code/data_preparation_May_08.R

@@ -267,7 +267,7 @@ clean_data$Month <- as.factor(clean_data$Month)
 clean_data <- clean_data%>%
   subset(Month!="September")
 
-clean_data$Month <- fct_relevel(clean_data$Month,levels=c("October","November","December","January"))
+clean_data$Month <- fct_relevel(clean_data$Month,"October","November","December","January")
 
 
 # Remove certain Health REgions due to low number of obs: only 254 in North East

### 2023-05-15 manuscript_v5/03-results.qmd
updated discussion

@@
 data_summary
-
-# test<-a1_rake$variables
-#
-
-# ds2<- a1_rake %>%
-#   tbl_svysummary(percent="row",
-#               by= first_dose,
-#               include=c(income_ord,
-#                         Age_group_ord,
-#                         Health_Region,
-#                         Month,
-#                         Race),
-#               label=list(Health_Region ~ "Health Region",
-#                  Age_group_ord ~ "Age Group",
-#                  income_ord="Income (CAD)")
-#   ) %>% add_p(test = everything () ~ "chisq.test")%>%
-#   bold_labels() %>%
-#   modify_table_styling(
-#     columns=label,
-#     rows=label=="Other",
-#     footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
-#   ) %>%
-#   modify_header(label = "**Variable**")
-# 
-# ds1<- a1 %>%
-#   tbl_svysummary(percent="row",
-#               by= first_dose,
-#               include=c(income_ord,
-#                         Age_group_ord,
-#                         Health_Region,
-#                         Month,
-#                         Race),
-#               label=list(Health_Region ~ "Health Region",
-#                  Age_group_ord ~ "Age Group",
-#                  income_ord="Income (CAD)")
-#   ) %>% add_p(test = everything () ~ "chisq.test")%>%
-#   bold_labels() %>%
-#   modify_table_styling(
-#     columns=label,
-#     rows=label=="Other",
-#     footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
-#   ) %>%
-#   modify_header(label = "**Variable**") 
-# 
-# 
-# tbl_merge(tbls=list(ds1,ds2)) %>%
-#   as_gt()
 
@@
 #table for comparison
-#tbl_merge(tbls = list(tbl_regression(m0,exponentiate=TRUE),tbl_regression(m2,exponentiate = TRUE)))
+# tbl_merge(tbls = list(tbl_regression(m0,exponentiate=TRUE),tbl_regression(m2,exponentiate = TRUE)))
+# 
+#  ds2<- a1_rake %>%
+#    tbl_svysummary(percent="row",
+#               by= first_dose,
+#               include=c(income_ord,
+#                          Age_group_ord,
+#                          Health_Region,
+#                          Month,
+#                         Race),
+#               label=list(Health_Region ~ "Health Region",                  Age_group_ord ~ "Age Group",
+#                   income_ord="Income (CAD)")
+#    ) %>% add_p(test = everything () ~ "chisq.test")%>%
+#   bold_labels() %>%
+#   modify_table_styling(
+#     columns=label,
+#      rows=label=="Other",
+#      footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
+#    ) %>%
+#   modify_header(label = "**Variable**")
+# 
+#  ds1<- a1 %>%
+#   tbl_svysummary(percent="row",
+#                by= first_dose,
+#               include=c(income_ord,
+#                          Age_group_ord,
+#                          Health_Region,
+#                          Month,
+#                          Race),
+#                label=list(Health_Region ~ "Health Region",
+#                   Age_group_ord ~ "Age Group",
+#                   income_ord="Income (CAD)")
+#   ) %>% add_p(test = everything () ~ "chisq.test")%>%
+#    bold_labels() %>%
+#    modify_table_styling(
+#      columns=label,
+#     rows=label=="Other",
+#      footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
+#    ) %>%
+#    modify_header(label = "**Variable**")
+# 
+# 
+#  tbl_merge(tbls=list(ds1,ds2)) %>%
+#    as_gt()
+#
+
 ```

### 2023-05-15 manuscript_v5/04-discussion.qmd
updated discussion

@@
 # Discussion
-
-West Health Region includes the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
-The Central Health Region includes the regions of Peel and Halton, the regional municipality of York, the district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of Etobicoke.
 
@@
 
-However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@fig-model-corr].
-In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-6).
+Additionally, the interaction effect of Health Region and race was also significant in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models].
+In these cases, vaccination odds were lower when compared to those of individuals in the Toronto Health Region (Supplementary Figure A-6).
+Although these findings are likely the result of multiple causes, we would like to indicate some contributing factors that might help provide context in each case.
+
+To contextualize these results, it is useful to disaggregate by LHINs, because in the literature, most of the studies that have analyzed health in Ontario use the LHINs as the base of their analyses.
+From this perspective, is interesting to note that most observations corresponding to Indigenous individuals and those with of mixed racial background in the West Health Region came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs, whereas for Arab/Middle Eastern Individuals in the East Health Region the LHINs with the highest number of observations corresponded to the Champlain and Central East.
+Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].
+
+Furthermore, the socio-demographic composition of the Health Regions is varied.
+For example, in the West Health Region Indigenous individuals and visible minorities account for 15.5\% of the population, whereas in the Central Health Region these groups account for 48\% of the population.
+Despite this, it seems that in both regions there continues to exist limitations  are only 3 Aboriginal Health Access Centres [@ontariohealth]
 
@@
 In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-6).
+
+I
+
+From one side, it is important to mention that historically, the social composition of the areas now covered by each Health Region (and previously by the LHINs) has been very different, and
+
+and we would like to focus on two areas that might help contextualize the observed reduced odds of vaccination. In the case of the West Health Region,
+
+The rationale behind these results its likely to be dual.
+
+The West Health Region is located in the leftmost southern part of the province, including the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+The Central Health Region includes the regions of Peel and Halton, the regional municipality of York, the district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of Etobicoke.
 

### 2023-05-15 manuscript_v5/main.tex
updated discussion

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-West Health Region includes the regions of Waterloo and Niagara, the
-counties of Wellington, Essex, and Lambton, and the cities of Hamilton,
-Haldimand, Brant, and Chatham-Kent. The Central Health Region includes
-the regions of Peel and Halton, the regional municipality of York, the
-district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the
-city of Etobicoke.
-
 had significantly higher odds of vaccination than those in the Health
@@
 
+Additionally, the interaction effect of Health Region and race was also
+significant in the case of individuals identifying as Indigenous or with
+mixed racial background in the West Health Region, Black individuals in
+the Central Health Region, and Arab/Middle Eastern individuals in the
+East Health Region Figure~\ref{fig-models}. In these cases, vaccination
+odds were lower when compared to those of individuals in the Toronto
+Health Region (Supplementary Figure A-6). Although these findings are
+likely the result of multiple causes, we would like to indicate some
+contributing factors that might help provide context in each case.
+
+First, it is interesting to note that theLHINs with most observations
+for Indigenous individuals and those with of mixed racial background in
+the West Health Region were the LHINs of Hamilton Niagara Haldimand
+Brant, South West, and Waterloo Wellington. Previous research has
+identified health disparities in these regions, such as unequal
+distribution of primary care providers, increased mortality, and low
+pharmacist
+availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.
+
+According to Ontario Health, 13.2\% of the population in the West Health
+Region identifies as a visible minority, whereas 2.5\% identifies as
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
+Therefore, the estimated lower odds are likely to be explained from a
+socio-economic perspective. In fact, 50\% of the answers from this
+region in the survey came from the former LHINs of Hamilton Niagara
+Haldimand Brant, and Erie St.~Clair, both which are among the regions of
+Ontario with the highest proportion of their population (more than 20\%)
+in the lowest income
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{63}}
+(Supplementary Table A-7). Interestingly, a disproportionate number of
+COVID-19 cases and low vaccination rate (under 50\%) have been also
+reported in the South Asian community of
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{64}}. These
+results provide context to the observed differences in vaccination among
+the Health Regions of the province, showing which equity-deserving and
+socio-demographic groups in certain regions need to be prioritized in
+future vaccination strategies. Moreover, these results provide a
+rationale for future studies that explore how vaccination uptake varies
+across different minority groups within Ontario and other Canadian
+provinces.
+
 Additionally, significant higher vaccination odds were identified in the
@@
 
+I
+
+From one side, it is important to mention that historically, the social
+composition of the areas now covered by each Health Region (and
+previously by the LHINs) has been very different, and
+
+and we would like to focus on two areas that might help contextualize
+the observed reduced odds of vaccination. In the case of the West Health
+Region,
+
+The rationale behind these results its likely to be dual.
+
+The West Health Region is located in the leftmost southern part of the
+province, including the regions of Waterloo and Niagara, the counties of
+Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
+Brant, and Chatham-Kent. The Central Health Region includes the regions
+of Peel and Halton, the regional municipality of York, the district of
+Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of
+Etobicoke.
+
 According to Ontario Health, 13.2\% of the population in the West Health
 Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{59}}.
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
 Therefore, the estimated lower odds are likely to be explained from a
@@
 in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{60}}
+quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{63}}
 (Supplementary Table A-7). Interestingly, a disproportionate number of
@@
 reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{61}}. These
+Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{64}}. These
 results provide context to the observed differences in vaccination among
@@
 are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{59}}, a
+Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}, a
 minority group that has historically suffered from reduced access to
@@
 reported by Public Health Ontario at the start of the period covered by
-the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{62}}.
+the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{65}}.
 However, we found good agreement between the estimates from the model
@@
 relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{63}}, and
+countries\textsuperscript{\protect\hyperlink{ref-dube2022}{66}}, and
 with vaccination uptake rates across different age groups presented in
 other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{64}}.
+studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{67}}.
 
@@
 across all age groups during the three months covered by the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{62}}. It is
+survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{65}}. It is
 also important to mention that to this day, differences in vaccination
@@
 still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{65}}, and
+rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{68}}, and
 although data for the period analyzed in this study is not
@@
 documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{66}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{69}}, we
 believe that our data was able to provide a valid assessment of
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-shah2019}{}}%
+\CSLLeftMargin{59. }%
+\CSLRightInline{Shah TI, Clark AF, Seabrook JA, Sibbald S, Gilliland JA.
+Geographic accessibility to primary care providers: Comparing rural and
+urban areas in southwestern ontario. \emph{The Canadian Geographer / Le
+G{é}ographe canadien}. 2019;64(1):65-78.
+doi:\href{https://doi.org/10.1111/cag.12557}{10.1111/cag.12557}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-crighton2015}{}}%
+\CSLLeftMargin{60. }%
+\CSLRightInline{Crighton EJ, Ragetlie R, Luo J, To T, Gershon A. A
+spatial analysis of {COPD} prevalence, incidence, mortality and health
+service use in ontario. \emph{Health Rep}. 2015;26(3):10-18.}
+
+\leavevmode\vadjust pre{\hypertarget{ref-timony2022}{}}%
+\CSLLeftMargin{61. }%
+\CSLRightInline{Timony P, Houle SKD, Gauthier A, Waite NM. Geographic
+distribution of ontario pharmacists: A focus on rural and northern
+communities. \emph{Canadian Pharmacists Journal / Revue des Pharmaciens
+du Canada}. 2022;155(5):267-276.
+doi:\href{https://doi.org/10.1177/17151635221115411}{10.1177/17151635221115411}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
@@
 \leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of

### 2023-05-15 code/complete_case_test.R

@@ -446,4 +446,41 @@ a1<-tbl_regression(m0,exponentiate=TRUE) %>%
 a2<-tbl_regression(m2,exponentiate = TRUE) %>% 
   bold_labels()
 
-tbl_merge(tbls=list(a1,a2), tab_spanner = c("Uncorrected","Corrected")) 
+tbl_merge(tbls=list(a1,a2), tab_spanner = c("Uncorrected","Corrected"))
+
+
+clean_data %>%
+  subset(Health_Region=="West" & Race=="Mixed")%>%
+  droplevels()%>%
+  tbl_summary(by= first_dose,
+              percent = "row",
+                 missing="no",
+                 sort = list(everything() ~ "frequency"),
+              include=c(income_ord,
+                        Age_group_ord,
+                        Health_Region,
+                        Month,
+                        Race,
+                        LHIN,
+                        city),
+              label=list(Health_Region ~ "Health Region",
+                         Age_group_ord ~ "Age Group",
+                         income_ord="Income (CAD)")
+  ) 
+
+a1_rake %>%
+  subset(Health_Region=="West" & Race=="Indigenous")%>%
+  tbl_svysummary(by= first_dose,
+              #missing="no",
+              sort = list(everything() ~ "frequency"),
+              include=c(income_ord,
+                        Age_group_ord,
+                        Health_Region,
+                        Month,
+                        Race,
+                        LHIN,
+                        city),
+              label=list(Health_Region ~ "Health Region",
+                         Age_group_ord ~ "Age Group",
+                         income_ord="Income (CAD)")
+  ) 

### 2023-05-16 manuscript_v5/01-background.qmd
updating discussion

@@
 
-Between 2007 and 2019, Ontario provided healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integrated Networks (LHINs).
-However, this approach was complex and bureaucratic, resulting in excessive expenditures, disparities in mortality rates, and the deterioration of certain performance indicators such as wait times and readmissions [@tsasis2012;@muratov2018;@dong2022].
-In an effort to improve the healthcare system of the province, in late 2019 the government dissolved the 14 LHINs  and the areas they covered were incorporated into 6 Health Regions (North East, North West, Central, Toronto, West, and East)  [@dong2022].
-Due to the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
+This need is specially important in the case of Ontario, the most populated province of Canada.
+Between 2007 and 2019, Ontario managed healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integration Networks (LHINs), which aimed to provide an integrated health system for the province.
+However, this approach was complex and bureaucratic, and resulted in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators such as wait times and hospital readmissions, fragmented electronic health systems, the decline of performance indicators, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016].
+Therefore, with the intent of better organizing and delivering care in late 2019 the provincial government eliminated the LHINs and incorporated the areas covered by them into six larger Health Regions (North East, North West, Central, Toronto, West, and East) [@dong2022].
+
+Because the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
 In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].

### 2023-05-16 manuscript_v5/04-discussion.qmd
updating discussion

--- 2023-05-14 manuscript_v5/04-discussion.qmd
@@
 
-Between 2007 and 2019, healthcare in Ontario was managed by 14 different LHINs, which aimed to provide an integrated health system for the province.
-However, this approach faced chronic problems such as fragmented electronic health systems [@lysyk2016], the decline of performance indicators (e.g., increased unplanned emergency visits, number of patient readmissions), and inequities in health services access [@lysyk2015].
-Based on these problematics and with the intent of better organizing and delivering care, the provincial government moved to the Health Region model in late 2019 [@dong2022].
+In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its missing of improving health access for the inhabitants of Ontario.
 
-In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its missing of improving health access for the inhabitants of Ontario.
-
-Our results show differences in odds of vaccination in individuals across certain socio-demographic groups.
+Our results show differences in odds of vaccination across Ontario in certain socio-demographic groups.
 Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models).
@@
 In the case of equity-deserving individuals, such barriers include complex scheduling systems, language barriers, lack of adequate public transportation, and lack of accessible vaccination sites [@njoku2021].
-In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with demand [@gill2022] .
-However, to the best of our knowledge, there seems to be a very limited amount of studies in this area, indicating an ongoing need of future studies that examine the longitudinal impact of vaccine accessibility and structural barriers that affected equity-deserving groups within the province.
+In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with demand [@gill2022].
+This suggests disparities in vaccine accessibility that affected in particular equity-deserving individuals in Ontario at the time of the survey.
+However, because to the best of our knowledge there seems to be a very limited amount of literature on this topic in the context of Ontario, there is an ongoing need of future studies that examine the longitudinal impact of vaccine accessibility and structural barriers that affect equity-deserving groups within the province.
 
-Although overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets (when compared to the CAD 60,000 and over bracket, Supplementary Figure A-5).
+Interestingly, whereas overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets when compared to the CAD 60,000 and over bracket (Supplementary Figure A-5).
 Specifically, individuals in low income brackets that belonged to Arab/Middle Eastern, Black, or other minority groups had higher odds of vaccination that their peers with an income above 60,000 CAD.
 
-This result is likely due in part to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include grocery store workers, gas station, warehouse, distribution, and manufacturing workers, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
-In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and interestingly, there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022].
+This result is likely reflects in part the fact that individuals in racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include grocery store, gas station, warehouse, distribution, and manufacturing workers, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
+In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022].
 These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation from individuals in equity-deserving groups played an important role in increasing the odds of vaccination in the province.
 
-However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@fig-model-corr].
-In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-6).
+However, the results also indicate that the place of habitation within the province for certain equity-deserving groups also affected the odds of vaccination (interaction terms of Health Region and Race, @fig-models).
+Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models].
+For these individuals, vaccination odds were lower when compared to the Toronto Health Region (Supplementary Figure A-6).
+Although it is likely that these findings have multiple causes, we indicate next some contributing factors that might help provide context in each case.
 
-According to Ontario Health, 13.2% of the population in the West Health Region identifies as a visible minority, whereas 2.5% identifies as Indigenous [@ontariohealth].
-Therefore, the estimated lower odds are likely to be explained from a socio-economic perspective.
-In fact, 50% of the answers from this region in the survey came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
-Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-7).
-Interestingly, a disproportionate number of COVID-19 cases and low vaccination rate (under 50%) have been also reported in the South Asian community of Ontario [@anand2022].
-These results provide context to the observed differences in vaccination among the Health Regions of the province, showing which equity-deserving and socio-demographic groups in certain regions need to be prioritized in future vaccination strategies.
-Moreover, these results provide a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.
+First, in the case of Indigenous, mixed, and Arab/Middle Eastern individuals is useful to disaggregate the data by LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses.
+Interestingly, for the first two groups most of the observations in the survey (for West Health Region) came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs.
+Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].
 
-Additionally, significant higher vaccination odds were identified in the West Health Region when compared to the Health Region of Toronto (@fig-model-corr).
-The West Health Region comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
-In this case, these results could reflect the fact that in the survey, about half of the entries for this Health Region corresponded to White/Caucasian individuals, who reported a vaccination rate of 83\% (Supplementary Table A-6).
-However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@fig-model-corr].
-In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-6).
+Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals.
+For example, the West Health Region has only two Aboriginal Health Access Centres in contrast to an Indigenous population of about 100,000 individuals [@ontariohealth].
+Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada where these rationales have been associated to lower vaccination among these groups [@smylie2022;@eissa2021].
+Taken together, these aspects indicate ongoing multiple healthcare disparities that continue to affect equity-deserving individuals across the province.
 
-According to Ontario Health, 13.2% of the population in the West Health Region identifies as a visible minority, whereas 2.5% identifies as Indigenous [@ontariohealth].
-Therefore, the estimated lower odds are likely to be explained from a socio-economic perspective.
-In fact, 50% of the answers from this region in the survey came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St.
-Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-7).
-Interestingly, a disproportionate number of COVID-19 cases and low vaccination rate (under 50%) have been also reported in the South Asian community of Ontario [@anand2022].
-These results provide context to the observed differences in vaccination among the Health Regions of the province, showing which equity-deserving and socio-demographic groups in certain regions need to be prioritized in future vaccination strategies.
-Moreover, these results provide a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.
+The West Health Region is located in the leftmost southern part of the province, including the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
+The Central Health Region includes the regions of Peel and Halton, the regional municipality of York, the district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of Etobicoke.
 
@@
 
-It is also worth mentioning that province-wide vaccination rates for the period of interest are somewhat different from those of the survey, particularly in the case of those 55 years of age and older, which in the survey had a vaccination rate of 72\%, against a rate of 88.4\% reported for the closest age bracket (50 years of age and older) reported by Public Health Ontario at the start of the period covered by the data [@ontario-covid].
-However, we found good agreement between the estimates from the model and overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], and with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022].
-
-In other words, although vaccination rates obtained from the survey were slightly lower than the provincial estimates, these values still represented a valid approximation to overall trends; this notion is reinforced by the consistency in the proportion of vaccination rates (@tbl-descriptive-stats) and vaccination odds (@fig-model-corr) across the period covered by the survey, which closely match the vaccination rates from Public Health Ontario and which indicate that due to the relatively high coverage achieved in the population at that point, vaccination rates increased by around 3% across all age groups during the three months covered by the survey [@ontario-covid].
-It is also important to mention that to this day, differences in vaccination rates within the province continue.
-As of March of 2023 some regions still have less than 75\% vaccination rate [@ontario-covid-map], and although data for the period analyzed in this study is not publicly-available, it is likely that differences in vaccination rates were higher at the time, being partially captured by the survey.
-
 The results in this study are based on self-reported data, where bias might be present.

### 2023-05-16 manuscript_v5/main.tex
updating discussion

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, interior hidden, sharp corners, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, interior hidden, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-Between 2007 and 2019, Ontario provided healthcare access to its
-inhabitants using 14 intra-provincial divisions called the Local Health
-Integrated Networks (LHINs). However, this approach was complex and
-bureaucratic, resulting in excessive expenditures, disparities in
-mortality rates, and the deterioration of certain performance indicators
-such as wait times and
-readmissions\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33}--\protect\hyperlink{ref-dong2022}{35}}.
-In an effort to improve the healthcare system of the province, in late
-2019 the government dissolved the 14 LHINs and the areas they covered
-were incorporated into 6 Health Regions (North East, North West,
-Central, Toronto, West, and
-East)\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}. Due to the
-relatively recent adoption of the Health Region model and its alignment
-with the onset of the COVID-19 pandemic, there is a need to analyze if
-there are ongoing disparities in health access under this approach that
-need to be addressed before they are exploited by a new disease or
-public health threat. In this regard, previous research has highlighted
-disparities in the level of activity of each Health
-Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{36}}.
+This need is specially important in the case of Ontario, the most
+populated province of Canada. Between 2007 and 2019, Ontario managed
+healthcare access to its inhabitants using 14 intra-provincial divisions
+called the Local Health Integration Networks (LHINs), which aimed to
+provide an integrated health system for the province. However, this
+approach was complex and bureaucratic, and resulted in excessive
+expenditures, disparities in mortality rates, the deterioration of
+certain performance indicators such as wait times and hospital
+readmissions, fragmented electronic health systems, the decline of
+performance indicators, and inequities in health services
+access\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33}--\protect\hyperlink{ref-lysyk2016}{37}}.
+Therefore, with the intent of better organizing and delivering care in
+late 2019 the provincial government eliminated the LHINs and
+incorporated the areas covered by them into six larger Health Regions
+(North East, North West, Central, Toronto, West, and
+East)\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}.
+
+Because the relatively recent adoption of the Health Region model and
+its alignment with the onset of the COVID-19 pandemic, there is a need
+to analyze if there are ongoing disparities in health access under this
+approach that need to be addressed before they are exploited by a new
+disease or public health threat. In this regard, previous research has
+highlighted disparities in the level of activity of each Health
+Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{38}}.
 Therefore, analyzing differences in vaccination uptake within the Health
@@
 of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{37}--\protect\hyperlink{ref-cnat2022a}{39}}.
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{39}--\protect\hyperlink{ref-cnat2022a}{41}}.
 Because we identified differences in representativity between the survey
@@
 proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{40}}
+(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{42}}
 to correct the data using data from the Census and Health Region
@@
 Appendix. All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{41}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{42}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{43}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{44}},
+\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{43}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{44}},
+\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{45}},
+\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{46}},
 and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{45}}.
+\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{47}}.
 
@@
 
-Between 2007 and 2019, healthcare in Ontario was managed by 14 different
-LHINs, which aimed to provide an integrated health system for the
-province. However, this approach faced chronic problems such as
-fragmented electronic health
-systems\textsuperscript{\protect\hyperlink{ref-lysyk2016}{46}}, the
-decline of performance indicators (e.g., increased unplanned emergency
-visits, number of patient readmissions), and inequities in health
-services access\textsuperscript{\protect\hyperlink{ref-lysyk2015}{47}}.
-Based on these problematics and with the intent of better organizing and
-delivering care, the provincial government moved to the Health Region
-model in late
-2019\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}.
-
 In this study we hypothesized that differences in COVID-19 vaccination
-uptake were present between the Health Regions, aiming at determining
-which socio-demographic groups could be impacted by these disparities in
-order to provide decision-makers with information that could be used to
-develop policies focused on reducing or eliminating these differences
-and ensure that the Health Region model is able to fulfill its missing
-of improving health access for the inhabitants of Ontario.
-
-Our results show differences in odds of vaccination in individuals
-across certain socio-demographic groups. Specifically, those who
-identified as Arab/Middle Eastern, Black, Latin American, having mixed
-racial or ethnic background, or that belonged to other groups not
-explicitly included in the survey (Southeast Asian, Filipino, West
-Asian, and minority groups not identified elsewhere) had vaccination
-odds that were between a third and a half of that of individuals that
-identified as White/Caucasian (Figure~\ref{fig-models}). These results
-are consistent with previous studies that have shown lower vaccination
-rates in individuals with the same socio-demographic
+uptake were present between the Health Regions during between late 2021
+and early 2022, aiming at determining which socio-demographic groups
+could be impacted by these disparities in order to provide
+decision-makers with information that could be used to develop policies
+focused on reducing or eliminating these differences and ensure that the
+Health Region model is able to fulfill its missing of improving health
+access for the inhabitants of Ontario.
+
+Our results show differences in odds of vaccination across Ontario in
+certain socio-demographic groups. Specifically, those who identified as
+Arab/Middle Eastern, Black, Latin American, having mixed racial or
+ethnic background, or that belonged to other groups not explicitly
+included in the survey (Southeast Asian, Filipino, West Asian, and
+minority groups not identified elsewhere) had vaccination odds that were
+between a third and a half of that of individuals that identified as
+White/Caucasian (Figure~\ref{fig-models}). These results are consistent
+with previous studies that have shown lower vaccination rates in
+individuals with the same socio-demographic
 characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
@@
 in health
-settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{39},\protect\hyperlink{ref-basta2022}{49}--\protect\hyperlink{ref-cnat2023}{51}}.
+settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{41},\protect\hyperlink{ref-basta2022}{49}--\protect\hyperlink{ref-cnat2023}{51}}.
 However, it has been shown that structural barriers also play an
@@
 households) could not keep up with
-demand\textsuperscript{\protect\hyperlink{ref-gill2022}{54}} . However,
-to the best of our knowledge, there seems to be a very limited amount of
-studies in this area, indicating an ongoing need of future studies that
-examine the longitudinal impact of vaccine accessibility and structural
-barriers that affected equity-deserving groups within the province.
-
-Although overall self-reported vaccination rates were found to be
-statistically significantly lower in various racial minority groups when
-compared to White/Caucasian individuals, the change in odds of
-vaccination within certain racial groups and income strata was actually
-positive, in contrast to the White/Caucasian group, where vaccination
-odds decreased in lower income brackets (when compared to the CAD 60,000
-and over bracket, Supplementary Figure A-5). Specifically, individuals
-in low income brackets that belonged to Arab/Middle Eastern, Black, or
-other minority groups had higher odds of vaccination that their peers
-with an income above 60,000 CAD.
-
-This result is likely due in part to the fact that individuals that
-belong racial minority groups tend to perform occupations that have been
-deemed as ``essential'' in the context of the
+demand\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. This
+suggests disparities in vaccine accessibility that affected in
+particular equity-deserving individuals in Ontario at the time of the
+survey. However, because to the best of our knowledge there seems to be
+a very limited amount of literature on this topic in the context of
+Ontario, there is an ongoing need of future studies that examine the
+longitudinal impact of vaccine accessibility and structural barriers
+that affect equity-deserving groups within the province.
+
+Interestingly, whereas overall self-reported vaccination rates were
+found to be statistically significantly lower in various racial minority
+groups when compared to White/Caucasian individuals, the change in odds
+of vaccination within certain racial groups and income strata was
+actually positive, in contrast to the White/Caucasian group, where
+vaccination odds decreased in lower income brackets when compared to the
+CAD 60,000 and over bracket (Supplementary Figure A-5). Specifically,
+individuals in low income brackets that belonged to Arab/Middle Eastern,
+Black, or other minority groups had higher odds of vaccination that
+their peers with an income above 60,000 CAD.
+
+This result is likely reflects in part the fact that individuals in
+racial minority groups tend to perform occupations that have been deemed
+as ``essential'' in the context of the
 pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{55},\protect\hyperlink{ref-ct2021}{56}},
-which include grocery store workers, gas station, warehouse,
-distribution, and manufacturing workers, all being occupations for which
-an income within the significant brackets identified in the analysis is
-to be expected. In Ontario, these workers had priority for COVID-19
+which include grocery store, gas station, warehouse, distribution, and
+manufacturing workers, all being occupations for which an income within
+the significant brackets identified in the analysis is to be expected.
+In Ontario, these workers had priority for COVID-19
 vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{57}}; and
-interestingly, there is evidence of interventions by vaccination staff
-in certain parts of the province to encourage vaccination uptake by
-these individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}.
-These facts, combined with evidence of increased trends in vaccination
-in this group
+there is evidence of interventions by vaccination staff in certain parts
+of the province to encourage vaccination uptake by these
+individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. These
+facts, combined with evidence of increased trends in vaccination in this
+group
 elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{58}},
@@
 
-Additionally, the interaction effect of Health Region and race was also
-significant in the case of individuals identifying as Indigenous or with
-mixed racial background in the West Health Region, Black individuals in
-the Central Health Region, and Arab/Middle Eastern individuals in the
-East Health Region Figure~\ref{fig-models}. In these cases, vaccination
-odds were lower when compared to those of individuals in the Toronto
-Health Region (Supplementary Figure A-6). Although these findings are
-likely the result of multiple causes, we would like to indicate some
-contributing factors that might help provide context in each case.
-
-First, it is interesting to note that theLHINs with most observations
-for Indigenous individuals and those with of mixed racial background in
-the West Health Region were the LHINs of Hamilton Niagara Haldimand
-Brant, South West, and Waterloo Wellington. Previous research has
-identified health disparities in these regions, such as unequal
-distribution of primary care providers, increased mortality, and low
-pharmacist
+However, the results also indicate that the place of habitation within
+the province for certain equity-deserving groups also affected the odds
+of vaccination (interaction terms of Health Region and Race,
+Figure~\ref{fig-models}). Specifically, this held true in the case of
+individuals identifying as Indigenous or with mixed racial background in
+the West Health Region, Black individuals in the Central Health Region,
+and Arab/Middle Eastern individuals in the East Health Region
+Figure~\ref{fig-models}. For these individuals, vaccination odds were
+lower when compared to the Toronto Health Region (Supplementary Figure
+A-6). Although it is likely that these findings have multiple causes, we
+indicate next some contributing factors that might help provide context
+in each case.
+
+First, in the case of Indigenous, mixed, and Arab/Middle Eastern
+individuals is useful to disaggregate the data by LHINs, because most of
+the studies in the literature that have analyzed health in Ontario use
+the LHINs as the base of their analyses. Interestingly, for the first
+two groups most of the observations in the survey (for West Health
+Region) came from the Hamilton Niagara Haldimand Brant, South West, and
+Waterloo Wellington LHINs (\textbf{add table in Appendix}), whereas for
+Arab/Middle Eastern Individuals in the East Health Region the highest
+number of observations corresponded to the Champlain and Central East
+LHINs. Previous research has identified health disparities in these
+(mostly rural) regions, such as unequal distribution of primary care
+providers, increased mortality, and low pharmacist
 availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.
 
-According to Ontario Health, 13.2\% of the population in the West Health
-Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
-Therefore, the estimated lower odds are likely to be explained from a
-socio-economic perspective. In fact, 50\% of the answers from this
-region in the survey came from the former LHINs of Hamilton Niagara
-Haldimand Brant, and Erie St.~Clair, both which are among the regions of
-Ontario with the highest proportion of their population (more than 20\%)
-in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{63}}
-(Supplementary Table A-7). Interestingly, a disproportionate number of
-COVID-19 cases and low vaccination rate (under 50\%) have been also
-reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{64}}. These
-results provide context to the observed differences in vaccination among
-the Health Regions of the province, showing which equity-deserving and
-socio-demographic groups in certain regions need to be prioritized in
-future vaccination strategies. Moreover, these results provide a
-rationale for future studies that explore how vaccination uptake varies
-across different minority groups within Ontario and other Canadian
-provinces.
-
-Additionally, significant higher vaccination odds were identified in the
-West Health Region when compared to the Health Region of Toronto
-(\textbf{?@fig-model-corr}). The West Health Region comprises the
-regions of Waterloo and Niagara, the counties of Wellington, Essex and
-Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
-In this case, these results could reflect the fact that in the survey,
-about half of the entries for this Health Region corresponded to
-White/Caucasian individuals, who reported a vaccination rate of 83\%
-(Supplementary Table A-6). However, the interaction effect of Health
-Region and race was also significant in the case of individuals
-identifying as South Asian or other minorities not included in the
-survey \textbf{?@fig-model-corr}. In this case, the results of the
-interaction term in the model indicate that the odds of vaccination for
-those within the South Asian and Other minority groups in the West
-Region decreased when compared to the other Health Regions
-(Supplementary Figure A-6).
-
-I
-
-From one side, it is important to mention that historically, the social
-composition of the areas now covered by each Health Region (and
-previously by the LHINs) has been very different, and
-
-and we would like to focus on two areas that might help contextualize
-the observed reduced odds of vaccination. In the case of the West Health
-Region,
-
-The rationale behind these results its likely to be dual.
+Furthermore, there is an ongoing challenge for the health system of the
+province with regard to personalized healthcare for marginalized
+individuals. For example, the West Health Region has only two Aboriginal
+Health Access Centres in contrast to an Indigenous population of about
+100,000
+individuals\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
+Lack of access to personalized healthcare affects individuals that may
+mistrust the traditional healthcare system due to systemic racism or
+oppression, which is known to be the case for Indigenous and Black
+individuals in Canada where these rationales have been associated to
+lower vaccination among these
+groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{63},\protect\hyperlink{ref-eissa2021}{64}}.
+Taken together, these aspects indicate ongoing multiple healthcare
+disparities that continue to affect equity-deserving individuals across
+the province.
 
@@
 Etobicoke.
-
-According to Ontario Health, 13.2\% of the population in the West Health
-Region identifies as a visible minority, whereas 2.5\% identifies as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
-Therefore, the estimated lower odds are likely to be explained from a
-socio-economic perspective. In fact, 50\% of the answers from this
-region in the survey came from the former LHINs of Hamilton Niagara
-Haldimand Brant, and Erie St.~Clair, both which are among the regions of
-Ontario with the highest proportion of their population (more than 20\%)
-in the lowest income
-quintile\textsuperscript{\protect\hyperlink{ref-buajitti2018}{63}}
-(Supplementary Table A-7). Interestingly, a disproportionate number of
-COVID-19 cases and low vaccination rate (under 50\%) have been also
-reported in the South Asian community of
-Ontario\textsuperscript{\protect\hyperlink{ref-anand2022}{64}}. These
-results provide context to the observed differences in vaccination among
-the Health Regions of the province, showing which equity-deserving and
-socio-demographic groups in certain regions need to be prioritized in
-future vaccination strategies. Moreover, these results provide a
-rationale for future studies that explore how vaccination uptake varies
-across different minority groups within Ontario and other Canadian
-provinces.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lysyk2015}{}}%
+\CSLLeftMargin{36. }%
+\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
+2015. In: \emph{Section 3.08: LHINs - Local Health Integration
+Networks}. Queen's Printer for Ontario; 2015. Accessed May 12, 2023.
+\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en15/3.08en15.pdf}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-lysyk2016}{}}%
+\CSLLeftMargin{37. }%
+\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
+2016. In: \emph{Section 3.03: Electronic Health Records' Implementation
+Status}. Queen's Printer for Ontario; 2016. Accessed May 12, 2023.
+\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en16/v1_303en16.pdf}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-sethuram2023}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Sethuram C, McCutcheon T, Liddy C. An environmental scan
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
@@
 \leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
@@
 \leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
@@
 doi:\href{https://doi.org/10.32614/RJ-2021-053}{10.32614/RJ-2021-053}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lysyk2016}{}}%
-\CSLLeftMargin{46. }%
-\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
-2016. In: \emph{Section 3.03: Electronic Health Records' Implementation
-Status}. Queen's Printer for Ontario; 2016. Accessed May 12, 2023.
-\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en16/v1_303en16.pdf}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lysyk2015}{}}%
-\CSLLeftMargin{47. }%
-\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
-2015. In: \emph{Section 3.08: LHINs - Local Health Integration
-Networks}. Queen's Printer for Ontario; 2015. Accessed May 12, 2023.
-\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en15/3.08en15.pdf}}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-buajitti2018}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
 \CSLLeftMargin{63. }%
-\CSLRightInline{Buajitti E, Watson T, Kornas K, Bornbaum C, Henry D,
-Rosella LC. Ontario atlas of adult mortality, 1992-2015: Trends in
-{L}ocal {H}ealth {I}ntegration {N}etworks. Published online 2018.
-\url{https://tspace.library.utoronto.ca/handle/1807/82836}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-anand2022}{}}%
+\CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
+{SARS}-{COV}-2 vaccine uptake and {COVID}-19 impacts among first
+nations, inuit and m{é}tis peoples living in toronto and london,
+ontario. \emph{Canadian Medical Association Journal}.
+2022;194(29):E1018-E1026.
+doi:\href{https://doi.org/10.1503/cmaj.212147}{10.1503/cmaj.212147}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
 \CSLLeftMargin{64. }%
-\CSLRightInline{Anand SS, Arnold C, Bangdiwala SI, et al. Seropositivity
-and risk factors for {SARS}-{CoV}-2 infection in a south asian community
-in ontario: A cross-sectional analysis of a prospective cohort study.
-\emph{{CMAJ} Open}. 2022;10(3):E599-E609.
-doi:\href{https://doi.org/10.9778/cmajo.20220031}{10.9778/cmajo.20220031}}
+\CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
+Increasing {SARS}-{CoV}-2 vaccination rates among black people in
+canada. \emph{Canadian Medical Association Journal}.
+2021;193(31):E1220-E1221.
+doi:\href{https://doi.org/10.1503/cmaj.210949}{10.1503/cmaj.210949}}
 

### 2023-05-16 code/complete_case_test.R

@@ -450,7 +450,7 @@ tbl_merge(tbls=list(a1,a2), tab_spanner = c("Uncorrected","Corrected"))
 
 
 clean_data %>%
-  subset(Health_Region=="West" & Race=="Mixed")%>%
+  subset(Health_Region=="Central" & Race=="Black")%>%
   droplevels()%>%
   tbl_summary(by= first_dose,
               percent = "row",

### 2023-05-17 manuscript_v5/04-discussion.qmd
discussion updated

@@
 
-had significantly higher odds of vaccination than those in the Health Region of Toronto
+In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its mission of improving health access for the inhabitants of Ontario.
 
-In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its missing of improving health access for the inhabitants of Ontario.
-
-Our results show differences in odds of vaccination across Ontario in certain socio-demographic groups.
+Our results show that indeed, there were differences in vaccination odds across Ontario in certain socio-demographic groups.
 Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models).
@@
 
-However, the results also indicate that the place of habitation within the province for certain equity-deserving groups also affected the odds of vaccination (interaction terms of Health Region and Race, @fig-models).
+However, the results also indicate that the place of habitation affected the odds of vaccination for certain equity-deserving groups (interaction term of Health Region and Race, @fig-models).
 Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models].
 For these individuals, vaccination odds were lower when compared to the Toronto Health Region (Supplementary Figure A-6).
-Although it is likely that these findings have multiple causes, we indicate next some contributing factors that might help provide context in each case.
+We indicate next some contributing factors that might help provide context in each case.
 
-First, in the case of Indigenous, mixed, and Arab/Middle Eastern individuals is useful to disaggregate the data by LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses.
-Interestingly, for the first two groups most of the observations in the survey (for West Health Region) came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs.
+First, it is useful to analyze the data using the LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses.
+Interestingly, for Indigenous and mixed individuals most of the observations in the survey for West Health Region came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs.
 Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].
@@
 Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals.
-For example, the West Health Region has only two Aboriginal Health Access Centres in contrast to an Indigenous population of about 100,000 individuals [@ontariohealth].
-Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada where these rationales have been associated to lower vaccination among these groups [@smylie2022;@eissa2021].
-Taken together, these aspects indicate ongoing multiple healthcare disparities that continue to affect equity-deserving individuals across the province.
-
-The West Health Region is located in the leftmost southern part of the province, including the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent.
-The Central Health Region includes the regions of Peel and Halton, the regional municipality of York, the district of Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of Etobicoke.
+For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated 100,000 Indigenous individuals living in the area [@ontariohealth].
+Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada.
+Indeed, these rationales have been associated to observed lower vaccination rates among these groups [@smylie2022;@eissa2021].
+Taken together, this suggests that healthcare disparities that are specific to certain socio-demographic groups are associated with lower vaccination uptake.
 
 There are some limitations to the present study.
-First, the data collection design, which allowed respondents to withdraw from the survey at any point, resulted in a high number of unique entries in the survey with multiple missing answers.
-Because we focused on entries that had complete observations in the covariates of interest for our analysis, it is possible that some information was not considered by excluding observations that had information in other variables (such as work from home, or number of persons in the household).
-However, we attempted to minimize this possibility by correcting the dataset using information from the Census.
-More granular corrections, which for example could be based on demographic information by municipality, could be used in the future to obtain a more accurate approximation to the population totals of the province.
-Moreover, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map).
-Although low representation from these areas is based on the fact that these regions only account for 5\% of the total population of Ontario, these regions are the home to more than 100,000 individuals that identify as Indigenous [@ontariohealth], a minority group that has historically suffered from reduced access to health care and discrimination [@mosby2021].
-Therefore, there is a need for additional studies that focus on these low-populated Health Regions in Ontario where disparities in vaccination might be significant and understudied.
+First, the data collection design, which allowed respondents to withdraw from the survey at any point, and that deployed the questions in a random manner resulted in an elevated number of missing observations without a definite pattern and complicated the implementation of sensitivity analyses.
+Therefore, we focused on entries that had complete answers, and corrected the data using population-wide information from the Census.
+However, more granular corrections would be needed to obtain more accurate estimates.
+For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates that other racial groups.
+Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates.
+Unfortunately, at the moment this correction cannot be implemented as such stratification is not currently available in the Census data.
+
+Additionally, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map).
+Low representation is expected as these regions only account for 5\% of the total population of Ontario, but in contrast, they have the highest proportion of Indigenous inhabitants [@ontariohealth].
+In the context of personalized care, there is a need for collecting data that focuses on these Health Regions where additional health disparities might be present and possibly understudied.
 
@@
 However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid assessment of vaccination in the province.
-Finally, this study focused on first-dose vaccination status  within a relatively short time window, and therefore can only provide a snapshot of the societal dynamics behind the pandemic.
+Finally, although higher vaccination odds identified for January of 2022 in the model are in accordance with province-wide trends reported by Public Health Ontario (which show a 4\% increase between early December and January, in contrast to a 2.5\% increase between October and November [@ontario-covid]), these results are only a snapshot when compared to the overall duration of the pandemic.
+
 Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.

### 2023-05-17 manuscript_v5/main.tex
discussion updated

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, interior hidden, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, sharp corners, interior hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-had significantly higher odds of vaccination than those in the Health
-Region of Toronto
-
 In this study we hypothesized that differences in COVID-19 vaccination
@@
 focused on reducing or eliminating these differences and ensure that the
-Health Region model is able to fulfill its missing of improving health
+Health Region model is able to fulfill its mission of improving health
 access for the inhabitants of Ontario.
 
-Our results show differences in odds of vaccination across Ontario in
-certain socio-demographic groups. Specifically, those who identified as
-Arab/Middle Eastern, Black, Latin American, having mixed racial or
-ethnic background, or that belonged to other groups not explicitly
-included in the survey (Southeast Asian, Filipino, West Asian, and
-minority groups not identified elsewhere) had vaccination odds that were
-between a third and a half of that of individuals that identified as
-White/Caucasian (Figure~\ref{fig-models}). These results are consistent
-with previous studies that have shown lower vaccination rates in
-individuals with the same socio-demographic
+Our results show that indeed, there were differences in vaccination odds
+across Ontario in certain socio-demographic groups. Specifically, those
+who identified as Arab/Middle Eastern, Black, Latin American, having
+mixed racial or ethnic background, or that belonged to other groups not
+explicitly included in the survey (Southeast Asian, Filipino, West
+Asian, and minority groups not identified elsewhere) had vaccination
+odds that were between a third and a half of that of individuals that
+identified as White/Caucasian (Figure~\ref{fig-models}). These results
+are consistent with previous studies that have shown lower vaccination
+rates in individuals with the same socio-demographic
 characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
@@
 
-However, the results also indicate that the place of habitation within
-the province for certain equity-deserving groups also affected the odds
-of vaccination (interaction terms of Health Region and Race,
-Figure~\ref{fig-models}). Specifically, this held true in the case of
-individuals identifying as Indigenous or with mixed racial background in
-the West Health Region, Black individuals in the Central Health Region,
-and Arab/Middle Eastern individuals in the East Health Region
-Figure~\ref{fig-models}. For these individuals, vaccination odds were
-lower when compared to the Toronto Health Region (Supplementary Figure
-A-6). Although it is likely that these findings have multiple causes, we
-indicate next some contributing factors that might help provide context
-in each case.
-
-First, in the case of Indigenous, mixed, and Arab/Middle Eastern
-individuals is useful to disaggregate the data by LHINs, because most of
+However, the results also indicate that the place of habitation affected
+the odds of vaccination for certain equity-deserving groups (interaction
+term of Health Region and Race, Figure~\ref{fig-models}). Specifically,
+this held true in the case of individuals identifying as Indigenous or
+with mixed racial background in the West Health Region, Black
+individuals in the Central Health Region, and Arab/Middle Eastern
+individuals in the East Health Region Figure~\ref{fig-models}. For these
+individuals, vaccination odds were lower when compared to the Toronto
+Health Region (Supplementary Figure A-6). We indicate next some
+contributing factors that might help provide context in each case.
+
+First, it is useful to analyze the data using the LHINs, because most of
 the studies in the literature that have analyzed health in Ontario use
-the LHINs as the base of their analyses. Interestingly, for the first
-two groups most of the observations in the survey (for West Health
-Region) came from the Hamilton Niagara Haldimand Brant, South West, and
-Waterloo Wellington LHINs (\textbf{add table in Appendix}), whereas for
-Arab/Middle Eastern Individuals in the East Health Region the highest
-number of observations corresponded to the Champlain and Central East
-LHINs. Previous research has identified health disparities in these
-(mostly rural) regions, such as unequal distribution of primary care
-providers, increased mortality, and low pharmacist
+the LHINs as the base of their analyses. Interestingly, for Indigenous
+and mixed individuals most of the observations in the survey for West
+Health Region came from the Hamilton Niagara Haldimand Brant, South
+West, and Waterloo Wellington LHINs (\textbf{add table in Appendix}),
+whereas for Arab/Middle Eastern Individuals in the East Health Region
+the highest number of observations corresponded to the Champlain and
+Central East LHINs. Previous research has identified health disparities
+in these (mostly rural) regions, such as unequal distribution of primary
+care providers, increased mortality, and low pharmacist
 availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.
@@
 individuals. For example, the West Health Region has only two Aboriginal
-Health Access Centres in contrast to an Indigenous population of about
-100,000
-individuals\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
-Lack of access to personalized healthcare affects individuals that may
-mistrust the traditional healthcare system due to systemic racism or
-oppression, which is known to be the case for Indigenous and Black
-individuals in Canada where these rationales have been associated to
-lower vaccination among these
+Health Access Centres (community-led primary healthcare organizations
+focused on First Nations, Métis, and Inuit communities) to provide care
+to an estimated 100,000 Indigenous individuals living in the
+area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}. Lack of
+access to personalized healthcare affects individuals that may mistrust
+the traditional healthcare system due to systemic racism or oppression,
+which is known to be the case for Indigenous and Black individuals in
+Canada. Indeed, these rationales have been associated to observed lower
+vaccination rates among these
 groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{63},\protect\hyperlink{ref-eissa2021}{64}}.
-Taken together, these aspects indicate ongoing multiple healthcare
-disparities that continue to affect equity-deserving individuals across
-the province.
-
-The West Health Region is located in the leftmost southern part of the
-province, including the regions of Waterloo and Niagara, the counties of
-Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand,
-Brant, and Chatham-Kent. The Central Health Region includes the regions
-of Peel and Halton, the regional municipality of York, the district of
-Muskoka, the counties of Dufferin, Simcoe, Grey, and the city of
-Etobicoke.
+Taken together, this suggests that healthcare disparities that are
+specific to certain socio-demographic groups are associated with lower
+vaccination uptake.
 
@@
 collection design, which allowed respondents to withdraw from the survey
-at any point, resulted in a high number of unique entries in the survey
-with multiple missing answers. Because we focused on entries that had
-complete observations in the covariates of interest for our analysis, it
-is possible that some information was not considered by excluding
-observations that had information in other variables (such as work from
-home, or number of persons in the household). However, we attempted to
-minimize this possibility by correcting the dataset using information
-from the Census. More granular corrections, which for example could be
-based on demographic information by municipality, could be used in the
-future to obtain a more accurate approximation to the population totals
-of the province. Moreover, our analysis did not consider the North West
-and North East Health Regions, due to the low number of entries from
-these areas in the survey (Figure~\ref{fig-map}). Although low
-representation from these areas is based on the fact that these regions
-only account for 5\% of the total population of Ontario, these regions
-are the home to more than 100,000 individuals that identify as
-Indigenous\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}, a
-minority group that has historically suffered from reduced access to
-health care and
-discrimination\textsuperscript{\protect\hyperlink{ref-mosby2021}{21}}.
-Therefore, there is a need for additional studies that focus on these
-low-populated Health Regions in Ontario where disparities in vaccination
-might be significant and understudied.
-
-It is also worth mentioning that province-wide vaccination rates for the
-period of interest are somewhat different from those of the survey,
-particularly in the case of those 55 years of age and older, which in
-the survey had a vaccination rate of 72\%, against a rate of 88.4\%
-reported for the closest age bracket (50 years of age and older)
-reported by Public Health Ontario at the start of the period covered by
-the data\textsuperscript{\protect\hyperlink{ref-ontario-covid}{65}}.
-However, we found good agreement between the estimates from the model
-and overall vaccination rates reported for Canada, which have been
-relatively higher when compared to other high income
-countries\textsuperscript{\protect\hyperlink{ref-dube2022}{66}}, and
-with vaccination uptake rates across different age groups presented in
-other
-studies\textsuperscript{\protect\hyperlink{ref-guay2022}{18},\protect\hyperlink{ref-macdonald2021}{67}}.
-
-In other words, although vaccination rates obtained from the survey were
-slightly lower than the provincial estimates, these values still
-represented a valid approximation to overall trends; this notion is
-reinforced by the consistency in the proportion of vaccination rates
-(Table~\ref{tbl-descriptive-stats}) and vaccination odds
-(\textbf{?@fig-model-corr}) across the period covered by the survey,
-which closely match the vaccination rates from Public Health Ontario and
-which indicate that due to the relatively high coverage achieved in the
-population at that point, vaccination rates increased by around 3\%
-across all age groups during the three months covered by the
-survey\textsuperscript{\protect\hyperlink{ref-ontario-covid}{65}}. It is
-also important to mention that to this day, differences in vaccination
-rates within the province continue. As of March of 2023 some regions
-still have less than 75\% vaccination
-rate\textsuperscript{\protect\hyperlink{ref-ontario-covid-map}{68}}, and
-although data for the period analyzed in this study is not
-publicly-available, it is likely that differences in vaccination rates
-were higher at the time, being partially captured by the survey.
+at any point, and that deployed the questions in a random manner
+resulted in an elevated number of missing observations without a
+definite pattern and complicated the implementation of sensitivity
+analyses. Therefore, we focused on entries that had complete answers,
+and corrected the data using population-wide information from the
+Census. However, more granular corrections would be needed to obtain
+more accurate estimates. For example, our analysis identified higher
+odds of vaccination in the Central and West Health Regions, but in this
+case these differences are likely to be driven by the proportion of
+White/Caucasian individuals, who had higher vaccination rates that other
+racial groups. Correcting for each racial/ethnic group in each Health
+Region can provide a more accurate estimation of region-wide vaccination
+rates. Unfortunately, at the moment this correction cannot be
+implemented as such stratification is not currently available in the
+Census data.
+
+Additionally, our analysis did not consider the North West and North
+East Health Regions, due to the low number of entries from these areas
+in the survey (Figure~\ref{fig-map}). Low representation is expected as
+these regions only account for 5\% of the total population of Ontario,
+but in contrast, they have the highest proportion of Indigenous
+inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
+In the context of personalized care, there is a need for collecting data
+that focuses on these Health Regions where additional health disparities
+might be present and possibly understudied.
 
@@
 documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{69}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{65}}, we
 believe that our data was able to provide a valid assessment of
-vaccination in the province. Finally, this study focused on first-dose
-vaccination status within a relatively short time window, and therefore
-can only provide a snapshot of the societal dynamics behind the
-pandemic. Nonetheless, the results presented here can serve as a
-starting point to motivate the collection of robust longitudinal data
-that can be used to quantify geographical and temporal differences
-within vulnerable segments of the population, and that can be used to
-inform the development of adequate public health policies within the
-province of Ontario or across other provinces that aim to minimize
-disparities in health access.
+vaccination in the province. Finally, although higher vaccination odds
+identified for January of 2022 in the model are in accordance with
+province-wide trends reported by Public Health Ontario (which show a 4\%
+increase between early December and January, in contrast to a 2.5\%
+increase between October and
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{66}}),
+these results are only a snapshot when compared to the overall duration
+of the pandemic.
+
+Nonetheless, the results presented here can serve as a starting point to
+motivate the collection of robust longitudinal data that can be used to
+quantify geographical and temporal differences within vulnerable
+segments of the population, and that can be used to inform the
+development of adequate public health policies within the province of
+Ontario or across other provinces that aim to minimize disparities in
+health access.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
 \CSLLeftMargin{65. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
-2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dube2022}{}}%
-\CSLLeftMargin{66. }%
-\CSLRightInline{Dubé E, Gagnon D, MacDonald N. Between persuasion and
-compulsion: The case of {COVID}-19 vaccination in canada.
-\emph{Vaccine}. 2022;40(29):3923-3926.
-doi:\href{https://doi.org/10.1016/j.vaccine.2022.05.053}{10.1016/j.vaccine.2022.05.053}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-macdonald2021}{}}%
-\CSLLeftMargin{67. }%
-\CSLRightInline{MacDonald NE, Comeau J, Dubé Ève, et al. Royal society
-of canada {COVID}-19 report: Enhancing {COVID}-19 vaccine acceptance in
-canada. Blais JM, ed. \emph{{FACETS}}. 2021;6:1184-1246.
-doi:\href{https://doi.org/10.1139/facets-2021-0037}{10.1139/facets-2021-0037}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid-map}{}}%
-\CSLLeftMargin{68. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed March 24, 2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=maps}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{69. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
+\CSLLeftMargin{66. }%
+\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
+2023.
+\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
+
 \end{CSLReferences}

### 2023-05-18 manuscript_v5/02-methods.qmd
discussion, appendix updated

@@
 The survey collected socio-economic information from participants (@tbl-descriptive-stats), their location (nearest municipality,  as shown in @fig-map), the date of access to the survey, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".
-The original dataset contained 39,029 observations (where each observation corresponded to a unique respondent).
+The original dataset contained 39,029 observations.
 
-Preliminary analyses of the data included the removal of outliers (**should we still do this?
-it's only 19 observations with income >110k and household of 1, but we are not even using such income bracket in the analysis because we re-grouped the data, and the household size variable has 90% missing rate**), of observations where respondents did not provide answers in all the covariates of interest, matching the city of each observations with its corresponding LHIN and Health Region, and removing observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions).
-After all the preliminary analyses indicated above, the total number of observations used for analysis was 3,549 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and December 12th, 2021.
+By design, the survey allowed respondents to exit at any time and deployed the questions randomly, which resulted in $\approx$ 84\% of the observations having multiple missing answers or being incomplete.
+Therefore, we selected 6,343 observations that were labeled as "complete" and that had complete answers for all covariates considered in our analysis.
+Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions).
+After all the preliminary analyses indicated above, the total number of observations used for analysis was 6,236 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and January 17, 2022.
 The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.
@@
 #| echo: false
-#| fig-height: 3
-#| figh-width: 2
+#| fig-height: 12
+#| figh-width: 10
 

### 2023-05-18 manuscript_v5/03-results.qmd
discussion, appendix updated

@@
 @tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis.
-The sample contained **6,236** observations, from which 24.8\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
-Vaccination rates ranged between 71-79\% across household income brackets, age groups, Health Regions, and the months considered in the survey.
+The sample contained 6,236 observations, from which 24.8\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
+Vaccination rates ranged between 71-79\% across all household income brackets, age groups, Health Regions, and the months considered in the survey.
 However, the highest vaccination rates in each category were reported by individuals in the highest income bracket (79\%), those between 16 and 34 years of age (77\%), individuals that lived in the East Health Region (77\%), and  during January of 2022 (78\%).
-Differences were higher between racial/ethnic groups, where the higher vaccination rate was reported by White/Caucasian individuals (84\%), against vaccination rates between 63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and those that reported belonging to "Other" racial groups, which included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere.
+Between racial/ethnic groups, the highest vaccination rate was reported by White/Caucasian individuals (84\%), against vaccination rates between 63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and those that reported belonging to "Other" racial groups, which included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere.
 
@@
 Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of equity-deserving groups.
-However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, certain estimates were not deemed to be statistically-significant after the correction, in contrast to the estimates from the uncorrected model.
-Specifically, the corrected model showed no significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-models,B).
+However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, there were differences in the statistical significance of certain estimates before and after the correction.
+Specifically, the uncorrected model showed significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-models,B) but these were deemed non statistically-significant after the correction.
+
 <!-- \newpage -->
@@
 Only statistically significant interaction terms are shown.
-Full interaction terms can be found in Supplementary Figure A-3."
+Full interaction terms can be found in Supplementary Figures A-3 and A-4."
 
@@
 Interestingly, individuals in equity-deserving groups with a household income below CAD 25,000 had higher odds of vaccination (when compared to those with a household income above CAD 60,000).
-This held true in the case of Arab/Middle Eastern (OR=34, CI=[1.70,6.79]), Black individuals (OR=3.81, CI=[2.05, 7.09]), or those in other racial or ethnic groups (OR=3.19, CI=[2.00,5.09]).
-Additionally, individuals with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern and other racial ethnic groups had higher odds of vaccination (ORs and CIs=6.96 [2.67,18.16], and 3.5 [1.85,6.62]).
+This held true in the case of Arab/Middle Eastern individuals (OR=34, CI=[1.70,6.79]), Black individuals (OR=3.81, CI=[2.05, 7.09]), and those in other racial or ethnic groups (OR=3.19, CI=[2.00,5.09]).
+Additionally, individuals with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern and other racial ethnic groups also had higher odds of vaccination than their high-income peers (ORs and CIs=6.96 [2.67,18.16], and 3.5 [1.85,6.62]).
 

### 2023-05-18 manuscript_v5/04-discussion.qmd
discussion, appendix updated

@@
 
-In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022, aiming at determining which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its mission of improving health access for the inhabitants of Ontario.
+In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022.
+Our goal was to determine which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its mission of improving health access for all Ontarians.
 
 Our results show that indeed, there were differences in vaccination odds across Ontario in certain socio-demographic groups.
-Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White/Caucasian (@fig-models).
+Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White or Caucasian (@fig-models).
 These results are consistent with previous studies that have shown lower vaccination rates in individuals with the same socio-demographic characteristics [@guay2022;@muhajarine2021;@carter2022;@hussain2022].
@@
 This suggests disparities in vaccine accessibility that affected in particular equity-deserving individuals in Ontario at the time of the survey.
-However, because to the best of our knowledge there seems to be a very limited amount of literature on this topic in the context of Ontario, there is an ongoing need of future studies that examine the longitudinal impact of vaccine accessibility and structural barriers that affect equity-deserving groups within the province.
+However, because to the best of our knowledge there seems to be a very limited amount of literature on vaccine accesibility for marginalized groups in the context of the province, there is an ongoing need of future studies this topic in the different Health Regions of the province.
 
@@
 
-However, the results also indicate that the place of habitation affected the odds of vaccination for certain equity-deserving groups (interaction term of Health Region and Race, @fig-models).
+However, the results also indicate that the place of habitation affected the odds of vaccination for certain equity-deserving groups (interaction term of Health Region and Race, @fig-models,B).
 Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models].
 For these individuals, vaccination odds were lower when compared to the Toronto Health Region (Supplementary Figure A-6).
-We indicate next some contributing factors that might help provide context in each case.
+We indicate next some contributing factors that might help provide context to these results.
 
-First, it is useful to analyze the data using the LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses.
+First, it is useful to dissagregate the data by LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses.
 Interestingly, for Indigenous and mixed individuals most of the observations in the survey for West Health Region came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs.

### 2023-05-18 manuscript_v5/appendix.qmd
discussion, appendix updated

@@
 
-  clean_data %>%
-  select(Race,Health_Region,first_dose) %>%
-  tbl_strata(
-    strata = Health_Region,
-    .tbl_fun =
-      ~ .x %>%
-        tbl_summary(by = first_dose, missing = "no",percent="row") %>%
-        add_n(),
-    .header = "**{strata}**, N = {n}"
-  )
+  # clean_data %>%
+  # select(Race,Health_Region,first_dose) %>%
+  # tbl_strata(
+  #   strata = Health_Region,
+  #   .tbl_fun =
+  #     ~ .x %>%
+  #       tbl_summary(by = first_dose, missing = "no",percent="row") %>%
+  #       add_n(),
+  #   .header = "**{strata}**, N = {n}"
+  # )%>%
+  # bold_labels()%>%
+  # as_kable_extra(booktabs = TRUE,linesep="")
 ```
@@
 
-```{r,Central-HR-LHIN}
-#| label: tbl-central-hr-lhin
-#| tbl-cap: "Proportion of answers by LHIN (Central Health Region)"
+```{r,West-HR-Indigenous}
+#| label: tbl-west-region-indigenous
+#| tbl-cap: "Proportion of answers by LHIN in the West Health Region for Indigenous Individuals"
 #| message: false
@@
 
-clean_data %>% subset (Health_Region=="Central")%>%
-tbl_summary(include=c(LHIN),
-            sort = all_categorical() ~ "frequency") %>% 
+clean_data %>%
+  subset(Health_Region=="West" & Race=="Indigenous")%>%
+  tbl_summary(by= first_dose,
+              #missing="no",
+              sort = list(everything() ~ "frequency"),
+              include=c(LHIN))%>% 
   bold_labels()%>%
@@
 ```
+
+```{r,West-HR-Mixed}
+#| label: tbl-west-region-mixed
+#| tbl-cap: "Proportion of answers by LHIN in the West Health Region for Mixed Individuals"
+#| message: false
+#| echo: false
+
+clean_data %>%
+  subset(Health_Region=="West" & Race=="Mixed")%>%
+  tbl_summary(by= first_dose,
+              #missing="no",
+              sort = list(everything() ~ "frequency"),
+              include=c(LHIN)) %>% 
+  bold_labels()%>%
+  as_kable_extra(booktabs = TRUE,linesep="")
+
+```
+
+```{r,East-HR-Arab}
+#| label: tbl-East-region-Arab
+#| tbl-cap: "Proportion of answers by LHIN in the East Health Region for Arab/Middle Eastern Individuals"
+#| message: false
+#| echo: false
+
+a1_rake %>%
+  subset(Health_Region=="East" & Race=="Arab/Middle Eastern")%>%
+  tbl_svysummary(by= first_dose,
+              #missing="no",
+              sort = list(everything() ~ "frequency"),
+              include=c(LHIN))%>% 
+  bold_labels()%>%
+  as_kable_extra(booktabs = TRUE,linesep="")
+
+```

### 2023-05-18 manuscript_v5/appendix.tex
discussion, appendix updated

@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, breakable, enhanced, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{tbl-central-hr-lhin}{}
+\hypertarget{tbl-west-region-indigenous}{}
 \begin{table}
-\caption{\label{tbl-central-hr-lhin}Proportion of answers by LHIN (Central Health Region) }\tabularnewline
+\caption{\label{tbl-west-region-indigenous}Proportion of answers by LHIN in the West Health Region for Indigenous
+Individuals }\tabularnewline
 
 \centering
-\begin{tabular}{lc}
+\begin{tabular}{lcc}
 \toprule
-\textbf{Characteristic} & \textbf{N = 1,455}\\
+\textbf{Characteristic} & \textbf{no}, N = 32 & \textbf{yes}, N = 42\\
 \midrule
-\textbf{LHIN} & \\
-\hspace{1em}Mississauga Halton & 495 (34\%)\\
-\hspace{1em}Central & 428 (29\%)\\
-\hspace{1em}Central West & 362 (25\%)\\
-\hspace{1em}North Simcoe Muskoka & 170 (12\%)\\
+\textbf{LHIN} &  & \\
+\hspace{1em}Hamilton Niagara Haldimand Brant (Hnhb) & 19 (59\%) & 9 (21\%)\\
+\hspace{1em}South West & 4 (13\%) & 18 (43\%)\\
+\hspace{1em}Waterloo Wellington & 7 (22\%) & 11 (26\%)\\
+\hspace{1em}Erie St. Clair & 2 (6.3\%) & 4 (9.5\%)\\
 \bottomrule
-\multicolumn{2}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\end{tabular}
+\end{table}
+
+\hypertarget{tbl-west-region-mixed}{}
+\begin{table}
+\caption{\label{tbl-west-region-mixed}Proportion of answers by LHIN in the West Health Region for Mixed
+Individuals }\tabularnewline
+
+\centering
+\begin{tabular}{lcc}
+\toprule
+\textbf{Characteristic} & \textbf{no}, N = 41 & \textbf{yes}, N = 76\\
+\midrule
+\textbf{LHIN} &  & \\
+\hspace{1em}Waterloo Wellington & 14 (34\%) & 20 (26\%)\\
+\hspace{1em}Hamilton Niagara Haldimand Brant (Hnhb) & 10 (24\%) & 22 (29\%)\\
+\hspace{1em}South West & 11 (27\%) & 20 (26\%)\\
+\hspace{1em}Erie St. Clair & 6 (15\%) & 14 (18\%)\\
+\bottomrule
+\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
+\end{tabular}
+\end{table}
+
+\hypertarget{tbl-East-region-Arab}{}
+\begin{table}
+\caption{\label{tbl-East-region-Arab}Proportion of answers by LHIN in the East Health Region for Arab/Middle
+Eastern Individuals }\tabularnewline
+
+\centering
+\begin{tabular}{lcc}
+\toprule
+\textbf{Characteristic} & \textbf{no}, N = 19,716 & \textbf{yes}, N = 23,499\\
+\midrule
+\textbf{LHIN} &  & \\
+\hspace{1em}Toronto \hspace{1em}Central & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}Champlain & 13,173 (67\%) & 19,825 (84\%)\\
+\hspace{1em}Mississauga Halton & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}Hamilton Niagara Haldimand Brant (Hnhb) & 0 (0\%) & 0 (0\%)\\
+Central & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}Waterloo Wellington & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}Central West & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}South West & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}Central East & 6,543 (33\%) & 3,674 (16\%)\\
+\hspace{1em}Erie St. Clair & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}North Simcoe Muskoka & 0 (0\%) & 0 (0\%)\\
+\hspace{1em}South East & 0 (0\%) & 0 (0\%)\\
+\bottomrule
+\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
 \end{tabular}

### 2023-05-18 manuscript_v5/main.tex
discussion, appendix updated

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, sharp corners, interior hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, boxrule=0pt, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi
 
@@
 vaccine?'', with possible answers ``yes'' and ``no''. The original
-dataset contained 39,029 observations (where each observation
-corresponded to a unique respondent).
-
-Preliminary analyses of the data included the removal of outliers
-(\textbf{should we still do this? it's only 19 observations with income
-\textgreater110k and household of 1, but we are not even using such
-income bracket in the analysis because we re-grouped the data, and the
-household size variable has 90\% missing rate}), of observations where
-respondents did not provide answers in all the covariates of interest,
-matching the city of each observations with its corresponding LHIN and
-Health Region, and removing observations from areas with low
-representation (107 observations corresponding to the North West and
-North East Health Regions). After all the preliminary analyses indicated
-above, the total number of observations used for analysis was 3,549
-which included the East, Central, Toronto, and West Health Regions
-covering the period between October 1st,2021 and December 12th, 2021.
-The original dataset, clean dataset, and details on the data cleaning
-process are described in detail in the
+dataset contained 39,029 observations.
+
+By design, the survey allowed respondents to exit at any time and
+deployed the questions randomly, which resulted in \(\approx\) 84\% of
+the observations having multiple missing answers or being incomplete.
+Therefore, we selected 6,343 observations that were labeled as
+``complete'' and that had complete answers for all covariates considered
+in our analysis. Later, we matched the city of each observation with its
+corresponding LHIN and Health Region, and removed observations from
+areas with low representation (107 observations corresponding to the
+North West and North East Health Regions). After all the preliminary
+analyses indicated above, the total number of observations used for
+analysis was 6,236 which included the East, Central, Toronto, and West
+Health Regions covering the period between October 1st,2021 and January
+17, 2022. The original dataset, clean dataset, and details on the data
+cleaning process are described in detail in the
 \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
@@
 from the Fields COVID-19 survey used for analysis. The sample contained
-\textbf{6,236} observations, from which 24.8\% (1,547) corresponded to
+6,236 observations, from which 24.8\% (1,547) corresponded to
 individuals that reported not having received the first dose of the

### 2023-05-18 code/corr_model.R

@@ -16,7 +16,13 @@ mplot2$factor<-c(rep("Age",2),
 #order factors for plot
 #mplot2$term_ord <- factor(mplot2$term,ordered=TRUE,levels=mplot2$term)
 
-mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age","Month","Income","Race/Ethnicity","Health Region","Income and\nRace/Ethnicity","Race/Ethnicity\nand\nHealth Region"))
+mplot2$factor <- factor(mplot2$factor,ordered=TRUE,levels=c("Age",
+                                                            "Month",
+                                                            "Income",
+                                                            "Race/Ethnicity",
+                                                            "Health Region",
+                                                            "Income and\nRace/Ethnicity",
+                                                            "Race/Ethnicity\nand\nHealth Region"))
 
 # remove non-significant interaction factors
 

### 2023-05-18 manuscript_v5/02-methods.qmd
discussion updated

@@
 By design, the survey allowed respondents to exit at any time and deployed the questions randomly, which resulted in $\approx$ 84\% of the observations having multiple missing answers or being incomplete.
-Therefore, we selected 6,343 observations that were labeled as "complete" and that had complete answers for all covariates considered in our analysis.
+Therefore, we selected 6,343 observations that were labeled as "complete" in the dataset and that had answers for all covariates considered in our analysis.
 Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions).
-After all the preliminary analyses indicated above, the total number of observations used for analysis was 6,236 which included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and January 17, 2022.
+After all the preliminary analyses indicated above, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and January 17, 2022.
 The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.

### 2023-05-18 manuscript_v5/04-discussion.qmd
discussion updated

@@
 In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with demand [@gill2022].
-This suggests disparities in vaccine accessibility that affected in particular equity-deserving individuals in Ontario at the time of the survey.
-However, because to the best of our knowledge there seems to be a very limited amount of literature on vaccine accesibility for marginalized groups in the context of the province, there is an ongoing need of future studies this topic in the different Health Regions of the province.
+This suggests that the observed differences are associated to disparities in vaccine access that were present during the period covered by the survey.
 
-Interestingly, whereas overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in lower income brackets when compared to the CAD 60,000 and over bracket (Supplementary Figure A-5).
+Interestingly, whereas overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in income brackets below CAD 60,000 (Supplementary Figure A-5).
 Specifically, individuals in low income brackets that belonged to Arab/Middle Eastern, Black, or other minority groups had higher odds of vaccination that their peers with an income above 60,000 CAD.
 
-This result is likely reflects in part the fact that individuals in racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include grocery store, gas station, warehouse, distribution, and manufacturing workers, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
+This result is likely reflects in part the fact that individuals in racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include workers in the areas of grocery stores, gas stations, warehouses, distribution, and manufacturing, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
 In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022].
-These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation from individuals in equity-deserving groups played an important role in increasing the odds of vaccination in the province.
+These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation in individuals of equity-deserving groups played an important role in increasing the odds of vaccination.
 
@@
 
-First, it is useful to dissagregate the data by LHINs, because most of the studies in the literature that have analyzed health in Ontario use the LHINs as the base of their analyses.
-Interestingly, for Indigenous and mixed individuals most of the observations in the survey for West Health Region came from the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs (**add table in Appendix**), whereas for Arab/Middle Eastern Individuals in the East Health Region the highest number of observations corresponded to the Champlain and Central East LHINs.
+First, in this case it is useful to analyze the data considering the LHINs in each Health Region, because most studies in the literature focused on Ontario use the LHINs as the base of their analyses.
+The West Health Region covers the area previously occupied by the Hamilton Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs, whereas the East Health Region covers the area of the former Champlain and Central East LHINs.
 Previous research has identified health disparities in these (mostly rural) regions, such as unequal distribution of primary care providers, increased mortality, and low pharmacist availability [@shah2019;@crighton2015;@timony2022].
@@
 Indeed, these rationales have been associated to observed lower vaccination rates among these groups [@smylie2022;@eissa2021].
-Taken together, this suggests that healthcare disparities that are specific to certain socio-demographic groups are associated with lower vaccination uptake.
+Taken together, this suggests that healthcare disparities specific to these equity-deserving groups in certain parts of the province impacted vaccination uptake, and highlights the need of policy-makers to invest in resources and infrastructure focused on personalized care, such as community-led advisors or race-concordant health professionals that have been shown to improve trust in vaccination in marginalized individuals [@schafferderoo2020].
 
@@
 Therefore, we focused on entries that had complete answers, and corrected the data using population-wide information from the Census.
-However, more granular corrections would be needed to obtain more accurate estimates.
+More granular corrections would be needed to obtain more accurate estimates.
 For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates that other racial groups.
-Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates.
-Unfortunately, at the moment this correction cannot be implemented as such stratification is not currently available in the Census data.
+Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates but unfortunately at the moment this correction cannot be implemented as such stratification is has not been implemented in the Census.
 
 Additionally, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map).
-Low representation is expected as these regions only account for 5\% of the total population of Ontario, but in contrast, they have the highest proportion of Indigenous inhabitants [@ontariohealth].
+Low representation is expected as these regions as they only account for 5\% of the total population of Ontario, but in contrast, they have the highest proportion of Indigenous inhabitants [@ontariohealth].
 In the context of personalized care, there is a need for collecting data that focuses on these Health Regions where additional health disparities might be present and possibly understudied.
@@
 The results in this study are based on self-reported data, where bias might be present.
-However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid assessment of vaccination in the province.
-Finally, although higher vaccination odds identified for January of 2022 in the model are in accordance with province-wide trends reported by Public Health Ontario (which show a 4\% increase between early December and January, in contrast to a 2.5\% increase between October and November [@ontario-covid]), these results are only a snapshot when compared to the overall duration of the pandemic.
+However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid sample of vaccination uptake in the province.
+This is supported by the statistically-significant higher vaccination odds identified for January of 2022 in the model, which are consistent with province-wide trends reported by Public Health Ontario (which show a 4\% increase between early December and January, in contrast to a 2.5\% increase between October and November [@ontario-covid]); however, the short time window constitutes essentially a "snapshot" view of the evolution of the disease, and additional data would be needed in order to obtain estimates per racial/ethnic group over time across all Health Regions that can help inform the existence of other health disparities.
 
-Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.
+Nonetheless, the results presented here can serve as a starting point to motivate the collection of robust longitudinal data that can be used to quantify geographical and temporal differences within vulnerable segments of the population, and that can be used to inform the development of adequate public health policies within the province of Ontario or across other provinces in Canada that aim to minimize disparities in health access.

### 2023-05-18 manuscript_v5/05-conclusion.qmd
discussion updated

--- 2023-04-28 manuscript_v4/05-conclusion.qmd
@@
 # Conclusion
+
+The implementation of the Health Region model in Ontario aimed at reducing the bureaucratic complexity and disparities in health identified under the LHINs.
+However, the complex socio-demographic landscape of the province poses multiple challenges to the aim of the Health Region model of improving healthcare for all Ontarians.
+In particular, because the Health Regions cover larger geographical areas with various degrees of rurality and representation of equity-deserving groups, there is a need for decision-makers to consider that the evidence collected during the COVID-19 indicates a lack of infrastructure that can support personalized care for marginalized individuals
 

### 2023-05-18 manuscript_v5/appendix.qmd
discussion updated

--- 2023-05-11 manuscript_v5/appendix.qmd
@@
 ```
-
-# Socio-economic Information from the Survey Data
-
-Tables A-7 and A-8 present summaries of information from the Fields survey (from the clean dataset analyzed in the main paper), showing responses by White/Caucasian individuals by Health Region and proportion of answers by LHIN in the West Health Region.
-
-```{r,health-region}
-#| label: tbl-hr-percentages
-#| tbl-cap: "Reported Vaccination Status from the Survey for White/Caucasian Individuals by Health Region"
-#| message: false
-#| echo: false
-
-clean_data %>% subset (Race=="White/Caucasian")%>%
-  tbl_summary(
-    percent="row",
-    by= first_dose,
-    include=c(Health_Region),
-    sort = all_categorical() ~ "frequency"
-  ) %>% bold_labels()%>%
-  as_kable_extra(booktabs = TRUE,linesep="")
-```
-
-```{r,West-HR-LHIN}
-#| label: tbl-west-hr-lhin
-#| tbl-cap: "Proportion of answers by LHIN (West Health Region)"
-#| message: false
-#| echo: false
-
-clean_data %>% subset (Health_Region=="West")%>%
-tbl_summary(include=c(LHIN),
-            sort = all_categorical() ~ "frequency") %>% 
-  bold_labels()%>%
-  as_kable_extra(booktabs = TRUE,linesep="")
-
-```

### 2023-05-18 manuscript_v5/appendix.tex
discussion updated

--- 2023-05-11 manuscript_v5/appendix.tex
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\hypertarget{socio-economic-information-from-the-survey-data}{%
-\section{Socio-economic Information from the Survey
-Data}\label{socio-economic-information-from-the-survey-data}}
-
-Tables A-7 and A-8 present summaries of information from the Fields
-survey (from the clean dataset analyzed in the main paper), showing
-responses by White/Caucasian individuals by Health Region and proportion
-of answers by LHIN in the West Health Region.
-
-\hypertarget{tbl-hr-percentages}{}
-\begin{table}
-\caption{\label{tbl-hr-percentages}Reported Vaccination Status from the Survey for White/Caucasian
-Individuals by Health Region }\tabularnewline
-
-\centering
-\begin{tabular}{lcc}
-\toprule
-\textbf{Characteristic} & \textbf{no}, N = 354 & \textbf{yes}, N = 1,871\\
-\midrule
-\textbf{Health\_Region} &  & \\
-\hspace{1em}West & 105 (15\%) & 582 (85\%)\\
-\hspace{1em}Toronto & 127 (19\%) & 533 (81\%)\\
-\hspace{1em}East & 66 (14\%) & 396 (86\%)\\
-\hspace{1em}Central & 56 (13\%) & 360 (87\%)\\
-\bottomrule
-\multicolumn{3}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
-\end{tabular}
-\end{table}
-
-\hypertarget{tbl-west-hr-lhin}{}
-\begin{table}
-\caption{\label{tbl-west-hr-lhin}Proportion of answers by LHIN (West Health Region) }\tabularnewline
-
-\centering
-\begin{tabular}{lc}
-\toprule
-\textbf{Characteristic} & \textbf{N = 1,460}\\
-\midrule
-\textbf{LHIN} & \\
-\hspace{1em}Hamilton Niagara Haldimand Brant (Hnhb) & 470 (32\%)\\
-\hspace{1em}Waterloo Wellington & 400 (27\%)\\
-\hspace{1em}South West & 345 (24\%)\\
-\hspace{1em}Erie St. Clair & 245 (17\%)\\
-\bottomrule
-\multicolumn{2}{l}{\rule{0pt}{1em}\textsuperscript{1} n (\%)}\\
-\end{tabular}
-\end{table}
-
 \end{document}

### 2023-05-18 manuscript_v5/main.tex
discussion updated

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, boxrule=0pt, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 Therefore, we selected 6,343 observations that were labeled as
-``complete'' and that had complete answers for all covariates considered
-in our analysis. Later, we matched the city of each observation with its
-corresponding LHIN and Health Region, and removed observations from
-areas with low representation (107 observations corresponding to the
-North West and North East Health Regions). After all the preliminary
-analyses indicated above, the total number of observations used for
-analysis was 6,236 which included the East, Central, Toronto, and West
-Health Regions covering the period between October 1st,2021 and January
-17, 2022. The original dataset, clean dataset, and details on the data
-cleaning process are described in detail in the
+``complete'' in the dataset and that had answers for all covariates
+considered in our analysis. Later, we matched the city of each
+observation with its corresponding LHIN and Health Region, and removed
+observations from areas with low representation (107 observations
+corresponding to the North West and North East Health Regions). After
+all the preliminary analyses indicated above, the total number of
+observations used for analysis was 6,236 and included the East, Central,
+Toronto, and West Health Regions covering the period between October
+1st,2021 and January 17, 2022. The original dataset, clean dataset, and
+details on the data cleaning process are described in detail in the
 \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
@@
 individuals that reported not having received the first dose of the
-vaccine. Vaccination rates ranged between 71-79\% across household
+vaccine. Vaccination rates ranged between 71-79\% across all household
 income brackets, age groups, Health Regions, and the months considered
@@
 between 16 and 34 years of age (77\%), individuals that lived in the
-East Health Region (77\%), and during January of 2022 (78\%).
-Differences were higher between racial/ethnic groups, where the higher
-vaccination rate was reported by White/Caucasian individuals (84\%),
-against vaccination rates between 63-66\% reported in the case of
-Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and
-those that reported belonging to ``Other'' racial groups, which included
-Southeast Asian, Filipino, West Asian, and minorities not identified
-elsewhere.
+East Health Region (77\%), and during January of 2022 (78\%). Between
+racial/ethnic groups, the highest vaccination rate was reported by
+White/Caucasian individuals (84\%), against vaccination rates between
+63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous,
+Latin American individuals, and those that reported belonging to
+``Other'' racial groups, which included Southeast Asian, Filipino, West
+Asian, and minorities not identified elsewhere.
 
@@
 the magnitude of the estimates differed between the uncorrected and
-corrected models and more importantly, certain estimates were not deemed
-to be statistically-significant after the correction, in contrast to the
-estimates from the uncorrected model. Specifically, the corrected model
-showed no significant differences in vaccination odds between the age
-groups considered, the East Health Region, Latin American individuals
-with a household income under CAD 25,000, and Indigenous individuals
-living in the Central Health Region (Figure~\ref{fig-models},B).
+corrected models and more importantly, there were differences in the
+statistical significance of certain estimates before and after the
+correction. Specifically, the uncorrected model showed significant
+differences in vaccination odds between the age groups considered, the
+East Health Region, Latin American individuals with a household income
+under CAD 25,000, and Indigenous individuals living in the Central
+Health Region (Figure~\ref{fig-models},B) but these were deemed non
+statistically-significant after the correction.
 
@@
 interaction terms are shown. Full interaction terms can be found in
-Supplementary Figure A-3.}
+Supplementary Figures A-3 and A-4.}
 
@@
 those with a household income above CAD 60,000). This held true in the
-case of Arab/Middle Eastern (OR=34, CI={[}1.70,6.79{]}), Black
-individuals (OR=3.81, CI={[}2.05, 7.09{]}), or those in other racial or
-ethnic groups (OR=3.19, CI={[}2.00,5.09{]}). Additionally, individuals
-with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern
-and other racial ethnic groups had higher odds of vaccination (ORs and
-CIs=6.96 {[}2.67,18.16{]}, and 3.5 {[}1.85,6.62{]}).
+case of Arab/Middle Eastern individuals (OR=34, CI={[}1.70,6.79{]}),
+Black individuals (OR=3.81, CI={[}2.05, 7.09{]}), and those in other
+racial or ethnic groups (OR=3.19, CI={[}2.00,5.09{]}). Additionally,
+individuals with an income between CAD 25,000 and 59,999 in the
+Arab/Middle Eastern and other racial ethnic groups also had higher odds
+of vaccination than their high-income peers (ORs and CIs=6.96
+{[}2.67,18.16{]}, and 3.5 {[}1.85,6.62{]}).
 
@@
 uptake were present between the Health Regions during between late 2021
-and early 2022, aiming at determining which socio-demographic groups
+and early 2022. Our goal was to determine which socio-demographic groups
 could be impacted by these disparities in order to provide
@@
 Health Region model is able to fulfill its mission of improving health
-access for the inhabitants of Ontario.
+access for all Ontarians.
 
@@
 odds that were between a third and a half of that of individuals that
-identified as White/Caucasian (Figure~\ref{fig-models}). These results
-are consistent with previous studies that have shown lower vaccination
-rates in individuals with the same socio-demographic
+identified as White or Caucasian (Figure~\ref{fig-models}). These
+results are consistent with previous studies that have shown lower
+vaccination rates in individuals with the same socio-demographic
 characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
@@
 demand\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. This
-suggests disparities in vaccine accessibility that affected in
-particular equity-deserving individuals in Ontario at the time of the
-survey. However, because to the best of our knowledge there seems to be
-a very limited amount of literature on this topic in the context of
-Ontario, there is an ongoing need of future studies that examine the
-longitudinal impact of vaccine accessibility and structural barriers
-that affect equity-deserving groups within the province.
+suggests that the observed differences are associated to disparities in
+vaccine access that were present during the period covered by the
+survey.
 
@@
 actually positive, in contrast to the White/Caucasian group, where
-vaccination odds decreased in lower income brackets when compared to the
-CAD 60,000 and over bracket (Supplementary Figure A-5). Specifically,
-individuals in low income brackets that belonged to Arab/Middle Eastern,
-Black, or other minority groups had higher odds of vaccination that
-their peers with an income above 60,000 CAD.
+vaccination odds decreased in income brackets below CAD 60,000
+(Supplementary Figure A-5). Specifically, individuals in low income
+brackets that belonged to Arab/Middle Eastern, Black, or other minority
+groups had higher odds of vaccination that their peers with an income
+above 60,000 CAD.
 
@@
 pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{55},\protect\hyperlink{ref-ct2021}{56}},
-which include grocery store, gas station, warehouse, distribution, and
-manufacturing workers, all being occupations for which an income within
-the significant brackets identified in the analysis is to be expected.
-In Ontario, these workers had priority for COVID-19
+which include workers in the areas of grocery stores, gas stations,
+warehouses, distribution, and manufacturing, all being occupations for
+which an income within the significant brackets identified in the
+analysis is to be expected. In Ontario, these workers had priority for
+COVID-19
 vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{57}}; and
@@
 elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{58}},
-suggest that the type of occupation from individuals in equity-deserving
-groups played an important role in increasing the odds of vaccination in
-the province.
+suggest that the type of occupation in individuals of equity-deserving
+groups played an important role in increasing the odds of vaccination.
 
@@
 the odds of vaccination for certain equity-deserving groups (interaction
-term of Health Region and Race, Figure~\ref{fig-models}). Specifically,
-this held true in the case of individuals identifying as Indigenous or
-with mixed racial background in the West Health Region, Black
-individuals in the Central Health Region, and Arab/Middle Eastern
+term of Health Region and Race, Figure~\ref{fig-models},B).
+Specifically, this held true in the case of individuals identifying as
+Indigenous or with mixed racial background in the West Health Region,
+Black individuals in the Central Health Region, and Arab/Middle Eastern
 individuals in the East Health Region Figure~\ref{fig-models}. For these
@@
 Health Region (Supplementary Figure A-6). We indicate next some
-contributing factors that might help provide context in each case.
-
-First, it is useful to analyze the data using the LHINs, because most of
-the studies in the literature that have analyzed health in Ontario use
-the LHINs as the base of their analyses. Interestingly, for Indigenous
-and mixed individuals most of the observations in the survey for West
-Health Region came from the Hamilton Niagara Haldimand Brant, South
-West, and Waterloo Wellington LHINs (\textbf{add table in Appendix}),
-whereas for Arab/Middle Eastern Individuals in the East Health Region
-the highest number of observations corresponded to the Champlain and
-Central East LHINs. Previous research has identified health disparities
-in these (mostly rural) regions, such as unequal distribution of primary
-care providers, increased mortality, and low pharmacist
+contributing factors that might help provide context to these results.
+
+First, in this case it is useful to analyze the data considering the
+LHINs in each Health Region, because most studies in the literature
+focused on Ontario use the LHINs as the base of their analyses. The West
+Health Region covers the area previously occupied by the Hamilton
+Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs,
+whereas the East Health Region covers the area of the former Champlain
+and Central East LHINs. Previous research has identified health
+disparities in these (mostly rural) regions, such as unequal
+distribution of primary care providers, increased mortality, and low
+pharmacist
 availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.
@@
 groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{63},\protect\hyperlink{ref-eissa2021}{64}}.
-Taken together, this suggests that healthcare disparities that are
-specific to certain socio-demographic groups are associated with lower
-vaccination uptake.
+Taken together, this suggests that healthcare disparities specific to
+these equity-deserving groups in certain parts of the province impacted
+vaccination uptake, and highlights the need of policy-makers to invest
+in resources and infrastructure focused on personalized care, such as
+community-led advisors or race-concordant health professionals that have
+been shown to improve trust in vaccination in marginalized
+individuals\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{65}}.
 
@@
 and corrected the data using population-wide information from the
-Census. However, more granular corrections would be needed to obtain
-more accurate estimates. For example, our analysis identified higher
-odds of vaccination in the Central and West Health Regions, but in this
-case these differences are likely to be driven by the proportion of
+Census. More granular corrections would be needed to obtain more
+accurate estimates. For example, our analysis identified higher odds of
+vaccination in the Central and West Health Regions, but in this case
+these differences are likely to be driven by the proportion of
 White/Caucasian individuals, who had higher vaccination rates that other
@@
 Region can provide a more accurate estimation of region-wide vaccination
-rates. Unfortunately, at the moment this correction cannot be
-implemented as such stratification is not currently available in the
-Census data.
+rates but unfortunately at the moment this correction cannot be
+implemented as such stratification is has not been implemented in the
+Census.
 
@@
 in the survey (Figure~\ref{fig-map}). Low representation is expected as
-these regions only account for 5\% of the total population of Ontario,
-but in contrast, they have the highest proportion of Indigenous
+these regions as they only account for 5\% of the total population of
+Ontario, but in contrast, they have the highest proportion of Indigenous
 inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
@@
 documented vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{65}}, we
-believe that our data was able to provide a valid assessment of
-vaccination in the province. Finally, although higher vaccination odds
-identified for January of 2022 in the model are in accordance with
-province-wide trends reported by Public Health Ontario (which show a 4\%
-increase between early December and January, in contrast to a 2.5\%
-increase between October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{66}}),
-these results are only a snapshot when compared to the overall duration
-of the pandemic.
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{66}}, we
+believe that our data was able to provide a valid sample of vaccination
+uptake in the province. This is supported by the
+statistically-significant higher vaccination odds identified for January
+of 2022 in the model, which are consistent with province-wide trends
+reported by Public Health Ontario (which show a 4\% increase between
+early December and January, in contrast to a 2.5\% increase between
+October and
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{67}});
+however, the short time window constitutes essentially a ``snapshot''
+view of the evolution of the disease, and additional data would be
+needed in order to obtain estimates per racial/ethnic group over time
+across all Health Regions that can help inform the existence of other
+health disparities.
 
@@
 development of adequate public health policies within the province of
-Ontario or across other provinces that aim to minimize disparities in
-health access.
+Ontario or across other provinces in Canada that aim to minimize
+disparities in health access.
 
@@
 \section{Conclusion}\label{conclusion}}
+
+The implementation of the Health Region model in Ontario aimed at
+reducing the bureaucratic complexity and disparities in health
+identified under the LHINs. However, the complex socio-demographic
+landscape of the province poses multiple challenges to the aim of the
+Health Region model of improving healthcare for all Ontarians. In
+particular, because the Health Regions cover larger geographical areas
+with various degrees of rurality and representation of equity-deserving
+groups, there is a need for decision-makers to consider that the
+evidence collected during the COVID-19 indicates a lack of
+infrastructure that can support personalized care for marginalized
+individuals
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
+\CSLLeftMargin{65. }%
+\CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
+vaccination program. \emph{{JAMA}}. 2020;323(24):2458.
+doi:\href{https://doi.org/10.1001/jama.2020.8711}{10.1001/jama.2020.8711}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,

### 2023-05-19 manuscript_v5/02-methods.qmd
updated map, need to fix author section and figure (sideways)

@@
 Therefore, we selected 6,343 observations that were labeled as "complete" in the dataset and that had answers for all covariates considered in our analysis.
-Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (107 observations corresponding to the North West and North East Health Regions).
+Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions).
+Finally, we removed outliers from the data (19 observations of individuals with household size of 1 and income above 110,000 CAD in the original dataset).
 After all the preliminary analyses indicated above, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and January 17, 2022.
@@
 #| fig-cap: "Geographic representation of the data collected by the _Survey of COVID-19 related Behaviours and Attitudes_, collected by the Fields Institute in Ontario.
-The municipalities (cities) from where survey participants provided answers (in the clean dataset) appear as points.
+The municipalities from where survey participants provided answers appear as points.
 The Health six Regions are color-coded.
@@
 #| echo: false
-#| fig-height: 12
-#| figh-width: 10
+#| out-height: 100%
 
-knitr::include_graphics(here("data","map_data","map.pdf"))
+knitr::include_graphics(here("data","map_data","map_cairo.pdf"))
 

### 2023-05-19 manuscript_v5/04-discussion.qmd
updated map, need to fix author section and figure (sideways)

@@
 For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated 100,000 Indigenous individuals living in the area [@ontariohealth].
-Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada.
-Indeed, these rationales have been associated to observed lower vaccination rates among these groups [@smylie2022;@eissa2021].
-Taken together, this suggests that healthcare disparities specific to these equity-deserving groups in certain parts of the province impacted vaccination uptake, and highlights the need of policy-makers to invest in resources and infrastructure focused on personalized care, such as community-led advisors or race-concordant health professionals that have been shown to improve trust in vaccination in marginalized individuals [@schafferderoo2020].
+Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada, as these rationales have been associated to observed lower vaccination rates among these groups [@smylie2022;@eissa2021].
+Taken together, this suggests that healthcare disparities specific to these equity-deserving groups in certain parts of the province impacted vaccination uptake, and highlights the need of investments in the Health Regions focused on resources, infrastructure, and specially personnel that can deliver personalized care to marginalized communities, as it has been shown that such efforts have improved trust in vaccination in visible minorities [@schafferderoo2020].
 

### 2023-05-19 manuscript_v5/05-conclusion.qmd
updated map, need to fix author section and figure (sideways)

--- 2023-04-28 manuscript_v4/05-conclusion.qmd
@@
 
-This study explored differences in COVID-19 vaccination across the province of Ontario during the last quarter of 2021 taking into consideration socio-economic factors, such as income and race, their interactions, and the Health Regions within the province.
-Our results show that during the period analyzed, differences in vaccination uptake existed across multiple equity-deserving groups in the province, and that these differences were significant in two of the Health Regions analyzed.
-It is important that future public health policies in Ontario take into consideration how to adequately reach individuals from equity-deserving groups that might live in areas of the province where access to healthcare might be difficult.
-Only in this way the goal of the Health Region model, which aims at reducing disparities, will become successful.
+The implementation of the Health Regions in Ontario aimed at reducing the bureaucratic complexity and health disparities identified under the LHIN model.
+However, there are currently multiple challenges that need to be addressed to ensure that the new model is able to improve healthcare for the inhabitants of the province.
+First, the fact that each Health Region now covers a large geographical area that was served by multiple LHINs in the past creates a complex socio-demographic landscape that is different in each case due the different levels of rurality and representation of equity-deserving groups that are now within each Health Region.
+So far, the evidence collected during the COVID-19 pandemic indicates that differences in vaccination uptake are associated to a lack of infrastructure and resources that can adequately support personalized care to marginalized individuals.
+In the near future, health decision-makers will need to consider the implementation of policies that are focused on addressing this problematic.
+
+Moreover, the recent nature in the adaption of the Health Region poses a challenge for researchers in the acquisition of data and information that can be used to analyze the performance of the new system.
+From one side, the Health Regions have not been incorporated as part of Census data (LHINs were considered before in the Census), and this impact the amount and level of detail of available information.
+Currently, the only demographic information available for each Health Region is provided by Ontario Health (the agency that administers the Health Regions) but this information only provides general estimates that do not allow for detailed analyses on performance indicators (such as hospitalizations, readmissions, and trends in chronic disease incidence) between the regions.
+Without open information, it is impossible to assess the level of success of the Health Region model, which is critical considering that such evaluations have not been part of the Annual Reports of the Auditor General of Ontario, which in the past analyzed the performance of the LHINs and pointed to ongoing needs and failures in the system.
+
+The Health Region model will only by successful if it ensures that healthcare improves across all segments of the population of Ontario, particularly in the event of a future public health emergency or pandemic where so far, based on the experience of the COVID-19 pandemic, equity-deserving individuals have been disproportionately affected.

### 2023-05-19 manuscript_v5/appendix.tex
updated map, need to fix author section and figure (sideways)

@@
 \usepackage{booktabs}
+\usepackage{caption}
 \usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-\usepackage{caption}
 \usepackage{placeins}
@@
 
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi
 

### 2023-05-19 manuscript_v5/main.tex
updated map, need to fix author section and figure (sideways)

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, sharp corners, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 observation with its corresponding LHIN and Health Region, and removed
-observations from areas with low representation (107 observations
-corresponding to the North West and North East Health Regions). After
-all the preliminary analyses indicated above, the total number of
-observations used for analysis was 6,236 and included the East, Central,
-Toronto, and West Health Regions covering the period between October
-1st,2021 and January 17, 2022. The original dataset, clean dataset, and
-details on the data cleaning process are described in detail in the
-\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
-for this paper.
+observations from areas with low representation (254 observations
+corresponding to the North West and North East Health Regions). Finally,
+we removed outliers from the data (19 observations of individuals with
+household size of 1 and income above 110,000 CAD in the original
+dataset). After all the preliminary analyses indicated above, the total
+number of observations used for analysis was 6,236 and included the
+East, Central, Toronto, and West Health Regions covering the period
+between October 1st,2021 and January 17, 2022. The original dataset,
+clean dataset, and details on the data cleaning process are described in
+detail in the \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub
+repository} for this paper.
 
@@
 
-{\centering \includegraphics{../data/map_data/map.pdf}
+{\centering \includegraphics[width=\textwidth,height=1\textheight]{../data/map_data/map_cairo.pdf}
 
@@
 by the \emph{Survey of COVID-19 related Behaviours and Attitudes},
-collected by the Fields Institute in Ontario. The municipalities
-(cities) from where survey participants provided answers (in the clean
-dataset) appear as points. The Health six Regions are color-coded.
-Internal boundaries within certain Health Regions indicate areas that
-belonged to the Local Integrated Health Networks (LHINs), the geographic
-areas for healthcare in Ontario before the adoption of the Health
-Regions.}
+collected by the Fields Institute in Ontario. The municipalities from
+where survey participants provided answers appear as points. The Health
+six Regions are color-coded. Internal boundaries within certain Health
+Regions indicate areas that belonged to the Local Integrated Health
+Networks (LHINs), the geographic areas for healthcare in Ontario before
+the adoption of the Health Regions.}
 
@@
 which is known to be the case for Indigenous and Black individuals in
-Canada. Indeed, these rationales have been associated to observed lower
+Canada, as these rationales have been associated to observed lower
 vaccination rates among these
@@
 these equity-deserving groups in certain parts of the province impacted
-vaccination uptake, and highlights the need of policy-makers to invest
-in resources and infrastructure focused on personalized care, such as
-community-led advisors or race-concordant health professionals that have
-been shown to improve trust in vaccination in marginalized
-individuals\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{65}}.
+vaccination uptake, and highlights the need of investments in the Health
+Regions focused on resources, infrastructure, and specially personnel
+that can deliver personalized care to marginalized communities, as it
+has been shown that such efforts have improved trust in vaccination in
+visible
+minorities\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{65}}.
 
@@
 
-The implementation of the Health Region model in Ontario aimed at
-reducing the bureaucratic complexity and disparities in health
-identified under the LHINs. However, the complex socio-demographic
-landscape of the province poses multiple challenges to the aim of the
-Health Region model of improving healthcare for all Ontarians. In
-particular, because the Health Regions cover larger geographical areas
-with various degrees of rurality and representation of equity-deserving
-groups, there is a need for decision-makers to consider that the
-evidence collected during the COVID-19 indicates a lack of
-infrastructure that can support personalized care for marginalized
-individuals
-
-This study explored differences in COVID-19 vaccination across the
-province of Ontario during the last quarter of 2021 taking into
-consideration socio-economic factors, such as income and race, their
-interactions, and the Health Regions within the province. Our results
-show that during the period analyzed, differences in vaccination uptake
-existed across multiple equity-deserving groups in the province, and
-that these differences were significant in two of the Health Regions
-analyzed. It is important that future public health policies in Ontario
-take into consideration how to adequately reach individuals from
-equity-deserving groups that might live in areas of the province where
-access to healthcare might be difficult. Only in this way the goal of
-the Health Region model, which aims at reducing disparities, will become
-successful.
+The implementation of the Health Regions in Ontario aimed at reducing
+the bureaucratic complexity and health disparities identified under the
+LHIN model. However, there are currently multiple challenges that need
+to be addressed to ensure that the new model is able to improve
+healthcare for the inhabitants of the province. First, the fact that
+each Health Region now covers a large geographical area that was served
+by multiple LHINs in the past creates a complex socio-demographic
+landscape that is different in each case due the different levels of
+rurality and representation of equity-deserving groups that are now
+within each Health Region. So far, the evidence collected during the
+COVID-19 pandemic indicates that differences in vaccination uptake are
+associated to a lack of infrastructure and resources that can adequately
+support personalized care to marginalized individuals. In the near
+future, health decision-makers will need to consider the implementation
+of policies that are focused on addressing this problematic.
+
+Moreover, the recent nature in the adaption of the Health Region poses a
+challenge for researchers in the acquisition of data and information
+that can be used to analyze the performance of the new system. From one
+side, the Health Regions have not been incorporated as part of Census
+data (LHINs were considered before in the Census), and this impact the
+amount and level of detail of available information. Currently, the only
+demographic information available for each Health Region is provided by
+Ontario Health (the agency that administers the Health Regions) but this
+information only provides general estimates that do not allow for
+detailed analyses on performance indicators (such as hospitalizations,
+readmissions, and trends in chronic disease incidence) between the
+regions. Without open information, it is impossible to assess the level
+of success of the Health Region model, which is critical considering
+that such evaluations have not been part of the Annual Reports of the
+Auditor General of Ontario, which in the past analyzed the performance
+of the LHINs and pointed to ongoing needs and failures in the system.
+
+The Health Region model will only by successful if it ensures that
+healthcare improves across all segments of the population of Ontario,
+particularly in the event of a future public health emergency or
+pandemic where so far, based on the experience of the COVID-19 pandemic,
+equity-deserving individuals have been disproportionately affected.
 

### 2023-05-19 code/data_preparation_May_08.R

@@ -16,13 +16,13 @@ rawdata<-read.csv(here("data","Fields_data.csv"),na.strings = c(""))
 rawdata <- rawdata %>%
   select(age,age_group,q09_hh_income,q16_race,city,q02_first_dose,complete,time_date_code,q13_hh_size)
 
-miss_raw<-miss_var_summary(rawdata, show_pct = TRUE)
+#miss_raw<-miss_var_summary(rawdata, show_pct = TRUE)
 
 ## summary of missing data in the raw dataset
 
-miss_raw %>%
-  kableExtra::kbl() %>%
-  kableExtra::kable_styling()
+# miss_raw %>%
+#   kableExtra::kbl() %>%
+#   kableExtra::kable_styling()
 
 # rates: first dose 
 
@@ -34,9 +34,9 @@ miss_raw %>%
 clean_data<-rawdata %>%
   subset(complete=="complete") 
 
-miss_var_summary(clean_data,show_pct=TRUE) %>%
-  kableExtra::kbl() %>%
-  kableExtra::kable_styling()
+# miss_var_summary(clean_data,show_pct=TRUE) %>%
+#   kableExtra::kbl() %>%
+#   kableExtra::kable_styling()
 
 # remove cities entered as "None"
 
@@ -270,9 +270,9 @@ clean_data <- clean_data%>%
 clean_data$Month <- fct_relevel(clean_data$Month,"October","November","December","January")
 
 
-# Remove certain Health REgions due to low number of obs: only 254 in North East
-# and North West Regions at this poing, in contrast with about a thousand from each 
-# othe region
+# Remove certain Health Regions due to low number of observations: only 254 in North East
+# and North West Regions at this point, in contrast with about a thousand from each 
+# other region
 
 #get health regions to eliminate
 clean_data <-clean_data %>%

### 2023-05-19 code/data_preparation_map.R

@@ -1,24 +1,230 @@
-## data preparation for map, including the North West and North East Health Regions
+## Script to clean and prepare the raw dataset to use for map creation
+## Updated on May 19, 2023
 
-# this script takes the clean dataset generated from the data cleaning process and does some preparatory steps such as 
-# variable recoding, creating a variable for the dependent variable in the regression, and creating 
-# a summary table of the dataset
+rawdata<-read.csv(here("data","Fields_data.csv"),na.strings = c(""))
+#clean_data <-read.csv(here("data","clean_dataset.csv"))
 
-data<-read.csv(here("data","clean_dataset.csv"))
+## select the variables used in the analysis
 
-# remove rows with missing observations for race
+## household income
+## race
+## city
+## age group
+## month
 
-clean_data<-data%>%drop_na(race)
 
-#remove "municipal status" as is not part of the analysis (to avoid unexpected errors)
+rawdata <- rawdata %>%
+  select(age,age_group,q09_hh_income,q16_race,city,q02_first_dose,complete,time_date_code,q13_hh_size)
+
+
+# The dataset had a column called "complete". The data dictionary was not
+# clear about what this meant, it seems that observations that are "complete" 
+# have complete observations for vaccination status, and all other covariates, except race, with a 4%
+# of missing obs.
+
+clean_data<-rawdata %>%
+  subset(complete=="complete") 
+
+# remove cities entered as "None"
 
 clean_data<-clean_data %>%
-  select(-Municipal.status) %>%
-  rename(Race=race,Age_group=age_group,Income=income_ord)
+  subset(city!="None")
 
-#make variables of interest factors
-clean_data<-clean_data %<>%
-  mutate_at(c("Income","Race","Age_group","Health_Region"),factor)
+#recode variable names
+
+clean_data <- clean_data %>%
+  rename(date=time_date_code,
+         first_dose=q02_first_dose,
+         income=q09_hh_income,
+         race=q16_race,
+  )%>%
+  mutate(across(c("age_group",
+                  "city",
+                  "income",
+                  "race"),as.factor))
+
+
+## Recode income
+
+clean_data<-clean_data%>%
+  mutate(income_ord = case_when(
[…]

### 2023-05-19 code/map.R

@@ -24,13 +24,19 @@ library(sf)
 library(viridis)
 library(here)
 library(tidyverse)
+library(lubridate)
 library(ggspatial)
 library(magrittr)
 library(scico)
+library(RColorBrewer)
 
 thm1<-scale_fill_viridis_d(option="mako",begin=0,end=1,direction = 1)
 
-thm2<-scale_fill_scico_d(palette = "nuuk",begin=0,end=1,direction = -1)
+thm2<-scale_fill_scico_d(palette = "bukavu",begin=0,end=1,direction = -1)
+
+thm3<-c("#e7d4e8","#d9f0d3","#762a83","#af8dc3","#1b7837","#7fbf7b")
+
+thm4<-scale_fill_brewer(palette="PRGn",direction =-1,type="diverging")
 
 mytheme<-theme(axis.line=element_blank(),
                axis.text.x=element_blank(),
@@ -38,7 +44,8 @@ mytheme<-theme(axis.line=element_blank(),
                axis.ticks=element_blank(),
                axis.title.x=element_blank(),
                axis.title.y=element_blank(),
-               legend.position="right",
+               legend.position="bottom",
+               legend.text=element_text(10),
                panel.background=element_blank(),
                panel.border=element_blank(),
                panel.grid.major=element_blank(),
@@ -124,35 +131,59 @@ source(here::here("code","data_preparation_map.R"))
 # get percentage of having received the vaccine by city
 
 
-perc<-clean_data %>% group_by(city,Health_Region) %>% 
+perc<-clean_data %>% 
+  group_by(city,Health_Region) %>% 
   summarise(Percent = sum(first_dose == "yes")/n())
 
-## for plotting the map, fix two names that are different between the clean dataset
-## and the data from Ontario:  Amherstburg(is mispelled Amhertsburg) and Big Trout Lake 
-## (called Kitchenuhmaykoosib Aaki 84 in the dataset)
+## for plotting the map, fix five names that are different between the clean dataset
+## and the data from Ontario:  Amherstburg(is mispelled Amhertsburg), Big Trout Lake 
+## (called Kitchenuhmaykoosib Aaki 84 in the dataset), Orleans (spelled Orléans),
+## Newcastle (appears as New Castle Village), Mitchell/Ontario (appears as "Mitchell"),
+## also, Fallingbrook did not appear in the dataset from Ontario and is not included in the
[…]

### 2023-05-23 manuscript_v5/01-background.qmd
minor changes

@@
 
-As of May of 2023 there have been 765 million confirmed cases of COVID-19 around the world, including 6.8 million deaths [@WHO-Covid].
+As of May of 2023, there have been 765 million confirmed cases of COVID-19 around the world, including 6.8 million deaths [@WHO-Covid].
 Although this disease is no longer categorized as a global health emergency by the World Health Organization (WHO) [@rigby2023], there is ongoing concern due to continued transmission, surges in cases and deaths due to new variants [@un2023], and weaknesses in health systems around the world that could be exploited by a novel virus or another public health emergency in the future [@mackey2021].
@@
 In particular, a major weakness that has received attention during the pandemic has been related to inequalities in vaccine uptake.
-The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic due to the start of vaccination campaigns in certain parts of the world toward the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020]) but inequalities in vaccine uptake made these pharmaceutical interventions ultimately unable to replicate the experience of smallpox, where vaccination on a global scale and was crucial to control this disease [@kayser2021].
+The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic due to the start of vaccination campaigns in certain parts of the world toward the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020]) but inequalities in vaccine uptake made these pharmaceutical interventions ultimately unable to replicate the success of the smallpox vaccination program, which was crucial to control this disease [@kayser2021].
 
 This problematic is a multifaceted issue resulting from a combination of factors, among which are failed public health measures [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021; @tamey2022], and vaccine hesitancy [@nafilyan2021].
-Furthermore, it is well established that this issue has affected in particular individuals in certain equity-deserving groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
+Furthermore, it is well established that this issue has affected in particular individuals in certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
 
 Reasons given for this inequality have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
-However, it is important to  also consider that vaccination uptake can be influenced by geographical (spatial) factors.
+However, it is important also to consider that vaccination uptake can be influenced by geographical (spatial) factors.
 In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], spatial differences in vaccine access and supply, vaccination location availability, and lack of prioritization of areas where vulnerable groups reside [@bogoch2022; @nguyen2021].
@@
 
-However, such analyses have been carried mostly in territories outside of Canada, where available studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022].
+However, such analyses have been carried out mostly in territories outside of Canada, where available studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022].
 Therefore, there is a need for studies that explore spatial differences in vaccination within the Canadian territory and that consequently, can help identify disparities that need to be addressed within specific areas in each province.
 
-This need is specially important in the case of Ontario, the most populated province of Canada.
+This need is especially important in the case of Ontario, the most populated province of Canada.
 Between 2007 and 2019, Ontario managed healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integration Networks (LHINs), which aimed to provide an integrated health system for the province.
-However, this approach was complex and bureaucratic, and resulted in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators such as wait times and hospital readmissions, fragmented electronic health systems, the decline of performance indicators, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016].
+However, this approach was complex, bureaucratic, and resulted in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators such as wait times and hospital readmissions, fragmented electronic health systems, the decline of performance indicators, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016].
 Therefore, with the intent of better organizing and delivering care in late 2019 the provincial government eliminated the LHINs and incorporated the areas covered by them into six larger Health Regions (North East, North West, Central, Toronto, West, and East) [@dong2022].
 
-Because the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
+Because of the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
 In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
@@
 
-Therefore, in this study we hypothesized that there were differences in vaccination uptake between the different Health Regions of Ontario between October of 2021 and January of 2022.
-By including socio-economic factors in our analysis, we aimed at identifying in which groups these differences were significant in order to provide an assessment of the current state of healthcare access in Ontario.
+Therefore, in this study we aim to understand the differences in vaccination uptake between the different Health Regions of Ontario between October of 2021 and January of 2022.
+By including socio-economic factors in our analysis, we aimed to identify in which groups these differences were significant in order to provide an assessment of the current state of healthcare access in Ontario.

### 2023-05-23 manuscript_v5/02-methods.qmd
minor changes

@@
 By design, the survey allowed respondents to exit at any time and deployed the questions randomly, which resulted in $\approx$ 84\% of the observations having multiple missing answers or being incomplete.
-Therefore, we selected 6,343 observations that were labeled as "complete" in the dataset and that had answers for all covariates considered in our analysis.
+Therefore, we selected 6,343 observations that were labelled as "complete" in the dataset and that had answers for all covariates considered in our analysis.
 Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions).
-Finally, we removed outliers from the data (19 observations of individuals with household size of 1 and income above 110,000 CAD in the original dataset).
-After all the preliminary analyses indicated above, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering the period between October 1st,2021 and January 17, 2022.
+Finally, we removed outliers from the data (19 observations of individuals under 25 years of age with household size of 1 and income above 110,000 CAD in the original dataset, that were demeed to lack accuracy).
+After all the preliminary analyses, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering between October 1st,2021 and January 17, 2022.
 The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.
@@
 
-We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
+We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
 Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data.

### 2023-05-23 manuscript_v5/03-results.qmd
minor changes

@@
 @fig-models presents the estimates (as odd ratios) from the logistic regression models for vaccination status using the socio-demographic factors collected by the survey, and their interactions.
-Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of equity-deserving groups.
+Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of underrepresented groups.
 However, the magnitude of the estimates differed between the uncorrected and corrected models and more importantly, there were differences in the statistical significance of certain estimates before and after the correction.
-Specifically, the uncorrected model showed significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-models,B) but these were deemed non statistically-significant after the correction.
+Specifically, the uncorrected model showed significant differences in vaccination odds between the age groups considered, the East Health Region, Latin American individuals with a household income under CAD 25,000, and Indigenous individuals living in the Central Health Region (@fig-models,B) but these were deemed non statistically significant after the correction.
 
@@
 
-Interestingly, individuals in equity-deserving groups with a household income below CAD 25,000 had higher odds of vaccination (when compared to those with a household income above CAD 60,000).
+Interestingly, individuals in underrepresented groups with a household income below CAD 25,000 had higher odds of vaccination (when compared to those with a household income above CAD 60,000).
 This held true in the case of Arab/Middle Eastern individuals (OR=34, CI=[1.70,6.79]), Black individuals (OR=3.81, CI=[2.05, 7.09]), and those in other racial or ethnic groups (OR=3.19, CI=[2.00,5.09]).
-Additionally, individuals with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern and other racial ethnic groups also had higher odds of vaccination than their high-income peers (ORs and CIs=6.96 [2.67,18.16], and 3.5 [1.85,6.62]).
+Additionally, individuals with an income between CAD 25,000 and 59,999 in the Arab/Middle Eastern and other racial or ethnic groups also had higher odds of vaccination than their high-income peers (ORs and CIs=6.96 [2.67,18.16], and 3.5 [1.85,6.62]).
 

### 2023-05-23 manuscript_v5/04-discussion.qmd
minor changes

@@
 
-In this study we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions during between late 2021 and early 2022.
+In this study, we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions between late 2021 and early 2022.
 Our goal was to determine which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its mission of improving health access for all Ontarians.
+**making corrections up to this point**
 

### 2023-05-23 manuscript_v5/main.qmd
minor changes

@@
 The COVID-19 pandemic continues to be a worldwide public health concern.
-Although vaccines against this disease were rapidly developed, vaccination uptake has not ben equal across all the segments of the population.
+Although vaccines against this disease were rapidly developed, vaccination uptake has not been equal across all the segments of the population.
 In particular, it has been shown that there have been differences in vaccine uptake across different segments of the population.
-However, there are also differences in vaccination across geographical areas, which might be important to consider in the development of future public health policies against COVID-19.
+However, there are also differences in vaccination across geographical areas, which might be important to consider in the development of future public health vaccination policies.
 In this study, we examined the relationship between vaccination status (having received the first dose of a COVID-19 vaccine), and different socio-economic and geographical factors.
-Our results show that during the last three months of 2021, individuals in certain equity-deserving groups (visible minorities) were three times less likely to be vaccinated than White/Caucasian individuals across the province and that in some cases, within these groups individuals in low income brackets had significantly higher odds of vaccination when compared to their peers in high income brackets.
-Finally, we identified significantly lower odds of vaccination in the West Health Region of Ontario within certain equity-deserving groups.
-This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population of Ontario that have been largely impacted by the pandemic.
+Our results show that between October of 2021 and January of 2022, individuals from underrepresented communities were three times less likely to be vaccinated than White/Caucasian individuals across the province of Ontario in Canada, and that in some cases, within these groups, individuals in low income brackets had significantly higher odds of vaccination when compared to their peers in high income brackets.
+Finally, we identified significantly lower odds of vaccination in the Central, East and West Health Regions of Ontario within certain underrepresented groups.
+This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population of Ontario that the pandemic has largely impacted.
 

### 2023-05-26 manuscript_v5/02-methods.qmd
working on map

@@
 
-We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto, and which ran between September 30th, 2021 and January 17th,2022.
+We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto (under protocol 00043317), and which ran between September 30th, 2021 and January 17th,2022.
 The survey collected socio-economic information from participants (@tbl-descriptive-stats), their location (nearest municipality,  as shown in @fig-map), the date of access to the survey, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".

### 2023-05-26 manuscript_v5/03-results.qmd
working on map

@@
 
-Finally, significantly lower odds of vaccination were identified (when compared to the Toronto Health Region) for Black individuals in the Central Health Region (OR=0.39, CI=[0.2,0.75]), Arab/Middle Eastern individuals in the East Health Region (OR=0.41 [0.17, 0.98]), and in the Indigenous and mixed groups in the West Health Region (ORs and CIs=[0.31 [0.14, 0.7] and 0.38 [0.19, 0.76], respectively).
+Finally, **the place of habitation affected the odds of vaccination for certain underrepresented groups, as significantly lower odds of vaccination were identified for the interaction between Health Region and race** in the case of Black individuals in the Central Health Region (OR=0.39, CI=[0.2,0.75]), Arab/Middle Eastern individuals in the East Health Region (OR=0.41 [0.17, 0.98]), and in the Indigenous and mixed groups in the West Health Region (ORs and CIs=[0.31 [0.14, 0.7] and 0.38 [0.19, 0.76], respectively).
 

### 2023-05-26 manuscript_v5/04-discussion.qmd
working on map

@@
 
-Lower vaccine uptake in the socio-demographic groups indicated above may be influenced in part, by vaccine hesitancy and refusal, which have been associated in equity-deserving Canadian individuals to concerns on vaccine safety, effectiveness, and experiences of racial discrimination in health settings [@basta2022;@cnat2022a;@cnat2022b;@cnat2023].
+Lower vaccine uptake in the socio-demographic groups indicated above may be influenced in part, by vaccine hesitancy and refusal, which have been associated in underrepresented Canadian individuals with concerns on vaccine safety, effectiveness, and experiences of racial discrimination in health settings [@basta2022;@cnat2022a;@cnat2022b;@cnat2023].
 However, it has been shown that structural barriers also play an important role in vaccination uptake.
-In the case of equity-deserving individuals, such barriers include complex scheduling systems, language barriers, lack of adequate public transportation, and lack of accessible vaccination sites [@njoku2021].
-In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with demand [@gill2022].
-This suggests that the observed differences are associated to disparities in vaccine access that were present during the period covered by the survey.
+In the case of underrepresented individuals, such barriers include complex scheduling systems, language barriers, lack of adequate public transportation, and lack of accessible vaccination sites [@njoku2021].
+In this regard, it is interesting to note that vaccination venues were scarce in low socio-economic areas that had the highest burden of COVID-19 in Toronto and other regions of Ontario around the time covered by the survey [@bogoch2022;@iveniuk2021], and that pharmacies in the Peel region (an area identified as a "hotspot" with high numbers of essential workers and multigenerational households) could not keep up with vaccine demand [@gill2022].
+This suggests that the observed differences are associated with disparities in vaccine access that were present during the period covered by the survey.
 
-Interestingly, whereas overall self-reported vaccination rates were found to be statistically significantly lower in various racial minority groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in income brackets below CAD 60,000 (Supplementary Figure A-5).
+Interestingly, whereas overall self-reported vaccination rates were found to be statistically significantly lower in various underrepresented groups when compared to White/Caucasian individuals, the change in odds of vaccination within certain racial groups and income strata was actually positive, in contrast to the White/Caucasian group, where vaccination odds decreased in income brackets below CAD 60,000 (Supplementary Figure A-5).
 Specifically, individuals in low income brackets that belonged to Arab/Middle Eastern, Black, or other minority groups had higher odds of vaccination that their peers with an income above 60,000 CAD.
 
-This result is likely reflects in part the fact that individuals in racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include workers in the areas of grocery stores, gas stations, warehouses, distribution, and manufacturing, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
+This result is likely reflects in part the fact that individuals in underrepresented groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include workers in the areas of grocery stores, gas stations, warehouses, distribution, and manufacturing, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
 In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022].
-These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation in individuals of equity-deserving groups played an important role in increasing the odds of vaccination.
+These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation of individuals in underrepresented groups played an important role in increasing the odds of vaccination.
 
-However, the results also indicate that the place of habitation affected the odds of vaccination for certain equity-deserving groups (interaction term of Health Region and Race, @fig-models,B).
+However, the results also indicate that the place of habitation affected the odds of vaccination for certain underrepresented groups (interaction term of Health Region and Race, @fig-models,B).
 Specifically, this held true in the case of individuals identifying as Indigenous or with mixed racial background in the West Health Region, Black individuals in the Central Health Region, and Arab/Middle Eastern individuals in the East Health Region [@fig-models].
@@
 For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated 100,000 Indigenous individuals living in the area [@ontariohealth].
-Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada, as these rationales have been associated to observed lower vaccination rates among these groups [@smylie2022;@eissa2021].
-Taken together, this suggests that healthcare disparities specific to these equity-deserving groups in certain parts of the province impacted vaccination uptake, and highlights the need of investments in the Health Regions focused on resources, infrastructure, and specially personnel that can deliver personalized care to marginalized communities, as it has been shown that such efforts have improved trust in vaccination in visible minorities [@schafferderoo2020].
+Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada, as these rationales have been associated with observed lower vaccination rates among these groups [@smylie2022;@eissa2021].
+Taken together, this suggests that healthcare disparities specific to these underrepresented groups in certain parts of the province impacted vaccination uptake, and highlights the need of investments in the Health Regions focused on resources, infrastructure, and specially personnel that can deliver personalized care to marginalized communities, as it has been shown that such efforts have improved trust in vaccination in underrepresented groups elsewhere [@schafferderoo2020].
 
@@
 More granular corrections would be needed to obtain more accurate estimates.
-For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates that other racial groups.
-Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates but unfortunately at the moment this correction cannot be implemented as such stratification is has not been implemented in the Census.
+For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates than other racial groups.
+Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates but unfortunately, at the moment this correction cannot be implemented as such stratification is has not been implemented in the Census.
 
 Additionally, our analysis did not consider the North West and North East Health Regions, due to the low number of entries from these areas in the survey (@fig-map).
-Low representation is expected as these regions as they only account for 5\% of the total population of Ontario, but in contrast, they have the highest proportion of Indigenous inhabitants [@ontariohealth].
-In the context of personalized care, there is a need for collecting data that focuses on these Health Regions where additional health disparities might be present and possibly understudied.
+Low representation is expected as these regions as they only account for 5\% of the total population of Ontario.
+However, these areas have the highest proportion of Indigenous inhabitants [@ontariohealth].
+In the context of personalized care, there is a need to collect data that focuses on these Health Regions where additional health disparities might be present and possibly understudied.
 
 The results in this study are based on self-reported data, where bias might be present.
-However, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid sample of vaccination uptake in the province.
-This is supported by the statistically-significant higher vaccination odds identified for January of 2022 in the model, which are consistent with province-wide trends reported by Public Health Ontario (which show a 4\% increase between early December and January, in contrast to a 2.5\% increase between October and November [@ontario-covid]); however, the short time window constitutes essentially a "snapshot" view of the evolution of the disease, and additional data would be needed in order to obtain estimates per racial/ethnic group over time across all Health Regions that can help inform the existence of other health disparities.
+However, in the context of COVID-19, it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], we believe that our data was able to provide a valid sample of vaccination uptake in the province.
+This is supported by the statistically significant higher vaccination odds that were identified for January of 2022 in the model, which are consistent with province-wide trends reported by Public Health Ontario (which show a 4\% increase between early December and January, in contrast to a 2.5\% increase between October and November [@ontario-covid]); however, the short time window constitutes essentially a "snapshot" view of the evolution of the disease, and additional data would be needed to obtain estimates per racial/ethnic group over time across all Health Regions that can help inform the existence of other health disparities.
 

### 2023-05-26 manuscript_v5/05-conclusion.qmd
working on map

@@
 The implementation of the Health Regions in Ontario aimed at reducing the bureaucratic complexity and health disparities identified under the LHIN model.
-However, there are currently multiple challenges that need to be addressed to ensure that the new model is able to improve healthcare for the inhabitants of the province.
+However, there are currently multiple challenges that need to be addressed to ensure that the new model can improve healthcare for the inhabitants of the province.
 First, the fact that each Health Region now covers a large geographical area that was served by multiple LHINs in the past creates a complex socio-demographic landscape that is different in each case due the different levels of rurality and representation of equity-deserving groups that are now within each Health Region.
@@
 From one side, the Health Regions have not been incorporated as part of Census data (LHINs were considered before in the Census), and this impact the amount and level of detail of available information.
-Currently, the only demographic information available for each Health Region is provided by Ontario Health (the agency that administers the Health Regions) but this information only provides general estimates that do not allow for detailed analyses on performance indicators (such as hospitalizations, readmissions, and trends in chronic disease incidence) between the regions.
+Currently, the only demographic information available for each Health Region is provided by Ontario Health (the agency that administers the Health Regions) but this information only provides general estimates that do not allow for detailed analyses of performance indicators (such as hospitalizations, readmissions, and trends in chronic disease incidence) between the regions.
 Without open information, it is impossible to assess the level of success of the Health Region model, which is critical considering that such evaluations have not been part of the Annual Reports of the Auditor General of Ontario, which in the past analyzed the performance of the LHINs and pointed to ongoing needs and failures in the system.
 
-The Health Region model will only by successful if it ensures that healthcare improves across all segments of the population of Ontario, particularly in the event of a future public health emergency or pandemic where so far, based on the experience of the COVID-19 pandemic, equity-deserving individuals have been disproportionately affected.
+The Health Region model will only by successful if it ensures that healthcare improves across all segments of the population of Ontario, particularly in the event of a future public health emergency or pandemic where so far, based on the experience of the COVID-19 pandemic, underrepresented individuals have been disproportionately affected.

### 2023-05-26 manuscript_v5/appendix.qmd
working on map

@@
 ---
-title: '**Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata**'
-subtitle: Appendix
-authors:
+title: '**Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata** \\ Appendix'
+author:
   - name: Ariel Mundo Ortiz
+    affil-id: 1,2,3
+  - name: Bouchra Nasri
+    affil-id: 1,2,3,\*
 affiliations: 
-      - ref: crm
-      - ref: espum
-    email: ariel.mundo.ortiz@umontreal.ca
-    orcid: 0000-0002-6014-4538
-  - name: Bouchra Nasri
-    affiliations:
-      - ref: crm
-      - ref: espum
-    email: bouchra.nasri@umontreal.ca
-    corresponding: true
-affiliations:
-  - id: crm
+  - id: 1
     name: Centre de Recherches Mathématiques, University of Montreal, Montréal, Canada
-  - id: espum
+  - id: 2
     name: Department of Social and Preventive Medicine, École de Santé Publique, University of Montreal, Montréal, Canada
-filters:
-  - authors-block
+  - id: 3
+    name: Centre de recherche en santé publique, University of Montreal, Montréal, Canada
+  - id: \*
+    name: Corresponding author, \href{mailto:bouchra.nasri@umontreal.ca}{bouchra.nasri@umontreal.ca}
 format: 
   pdf:
+    keep-tex: true
+    template-partials: 
+      - latex/title.tex
     include-in-header:
       - text: |
-         \usepackage{placeins}
-    include-before-body: 
+         \usepackage[noblocks]{authblk}
+         \renewcommand*{\Authsep}{, }
+         \renewcommand*{\Authand}{, }
+         \renewcommand*{\Authands}{, }
+         \renewcommand\Affilfont{\small}
       - file: latex/appendix_preamble.sty
 linestretch: 2
-keep-tex: true
 ---

### 2023-05-26 manuscript_v5/appendix.tex
working on map

@@
 \documentclass[
-  letterpaper,
-  DIV=11,
-  numbers=noendperiod]{scrartcl}
+]{article}
 
@@
 \usepackage{longtable}
+%start sections and page numbers with A
 \usepackage{placeins}
-\KOMAoption{captions}{tableheading}
+\setcounter{page}{1}
+\renewcommand\thesection{A}
+\renewcommand\thesubsection{\thesection.\arabic{subsection}}
+\renewcommand{\thepage}{A-\arabic{page}}
+\renewcommand{\thetable}{A-\arabic{table}}
+\renewcommand\thefigure{\thesection-\arabic{figure}}
+
+\usepackage[noblocks]{authblk}
+\renewcommand*{\Authsep}{, }
+\renewcommand*{\Authand}{, }
+\renewcommand*{\Authands}{, }
+\renewcommand\Affilfont{\small}
 \makeatletter
@@
 \hypersetup{
-  pdftitle={Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata},
-  pdfauthor={Ariel Mundo Ortiz1,2; Bouchra Nasri1,2,},
+  pdftitle={Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata \textbackslash{} Appendix},
+  pdfauthor={Ariel Mundo Ortiz; Bouchra Nasri},
   colorlinks=true,
@@
 \title{\textbf{Differences in COVID-19 vaccination in the province of
-Ontario across Health Regions and socio-economic strata}}
-\usepackage{etoolbox}
-\makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
-\makeatother
-\subtitle{Appendix}
-\author{}
+Ontario across Health Regions and socio-economic strata}
+\textbackslash{} Appendix}
+
+\author[1,2,3]{Ariel Mundo Ortiz}
+\author[1,2,3,*]{Bouchra Nasri}
+
+\affil[1]{Centre de Recherches Mathématiques, University of Montreal,
+Montréal, Canada}
+\affil[2]{Department of Social and Preventive Medicine, École de Santé
+Publique, University of Montreal, Montréal, Canada}
+\affil[3]{Centre de recherche en santé publique, University of Montreal,
+Montréal, Canada}
+\affil[*]{Corresponding author, \url{bouchra.nasri@umontreal.ca}}
+
 \date{}
-
 \begin{document}
 \maketitle
-%start sections and page numbers with A
-\setcounter{page}{1}
-\renewcommand\thesection{A}
-\renewcommand\thesubsection{\thesection.\arabic{subsection}}
-\renewcommand{\thepage}{A-\arabic{page}}
-\renewcommand{\thetable}{A-\arabic{table}}
-\renewcommand\thefigure{\thesection-\arabic{figure}}
-
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, frame hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
 \setstretch{2}
-\textsuperscript{1} Centre de Recherches Mathématiques, University of
-Montreal, Montréal, Canada\\
-\textsuperscript{2} Department of Social and Preventive Medicine, École
-de Santé Publique, University of Montreal, Montréal, Canada
-
-\textsuperscript{*} Correspondence:
-\href{mailto:bouchra.nasri@umontreal.ca}{Bouchra Nasri
-\textless{}bouchra.nasri@umontreal.ca\textgreater{}}
-
 This document provides details on assessing the socio-economic

### 2023-05-26 manuscript_v5/main.qmd
working on map

@@
 title: '**Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata**'
-authors:
+author:
   - name: Ariel Mundo Ortiz
+    affil-id: 1,2,3
+  - name: Bouchra Nasri
+    affil-id: 1,2,3,\*
 affiliations: 
-      - ref: crm
-      - ref: espum
-    email: ariel.mundo.ortiz@umontreal.ca
-    orcid: 0000-0002-6014-4538
-  - name: Bouchra Nasri
-    affiliations:
-      - ref: crm
-      - ref: espum
-    email: bouchra.nasri@umontreal.ca
-    corresponding: true
-affiliations:
-  - id: crm
+  - id: 1
     name: Centre de Recherches Mathématiques, University of Montreal, Montréal, Canada
-  - id: espum
+  - id: 2
     name: Department of Social and Preventive Medicine, École de Santé Publique, University of Montreal, Montréal, Canada
-filters:
-  - authors-block
+  - id: 3
+    name: Centre de recherche en santé publique, University of Montreal, Montréal, Canada
+  - id: \*
+    name: Corresponding author, \url{bouchra.nasri@umontreal.ca}
 format: 
   pdf:
+    keep-tex: true
+    template-partials: 
+      - latex/title.tex
     include-in-header:
       - text: |
-          \usepackage{rotfloat}
+         \usepackage[noblocks]{authblk}
+         \renewcommand*{\Authsep}{, }
+         \renewcommand*{\Authand}{, }
+         \renewcommand*{\Authands}{, }
+         \renewcommand\Affilfont{\small}
       - file: latex/preamble.sty
-    #mainfont: "Palatino Linotype"
-    keep-tex: true
+# authors:
+#   - name: Ariel Mundo Ortiz
+#     affiliations:
+#       - ref: crm
+#       - ref: espum
+#     email: ariel.mundo.ortiz@umontreal.ca
+#     orcid: 0000-0002-6014-4538
+#   - name: Bouchra Nasri
+#     affiliations:
+#       - ref: crm
+#       - ref: espum
+#     email: bouchra.nasri@umontreal.ca
+#     corresponding: true
+# affiliations:
+#   - id: crm
+#     name: Centre de Recherches Mathématiques, University of Montreal, Montréal, Canada
+#   - id: espum
+#     name: Department of Social and Preventive Medicine, École de Santé Publique, University of Montreal, Montréal, Canada 
+# filters:
+#   - authors-block
+# format: 
+#   pdf:
+    # include-in-header:
+    #   - text: |
+    #       \usepackage{rotfloat}
+    #   - file: latex/preamble.sty
+    # #mainfont: "Palatino Linotype"
+    # keep-tex: true
     link-citations: true #adds links to the citations
@@
 
+# Acknowledgments
+
+This work was supported by the Fonds de recherche du Québec Scholar Program (J1 in Artificial Intelligence and Digital Health, BN), the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant Program (RGPID-560523-2020, BN), the Mathematics for Public Health (MfPH) Emerging Infectious Diseases Modelling Initiative (BN, AM), and the OMNI Emerging Infectious Disease Modelling Initiative (BN).
+
+# Conflicts of Interest
+
+The authors declare no conflict of interest.
+
 # References

### 2023-05-26 manuscript_v5/main.tex
working on map

@@
 
-\usepackage{rotfloat}
-\KOMAoption{captions}{tableheading}
+\usepackage[noblocks]{authblk}
+\renewcommand*{\Authsep}{, }
+\renewcommand*{\Authand}{, }
+\renewcommand*{\Authands}{, }
+\renewcommand\Affilfont{\small}
 \makeatletter
@@
   pdftitle={Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata},
-  pdfauthor={Ariel Mundo Ortiz1,2; Bouchra Nasri1,2,},
+  pdfauthor={Ariel Mundo Ortiz; Bouchra Nasri},
   colorlinks=true,
@@
 Ontario across Health Regions and socio-economic strata}}
-\author{}
+
+\author[1,2,3]{Ariel Mundo Ortiz}
+\author[1,2,3,*]{Bouchra Nasri}
+
+\affil[1]{Centre de Recherches Mathématiques, University of Montreal,
+Montréal, Canada}
+\affil[2]{Department of Social and Preventive Medicine, École de Santé
+Publique, University of Montreal, Montréal, Canada}
+\affil[3]{Centre de recherche en santé publique, University of Montreal,
+Montréal, Canada}
+\affil[*]{Corresponding author, \url{bouchra.nasri@umontreal.ca}}
+
 \date{}
-
 \begin{document}
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, breakable, interior hidden, sharp corners, enhanced, frame hidden]}{\end{tcolorbox}}\fi
-
-\textsuperscript{1} Centre de Recherches Mathématiques, University of
-Montreal, Montréal, Canada\\
-\textsuperscript{2} Department of Social and Preventive Medicine, École
-de Santé Publique, University of Montreal, Montréal, Canada
-
-\textsuperscript{*} Correspondence:
-\href{mailto:bouchra.nasri@umontreal.ca}{Bouchra Nasri
-\textless{}bouchra.nasri@umontreal.ca\textgreater{}}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 Although vaccines against this disease were rapidly developed,
-vaccination uptake has not ben equal across all the segments of the
+vaccination uptake has not been equal across all the segments of the
 population. In particular, it has been shown that there have been
@@
 geographical areas, which might be important to consider in the
-development of future public health policies against COVID-19. In this
-study, we examined the relationship between vaccination status (having
-received the first dose of a COVID-19 vaccine), and different
-socio-economic and geographical factors. Our results show that during
-the last three months of 2021, individuals in certain equity-deserving
-groups (visible minorities) were three times less likely to be
-vaccinated than White/Caucasian individuals across the province and that
-in some cases, within these groups individuals in low income brackets
-had significantly higher odds of vaccination when compared to their
-peers in high income brackets. Finally, we identified significantly
-lower odds of vaccination in the West Health Region of Ontario within
-certain equity-deserving groups. This study shows that there is an
-ongoing need to better understand and address differences in vaccination
-uptake across diverse segments of the population of Ontario that have
-been largely impacted by the pandemic.
+development of future public health vaccination policies. In this study,
+we examined the relationship between vaccination status (having received
+the first dose of a COVID-19 vaccine), and different socio-economic and
+geographical factors. Our results show that between October of 2021 and
+January of 2022, individuals from underrepresented communities were
+three times less likely to be vaccinated than White/Caucasian
+individuals across the province of Ontario in Canada, and that in some
+cases, within these groups, individuals in low income brackets had
+significantly higher odds of vaccination when compared to their peers in
+high income brackets. Finally, we identified significantly lower odds of
+vaccination in the Central, East and West Health Regions of Ontario
+within certain underrepresented groups. This study shows that there is
+an ongoing need to better understand and address differences in
+vaccination uptake across diverse segments of the population of Ontario
+that the pandemic has largely impacted.
 
@@
 
-As of May of 2023 there have been 765 million confirmed cases of
+As of May of 2023, there have been 765 million confirmed cases of
 COVID-19 around the world, including 6.8 million
@@
 but inequalities in vaccine uptake made these pharmaceutical
-interventions ultimately unable to replicate the experience of smallpox,
-where vaccination on a global scale and was crucial to control this
+interventions ultimately unable to replicate the success of the smallpox
+vaccination program, which was crucial to control this
 disease\textsuperscript{\protect\hyperlink{ref-kayser2021}{9}}.
@@
 Furthermore, it is well established that this issue has affected in
-particular individuals in certain equity-deserving groups (e.g., Black,
+particular individuals in certain underrepresented groups (e.g., Black,
 Asian, or Indigenous) as well as individuals with socio-economic
@@
 theories\textsuperscript{\protect\hyperlink{ref-mosby2021}{21}--\protect\hyperlink{ref-freeman2020}{23}}.
-However, it is important to also consider that vaccination uptake can be
+However, it is important also to consider that vaccination uptake can be
 influenced by geographical (spatial) factors. In this regard,
@@
 
-However, such analyses have been carried mostly in territories outside
-of Canada, where available studies have been focused in certain cities
-(such as Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{31}},
-or Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{32}}),
-or have explored differences at a province-wide
+However, such analyses have been carried out mostly in territories
+outside of Canada, where available studies have been focused in certain
+cities (such as
+Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{31}}, or
+Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{32}}), or
+have explored differences at a province-wide
 level\textsuperscript{\protect\hyperlink{ref-guay2022}{18}}. Therefore,
@@
 
-This need is specially important in the case of Ontario, the most
+This need is especially important in the case of Ontario, the most
 populated province of Canada. Between 2007 and 2019, Ontario managed
@@
 provide an integrated health system for the province. However, this
-approach was complex and bureaucratic, and resulted in excessive
+approach was complex, bureaucratic, and resulted in excessive
 expenditures, disparities in mortality rates, the deterioration of
@@
 
-Because the relatively recent adoption of the Health Region model and
+Because of the relatively recent adoption of the Health Region model and
 its alignment with the onset of the COVID-19 pandemic, there is a need
@@
 
-Therefore, in this study we hypothesized that there were differences in
+Therefore, in this study we aim to understand the differences in
 vaccination uptake between the different Health Regions of Ontario
 between October of 2021 and January of 2022. By including socio-economic
-factors in our analysis, we aimed at identifying in which groups these
+factors in our analysis, we aimed to identify in which groups these
 differences were significant in order to provide an assessment of the
@@
 COVID-19 Task Force under ethical guidance from the University of
-Toronto, and which ran between September 30th, 2021 and January
-17th,2022. The survey collected socio-economic information from
-participants (Table~\ref{tbl-descriptive-stats}), their location
-(nearest municipality, as shown in Figure~\ref{fig-map}), the date of
-access to the survey, and asked information on vaccination status by
-using the question ``Have you received the first dose of the COVID
-vaccine?'', with possible answers ``yes'' and ``no''. The original
+Toronto (under protocol 00043317), and which ran between September 30th,
+2021 and January 17th,2022. The survey collected socio-economic
+information from participants (Table~\ref{tbl-descriptive-stats}), their
+location (nearest municipality, as shown in Figure~\ref{fig-map}), the
+date of access to the survey, and asked information on vaccination
+status by using the question ``Have you received the first dose of the
+COVID vaccine?'', with possible answers ``yes'' and ``no''. The original
 dataset contained 39,029 observations.
@@
 the observations having multiple missing answers or being incomplete.
-Therefore, we selected 6,343 observations that were labeled as
+Therefore, we selected 6,343 observations that were labelled as
 ``complete'' in the dataset and that had answers for all covariates
@@
 corresponding to the North West and North East Health Regions). Finally,
-we removed outliers from the data (19 observations of individuals with
-household size of 1 and income above 110,000 CAD in the original
-dataset). After all the preliminary analyses indicated above, the total
-number of observations used for analysis was 6,236 and included the
-East, Central, Toronto, and West Health Regions covering the period
-between October 1st,2021 and January 17, 2022. The original dataset,
-clean dataset, and details on the data cleaning process are described in
-detail in the \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub
-repository} for this paper.
+we removed outliers from the data (19 observations of individuals under
+25 years of age with household size of 1 and income above 110,000 CAD in
+the original dataset, that were demeed to lack accuracy). After all the
+preliminary analyses, the total number of observations used for analysis
+was 6,236 and included the East, Central, Toronto, and West Health
+Regions covering between October 1st,2021 and January 17, 2022. The
+original dataset, clean dataset, and details on the data cleaning
+process are described in detail in the
+\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
+for this paper.
 
@@
 Regions in vaccination rates while considering the socio-economic
-factors and and months covered by the survey
+factors and months covered by the survey
 (Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
@@
 identified in both cases in individuals characterized by a low household
-income, or that identified as part of equity-deserving groups. However,
+income, or that identified as part of underrepresented groups. However,
 the magnitude of the estimates differed between the uncorrected and
@@
 Health Region (Figure~\ref{fig-models},B) but these were deemed non
-statistically-significant after the correction.
+statistically significant after the correction.
 
@@
 
-Interestingly, individuals in equity-deserving groups with a household
+Interestingly, individuals in underrepresented groups with a household
 income below CAD 25,000 had higher odds of vaccination (when compared to
@@
 individuals with an income between CAD 25,000 and 59,999 in the
-Arab/Middle Eastern and other racial ethnic groups also had higher odds
-of vaccination than their high-income peers (ORs and CIs=6.96
+Arab/Middle Eastern and other racial or ethnic groups also had higher
+odds of vaccination than their high-income peers (ORs and CIs=6.96
 {[}2.67,18.16{]}, and 3.5 {[}1.85,6.62{]}).
 
-Finally, significantly lower odds of vaccination were identified (when
-compared to the Toronto Health Region) for Black individuals in the
-Central Health Region (OR=0.39, CI={[}0.2,0.75{]}), Arab/Middle Eastern
-individuals in the East Health Region (OR=0.41 {[}0.17, 0.98{]}), and in
-the Indigenous and mixed groups in the West Health Region (ORs and
-CIs={[}0.31 {[}0.14, 0.7{]} and 0.38 {[}0.19, 0.76{]}, respectively).
+Finally, \textbf{the place of habitation affected the odds of
+vaccination for certain underrepresented groups, as significantly lower
+odds of vaccination were identified for the interaction between Health
+Region and race} in the case of Black individuals in the Central Health
+Region (OR=0.39, CI={[}0.2,0.75{]}), Arab/Middle Eastern individuals in
+the East Health Region (OR=0.41 {[}0.17, 0.98{]}), and in the Indigenous
+and mixed groups in the West Health Region (ORs and CIs={[}0.31 {[}0.14,
+0.7{]} and 0.38 {[}0.19, 0.76{]}, respectively).
 
@@
 
-In this study we hypothesized that differences in COVID-19 vaccination
-uptake were present between the Health Regions during between late 2021
-and early 2022. Our goal was to determine which socio-demographic groups
+In this study, we hypothesized that differences in COVID-19 vaccination
+uptake were present between the Health Regions between late 2021 and
+early 2022. Our goal was to determine which socio-demographic groups
 could be impacted by these disparities in order to provide
@@
 Health Region model is able to fulfill its mission of improving health
-access for all Ontarians.
+access for all Ontarians. \textbf{making corrections up to this point}
 
@@
 be influenced in part, by vaccine hesitancy and refusal, which have been
-associated in equity-deserving Canadian individuals to concerns on
+associated in underrepresented Canadian individuals with concerns on
 vaccine safety, effectiveness, and experiences of racial discrimination
@@
 However, it has been shown that structural barriers also play an
-important role in vaccination uptake. In the case of equity-deserving
+important role in vaccination uptake. In the case of underrepresented
 individuals, such barriers include complex scheduling systems, language
@@
 ``hotspot'' with high numbers of essential workers and multigenerational
-households) could not keep up with
+households) could not keep up with vaccine
 demand\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. This
-suggests that the observed differences are associated to disparities in
-vaccine access that were present during the period covered by the
+suggests that the observed differences are associated with disparities
+in vaccine access that were present during the period covered by the
 survey.
@@
 Interestingly, whereas overall self-reported vaccination rates were
-found to be statistically significantly lower in various racial minority
-groups when compared to White/Caucasian individuals, the change in odds
-of vaccination within certain racial groups and income strata was
-actually positive, in contrast to the White/Caucasian group, where
-vaccination odds decreased in income brackets below CAD 60,000
-(Supplementary Figure A-5). Specifically, individuals in low income
-brackets that belonged to Arab/Middle Eastern, Black, or other minority
-groups had higher odds of vaccination that their peers with an income
-above 60,000 CAD.
+found to be statistically significantly lower in various
+underrepresented groups when compared to White/Caucasian individuals,
+the change in odds of vaccination within certain racial groups and
+income strata was actually positive, in contrast to the White/Caucasian
+group, where vaccination odds decreased in income brackets below CAD
+60,000 (Supplementary Figure A-5). Specifically, individuals in low
+income brackets that belonged to Arab/Middle Eastern, Black, or other
+minority groups had higher odds of vaccination that their peers with an
+income above 60,000 CAD.
 
 This result is likely reflects in part the fact that individuals in
-racial minority groups tend to perform occupations that have been deemed
-as ``essential'' in the context of the
+underrepresented groups tend to perform occupations that have been
+deemed as ``essential'' in the context of the
 pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{55},\protect\hyperlink{ref-ct2021}{56}},
@@
 elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{58}},
-suggest that the type of occupation in individuals of equity-deserving
+suggest that the type of occupation of individuals in underrepresented
 groups played an important role in increasing the odds of vaccination.
@@
 However, the results also indicate that the place of habitation affected
-the odds of vaccination for certain equity-deserving groups (interaction
+the odds of vaccination for certain underrepresented groups (interaction
 term of Health Region and Race, Figure~\ref{fig-models},B).
@@
 which is known to be the case for Indigenous and Black individuals in
-Canada, as these rationales have been associated to observed lower
+Canada, as these rationales have been associated with observed lower
 vaccination rates among these
@@
 Taken together, this suggests that healthcare disparities specific to
-these equity-deserving groups in certain parts of the province impacted
+these underrepresented groups in certain parts of the province impacted
 vaccination uptake, and highlights the need of investments in the Health
@@
 has been shown that such efforts have improved trust in vaccination in
-visible
-minorities\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{65}}.
+underrepresented groups
+elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{65}}.
 
@@
 these differences are likely to be driven by the proportion of
-White/Caucasian individuals, who had higher vaccination rates that other
+White/Caucasian individuals, who had higher vaccination rates than other
 racial groups. Correcting for each racial/ethnic group in each Health
 Region can provide a more accurate estimation of region-wide vaccination
-rates but unfortunately at the moment this correction cannot be
+rates but unfortunately, at the moment this correction cannot be
 implemented as such stratification is has not been implemented in the
@@
 these regions as they only account for 5\% of the total population of
-Ontario, but in contrast, they have the highest proportion of Indigenous
+Ontario. However, these areas have the highest proportion of Indigenous
 inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
-In the context of personalized care, there is a need for collecting data
+In the context of personalized care, there is a need to collect data
 that focuses on these Health Regions where additional health disparities
@@
 The results in this study are based on self-reported data, where bias
-might be present. However, because in the context of COVID-19 it has
-been shown that good agreement exists between self-reported and
-documented vaccination
+might be present. However, in the context of COVID-19, it has been shown
+that good agreement exists between self-reported and documented
+vaccination
 status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{66}}, we
 believe that our data was able to provide a valid sample of vaccination
-uptake in the province. This is supported by the
-statistically-significant higher vaccination odds identified for January
-of 2022 in the model, which are consistent with province-wide trends
+uptake in the province. This is supported by the statistically
+significant higher vaccination odds that were identified for January of
+2022 in the model, which are consistent with province-wide trends
 reported by Public Health Ontario (which show a 4\% increase between
@@
 view of the evolution of the disease, and additional data would be
-needed in order to obtain estimates per racial/ethnic group over time
-across all Health Regions that can help inform the existence of other
-health disparities.
+needed to obtain estimates per racial/ethnic group over time across all
+Health Regions that can help inform the existence of other health
+disparities.
 
@@
 LHIN model. However, there are currently multiple challenges that need
-to be addressed to ensure that the new model is able to improve
-healthcare for the inhabitants of the province. First, the fact that
-each Health Region now covers a large geographical area that was served
-by multiple LHINs in the past creates a complex socio-demographic
-landscape that is different in each case due the different levels of
-rurality and representation of equity-deserving groups that are now
-within each Health Region. So far, the evidence collected during the
-COVID-19 pandemic indicates that differences in vaccination uptake are
-associated to a lack of infrastructure and resources that can adequately
-support personalized care to marginalized individuals. In the near
-future, health decision-makers will need to consider the implementation
-of policies that are focused on addressing this problematic.
+to be addressed to ensure that the new model can improve healthcare for
+the inhabitants of the province. First, the fact that each Health Region
+now covers a large geographical area that was served by multiple LHINs
+in the past creates a complex socio-demographic landscape that is
+different in each case due the different levels of rurality and
+representation of equity-deserving groups that are now within each
+Health Region. So far, the evidence collected during the COVID-19
+pandemic indicates that differences in vaccination uptake are associated
+to a lack of infrastructure and resources that can adequately support
+personalized care to marginalized individuals. In the near future,
+health decision-makers will need to consider the implementation of
+policies that are focused on addressing this problematic.
 
@@
 information only provides general estimates that do not allow for
-detailed analyses on performance indicators (such as hospitalizations,
+detailed analyses of performance indicators (such as hospitalizations,
 readmissions, and trends in chronic disease incidence) between the
@@
 pandemic where so far, based on the experience of the COVID-19 pandemic,
-equity-deserving individuals have been disproportionately affected.
+underrepresented individuals have been disproportionately affected.
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+This work was supported by the Fonds de recherche du Québec Scholar
+Program (J1 in Artificial Intelligence and Digital Health, BN), the
+Natural Sciences and Engineering Research Council of Canada through the
+Discovery Grant Program (RGPID-560523-2020, BN), the Mathematics for
+Public Health (MfPH) Emerging Infectious Diseases Modelling Initiative
+(BN, AM), and the OMNI Emerging Infectious Disease Modelling Initiative
+(BN).
+
+\hypertarget{conflicts-of-interest}{%
+\section{Conflicts of Interest}\label{conflicts-of-interest}}
+
+The authors declare no conflict of interest.
 

### 2023-05-26 code/map.R

@@ -29,6 +29,7 @@ library(ggspatial)
 library(magrittr)
 library(scico)
 library(RColorBrewer)
+library(rayshader)
 
 thm1<-scale_fill_viridis_d(option="mako",begin=0,end=1,direction = 1)
 
@@ -158,27 +159,62 @@ perc$city[perc$city=="Mitchell/Ontario"] <- "Mitchell"
 #join to get coordinates for each city
 test<-left_join(perc,fort_on, by=c( "city" = "GEONAME"))
 
+## get centroid coordinates for LHINs
+
+centroids.df <- as.data.frame(coordinates(lhins1))
+
+# add labels and change order so only the ones to be plotted are in sequence
+
+centroids.df$label <- c(0,2,0,0,0,0,1,0,0,4,0,3,5,6)
+
+centroids.df<-centroids.df %>%
+  filter(label!=0)
+
+## add names of Health Regions
+
+centroids.df$names<-c("West","Toronto","East","Central", "North East","North West")
+  
+
+fort_lhin$z<-rep(100,times=nrow(fort_lhin))
 ## to label Health Regions
 # dff <- fort_lhin %>%
 #   group_by(Health_Region) %>%
 #   summarize(long = mean(long, na.rm = T), lat = mean(lat, na.rm = T), group = group)
-#   
+#
+
+#add height to the map
+
 m1<-ggplot(data=fort_lhin,aes(x=long, 
                               y=lat,
                               group=group,
                               fill=Health_Region))+
-  geom_polygon(color="black",size=0.2)+
+  geom_polygon(color="black",size=0.2,show.legend = FALSE)+
   geom_point(data=test,
              inherit.aes = FALSE,
              show.legend = FALSE,
-             color="blue",
-             fill="#e0ff33",
+             color="black",
+             fill="#fde725ff",
              shape=21,
              size=2.0,
              stroke=0.5,
              aes(x=long,
                  y=lat,
-                 alpha=0.9))+
+                 alpha=0.8))+
+  # geom_text(data=centroids.df, 
+  #           aes(label = label, x = V1, y = V2),
+  #           inherit.aes = FALSE,
+  #           size=8,
+  #           color="red") + 
+  geom_path(data=fort_lhin,aes(x=long, 
+                               y=lat,
+                               group=group,
+                               fill=Health_Region),
+            color="black",size=0.2)+
+  geom_contour(data=fort_lhin,
[…]

### 2023-06-01 manuscript_v5/02-methods.qmd
map updated, majority of changes made

@@
 #| fig-cap: "Geographic representation of the data collected by the _Survey of COVID-19 related Behaviours and Attitudes_, collected by the Fields Institute in Ontario.
-The municipalities from where survey participants provided answers appear as points.
-The Health six Regions are color-coded.
-Internal boundaries within certain Health Regions indicate areas that belonged to the Local Integrated Health Networks (LHINs), the geographic areas for healthcare in Ontario before the adoption of the Health Regions."
+Municipalities from where survey participants provided answers appear as points, color indicates number of observation obtained from each city.
+The Health six Regions are color-coded and labelled sequentially.
+Internal boundaries within certain Health Regions indicate areas previously covered by the Local Integrated Health Networks (LHINs)."
 #| message: false
@@
 #| out-height: 100%
+#| out-width: 100%
 
-knitr::include_graphics(here("data","map_data","map_cairo.pdf"))
+knitr::include_graphics(here("data","map_data","map_June_01.pdf"))
 

### 2023-06-01 manuscript_v5/main.qmd
map updated, majority of changes made

@@
   pdf:
+    geometry:
+      - top=20mm
+      - left=20mm
+      - right=20mm
+      - bottom=20mm
     keep-tex: true

### 2023-06-01 manuscript_v5/main.tex
map updated, majority of changes made

@@
 \usepackage{xcolor}
+\usepackage[top=20mm,left=20mm,right=20mm,bottom=20mm]{geometry}
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 
-{\centering \includegraphics[width=\textwidth,height=1\textheight]{../data/map_data/map_cairo.pdf}
+{\centering \includegraphics[width=1\textwidth,height=1\textheight]{../data/map_data/map_June_01.pdf}
 
@@
 by the \emph{Survey of COVID-19 related Behaviours and Attitudes},
-collected by the Fields Institute in Ontario. The municipalities from
-where survey participants provided answers appear as points. The Health
-six Regions are color-coded. Internal boundaries within certain Health
-Regions indicate areas that belonged to the Local Integrated Health
-Networks (LHINs), the geographic areas for healthcare in Ontario before
-the adoption of the Health Regions.}
+collected by the Fields Institute in Ontario. Municipalities from where
+survey participants provided answers appear as points, color indicates
+number of observation obtained from each city. The Health six Regions
+are color-coded and labelled sequentially. Internal boundaries within
+certain Health Regions indicate areas previously covered by the Local
+Integrated Health Networks (LHINs).}
 
@@
 
-Finally, \textbf{the place of habitation affected the odds of
-vaccination for certain underrepresented groups, as significantly lower
-odds of vaccination were identified for the interaction between Health
-Region and race} in the case of Black individuals in the Central Health
-Region (OR=0.39, CI={[}0.2,0.75{]}), Arab/Middle Eastern individuals in
-the East Health Region (OR=0.41 {[}0.17, 0.98{]}), and in the Indigenous
-and mixed groups in the West Health Region (ORs and CIs={[}0.31 {[}0.14,
+Finally, the place of habitation affected the odds of vaccination for
+certain underrepresented groups, as significantly lower odds of
+vaccination were identified for the interaction between Health Region
+and race in the case of Black individuals in the Central Health Region
+(OR=0.39, CI={[}0.2,0.75{]}), Arab/Middle Eastern individuals in the
+East Health Region (OR=0.41 {[}0.17, 0.98{]}), and in the Indigenous and
+mixed groups in the West Health Region (ORs and CIs={[}0.31 {[}0.14,
 0.7{]} and 0.38 {[}0.19, 0.76{]}, respectively).
@@
 decision-makers with information that could be used to develop policies
-focused on reducing or eliminating these differences and ensure that the
-Health Region model is able to fulfill its mission of improving health
-access for all Ontarians. \textbf{making corrections up to this point}
+focused on reducing or eliminating these differences and ensuring that
+the Health Region model is able to fulfill its mission of improving
+health access for all Ontarians.
 

### 2023-06-01 code/data_preparation_map.R

@@ -201,20 +201,20 @@ clean_data<-clean_data %>%
   mutate(LHIN = coalesce(LHIN.x,LHIN.y)) %>%
   select(-c(LHIN.x,LHIN.y,Full_title.x,Geographic.area.y,Geographic_area_title.y))%>%
   mutate(Health_Region=
-           case_when(LHIN=="Central"~ "Central",
-                     LHIN=="Central West"~"Central",
-                     LHIN=="Mississauga Halton"~"Central",
-                     LHIN=="North Simcoe Muskoka"~"Central",
-                     LHIN=="Central East"~"East",
-                     LHIN=="South East"~"East",
-                     LHIN=="Champlain"~"East",
-                     LHIN=="North East"~"North East",
-                     LHIN=="North West"~"North West",
-                     LHIN=="Toronto Central"~"Toronto",
-                     LHIN=="South West"~"West",
-                     LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)"~"West",
-                     LHIN=="Waterloo Wellington"~"West",
-                     LHIN=="Erie St. Clair"~"West"
+           case_when(Health_Region=="Central"~ "3.Central",
+                     Health_Region=="Central West"~"3.Central",
+                     Health_Region=="Mississauga Halton"~"3.Central",
+                     Health_Region=="North Simcoe Muskoka"~"3.Central",
+                     Health_Region=="Central East"~"4.East",
+                     Health_Region=="South East"~"4.East",
+                     Health_Region=="Champlain"~"4.East",
+                     Health_Region=="North East"~"5.North East",
+                     Health_Region=="North West"~"6.North West",
+                     Health_Region=="Toronto Central"~"1.Toronto",
+                     Health_Region=="South West"~"2.West",
+                     Health_Region=="Hamilton Niagara Haldimand Brant"~"2.West",
+                     Health_Region=="Waterloo Wellington"~"2.West",
+                     Health_Region=="Erie St. Clair"~"2.West"
            ))%>%
   rename(Geographic_area=Geographic.area.x,Geographic_area_title=Geographic_area_title.x)
 

### 2023-06-01 code/map_June_01.R

@@ -186,14 +186,6 @@ centroids.df$names<-c("West","Toronto","East","Central", "North East","North Wes
 centroids.df$names <- as.factor(paste(centroids.df$label, ".", centroids.df$names))
 # 
 
-#fort_lhin$z<-rep(100,times=nrow(fort_lhin))
-## to label Health Regions
-# dff <- fort_lhin %>%
-#   group_by(Health_Region) %>%
-#   summarize(long = mean(long, na.rm = T), lat = mean(lat, na.rm = T), group = group)
-#
-
-#add height to the map
 breaks <- c(1,10,100,2000)
 
 m1<-ggplot(data=fort_lhin,aes(x=long, 
@@ -205,20 +197,6 @@ m1<-ggplot(data=fort_lhin,aes(x=long,
                size=0.2,
                alpha=0.8,
                show.legend = FALSE)+
-  # geom_point(data=test,
-  #            aes(x=long,
-  #                y=lat,
-  #                color=Percent
-  #                ),
-  #            #alpha=0.8,
-  #            inherit.aes = FALSE,
-  #            show.legend = TRUE,
-  #            #color="black",
-  #            #fill="#fde725ff",
-  #            shape=21,
-  #            size=2.0,
-  #            stroke=0.5,
-  #            )+
   geom_point(data=test,
              aes(x=long,
                  y=lat,
@@ -226,10 +204,7 @@ m1<-ggplot(data=fort_lhin,aes(x=long,
              ),
              alpha=1,
              inherit.aes = FALSE,
-             show.legend = TRUE,
-             #color="black",
-             #shape=1,
-             #fill="cyan"
+             show.legend = TRUE
              )+
   geom_path(data=fort_lhin,aes(x=long,
                                y=lat,
@@ -251,8 +226,7 @@ m1<-ggplot(data=fort_lhin,aes(x=long,
                    inherit.aes = FALSE,
                    aes(label = label,
                        x = V1,
-                       y = V2,
-                       #fill=names
+                       y = V2
                    ),
                    nudge_y = 10,
                    size=3,
@@ -268,9 +242,7 @@ m1<-ggplot(data=fort_lhin,aes(x=long,
              alpha=1,
              inherit.aes = FALSE,
              show.legend = FALSE,
-             color="#f2ea0a",
-             #shape=1,
-             #fill="cyan"
+             color="#f2ea0a"
   )+
   annotation_scale(
                    location = "tr",
[…]

### 2023-06-09 manuscript_v5/02-methods.qmd
updated manuscript following BN comments

@@
 
-By design, the survey allowed respondents to exit at any time and deployed the questions randomly, which resulted in $\approx$ 84\% of the observations having multiple missing answers or being incomplete.
+By design, the survey allowed respondents to exit at any time and deployed the questions randomly.
+This resulted in $\approx$ 84\% of the observations having multiple missing answers or being incomplete due to participants leaving the survey at different stages.
 Therefore, we selected 6,343 observations that were labelled as "complete" in the dataset and that had answers for all covariates considered in our analysis.
 Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions).
-Finally, we removed outliers from the data (19 observations of individuals under 25 years of age with household size of 1 and income above 110,000 CAD in the original dataset, that were demeed to lack accuracy).
 After all the preliminary analyses, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering between October 1st,2021 and January 17, 2022.

### 2023-06-09 manuscript_v5/main.tex
updated manuscript following BN comments

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 By design, the survey allowed respondents to exit at any time and
-deployed the questions randomly, which resulted in \(\approx\) 84\% of
-the observations having multiple missing answers or being incomplete.
-Therefore, we selected 6,343 observations that were labelled as
-``complete'' in the dataset and that had answers for all covariates
-considered in our analysis. Later, we matched the city of each
-observation with its corresponding LHIN and Health Region, and removed
-observations from areas with low representation (254 observations
-corresponding to the North West and North East Health Regions). Finally,
-we removed outliers from the data (19 observations of individuals under
-25 years of age with household size of 1 and income above 110,000 CAD in
-the original dataset, that were demeed to lack accuracy). After all the
-preliminary analyses, the total number of observations used for analysis
-was 6,236 and included the East, Central, Toronto, and West Health
-Regions covering between October 1st,2021 and January 17, 2022. The
-original dataset, clean dataset, and details on the data cleaning
-process are described in detail in the
+deployed the questions randomly. This resulted in \(\approx\) 84\% of
+the observations having multiple missing answers or being incomplete due
+to participants leaving the survey at different stages. Therefore, we
+selected 6,343 observations that were labelled as ``complete'' in the
+dataset and that had answers for all covariates considered in our
+analysis. Later, we matched the city of each observation with its
+corresponding LHIN and Health Region, and removed observations from
+areas with low representation (254 observations corresponding to the
+North West and North East Health Regions). After all the preliminary
+analyses, the total number of observations used for analysis was 6,236
+and included the East, Central, Toronto, and West Health Regions
+covering between October 1st,2021 and January 17, 2022. The original
+dataset, clean dataset, and details on the data cleaning process are
+described in detail in the
 \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
@@
 challenge for researchers in the acquisition of data and information
-that can be used to analyze the performance of the new system. From one
+that can be used to analyze the performance of the new model. From one
 side, the Health Regions have not been incorporated as part of Census
 data (LHINs were considered before in the Census), and this impact the
-amount and level of detail of available information. Currently, the only
-demographic information available for each Health Region is provided by
-Ontario Health (the agency that administers the Health Regions) but this
-information only provides general estimates that do not allow for
-detailed analyses of performance indicators (such as hospitalizations,
-readmissions, and trends in chronic disease incidence) between the
-regions. Without open information, it is impossible to assess the level
-of success of the Health Region model, which is critical considering
-that such evaluations have not been part of the Annual Reports of the
-Auditor General of Ontario, which in the past analyzed the performance
-of the LHINs and pointed to ongoing needs and failures in the system.
+amount and level of detail of available information. On the other hand,
+evaluations of the Health Region model are at the moment not considered
+in the Annual Reports of the Auditor General of Ontario, which was
+critical in the past to identify limitations in the LHIN model.
+Currently, the only demographic information available for each Health
+Region is provided by Ontario Health (the agency that administers the
+Health Regions) but this information only provides general estimates
+that do not allow for detailed analyses of performance indicators (such
+as hospitalizations, readmissions, and trends in chronic disease
+incidence) between the Regions. Therefore, there is a pressing need for
+open data that can be used by researchers and decision-makers to examine
+the performance of the Health Region model.
 

### 2023-06-27 manuscript_v5/01-background.qmd
v6 after SW's comments

@@
 
-In particular, a major weakness that has received attention during the pandemic has been related to inequalities in vaccine uptake.
-The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic due to the start of vaccination campaigns in certain parts of the world toward the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020]) but inequalities in vaccine uptake made these pharmaceutical interventions ultimately unable to replicate the success of the smallpox vaccination program, which was crucial to control this disease [@kayser2021].
+In particular, a major weakness that has received attention during the pandemic has been related to vaccination.
+The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic as vaccination campaigns in certain parts of the world started toward the end of 2020[@thelancet2021;@davis2022;@bogoch2022;@tanne2020].
+However, although it has been estimated that COVID-19 vaccines have been able to prevent millions of deaths worldwide [@watson2022], they were ultimately unable to bring the pandemic to a rapid end, in contrast to previous experiences (such as in the case of smallpox), where the global vaccination program was crucial to control this disease[@kayser2021].
 
-This problematic is a multifaceted issue resulting from a combination of factors, among which are failed public health measures [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021; @tamey2022], and vaccine hesitancy [@nafilyan2021].
-Furthermore, it is well established that this issue has affected in particular individuals in certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
+This failure of vaccines to stop the pandemic is a complex issue resulting from a combination of factors, among which are inadequate public health measures and the emergence of new variants [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population[@willis2021;@skirrow2022].
+In particular, it is well established that lower vaccination uptake has been observed in individuals within certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
 
-Reasons given for this inequality have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
+Reasons given for this inequality in vaccination uptake have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
 However, it is important also to consider that vaccination uptake can be influenced by geographical (spatial) factors.

### 2023-06-27 manuscript_v6/01-background.qmd
v6 after SW's comments

--- 2023-06-27 manuscript_v5/01-background.qmd
@@
 In particular, a major weakness that has received attention during the pandemic has been related to vaccination.
-The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic as vaccination campaigns in certain parts of the world started toward the end of 2020[@thelancet2021;@davis2022;@bogoch2022;@tanne2020].
-However, although it has been estimated that COVID-19 vaccines have been able to prevent millions of deaths worldwide [@watson2022], they were ultimately unable to bring the pandemic to a rapid end, in contrast to previous experiences (such as in the case of smallpox), where the global vaccination program was crucial to control this disease[@kayser2021].
+The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic as vaccination campaigns in certain parts of the world started as early as the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020].
+However, although it has been estimated that COVID-19 vaccines have been able to prevent millions of deaths worldwide [@watson2022], their original goal, which was to bring the pandemic to a rapid end, was not accomplished.
+This contrast with previous vaccination experiences, such as in the case of smallpox, where the implementation of a global vaccination program was crucial to control this disease[@kayser2021].
 
-This failure of vaccines to stop the pandemic is a complex issue resulting from a combination of factors, among which are inadequate public health measures and the emergence of new variants [@li2021], inequality in vaccine access between high- and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population[@willis2021;@skirrow2022].
+The failure of vaccines to stop the pandemic is a complex issue resulting from a combination of factors, among which are inadequate public health measures and the emergence of new variants [@li2021], inequality in vaccine access between high-income and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population[@willis2021;@skirrow2022].
 In particular, it is well established that lower vaccination uptake has been observed in individuals within certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
 
-Reasons given for this inequality in vaccination uptake have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of conspiracy theories [@bogart2021;@mosby2021;@freeman2020].
+Reasons given for this inequality in vaccination uptake have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of misinformation and disinformation [@bogart2021;@mosby2021;@freeman2020].
 However, it is important also to consider that vaccination uptake can be influenced by geographical (spatial) factors.
-In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], spatial differences in vaccine access and supply, vaccination location availability, and lack of prioritization of areas where vulnerable groups reside [@bogoch2022; @nguyen2021].
-Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], and that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022].
+In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], lack of vaccine access in areas where underrepresented groups reside[@bogoch2022], as well as an absence of prioritization of areas inhabited by vulnerable groups [@nguyen2021].
+Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], indicating that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022].
 
@@
 
+On the other hand, public health is managed in the province by Public Health Ontario, which was established in late 2007 and is currently composed of 34 Public Health Units (PHUs).
+Importantly, during the COVID-19 pandemic, these PHUS led the distribution of the vaccine within the areas they covered [@rotenberg2021]
+
+- add about intersection of Health Regions and PHUs. How vaccination presented a challenge as vaccine was ultimately administered in facilites managed by OHT?
+
+https://www.mcmasterforum.org/docs/default-source/product-documents/rapid-responses/examining-the-intersections-between-ontario-health-teams-and-public-health.pdf?sfvrsn=69caa8c9_11
+
 Because of the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.

### 2023-06-27 manuscript_v6/03-results.qmd
v6 after SW's comments

--- 2023-05-23 manuscript_v5/03-results.qmd
@@
 
-Finally, significantly lower odds of vaccination were identified (when compared to the Toronto Health Region) for Black individuals in the Central Health Region (OR=0.39, CI=[0.2,0.75]), Arab/Middle Eastern individuals in the East Health Region (OR=0.41 [0.17, 0.98]), and in the Indigenous and mixed groups in the West Health Region (ORs and CIs=[0.31 [0.14, 0.7] and 0.38 [0.19, 0.76], respectively).
+Finally, the place of habitation affected the odds of vaccination for certain underrepresented groups, as significantly lower odds of vaccination were identified for the interaction between Health Region and race in the case of Black individuals in the Central Health Region (OR=0.39, CI=[0.2,0.75]), Arab/Middle Eastern individuals in the East Health Region (OR=0.41 [0.17, 0.98]), and in the Indigenous and mixed groups in the West Health Region (ORs and CIs=[0.31 [0.14, 0.7] and 0.38 [0.19, 0.76], respectively).
 

### 2023-06-27 manuscript_v6/04-discussion.qmd
v6 after SW's comments

--- 2023-05-26 manuscript_v5/04-discussion.qmd
@@
 In this study, we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions between late 2021 and early 2022.
-Our goal was to determine which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensure that the Health Region model is able to fulfill its mission of improving health access for all Ontarians.
-**making corrections up to this point**
+Our goal was to determine which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensuring that the Health Region model is able to fulfill its mission of improving health access for all Ontarians.
 

### 2023-06-27 manuscript_v6/05-conclusion.qmd
v6 after SW's comments

--- 2023-06-09 manuscript_v5/main.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
-\documentclass[
-]{article}
+# Conclusion
 
-\usepackage{amsmath,amssymb}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
-\usepackage{xcolor}
-\usepackage[top=20mm,left=20mm,right=20mm,bottom=20mm]{geometry}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+The implementation of the Health Regions in Ontario aimed at reducing the bureaucratic complexity and health disparities identified under the LHIN model.
+However, there are currently multiple challenges that need to be addressed to ensure that the new model can improve healthcare for the inhabitants of the province.
+First, the fact that each Health Region now covers a large geographical area that was served by multiple LHINs in the past creates a complex socio-demographic landscape that is different in each case due the different levels of rurality and representation of equity-deserving groups that are now within each Health Region.
+So far, the evidence collected during the COVID-19 pandemic indicates that differences in vaccination uptake are associated to a lack of infrastructure and resources that can adequately support personalized care to marginalized individuals.
+In the near future, health decision-makers will need to consider the implementation of policies that are focused on addressing this problematic.
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-\usepackage{graphicx}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-% Set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+Moreover, the recent nature in the adaption of the Health Region poses a challenge for researchers in the acquisition of data and information that can be used to analyze the performance of the new model.
+From one side, the Health Regions have not been incorporated as part of Census data (LHINs were considered before in the Census), and this impact the amount and level of detail of available information.
+On the other hand, evaluations of the Health Region model are at the moment not considered in the Annual Reports of the Auditor General of Ontario, which was critical in the past to identify limitations in the LHIN model.
+Currently, the only demographic information available for each Health Region is provided by Ontario Health (the agency that administers the Health Regions) but this information only provides general estimates that do not allow for detailed analyses of performance indicators (such as hospitalizations, readmissions, and trends in chronic disease incidence) between the Regions.
+Therefore, there is a pressing need for open data that can be used by researchers and decision-makers to examine the performance of the Health Region model.
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-\usepackage{caption}
-%line numbers
-%\usepackage{mathpazo}
-%\usepackage{rotating}
-\usepackage{lineno}
-\linenumbers
-
-\usepackage[noblocks]{authblk}
-\renewcommand*{\Authsep}{, }
-\renewcommand*{\Authand}{, }
-\renewcommand*{\Authands}{, }
-\renewcommand\Affilfont{\small}
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\AtBeginDocument{%
-\ifdefined\contentsname
-  \renewcommand*\contentsname{Table of contents}
-\else
-  \newcommand\contentsname{Table of contents}
-\fi
-\ifdefined\listfigurename
-  \renewcommand*\listfigurename{List of Figures}
-\else
-  \newcommand\listfigurename{List of Figures}
-\fi
-\ifdefined\listtablename
-  \renewcommand*\listtablename{List of Tables}
-\else
-  \newcommand\listtablename{List of Tables}
-\fi
-\ifdefined\figurename
-  \renewcommand*\figurename{Figure}
-\else
-  \newcommand\figurename{Figure}
-\fi
-\ifdefined\tablename
-  \renewcommand*\tablename{Table}
-\else
-  \newcommand\tablename{Table}
-\fi
-}
-\@ifpackageloaded{float}{}{\usepackage{float}}
-\floatstyle{ruled}
-\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
-\floatname{codelisting}{Listing}
-\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata},
-  pdfauthor={Ariel Mundo Ortiz; Bouchra Nasri},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-
-\title{\textbf{Differences in COVID-19 vaccination in the province of
-Ontario across Health Regions and socio-economic strata}}
-
-\author[1,2,3]{Ariel Mundo Ortiz}
-\author[1,2,3,*]{Bouchra Nasri}
-
-\affil[1]{Centre de Recherches Mathématiques, University of Montreal,
-Montréal, Canada}
-\affil[2]{Department of Social and Preventive Medicine, École de Santé
-Publique, University of Montreal, Montréal, Canada}
-\affil[3]{Centre de recherche en santé publique, University of Montreal,
-Montréal, Canada}
-\affil[*]{Corresponding author, \url{bouchra.nasri@umontreal.ca}}
-
-\date{}
-\begin{document}
-\maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
-
-\hypertarget{abstract}{%
-\section{Abstract}\label{abstract}}
-
-The COVID-19 pandemic continues to be a worldwide public health concern.
-Although vaccines against this disease were rapidly developed,
-vaccination uptake has not been equal across all the segments of the
-population. In particular, it has been shown that there have been
-differences in vaccine uptake across different segments of the
-population. However, there are also differences in vaccination across
-geographical areas, which might be important to consider in the
-development of future public health vaccination policies. In this study,
-we examined the relationship between vaccination status (having received
-the first dose of a COVID-19 vaccine), and different socio-economic and
-geographical factors. Our results show that between October of 2021 and
-January of 2022, individuals from underrepresented communities were
-three times less likely to be vaccinated than White/Caucasian
-individuals across the province of Ontario in Canada, and that in some
-cases, within these groups, individuals in low income brackets had
-significantly higher odds of vaccination when compared to their peers in
-high income brackets. Finally, we identified significantly lower odds of
-vaccination in the Central, East and West Health Regions of Ontario
-within certain underrepresented groups. This study shows that there is
-an ongoing need to better understand and address differences in
-vaccination uptake across diverse segments of the population of Ontario
-that the pandemic has largely impacted.
-
-\hypertarget{keywords}{%
-\section*{Keywords}\label{keywords}}
-\addcontentsline{toc}{section}{Keywords}
-
-Covid-19, vaccination, survey, socio-economic factors, visible
-minorities.
-
-\hypertarget{background}{%
-\section{Background}\label{background}}
-
-As of May of 2023, there have been 765 million confirmed cases of
-COVID-19 around the world, including 6.8 million
-deaths\textsuperscript{\protect\hyperlink{ref-WHO-Covid}{1}}. Although
-this disease is no longer categorized as a global health emergency by
-the World Health Organization
-(WHO)\textsuperscript{\protect\hyperlink{ref-rigby2023}{2}}, there is
-ongoing concern due to continued transmission, surges in cases and
-deaths due to new
-variants\textsuperscript{\protect\hyperlink{ref-un2023}{3}}, and
-weaknesses in health systems around the world that could be exploited by
-a novel virus or another public health emergency in the
-future\textsuperscript{\protect\hyperlink{ref-mackey2021}{4}}.
-
-In particular, a major weakness that has received attention during the
-pandemic has been related to inequalities in vaccine uptake. The rapid
-development of vaccines against COVID-19 initially brought the hope of a
-rapid end to the pandemic due to the start of vaccination campaigns in
-certain parts of the world toward the end of
-2020\textsuperscript{\protect\hyperlink{ref-thelancet2021}{5}--\protect\hyperlink{ref-tanne2020}{8}})
-but inequalities in vaccine uptake made these pharmaceutical
-interventions ultimately unable to replicate the success of the smallpox
-vaccination program, which was crucial to control this
-disease\textsuperscript{\protect\hyperlink{ref-kayser2021}{9}}.
-
-This problematic is a multifaceted issue resulting from a combination of
-factors, among which are failed public health
-measures\textsuperscript{\protect\hyperlink{ref-li2021}{10}}, inequality
-in vaccine access between high- and low-income
-countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{11},\protect\hyperlink{ref-tamey2022}{12}},
-and vaccine
-hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{13}}.
-Furthermore, it is well established that this issue has affected in
-particular individuals in certain underrepresented groups (e.g., Black,
-Asian, or Indigenous) as well as individuals with socio-economic
-disadvantages\textsuperscript{\protect\hyperlink{ref-willis2021}{14}--\protect\hyperlink{ref-hussain2022}{20}}.
-
-Reasons given for this inequality have included medical mistrust due to
-systemic medical
-racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{16},\protect\hyperlink{ref-mosby2021}{21}},
-mistrust in
-vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{14}}, and
-the influence of conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-mosby2021}{21}--\protect\hyperlink{ref-freeman2020}{23}}.
-However, it is important also to consider that vaccination uptake can be
-influenced by geographical (spatial) factors. In this regard,
-differences in COVID-19 vaccination rates have been associated with
-varied regional attitudes towards
-vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{24}},
-spatial differences in vaccine access and supply, vaccination location
-availability, and lack of prioritization of areas where vulnerable
-groups
-reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-nguyen2021}{25}}.
-Other studies have also shown heterogeneity in vaccine uptake within
-small governmental administrative units such as
-counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{26}--\protect\hyperlink{ref-bhuiyan2022}{29}},
-and that accounting for geographical differences in vaccination can help
-predict patterns of booster
-uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{30}}.
-
-However, such analyses have been carried out mostly in territories
-outside of Canada, where available studies have been focused in certain
-cities (such as
-Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{31}}, or
-Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{32}}), or
-have explored differences at a province-wide
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{18}}. Therefore,
-there is a need for studies that explore spatial differences in
-vaccination within the Canadian territory and that consequently, can
-help identify disparities that need to be addressed within specific
-areas in each province.
-
-This need is especially important in the case of Ontario, the most
-populated province of Canada. Between 2007 and 2019, Ontario managed
-healthcare access to its inhabitants using 14 intra-provincial divisions
-called the Local Health Integration Networks (LHINs), which aimed to
-provide an integrated health system for the province. However, this
-approach was complex, bureaucratic, and resulted in excessive
-expenditures, disparities in mortality rates, the deterioration of
-certain performance indicators such as wait times and hospital
-readmissions, fragmented electronic health systems, the decline of
-performance indicators, and inequities in health services
-access\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33}--\protect\hyperlink{ref-lysyk2016}{37}}.
-Therefore, with the intent of better organizing and delivering care in
-late 2019 the provincial government eliminated the LHINs and
-incorporated the areas covered by them into six larger Health Regions
-(North East, North West, Central, Toronto, West, and
-East)\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}.
-
-Because of the relatively recent adoption of the Health Region model and
-its alignment with the onset of the COVID-19 pandemic, there is a need
-to analyze if there are ongoing disparities in health access under this
-approach that need to be addressed before they are exploited by a new
-disease or public health threat. In this regard, previous research has
-highlighted disparities in the level of activity of each Health
-Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{38}}.
-Therefore, analyzing differences in vaccination uptake within the Health
-Regions and can help identify which socio-demographic groups are the
-most vulnerable and what areas of the province deserve special attention
-by decision-makers.
-
-Therefore, in this study we aim to understand the differences in
-vaccination uptake between the different Health Regions of Ontario
-between October of 2021 and January of 2022. By including socio-economic
-factors in our analysis, we aimed to identify in which groups these
-differences were significant in order to provide an assessment of the
-current state of healthcare access in Ontario.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{sec-data}{%
-\subsection{Data and Methods}\label{sec-data}}
-
-We used data from the \emph{Survey of COVID-19 related Behaviours and
-Attitudes}, a repeated cross sectional survey focused on the Canadian
-province of Ontario that was commissioned by the Fields Institute for
-Research in Mathematical Sciences and the Mathematical Modelling of
-COVID-19 Task Force under ethical guidance from the University of
-Toronto (under protocol 00043317), and which ran between September 30th,
-2021 and January 17th,2022. The survey collected socio-economic
-information from participants (Table~\ref{tbl-descriptive-stats}), their
-location (nearest municipality, as shown in Figure~\ref{fig-map}), the
-date of access to the survey, and asked information on vaccination
-status by using the question ``Have you received the first dose of the
-COVID vaccine?'', with possible answers ``yes'' and ``no''. The original
-dataset contained 39,029 observations.
-
-By design, the survey allowed respondents to exit at any time and
-deployed the questions randomly. This resulted in \(\approx\) 84\% of
-the observations having multiple missing answers or being incomplete due
-to participants leaving the survey at different stages. Therefore, we
-selected 6,343 observations that were labelled as ``complete'' in the
-dataset and that had answers for all covariates considered in our
-analysis. Later, we matched the city of each observation with its
-corresponding LHIN and Health Region, and removed observations from
-areas with low representation (254 observations corresponding to the
-North West and North East Health Regions). After all the preliminary
-analyses, the total number of observations used for analysis was 6,236
-and included the East, Central, Toronto, and West Health Regions
-covering between October 1st,2021 and January 17, 2022. The original
-dataset, clean dataset, and details on the data cleaning process are
-described in detail in the
-\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
-for this paper.
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\textwidth,height=1\textheight]{../data/map_data/map_June_01.pdf}
-
-}
-
-\caption{\label{fig-map}Geographic representation of the data collected
-by the \emph{Survey of COVID-19 related Behaviours and Attitudes},
-collected by the Fields Institute in Ontario. Municipalities from where
-survey participants provided answers appear as points, color indicates
-number of observation obtained from each city. The Health six Regions
-are color-coded and labelled sequentially. Internal boundaries within
-certain Health Regions indicate areas previously covered by the Local
-Integrated Health Networks (LHINs).}
-
-\end{figure}
-
-\hypertarget{statistical-analyses}{%
-\subsection{Statistical analyses}\label{statistical-analyses}}
-
-We used a logistic regression model to examine the impact of the Health
-Regions in vaccination rates while considering the socio-economic
-factors and months covered by the survey
-(Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
-Health Region and Race and income), as previous studies have shown that
-socio-economic factors and their interactions are significant predictors
-of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{39}--\protect\hyperlink{ref-cnat2022a}{41}}.
-Because we identified differences in representativity between the survey
-data and the estimates from the Census, we used an iterative
-proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{42}}
-to correct the data using data from the Census and Health Region
-population totals; and fitted the regression model to the uncorrected
-and corrected data. Details regarding the correction can be found in the
-Appendix. All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{43}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{44}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{45}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{46}},
-and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{47}}.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{sample-characteristics}{%
-\subsection{Sample Characteristics}\label{sample-characteristics}}
-
-Table~\ref{tbl-descriptive-stats} shows the characteristics of the data
-from the Fields COVID-19 survey used for analysis. The sample contained
-6,236 observations, from which 24.8\% (1,547) corresponded to
-individuals that reported not having received the first dose of the
-vaccine. Vaccination rates ranged between 71-79\% across all household
-income brackets, age groups, Health Regions, and the months considered
-in the survey. However, the highest vaccination rates in each category
-were reported by individuals in the highest income bracket (79\%), those
-between 16 and 34 years of age (77\%), individuals that lived in the
-East Health Region (77\%), and during January of 2022 (78\%). Between
-racial/ethnic groups, the highest vaccination rate was reported by
-White/Caucasian individuals (84\%), against vaccination rates between
-63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous,
-Latin American individuals, and those that reported belonging to
-``Other'' racial groups, which included Southeast Asian, Filipino, West
-Asian, and minorities not identified elsewhere.
-
-\hypertarget{tbl-descriptive-stats}{}
-\setlength{\LTpost}{0mm}
-\begin{longtable}{lccc}
-\caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination
-Status) }\tabularnewline
-
-\toprule
-\textbf{Variable} & \textbf{no}, N = 1,547\textsuperscript{\textit{1}} & \textbf{yes}, N = 4,689\textsuperscript{\textit{1}} & \textbf{p-value}\textsuperscript{\textit{2}} \\ 
-\midrule
-Income (CAD) &  &  & <0.001 \\ 
-    60000 and above & 542 (21\%) & 1,996 (79\%) &  \\ 
-    25000-59999 & 347 (25\%) & 1,046 (75\%) &  \\ 
-    under 25000 & 658 (29\%) & 1,647 (71\%) &  \\ 
-Age Group &  &  & 0.002 \\ 
-    16-34 & 645 (23\%) & 2,117 (77\%) &  \\ 
-    35-54 & 411 (24\%) & 1,305 (76\%) &  \\ 
-    55 and over & 491 (28\%) & 1,267 (72\%) &  \\ 
-Health Region &  &  & 0.3 \\ 
-    Toronto & 593 (26\%) & 1,709 (74\%) &  \\ 
-    Central & 372 (26\%) & 1,083 (74\%) &  \\ 
-    East & 236 (23\%) & 783 (77\%) &  \\ 
-    West & 346 (24\%) & 1,114 (76\%) &  \\ 
-Month &  &  & <0.001 \\ 
-    October & 469 (27\%) & 1,263 (73\%) &  \\ 
-    November & 376 (28\%) & 980 (72\%) &  \\ 
-    December & 181 (24\%) & 565 (76\%) &  \\ 
-    January & 521 (22\%) & 1,881 (78\%) &  \\ 
-Race &  &  & <0.001 \\ 
-    White/Caucasian & 354 (16\%) & 1,871 (84\%) &  \\ 
-    Arab/Middle Eastern & 111 (34\%) & 220 (66\%) &  \\ 
-    Black & 159 (34\%) & 303 (66\%) &  \\ 
-    East Asian/Pacific Islander & 94 (19\%) & 404 (81\%) &  \\ 
-    Indigenous & 112 (37\%) & 194 (63\%) &  \\ 
-    Latin American & 99 (34\%) & 195 (66\%) &  \\ 
-    Mixed & 177 (30\%) & 411 (70\%) &  \\ 
-    Other\textsuperscript{\textit{3}} & 315 (34\%) & 606 (66\%) &  \\ 
-    South Asian & 126 (21\%) & 485 (79\%) &  \\ 
-\bottomrule
-\end{longtable}
-\begin{minipage}{\linewidth}
-\textsuperscript{\textit{1}}n (\%)\\
-\textsuperscript{\textit{2}}Pearson's Chi-squared test\\
-\textsuperscript{\textit{3}}Southeast Asian, Filipino, West Asian,
-and minorities not identified elsewhere according to the Census.\\
-\end{minipage}
-
-\hypertarget{multivariate-regression}{%
-\subsection{Multivariate Regression}\label{multivariate-regression}}
-
-Figure~\ref{fig-models} presents the estimates (as odd ratios) from the
-logistic regression models for vaccination status using the
-socio-demographic factors collected by the survey, and their
-interactions. Generally speaking, lower odds of vaccination were
-identified in both cases in individuals characterized by a low household
-income, or that identified as part of underrepresented groups. However,
-the magnitude of the estimates differed between the uncorrected and
-corrected models and more importantly, there were differences in the
-statistical significance of certain estimates before and after the
-correction. Specifically, the uncorrected model showed significant
-differences in vaccination odds between the age groups considered, the
-East Health Region, Latin American individuals with a household income
-under CAD 25,000, and Indigenous individuals living in the Central
-Health Region (Figure~\ref{fig-models},B) but these were deemed non
-statistically significant after the correction.
-
-\begin{figure}
-
-\includegraphics{main_files/figure-pdf/fig-models-1.pdf} \hfill{}
-
-\caption{\label{fig-models}Coefficient estimates and confidence
-intervals for the uncorrected model. Only statistically significant
-interaction terms are shown. Full interaction terms can be found in
-Supplementary Figures A-3 and A-4.}
-
-\end{figure}
-
-However, significantly lower odds of vaccination were identified in the
-corrected model for those with a household income under CAD 25,000
-(OR=0.37, CI={[}0.27,0.51{]}) and those with an income between CAD
-25,000 and 59,999 (OR=0.58, CI={[}0.42,0.81{]}). Additionally,
-individuals who identified as Arab/Middle Eastern, Black, Latin
-American, of mixed background, or that belonged to other racial groups
-(a category that included Southeast Asian, Filipino, West Asian, and
-minorities not identified elsewhere), had significantly lower odds of
-vaccination than those in the White/Caucasian group (ORs and CIs=0.28
-{[}0.16,0.51{]}, 0.27 {[}0.16,0.45{]}, 0.40 {[}0.21,0.76{]}, 0.53
-{[}0.30,0.92{]}, 0.23 {[}0.15,0.36{]}). Additionally, individuals that
-reported living in the Central and West Health Regions had higher odds
-of vaccination than those in the Health Region of Toronto (ORs and
-CIs=1.61 {[}1.10,2.34{]}, and 1.59 {[}1.16,2.19{]}, respectively).
-
-Interestingly, individuals in underrepresented groups with a household
-income below CAD 25,000 had higher odds of vaccination (when compared to
-those with a household income above CAD 60,000). This held true in the
-case of Arab/Middle Eastern individuals (OR=34, CI={[}1.70,6.79{]}),
-Black individuals (OR=3.81, CI={[}2.05, 7.09{]}), and those in other
-racial or ethnic groups (OR=3.19, CI={[}2.00,5.09{]}). Additionally,
-individuals with an income between CAD 25,000 and 59,999 in the
-Arab/Middle Eastern and other racial or ethnic groups also had higher
-odds of vaccination than their high-income peers (ORs and CIs=6.96
-{[}2.67,18.16{]}, and 3.5 {[}1.85,6.62{]}).
-
-Finally, the place of habitation affected the odds of vaccination for
-certain underrepresented groups, as significantly lower odds of
-vaccination were identified for the interaction between Health Region
-and race in the case of Black individuals in the Central Health Region
-(OR=0.39, CI={[}0.2,0.75{]}), Arab/Middle Eastern individuals in the
-East Health Region (OR=0.41 {[}0.17, 0.98{]}), and in the Indigenous and
-mixed groups in the West Health Region (ORs and CIs={[}0.31 {[}0.14,
-0.7{]} and 0.38 {[}0.19, 0.76{]}, respectively).
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-In this study, we hypothesized that differences in COVID-19 vaccination
-uptake were present between the Health Regions between late 2021 and
-early 2022. Our goal was to determine which socio-demographic groups
-could be impacted by these disparities in order to provide
-decision-makers with information that could be used to develop policies
-focused on reducing or eliminating these differences and ensuring that
-the Health Region model is able to fulfill its mission of improving
-health access for all Ontarians.
-
-Our results show that indeed, there were differences in vaccination odds
-across Ontario in certain socio-demographic groups. Specifically, those
-who identified as Arab/Middle Eastern, Black, Latin American, having
-mixed racial or ethnic background, or that belonged to other groups not
-explicitly included in the survey (Southeast Asian, Filipino, West
-Asian, and minority groups not identified elsewhere) had vaccination
-odds that were between a third and a half of that of individuals that
-identified as White or Caucasian (Figure~\ref{fig-models}). These
-results are consistent with previous studies that have shown lower
-vaccination rates in individuals with the same socio-demographic
-characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
-
-Lower vaccine uptake in the socio-demographic groups indicated above may
-be influenced in part, by vaccine hesitancy and refusal, which have been
-associated in underrepresented Canadian individuals with concerns on
-vaccine safety, effectiveness, and experiences of racial discrimination
-in health
-settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{41},\protect\hyperlink{ref-basta2022}{49}--\protect\hyperlink{ref-cnat2023}{51}}.
-However, it has been shown that structural barriers also play an
-important role in vaccination uptake. In the case of underrepresented
-individuals, such barriers include complex scheduling systems, language
-barriers, lack of adequate public transportation, and lack of accessible
-vaccination
-sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{52}}. In this
-regard, it is interesting to note that vaccination venues were scarce in
-low socio-economic areas that had the highest burden of COVID-19 in
-Toronto and other regions of Ontario around the time covered by the
-survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{53}},
-and that pharmacies in the Peel region (an area identified as a
-``hotspot'' with high numbers of essential workers and multigenerational
-households) could not keep up with vaccine
-demand\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. This
-suggests that the observed differences are associated with disparities
-in vaccine access that were present during the period covered by the
-survey.
-
-Interestingly, whereas overall self-reported vaccination rates were
-found to be statistically significantly lower in various
-underrepresented groups when compared to White/Caucasian individuals,
-the change in odds of vaccination within certain racial groups and
-income strata was actually positive, in contrast to the White/Caucasian
-group, where vaccination odds decreased in income brackets below CAD
-60,000 (Supplementary Figure A-5). Specifically, individuals in low
-income brackets that belonged to Arab/Middle Eastern, Black, or other
-minority groups had higher odds of vaccination that their peers with an
-income above 60,000 CAD.
-
-This result is likely reflects in part the fact that individuals in
-underrepresented groups tend to perform occupations that have been
-deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{55},\protect\hyperlink{ref-ct2021}{56}},
-which include workers in the areas of grocery stores, gas stations,
-warehouses, distribution, and manufacturing, all being occupations for
-which an income within the significant brackets identified in the
-analysis is to be expected. In Ontario, these workers had priority for
-COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{57}}; and
-there is evidence of interventions by vaccination staff in certain parts
-of the province to encourage vaccination uptake by these
-individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. These
-facts, combined with evidence of increased trends in vaccination in this
-group
-elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{58}},
-suggest that the type of occupation of individuals in underrepresented
-groups played an important role in increasing the odds of vaccination.
-
-However, the results also indicate that the place of habitation affected
-the odds of vaccination for certain underrepresented groups (interaction
-term of Health Region and Race, Figure~\ref{fig-models},B).
-Specifically, this held true in the case of individuals identifying as
-Indigenous or with mixed racial background in the West Health Region,
-Black individuals in the Central Health Region, and Arab/Middle Eastern
-individuals in the East Health Region Figure~\ref{fig-models}. For these
-individuals, vaccination odds were lower when compared to the Toronto
-Health Region (Supplementary Figure A-6). We indicate next some
-contributing factors that might help provide context to these results.
-
-First, in this case it is useful to analyze the data considering the
-LHINs in each Health Region, because most studies in the literature
-focused on Ontario use the LHINs as the base of their analyses. The West
-Health Region covers the area previously occupied by the Hamilton
-Niagara Haldimand Brant, South West, and Waterloo Wellington LHINs,
-whereas the East Health Region covers the area of the former Champlain
-and Central East LHINs. Previous research has identified health
-disparities in these (mostly rural) regions, such as unequal
-distribution of primary care providers, increased mortality, and low
-pharmacist
-availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.
-
-Furthermore, there is an ongoing challenge for the health system of the
-province with regard to personalized healthcare for marginalized
-individuals. For example, the West Health Region has only two Aboriginal
-Health Access Centres (community-led primary healthcare organizations
-focused on First Nations, Métis, and Inuit communities) to provide care
-to an estimated 100,000 Indigenous individuals living in the
-area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}. Lack of
-access to personalized healthcare affects individuals that may mistrust
-the traditional healthcare system due to systemic racism or oppression,
-which is known to be the case for Indigenous and Black individuals in
-Canada, as these rationales have been associated with observed lower
-vaccination rates among these
-groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{63},\protect\hyperlink{ref-eissa2021}{64}}.
-Taken together, this suggests that healthcare disparities specific to
-these underrepresented groups in certain parts of the province impacted
-vaccination uptake, and highlights the need of investments in the Health
-Regions focused on resources, infrastructure, and specially personnel
-that can deliver personalized care to marginalized communities, as it
-has been shown that such efforts have improved trust in vaccination in
-underrepresented groups
-elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{65}}.
-
-There are some limitations to the present study. First, the data
-collection design, which allowed respondents to withdraw from the survey
-at any point, and that deployed the questions in a random manner
-resulted in an elevated number of missing observations without a
-definite pattern and complicated the implementation of sensitivity
-analyses. Therefore, we focused on entries that had complete answers,
-and corrected the data using population-wide information from the
-Census. More granular corrections would be needed to obtain more
-accurate estimates. For example, our analysis identified higher odds of
-vaccination in the Central and West Health Regions, but in this case
-these differences are likely to be driven by the proportion of
-White/Caucasian individuals, who had higher vaccination rates than other
-racial groups. Correcting for each racial/ethnic group in each Health
-Region can provide a more accurate estimation of region-wide vaccination
-rates but unfortunately, at the moment this correction cannot be
-implemented as such stratification is has not been implemented in the
-Census.
-
-Additionally, our analysis did not consider the North West and North
-East Health Regions, due to the low number of entries from these areas
-in the survey (Figure~\ref{fig-map}). Low representation is expected as
-these regions as they only account for 5\% of the total population of
-Ontario. However, these areas have the highest proportion of Indigenous
-inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
-In the context of personalized care, there is a need to collect data
-that focuses on these Health Regions where additional health disparities
-might be present and possibly understudied.
-
-The results in this study are based on self-reported data, where bias
-might be present. However, in the context of COVID-19, it has been shown
-that good agreement exists between self-reported and documented
-vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{66}}, we
-believe that our data was able to provide a valid sample of vaccination
-uptake in the province. This is supported by the statistically
-significant higher vaccination odds that were identified for January of
-2022 in the model, which are consistent with province-wide trends
-reported by Public Health Ontario (which show a 4\% increase between
-early December and January, in contrast to a 2.5\% increase between
-October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{67}});
-however, the short time window constitutes essentially a ``snapshot''
-view of the evolution of the disease, and additional data would be
-needed to obtain estimates per racial/ethnic group over time across all
-Health Regions that can help inform the existence of other health
-disparities.
-
-Nonetheless, the results presented here can serve as a starting point to
-motivate the collection of robust longitudinal data that can be used to
-quantify geographical and temporal differences within vulnerable
-segments of the population, and that can be used to inform the
-development of adequate public health policies within the province of
-Ontario or across other provinces in Canada that aim to minimize
-disparities in health access.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-The implementation of the Health Regions in Ontario aimed at reducing
-the bureaucratic complexity and health disparities identified under the
-LHIN model. However, there are currently multiple challenges that need
-to be addressed to ensure that the new model can improve healthcare for
-the inhabitants of the province. First, the fact that each Health Region
-now covers a large geographical area that was served by multiple LHINs
-in the past creates a complex socio-demographic landscape that is
-different in each case due the different levels of rurality and
-representation of equity-deserving groups that are now within each
-Health Region. So far, the evidence collected during the COVID-19
-pandemic indicates that differences in vaccination uptake are associated
-to a lack of infrastructure and resources that can adequately support
-personalized care to marginalized individuals. In the near future,
-health decision-makers will need to consider the implementation of
-policies that are focused on addressing this problematic.
-
-Moreover, the recent nature in the adaption of the Health Region poses a
-challenge for researchers in the acquisition of data and information
-that can be used to analyze the performance of the new model. From one
-side, the Health Regions have not been incorporated as part of Census
-data (LHINs were considered before in the Census), and this impact the
-amount and level of detail of available information. On the other hand,
-evaluations of the Health Region model are at the moment not considered
-in the Annual Reports of the Auditor General of Ontario, which was
-critical in the past to identify limitations in the LHIN model.
-Currently, the only demographic information available for each Health
-Region is provided by Ontario Health (the agency that administers the
-Health Regions) but this information only provides general estimates
-that do not allow for detailed analyses of performance indicators (such
-as hospitalizations, readmissions, and trends in chronic disease
-incidence) between the Regions. Therefore, there is a pressing need for
-open data that can be used by researchers and decision-makers to examine
-the performance of the Health Region model.
-
-The Health Region model will only by successful if it ensures that
-healthcare improves across all segments of the population of Ontario,
-particularly in the event of a future public health emergency or
-pandemic where so far, based on the experience of the COVID-19 pandemic,
-underrepresented individuals have been disproportionately affected.
-
-\hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
-
-This work was supported by the Fonds de recherche du Québec Scholar
-Program (J1 in Artificial Intelligence and Digital Health, BN), the
-Natural Sciences and Engineering Research Council of Canada through the
-Discovery Grant Program (RGPID-560523-2020, BN), the Mathematics for
-Public Health (MfPH) Emerging Infectious Diseases Modelling Initiative
-(BN, AM), and the OMNI Emerging Infectious Disease Modelling Initiative
-(BN).
-
-\hypertarget{conflicts-of-interest}{%
-\section{Conflicts of Interest}\label{conflicts-of-interest}}
-
-The authors declare no conflict of interest.
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-WHO-Covid}{}}%
-\CSLLeftMargin{1. }%
-\CSLRightInline{{World Health Organization Coronavirus (COVID-19)
-Dashboard}. Accessed May 11, 2023. \url{https://covid19.who.int/}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-rigby2023}{}}%
-\CSLLeftMargin{2. }%
-\CSLRightInline{Rigby J, Satija B. {WHO} declares end to COVID global
-health emergency. \emph{Reuters}. Published online May 8, 2023. Accessed
-May 11, 2022.
-\url{https://www.reuters.com/business/healthcare-pharmaceuticals/covid-is-no-longer-global-health-emergency-who-2023-05-05/}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-un2023}{}}%
-\CSLLeftMargin{3. }%
-\CSLRightInline{Nations U. {WHO} chief declares end to COVID-19 as a
-global health emergency. \emph{UN News}. Published online May 5, 2023.
-Accessed May 11, 2022.
-\url{https://news.un.org/en/story/2023/05/1136367}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mackey2021}{}}%
-\CSLLeftMargin{4. }%
-\CSLRightInline{Mackey K, Ayers CK, Kondo KK, et al. Racial and ethnic
-disparities in {COVID}-19{\textendash}related infections,
-hospitalizations, and deaths. \emph{Annals of Internal Medicine}.
-2021;174(3):362-373.
-doi:\href{https://doi.org/10.7326/m20-6306}{10.7326/m20-6306}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-thelancet2021}{}}%
-\CSLLeftMargin{5. }%
-\CSLRightInline{Microbe TL. {COVID}-19 vaccines: The pandemic will not
-end overnight. \emph{The Lancet Microbe}. 2021;2(1):e1.
-doi:\href{https://doi.org/10.1016/s2666-5247(20)30226-3}{10.1016/s2666-5247(20)30226-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-davis2022}{}}%
-\CSLLeftMargin{6. }%
-\CSLRightInline{Davis CJ, Golding M, McKay R. Efficacy information
-influences intention to take COVID-19 vaccine. \emph{British Journal of
-Health Psychology}. 2022;27(2):300-319.
-doi:\url{https://doi.org/10.1111/bjhp.12546}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogoch2022}{}}%
-\CSLLeftMargin{7. }%
-\CSLRightInline{Bogoch II, Halani S. {COVID}-19 vaccines: A geographic,
-social and policy view of vaccination efforts in ontario, canada.
-\emph{Cambridge Journal of Regions, Economy and Society}. Published
-online November 2022.
-doi:\href{https://doi.org/10.1093/cjres/rsac043}{10.1093/cjres/rsac043}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tanne2020}{}}%
-\CSLLeftMargin{8. }%
-\CSLRightInline{Tanne JH. Covid-19: {FDA} panel votes to authorise
-pfizer {BioNTech} vaccine. \emph{{BMJ}}. Published online December
-2020:m4799.
-doi:\href{https://doi.org/10.1136/bmj.m4799}{10.1136/bmj.m4799}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
-\CSLLeftMargin{9. }%
-\CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
-and emerging issues. \emph{Human Vaccines {\&}amp\(\mathsemicolon\)
-Immunotherapeutics}. 2021;17(12):5255-5268.
-doi:\href{https://doi.org/10.1080/21645515.2021.1977057}{10.1080/21645515.2021.1977057}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-li2021}{}}%
-\CSLLeftMargin{10. }%
-\CSLRightInline{Li Q, Wang J, Tang Y, Lu H. Next-generation COVID-19
-vaccines: Opportunities for vaccine development and challenges in
-tackling COVID-19. \emph{Drug Discoveries \& Therapeutics}.
-2021;15(3):118-123.
-doi:\href{https://doi.org/10.5582/ddt.2021.0105}{10.5582/ddt.2021.0105}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
-\CSLLeftMargin{11. }%
-\CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
-determinants of {COVID}-19 vaccine hesitancy. Inbaraj LR, ed.
-\emph{{PLOS} {ONE}}. 2021;16(11):e0258462.
-doi:\href{https://doi.org/10.1371/journal.pone.0258462}{10.1371/journal.pone.0258462}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
-\CSLLeftMargin{12. }%
-\CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
-to achieve global covid-19 vaccine equity. \emph{{BMJ}}. Published
-online March 2022:e070650.
-doi:\href{https://doi.org/10.1136/bmj-2022-070650}{10.1136/bmj-2022-070650}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
-\CSLLeftMargin{13. }%
-\CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
-inequality in {COVID}-19 vaccination coverage among elderly adults in
-england: A national linked data study. \emph{{BMJ} Open}.
-2021;11(7):e053402.
-doi:\href{https://doi.org/10.1136/bmjopen-2021-053402}{10.1136/bmjopen-2021-053402}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
-\CSLLeftMargin{14. }%
-\CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
-{COVID}-19 vaccine hesitancy: Race/ethnicity, trust, and fear.
-\emph{Clinical and Translational Science}. 2021;14(6):2200-2207.
-doi:\href{https://doi.org/10.1111/cts.13077}{10.1111/cts.13077}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
-\CSLLeftMargin{15. }%
-\CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
-accepting {COVID}-19 vaccination during and after pregnancy, and for
-their babies: A multi-methods study in the {UK}. \emph{{BMC} Pregnancy
-and Childbirth}. 2022;22(1).
-doi:\href{https://doi.org/10.1186/s12884-021-04321-3}{10.1186/s12884-021-04321-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
-\CSLLeftMargin{16. }%
-\CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
-limits of medical trust in mitigating {COVID}-19 vaccine hesitancy among
-black americans. \emph{Journal of General Internal Medicine}.
-2021;36(11):3629-3631.
-doi:\href{https://doi.org/10.1007/s11606-021-06743-3}{10.1007/s11606-021-06743-3}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
-\CSLLeftMargin{17. }%
-\CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
-Sharma M, Webb FJ. {COVID}-19 vaccination hesitancy in the united
-states: A rapid national assessment. \emph{Journal of Community Health}.
-2021;46(2):270-277.
-doi:\href{https://doi.org/10.1007/s10900-020-00958-x}{10.1007/s10900-020-00958-x}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
-\CSLLeftMargin{18. }%
-\CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
-inequalities in {COVID}-19 vaccination uptake and intent: Results from
-the canadian community health survey 2021. \emph{{BMC} Public Health}.
-2022;22(1).
-doi:\href{https://doi.org/10.1186/s12889-022-14090-z}{10.1186/s12889-022-14090-z}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
-\CSLLeftMargin{19. }%
-\CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
-Fahlman M, Kallio N. {COVID}-19 vaccine hesitancy and refusal and
-associated factors in an adult population in saskatchewan, canada:
-Evidence from predictive modelling. Gesser-Edelsburg A, ed. \emph{{PLOS}
-{ONE}}. 2021;16(11):e0259513.
-doi:\href{https://doi.org/10.1371/journal.pone.0259513}{10.1371/journal.pone.0259513}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
-\CSLLeftMargin{20. }%
-\CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
-Overcoming {COVID}-19 vaccine hesitancy among ethnic minorities: A
-systematic review of {UK} studies. \emph{Vaccine}.
-2022;40(25):3413-3432.
-doi:\href{https://doi.org/10.1016/j.vaccine.2022.04.030}{10.1016/j.vaccine.2022.04.030}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
-\CSLLeftMargin{21. }%
-\CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
-roots of {COVID}-19 vaccine hesitancy among indigenous peoples in
-canada. \emph{Canadian Medical Association Journal}.
-2021;193(11):E381-E383.
-doi:\href{https://doi.org/10.1503/cmaj.210112}{10.1503/cmaj.210112}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
-\CSLLeftMargin{22. }%
-\CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
-related medical mistrust, health impacts, and potential vaccine
-hesitancy among black americans living with {HIV}. \emph{{JAIDS} Journal
-of Acquired Immune Deficiency Syndromes}. 2021;86(2):200-207.
-doi:\href{https://doi.org/10.1097/qai.0000000000002570}{10.1097/qai.0000000000002570}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
-\CSLLeftMargin{23. }%
-\CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
-hesitancy in the {UK}: The oxford coronavirus explanations, attitudes,
-and narratives survey (oceans) {II}. \emph{Psychological Medicine}.
-2020;52(14):3127-3141.
-doi:\href{https://doi.org/10.1017/s0033291720005188}{10.1017/s0033291720005188}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
-\CSLLeftMargin{24. }%
-\CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
-of {COVID}-19 vaccine acceptance in the {US}.
-\emph{{EClinicalMedicine}}. 2020;26:100495.
-doi:\href{https://doi.org/10.1016/j.eclinm.2020.100495}{10.1016/j.eclinm.2020.100495}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
-\CSLLeftMargin{25. }%
-\CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
-Changes in {COVID}-19 vaccination receipt and intention to vaccinate by
-socioeconomic characteristics and geographic area, united states,
-january 6 {\textendash} march 29, 2021. \emph{Annals of Medicine}.
-2021;53(1):1419-1428.
-doi:\href{https://doi.org/10.1080/07853890.2021.1957998}{10.1080/07853890.2021.1957998}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
-\CSLLeftMargin{26. }%
-\CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
-vaccine hesitancy in the united states. \emph{International Journal of
-Environmental Research and Public Health}. 2021;18(18):9488.
-doi:\href{https://doi.org/10.3390/ijerph18189488}{10.3390/ijerph18189488}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
-\CSLLeftMargin{27. }%
-\CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
-spatial process: {COVID}-19 fully vaccinated rates in u.s. counties.
-\emph{American Journal of Preventive Medicine}. 2022;63(6):954-961.
-doi:\href{https://doi.org/10.1016/j.amepre.2022.06.006}{10.1016/j.amepre.2022.06.006}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
-\CSLLeftMargin{28. }%
-\CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
-the spatiotemporal heterogeneity of the {COVID}-19 vaccination
-landscape. \emph{American Journal of Epidemiology}.
-2022;191(10):1792-1802.
-doi:\href{https://doi.org/10.1093/aje/kwac080}{10.1093/aje/kwac080}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
-\CSLLeftMargin{29. }%
-\CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
-social vulnerability index to assess {COVID}-19 vaccine uptake in
-louisiana. \emph{{GeoJournal}}. Published online December 2022.
-doi:\href{https://doi.org/10.1007/s10708-022-10802-5}{10.1007/s10708-022-10802-5}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
-\CSLLeftMargin{30. }%
-\CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
-future spatial patterns in {COVID}-19 booster vaccine uptake. Published
-online September 2022.
-doi:\href{https://doi.org/10.1101/2022.08.30.22279415}{10.1101/2022.08.30.22279415}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
-\CSLLeftMargin{31. }%
-\CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
-trajectories. \emph{Socius: Sociological Research for a Dynamic World}.
-2021;7:237802312110529.
-doi:\href{https://doi.org/10.1177/23780231211052946}{10.1177/23780231211052946}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
-\CSLLeftMargin{32. }%
-\CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
-Social inequalities in {COVID}-19 vaccine acceptance and uptake for
-children and adolescents in montreal, canada. \emph{Vaccine}.
-2021;39(49):7140-7145.
-doi:\href{https://doi.org/10.1016/j.vaccine.2021.10.077}{10.1016/j.vaccine.2021.10.077}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
-\CSLLeftMargin{33. }%
-\CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
-integrated care: A complex-adaptive systems perspective.
-\emph{International Journal of Integrated Care}. 2012;12(5).
-doi:\href{https://doi.org/10.5334/ijic.843}{10.5334/ijic.843}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-muratov2018}{}}%
-\CSLLeftMargin{34. }%
-\CSLRightInline{Muratov S, Lee J, Holbrook A, et al. Regional variation
-in healthcare spending and mortality among senior high-cost healthcare
-users in ontario, canada: A retrospective matched cohort study.
-\emph{{BMC} Geriatrics}. 2018;18(1).
-doi:\href{https://doi.org/10.1186/s12877-018-0952-7}{10.1186/s12877-018-0952-7}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
-\CSLLeftMargin{35. }%
-\CSLRightInline{Dong L, Sahu R, Black R. Governance in the
-transformational journey toward integrated healthcare: The case of
-ontario. \emph{Journal of Information Technology Teaching Cases}.
-Published online December 2022:204388692211473.
-doi:\href{https://doi.org/10.1177/20438869221147313}{10.1177/20438869221147313}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lysyk2015}{}}%
-\CSLLeftMargin{36. }%
-\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
-2015. In: \emph{Section 3.08: LHINs - Local Health Integration
-Networks}. Queen's Printer for Ontario; 2015. Accessed May 12, 2023.
-\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en15/3.08en15.pdf}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lysyk2016}{}}%
-\CSLLeftMargin{37. }%
-\CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
-2016. In: \emph{Section 3.03: Electronic Health Records' Implementation
-Status}. Queen's Printer for Ontario; 2016. Accessed May 12, 2023.
-\url{https://www.auditor.on.ca/en/content/annualreports/arreports/en16/v1_303en16.pdf}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sethuram2023}{}}%
-\CSLLeftMargin{38. }%
-\CSLRightInline{Sethuram C, McCutcheon T, Liddy C. An environmental scan
-of ontario health teams: A descriptive study. \emph{{BMC} Health
-Services Research}. 2023;23(1).
-doi:\href{https://doi.org/10.1186/s12913-023-09102-6}{10.1186/s12913-023-09102-6}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{39. }%
-\CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
-Corlin L. Disparities in national and state estimates of {COVID}-19
-vaccination receipt and intent to vaccinate by race/ethnicity, income,
-and age group among adults~\(\geq\)~18~years, united states.
-\emph{Vaccine}. 2022;40(1):107-113.
-doi:\href{https://doi.org/10.1016/j.vaccine.2021.11.040}{10.1016/j.vaccine.2021.11.040}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{40. }%
-\CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
-Zikmund-Fisher BJ. Vaccine hesitancy and rejection of a vaccine for the
-novel coronavirus in the united states. \emph{Frontiers in Immunology}.
-2021;12.
-doi:\href{https://doi.org/10.3389/fimmu.2021.558270}{10.3389/fimmu.2021.558270}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
-and factors related to {COVID}-19 vaccine hesitancy and unwillingness in
-canada: A systematic review and meta-analysis. \emph{Journal of Medical
-Virology}. 2022;95(1).
-doi:\href{https://doi.org/10.1002/jmv.28156}{10.1002/jmv.28156}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{42. }%
-\CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
-a sampled frequency table when the expected marginal totals are known.
-\emph{The Annals of Mathematical Statistics}. 1940;11(4):427-444.
-doi:\href{https://doi.org/10.1214/aoms/1177731829}{10.1214/aoms/1177731829}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{43. }%
-\CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
-2011.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{44. }%
-\CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
-{tidyverse}. \emph{Journal of Open Source Software}. 2019;4(43):1686.
-doi:\href{https://doi.org/10.21105/joss.01686}{10.21105/joss.01686}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{45. }%
-\CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
-Markdown Publishing System}.; 2022.
-\url{https://CRAN.R-project.org/package=quarto}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{46. }%
-\CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
-in {R}. \emph{Journal of Statistical Software}. 2022;103(1):1-23.
-doi:\href{https://doi.org/10.18637/jss.v103.i01}{10.18637/jss.v103.i01}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{47. }%
-\CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
-Reproducible summary tables with the gtsummary package. \emph{{The R
-Journal}}. 2021;13:570-580.
-doi:\href{https://doi.org/10.32614/RJ-2021-053}{10.32614/RJ-2021-053}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{48. }%
-\CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
-{COVID}-19 vaccine uptake in southeastern ontario, canada: Monitoring
-and addressing health inequities. \emph{Journal of Public Health
-Management and Practice}. 2022;28(6):615-623.
-doi:\href{https://doi.org/10.1097/phh.0000000000001565}{10.1097/phh.0000000000001565}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-basta2022}{}}%
-\CSLLeftMargin{49. }%
-\CSLRightInline{Basta NE, Sohel N, Sulis G, et al. Factors associated
-with willingness to receive a {COVID}-19 vaccine among 23, 819 adults
-aged 50 years or older: An analysis of the canadian longitudinal study
-on aging. \emph{American Journal of Epidemiology}. 2022;191(6):987-998.
-doi:\href{https://doi.org/10.1093/aje/kwac029}{10.1093/aje/kwac029}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
-\CSLLeftMargin{50. }%
-\CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
-review on vaccine hesitancy in black communities in canada: Critical
-issues and research failures. \emph{Vaccines}. 2022;10(11):1937.
-doi:\href{https://doi.org/10.3390/vaccines10111937}{10.3390/vaccines10111937}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cnat2023}{}}%
-\CSLLeftMargin{51. }%
-\CSLRightInline{Cénat JM, Farahi SMMM, Bakombo SM, et al. Vaccine
-mistrust among black individuals in canada: The major role of health
-literacy, conspiracy theories, and racial discrimination in the
-healthcare system. \emph{Journal of Medical Virology}. 2023;95(4).
-doi:\href{https://doi.org/10.1002/jmv.28738}{10.1002/jmv.28738}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{52. }%
-\CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
-Structural barriers and racial and ethnic inequities in {COVID}-19
-vaccination. \emph{International Journal of Environmental Research and
-Public Health}. 2021;18(18):9904.
-doi:\href{https://doi.org/10.3390/ijerph18189904}{10.3390/ijerph18189904}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-iveniuk2021}{}}%
-\CSLLeftMargin{53. }%
-\CSLRightInline{Iveniuk J, Leon S. \emph{Uneven Recovery: Measuring
-Covid-19 Vaccine Equity in Ontario}. Wellesley Institute; 2021. Accessed
-May 12, 2023.
-\url{https://www.wellesleyinstitute.com/wp-content/uploads/2021/04/An-uneven-recovery-Measuring-COVID-19-vaccine-equity-in-Ontario.pdf}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gill2022}{}}%
-\CSLLeftMargin{54. }%
-\CSLRightInline{Gill M, Datta D, Gregory P, Austin Z. {COVID}-19
-vaccination in high-risk communities: Case study of brampton, ontario.
-\emph{Canadian Pharmacists Journal / Revue des Pharmaciens du Canada}.
-2022;155(6):345-351.
-doi:\href{https://doi.org/10.1177/17151635221123042}{10.1177/17151635221123042}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{55. }%
-\CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
-and other infection exposure according to race and ethnicity.
-\emph{American Journal of Industrial Medicine}. 2020;63(9):817-820.
-doi:\href{https://doi.org/10.1002/ajim.23145}{10.1002/ajim.23145}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{56. }%
-\CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
-review of {COVID}-19 and vulnerable workers: Intersecting occupational
-and public health issues. \emph{American Journal of Industrial
-Medicine}. 2021;64(7):551-566.
-doi:\href{https://doi.org/10.1002/ajim.23256}{10.1002/ajim.23256}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{57. }%
-\CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
-Strategy for Ontario {COVID}-19 Hotspots and Essential Workers}. Ontario
-{COVID}-19 Science Advisory Table; 2021.
-doi:\href{https://doi.org/10.47326/ocsat.2021.02.26.1.0}{10.47326/ocsat.2021.02.26.1.0}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nguyen2021b}{}}%
-\CSLLeftMargin{58. }%
-\CSLRightInline{Nguyen KH, Yankey D, Coy KC, et al. {COVID}-19
-vaccination coverage, intent, knowledge, attitudes, and beliefs among
-essential workers, united states. \emph{Emerging Infectious Diseases}.
-2021;27(11):2908-2913.
-doi:\href{https://doi.org/10.3201/eid2711.211557}{10.3201/eid2711.211557}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-shah2019}{}}%
-\CSLLeftMargin{59. }%
-\CSLRightInline{Shah TI, Clark AF, Seabrook JA, Sibbald S, Gilliland JA.
-Geographic accessibility to primary care providers: Comparing rural and
-urban areas in southwestern ontario. \emph{The Canadian Geographer / Le
-G{é}ographe canadien}. 2019;64(1):65-78.
-doi:\href{https://doi.org/10.1111/cag.12557}{10.1111/cag.12557}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-crighton2015}{}}%
-\CSLLeftMargin{60. }%
-\CSLRightInline{Crighton EJ, Ragetlie R, Luo J, To T, Gershon A. A
-spatial analysis of {COPD} prevalence, incidence, mortality and health
-service use in ontario. \emph{Health Rep}. 2015;26(3):10-18.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-timony2022}{}}%
-\CSLLeftMargin{61. }%
-\CSLRightInline{Timony P, Houle SKD, Gauthier A, Waite NM. Geographic
-distribution of ontario pharmacists: A focus on rural and northern
-communities. \emph{Canadian Pharmacists Journal / Revue des Pharmaciens
-du Canada}. 2022;155(5):267-276.
-doi:\href{https://doi.org/10.1177/17151635221115411}{10.1177/17151635221115411}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{62. }%
-\CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
-Health;
-\url{https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf};
-2022.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
-\CSLLeftMargin{63. }%
-\CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
-{SARS}-{COV}-2 vaccine uptake and {COVID}-19 impacts among first
-nations, inuit and m{é}tis peoples living in toronto and london,
-ontario. \emph{Canadian Medical Association Journal}.
-2022;194(29):E1018-E1026.
-doi:\href{https://doi.org/10.1503/cmaj.212147}{10.1503/cmaj.212147}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
-\CSLLeftMargin{64. }%
-\CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
-Increasing {SARS}-{CoV}-2 vaccination rates among black people in
-canada. \emph{Canadian Medical Association Journal}.
-2021;193(31):E1220-E1221.
-doi:\href{https://doi.org/10.1503/cmaj.210949}{10.1503/cmaj.210949}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
-\CSLLeftMargin{65. }%
-\CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
-vaccination program. \emph{{JAMA}}. 2020;323(24):2458.
-doi:\href{https://doi.org/10.1001/jama.2020.8711}{10.1001/jama.2020.8711}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{66. }%
-\CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
-vaccination status by self-report versus source documentation: Impact on
-measuring {COVID}-19 vaccine effectiveness. \emph{Influenza and Other
-Respiratory Viruses}. 2022;16(6):1101-1111.
-doi:\href{https://doi.org/10.1111/irv.13023}{10.1111/irv.13023}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{67. }%
-\CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
-2023.
-\url{https://www.publichealthontario.ca/en/data-and-analysis/infectious-disease/covid-19-data-surveillance/covid-19-data-tool?tab=vaccine}}
-
-\end{CSLReferences}
-
-\end{document}
+The Health Region model will only by successful if it ensures that healthcare improves across all segments of the population of Ontario, particularly in the event of a future public health emergency or pandemic where so far, based on the experience of the COVID-19 pandemic, underrepresented individuals have been disproportionately affected.

### 2023-06-27 manuscript_v6/main.qmd
v6 after SW's comments

--- 2023-05-26 manuscript_v5/main.qmd
@@
 format: 
+  docx:
+    reference-doc: custom-reference-doc.docx
   pdf:
+    geometry:
+      - top=20mm
+      - left=20mm
+      - right=20mm
+      - bottom=20mm
     keep-tex: true
@@
 In this study, we examined the relationship between vaccination status (having received the first dose of a COVID-19 vaccine), and different socio-economic and geographical factors.
-Our results show that between October of 2021 and January of 2022, individuals from underrepresented communities were three times less likely to be vaccinated than White/Caucasian individuals across the province of Ontario in Canada, and that in some cases, within these groups, individuals in low income brackets had significantly higher odds of vaccination when compared to their peers in high income brackets.
+Our results show that between October of 2021 and January of 2022, individuals from underrepresented communities were three times less likely to be vaccinated than White/Caucasian individuals across the province of Ontario in Canada, and that in some cases, within these groups, individuals in low-income brackets had significantly higher odds of vaccination when compared to their peers in high income brackets.
 Finally, we identified significantly lower odds of vaccination in the Central, East and West Health Regions of Ontario within certain underrepresented groups.

### 2023-06-27 manuscript_v6/main.tex
v6 after SW's comments

--- 2023-06-09 manuscript_v5/main.tex
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, boxrule=0pt, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 individuals across the province of Ontario in Canada, and that in some
-cases, within these groups, individuals in low income brackets had
+cases, within these groups, individuals in low-income brackets had
 significantly higher odds of vaccination when compared to their peers in

### 2023-06-28 manuscript_v6/01-background.qmd
updating background

@@
 As of May of 2023, there have been 765 million confirmed cases of COVID-19 around the world, including 6.8 million deaths [@WHO-Covid].
-Although this disease is no longer categorized as a global health emergency by the World Health Organization (WHO) [@rigby2023], there is ongoing concern due to continued transmission, surges in cases and deaths due to new variants [@un2023], and weaknesses in health systems around the world that could be exploited by a novel virus or another public health emergency in the future [@mackey2021].
+Although this disease is no longer categorized as a global health emergency by the World Health Organization (WHO) [@rigby2023], there is ongoing concern due to continued transmission, the possibility of surges in cases and deaths due to new virus variants [@un2023], and ongoing issues in health systems around the world that could be exploited by a novel virus or another public health emergency in the future [@mackey2021].
 
-In particular, a major weakness that has received attention during the pandemic has been related to vaccination.
-The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic as vaccination campaigns in certain parts of the world started as early as the end of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020].
-However, although it has been estimated that COVID-19 vaccines have been able to prevent millions of deaths worldwide [@watson2022], their original goal, which was to bring the pandemic to a rapid end, was not accomplished.
-This contrast with previous vaccination experiences, such as in the case of smallpox, where the implementation of a global vaccination program was crucial to control this disease[@kayser2021].
-
-The failure of vaccines to stop the pandemic is a complex issue resulting from a combination of factors, among which are inadequate public health measures and the emergence of new variants [@li2021], inequality in vaccine access between high-income and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population[@willis2021;@skirrow2022].
+In particular, the pandemic showed multiple challenges with regard to vaccination.
+The rapid development of vaccines against COVID-19 initially brought the hope of a rapid end to the pandemic as vaccination campaigns in certain parts of the world started as early as December of 2020 [@thelancet2021;@davis2022;@bogoch2022;@tanne2020].
+Although it has been estimated that so far COVID-19 vaccines have been able to prevent millions of deaths worldwide [@watson2022], their implementation has faced multiple challenges across the world.
+These challenges included the rapid emergence of new virus variants [@li2021], the waning of vaccine protection [@dolgin2021], inequality in vaccine access between high-income and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population [@willis2021;@skirrow2022].
 In particular, it is well established that lower vaccination uptake has been observed in individuals within certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
 
-Reasons given for this inequality in vaccination uptake have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the influence of misinformation and disinformation [@bogart2021;@mosby2021;@freeman2020].
+<!-- pharmaceutical interventions were ultimately not determinant to bring the pandemic to a rapid end.  -->
+
+<!-- This contrast with previous vaccination experiences, such as in the case of smallpox, where the implementation of a global vaccination program was crucial to control this disease[@kayser2021].
+-->
+
+<!-- The failure of vaccines to stop the pandemic is a complex issue resulting from a combination of factors, among which are inadequate public health measures and the emergence of new variants [@li2021], inequality in vaccine access between high-income and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population[@willis2021;@skirrow2022].
+In particular, it is well established that lower vaccination uptake has been observed in individuals within certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
+-->
+
+Reasons given for this inequality in vaccination uptake have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the existence of health misinformation and disinformation [@bogart2021;@mosby2021;@freeman2020].
 However, it is important also to consider that vaccination uptake can be influenced by geographical (spatial) factors.
@@
 
-However, such analyses have been carried out mostly in territories outside of Canada, where available studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022].
-Therefore, there is a need for studies that explore spatial differences in vaccination within the Canadian territory and that consequently, can help identify disparities that need to be addressed within specific areas in each province.
+However, the majority of such analyses have been carried out outside of Canada, and existing studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022].
+**Canada’s pandemic response goals, were to reduce morbidity/mortality and to reduce societal disruption from COVID-19**.
+
+- How does this integrate with vaccination? We need to study vaccination because 
+-
+
+- -
+
+and that consequently, can help identify disparities that need to be addressed within specific areas in each province.
 
@@
 Between 2007 and 2019, Ontario managed healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integration Networks (LHINs), which aimed to provide an integrated health system for the province.
-However, this approach was complex, bureaucratic, and resulted in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators such as wait times and hospital readmissions, fragmented electronic health systems, the decline of performance indicators, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016].
-Therefore, with the intent of better organizing and delivering care in late 2019 the provincial government eliminated the LHINs and incorporated the areas covered by them into six larger Health Regions (North East, North West, Central, Toronto, West, and East) [@dong2022].
+However, this approach was complex, bureaucratic, and resulted in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators such as wait times and hospital readmissions, fragmented electronic health systems, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016].
+Therefore, with the intent of better organizing and delivering care in late 2019 the provincial government eliminated the LHINs and incorporated the areas covered by them into six larger Health Regions (North East, North West, Central, Toronto, West, and East), which are managed by the new government agency Ontario Health (OH) [@dong2022].
 
-On the other hand, public health is managed in the province by Public Health Ontario, which was established in late 2007 and is currently composed of 34 Public Health Units (PHUs).
-Importantly, during the COVID-19 pandemic, these PHUS led the distribution of the vaccine within the areas they covered [@rotenberg2021]
+On the other hand, public health in Ontario is administered by Public Health Ontario (PHO), a government agency established in late 2007 and currently composed of 34 Public Health Units (PHUs) that cover the entire geography of the province.
+Although these PHUS were commissioned with leading the distribution of the COVID-19 vaccine within the areas they covered [@rotenberg2021], there was significant overlap between PHO and OH in the actual vaccine rollout, and in many instances vaccines were administered in vaccination clinics ran by members of OH working alone or in conjunction with members of PHO [@ashcroft2023].
 

### 2023-06-28 manuscript_v6/main.tex
updating background

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, boxrule=0pt, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 (WHO)\textsuperscript{\protect\hyperlink{ref-rigby2023}{2}}, there is
-ongoing concern due to continued transmission, surges in cases and
-deaths due to new
-variants\textsuperscript{\protect\hyperlink{ref-un2023}{3}}, and
-weaknesses in health systems around the world that could be exploited by
-a novel virus or another public health emergency in the
+ongoing concern due to continued transmission, the possibility of surges
+in cases and deaths due to new virus
+variants\textsuperscript{\protect\hyperlink{ref-un2023}{3}}, and ongoing
+issues in health systems around the world that could be exploited by a
+novel virus or another public health emergency in the
 future\textsuperscript{\protect\hyperlink{ref-mackey2021}{4}}.
 
-In particular, a major weakness that has received attention during the
-pandemic has been related to inequalities in vaccine uptake. The rapid
-development of vaccines against COVID-19 initially brought the hope of a
-rapid end to the pandemic due to the start of vaccination campaigns in
-certain parts of the world toward the end of
-2020\textsuperscript{\protect\hyperlink{ref-thelancet2021}{5}--\protect\hyperlink{ref-tanne2020}{8}})
-but inequalities in vaccine uptake made these pharmaceutical
-interventions ultimately unable to replicate the success of the smallpox
-vaccination program, which was crucial to control this
-disease\textsuperscript{\protect\hyperlink{ref-kayser2021}{9}}.
-
-This problematic is a multifaceted issue resulting from a combination of
-factors, among which are failed public health
-measures\textsuperscript{\protect\hyperlink{ref-li2021}{10}}, inequality
-in vaccine access between high- and low-income
-countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{11},\protect\hyperlink{ref-tamey2022}{12}},
-and vaccine
-hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{13}}.
-Furthermore, it is well established that this issue has affected in
-particular individuals in certain underrepresented groups (e.g., Black,
-Asian, or Indigenous) as well as individuals with socio-economic
-disadvantages\textsuperscript{\protect\hyperlink{ref-willis2021}{14}--\protect\hyperlink{ref-hussain2022}{20}}.
-
-Reasons given for this inequality have included medical mistrust due to
-systemic medical
-racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{16},\protect\hyperlink{ref-mosby2021}{21}},
+In particular, the pandemic showed multiple challenges with regard to
+vaccination. The rapid development of vaccines against COVID-19
+initially brought the hope of a rapid end to the pandemic as vaccination
+campaigns in certain parts of the world started as early as December of
+2020\textsuperscript{\protect\hyperlink{ref-thelancet2021}{5}--\protect\hyperlink{ref-tanne2020}{8}}.
+Although it has been estimated that so far COVID-19 vaccines have been
+able to prevent millions of deaths
+worldwide\textsuperscript{\protect\hyperlink{ref-watson2022}{9}}, their
+implementation has faced multiple challenges across the world. These
+challenges included the rapid emergence of new virus
+variants\textsuperscript{\protect\hyperlink{ref-li2021}{10}}, the waning
+of vaccine
+protection\textsuperscript{\protect\hyperlink{ref-dolgin2021}{11}},
+inequality in vaccine access between high-income and low-income
+countries\textsuperscript{\protect\hyperlink{ref-gerretsen2021}{12},\protect\hyperlink{ref-tamey2022}{13}},
+vaccine
+hesitancy\textsuperscript{\protect\hyperlink{ref-nafilyan2021}{14}}, and
+differences in vaccine uptake across the
+population\textsuperscript{\protect\hyperlink{ref-willis2021}{15},\protect\hyperlink{ref-skirrow2022}{16}}.
+In particular, it is well established that lower vaccination uptake has
+been observed in individuals within certain underrepresented groups
+(e.g., Black, Asian, or Indigenous) as well as individuals with
+socio-economic
+disadvantages\textsuperscript{\protect\hyperlink{ref-willis2021}{15}--\protect\hyperlink{ref-hussain2022}{21}}.
+
+Reasons given for this inequality in vaccination uptake have included
+medical mistrust due to systemic medical
+racism\textsuperscript{\protect\hyperlink{ref-stoler2021}{17},\protect\hyperlink{ref-mosby2021}{22}},
 mistrust in
-vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{14}}, and
-the influence of conspiracy
-theories\textsuperscript{\protect\hyperlink{ref-mosby2021}{21}--\protect\hyperlink{ref-freeman2020}{23}}.
+vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{15}}, and
+the influence of misinformation and
+disinformation\textsuperscript{\protect\hyperlink{ref-mosby2021}{22}--\protect\hyperlink{ref-freeman2020}{24}}.
 However, it is important also to consider that vaccination uptake can be
@@
 varied regional attitudes towards
-vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{24}},
-spatial differences in vaccine access and supply, vaccination location
-availability, and lack of prioritization of areas where vulnerable
-groups
-reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-nguyen2021}{25}}.
-Other studies have also shown heterogeneity in vaccine uptake within
-small governmental administrative units such as
-counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{26}--\protect\hyperlink{ref-bhuiyan2022}{29}},
-and that accounting for geographical differences in vaccination can help
-predict patterns of booster
-uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{30}}.
+vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{25}}, lack
+of vaccine access in areas where underrepresented groups
+reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as well
+as an absence of prioritization of areas inhabited by vulnerable
+groups\textsuperscript{\protect\hyperlink{ref-nguyen2021}{26}}. Other
+studies have also shown heterogeneity in vaccine uptake within small
+governmental administrative units such as
+counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{27}--\protect\hyperlink{ref-bhuiyan2022}{30}},
+indicating that accounting for geographical differences in vaccination
+can help predict patterns of booster
+uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{31}}.
 
@@
 cities (such as
-Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{31}}, or
-Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{32}}), or
+Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{32}}, or
+Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{33}}), or
 have explored differences at a province-wide
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{18}}. Therefore,
+level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. Therefore,
 there is a need for studies that explore spatial differences in
@@
 certain performance indicators such as wait times and hospital
-readmissions, fragmented electronic health systems, the decline of
-performance indicators, and inequities in health services
-access\textsuperscript{\protect\hyperlink{ref-tsasis2012}{33}--\protect\hyperlink{ref-lysyk2016}{37}}.
+readmissions, fragmented electronic health systems, and inequities in
+health services
+access\textsuperscript{\protect\hyperlink{ref-tsasis2012}{34}--\protect\hyperlink{ref-lysyk2016}{38}}.
 Therefore, with the intent of better organizing and delivering care in
@@
 incorporated the areas covered by them into six larger Health Regions
-(North East, North West, Central, Toronto, West, and
-East)\textsuperscript{\protect\hyperlink{ref-dong2022}{35}}.
+(North East, North West, Central, Toronto, West, and East), which are
+managed by the new government agency Ontario Health
+(OH)\textsuperscript{\protect\hyperlink{ref-dong2022}{36}}.
+
+On the other hand, public health in Ontario is administered by Public
+Health Ontario (PHO), a government agency established in late 2007 and
+currently composed of 34 Public Health Units (PHUs) that cover the
+entire geography of the province. Although these PHUS were commissioned
+with leading the distribution of the COVID-19 vaccine within the areas
+they covered\textsuperscript{\protect\hyperlink{ref-rotenberg2021}{39}},
+there was significant overlap between PHO and OH in the actual vaccine
+rollout, and in many instances vaccines were administered in vaccination
+clinics ran by members of OH working alone or in conjunction with
+members of
+PHO\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{40}}.
+
+\begin{itemize}
+\tightlist
+\item
+  add about intersection of Health Regions and PHUs. How vaccination
+  presented a challenge as vaccine was ultimately administered in
+  facilites managed by OHT?
+\end{itemize}
+
+https://www.mcmasterforum.org/docs/default-source/product-documents/rapid-responses/examining-the-intersections-between-ontario-health-teams-and-public-health.pdf?sfvrsn=69caa8c9\_11
 
@@
 highlighted disparities in the level of activity of each Health
-Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{38}}.
+Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{41}}.
 Therefore, analyzing differences in vaccination uptake within the Health
@@
 of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{39}--\protect\hyperlink{ref-cnat2022a}{41}}.
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{42}--\protect\hyperlink{ref-cnat2022a}{44}}.
 Because we identified differences in representativity between the survey
@@
 proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{42}}
+(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{45}}
 to correct the data using data from the Census and Health Region
@@
 Appendix. All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{43}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{44}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{45}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{46}},
+\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{46}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{47}},
+\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{48}},
+\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{49}},
 and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{47}}.
+\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{50}}.
 
@@
 vaccination rates in individuals with the same socio-demographic
-characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{18}--\protect\hyperlink{ref-hussain2022}{20},\protect\hyperlink{ref-carter2022}{48}}.
+characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{51}}.
 
@@
 in health
-settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{41},\protect\hyperlink{ref-basta2022}{49}--\protect\hyperlink{ref-cnat2023}{51}}.
+settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{44},\protect\hyperlink{ref-basta2022}{52}--\protect\hyperlink{ref-cnat2023}{54}}.
 However, it has been shown that structural barriers also play an
@@
 vaccination
-sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{52}}. In this
+sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{55}}. In this
 regard, it is interesting to note that vaccination venues were scarce in
@@
 Toronto and other regions of Ontario around the time covered by the
-survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{53}},
+survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{56}},
 and that pharmacies in the Peel region (an area identified as a
@@
 households) could not keep up with vaccine
-demand\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. This
+demand\textsuperscript{\protect\hyperlink{ref-gill2022}{57}}. This
 suggests that the observed differences are associated with disparities
@@
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{55},\protect\hyperlink{ref-ct2021}{56}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{58},\protect\hyperlink{ref-ct2021}{59}},
 which include workers in the areas of grocery stores, gas stations,
@@
 COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{57}}; and
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{60}}; and
 there is evidence of interventions by vaccination staff in certain parts
 of the province to encourage vaccination uptake by these
-individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{54}}. These
+individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{57}}. These
 facts, combined with evidence of increased trends in vaccination in this
 group
-elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{58}},
+elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{61}},
 suggest that the type of occupation of individuals in underrepresented
@@
 pharmacist
-availability\textsuperscript{\protect\hyperlink{ref-shah2019}{59}--\protect\hyperlink{ref-timony2022}{61}}.
+availability\textsuperscript{\protect\hyperlink{ref-shah2019}{62}--\protect\hyperlink{ref-timony2022}{64}}.
 
@@
 to an estimated 100,000 Indigenous individuals living in the
-area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}. Lack of
+area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{65}}. Lack of
 access to personalized healthcare affects individuals that may mistrust
@@
 vaccination rates among these
-groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{63},\protect\hyperlink{ref-eissa2021}{64}}.
+groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{66},\protect\hyperlink{ref-eissa2021}{67}}.
 Taken together, this suggests that healthcare disparities specific to
@@
 underrepresented groups
-elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{65}}.
+elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{68}}.
 
@@
 Ontario. However, these areas have the highest proportion of Indigenous
-inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{62}}.
+inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{65}}.
 In the context of personalized care, there is a need to collect data
@@
 vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{66}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{69}}, we
 believe that our data was able to provide a valid sample of vaccination
@@
 October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{67}});
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{70}});
 however, the short time window constitutes essentially a ``snapshot''
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kayser2021}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-watson2022}{}}%
 \CSLLeftMargin{9. }%
-\CSLRightInline{Kayser V, Ramzan I. Vaccines and vaccination: History
-and emerging issues. \emph{Human Vaccines {\&}amp\(\mathsemicolon\)
-Immunotherapeutics}. 2021;17(12):5255-5268.
-doi:\href{https://doi.org/10.1080/21645515.2021.1977057}{10.1080/21645515.2021.1977057}}
+\CSLRightInline{Watson OJ, Barnsley G, Toor J, Hogan AB, Winskill P,
+Ghani AC. Global impact of the first year of {COVID}-19 vaccination: A
+mathematical modelling study. \emph{The Lancet Infectious Diseases}.
+2022;22(9):1293-1302.
+doi:\href{https://doi.org/10.1016/s1473-3099(22)00320-6}{10.1016/s1473-3099(22)00320-6}}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-dolgin2021}{}}%
+\CSLLeftMargin{11. }%
+\CSLRightInline{Dolgin E. {COVID} vaccine immunity is waning
+{\textemdash} how much does that matter? \emph{Nature}.
+2021;597(7878):606-607.
+doi:\href{https://doi.org/10.1038/d41586-021-02532-4}{10.1038/d41586-021-02532-4}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-gerretsen2021}{}}%
-\CSLLeftMargin{11. }%
+\CSLLeftMargin{12. }%
 \CSLRightInline{Gerretsen P, Kim J, Caravaggio F, et al. Individual
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tamey2022}{}}%
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{13. }%
 \CSLRightInline{Yamey G, Garcia P, Hassan F, et al. It is not too late
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nafilyan2021}{}}%
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{14. }%
 \CSLRightInline{Nafilyan V, Dolby T, Razieh C, et al. Sociodemographic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-willis2021}{}}%
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{15. }%
 \CSLRightInline{Willis DE, Andersen JA, Bryant-Moore K, et al.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-skirrow2022}{}}%
-\CSLLeftMargin{15. }%
+\CSLLeftMargin{16. }%
 \CSLRightInline{Skirrow H, Barnett S, Bell S, et al. Women's views on
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stoler2021}{}}%
-\CSLLeftMargin{16. }%
+\CSLLeftMargin{17. }%
 \CSLRightInline{Stoler J, Enders AM, Klofstad CA, Uscinski JE. The
@@
 \leavevmode\vadjust pre{\hypertarget{ref-khubchandani2021}{}}%
-\CSLLeftMargin{17. }%
+\CSLLeftMargin{18. }%
 \CSLRightInline{Khubchandani J, Sharma S, Price JH, Wiblishauser MJ,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-guay2022}{}}%
-\CSLLeftMargin{18. }%
+\CSLLeftMargin{19. }%
 \CSLRightInline{Guay M, Maquiling A, Chen R, et al. Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-muhajarine2021}{}}%
-\CSLLeftMargin{19. }%
+\CSLLeftMargin{20. }%
 \CSLRightInline{Muhajarine N, Adeyinka DA, McCutcheon J, Green KL,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hussain2022}{}}%
-\CSLLeftMargin{20. }%
+\CSLLeftMargin{21. }%
 \CSLRightInline{Hussain B, Latif A, Timmons S, Nkhoma K, Nellums LB.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mosby2021}{}}%
-\CSLLeftMargin{21. }%
+\CSLLeftMargin{22. }%
 \CSLRightInline{Mosby I, Swidrovich J. Medical experimentation and the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bogart2021}{}}%
-\CSLLeftMargin{22. }%
+\CSLLeftMargin{23. }%
 \CSLRightInline{Bogart LM, Ojikutu BO, Tyagi K, et al. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-freeman2020}{}}%
-\CSLLeftMargin{23. }%
+\CSLLeftMargin{24. }%
 \CSLRightInline{Freeman D, Loe BS, Chadwick A, et al. {COVID}-19 vaccine
@@
 \leavevmode\vadjust pre{\hypertarget{ref-malik2020}{}}%
-\CSLLeftMargin{24. }%
+\CSLLeftMargin{25. }%
 \CSLRightInline{Malik AA, McFadden SM, Elharake J, Omer SB. Determinants
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021}{}}%
-\CSLLeftMargin{25. }%
+\CSLLeftMargin{26. }%
 \CSLRightInline{Nguyen KH, Nguyen K, Corlin L, Allen JD, Chung M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mollalo2021}{}}%
-\CSLLeftMargin{26. }%
+\CSLLeftMargin{27. }%
 \CSLRightInline{Mollalo A, Tatar M. Spatial modeling of {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-yang2022}{}}%
-\CSLLeftMargin{27. }%
+\CSLLeftMargin{28. }%
 \CSLRightInline{Yang TC, Matthews SA, Sun F. Multiscale dimensions of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tiu2022}{}}%
-\CSLLeftMargin{28. }%
+\CSLLeftMargin{29. }%
 \CSLRightInline{Tiu A, Susswein Z, Merritt A, Bansal S. Characterizing
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bhuiyan2022}{}}%
-\CSLLeftMargin{29. }%
+\CSLLeftMargin{30. }%
 \CSLRightInline{Bhuiyan MAN, Davis TC, Arnold CL, et al. Using the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wood2022}{}}%
-\CSLLeftMargin{30. }%
+\CSLLeftMargin{31. }%
 \CSLRightInline{Wood AJ, MacKintosh AM, Stead M, Kao RR. Predicting
@@
 \leavevmode\vadjust pre{\hypertarget{ref-choi2021}{}}%
-\CSLLeftMargin{31. }%
+\CSLLeftMargin{32. }%
 \CSLRightInline{Choi KH, Denice PA, Ramaj S. Vaccine and {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mckinnon2021}{}}%
-\CSLLeftMargin{32. }%
+\CSLLeftMargin{33. }%
 \CSLRightInline{McKinnon B, Quach C, Dubé Ève, Nguyen CT, Zinszer K.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
@@
 \leavevmode\vadjust pre{\hypertarget{ref-muratov2018}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Muratov S, Lee J, Holbrook A, et al. Regional variation
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Dong L, Sahu R, Black R. Governance in the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lysyk2015}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lysyk2016}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-rotenberg2021}{}}%
+\CSLLeftMargin{39. }%
+\CSLRightInline{Rotenberg S, Downer MB, Brown H, et al. \emph{{COVID}-19
+Vaccination for People with Disabilities}. Ontario {COVID}-19 Science
+Advisory Table; 2021.
+doi:\href{https://doi.org/10.47326/ocsat.2021.02.35.1.0}{10.47326/ocsat.2021.02.35.1.0}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ashcroft2023}{}}%
+\CSLLeftMargin{40. }%
+\CSLRightInline{Ashcroft R, Donnelly C, Lam S, et al. A qualitative
+examination of primary care team's participation in the distribution of
+the {COVID}-19 vaccination. Published online June 2023.
+doi:\href{https://doi.org/10.21203/rs.3.rs-3016276/v1}{10.21203/rs.3.rs-3016276/v1}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-sethuram2023}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Sethuram C, McCutcheon T, Liddy C. An environmental scan
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
@@
 \leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
@@
 \leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-basta2022}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Basta NE, Sohel N, Sulis G, et al. Factors associated
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2023}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Cénat JM, Farahi SMMM, Bakombo SM, et al. Vaccine
@@
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-iveniuk2021}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{Iveniuk J, Leon S. \emph{Uneven Recovery: Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gill2022}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Gill M, Datta D, Gregory P, Austin Z. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021b}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Nguyen KH, Yankey D, Coy KC, et al. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shah2019}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Shah TI, Clark AF, Seabrook JA, Sibbald S, Gilliland JA.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-crighton2015}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Crighton EJ, Ragetlie R, Luo J, To T, Gershon A. A
@@
 \leavevmode\vadjust pre{\hypertarget{ref-timony2022}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Timony P, Houle SKD, Gauthier A, Waite NM. Geographic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
@@
 \leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,

### 2023-07-02 manuscript_v6/01-background.qmd
updated methods

@@
 Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], indicating that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022].
+However, there is a limited amount of studies that have analyzed geographical differences in vaccination in Canada.
+Existing studies in this area have focused on differences within certain cities, such as Toronto [@choi2021], or Montréal [@mckinnon2021]), or have explored differences between provinces [@guay2022], but to our knowledge, there are no studies that analyze  differences in vaccination uptake at the provincial level.
+Such need is specially important in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023].
+Analyzing differences in vaccination uptake within the provinces can aid to identify inequalities that might exist and that need to be addressed before the advent of another pandemic.
 
-However, the majority of such analyses have been carried out outside of Canada, and existing studies have been focused in certain cities (such as Toronto [@choi2021], or Montreal [@mckinnon2021]), or have explored differences at a province-wide level [@guay2022].
-**Canada’s pandemic response goals, were to reduce morbidity/mortality and to reduce societal disruption from COVID-19**.
+<!-- However, in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
+-->
 
-- How does this integrate with vaccination? We need to study vaccination because 
--
+<!-- **Canada’s pandemic response goals, were to reduce morbidity/mortality and to reduce societal disruption from COVID-19**.   -->
 
-- -
+<!-- - How does this integrate with vaccination? We need to study vaccination because  -->
+<!-- -  -->
 
-and that consequently, can help identify disparities that need to be addressed within specific areas in each province.
+<!-- - - -->
 
-This need is especially important in the case of Ontario, the most populated province of Canada.
-Between 2007 and 2019, Ontario managed healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integration Networks (LHINs), which aimed to provide an integrated health system for the province.
-However, this approach was complex, bureaucratic, and resulted in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators such as wait times and hospital readmissions, fragmented electronic health systems, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016].
-Therefore, with the intent of better organizing and delivering care in late 2019 the provincial government eliminated the LHINs and incorporated the areas covered by them into six larger Health Regions (North East, North West, Central, Toronto, West, and East), which are managed by the new government agency Ontario Health (OH) [@dong2022].
+<!-- and that consequently, can help identify disparities that need to be addressed within specific areas in each province. -->
 
-On the other hand, public health in Ontario is administered by Public Health Ontario (PHO), a government agency established in late 2007 and currently composed of 34 Public Health Units (PHUs) that cover the entire geography of the province.
-Although these PHUS were commissioned with leading the distribution of the COVID-19 vaccine within the areas they covered [@rotenberg2021], there was significant overlap between PHO and OH in the actual vaccine rollout, and in many instances vaccines were administered in vaccination clinics ran by members of OH working alone or in conjunction with members of PHO [@ashcroft2023].
+<!-- However, in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
+-->
 
-- add about intersection of Health Regions and PHUs. How vaccination presented a challenge as vaccine was ultimately administered in facilites managed by OHT?
+This need is especially important in the case of Ontario, the most populated province of Canada, and which as has a complex healthcare system.
+Between 2007 and 2019, Ontario managed healthcare access to its inhabitants using 14 intra-provincial divisions called the Local Health Integration Networks (LHINs), which aimed to provide an integrated health system.
+However, this approach was complex, bureaucratic, resulting in excessive expenditures, disparities in mortality rates, the deterioration of certain performance indicators (such as wait times and hospital readmissions), fragmented electronic health systems, and inequities in health services access [@tsasis2012;@muratov2018;@dong2022;@lysyk2015;@lysyk2016].
+With the intent of better organizing and delivering care, in late 2019 the provincial government eliminated the LHINs and incorporated the areas covered by them into six larger Health Regions (North East, North West, Central, Toronto, West, and East), which are managed by a new government agency, Ontario Health (OH) [@dong2022].
 
-https://www.mcmasterforum.org/docs/default-source/product-documents/rapid-responses/examining-the-intersections-between-ontario-health-teams-and-public-health.pdf?sfvrsn=69caa8c9_11
+On the other hand, public health in Ontario is administered by Public Health Ontario (PHO), a government agency established in late 2007 and that is currently composed of 34 Public Health Units (PHUs) that cover the entire geography of the province.
+Although PHUS were commissioned with leading the distribution of the COVID-19 vaccine within their respective areas [@rotenberg2021], the vaccine rollout occurred with significant interaction between PHO and OH; in many instances both agencies had to work together to organize vaccination clinics, with personnel associated with OH also being actively involved when demand exceeded the capacity of PHO personnel, especially in rural areas [@ashcroft2023].
+Indeed, based on the experience of COVID-19, there is an ongoing debate on how OH and PHO will interact in the future, and the challenges that this will entail for the healthcare system of the province [@dion2022].
 
-Because of the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
+Therefore, considering the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic, there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within the Health Regions and identify the socio-demographic groups that might be affected, as this can serve as an indirect assessment of the state of the implementation of the new model while helping identify ongoing challenges that decision-makers might need to address to ensure the long-term  success of this model and its interaction with PHO, which is specially important considering that previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
+
+<!-- before the advent of a future pandemic -->
+
+<!-- In the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within Ontario and identify the socio-demographic groups that might be affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
+-->
+
+<!-- there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
 In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
 Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
+-->
 
-Therefore, in this study we aim to understand the differences in vaccination uptake between the different Health Regions of Ontario between October of 2021 and January of 2022.
-By including socio-economic factors in our analysis, we aimed to identify in which groups these differences were significant in order to provide an assessment of the current state of healthcare access in Ontario.
+<!-- Due to the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic,  -->
+
+<!-- In the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within Ontario and identify the socio-demographic groups that might be affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
+-->
+
+<!-- how the relatively recent adoption of the Health Region model the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
+-->
+
+<!-- there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
+In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
+Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
+-->
+
+<!-- However, in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
+-->
+
+<!-- there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
+In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
+Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
+-->
+
+Therefore, in this study we hypothesized that there were differences in COVID-19 vaccination rates between the Health Regions between October 2021 and January of 2022.
+To understand and contextualize these differences, we by included socio-economic factors in our analysis, aiming to identify in which demographic groups were particularly impacted, in order to provide an assessment of the current state of healthcare access in Ontario.

### 2023-07-02 manuscript_v6/02-methods.qmd
updated methods

--- 2023-06-09 manuscript_v5/02-methods.qmd
@@
 We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto (under protocol 00043317), and which ran between September 30th, 2021 and January 17th,2022.
-The survey collected socio-economic information from participants (@tbl-descriptive-stats), their location (nearest municipality,  as shown in @fig-map), the date of access to the survey, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no".
-The original dataset contained 39,029 observations.
 
-By design, the survey allowed respondents to exit at any time and deployed the questions randomly.
-This resulted in $\approx$ 84\% of the observations having multiple missing answers or being incomplete due to participants leaving the survey at different stages.
-Therefore, we selected 6,343 observations that were labelled as "complete" in the dataset and that had answers for all covariates considered in our analysis.
+Briefly, the survey was deployed using random domain intercept technology, a methodology where internet users whose device meta-data indicated their presence in the province of Ontario had a random chance of being redirected to the survey after they had clicked on a registered but commercially inactive web link, or typed in a web address for a site that was dormant but that was temporarily managed by the company that administered the survey (RIWI Corp) [@sargent2022].
+Users then decided whether to anonymously participate in the survey, and those that participated were able to exit the survey at any time.
+After the survey closed, regardless if it was complete or incomplete, access was denied to any further users with the same internet protocol address (IP), effectively allowing each user only one opportunity to participate in the survey.
+Users who indicated they were under the age of 16 were exited from the survey without creating a record.
+Finally, the personal identifier information from each user that participated in the survey was automatically scrubbed and replaced by a unique ID.
+
+The survey allowed users to enter their socio-economic information, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no" (@tbl-descriptive-stats).
+Additionally, the survey automatically collected the geographical location of the respondent (using the nearest municipality,  as shown in @fig-map), and the date of access to the survey.
+The original dataset contained 39,029 observations with a response rate of 16.25% (observations where users had complete answers), due to the survey design that allowed respondents to exit at any time and deployed the questions randomly.
+We selected these  observations with complete answers (6,343 observations) for our analysis.
 Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions).

### 2023-07-02 manuscript_v6/main.tex
updated methods

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, boxrule=0pt, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, boxrule=0pt, sharp corners, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 vaccines\textsuperscript{\protect\hyperlink{ref-willis2021}{15}}, and
-the influence of misinformation and
+the existence of health misinformation and
 disinformation\textsuperscript{\protect\hyperlink{ref-mosby2021}{22}--\protect\hyperlink{ref-freeman2020}{24}}.
@@
 can help predict patterns of booster
-uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{31}}.
-
-However, such analyses have been carried out mostly in territories
-outside of Canada, where available studies have been focused in certain
-cities (such as
+uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{31}}. However,
+there is a limited amount of studies that have analyzed geographical
+differences in vaccination in Canada. Existing studies in this area have
+focused on differences within certain cities, such as
 Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{32}}, or
-Montreal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{33}}), or
-have explored differences at a province-wide
-level\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}. Therefore,
-there is a need for studies that explore spatial differences in
-vaccination within the Canadian territory and that consequently, can
-help identify disparities that need to be addressed within specific
-areas in each province.
+Montréal\textsuperscript{\protect\hyperlink{ref-mckinnon2021}{33}}), or
+have explored differences between
+provinces\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}, but to
+our knowledge, there are no studies that analyze differences in
+vaccination uptake at the provincial level. Such need is specially
+important in the context of Canada's pandemic response goals, which have
+been to minimize serious illness and deaths while minimizing societal
+disruption\textsuperscript{\protect\hyperlink{ref-naci2023}{34}}.
+Analyzing differences in vaccination uptake within the provinces can aid
+to identify inequalities that might exist and that need to be addressed
+before the advent of another pandemic.
 
 This need is especially important in the case of Ontario, the most
-populated province of Canada. Between 2007 and 2019, Ontario managed
-healthcare access to its inhabitants using 14 intra-provincial divisions
-called the Local Health Integration Networks (LHINs), which aimed to
-provide an integrated health system for the province. However, this
-approach was complex, bureaucratic, and resulted in excessive
-expenditures, disparities in mortality rates, the deterioration of
-certain performance indicators such as wait times and hospital
-readmissions, fragmented electronic health systems, and inequities in
-health services
-access\textsuperscript{\protect\hyperlink{ref-tsasis2012}{34}--\protect\hyperlink{ref-lysyk2016}{38}}.
-Therefore, with the intent of better organizing and delivering care in
-late 2019 the provincial government eliminated the LHINs and
-incorporated the areas covered by them into six larger Health Regions
-(North East, North West, Central, Toronto, West, and East), which are
-managed by the new government agency Ontario Health
-(OH)\textsuperscript{\protect\hyperlink{ref-dong2022}{36}}.
+populated province of Canada, and which as has a complex healthcare
+system. Between 2007 and 2019, Ontario managed healthcare access to its
+inhabitants using 14 intra-provincial divisions called the Local Health
+Integration Networks (LHINs), which aimed to provide an integrated
+health system. However, this approach was complex, bureaucratic,
+resulting in excessive expenditures, disparities in mortality rates, the
+deterioration of certain performance indicators (such as wait times and
+hospital readmissions), fragmented electronic health systems, and
+inequities in health services
+access\textsuperscript{\protect\hyperlink{ref-tsasis2012}{35}--\protect\hyperlink{ref-lysyk2016}{39}}.
+With the intent of better organizing and delivering care, in late 2019
+the provincial government eliminated the LHINs and incorporated the
+areas covered by them into six larger Health Regions (North East, North
+West, Central, Toronto, West, and East), which are managed by a new
+government agency, Ontario Health
+(OH)\textsuperscript{\protect\hyperlink{ref-dong2022}{37}}.
 
@@
 Health Ontario (PHO), a government agency established in late 2007 and
-currently composed of 34 Public Health Units (PHUs) that cover the
-entire geography of the province. Although these PHUS were commissioned
-with leading the distribution of the COVID-19 vaccine within the areas
-they covered\textsuperscript{\protect\hyperlink{ref-rotenberg2021}{39}},
-there was significant overlap between PHO and OH in the actual vaccine
-rollout, and in many instances vaccines were administered in vaccination
-clinics ran by members of OH working alone or in conjunction with
-members of
-PHO\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{40}}.
-
-\begin{itemize}
-\tightlist
-\item
-  add about intersection of Health Regions and PHUs. How vaccination
-  presented a challenge as vaccine was ultimately administered in
-  facilites managed by OHT?
-\end{itemize}
-
-https://www.mcmasterforum.org/docs/default-source/product-documents/rapid-responses/examining-the-intersections-between-ontario-health-teams-and-public-health.pdf?sfvrsn=69caa8c9\_11
-
-Because of the relatively recent adoption of the Health Region model and
-its alignment with the onset of the COVID-19 pandemic, there is a need
-to analyze if there are ongoing disparities in health access under this
-approach that need to be addressed before they are exploited by a new
-disease or public health threat. In this regard, previous research has
-highlighted disparities in the level of activity of each Health
-Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{41}}.
-Therefore, analyzing differences in vaccination uptake within the Health
-Regions and can help identify which socio-demographic groups are the
-most vulnerable and what areas of the province deserve special attention
-by decision-makers.
-
-Therefore, in this study we aim to understand the differences in
-vaccination uptake between the different Health Regions of Ontario
-between October of 2021 and January of 2022. By including socio-economic
-factors in our analysis, we aimed to identify in which groups these
-differences were significant in order to provide an assessment of the
-current state of healthcare access in Ontario.
+that is currently composed of 34 Public Health Units (PHUs) that cover
+the entire geography of the province. Although PHUS were commissioned
+with leading the distribution of the COVID-19 vaccine within their
+respective
+areas\textsuperscript{\protect\hyperlink{ref-rotenberg2021}{40}}, the
+vaccine rollout occurred with significant interaction between PHO and
+OH; in many instances both agencies had to work together to organize
+vaccination clinics, with personnel associated with OH also being
+actively involved when demand exceeded the capacity of PHO personnel,
+especially in rural
+areas\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}}. Indeed,
+based on the experience of COVID-19, there is an ongoing debate on how
+OH and PHO will interact in the future, and the challenges that this
+will entail for the healthcare system of the
+province\textsuperscript{\protect\hyperlink{ref-dion2022}{42}}.
+
+Therefore, considering the relatively recent adoption of the Health
+Region model and its alignment with the onset of the COVID-19 pandemic,
+there is an ongoing need to analyze the existence of geographical
+disparities in vaccination uptake within the Health Regions and identify
+the socio-demographic groups that might be affected, as this can serve
+as an indirect assessment of the state of the implementation of the new
+model while helping identify ongoing challenges that decision-makers
+might need to address to ensure the long-term success of this model and
+its interaction with PHO, which is specially important considering that
+previous research has highlighted disparities in the level of activity
+of each Health
+Region\textsuperscript{\protect\hyperlink{ref-sethuram2023}{43}}.
+
+Therefore, in this study we hypothesized that there were differences in
+COVID-19 vaccination rates between the Health Regions between October
+2021 and January of 2022. To understand and contextualize these
+differences, we by included socio-economic factors in our analysis,
+aiming to identify in which demographic groups were particularly
+impacted, in order to provide an assessment of the current state of
+healthcare access in Ontario.
 
@@
 Toronto (under protocol 00043317), and which ran between September 30th,
-2021 and January 17th,2022. The survey collected socio-economic
-information from participants (Table~\ref{tbl-descriptive-stats}), their
-location (nearest municipality, as shown in Figure~\ref{fig-map}), the
-date of access to the survey, and asked information on vaccination
-status by using the question ``Have you received the first dose of the
-COVID vaccine?'', with possible answers ``yes'' and ``no''. The original
-dataset contained 39,029 observations.
-
-By design, the survey allowed respondents to exit at any time and
-deployed the questions randomly. This resulted in \(\approx\) 84\% of
-the observations having multiple missing answers or being incomplete due
-to participants leaving the survey at different stages. Therefore, we
-selected 6,343 observations that were labelled as ``complete'' in the
-dataset and that had answers for all covariates considered in our
-analysis. Later, we matched the city of each observation with its
-corresponding LHIN and Health Region, and removed observations from
-areas with low representation (254 observations corresponding to the
-North West and North East Health Regions). After all the preliminary
-analyses, the total number of observations used for analysis was 6,236
-and included the East, Central, Toronto, and West Health Regions
-covering between October 1st,2021 and January 17, 2022. The original
-dataset, clean dataset, and details on the data cleaning process are
-described in detail in the
-\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
-for this paper.
+2021 and January 17th,2022.
+
+Briefly, the survey was deployed using random domain intercept
+technology, a methodology where internet users whose device meta-data
+indicated their presence in the province of Ontario had a random chance
+of being redirected to the survey after they had clicked on a registered
+but commercially inactive web link, or typed in a web address for a site
+that was dormant but that was temporarily managed by the company that
+administered the survey (RIWI
+Corp)\textsuperscript{\protect\hyperlink{ref-sargent2022}{44}}. Users
+then decided whether to anonymously participate in the survey, and those
+that participated were able to exit the survey at any time. After the
+survey closed, regardless if it was complete or incomplete, access was
+denied to any further users with the same internet protocol address
+(IP), effectively allowing each user only one opportunity to participate
+in the survey. Users who indicated they were under the age of 16 were
+exited from the survey without creating a record. Finally, the personal
+identifier information from each user that participated in the survey
+was automatically scrubbed and replaced by a unique ID.
+
+The survey allowed users to enter their socio-economic information , and
+asked information on vaccination status by using the question ``Have you
+received the first dose of the COVID vaccine?'', with possible answers
+``yes'' and ``no'' (Table~\ref{tbl-descriptive-stats}). Additionally,
+the survey automatically collected the geographical location of the
+respondent (using the nearest municipality, as shown in
+Figure~\ref{fig-map}), and the date of access to the survey. The
+original dataset contained 39,029 observations with a response rate of
+16.25\% (defined as observations that had complete answers), due to the
+survey design that allowed respondents to exit at any time and deployed
+the questions randomly. We selected 6,343 observations that were
+labelled as ``complete'' in the dataset and that corresponded to those
+that had answers for all covariates considered in our analysis. Later,
+we matched the city of each observation with its corresponding LHIN and
+Health Region, and removed observations from areas with low
+representation (254 observations corresponding to the North West and
+North East Health Regions). After all the preliminary analyses, the
+total number of observations used for analysis was 6,236 and included
+the East, Central, Toronto, and West Health Regions covering between
+October 1st,2021 and January 17, 2022. The original dataset, clean
+dataset, and details on the data cleaning process are described in
+detail in the \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub
+repository} for this paper.
 
@@
 of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{42}--\protect\hyperlink{ref-cnat2022a}{44}}.
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{45}--\protect\hyperlink{ref-cnat2022a}{47}}.
 Because we identified differences in representativity between the survey
@@
 proportional fitting procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{45}}
+(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{48}}
 to correct the data using data from the Census and Health Region
@@
 Appendix. All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{46}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{47}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{48}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{49}},
+\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{49}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{50}},
+\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{51}},
+\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{52}},
 and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{50}}.
+\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{53}}.
 
@@
 vaccination rates in individuals with the same socio-demographic
-characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{51}}.
+characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{54}}.
 
@@
 in health
-settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{44},\protect\hyperlink{ref-basta2022}{52}--\protect\hyperlink{ref-cnat2023}{54}}.
+settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{47},\protect\hyperlink{ref-basta2022}{55}--\protect\hyperlink{ref-cnat2023}{57}}.
 However, it has been shown that structural barriers also play an
@@
 vaccination
-sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{55}}. In this
+sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{58}}. In this
 regard, it is interesting to note that vaccination venues were scarce in
@@
 Toronto and other regions of Ontario around the time covered by the
-survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{56}},
+survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{59}},
 and that pharmacies in the Peel region (an area identified as a
@@
 households) could not keep up with vaccine
-demand\textsuperscript{\protect\hyperlink{ref-gill2022}{57}}. This
+demand\textsuperscript{\protect\hyperlink{ref-gill2022}{60}}. This
 suggests that the observed differences are associated with disparities
@@
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{58},\protect\hyperlink{ref-ct2021}{59}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{61},\protect\hyperlink{ref-ct2021}{62}},
 which include workers in the areas of grocery stores, gas stations,
@@
 COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{60}}; and
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{63}}; and
 there is evidence of interventions by vaccination staff in certain parts
 of the province to encourage vaccination uptake by these
-individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{57}}. These
+individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{60}}. These
 facts, combined with evidence of increased trends in vaccination in this
 group
-elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{61}},
+elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{64}},
 suggest that the type of occupation of individuals in underrepresented
@@
 pharmacist
-availability\textsuperscript{\protect\hyperlink{ref-shah2019}{62}--\protect\hyperlink{ref-timony2022}{64}}.
+availability\textsuperscript{\protect\hyperlink{ref-shah2019}{65}--\protect\hyperlink{ref-timony2022}{67}}.
 
@@
 to an estimated 100,000 Indigenous individuals living in the
-area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{65}}. Lack of
+area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{68}}. Lack of
 access to personalized healthcare affects individuals that may mistrust
@@
 vaccination rates among these
-groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{66},\protect\hyperlink{ref-eissa2021}{67}}.
+groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{69},\protect\hyperlink{ref-eissa2021}{70}}.
 Taken together, this suggests that healthcare disparities specific to
@@
 underrepresented groups
-elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{68}}.
+elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{71}}.
 
@@
 Ontario. However, these areas have the highest proportion of Indigenous
-inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{65}}.
+inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{68}}.
 In the context of personalized care, there is a need to collect data
@@
 vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{69}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{72}}, we
 believe that our data was able to provide a valid sample of vaccination
@@
 October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{70}});
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{73}});
 however, the short time window constitutes essentially a ``snapshot''
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-naci2023}{}}%
+\CSLLeftMargin{34. }%
+\CSLRightInline{National Advisory Committee on Immunization (NACI).
+\emph{Guidance on an Additional COVID-19 Booster Dose in the Spring of
+2023 for Individuals at High Risk of Severe Illness Due to COVID-19}.
+Public Health Agency of Canada; 2023.
+\url{https://www.canada.ca/en/public-health/services/publications/vaccines-immunization/national-advisory-committee-immunization-guidance-additional-covid-19-booster-dose-spring-2023-individuals-high-risk-severe-illness-due-covid-19.html}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-tsasis2012}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Tsasis P, Evans JM, Owen S. Reframing the challenges to
@@
 \leavevmode\vadjust pre{\hypertarget{ref-muratov2018}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Muratov S, Lee J, Holbrook A, et al. Regional variation
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dong2022}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Dong L, Sahu R, Black R. Governance in the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lysyk2015}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lysyk2016}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Auditor General of Ontario O of the, ed. Annual {R}eport
@@
 \leavevmode\vadjust pre{\hypertarget{ref-rotenberg2021}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Rotenberg S, Downer MB, Brown H, et al. \emph{{COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ashcroft2023}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Ashcroft R, Donnelly C, Lam S, et al. A qualitative
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-dion2022}{}}%
+\CSLLeftMargin{42. }%
+\CSLRightInline{Dion A, Waddell K, Moat K, Reid R, Lavis J. \emph{Rapid
+Synthesis: Examining the Intersections Between Ontario Health Teams and
+Public Health}. McMaster Health Forum; 2022.
+\url{https://www.mcmasterforum.org/docs/default-source/product-documents/rapid-responses/examining-the-intersections-between-ontario-health-teams-and-public-health.pdf?sfvrsn=69caa8c9_11}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-sethuram2023}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Sethuram C, McCutcheon T, Liddy C. An environmental scan
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-sargent2022}{}}%
+\CSLLeftMargin{44. }%
+\CSLRightInline{Sargent RH, Laurie S, Weakland LF, et al. Use of random
+domain intercept technology to track {COVID}-19 vaccination rates in
+real time across the united states: Survey study. \emph{Journal of
+Medical Internet Research}. 2022;24(7):e37920.
+doi:\href{https://doi.org/10.2196/37920}{10.2196/37920}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
@@
 \leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
@@
 \leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-basta2022}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Basta NE, Sohel N, Sulis G, et al. Factors associated
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2023}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Cénat JM, Farahi SMMM, Bakombo SM, et al. Vaccine
@@
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-iveniuk2021}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Iveniuk J, Leon S. \emph{Uneven Recovery: Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gill2022}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Gill M, Datta D, Gregory P, Austin Z. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021b}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Nguyen KH, Yankey D, Coy KC, et al. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shah2019}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Shah TI, Clark AF, Seabrook JA, Sibbald S, Gilliland JA.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-crighton2015}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Crighton EJ, Ragetlie R, Luo J, To T, Gershon A. A
@@
 \leavevmode\vadjust pre{\hypertarget{ref-timony2022}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Timony P, Houle SKD, Gauthier A, Waite NM. Geographic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
@@
 \leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{71. }%
 \CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{69. }%
+\CSLLeftMargin{72. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{70. }%
+\CSLLeftMargin{73. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,

### 2023-07-04 manuscript_v6/02-methods.qmd
updated methods and results

@@
 
-We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto (under protocol 00043317), and which ran between September 30th, 2021 and January 17th,2022.
+We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto (under Research Ethics Board approval #40999), and which ran between September 30th, 2021 and January 17th,2022.
 
@@
 
-The survey allowed users to enter their socio-economic information, and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no" (@tbl-descriptive-stats).
+Survey users entered their socio-economic information (age, income, and racial/ethnic group), and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no" (@tbl-descriptive-stats).
+Of notice, because the racial/ethnic categories of the survey did not match exactly the categories from the 2016 Census, we used a combination of sources to re-group certain categories in order to obtain estimates that could be used to correct the data.
+Further details can be found in the Appendix.
+
 Additionally, the survey automatically collected the geographical location of the respondent (using the nearest municipality,  as shown in @fig-map), and the date of access to the survey.
@@
 We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
-Because we identified differences in representativity between the survey data and the estimates from the Census, we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data.
+Because we identified differences in representativity between the survey data and the estimates from the 2016 Census for the variables considered in the analysis, and used used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data.
 Details regarding the correction can be found in the Appendix.

### 2023-07-04 manuscript_v6/03-results.qmd
updated methods and results

@@
 @tbl-descriptive-stats shows the characteristics of the data from the Fields COVID-19 survey used for analysis.
-The sample contained 6,236 observations, from which 24.8\% (1,547) corresponded to individuals that reported not having received the first dose of the vaccine.
-Vaccination rates ranged between 71-79\% across all household income brackets, age groups, Health Regions, and the months considered in the survey.
-However, the highest vaccination rates in each category were reported by individuals in the highest income bracket (79\%), those between 16 and 34 years of age (77\%), individuals that lived in the East Health Region (77\%), and  during January of 2022 (78\%).
-Between racial/ethnic groups, the highest vaccination rate was reported by White/Caucasian individuals (84\%), against vaccination rates between 63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and those that reported belonging to "Other" racial groups, which included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere.
+The sample contained 6,236 observations, from which 24.8\% (1,547) corresponded to individuals that reported not having initiated a COVID-19 vaccine primary series (in other words, not having received the first dose of the vaccine).
+The rate for the first dose of the vaccine ranged between 71-79\% across all household income brackets, age groups, Health Regions, and the months considered in the survey.
+However, the highest rate for the uptake of the first dose of the vaccine were reported by individuals in the highest income bracket (79\%), those between 16 and 34 years of age (77\%), individuals that lived in the East Health Region (77\%), and  during January of 2022 (78\%).
+Between racial/ethnic groups, White/Caucasian individuals reported the highest uptake of the first dose of the vaccine (84\%), against values that ranged between 63 and 66\% in the case of Arab/Middle Eastern, Black, Indigenous, Latin American individuals, and those that belonged to the "other" racial group category (which included Southeast Asian, Filipino, West Asian, and minorities not identified elsewhere percentages ).
 
@@
     rows=label=="Other",
-    footnote="Southeast Asian, Filipino, West Asian, \nand minorities not identified elsewhere according to the Census."
+    footnote="This category included Southeast Asian, Filipino, and West Asian individuals, \nand minorities not identified elsewhere according to the 2016 Census."
   ) %>%

### 2023-07-04 manuscript_v6/04-discussion.qmd
updated methods and results

@@
 
-In this study, we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions between late 2021 and early 2022.
+In this study, we hypothesized that differences in the uptake of the first dose of the COVID-19 vaccine were present between the Health Regions between late 2021 and early 2022.
 Our goal was to determine which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensuring that the Health Region model is able to fulfill its mission of improving health access for all Ontarians.

### 2023-07-07 manuscript_v6/01-background.qmd
updated discussion and conclusion, acknowlegements

@@
 However, there is a limited amount of studies that have analyzed geographical differences in vaccination in Canada.
-Existing studies in this area have focused on differences within certain cities, such as Toronto [@choi2021], or Montréal [@mckinnon2021]), or have explored differences between provinces [@guay2022], but to our knowledge, there are no studies that analyze  differences in vaccination uptake at the provincial level.
+Existing studies in this area have focused on differences within certain cities, such as Toronto [@choi2021], or Montréal [@mckinnon2021]), or have explored differences between provinces [@guay2022], but to our knowledge, there are no studies that analyze differences in vaccination uptake at the intra-provincial level.
 Such need is specially important in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023].

### 2023-07-07 manuscript_v6/04-discussion.qmd
updated discussion and conclusion, acknowlegements

--- 2023-06-27 manuscript_v6/04-discussion.qmd
@@
 
-In this study, we hypothesized that differences in COVID-19 vaccination uptake were present between the Health Regions between late 2021 and early 2022.
+The COVID-19 pandemic has presented many significant challenges to the healthcare systems of the Canadian provinces [@alami2021].
+In the case of Ontario, the recent adoption of the Health Region model at the onset of the pandemic signified an additional level of complexity because although PHO (through the 34 Public Health Units of the province) was the lead in vaccine distribution, in reality the vaccine rollout required active collaboration between PHO and OH in multiple instances [@ashcroft2023].
+
+In this study, we hypothesized that there were differences in the uptake of the first dose of the COVID-19 vaccine between individuals living in the different Health Regions of Ontario during late 2021 and early 2022.
+Using the Health Regions as the base of our analysis was advantageous, as these new Health Regions match closely the geographical boundaries of the Health Regions that have been historically used to group Public Health Units in the province [@bolotin2018;@ohesi], thus providing information with relevance at both the public health and healthcare system level.
+Therefore, this analysis could provide an overall assessment of intra-provincial disparities that might need to be addressed moving forward by decision-makers, in order to ensure that in the future, and specially in the event of a future public health emergency, OH and PHO will able to collaboratively work in an efficient way and thus ensure that the population of the province benefits in the long term of the newly implemented Health Region model.
+
+<!-- Specifically, our goal was to determine which socio-demographic groups could be impacted by these differences in vaccination rates in order to highlight challenges that will need to be addressed by OH to determine the exact nature of its future interactions with PHO in the event of a future public health emergency.
+-->
+
+<!-- The complex relationships within the health system of Ontario, with the recent implementation of OH make necessary to determine  COVID-19 pandemic highlighted the need of an active collaboration between PHO and OH -->
+
+<!-- The recent implementation of the Health Region model  vaccination campaign in Ontario  -->
+
+<!-- Due to the recent implementation of the Health Region model in the province, and the Our goal was to determine which socio-demographic groups could be impacted by these disparities, in order to provide decision-makers with information that could be used to inform  -->
+
+<!-- develop policies focused on reducing or eliminating these differences and ensuring that the Health Region model is able to fulfill its mission of improving health access for all Ontarians. -->
+
+<!-- In this study, we hypothesized that differences in the uptake of the first dose of the COVID-19 vaccine were present between the Health Regions between late 2021 and early 2022.
 Our goal was to determine which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensuring that the Health Region model is able to fulfill its mission of improving health access for all Ontarians.
+-->
 
-Our results show that indeed, there were differences in vaccination odds across Ontario in certain socio-demographic groups.
+Our results show that indeed, there were differences in the uptake of the first dose of the vaccine across Ontario in certain socio-demographic groups.
 Specifically, those who identified as Arab/Middle Eastern, Black, Latin American, having mixed racial or ethnic background, or that belonged to other groups not explicitly included in the survey (Southeast Asian, Filipino, West Asian, and minority groups not identified elsewhere) had vaccination odds that were between a third and a half of that of individuals that identified as White or Caucasian (@fig-models).
@@
 
-This result is likely reflects in part the fact that individuals in underrepresented groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include workers in the areas of grocery stores, gas stations, warehouses, distribution, and manufacturing, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
-In Ontario, these workers had priority for COVID-19 vaccination [@mishra2021]; and there is evidence of interventions by vaccination staff in certain parts of the province to encourage vaccination uptake by these individuals [@gill2022].
-These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that the type of occupation of individuals in underrepresented groups played an important role in increasing the odds of vaccination.
+This result likely reflects in part the fact that individuals in underrepresented groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include workers in the areas of grocery stores, gas stations, warehouses, distribution, and manufacturing, all being occupations for which an income within the significant brackets identified in the analysis is to be expected.
+From one side, individuals in essential occupations in the province experienced higher rates of morbidity and mortality during the first year of the pandemic [@rao2021], but later on, they had priority for COVID-19 vaccination [@mishra2021].
+Additionally, it is known that vaccination uptake in these individuals was encouraged by vaccination staff in certain parts of the province [@gill2022].
+These facts, combined with evidence of increased trends in vaccination in this group elsewhere [@nguyen2021b], suggest that in Ontario, the type of occupation of individuals in underrepresented groups (which might have also affected their decision to get a vaccine based on their knowledge of increased risk), played an important role in the higher the odds of vaccination observed in these individuals.
 
@@
 Therefore, we focused on entries that had complete answers, and corrected the data using population-wide information from the Census.
-More granular corrections would be needed to obtain more accurate estimates.
-For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates than other racial groups.
-Correcting for each racial/ethnic group in each Health Region can provide a more accurate estimation of region-wide vaccination rates but unfortunately, at the moment this correction cannot be implemented as such stratification is has not been implemented in the Census.
+However, more granular corrections would be needed to obtain more accurate estimates: For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates than other racial groups.
+Correcting for each racial/ethnic group in each Health Region would provide a more accurate estimation of region-wide vaccination rates but unfortunately, presently this correction cannot be implemented as such stratification has not been implemented in the data that can be obtained from the Census.
 

### 2023-07-07 manuscript_v6/05-conclusion.qmd
updated discussion and conclusion, acknowlegements

@@
 The implementation of the Health Regions in Ontario aimed at reducing the bureaucratic complexity and health disparities identified under the LHIN model.
-However, there are currently multiple challenges that need to be addressed to ensure that the new model can improve healthcare for the inhabitants of the province.
-First, the fact that each Health Region now covers a large geographical area that was served by multiple LHINs in the past creates a complex socio-demographic landscape that is different in each case due the different levels of rurality and representation of equity-deserving groups that are now within each Health Region.
-So far, the evidence collected during the COVID-19 pandemic indicates that differences in vaccination uptake are associated to a lack of infrastructure and resources that can adequately support personalized care to marginalized individuals.
+However, there are certain challenges that this new model will need to overcome in order to work syngergistically with the rest of the healthcare system of the province.
+First, each Health Region now covers a large geographical area that was served in the past by multiple LHINs.
+This in turn creates a complex socio-demographic landscape within each Health Region that is likely to be different in each case due the different levels of rurality and the proportion of underrepresented groups in each Health Region.
+
+Second, it is likely that a future pandemic or public health emergency will impact the province in the years ahead.
+Although broadly speaking, PHO and OH have been able to collaborate on the COVID-19 vaccination rollout, future events will require a unified response from both the public health arm of the government (represented by PHO), and the healthcare system agency (OH).
+So far, the COVID-19 pandemic has as a stress test for both agencies, showing the necessity from decision makers to clearly define how PHO and OH will interact in the future in order to ensure that in the future, Ontarians can benefit by virtue of a robust collaboration between both agencies.
+The evidence collected in this study shows that healthcare disparities experienced by underrepresented individuals in the province are the result of complex relationships between infrastructure, resources, occupations, geography, and socio-economic factors, among others.
+Therefore, this complexity presents an ongoing challenge for the healthcare system of Ontario, which will need to adapt and continue to aim for an integrative approach that can serve those that are most vulnerable.
+
+<!-- So far, the evidence collected during the COVID-19 pandemic indicates that differences in vaccination uptake are associated to a lack of infrastructure and resources that can adequately support personalized care to marginalized individuals.
 In the near future, health decision-makers will need to consider the implementation of policies that are focused on addressing this problematic.
+-->
 
-Moreover, the recent nature in the adaption of the Health Region poses a challenge for researchers in the acquisition of data and information that can be used to analyze the performance of the new model.
+Finally, the recent nature in the adaption of the Health Region poses a challenge for researchers in the acquisition of data and information that can be used to analyze the performance of the new model.
 From one side, the Health Regions have not been incorporated as part of Census data (LHINs were considered before in the Census), and this impact the amount and level of detail of available information.

### 2023-07-07 manuscript_v6/main.tex
updated discussion and conclusion, acknowlegements

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, frame hidden, boxrule=0pt, sharp corners, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 our knowledge, there are no studies that analyze differences in
-vaccination uptake at the provincial level. Such need is specially
+vaccination uptake at the intra-provincial level. Such need is specially
 important in the context of Canada's pandemic response goals, which have
@@
 COVID-19 Task Force under ethical guidance from the University of
-Toronto (under protocol 00043317), and which ran between September 30th,
-2021 and January 17th,2022.
+Toronto (under Research Ethics Board approval \#40999), and which ran
+between September 30th, 2021 and January 17th,2022.
 
@@
 
-The survey allowed users to enter their socio-economic information , and
-asked information on vaccination status by using the question ``Have you
-received the first dose of the COVID vaccine?'', with possible answers
-``yes'' and ``no'' (Table~\ref{tbl-descriptive-stats}). Additionally,
-the survey automatically collected the geographical location of the
-respondent (using the nearest municipality, as shown in
+Survey users entered their socio-economic information (age, income, and
+racial/ethnic group), and asked information on vaccination status by
+using the question ``Have you received the first dose of the COVID
+vaccine?'', with possible answers ``yes'' and ``no''
+(Table~\ref{tbl-descriptive-stats}). Of notice, because the
+racial/ethnic categories of the survey did not match exactly the
+categories from the 2016 Census, we used a combination of sources to
+re-group certain categories in order to obtain estimates that could be
+used to correct the data. Further details can be found in the Appendix.
+
+Additionally, the survey automatically collected the geographical
+location of the respondent (using the nearest municipality, as shown in
 Figure~\ref{fig-map}), and the date of access to the survey. The
 original dataset contained 39,029 observations with a response rate of
-16.25\% (defined as observations that had complete answers), due to the
+16.25\% (observations where users had complete answers), due to the
 survey design that allowed respondents to exit at any time and deployed
-the questions randomly. We selected 6,343 observations that were
-labelled as ``complete'' in the dataset and that corresponded to those
-that had answers for all covariates considered in our analysis. Later,
-we matched the city of each observation with its corresponding LHIN and
-Health Region, and removed observations from areas with low
-representation (254 observations corresponding to the North West and
-North East Health Regions). After all the preliminary analyses, the
-total number of observations used for analysis was 6,236 and included
-the East, Central, Toronto, and West Health Regions covering between
-October 1st,2021 and January 17, 2022. The original dataset, clean
-dataset, and details on the data cleaning process are described in
-detail in the \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub
-repository} for this paper.
+the questions randomly. We selected these observations with complete
+answers (6,343 observations) for our analysis. Later, we matched the
+city of each observation with its corresponding LHIN and Health Region,
+and removed observations from areas with low representation (254
+observations corresponding to the North West and North East Health
+Regions). After all the preliminary analyses, the total number of
+observations used for analysis was 6,236 and included the East, Central,
+Toronto, and West Health Regions covering between October 1st,2021 and
+January 17, 2022. The original dataset, clean dataset, and details on
+the data cleaning process are described in detail in the
+\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
+for this paper.
 
@@
 Because we identified differences in representativity between the survey
-data and the estimates from the Census, we used an iterative
-proportional fitting procedure
+data and the estimates from the 2016 Census for the variables considered
+in the analysis, and used used an iterative proportional fitting
+procedure
 (\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{48}}
@@
 6,236 observations, from which 24.8\% (1,547) corresponded to
-individuals that reported not having received the first dose of the
-vaccine. Vaccination rates ranged between 71-79\% across all household
-income brackets, age groups, Health Regions, and the months considered
-in the survey. However, the highest vaccination rates in each category
-were reported by individuals in the highest income bracket (79\%), those
-between 16 and 34 years of age (77\%), individuals that lived in the
-East Health Region (77\%), and during January of 2022 (78\%). Between
-racial/ethnic groups, the highest vaccination rate was reported by
-White/Caucasian individuals (84\%), against vaccination rates between
-63-66\% reported in the case of Arab/Middle Eastern, Black, Indigenous,
-Latin American individuals, and those that reported belonging to
-``Other'' racial groups, which included Southeast Asian, Filipino, West
-Asian, and minorities not identified elsewhere.
+individuals that reported not having initiated a COVID-19 vaccine
+primary series (in other words, not having received the first dose of
+the vaccine). The rate for the first dose of the vaccine ranged between
+71-79\% across all household income brackets, age groups, Health
+Regions, and the months considered in the survey. However, the highest
+rate for the uptake of the first dose of the vaccine were reported by
+individuals in the highest income bracket (79\%), those between 16 and
+34 years of age (77\%), individuals that lived in the East Health Region
+(77\%), and during January of 2022 (78\%). Between racial/ethnic groups,
+White/Caucasian individuals reported the highest uptake of the first
+dose of the vaccine (84\%), against values that ranged between 63 and
+66\% in the case of Arab/Middle Eastern, Black, Indigenous, Latin
+American individuals, and those that belonged to the ``other'' racial
+group category (which included Southeast Asian, Filipino, West Asian,
+and minorities not identified elsewhere percentages ).
 
@@
 \textsuperscript{\textit{2}}Pearson's Chi-squared test\\
-\textsuperscript{\textit{3}}Southeast Asian, Filipino, West Asian,
-and minorities not identified elsewhere according to the Census.\\
+\textsuperscript{\textit{3}}This category included Southeast Asian, Filipino, and West Asian individuals,
+and minorities not identified elsewhere according to the 2016 Census.\\
 \end{minipage}
@@
 
-In this study, we hypothesized that differences in COVID-19 vaccination
-uptake were present between the Health Regions between late 2021 and
-early 2022. Our goal was to determine which socio-demographic groups
-could be impacted by these disparities in order to provide
-decision-makers with information that could be used to develop policies
-focused on reducing or eliminating these differences and ensuring that
-the Health Region model is able to fulfill its mission of improving
-health access for all Ontarians.
-
-Our results show that indeed, there were differences in vaccination odds
-across Ontario in certain socio-demographic groups. Specifically, those
-who identified as Arab/Middle Eastern, Black, Latin American, having
-mixed racial or ethnic background, or that belonged to other groups not
-explicitly included in the survey (Southeast Asian, Filipino, West
-Asian, and minority groups not identified elsewhere) had vaccination
-odds that were between a third and a half of that of individuals that
-identified as White or Caucasian (Figure~\ref{fig-models}). These
-results are consistent with previous studies that have shown lower
-vaccination rates in individuals with the same socio-demographic
-characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{54}}.
+The COVID-19 pandemic has presented many significant challenges to the
+healthcare systems of the Canadian
+provinces\textsuperscript{\protect\hyperlink{ref-alami2021}{54}}. In the
+case of Ontario, the recent adoption of the Health Region model at the
+onset of the pandemic signified an additional level of complexity
+because although PHO (through the 34 Public Health Units of the
+province) was the lead in vaccine distribution, in reality the vaccine
+rollout required active collaboration between PHO and OH in multiple
+instances\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}}.
+
+In this study, we hypothesized that there were differences in the uptake
+of the first dose of the COVID-19 vaccine between individuals living in
+the different Health Regions of Ontario during late 2021 and early 2022.
+Using the Health Regions as the base of our analysis was advantageous,
+as these new Health Regions match closely the geographical boundaries of
+the Health Regions that have been historically used to group Public
+Health Units in the
+province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{55},\protect\hyperlink{ref-ohesi}{56}},
+thus providing information with relevance at both the public health and
+healthcare system level. Therefore, this analysis could provide an
+overall assessment of intra-provincial disparities that might need to be
+addressed moving forward by decision-makers, in order to ensure that in
+the future, and specially in the event of a future public health
+emergency, OH and PHO will able to collaboratively work in an efficient
+way and thus ensure that the population of the province benefits in the
+long term of the newly implemented Health Region model.
+
+Our results show that indeed, there were differences in the uptake of
+the first dose of the vaccine across Ontario in certain
+socio-demographic groups. Specifically, those who identified as
+Arab/Middle Eastern, Black, Latin American, having mixed racial or
+ethnic background, or that belonged to other groups not explicitly
+included in the survey (Southeast Asian, Filipino, West Asian, and
+minority groups not identified elsewhere) had vaccination odds that were
+between a third and a half of that of individuals that identified as
+White or Caucasian (Figure~\ref{fig-models}). These results are
+consistent with previous studies that have shown lower vaccination rates
+in individuals with the same socio-demographic
+characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{57}}.
 
@@
 in health
-settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{47},\protect\hyperlink{ref-basta2022}{55}--\protect\hyperlink{ref-cnat2023}{57}}.
+settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{47},\protect\hyperlink{ref-basta2022}{58}--\protect\hyperlink{ref-cnat2023}{60}}.
 However, it has been shown that structural barriers also play an
@@
 vaccination
-sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{58}}. In this
+sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{61}}. In this
 regard, it is interesting to note that vaccination venues were scarce in
@@
 Toronto and other regions of Ontario around the time covered by the
-survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{59}},
+survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{62}},
 and that pharmacies in the Peel region (an area identified as a
@@
 households) could not keep up with vaccine
-demand\textsuperscript{\protect\hyperlink{ref-gill2022}{60}}. This
+demand\textsuperscript{\protect\hyperlink{ref-gill2022}{63}}. This
 suggests that the observed differences are associated with disparities
@@
 
-This result is likely reflects in part the fact that individuals in
+This result likely reflects in part the fact that individuals in
 underrepresented groups tend to perform occupations that have been
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{61},\protect\hyperlink{ref-ct2021}{62}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{64},\protect\hyperlink{ref-ct2021}{65}},
 which include workers in the areas of grocery stores, gas stations,
@@
 which an income within the significant brackets identified in the
-analysis is to be expected. In Ontario, these workers had priority for
-COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{63}}; and
-there is evidence of interventions by vaccination staff in certain parts
-of the province to encourage vaccination uptake by these
-individuals\textsuperscript{\protect\hyperlink{ref-gill2022}{60}}. These
+analysis is to be expected. From one side, individuals in essential
+occupations in the province experienced higher rates of morbidity and
+mortality during the first year of the
+pandemic\textsuperscript{\protect\hyperlink{ref-rao2021}{66}}, but later
+on, they had priority for COVID-19
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{67}}.
+Additionally, it is known that vaccination uptake in these individuals
+was encouraged by vaccination staff in certain parts of the
+province\textsuperscript{\protect\hyperlink{ref-gill2022}{63}}. These
 facts, combined with evidence of increased trends in vaccination in this
 group
-elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{64}},
-suggest that the type of occupation of individuals in underrepresented
-groups played an important role in increasing the odds of vaccination.
+elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{68}},
+suggest that in Ontario, the type of occupation of individuals in
+underrepresented groups (which might have also affected their decision
+to get a vaccine based on their knowledge of increased risk), played an
+important role in the higher the odds of vaccination observed in these
+individuals.
 
@@
 pharmacist
-availability\textsuperscript{\protect\hyperlink{ref-shah2019}{65}--\protect\hyperlink{ref-timony2022}{67}}.
+availability\textsuperscript{\protect\hyperlink{ref-shah2019}{69}--\protect\hyperlink{ref-timony2022}{71}}.
 
@@
 to an estimated 100,000 Indigenous individuals living in the
-area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{68}}. Lack of
+area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{72}}. Lack of
 access to personalized healthcare affects individuals that may mistrust
@@
 vaccination rates among these
-groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{69},\protect\hyperlink{ref-eissa2021}{70}}.
+groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{73},\protect\hyperlink{ref-eissa2021}{74}}.
 Taken together, this suggests that healthcare disparities specific to
@@
 underrepresented groups
-elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{71}}.
+elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{75}}.
 
@@
 and corrected the data using population-wide information from the
-Census. More granular corrections would be needed to obtain more
-accurate estimates. For example, our analysis identified higher odds of
-vaccination in the Central and West Health Regions, but in this case
-these differences are likely to be driven by the proportion of
+Census. However, more granular corrections would be needed to obtain
+more accurate estimates: For example, our analysis identified higher
+odds of vaccination in the Central and West Health Regions, but in this
+case these differences are likely to be driven by the proportion of
 White/Caucasian individuals, who had higher vaccination rates than other
 racial groups. Correcting for each racial/ethnic group in each Health
-Region can provide a more accurate estimation of region-wide vaccination
-rates but unfortunately, at the moment this correction cannot be
-implemented as such stratification is has not been implemented in the
-Census.
+Region would provide a more accurate estimation of region-wide
+vaccination rates but unfortunately, presently this correction cannot be
+implemented as such stratification has not been implemented in the data
+that can be obtained from the Census.
 
@@
 Ontario. However, these areas have the highest proportion of Indigenous
-inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{68}}.
+inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{72}}.
 In the context of personalized care, there is a need to collect data
@@
 vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{72}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{76}}, we
 believe that our data was able to provide a valid sample of vaccination
@@
 October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{73}});
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{77}});
 however, the short time window constitutes essentially a ``snapshot''
@@
 the bureaucratic complexity and health disparities identified under the
-LHIN model. However, there are currently multiple challenges that need
-to be addressed to ensure that the new model can improve healthcare for
-the inhabitants of the province. First, the fact that each Health Region
-now covers a large geographical area that was served by multiple LHINs
-in the past creates a complex socio-demographic landscape that is
-different in each case due the different levels of rurality and
-representation of equity-deserving groups that are now within each
-Health Region. So far, the evidence collected during the COVID-19
-pandemic indicates that differences in vaccination uptake are associated
-to a lack of infrastructure and resources that can adequately support
-personalized care to marginalized individuals. In the near future,
-health decision-makers will need to consider the implementation of
-policies that are focused on addressing this problematic.
-
-Moreover, the recent nature in the adaption of the Health Region poses a
+LHIN model. However, there are certain challenges that this new model
+will need to overcome in order to work syngergistically with the rest of
+the healthcare system of the province. First, each Health Region now
+covers a large geographical area that was served in the past by multiple
+LHINs. This in turn creates a complex socio-demographic landscape within
+each Health Region that is likely to be different in each case due the
+different levels of rurality and the proportion of underrepresented
+groups in each Health Region.
+
+Second, it is likely that a future pandemic or public health emergency
+will impact the province in the years ahead. Although broadly speaking,
+PHO and OH have been able to collaborate on the COVID-19 vaccination
+rollout, future events will require a unified response from both the
+public health arm of the government (represented by PHO), and the
+healthcare system agency (OH). So far, the COVID-19 pandemic has as a
+stress test for both agencies, showing the necessity from decision
+makers to clearly define how PHO and OH will interact in the future in
+order to ensure that in the future, Ontarians can benefit by virtue of a
+robust collaboration between both agencies. The evidence collected in
+this study shows that healthcare disparities experienced by
+underrepresented individuals in the province are the result of complex
+relationships between infrastructure, resources, occupations, geography,
+and socio-economic factors, among others. Therefore, this complexity
+presents an ongoing challenge for the healthcare system of Ontario,
+which will need to adapt and continue to aim for an integrative approach
+that can serve those that are most vulnerable.
+
+Finally, the recent nature in the adaption of the Health Region poses a
 challenge for researchers in the acquisition of data and information
@@
 
+The authors thank Dr.~Sarah Wilson, Medical Epidemiologist with Public
+Health Ontario, for her valuable comments and feedback.
+
 \hypertarget{conflicts-of-interest}{%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-alami2021}{}}%
+\CSLLeftMargin{54. }%
+\CSLRightInline{Alami H, Lehoux P, Fleet R, et al. How can health
+systems better prepare for the next pandemic? Lessons learned from the
+management of {COVID}-19 in quebec (canada). \emph{Frontiers in Public
+Health}. 2021;9.
+doi:\href{https://doi.org/10.3389/fpubh.2021.671833}{10.3389/fpubh.2021.671833}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bolotin2018}{}}%
+\CSLLeftMargin{55. }%
+\CSLRightInline{Bolotin JJAG Shelly AND Feld. Population-based estimate
+of hepatitis c virus prevalence in ontario, canada. \emph{PLOS ONE}.
+2018;13(1):1-10.
+doi:\href{https://doi.org/10.1371/journal.pone.0191184}{10.1371/journal.pone.0191184}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ohesi}{}}%
+\CSLLeftMargin{56. }%
+\CSLRightInline{Ontario HIV Epidemiology and Surveillance Initiative.
+{T}echnical {N}otes: {H}ealth regions - {O}{H}{E}{S}{I} --- ohesi.ca.
+Accessed July 7, 2023.
+\url{https://www.ohesi.ca/technical-notes/health-regions/}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-basta2022}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{Basta NE, Sohel N, Sulis G, et al. Factors associated
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2023}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Cénat JM, Farahi SMMM, Bakombo SM, et al. Vaccine
@@
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-iveniuk2021}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Iveniuk J, Leon S. \emph{Uneven Recovery: Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gill2022}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Gill M, Datta D, Gregory P, Austin Z. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-rao2021}{}}%
+\CSLLeftMargin{66. }%
+\CSLRightInline{Rao A, Ma H, Moloney G, et al. A disproportionate
+epidemic: {COVID}-19 cases and deaths among essential workers in
+toronto, canada. \emph{Annals of Epidemiology}. 2021;63:63-67.
+doi:\href{https://doi.org/10.1016/j.annepidem.2021.07.010}{10.1016/j.annepidem.2021.07.010}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021b}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{Nguyen KH, Yankey D, Coy KC, et al. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shah2019}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Shah TI, Clark AF, Seabrook JA, Sibbald S, Gilliland JA.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-crighton2015}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{Crighton EJ, Ragetlie R, Luo J, To T, Gershon A. A
@@
 \leavevmode\vadjust pre{\hypertarget{ref-timony2022}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{71. }%
 \CSLRightInline{Timony P, Houle SKD, Gauthier A, Waite NM. Geographic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{72. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
-\CSLLeftMargin{69. }%
+\CSLLeftMargin{73. }%
 \CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
@@
 \leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
-\CSLLeftMargin{70. }%
+\CSLLeftMargin{74. }%
 \CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
-\CSLLeftMargin{71. }%
+\CSLLeftMargin{75. }%
 \CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{72. }%
+\CSLLeftMargin{76. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{73. }%
+\CSLLeftMargin{77. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,

### 2023-07-10 manuscript_v6/02-methods.qmd
updated files, sent to Overleaf

@@
 
-Briefly, the survey was deployed using random domain intercept technology, a methodology where internet users whose device meta-data indicated their presence in the province of Ontario had a random chance of being redirected to the survey after they had clicked on a registered but commercially inactive web link, or typed in a web address for a site that was dormant but that was temporarily managed by the company that administered the survey (RIWI Corp) [@sargent2022].
+Briefly, the survey was deployed using Random Domain Intercept Technology (RDIT), a methodology for internet surveys developed by the company commissioned to run the survey (RIWI Corp., Toronto, Canada), and that has been used in the area of public health research to examine trends in vaccination rates[@sargent2022], ratings of care quality[@kim2021], and perceptions of vaccine efficacy[@seeman2010].
+In the case of the survey used for this study, internet users whose device meta-data indicated their presence in the province of Ontario had a random chance of being redirected to the survey after they had clicked on a registered but commercially inactive web link, or after they had typed in a web address for a site that was dormant but that was temporarily managed by RIWI.
 Users then decided whether to anonymously participate in the survey, and those that participated were able to exit the survey at any time.

### 2023-07-10 manuscript_v6/appendix.tex
updated files, sent to Overleaf

--- 2023-05-26 manuscript_v5/appendix.tex
@@
 \usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \usepackage{caption}
-\usepackage{longtable}
 %start sections and page numbers with A
@@
 Montréal, Canada}
-\affil[*]{Corresponding author, \url{bouchra.nasri@umontreal.ca}}
+\affil[*]{Corresponding author,
+\href{mailto:bouchra.nasri@umontreal.ca}{bouchra.nasri@umontreal.ca}}
 
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 

### 2023-07-10 manuscript_v6/main.tex
updated files, sent to Overleaf

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-Briefly, the survey was deployed using random domain intercept
-technology, a methodology where internet users whose device meta-data
-indicated their presence in the province of Ontario had a random chance
-of being redirected to the survey after they had clicked on a registered
-but commercially inactive web link, or typed in a web address for a site
-that was dormant but that was temporarily managed by the company that
-administered the survey (RIWI
-Corp)\textsuperscript{\protect\hyperlink{ref-sargent2022}{44}}. Users
-then decided whether to anonymously participate in the survey, and those
-that participated were able to exit the survey at any time. After the
-survey closed, regardless if it was complete or incomplete, access was
-denied to any further users with the same internet protocol address
-(IP), effectively allowing each user only one opportunity to participate
-in the survey. Users who indicated they were under the age of 16 were
-exited from the survey without creating a record. Finally, the personal
-identifier information from each user that participated in the survey
-was automatically scrubbed and replaced by a unique ID.
+Briefly, the survey was deployed using Random Domain Intercept
+Technology (RDIT), a methodology for internet surveys developed by the
+company commissioned to run the survey (RIWI Corp., Toronto, Canada),
+and that has been used in the area of public health research to examine
+trends in vaccination
+rates\textsuperscript{\protect\hyperlink{ref-sargent2022}{44}}, ratings
+of care quality\textsuperscript{\protect\hyperlink{ref-kim2021}{45}},
+and perceptions of vaccine
+efficacy\textsuperscript{\protect\hyperlink{ref-seeman2010}{46}}. In the
+case of the survey used for this study, internet users whose device
+meta-data indicated their presence in the province of Ontario had a
+random chance of being redirected to the survey after they had clicked
+on a registered but commercially inactive web link, or after they had
+typed in a web address for a site that was dormant but that was
+temporarily managed by RIWI. Users then decided whether to anonymously
+participate in the survey, and those that participated were able to exit
+the survey at any time. After the survey closed, regardless if it was
+complete or incomplete, access was denied to any further users with the
+same internet protocol address (IP), effectively allowing each user only
+one opportunity to participate in the survey. Users who indicated they
+were under the age of 16 were exited from the survey without creating a
+record. Finally, the personal identifier information from each user that
+participated in the survey was automatically scrubbed and replaced by a
+unique ID.
 
@@
 of intent of vaccination and vaccination
-status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{45}--\protect\hyperlink{ref-cnat2022a}{47}}.
+status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{47}--\protect\hyperlink{ref-cnat2022a}{49}}.
 Because we identified differences in representativity between the survey
@@
 procedure
-(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{48}}
+(\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{50}}
 to correct the data using data from the Census and Health Region
@@
 Appendix. All analyses were conducted in R 4.2.2 using the packages
-\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{49}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{50}},
-\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{51}},
-\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{52}},
+\texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{51}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{52}},
+\texttt{quarto}\textsuperscript{\protect\hyperlink{ref-quarto}{53}},
+\texttt{modelsummary}\textsuperscript{\protect\hyperlink{ref-modelsummary}{54}},
 and
-\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{53}}.
+\texttt{gtsummary}\textsuperscript{\protect\hyperlink{ref-gtsummary}{55}}.
 
@@
 healthcare systems of the Canadian
-provinces\textsuperscript{\protect\hyperlink{ref-alami2021}{54}}. In the
+provinces\textsuperscript{\protect\hyperlink{ref-alami2021}{56}}. In the
 case of Ontario, the recent adoption of the Health Region model at the
@@
 Health Units in the
-province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{55},\protect\hyperlink{ref-ohesi}{56}},
+province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{57},\protect\hyperlink{ref-ohesi}{58}},
 thus providing information with relevance at both the public health and
@@
 in individuals with the same socio-demographic
-characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{57}}.
+characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{59}}.
 
@@
 in health
-settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{47},\protect\hyperlink{ref-basta2022}{58}--\protect\hyperlink{ref-cnat2023}{60}}.
+settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{49},\protect\hyperlink{ref-basta2022}{60}--\protect\hyperlink{ref-cnat2023}{62}}.
 However, it has been shown that structural barriers also play an
@@
 vaccination
-sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{61}}. In this
+sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{63}}. In this
 regard, it is interesting to note that vaccination venues were scarce in
@@
 Toronto and other regions of Ontario around the time covered by the
-survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{62}},
+survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{64}},
 and that pharmacies in the Peel region (an area identified as a
@@
 households) could not keep up with vaccine
-demand\textsuperscript{\protect\hyperlink{ref-gill2022}{63}}. This
+demand\textsuperscript{\protect\hyperlink{ref-gill2022}{65}}. This
 suggests that the observed differences are associated with disparities
@@
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{64},\protect\hyperlink{ref-ct2021}{65}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{66},\protect\hyperlink{ref-ct2021}{67}},
 which include workers in the areas of grocery stores, gas stations,
@@
 mortality during the first year of the
-pandemic\textsuperscript{\protect\hyperlink{ref-rao2021}{66}}, but later
+pandemic\textsuperscript{\protect\hyperlink{ref-rao2021}{68}}, but later
 on, they had priority for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{67}}.
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{69}}.
 Additionally, it is known that vaccination uptake in these individuals
 was encouraged by vaccination staff in certain parts of the
-province\textsuperscript{\protect\hyperlink{ref-gill2022}{63}}. These
+province\textsuperscript{\protect\hyperlink{ref-gill2022}{65}}. These
 facts, combined with evidence of increased trends in vaccination in this
 group
-elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{68}},
+elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{70}},
 suggest that in Ontario, the type of occupation of individuals in
@@
 pharmacist
-availability\textsuperscript{\protect\hyperlink{ref-shah2019}{69}--\protect\hyperlink{ref-timony2022}{71}}.
+availability\textsuperscript{\protect\hyperlink{ref-shah2019}{71}--\protect\hyperlink{ref-timony2022}{73}}.
 
@@
 to an estimated 100,000 Indigenous individuals living in the
-area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{72}}. Lack of
+area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{74}}. Lack of
 access to personalized healthcare affects individuals that may mistrust
@@
 vaccination rates among these
-groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{73},\protect\hyperlink{ref-eissa2021}{74}}.
+groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{75},\protect\hyperlink{ref-eissa2021}{76}}.
 Taken together, this suggests that healthcare disparities specific to
@@
 underrepresented groups
-elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{75}}.
+elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{77}}.
 
@@
 Ontario. However, these areas have the highest proportion of Indigenous
-inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{72}}.
+inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{74}}.
 In the context of personalized care, there is a need to collect data
@@
 vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{76}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{78}}, we
 believe that our data was able to provide a valid sample of vaccination
@@
 October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{77}});
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{79}});
 however, the short time window constitutes essentially a ``snapshot''
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kim2021}{}}%
+\CSLLeftMargin{45. }%
+\CSLRightInline{Kim JH, Bell GA, Ratcliffe HL, et al. Predictors of
+patient-reported quality of care in low- and middle-income countries: A
+four-country survey of person-centered care. \emph{International Journal
+for Quality in Health Care}. 2021;33(3).
+doi:\href{https://doi.org/10.1093/intqhc/mzab110}{10.1093/intqhc/mzab110}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-seeman2010}{}}%
+\CSLLeftMargin{46. }%
+\CSLRightInline{Seeman N, Ing A, Rizo C. Assessing and responding in
+real time to online anti-vaccine sentiment during a flu pandemic.
+\emph{Healthcare Quarterly}. 2010;13(sp):8-15.
+doi:\href{https://doi.org/10.12927/hcq.2010.21923}{10.12927/hcq.2010.21923}}
+
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2022}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Nguyen KH, Anneser E, Toppo A, Allen JD, Parott JS,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shih2021}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Shih SF, Wagner AL, Masters NB, Prosser LA, Lu Y,
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022a}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Farahi SMMM, et al. Prevalence
@@
 \leavevmode\vadjust pre{\hypertarget{ref-deming1940}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{50. }%
 \CSLRightInline{Deming WE, Stephan FF. On a least squares adjustment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-lumley2011}{}}%
-\CSLLeftMargin{49. }%
+\CSLLeftMargin{51. }%
 \CSLRightInline{Lumley T. \emph{Complex Surveys}. John Wiley \& Sons;
@@
 \leavevmode\vadjust pre{\hypertarget{ref-wickham2019}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Wickham H, Averick M, Bryan J, et al. Welcome to the
@@
 \leavevmode\vadjust pre{\hypertarget{ref-quarto}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Allaire J. \emph{Quarto: R Interface to 'Quarto'
@@
 \leavevmode\vadjust pre{\hypertarget{ref-modelsummary}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Arel-Bundock V. {modelsummary}: Data and model summaries
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gtsummary}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Sjoberg DD, Whiting K, Curry M, Lavery JA, Larmarange J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-alami2021}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{Alami H, Lehoux P, Fleet R, et al. How can health
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bolotin2018}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Bolotin JJAG Shelly AND Feld. Population-based estimate
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ohesi}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{Ontario HIV Epidemiology and Surveillance Initiative.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-basta2022}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Basta NE, Sohel N, Sulis G, et al. Factors associated
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2023}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Cénat JM, Farahi SMMM, Bakombo SM, et al. Vaccine
@@
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-iveniuk2021}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Iveniuk J, Leon S. \emph{Uneven Recovery: Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gill2022}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Gill M, Datta D, Gregory P, Austin Z. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-rao2021}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{Rao A, Ma H, Moloney G, et al. A disproportionate
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021b}{}}%
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{Nguyen KH, Yankey D, Coy KC, et al. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shah2019}{}}%
-\CSLLeftMargin{69. }%
+\CSLLeftMargin{71. }%
 \CSLRightInline{Shah TI, Clark AF, Seabrook JA, Sibbald S, Gilliland JA.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-crighton2015}{}}%
-\CSLLeftMargin{70. }%
+\CSLLeftMargin{72. }%
 \CSLRightInline{Crighton EJ, Ragetlie R, Luo J, To T, Gershon A. A
@@
 \leavevmode\vadjust pre{\hypertarget{ref-timony2022}{}}%
-\CSLLeftMargin{71. }%
+\CSLLeftMargin{73. }%
 \CSLRightInline{Timony P, Houle SKD, Gauthier A, Waite NM. Geographic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{72. }%
+\CSLLeftMargin{74. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
-\CSLLeftMargin{73. }%
+\CSLLeftMargin{75. }%
 \CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
@@
 \leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
-\CSLLeftMargin{74. }%
+\CSLLeftMargin{76. }%
 \CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
-\CSLLeftMargin{75. }%
+\CSLLeftMargin{77. }%
 \CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{76. }%
+\CSLLeftMargin{78. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{77. }%
+\CSLLeftMargin{79. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,

### 2023-07-12 manuscript_v6/01-background.qmd
implemented changes suggested by BN

--- 2023-07-02 manuscript_v6/01-background.qmd
@@
 These challenges included the rapid emergence of new virus variants [@li2021], the waning of vaccine protection [@dolgin2021], inequality in vaccine access between high-income and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population [@willis2021;@skirrow2022].
-In particular, it is well established that lower vaccination uptake has been observed in individuals within certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
-
-<!-- pharmaceutical interventions were ultimately not determinant to bring the pandemic to a rapid end.  -->
-
-<!-- This contrast with previous vaccination experiences, such as in the case of smallpox, where the implementation of a global vaccination program was crucial to control this disease[@kayser2021].
--->
-
-<!-- The failure of vaccines to stop the pandemic is a complex issue resulting from a combination of factors, among which are inadequate public health measures and the emergence of new variants [@li2021], inequality in vaccine access between high-income and low-income countries [@gerretsen2021;@tamey2022], vaccine hesitancy [@nafilyan2021], and differences in vaccine uptake across the population[@willis2021;@skirrow2022].
-In particular, it is well established that lower vaccination uptake has been observed in individuals within certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
--->
+More specifically, it is well established that lower vaccination uptake has been observed in individuals within certain underrepresented groups (e.g., Black, Asian, or Indigenous) as well as individuals with socio-economic disadvantages [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021;@guay2022;@muhajarine2021;@hussain2022].
 
 Reasons given for this inequality in vaccination uptake have included medical mistrust due to systemic medical racism [@stoler2021;@mosby2021], mistrust in vaccines [@willis2021], and the existence of health misinformation and disinformation [@bogart2021;@mosby2021;@freeman2020].
-However, it is important also to consider that vaccination uptake can be influenced by geographical (spatial) factors.
-In this regard, differences in COVID-19 vaccination rates have been associated with varied regional attitudes towards vaccination [@malik2020], lack of vaccine access in areas where underrepresented groups reside[@bogoch2022], as well as an absence of prioritization of areas inhabited by vulnerable groups [@nguyen2021].
-Other studies have also shown heterogeneity in vaccine uptake within small governmental administrative units such as counties [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], indicating that accounting for geographical differences in vaccination can help predict patterns of booster uptake [@wood2022].
-However, there is a limited amount of studies that have analyzed geographical differences in vaccination in Canada.
-Existing studies in this area have focused on differences within certain cities, such as Toronto [@choi2021], or Montréal [@mckinnon2021]), or have explored differences between provinces [@guay2022], but to our knowledge, there are no studies that analyze  differences in vaccination uptake at the provincial level.
+However, it is important to consider that vaccination uptake can also vary geographically across the administrative or political subdivisions within a country.
+For example, differences in COVID-19 vaccination rates have been associated with differences in attitudes towards vaccination in administrative areas in New York and Chicago[@malik2020], lack of vaccine access in areas where underrepresented groups reside in Ontario[@bogoch2022], as well as an absence of prioritization of areas inhabited by vulnerable groups in the southeastern region of the US[@nguyen2021].
+Other studies have also shown heterogeneity in vaccine uptake within counties in the US [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], indicating that accounting for geographical differences in vaccination can help predict patterns of booster uptake in Scottish communities [@wood2022].
+Therefore, the study of vaccination uptake across administrative of political areas can be useful to highlight differences that governments or public health agencies might need to address.
+However, in the case of Canada, there is a limited amount of studies that have analyzed geographical differences in vaccination.
+Existing studies in this area have focused on differences within certain cities, such as Toronto [@choi2021], or Montréal [@mckinnon2021]), or have explored differences between provinces-at-large [@guay2022], but to our knowledge, there are no studies that analyze differences in vaccination uptake at the intra-provincial level.
 Such need is specially important in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023].
 Analyzing differences in vaccination uptake within the provinces can aid to identify inequalities that might exist and that need to be addressed before the advent of another pandemic.
-
-<!-- However, in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
--->
-
-<!-- **Canada’s pandemic response goals, were to reduce morbidity/mortality and to reduce societal disruption from COVID-19**.   -->
-
-<!-- - How does this integrate with vaccination? We need to study vaccination because  -->
-<!-- -  -->
-
-<!-- - - -->
-
-<!-- and that consequently, can help identify disparities that need to be addressed within specific areas in each province. -->
-
-<!-- However, in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
--->
 
@@
 On the other hand, public health in Ontario is administered by Public Health Ontario (PHO), a government agency established in late 2007 and that is currently composed of 34 Public Health Units (PHUs) that cover the entire geography of the province.
-Although PHUS were commissioned with leading the distribution of the COVID-19 vaccine within their respective areas [@rotenberg2021], the vaccine rollout occurred with significant interaction between PHO and OH; in many instances both agencies had to work together to organize vaccination clinics, with personnel associated with OH also being actively involved when demand exceeded the capacity of PHO personnel, especially in rural areas [@ashcroft2023].
+Although PHUs were commissioned with leading the distribution of the COVID-19 vaccine within their respective areas [@rotenberg2021], the vaccine rollout occurred with significant interaction between PHO and OH; in many instances both agencies had to work together to organize vaccination clinics, with personnel associated with OH also being actively involved when demand exceeded the capacity of PHO personnel, especially in rural areas [@ashcroft2023].
 Indeed, based on the experience of COVID-19, there is an ongoing debate on how OH and PHO will interact in the future, and the challenges that this will entail for the healthcare system of the province [@dion2022].

### 2023-07-12 manuscript_v6/02-methods.qmd
implemented changes suggested by BN

@@
 Additionally, the survey automatically collected the geographical location of the respondent (using the nearest municipality,  as shown in @fig-map), and the date of access to the survey.
-The original dataset contained 39,029 observations with a response rate of 16.25% (observations where users had complete answers), due to the survey design that allowed respondents to exit at any time and deployed the questions randomly.
-We selected these  observations with complete answers (6,343 observations) for our analysis.
-Later, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions).
-After all the preliminary analyses, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering between October 1st,2021 and January 17, 2022.
+The original dataset contained 39,029 observations.
+However, the number of complete observations was much lower that the total number of observations due to the survey design, which allowed respondents to exit at any time and deployed the questions randomly.
+We selected the observations with complete answers (6,343 observations, or 16.25\% of the total) for our analysis.
+It should be noted that this response rate is similar to response rates observed in previous studies that have used the same type of survey instrument and that have ranged between 15\% and 22\%[@seeman2010;@kim2021].
+Next, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions).
+After all the preliminary analyses, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering between October 1st, 2021 and January 17, 2022.
 The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.

### 2023-07-12 manuscript_v6/03-results.qmd
implemented changes suggested by BN

@@
                  income_ord="Income (CAD)")
-  ) %>% add_p(test = everything () ~ "chisq.test")%>%
+  ) %>%
   bold_labels() %>%
@@
 #| fig-width: 15
-#| fig-height: 14
+#| fig-height: 17
 #| label: fig-models

### 2023-07-12 manuscript_v6/04-discussion.qmd
implemented changes suggested by BN

@@
 
-The COVID-19 pandemic has presented many significant challenges to the healthcare systems of the Canadian provinces [@alami2021].
+<!-- The COVID-19 pandemic has presented many significant challenges to the healthcare systems of the Canadian provinces [@alami2021].
 In the case of Ontario, the recent adoption of the Health Region model at the onset of the pandemic signified an additional level of complexity because although PHO (through the 34 Public Health Units of the province) was the lead in vaccine distribution, in reality the vaccine rollout required active collaboration between PHO and OH in multiple instances [@ashcroft2023].
+-->
 
 In this study, we hypothesized that there were differences in the uptake of the first dose of the COVID-19 vaccine between individuals living in the different Health Regions of Ontario during late 2021 and early 2022.
-Using the Health Regions as the base of our analysis was advantageous, as these new Health Regions match closely the geographical boundaries of the Health Regions that have been historically used to group Public Health Units in the province [@bolotin2018;@ohesi], thus providing information with relevance at both the public health and healthcare system level.
-Therefore, this analysis could provide an overall assessment of intra-provincial disparities that might need to be addressed moving forward by decision-makers, in order to ensure that in the future, and specially in the event of a future public health emergency, OH and PHO will able to collaboratively work in an efficient way and thus ensure that the population of the province benefits in the long term of the newly implemented Health Region model.
-
-<!-- Specifically, our goal was to determine which socio-demographic groups could be impacted by these differences in vaccination rates in order to highlight challenges that will need to be addressed by OH to determine the exact nature of its future interactions with PHO in the event of a future public health emergency.
--->
-
-<!-- The complex relationships within the health system of Ontario, with the recent implementation of OH make necessary to determine  COVID-19 pandemic highlighted the need of an active collaboration between PHO and OH -->
-
-<!-- The recent implementation of the Health Region model  vaccination campaign in Ontario  -->
-
-<!-- Due to the recent implementation of the Health Region model in the province, and the Our goal was to determine which socio-demographic groups could be impacted by these disparities, in order to provide decision-makers with information that could be used to inform  -->
-
-<!-- develop policies focused on reducing or eliminating these differences and ensuring that the Health Region model is able to fulfill its mission of improving health access for all Ontarians. -->
-
-<!-- In this study, we hypothesized that differences in the uptake of the first dose of the COVID-19 vaccine were present between the Health Regions between late 2021 and early 2022.
-Our goal was to determine which socio-demographic groups could be impacted by these disparities in order to provide decision-makers with information that could be used to develop policies focused on reducing or eliminating these differences and ensuring that the Health Region model is able to fulfill its mission of improving health access for all Ontarians.
--->
+Using the Health Regions as the base of our analysis was advantageous, as these newly implemented administrative areas match closely the geographical boundaries of the Health Regions that have been historically used to group Public Health Units in the province [@bolotin2018;@ohesi].
+Thus, such analysis can provide information on vaccination rates with relevance at both the public health and healthcare system level.
+Moreover, such analysis could provide an overall assessment of intra-provincial disparities that might need to be addressed moving forward by decision-makers in order to ensure that in the event of a future public health emergency or pandemic, OH and PHO will able to work collaboratively in an efficient way having learned from the COVID-19 experience[@ashcroft2023], thus ensuring that the population of the province benefits in the long term from a robust health system.
 
@@
 Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals.
-For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated 100,000 Indigenous individuals living in the area [@ontariohealth].
+For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated population of 100,000 Indigenous individuals living in the area [@ontariohealth].
 Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada, as these rationales have been associated with observed lower vaccination rates among these groups [@smylie2022;@eissa2021].

### 2023-07-12 manuscript_v6/05-conclusion.qmd
implemented changes suggested by BN

@@
 Although broadly speaking, PHO and OH have been able to collaborate on the COVID-19 vaccination rollout, future events will require a unified response from both the public health arm of the government (represented by PHO), and the healthcare system agency (OH).
-So far, the COVID-19 pandemic has as a stress test for both agencies, showing the necessity from decision makers to clearly define how PHO and OH will interact in the future in order to ensure that in the future, Ontarians can benefit by virtue of a robust collaboration between both agencies.
+So far, the COVID-19 pandemic has served as a stress test for both agencies, showing the necessity from decision makers to clearly define how PHO and OH will interact in the future in order to ensure that Ontarians can benefit by virtue of a robust collaboration between both agencies.
 The evidence collected in this study shows that healthcare disparities experienced by underrepresented individuals in the province are the result of complex relationships between infrastructure, resources, occupations, geography, and socio-economic factors, among others.

### 2023-07-12 manuscript_v6/main.qmd
updated discussion and conclusion, acknowlegements
implemented changes suggested by BN

@@
 ---
-title: '**Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata**'
+title: '**Socio-demographic determinants of COVID-19 vaccine uptake in Ontario: Exploring differences across the Health Region model**'
 author:
@@
   - id: 3
-    name: Centre de recherche en santé publique, University of Montreal, Montréal, Canada
+    name: Centre de recherche en santé publique, University of Montréal, Montréal, Canada
   - id: \*
@@
 
+The authors thank Dr. Sarah Wilson, Medical Epidemiologist with Public Health Ontario, for her valuable comments and feedback.
+
 # Conflicts of Interest

### 2023-07-12 manuscript_v6/main.tex
implemented changes suggested by BN

@@
 \hypersetup{
-  pdftitle={Differences in COVID-19 vaccination in the province of Ontario across Health Regions and socio-economic strata},
+  pdftitle={Socio-demographic determinants of COVID-19 vaccine uptake in Ontario: Exploring differences across the Health Region model},
   pdfauthor={Ariel Mundo Ortiz; Bouchra Nasri},
@@
 
-\title{\textbf{Differences in COVID-19 vaccination in the province of
-Ontario across Health Regions and socio-economic strata}}
+\title{\textbf{Socio-demographic determinants of COVID-19 vaccine uptake
+in Ontario: Exploring differences across the Health Region model}}
 
@@
 Publique, University of Montreal, Montréal, Canada}
-\affil[3]{Centre de recherche en santé publique, University of Montreal,
+\affil[3]{Centre de recherche en santé publique, University of Montréal,
 Montréal, Canada}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi
 
@@
 population\textsuperscript{\protect\hyperlink{ref-willis2021}{15},\protect\hyperlink{ref-skirrow2022}{16}}.
-In particular, it is well established that lower vaccination uptake has
-been observed in individuals within certain underrepresented groups
+More specifically, it is well established that lower vaccination uptake
+has been observed in individuals within certain underrepresented groups
 (e.g., Black, Asian, or Indigenous) as well as individuals with
@@
 disinformation\textsuperscript{\protect\hyperlink{ref-mosby2021}{22}--\protect\hyperlink{ref-freeman2020}{24}}.
-However, it is important also to consider that vaccination uptake can be
-influenced by geographical (spatial) factors. In this regard,
-differences in COVID-19 vaccination rates have been associated with
-varied regional attitudes towards
-vaccination\textsuperscript{\protect\hyperlink{ref-malik2020}{25}}, lack
-of vaccine access in areas where underrepresented groups
-reside\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as well
-as an absence of prioritization of areas inhabited by vulnerable
-groups\textsuperscript{\protect\hyperlink{ref-nguyen2021}{26}}. Other
-studies have also shown heterogeneity in vaccine uptake within small
-governmental administrative units such as
-counties\textsuperscript{\protect\hyperlink{ref-mollalo2021}{27}--\protect\hyperlink{ref-bhuiyan2022}{30}},
+However, it is important to consider that vaccination uptake can also
+vary geographically across the administrative or political subdivisions
+within a country. For example, differences in COVID-19 vaccination rates
+have been associated with differences in attitudes towards vaccination
+in administrative areas in New York and
+Chicago\textsuperscript{\protect\hyperlink{ref-malik2020}{25}}, lack of
+vaccine access in areas where underrepresented groups reside in
+Ontario\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as well
+as an absence of prioritization of areas inhabited by vulnerable groups
+in the southeastern region of the
+US\textsuperscript{\protect\hyperlink{ref-nguyen2021}{26}}. Other
+studies have also shown heterogeneity in vaccine uptake within counties
+in the
+US\textsuperscript{\protect\hyperlink{ref-mollalo2021}{27}--\protect\hyperlink{ref-bhuiyan2022}{30}},
 indicating that accounting for geographical differences in vaccination
-can help predict patterns of booster
-uptake\textsuperscript{\protect\hyperlink{ref-wood2022}{31}}. However,
-there is a limited amount of studies that have analyzed geographical
-differences in vaccination in Canada. Existing studies in this area have
-focused on differences within certain cities, such as
+can help predict patterns of booster uptake in Scottish
+communities\textsuperscript{\protect\hyperlink{ref-wood2022}{31}}.
+Therefore, the study of vaccination uptake across administrative of
+political areas can be useful to highlight differences that governments
+or public health agencies might need to address. However, in the case of
+Canada, there is a limited amount of studies that have analyzed
+geographical differences in vaccination. Existing studies in this area
+have focused on differences within certain cities, such as
 Toronto\textsuperscript{\protect\hyperlink{ref-choi2021}{32}}, or
@@
 have explored differences between
-provinces\textsuperscript{\protect\hyperlink{ref-guay2022}{19}}, but to
-our knowledge, there are no studies that analyze differences in
+provinces-at-large\textsuperscript{\protect\hyperlink{ref-guay2022}{19}},
+but to our knowledge, there are no studies that analyze differences in
 vaccination uptake at the intra-provincial level. Such need is specially
@@
 that is currently composed of 34 Public Health Units (PHUs) that cover
-the entire geography of the province. Although PHUS were commissioned
+the entire geography of the province. Although PHUs were commissioned
 with leading the distribution of the COVID-19 vaccine within their
@@
 Figure~\ref{fig-map}), and the date of access to the survey. The
-original dataset contained 39,029 observations with a response rate of
-16.25\% (observations where users had complete answers), due to the
-survey design that allowed respondents to exit at any time and deployed
-the questions randomly. We selected these observations with complete
-answers (6,343 observations) for our analysis. Later, we matched the
-city of each observation with its corresponding LHIN and Health Region,
-and removed observations from areas with low representation (254
-observations corresponding to the North West and North East Health
-Regions). After all the preliminary analyses, the total number of
-observations used for analysis was 6,236 and included the East, Central,
-Toronto, and West Health Regions covering between October 1st,2021 and
-January 17, 2022. The original dataset, clean dataset, and details on
-the data cleaning process are described in detail in the
-\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
-for this paper.
+original dataset contained 39,029 observations. However, the number of
+complete observations was much lower that the total number of
+observations due to the survey design, which allowed respondents to exit
+at any time and deployed the questions randomly. We selected the
+observations with complete answers (6,343 observations, or 16.25\% of
+the total) for our analysis. It should be noted that this response rate
+is similar to response rates observed in previous studies that have used
+the same type of survey instrument and that have ranged between 15\% and
+22\%\textsuperscript{\protect\hyperlink{ref-kim2021}{45},\protect\hyperlink{ref-seeman2010}{46}}.
+Next, we matched the city of each observation with its corresponding
+LHIN and Health Region, and removed observations from areas with low
+representation (254 observations corresponding to the North West and
+North East Health Regions). After all the preliminary analyses, the
+total number of observations used for analysis was 6,236 and included
+the East, Central, Toronto, and West Health Regions covering between
+October 1st, 2021 and January 17, 2022. The original dataset, clean
+dataset, and details on the data cleaning process are described in
+detail in the \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub
+repository} for this paper.
 
@@
 \setlength{\LTpost}{0mm}
-\begin{longtable}{lccc}
+\begin{longtable}{lcc}
 \caption{\label{tbl-descriptive-stats}Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination
@@
 \toprule
-\textbf{Variable} & \textbf{no}, N = 1,547\textsuperscript{\textit{1}} & \textbf{yes}, N = 4,689\textsuperscript{\textit{1}} & \textbf{p-value}\textsuperscript{\textit{2}} \\ 
+\textbf{Variable} & \textbf{no}, N = 1,547\textsuperscript{\textit{1}} & \textbf{yes}, N = 4,689\textsuperscript{\textit{1}} \\ 
 \midrule
-Income (CAD) &  &  & <0.001 \\ 
-    60000 and above & 542 (21\%) & 1,996 (79\%) &  \\ 
-    25000-59999 & 347 (25\%) & 1,046 (75\%) &  \\ 
-    under 25000 & 658 (29\%) & 1,647 (71\%) &  \\ 
-Age Group &  &  & 0.002 \\ 
-    16-34 & 645 (23\%) & 2,117 (77\%) &  \\ 
-    35-54 & 411 (24\%) & 1,305 (76\%) &  \\ 
-    55 and over & 491 (28\%) & 1,267 (72\%) &  \\ 
-Health Region &  &  & 0.3 \\ 
-    Toronto & 593 (26\%) & 1,709 (74\%) &  \\ 
-    Central & 372 (26\%) & 1,083 (74\%) &  \\ 
-    East & 236 (23\%) & 783 (77\%) &  \\ 
-    West & 346 (24\%) & 1,114 (76\%) &  \\ 
-Month &  &  & <0.001 \\ 
-    October & 469 (27\%) & 1,263 (73\%) &  \\ 
-    November & 376 (28\%) & 980 (72\%) &  \\ 
-    December & 181 (24\%) & 565 (76\%) &  \\ 
-    January & 521 (22\%) & 1,881 (78\%) &  \\ 
-Race &  &  & <0.001 \\ 
-    White/Caucasian & 354 (16\%) & 1,871 (84\%) &  \\ 
-    Arab/Middle Eastern & 111 (34\%) & 220 (66\%) &  \\ 
-    Black & 159 (34\%) & 303 (66\%) &  \\ 
-    East Asian/Pacific Islander & 94 (19\%) & 404 (81\%) &  \\ 
-    Indigenous & 112 (37\%) & 194 (63\%) &  \\ 
-    Latin American & 99 (34\%) & 195 (66\%) &  \\ 
-    Mixed & 177 (30\%) & 411 (70\%) &  \\ 
-    Other\textsuperscript{\textit{3}} & 315 (34\%) & 606 (66\%) &  \\ 
-    South Asian & 126 (21\%) & 485 (79\%) &  \\ 
+Income (CAD) &  &  \\ 
+    60000 and above & 542 (21\%) & 1,996 (79\%) \\ 
+    25000-59999 & 347 (25\%) & 1,046 (75\%) \\ 
+    under 25000 & 658 (29\%) & 1,647 (71\%) \\ 
+Age Group &  &  \\ 
+    16-34 & 645 (23\%) & 2,117 (77\%) \\ 
+    35-54 & 411 (24\%) & 1,305 (76\%) \\ 
+    55 and over & 491 (28\%) & 1,267 (72\%) \\ 
+Health Region &  &  \\ 
+    Toronto & 593 (26\%) & 1,709 (74\%) \\ 
+    Central & 372 (26\%) & 1,083 (74\%) \\ 
+    East & 236 (23\%) & 783 (77\%) \\ 
+    West & 346 (24\%) & 1,114 (76\%) \\ 
+Month &  &  \\ 
+    October 2021 & 469 (27\%) & 1,263 (73\%) \\ 
+    November 2021 & 376 (28\%) & 980 (72\%) \\ 
+    December 2021 & 181 (24\%) & 565 (76\%) \\ 
+    January 2022 & 521 (22\%) & 1,881 (78\%) \\ 
+Race &  &  \\ 
+    White/Caucasian & 354 (16\%) & 1,871 (84\%) \\ 
+    Arab/Middle Eastern & 111 (34\%) & 220 (66\%) \\ 
+    Black & 159 (34\%) & 303 (66\%) \\ 
+    East Asian/Pacific Islander & 94 (19\%) & 404 (81\%) \\ 
+    Indigenous & 112 (37\%) & 194 (63\%) \\ 
+    Latin American & 99 (34\%) & 195 (66\%) \\ 
+    Mixed & 177 (30\%) & 411 (70\%) \\ 
+    Other\textsuperscript{\textit{2}} & 315 (34\%) & 606 (66\%) \\ 
+    South Asian & 126 (21\%) & 485 (79\%) \\ 
 \bottomrule
@@
 \textsuperscript{\textit{1}}n (\%)\\
-\textsuperscript{\textit{2}}Pearson's Chi-squared test\\
-\textsuperscript{\textit{3}}This category included Southeast Asian, Filipino, and West Asian individuals,
+\textsuperscript{\textit{2}}This category included Southeast Asian, Filipino, and West Asian individuals,
 and minorities not identified elsewhere according to the 2016 Census.\\
@@
 
-The COVID-19 pandemic has presented many significant challenges to the
-healthcare systems of the Canadian
-provinces\textsuperscript{\protect\hyperlink{ref-alami2021}{56}}. In the
-case of Ontario, the recent adoption of the Health Region model at the
-onset of the pandemic signified an additional level of complexity
-because although PHO (through the 34 Public Health Units of the
-province) was the lead in vaccine distribution, in reality the vaccine
-rollout required active collaboration between PHO and OH in multiple
-instances\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}}.
-
 In this study, we hypothesized that there were differences in the uptake
@@
 Using the Health Regions as the base of our analysis was advantageous,
-as these new Health Regions match closely the geographical boundaries of
-the Health Regions that have been historically used to group Public
-Health Units in the
-province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{57},\protect\hyperlink{ref-ohesi}{58}},
-thus providing information with relevance at both the public health and
-healthcare system level. Therefore, this analysis could provide an
-overall assessment of intra-provincial disparities that might need to be
-addressed moving forward by decision-makers, in order to ensure that in
-the future, and specially in the event of a future public health
-emergency, OH and PHO will able to collaboratively work in an efficient
-way and thus ensure that the population of the province benefits in the
-long term of the newly implemented Health Region model.
+as these newly implemented administrative areas match closely the
+geographical boundaries of the Health Regions that have been
+historically used to group Public Health Units in the
+province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{56},\protect\hyperlink{ref-ohesi}{57}}.
+Thus, such analysis can provide information on vaccination rates with
+relevance at both the public health and healthcare system level.
+Moreover, such analysis could provide an overall assessment of
+intra-provincial disparities that might need to be addressed moving
+forward by decision-makers in order to ensure that in the event of a
+future public health emergency or pandemic, OH and PHO will able to work
+collaboratively in an efficient way having learned from the COVID-19
+experience\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}},
+thus ensuring that the population of the province benefits in the long
+term from a robust health system.
 
@@
 in individuals with the same socio-demographic
-characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{59}}.
+characteristics\textsuperscript{\protect\hyperlink{ref-guay2022}{19}--\protect\hyperlink{ref-hussain2022}{21},\protect\hyperlink{ref-carter2022}{58}}.
 
@@
 in health
-settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{49},\protect\hyperlink{ref-basta2022}{60}--\protect\hyperlink{ref-cnat2023}{62}}.
+settings\textsuperscript{\protect\hyperlink{ref-cnat2022a}{49},\protect\hyperlink{ref-basta2022}{59}--\protect\hyperlink{ref-cnat2023}{61}}.
 However, it has been shown that structural barriers also play an
@@
 vaccination
-sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{63}}. In this
+sites\textsuperscript{\protect\hyperlink{ref-njoku2021}{62}}. In this
 regard, it is interesting to note that vaccination venues were scarce in
@@
 Toronto and other regions of Ontario around the time covered by the
-survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{64}},
+survey\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7},\protect\hyperlink{ref-iveniuk2021}{63}},
 and that pharmacies in the Peel region (an area identified as a
@@
 households) could not keep up with vaccine
-demand\textsuperscript{\protect\hyperlink{ref-gill2022}{65}}. This
+demand\textsuperscript{\protect\hyperlink{ref-gill2022}{64}}. This
 suggests that the observed differences are associated with disparities
@@
 deemed as ``essential'' in the context of the
-pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{66},\protect\hyperlink{ref-ct2021}{67}},
+pandemic\textsuperscript{\protect\hyperlink{ref-hawkins2020}{65},\protect\hyperlink{ref-ct2021}{66}},
 which include workers in the areas of grocery stores, gas stations,
@@
 mortality during the first year of the
-pandemic\textsuperscript{\protect\hyperlink{ref-rao2021}{68}}, but later
+pandemic\textsuperscript{\protect\hyperlink{ref-rao2021}{67}}, but later
 on, they had priority for COVID-19
-vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{69}}.
+vaccination\textsuperscript{\protect\hyperlink{ref-mishra2021}{68}}.
 Additionally, it is known that vaccination uptake in these individuals
 was encouraged by vaccination staff in certain parts of the
-province\textsuperscript{\protect\hyperlink{ref-gill2022}{65}}. These
+province\textsuperscript{\protect\hyperlink{ref-gill2022}{64}}. These
 facts, combined with evidence of increased trends in vaccination in this
 group
-elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{70}},
+elsewhere\textsuperscript{\protect\hyperlink{ref-nguyen2021b}{69}},
 suggest that in Ontario, the type of occupation of individuals in
@@
 pharmacist
-availability\textsuperscript{\protect\hyperlink{ref-shah2019}{71}--\protect\hyperlink{ref-timony2022}{73}}.
+availability\textsuperscript{\protect\hyperlink{ref-shah2019}{70}--\protect\hyperlink{ref-timony2022}{72}}.
 
@@
 focused on First Nations, Métis, and Inuit communities) to provide care
-to an estimated 100,000 Indigenous individuals living in the
-area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{74}}. Lack of
-access to personalized healthcare affects individuals that may mistrust
-the traditional healthcare system due to systemic racism or oppression,
-which is known to be the case for Indigenous and Black individuals in
-Canada, as these rationales have been associated with observed lower
-vaccination rates among these
-groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{75},\protect\hyperlink{ref-eissa2021}{76}}.
+to an estimated population of 100,000 Indigenous individuals living in
+the area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{73}}.
+Lack of access to personalized healthcare affects individuals that may
+mistrust the traditional healthcare system due to systemic racism or
+oppression, which is known to be the case for Indigenous and Black
+individuals in Canada, as these rationales have been associated with
+observed lower vaccination rates among these
+groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{74},\protect\hyperlink{ref-eissa2021}{75}}.
 Taken together, this suggests that healthcare disparities specific to
@@
 underrepresented groups
-elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{77}}.
+elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{76}}.
 
@@
 Ontario. However, these areas have the highest proportion of Indigenous
-inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{74}}.
+inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{73}}.
 In the context of personalized care, there is a need to collect data
@@
 vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{78}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{77}}, we
 believe that our data was able to provide a valid sample of vaccination
@@
 October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{79}});
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{78}});
 however, the short time window constitutes essentially a ``snapshot''
@@
 public health arm of the government (represented by PHO), and the
-healthcare system agency (OH). So far, the COVID-19 pandemic has as a
-stress test for both agencies, showing the necessity from decision
+healthcare system agency (OH). So far, the COVID-19 pandemic has served
+as a stress test for both agencies, showing the necessity from decision
 makers to clearly define how PHO and OH will interact in the future in
-order to ensure that in the future, Ontarians can benefit by virtue of a
-robust collaboration between both agencies. The evidence collected in
-this study shows that healthcare disparities experienced by
-underrepresented individuals in the province are the result of complex
-relationships between infrastructure, resources, occupations, geography,
-and socio-economic factors, among others. Therefore, this complexity
+order to ensure that Ontarians can benefit by virtue of a robust
+collaboration between both agencies. The evidence collected in this
+study shows that healthcare disparities experienced by underrepresented
+individuals in the province are the result of complex relationships
+between infrastructure, resources, occupations, geography, and
+socio-economic factors, among others. Therefore, this complexity
 presents an ongoing challenge for the healthcare system of Ontario,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-alami2021}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-bolotin2018}{}}%
 \CSLLeftMargin{56. }%
-\CSLRightInline{Alami H, Lehoux P, Fleet R, et al. How can health
-systems better prepare for the next pandemic? Lessons learned from the
-management of {COVID}-19 in quebec (canada). \emph{Frontiers in Public
-Health}. 2021;9.
-doi:\href{https://doi.org/10.3389/fpubh.2021.671833}{10.3389/fpubh.2021.671833}}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bolotin2018}{}}%
-\CSLLeftMargin{57. }%
 \CSLRightInline{Bolotin JJAG Shelly AND Feld. Population-based estimate
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ohesi}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Ontario HIV Epidemiology and Surveillance Initiative.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-carter2022}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{58. }%
 \CSLRightInline{Carter MA, Biro S, Maier A, Shingler C, Guan TH.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-basta2022}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Basta NE, Sohel N, Sulis G, et al. Factors associated
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2022b}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Cénat JM, Noorishad PG, Bakombo SM, et al. A systematic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-cnat2023}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Cénat JM, Farahi SMMM, Bakombo SM, et al. Vaccine
@@
 \leavevmode\vadjust pre{\hypertarget{ref-njoku2021}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Njoku A, Joseph M, Felix R. Changing the narrative:
@@
 \leavevmode\vadjust pre{\hypertarget{ref-iveniuk2021}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Iveniuk J, Leon S. \emph{Uneven Recovery: Measuring
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gill2022}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Gill M, Datta D, Gregory P, Austin Z. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-hawkins2020}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{Hawkins D. Differential occupational risk for {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ct2021}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Côté D, Durant S, MacEachen E, et al. A rapid scoping
@@
 \leavevmode\vadjust pre{\hypertarget{ref-rao2021}{}}%
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{Rao A, Ma H, Moloney G, et al. A disproportionate
@@
 \leavevmode\vadjust pre{\hypertarget{ref-mishra2021}{}}%
-\CSLLeftMargin{69. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{Mishra S, Stall NM, Ma H, et al. \emph{A Vaccination
@@
 \leavevmode\vadjust pre{\hypertarget{ref-nguyen2021b}{}}%
-\CSLLeftMargin{70. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Nguyen KH, Yankey D, Coy KC, et al. {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-shah2019}{}}%
-\CSLLeftMargin{71. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{Shah TI, Clark AF, Seabrook JA, Sibbald S, Gilliland JA.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-crighton2015}{}}%
-\CSLLeftMargin{72. }%
+\CSLLeftMargin{71. }%
 \CSLRightInline{Crighton EJ, Ragetlie R, Luo J, To T, Gershon A. A
@@
 \leavevmode\vadjust pre{\hypertarget{ref-timony2022}{}}%
-\CSLLeftMargin{73. }%
+\CSLLeftMargin{72. }%
 \CSLRightInline{Timony P, Houle SKD, Gauthier A, Waite NM. Geographic
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
-\CSLLeftMargin{74. }%
+\CSLLeftMargin{73. }%
 \CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
@@
 \leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
-\CSLLeftMargin{75. }%
+\CSLLeftMargin{74. }%
 \CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
@@
 \leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
-\CSLLeftMargin{76. }%
+\CSLLeftMargin{75. }%
 \CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
-\CSLLeftMargin{77. }%
+\CSLLeftMargin{76. }%
 \CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{78. }%
+\CSLLeftMargin{77. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{79. }%
+\CSLLeftMargin{78. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,

### 2023-07-12 code/data_preparation_May_08.R

@@ -269,6 +269,16 @@ clean_data <- clean_data%>%
 
 clean_data$Month <- fct_relevel(clean_data$Month,"October","November","December","January")
 
+clean_data<-clean_data%>%
+  mutate(Month=
+           case_when(Month=="October"~ "October 2021",
+                     Month=="November"~"November 2021",
+                     Month=="December"~"December 2021",
+                     Month=="January"~"January 2022"
+           ))%>%
+  mutate(Month=as.factor(Month))
+
+clean_data$Month <- fct_relevel(clean_data$Month,"October 2021","November 2021","December 2021","January 2022")
 
 # Remove certain Health Regions due to low number of observations: only 254 in North East
 # and North West Regions at this point, in contrast with about a thousand from each 
@@ -305,4 +315,4 @@ clean_data<- within(clean_data, Age_group_ord <- relevel(Age_group_ord, ref = "1
 clean_data<- within(clean_data, Race <- relevel(Race, ref = "White/Caucasian"))
 clean_data<- within(clean_data, Health_Region <- relevel(Health_Region, ref = "Toronto"))
 clean_data<- within(clean_data, income_ord <- relevel(income_ord, ref = "60000 and above"))
-clean_data<- within(clean_data, Month <- relevel(Month, ref = "October"))
+clean_data<- within(clean_data, Month <- relevel(Month, ref = "October 2021"))

### 2023-07-18 manuscript_v7/01-background.qmd
new version from BN's comments

--- 2023-07-12 manuscript_v6/01-background.qmd
@@
 However, it is important to consider that vaccination uptake can also vary geographically across the administrative or political subdivisions within a country.
-For example, differences in COVID-19 vaccination rates have been associated with differences in attitudes towards vaccination in administrative areas in New York and Chicago[@malik2020], lack of vaccine access in areas where underrepresented groups reside in Ontario[@bogoch2022], as well as an absence of prioritization of areas inhabited by vulnerable groups in the southeastern region of the US[@nguyen2021].
+For example, differences in COVID-19 vaccination rates have been associated with differences in attitudes towards vaccination in administrative areas in New York and Chicago[@malik2020], scarcity of vaccination facilities in areas where underrepresented groups reside in Toronto[@bogoch2022], as well as an absence of prioritization of areas inhabited by vulnerable groups in the southeastern region of the US[@nguyen2021].
 Other studies have also shown heterogeneity in vaccine uptake within counties in the US [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], indicating that accounting for geographical differences in vaccination can help predict patterns of booster uptake in Scottish communities [@wood2022].
@@
 Although PHUs were commissioned with leading the distribution of the COVID-19 vaccine within their respective areas [@rotenberg2021], the vaccine rollout occurred with significant interaction between PHO and OH; in many instances both agencies had to work together to organize vaccination clinics, with personnel associated with OH also being actively involved when demand exceeded the capacity of PHO personnel, especially in rural areas [@ashcroft2023].
-Indeed, based on the experience of COVID-19, there is an ongoing debate on how OH and PHO will interact in the future, and the challenges that this will entail for the healthcare system of the province [@dion2022].
+Indeed, based on the experience of COVID-19, there are ongoing discussions on how OH and PHO will interact in the future, and the challenges that this will entail for the healthcare system of the province [@dion2022].
 
@@
 
-<!-- before the advent of a future pandemic -->
-
-<!-- In the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within Ontario and identify the socio-demographic groups that might be affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
--->
-
-<!-- there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
-In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
-Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
--->
-
-<!-- Due to the relatively recent adoption of the Health Region model and its alignment with the onset of the COVID-19 pandemic,  -->
-
-<!-- In the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within Ontario and identify the socio-demographic groups that might be affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
--->
-
-<!-- how the relatively recent adoption of the Health Region model the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
--->
-
-<!-- there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
-In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
-Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
--->
-
-<!-- However, in the context of Canada's pandemic response goals, which have been to minimize serious illness and deaths while minimizing societal disruption [@naci2023], there is an ongoing need to analyze the existence of geographical disparities in vaccination uptake within the Canadian provinces and to identify the socio-demographic groups affected, as this can help identify inequalities that might exist and that need to be addressed before the advent of a future pandemic.
--->
-
-<!-- there is a need to analyze if there are ongoing disparities in health access under this approach that need to be addressed before they are exploited by a new disease or public health threat.
-In this regard, previous research has highlighted disparities in the level of activity of each Health Region [@sethuram2023].
-Therefore, analyzing differences in vaccination uptake within the Health Regions and can help identify which socio-demographic groups are the most vulnerable and what areas of the province deserve special attention by decision-makers.
--->
-
-Therefore, in this study we hypothesized that there were differences in COVID-19 vaccination rates between the Health Regions between October 2021 and January of 2022.
-To understand and contextualize these differences, we by included socio-economic factors in our analysis, aiming to identify in which demographic groups were particularly impacted, in order to provide an assessment of the current state of healthcare access in Ontario.
+Therefore, in this study we wanted to understand if there were differences in COVID-19 vaccination rates among the Health Regions between October 2021 and January of 2022.
+To contextualize these differences, we included socio-economic factors in our analysis aiming to identify which demographic groups were particularly impacted, in order to provide an assessment of the current state of healthcare access in Ontario.

### 2023-07-18 manuscript_v7/02-methods.qmd
new version from BN's comments

--- 2023-07-12 manuscript_v6/02-methods.qmd
@@
 
-We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto (under Research Ethics Board approval #40999), and which ran between September 30th, 2021 and January 17th,2022.
+We used data from the *Survey of COVID-19 related Behaviours and Attitudes*, a repeated cross-sectional survey focused on the Canadian province of Ontario that was commissioned by the Fields Institute for Research in Mathematical Sciences and the  Mathematical Modelling of COVID-19 Task Force under ethical guidance from the University of Toronto (under Research Ethics Board approval #40999), and which ran between September 30th, 2021 and January 17th, 2022.
 
@@
 
-Survey users entered their socio-economic information (age, income, and racial/ethnic group), and asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no" (@tbl-descriptive-stats).
-Of notice, because the racial/ethnic categories of the survey did not match exactly the categories from the 2016 Census, we used a combination of sources to re-group certain categories in order to obtain estimates that could be used to correct the data.
+Survey users entered their socio-economic information (age, income, and racial/ethnic group), and were asked information on vaccination status by using the question "Have you received the first dose of the COVID vaccine?", with possible answers "yes" and "no" (@tbl-descriptive-stats).
+Of notice, the racial/ethnic and  socio-economic categories from the survey did not match exactly the categories from the 2016 Census (e.g., the categories "East Asian/Pacific Islander" as well as the income bracket "over CAD 110,000" do not exist in the Census data).
+Therefore, we used a combination of sources to re-group certain categories and socio-economic strata in order to obtain estimates from the Census that could be used to correct the data.
 Further details can be found in the Appendix.
@@
 The original dataset contained 39,029 observations.
-However, the number of complete observations was much lower that the total number of observations due to the survey design, which allowed respondents to exit at any time and deployed the questions randomly.
+However, the number of complete observations was much lower than the total number of observations due to the survey design, which allowed respondents to exit at any time and deployed the questions randomly.
 We selected the observations with complete answers (6,343 observations, or 16.25\% of the total) for our analysis.
-It should be noted that this response rate is similar to response rates observed in previous studies that have used the same type of survey instrument and that have ranged between 15\% and 22\%[@seeman2010;@kim2021].
+It should be noted that this response rate is similar to response rates observed in previous studies that have used the same type of survey instrument, with response rates between 15\% and 22\%[@seeman2010;@kim2021].
 Next, we matched the city of each observation with its corresponding LHIN and Health Region, and removed observations from areas with low representation (254 observations corresponding to the North West and North East Health Regions).
 After all the preliminary analyses, the total number of observations used for analysis was 6,236 and included the East, Central, Toronto, and West Health Regions covering between October 1st, 2021 and January 17, 2022.
-The original dataset, clean dataset, and details on the data cleaning process are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.
+The original dataset, clean dataset, and details on the data cleaning process, and data preparation are described in detail in the [GitHub repository](https://github.com/aimundo/Fields_COVID-19/) for this paper.
 
@@
 
-## Statistical analyses
+<!-- ## Statistical analyses -->
 
-We used a logistic regression model to examine the impact of the Health Regions in vaccination rates while considering the socio-economic factors and months covered by the survey (@tbl-descriptive-stats) and certain interactions (Race and Health Region and Race and income), as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
-Because we identified differences in representativity between the survey data and the estimates from the 2016 Census for the variables considered in the analysis, and used used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the data using data from the Census and Health Region population totals; and fitted the regression model to the uncorrected and corrected data.
+To analyze the survey data, we used a logistic regression model to examine the impact of the Health Regions in the uptake of the first dose of the vaccine, using as the variable of interest the question "Have you received the first dose of the COVID vaccine?" (with possible answers "yes" or "no"), while considering as independent variables the socio-economic factors of the participants, months covered by the survey (@tbl-descriptive-stats), and the Health Regions of OH.
+Furthermore, we included certain interactions (Race and Health Region and Race and income) in our model as previous studies have shown that socio-economic factors and their interactions are significant predictors of intent of vaccination and vaccination status [@nguyen2022; @shih2021; @cnat2022a].
+
+Because we identified differences between the percentage of participation within the socio-demographic factors of the survey considered in in our analysis and the percentage within these factors in the 2016 Census (such as participation within age groups and racial/ethnic groups), we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the percentages within each socio-demographic variable using population totals obtained from the Census and OH; later, we fitted the regression model to the uncorrected and corrected data to determine if there were any differences in the obtanied estimates.
 Details regarding the correction can be found in the Appendix.

### 2023-07-18 manuscript_v7/03-results.qmd
new version from BN's comments

--- 2023-07-12 manuscript_v6/03-results.qmd
@@
 
-@fig-models presents the estimates (as odd ratios) from the logistic regression models for vaccination status using the socio-demographic factors collected by the survey, and their interactions.
+@fig-models presents the estimates (as odd ratios) from the logistic regression models for vaccination status using the socio-demographic factors collected by the survey, and their interactions for the corrected and uncorrected data.
 Generally speaking, lower odds of vaccination were identified in both cases in individuals characterized by a low household income, or that identified as part of underrepresented groups.
@@
 #| label: fig-models
-#| fig-cap: "Coefficient estimates and confidence intervals for the uncorrected model.
-Only statistically significant interaction terms are shown.
+#| fig-cap: "**A**.
+Coefficient estimates and confidence intervals for the uncorrected model.
+**B**.
+Coefficient estimates and confidence intervals for the corrected model.
+Only statistically significant interaction terms are shown in both cases.
 Full interaction terms can be found in Supplementary Figures A-3 and A-4."

### 2023-07-18 manuscript_v7/04-discussion.qmd
new version from BN's comments

--- 2023-07-12 manuscript_v6/04-discussion.qmd
@@
 
-In this study, we hypothesized that there were differences in the uptake of the first dose of the COVID-19 vaccine between individuals living in the different Health Regions of Ontario during late 2021 and early 2022.
+In this study, we wanted to understand if there were differences in COVID-19 vaccination rates among the Health Regions between October 2021 and January of 2022 while identifying the socio-demographic determinants (and their interactions) that could help predict these differences.
 Using the Health Regions as the base of our analysis was advantageous, as these newly implemented administrative areas match closely the geographical boundaries of the Health Regions that have been historically used to group Public Health Units in the province [@bolotin2018;@ohesi].
-Thus, such analysis can provide information on vaccination rates with relevance at both the public health and healthcare system level.
+Thus, a Health Region-based analysis could provide information on vaccination uptake that could be useful to understand patterns within the realm of PHO and OH.
 Moreover, such analysis could provide an overall assessment of intra-provincial disparities that might need to be addressed moving forward by decision-makers in order to ensure that in the event of a future public health emergency or pandemic, OH and PHO will able to work collaboratively in an efficient way having learned from the COVID-19 experience[@ashcroft2023], thus ensuring that the population of the province benefits in the long term from a robust health system.
@@
 However, more granular corrections would be needed to obtain more accurate estimates: For example, our analysis identified higher odds of vaccination in the Central and West Health Regions, but in this case these differences are likely to be driven by the proportion of White/Caucasian individuals, who had higher vaccination rates than other racial groups.
-Correcting for each racial/ethnic group in each Health Region would provide a more accurate estimation of region-wide vaccination rates but unfortunately, presently this correction cannot be implemented as such stratification has not been implemented in the data that can be obtained from the Census.
+Correcting for each racial/ethnic group in each Health Region would provide a more accurate estimation of region-wide vaccination rates but unfortunately, presently this correction cannot be implemented as such stratification has not been implemented in the open data that can be obtained from the Census.
 

### 2023-07-18 manuscript_v7/05-conclusion.qmd
new version from BN's comments

# Conclusion

The evidence collected in this study shows that differences in the uptake of the first dose of the COVID-19 vaccine by underrepresented individuals in the province of Ontario were the result of complex relationships between infrastructure, resources, occupations, geography, and socio-economic factors. Specifically, we identified that between the last trimester of 2021 and January of 2022 individuals that identified as Arab or Middle Eastern, Black, Latin American, of a mixed background, or that belonged to other racial groups had significantly lower odds of vaccination when compared to White/Caucasian individuals in Ontario.

However, although overall vaccination uptake was lower in these underrepresented individuals in the province as a whole, the odds of vaccination differed within certain income levels and Health Regions in these groups, specially for those that reported a low household income and that had higher vaccination odds that individuals in the same racial/ethnic group with a higher income. These results highlight the complex landscape of the province, where varying degrees of rurality exist in conjunction with a socio-demographic makeup that is unique to each of the Health Regions.

Such complexity presents an ongoing challenge for the healthcare system of Ontario, and in particular, to the Health Region model which will need to adapt and build on the LHIN model with the aim of providing an integrative approach that can serve those that are most vulnerable across all the Health Regions of the province. In particular, personalized care is an area that could be further developed to improve vaccination uptake in underrepresented individuals in the future. As we have shown above, there are currently a limited number of centers focused on community-led healthcare in the province, and our results showed the existence of differences in vaccination uptake between certain Health Regions. Improving personalized care for underrepresented individuals then, can serve as a point to improve trust and facilitate vaccine access in these marginalized communities.

<!-- The implementation of the Health Regions in Ontario aimed at reducing the bureaucratic complexity and health disparities identified under the LHIN model. However, there are certain challenges that this new model will need to overcome in order to work syngergistically with the rest of the healthcare system of the province. First, each Health Region now covers a large geographical area that was served in the past by multiple LHINs. This in turn creates a complex socio-demographic landscape within each Health Region that is likely to be different in each case due the different levels of rurality and the proportion of underrepresented groups in each Health Region.  -->

Furthermore, in the event of a future pandemic or public health emergency in the province in the years ahead, PHO and OH will need to collaborate on the implementation of public health measures (such as vaccination campaigns) building on the experience gained during the COVID-19 vaccination rollout. So far, the COVID-19 pandemic served as a stress test for both agencies, showing the necessity from decision makers to clearly define how PHO and OH will interact in the future in order to ensure that Ontarians can benefit by virtue of a robust collaboration between both agencies.

<!-- The evidence collected in this study shows that healthcare disparities experienced by underrepresented individuals in the province are the result of complex relationships between infrastructure, resources, occupations, geography, and socio-economic factors, among others. Therefore, this complexity presents an ongoing challenge for the healthcare system of Ontario, which will need to adapt and continue to aim for an integrative approach that can serve those that are most vulnerable. -->

<!-- So far, the evidence collected during the COVID-19 pandemic indicates that differences in vaccination uptake are associated to a lack of infrastructure and resources that can adequately support personalized care to marginalized individuals. In the near future, health decision-makers will need to consider the implementation of policies that are focused on addressing this problematic. -->

Finally, the recent nature in the adaption of the Health Region poses a challenge for researchers in the acquisition of data and information that can be used to analyze the performance of the new model. From one side, the Health Regions have not been incorporated as part of Census data (LHINs were considered before in the Census), and this impacts the amount of data that can be obtained. On the other hand, evaluations of the Health Region model are at the moment not considered in the Annual Reports of the Auditor General of Ontario, which was critical in the past to identify limitations in the LHIN model. Currently, the only demographic information available for each Health Region is provided by Ontario Health (the agency that administers the Health Regions) but this information only provides general estimates that do not allow for detailed analyses of performance indicators (such as hospitalizations, readmissions, and trends in chronic disease incidence) between the Regions. Therefore, there is a pressing need for open data that can be used by researchers and decision-makers to assess the performance of the Health Region model.

<!-- The Health Region model will only by successful if it ensures that healthcare improves across all segments of the population of Ontario, particularly in the event of a future public health emergency or pandemic where so far, based on the experience of the COVID-19 pandemic, underrepresented individuals have been disproportionately affected. -->

### 2023-07-18 manuscript_v7/main.tex
new version from BN's comments

--- 2023-07-12 manuscript_v6/main.tex
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 Covid-19, vaccination, survey, socio-economic factors, visible
-minorities.
+minorities, race and ethnicity
 
@@
 in administrative areas in New York and
-Chicago\textsuperscript{\protect\hyperlink{ref-malik2020}{25}}, lack of
-vaccine access in areas where underrepresented groups reside in
-Ontario\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as well
-as an absence of prioritization of areas inhabited by vulnerable groups
-in the southeastern region of the
+Chicago\textsuperscript{\protect\hyperlink{ref-malik2020}{25}}, scarcity
+of vaccination facilities in areas where underrepresented groups reside
+in Toronto\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as
+well as an absence of prioritization of areas inhabited by vulnerable
+groups in the southeastern region of the
 US\textsuperscript{\protect\hyperlink{ref-nguyen2021}{26}}. Other
@@
 areas\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}}. Indeed,
-based on the experience of COVID-19, there is an ongoing debate on how
-OH and PHO will interact in the future, and the challenges that this
+based on the experience of COVID-19, there are ongoing discussions on
+how OH and PHO will interact in the future, and the challenges that this
 will entail for the healthcare system of the
@@
 
-Therefore, in this study we hypothesized that there were differences in
-COVID-19 vaccination rates between the Health Regions between October
-2021 and January of 2022. To understand and contextualize these
-differences, we by included socio-economic factors in our analysis,
-aiming to identify in which demographic groups were particularly
-impacted, in order to provide an assessment of the current state of
-healthcare access in Ontario.
+Therefore, in this study we wanted to understand if there were
+differences in COVID-19 vaccination rates among the Health Regions
+between October 2021 and January of 2022. To contextualize these
+differences, we included socio-economic factors in our analysis aiming
+to identify which demographic groups were particularly impacted, in
+order to provide an assessment of the current state of healthcare access
+in Ontario.
 
@@
 We used data from the \emph{Survey of COVID-19 related Behaviours and
-Attitudes}, a repeated cross sectional survey focused on the Canadian
+Attitudes}, a repeated cross-sectional survey focused on the Canadian
 province of Ontario that was commissioned by the Fields Institute for
@@
 Toronto (under Research Ethics Board approval \#40999), and which ran
-between September 30th, 2021 and January 17th,2022.
+between September 30th, 2021 and January 17th, 2022.
 
@@
 Survey users entered their socio-economic information (age, income, and
-racial/ethnic group), and asked information on vaccination status by
-using the question ``Have you received the first dose of the COVID
+racial/ethnic group), and were asked information on vaccination status
+by using the question ``Have you received the first dose of the COVID
 vaccine?'', with possible answers ``yes'' and ``no''
-(Table~\ref{tbl-descriptive-stats}). Of notice, because the
-racial/ethnic categories of the survey did not match exactly the
-categories from the 2016 Census, we used a combination of sources to
-re-group certain categories in order to obtain estimates that could be
-used to correct the data. Further details can be found in the Appendix.
+(Table~\ref{tbl-descriptive-stats}). Of notice, the racial/ethnic and
+socio-economic categories from the survey did not match exactly the
+categories from the 2016 Census (e.g., the categories ``East
+Asian/Pacific Islander'' as well as the income bracket ``over CAD
+110,000'' do not exist in the Census data). Therefore, we used a
+combination of sources to re-group certain categories and socio-economic
+strata in order to obtain estimates from the Census that could be used
+to correct the data. Further details can be found in the Appendix.
 
@@
 original dataset contained 39,029 observations. However, the number of
-complete observations was much lower that the total number of
+complete observations was much lower than the total number of
 observations due to the survey design, which allowed respondents to exit
@@
 is similar to response rates observed in previous studies that have used
-the same type of survey instrument and that have ranged between 15\% and
+the same type of survey instrument, with response rates between 15\% and
 22\%\textsuperscript{\protect\hyperlink{ref-kim2021}{45},\protect\hyperlink{ref-seeman2010}{46}}.
@@
 October 1st, 2021 and January 17, 2022. The original dataset, clean
-dataset, and details on the data cleaning process are described in
-detail in the \href{https://github.com/aimundo/Fields_COVID-19/}{GitHub
-repository} for this paper.
+dataset, and details on the data cleaning process, and data preparation
+are described in detail in the
+\href{https://github.com/aimundo/Fields_COVID-19/}{GitHub repository}
+for this paper.
 
@@
 
-\hypertarget{statistical-analyses}{%
-\subsection{Statistical analyses}\label{statistical-analyses}}
-
-We used a logistic regression model to examine the impact of the Health
-Regions in vaccination rates while considering the socio-economic
-factors and months covered by the survey
-(Table~\ref{tbl-descriptive-stats}) and certain interactions (Race and
-Health Region and Race and income), as previous studies have shown that
-socio-economic factors and their interactions are significant predictors
-of intent of vaccination and vaccination
+To analyze the survey data, we used a logistic regression model to
+examine the impact of the Health Regions in the uptake of the first dose
+of the vaccine, using as the variable of interest the question ``Have
+you received the first dose of the COVID vaccine?'' (with possible
+answers ``yes'' or ``no''), while considering as independent variables
+the socio-economic factors of the participants, months covered by the
+survey (Table~\ref{tbl-descriptive-stats}), and the Health Regions of
+OH. Furthermore, we included certain interactions (Race and Health
+Region and Race and income) in our model as previous studies have shown
+that socio-economic factors and their interactions are significant
+predictors of intent of vaccination and vaccination
 status\textsuperscript{\protect\hyperlink{ref-nguyen2022}{47}--\protect\hyperlink{ref-cnat2022a}{49}}.
-Because we identified differences in representativity between the survey
-data and the estimates from the 2016 Census for the variables considered
-in the analysis, and used used an iterative proportional fitting
+
+Because we identified differences between the percentage of
+participation within the socio-demographic factors of the survey
+considered in in our analysis and the percentage within these factors in
+the 2016 Census (such as participation within age groups and
+racial/ethnic groups), we used an iterative proportional fitting
 procedure
 (\emph{raking})\textsuperscript{\protect\hyperlink{ref-deming1940}{50}}
-to correct the data using data from the Census and Health Region
-population totals; and fitted the regression model to the uncorrected
-and corrected data. Details regarding the correction can be found in the
-Appendix. All analyses were conducted in R 4.2.2 using the packages
+to correct the percentages within each socio-demographic variable using
+population totals obtained from the Census and OH; later, we fitted the
+regression model to the uncorrected and corrected data to determine if
+there were any differences in the obtanied estimates. Details regarding
+the correction can be found in the Appendix. All analyses were conducted
+in R 4.2.2 using the packages
 \texttt{survey}\textsuperscript{\protect\hyperlink{ref-lumley2011}{51}},\texttt{tidyverse}\textsuperscript{\protect\hyperlink{ref-wickham2019}{52}},
@@
 socio-demographic factors collected by the survey, and their
-interactions. Generally speaking, lower odds of vaccination were
-identified in both cases in individuals characterized by a low household
-income, or that identified as part of underrepresented groups. However,
-the magnitude of the estimates differed between the uncorrected and
-corrected models and more importantly, there were differences in the
-statistical significance of certain estimates before and after the
-correction. Specifically, the uncorrected model showed significant
-differences in vaccination odds between the age groups considered, the
-East Health Region, Latin American individuals with a household income
-under CAD 25,000, and Indigenous individuals living in the Central
-Health Region (Figure~\ref{fig-models},B) but these were deemed non
-statistically significant after the correction.
+interactions for the corrected and uncorrected data. Generally speaking,
+lower odds of vaccination were identified in both cases in individuals
+characterized by a low household income, or that identified as part of
+underrepresented groups. However, the magnitude of the estimates
+differed between the uncorrected and corrected models and more
+importantly, there were differences in the statistical significance of
+certain estimates before and after the correction. Specifically, the
+uncorrected model showed significant differences in vaccination odds
+between the age groups considered, the East Health Region, Latin
+American individuals with a household income under CAD 25,000, and
+Indigenous individuals living in the Central Health Region
+(Figure~\ref{fig-models},B) but these were deemed non statistically
+significant after the correction.
 
@@
 
-\caption{\label{fig-models}Coefficient estimates and confidence
-intervals for the uncorrected model. Only statistically significant
-interaction terms are shown. Full interaction terms can be found in
-Supplementary Figures A-3 and A-4.}
+\caption{\label{fig-models}\textbf{A}. Coefficient estimates and
+confidence intervals for the uncorrected model. \textbf{B}. Coefficient
+estimates and confidence intervals for the corrected model. Only
+statistically significant interaction terms are shown in both cases.
+Full interaction terms can be found in Supplementary Figures A-3 and
+A-4.}
 
@@
 
-In this study, we hypothesized that there were differences in the uptake
-of the first dose of the COVID-19 vaccine between individuals living in
-the different Health Regions of Ontario during late 2021 and early 2022.
+In this study, we wanted to understand if there were differences in
+COVID-19 vaccination rates among the Health Regions between October 2021
+and January of 2022 while identifying the socio-demographic determinants
+(and their interactions) that could help predict these differences.
 Using the Health Regions as the base of our analysis was advantageous,
@@
 province\textsuperscript{\protect\hyperlink{ref-bolotin2018}{56},\protect\hyperlink{ref-ohesi}{57}}.
-Thus, such analysis can provide information on vaccination rates with
-relevance at both the public health and healthcare system level.
-Moreover, such analysis could provide an overall assessment of
-intra-provincial disparities that might need to be addressed moving
-forward by decision-makers in order to ensure that in the event of a
-future public health emergency or pandemic, OH and PHO will able to work
-collaboratively in an efficient way having learned from the COVID-19
+Thus, a Health Region-based analysis could provide information on
+vaccination uptake that could be useful to understand patterns within
+the realm of PHO and OH. Moreover, such analysis could provide an
+overall assessment of intra-provincial disparities that might need to be
+addressed moving forward by decision-makers in order to ensure that in
+the event of a future public health emergency or pandemic, OH and PHO
+will able to work collaboratively in an efficient way having learned
+from the COVID-19
 experience\textsuperscript{\protect\hyperlink{ref-ashcroft2023}{41}},
@@
 vaccination rates but unfortunately, presently this correction cannot be
-implemented as such stratification has not been implemented in the data
-that can be obtained from the Census.
+implemented as such stratification has not been implemented in the open
+data that can be obtained from the Census.
 
@@
 
-The implementation of the Health Regions in Ontario aimed at reducing
-the bureaucratic complexity and health disparities identified under the
-LHIN model. However, there are certain challenges that this new model
-will need to overcome in order to work syngergistically with the rest of
-the healthcare system of the province. First, each Health Region now
-covers a large geographical area that was served in the past by multiple
-LHINs. This in turn creates a complex socio-demographic landscape within
-each Health Region that is likely to be different in each case due the
-different levels of rurality and the proportion of underrepresented
-groups in each Health Region.
-
-Second, it is likely that a future pandemic or public health emergency
-will impact the province in the years ahead. Although broadly speaking,
-PHO and OH have been able to collaborate on the COVID-19 vaccination
-rollout, future events will require a unified response from both the
-public health arm of the government (represented by PHO), and the
-healthcare system agency (OH). So far, the COVID-19 pandemic has served
-as a stress test for both agencies, showing the necessity from decision
+The evidence collected in this study shows that differences in the
+uptake of the first dose of the COVID-19 vaccine by underrepresented
+individuals in the province of Ontario were the result of complex
+relationships between infrastructure, resources, occupations, geography,
+and socio-economic factors. Specifically, we identified that between the
+last trimester of 2021 and January of 2022 individuals that identified
+as Arab or Middle Eastern, Black, Latin American, of a mixed background,
+or that belonged to other racial groups had significantly lower odds of
+vaccination when compared to White/Caucasian individuals in Ontario.
+
+However, although overall vaccination uptake was lower in these
+underrepresented individuals in the province as a whole, the odds of
+vaccination differed within certain income levels and Health Regions in
+these groups, specially for those that reported a low household income
+and that had higher vaccination odds that individuals in the same
+racial/ethnic group with a higher income. These results highlight the
+complex landscape of the province, where varying degrees of rurality
+exist in conjunction with a socio-demographic makeup that is unique to
+each of the Health Regions.
+
+Such complexity presents an ongoing challenge for the healthcare system
+of Ontario, and in particular, to the Health Region model which will
+need to adapt and build on the LHIN model with the aim of providing an
+integrative approach that can serve those that are most vulnerable
+across all the Health Regions of the province. In particular,
+personalized care is an area that could be further developed to improve
+vaccination uptake in underrepresented individuals in the future. As we
+have shown above, there are currently a limited number of centers
+focused on community-led healthcare in the province, and our results
+showed the existence of differences in vaccination uptake between
+certain Health Regions. Improving personalized care for underrepresented
+individuals then, can serve as a point to improve trust and facilitate
+vaccine access in these marginalized communities.
+
+Furthermore, in the event of a future pandemic or public health
+emergency in the province in the years ahead, PHO and OH will need to
+collaborate on the implementation of public health measures (such as
+vaccination campaigns) building on the experience gained during the
+COVID-19 vaccination rollout. So far, the COVID-19 pandemic served as a
+stress test for both agencies, showing the necessity from decision
 makers to clearly define how PHO and OH will interact in the future in
 order to ensure that Ontarians can benefit by virtue of a robust
-collaboration between both agencies. The evidence collected in this
-study shows that healthcare disparities experienced by underrepresented
-individuals in the province are the result of complex relationships
-between infrastructure, resources, occupations, geography, and
-socio-economic factors, among others. Therefore, this complexity
-presents an ongoing challenge for the healthcare system of Ontario,
-which will need to adapt and continue to aim for an integrative approach
-that can serve those that are most vulnerable.
+collaboration between both agencies.
 
@@
 side, the Health Regions have not been incorporated as part of Census
-data (LHINs were considered before in the Census), and this impact the
-amount and level of detail of available information. On the other hand,
-evaluations of the Health Region model are at the moment not considered
-in the Annual Reports of the Auditor General of Ontario, which was
-critical in the past to identify limitations in the LHIN model.
-Currently, the only demographic information available for each Health
-Region is provided by Ontario Health (the agency that administers the
-Health Regions) but this information only provides general estimates
-that do not allow for detailed analyses of performance indicators (such
-as hospitalizations, readmissions, and trends in chronic disease
-incidence) between the Regions. Therefore, there is a pressing need for
-open data that can be used by researchers and decision-makers to examine
-the performance of the Health Region model.
-
-The Health Region model will only by successful if it ensures that
-healthcare improves across all segments of the population of Ontario,
-particularly in the event of a future public health emergency or
-pandemic where so far, based on the experience of the COVID-19 pandemic,
-underrepresented individuals have been disproportionately affected.
+data (LHINs were considered before in the Census), and this impacts the
+amount of data that can be obtained. On the other hand, evaluations of
+the Health Region model are at the moment not considered in the Annual
+Reports of the Auditor General of Ontario, which was critical in the
+past to identify limitations in the LHIN model. Currently, the only
+demographic information available for each Health Region is provided by
+Ontario Health (the agency that administers the Health Regions) but this
+information only provides general estimates that do not allow for
+detailed analyses of performance indicators (such as hospitalizations,
+readmissions, and trends in chronic disease incidence) between the
+Regions. Therefore, there is a pressing need for open data that can be
+used by researchers and decision-makers to assess the performance of the
+Health Region model.
 

### 2023-07-18 manuscript_v7/02-methods.qmd
updated

@@
 
-Because we identified differences between the percentage of participation within the socio-demographic factors of the survey considered in in our analysis and the percentage within these factors in the 2016 Census (such as participation within age groups and racial/ethnic groups), we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the percentages within each socio-demographic variable using population totals obtained from the Census and OH; later, we fitted the regression model to the uncorrected and corrected data to determine if there were any differences in the obtanied estimates.
+Because we identified differences between the percentage of participation within each of the socio-demographic variables collected by the survey that were considered in our analysis and the percentage within these factors in the 2016 Census (such as participation by age and racial/ethnic groups), we used an iterative proportional fitting procedure (_raking_) [@deming1940] to correct the percentages within each socio-demographic variable using population totals obtained from the Census and OH; later, we fitted the regression model to the uncorrected and corrected data to determine if there were any differences in the obtanied estimates.
 Details regarding the correction can be found in the Appendix.

### 2023-07-18 manuscript_v7/main.tex
updated

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden]}{\end{tcolorbox}}\fi
 
@@
 Because we identified differences between the percentage of
-participation within the socio-demographic factors of the survey
-considered in in our analysis and the percentage within these factors in
-the 2016 Census (such as participation within age groups and
-racial/ethnic groups), we used an iterative proportional fitting
+participation within each of the socio-demographic variables collected
+by the survey that were considered in our analysis and the percentage
+within these factors in the 2016 Census (such as participation by age
+and racial/ethnic groups), we used an iterative proportional fitting
 procedure

### 2023-07-28 manuscript_v7/01-background.qmd
updated conclusion and discussion

@@
 However, it is important to consider that vaccination uptake can also vary geographically across the administrative or political subdivisions within a country.
-For example, differences in COVID-19 vaccination rates have been associated with differences in attitudes towards vaccination in administrative areas in New York and Chicago[@malik2020], scarcity of vaccination facilities in areas where underrepresented groups reside in Toronto[@bogoch2022], as well as an absence of prioritization of areas inhabited by vulnerable groups in the southeastern region of the US[@nguyen2021].
+For example, differences in COVID-19 vaccination rates have been associated with differences in attitudes towards vaccination in administrative areas in New York and Chicago[@malik2020], scarcity of vaccination facilities in areas where underrepresented groups reside in Toronto when the provincial vaccination program started[@bogoch2022], as well as an absence of prioritization of areas inhabited by vulnerable groups in the southeastern region of the US[@nguyen2021].
 Other studies have also shown heterogeneity in vaccine uptake within counties in the US [@mollalo2021; @yang2022; @tiu2022; @bhuiyan2022], indicating that accounting for geographical differences in vaccination can help predict patterns of booster uptake in Scottish communities [@wood2022].

### 2023-07-28 manuscript_v7/04-discussion.qmd
updated conclusion and discussion

@@
 Furthermore, there is an ongoing challenge for the health system of the province with regard to personalized healthcare for marginalized individuals.
-For example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated population of 100,000 Indigenous individuals living in the area [@ontariohealth].
-Lack of access to personalized healthcare affects individuals that may mistrust the traditional healthcare system due to systemic racism or oppression, which is known to be the case for Indigenous and Black individuals in Canada, as these rationales have been associated with observed lower vaccination rates among these groups [@smylie2022;@eissa2021].
-Taken together, this suggests that healthcare disparities specific to these underrepresented groups in certain parts of the province impacted vaccination uptake, and highlights the need of investments in the Health Regions focused on resources, infrastructure, and specially personnel that can deliver personalized care to marginalized communities, as it has been shown that such efforts have improved trust in vaccination in underrepresented groups elsewhere [@schafferderoo2020].
+Indeed, previous studies have pointed out mistrust in the traditional healthcare system (due to systemic racism or oppression) as a rationale for lower vaccination rates in the case of Indigenous and Black individuals in Canada [@smylie2022;@eissa2021].
+Other studies have also shown a need for increased culturally responsive care for other underrepresented groups such as Black women and and other immigrant groups in Canada and Ontario [@etowa2021;@patel2019].
+Considering that for example, the West Health Region has only two Aboriginal Health Access Centres (community-led primary healthcare organizations focused on First Nations, Métis, and Inuit communities) to provide care to an estimated population of 100,000 Indigenous individuals living in the area [@ontariohealth], it is possible that the limited personalized healthcare for underrepresented groups in certain parts of the province impacted vaccination uptake as well, and highlights the need of investments in the Health Regions focused on resources, infrastructure, and specially personnel that can deliver personalized care to marginalized communities, as it has been shown that such efforts have improved trust in vaccination in underrepresented groups elsewhere [@schafferderoo2020].
 

### 2023-07-28 manuscript_v7/05-conclusion.qmd
updated conclusion and discussion

@@
 
-The evidence collected in this study shows that differences in the uptake of the first dose of the COVID-19 vaccine by underrepresented individuals in the province of Ontario were the result of complex relationships between infrastructure, resources, occupations, geography, and socio-economic factors.
-Specifically, we identified that between the last trimester of 2021 and January of 2022 individuals that identified as Arab or Middle Eastern, Black, Latin American, of a mixed background, or that belonged to other racial groups had significantly lower odds of vaccination when compared to White/Caucasian individuals in Ontario.
-
-However, although overall vaccination uptake was lower in these underrepresented individuals in the province as a whole, the odds of vaccination differed within certain income levels and Health Regions in these groups, specially for those that reported a low household income and that had higher vaccination odds that individuals in the same racial/ethnic group with a higher income.
+The evidence collected in this study shows differences in the uptake of the first dose of the COVID-19 vaccine in Ontario between October 2021 and January 2022 in underrepresented groups, which had significantly lower odds of vaccination when compared to White/Caucasian individuals.
+However, although overall vaccination uptake was lower in underrepresented individuals in the province as a whole, the odds of vaccination differed within certain income levels and Health Regions in these groups.
+Specifically, those that reported a low household income had higher vaccination odds when compared to individuals in the same racial/ethnic group in a higher income bracket.
 These results highlight the complex landscape of the province, where varying degrees of rurality exist in conjunction with a socio-demographic makeup that is unique to each of the Health Regions.
 
-Such complexity presents an ongoing challenge for the healthcare system of Ontario, and in particular, to the Health Region model which will need to adapt and build on the LHIN model with the aim of providing an integrative approach that can serve those that are most vulnerable across all the Health Regions of the province.
-In particular, personalized care is an area that could be further developed to improve vaccination uptake in underrepresented individuals in the future.
-As we have shown above, there are currently a limited number of centers focused on community-led healthcare in the province, and our results showed the existence of differences in vaccination uptake between certain Health Regions.
-Improving personalized care for underrepresented individuals then, can serve as a point to improve trust and facilitate vaccine access in these marginalized communities.
+Personalized care is an area that could be further developed to improve vaccination uptake in underrepresented individuals in the future.
+As we have shown in the discussion, there are currently a limited number of centers focused on community-led healthcare in the province.
+Our results showed the existence of differences in vaccination uptake between certain Health Regions and therefore, improving personalized care for underrepresented individuals can serve as a point to improve trust and facilitate vaccine access in these marginalized communities.
 
-<!-- The implementation of the Health Regions in Ontario aimed at reducing the bureaucratic complexity and health disparities identified under the LHIN model.
-However, there are certain challenges that this new model will need to overcome in order to work syngergistically with the rest of the healthcare system of the province.
-First, each Health Region now covers a large geographical area that was served in the past by multiple LHINs.
-This in turn creates a complex socio-demographic landscape within each Health Region that is likely to be different in each case due the different levels of rurality and the proportion of underrepresented groups in each Health Region.
--->
-
-Furthermore, in the event of a future pandemic or public health emergency in the province in the years ahead, PHO and OH will need to collaborate on the implementation of public health measures (such as vaccination campaigns) building on the experience gained during the COVID-19 vaccination rollout.
-So far, the COVID-19 pandemic served as a stress test for both agencies, showing the necessity from decision makers to clearly define how PHO and OH will interact in the future in order to ensure that Ontarians can benefit by virtue of a robust collaboration between both agencies.
-
-<!-- The evidence collected in this study shows that healthcare disparities experienced by underrepresented individuals in the province are the result of complex relationships between infrastructure, resources, occupations, geography, and socio-economic factors, among others.
-Therefore, this complexity presents an ongoing challenge for the healthcare system of Ontario, which will need to adapt and continue to aim for an integrative approach that can serve those that are most vulnerable.
--->
-
-<!-- So far, the evidence collected during the COVID-19 pandemic indicates that differences in vaccination uptake are associated to a lack of infrastructure and resources that can adequately support personalized care to marginalized individuals.
-In the near future, health decision-makers will need to consider the implementation of policies that are focused on addressing this problematic.
--->
-
-Finally, the recent nature in the adaption of the Health Region poses a challenge for researchers in the acquisition of data and information that can be used to analyze the performance of the new model.
-From one side, the Health Regions have not been incorporated as part of Census data (LHINs were considered before in the Census), and this impacts the amount of data that can be obtained.
-On the other hand, evaluations of the Health Region model are at the moment not considered in the Annual Reports of the Auditor General of Ontario, which was critical in the past to identify limitations in the LHIN model.
-Currently, the only demographic information available for each Health Region is provided by Ontario Health (the agency that administers the Health Regions) but this information only provides general estimates that do not allow for detailed analyses of performance indicators (such as hospitalizations, readmissions, and trends in chronic disease incidence) between the Regions.
-Therefore, there is a pressing need for open data that can be used by researchers and decision-makers to assess the performance of the Health Region model.
-
-<!-- The Health Region model will only by successful if it ensures that healthcare improves across all segments of the population of Ontario, particularly in the event of a future public health emergency or pandemic where so far, based on the experience of the COVID-19 pandemic, underrepresented individuals have been disproportionately affected.
--->
+Overall, there are ongoing challenges for healthcare access across different segments of the population of Ontario.
+Future studies can be focused at more in-depth analyses of vaccination uptake within the different socio-demographic groups in each Health Region, in order to provide decision-makers with information that can serve to carefully consider and address how to improve vaccine access to marginalized communities in Ontario in the event of a future public health emergency.

### 2023-07-28 manuscript_v7/appendix.qmd
updated conclusion and discussion

--- 2023-05-26 manuscript_v5/appendix.qmd
@@
       - file: latex/appendix_preamble.sty
-linestretch: 2
+#linestretch: 2
 ---

### 2023-07-28 manuscript_v7/appendix.tex
updated conclusion and discussion

--- 2023-07-10 manuscript_v6/appendix.tex
@@
 \usepackage{amsmath,amssymb}
-\usepackage{setspace}
 \usepackage{iftex}
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, interior hidden, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt]}{\end{tcolorbox}}\fi
 
-\setstretch{2}
 This document provides details on assessing the socio-economic

### 2023-07-28 manuscript_v7/main.tex
updated conclusion and discussion

@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 of vaccination facilities in areas where underrepresented groups reside
-in Toronto\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as
-well as an absence of prioritization of areas inhabited by vulnerable
-groups in the southeastern region of the
+in Toronto when the provincial vaccination program
+started\textsuperscript{\protect\hyperlink{ref-bogoch2022}{7}}, as well
+as an absence of prioritization of areas inhabited by vulnerable groups
+in the southeastern region of the
 US\textsuperscript{\protect\hyperlink{ref-nguyen2021}{26}}. Other
@@
 province with regard to personalized healthcare for marginalized
-individuals. For example, the West Health Region has only two Aboriginal
-Health Access Centres (community-led primary healthcare organizations
-focused on First Nations, Métis, and Inuit communities) to provide care
-to an estimated population of 100,000 Indigenous individuals living in
-the area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{73}}.
-Lack of access to personalized healthcare affects individuals that may
-mistrust the traditional healthcare system due to systemic racism or
-oppression, which is known to be the case for Indigenous and Black
-individuals in Canada, as these rationales have been associated with
-observed lower vaccination rates among these
-groups\textsuperscript{\protect\hyperlink{ref-smylie2022}{74},\protect\hyperlink{ref-eissa2021}{75}}.
-Taken together, this suggests that healthcare disparities specific to
-these underrepresented groups in certain parts of the province impacted
-vaccination uptake, and highlights the need of investments in the Health
-Regions focused on resources, infrastructure, and specially personnel
-that can deliver personalized care to marginalized communities, as it
-has been shown that such efforts have improved trust in vaccination in
+individuals. Indeed, previous studies have pointed out mistrust in the
+traditional healthcare system (due to systemic racism or oppression) as
+a rationale for lower vaccination rates in the case of Indigenous and
+Black individuals in
+Canada\textsuperscript{\protect\hyperlink{ref-smylie2022}{73},\protect\hyperlink{ref-eissa2021}{74}}.
+Other studies have also shown a need for increased culturally responsive
+care for other underrepresented groups such as Black women and and other
+immigrant groups in Canada and
+Ontario\textsuperscript{\protect\hyperlink{ref-etowa2021}{75},\protect\hyperlink{ref-patel2019}{76}}.
+Considering that for example, the West Health Region has only two
+Aboriginal Health Access Centres (community-led primary healthcare
+organizations focused on First Nations, Métis, and Inuit communities) to
+provide care to an estimated population of 100,000 Indigenous
+individuals living in the
+area\textsuperscript{\protect\hyperlink{ref-ontariohealth}{77}}, it is
+possible that the limited personalized healthcare for underrepresented
+groups in certain parts of the province impacted vaccination uptake as
+well, and highlights the need of investments in the Health Regions
+focused on resources, infrastructure, and specially personnel that can
+deliver personalized care to marginalized communities, as it has been
+shown that such efforts have improved trust in vaccination in
 underrepresented groups
-elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{76}}.
+elsewhere\textsuperscript{\protect\hyperlink{ref-schafferderoo2020}{78}}.
 
@@
 Ontario. However, these areas have the highest proportion of Indigenous
-inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{73}}.
+inhabitants\textsuperscript{\protect\hyperlink{ref-ontariohealth}{77}}.
 In the context of personalized care, there is a need to collect data
@@
 vaccination
-status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{77}}, we
+status\textsuperscript{\protect\hyperlink{ref-stephenson2022}{79}}, we
 believe that our data was able to provide a valid sample of vaccination
@@
 October and
-November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{78}});
+November\textsuperscript{\protect\hyperlink{ref-ontario-covid}{80}});
 however, the short time window constitutes essentially a ``snapshot''
@@
 
-The evidence collected in this study shows that differences in the
-uptake of the first dose of the COVID-19 vaccine by underrepresented
-individuals in the province of Ontario were the result of complex
-relationships between infrastructure, resources, occupations, geography,
-and socio-economic factors. Specifically, we identified that between the
-last trimester of 2021 and January of 2022 individuals that identified
-as Arab or Middle Eastern, Black, Latin American, of a mixed background,
-or that belonged to other racial groups had significantly lower odds of
-vaccination when compared to White/Caucasian individuals in Ontario.
-
-However, although overall vaccination uptake was lower in these
+The evidence collected in this study shows differences in the uptake of
+the first dose of the COVID-19 vaccine in Ontario between October 2021
+and January 2022 in underrepresented groups, which had significantly
+lower odds of vaccination when compared to White/Caucasian individuals.
+However, although overall vaccination uptake was lower in
 underrepresented individuals in the province as a whole, the odds of
 vaccination differed within certain income levels and Health Regions in
-these groups, specially for those that reported a low household income
-and that had higher vaccination odds that individuals in the same
-racial/ethnic group with a higher income. These results highlight the
-complex landscape of the province, where varying degrees of rurality
+these groups. Specifically, those that reported a low household income
+had higher vaccination odds when compared to individuals in the same
+racial/ethnic group in a higher income bracket. These results highlight
+the complex landscape of the province, where varying degrees of rurality
 exist in conjunction with a socio-demographic makeup that is unique to
@@
 
-Such complexity presents an ongoing challenge for the healthcare system
-of Ontario, and in particular, to the Health Region model which will
-need to adapt and build on the LHIN model with the aim of providing an
-integrative approach that can serve those that are most vulnerable
-across all the Health Regions of the province. In particular,
-personalized care is an area that could be further developed to improve
+Personalized care is an area that could be further developed to improve
 vaccination uptake in underrepresented individuals in the future. As we
-have shown above, there are currently a limited number of centers
-focused on community-led healthcare in the province, and our results
+have shown in the discussion, there are currently a limited number of
+centers focused on community-led healthcare in the province. Our results
 showed the existence of differences in vaccination uptake between
-certain Health Regions. Improving personalized care for underrepresented
-individuals then, can serve as a point to improve trust and facilitate
-vaccine access in these marginalized communities.
-
-Furthermore, in the event of a future pandemic or public health
-emergency in the province in the years ahead, PHO and OH will need to
-collaborate on the implementation of public health measures (such as
-vaccination campaigns) building on the experience gained during the
-COVID-19 vaccination rollout. So far, the COVID-19 pandemic served as a
-stress test for both agencies, showing the necessity from decision
-makers to clearly define how PHO and OH will interact in the future in
-order to ensure that Ontarians can benefit by virtue of a robust
-collaboration between both agencies.
-
-Finally, the recent nature in the adaption of the Health Region poses a
-challenge for researchers in the acquisition of data and information
-that can be used to analyze the performance of the new model. From one
-side, the Health Regions have not been incorporated as part of Census
-data (LHINs were considered before in the Census), and this impacts the
-amount of data that can be obtained. On the other hand, evaluations of
-the Health Region model are at the moment not considered in the Annual
-Reports of the Auditor General of Ontario, which was critical in the
-past to identify limitations in the LHIN model. Currently, the only
-demographic information available for each Health Region is provided by
-Ontario Health (the agency that administers the Health Regions) but this
-information only provides general estimates that do not allow for
-detailed analyses of performance indicators (such as hospitalizations,
-readmissions, and trends in chronic disease incidence) between the
-Regions. Therefore, there is a pressing need for open data that can be
-used by researchers and decision-makers to assess the performance of the
-Health Region model.
+certain Health Regions and therefore, improving personalized care for
+underrepresented individuals can serve as a point to improve trust and
+facilitate vaccine access in these marginalized communities.
+
+Overall, there are ongoing challenges for healthcare access across
+different segments of the population of Ontario. Future studies can be
+focused at more in-depth analyses of vaccination uptake within the
+different socio-demographic groups in each Health Region, in order to
+provide decision-makers with information that can serve to carefully
+consider and address how to improve vaccine access to marginalized
+communities in Ontario.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
 \CSLLeftMargin{73. }%
-\CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
-Health;
-\url{https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf};
-2022.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-smylie2022}{}}%
-\CSLLeftMargin{74. }%
 \CSLRightInline{Smylie J, McConkey S, Rachlis B, et al. Uncovering
@@
 \leavevmode\vadjust pre{\hypertarget{ref-eissa2021}{}}%
-\CSLLeftMargin{75. }%
+\CSLLeftMargin{74. }%
 \CSLRightInline{Eissa A, Lofters A, Akor N, Prescod C, Nnorom O.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-etowa2021}{}}%
+\CSLLeftMargin{75. }%
+\CSLRightInline{Etowa J, Hyman I. Unpacking the health and social
+consequences of {COVID}-19 through a race, migration and gender lens.
+\emph{Canadian Journal of Public Health}. 2021;112(1):8-11.
+doi:\href{https://doi.org/10.17269/s41997-020-00456-6}{10.17269/s41997-020-00456-6}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-patel2019}{}}%
+\CSLLeftMargin{76. }%
+\CSLRightInline{Patel A, Dean J, Edge S, Wilson K, Ghassemi E. Double
+burden of rural migration in canada? Considering the social determinants
+of health related to immigrant settlement outside the cosmopolis.
+\emph{International Journal of Environmental Research and Public
+Health}. 2019;16(5):678.
+doi:\href{https://doi.org/10.3390/ijerph16050678}{10.3390/ijerph16050678}}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ontariohealth}{}}%
+\CSLLeftMargin{77. }%
+\CSLRightInline{\emph{{A}nnual {B}usiness {P}lan 2022/23}. Ontario
+Health;
+\url{https://www.ontariohealth.ca/sites/ontariohealth/files/2022-05/OHBusinessPlan22_23.pdf};
+2022.}
+
 \leavevmode\vadjust pre{\hypertarget{ref-schafferderoo2020}{}}%
-\CSLLeftMargin{76. }%
+\CSLLeftMargin{78. }%
 \CSLRightInline{DeRoo SS, Pudalov NJ, Fu LY. Planning for a {COVID}-19
@@
 \leavevmode\vadjust pre{\hypertarget{ref-stephenson2022}{}}%
-\CSLLeftMargin{77. }%
+\CSLLeftMargin{79. }%
 \CSLRightInline{Stephenson M, Olson SM, Self WH, et al. Ascertainment of
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ontario-covid}{}}%
-\CSLLeftMargin{78. }%
+\CSLLeftMargin{80. }%
 \CSLRightInline{{Ontario COVID-19 Data Tool}. Accessed February 27,
